### 2023-02-10 manuscript.Rmd
initialize

---
title             : "Full-day, in home validation of infant body position measurements from inertial sensors"
shorttitle        : "Infant position in the home"

author: 
  - name          : "John M. Franchak"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "UC Riverside Department of Psychology, 900 University Avenue, Riverside, CA 92521"
    email         : "franchak@ucr.edu"
  - name          : "Maximilian Tang"
    affiliation   : "1"
  - name          : "Hailey Rousey"
    affiliation   : "1"
  - name          : "Chuan Luo"
    affiliation   : "1"

authornote: |
  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

  Enter author note here.

abstract: |
  Abstract

keywords          : "body position, motor development, everyday experiences, sitting, machine learning"
wordcount         : "X"

bibliography      : "master.bib"

floatsintext      : no
linenumbers       : no
draft             : no
mask              : no
figurelist        : no
tablelist         : no
footnotelist      : no
csl               : apa.csl
classoption       : "man"
output            : papaja::apa6_pdf
header-includes:
  - \raggedbottom
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
library("papaja")
library(knitr)
library(patchwork)
library(tidyverse)
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

# Current

# Methods

## Participants

## Apparatus

## Procedure

## Body position annotation

## Body position classification

# Goal 1: Optimize and validate body position classification model

# Goal 2: Assess classification accuracy across a long delay

# Goal 3: Examine wear time and compliance in full-day data collection

# Goal 4: Compare classification estimates to prior literature

# Discussion

\newpage

# References

::: {#refs custom-style="Bibliography"}
:::

### 2023-02-11 manuscript.Rmd
data files organized

@@
 # Goal 1: Optimize and validate body position classification model
+<!--resampled-metrics, splithalf-metrics, predictions CSVs generated from tune_ml/classification_tidymodels.R -->
 
-# Goal 2: Assess classification accuracy across a long delay
+# Goal 2: Assess classification accuracy over long recordings
+<!-- compiled_agreement.csv, generated from analysis-pt2/compile_agreement.R -->
 
-# Goal 3: Examine wear time and compliance in full-day data collection
+# Goal 3: Compare classification estimates to prior literature
+<!-- imu-compiled.csv, generated from analysis-lena-imu/compile_lena_imu.R -->
 
-# Goal 4: Compare classification estimates to prior literature
+# Goal 4: Examine wear time and compliance in full-day data collection
+<!-- imu-compiled.csv, generated from analysis-lena-imu/compile_lena_imu.R -->
 

### 2023-02-11 manuscript.tex
data files organized

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  man]{apa6}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\ifLuaTeX
\usepackage[bidi=basic]{babel}
\else
\usepackage[bidi=default]{babel}
\fi
\babelprovide[main,import]{english}
% get rid of language-specific shorthands (see #6817):
\let\LanguageShortHands\languageshorthands
\def\languageshorthands#1{}
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

% Table formatting
\usepackage{longtable}
\usepackage{lscape}
% \usepackage[counterclockwise]{rotating}   % Landscape page setup for large tables
\usepackage{multirow}		% Table styling
\usepackage{tabularx}		% Control Column width
\usepackage[flushleft]{threeparttable}	% Allows for three part tables with a specified notes section
\usepackage{threeparttablex}            % Lets threeparttable work with longtable

% Create new environments so endfloat can handle them
% \newenvironment{ltable}
%   {\begin{landscape}\centering\begin{threeparttable}}
%   {\end{threeparttable}\end{landscape}}
\newenvironment{lltable}{\begin{landscape}\centering\begin{ThreePartTable}}{\end{ThreePartTable}\end{landscape}}

% Enables adjusting longtable caption width to table width
% Solution found at http://golatex.de/longtable-mit-caption-so-breit-wie-die-tabelle-t15767.html
\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}

% \setlength{\parindent}{0.5in}
% \setlength{\parskip}{0pt plus 0pt minus 0pt}

% Overwrite redefinition of paragraph and subparagraph by the default LaTeX template
% See https://github.com/crsh/papaja/issues/292
\makeatletter
\renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
  {-1em}%
  {\normalfont\normalsize\bfseries\itshape\typesectitle}}

\renewcommand{\subparagraph}[1]{\@startsection{subparagraph}{5}{1em}%
  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
  {-\z@\relax}%
  {\normalfont\normalsize\itshape\hspace{\parindent}{#1}\textit{\addperi}}{\relax}}
\makeatother

% \usepackage{etoolbox}
\makeatletter
\patchcmd{\HyOrg@maketitle}
  {\section{\normalfont\normalsize\abstractname}}
  {\section*{\normalfont\normalsize\abstractname}}
  {}{\typeout{Failed to patch abstract.}}
\patchcmd{\HyOrg@maketitle}
  {\section{\protect\normalfont{\@title}}}
  {\section*{\protect\normalfont{\@title}}}
  {}{\typeout{Failed to patch title.}}
\makeatother

\usepackage{xpatch}
\makeatletter
\xapptocmd\appendix
  {\xapptocmd\section
    {\addcontentsline{toc}{section}{\appendixname\ifoneappendix\else~\theappendix\fi\\: #1}}
    {}{\InnerPatchFailed}%
  }
{}{\PatchFailed}
\keywords{body position, motor development, everyday experiences, sitting, machine learning\newline\indent Word count: X}
\DeclareDelayedFloatFlavor{ThreePartTable}{table}
\DeclareDelayedFloatFlavor{lltable}{table}
\DeclareDelayedFloatFlavor*{longtable}{table}
\makeatletter
\renewcommand{\efloat@iwrite}[1]{\immediate\expandafter\protected@write\csname efloat@post#1\endcsname{}}
\makeatother
\usepackage{csquotes}
\raggedbottom
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Full-day, in home validation of infant body position measurements from inertial sensors},
  pdfauthor={John M. Franchak1, Maximilian Tang1, Hailey Rousey1, \& Chuan Luo1},
  pdflang={en-EN},
  pdfkeywords={body position, motor development, everyday experiences, sitting, machine learning},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{Full-day, in home validation of infant body position measurements from inertial sensors}
\author{John M. Franchak\textsuperscript{1}, Maximilian Tang\textsuperscript{1}, Hailey Rousey\textsuperscript{1}, \& Chuan Luo\textsuperscript{1}}
\date{}

\shorttitle{Infant position in the home}

\authornote{

Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

Enter author note here.

Correspondence concerning this article should be addressed to John M. Franchak, UC Riverside Department of Psychology, 900 University Avenue, Riverside, CA 92521. E-mail: \href{mailto:franchak@ucr.edu}{\nolinkurl{franchak@ucr.edu}}

}

\affiliation{\phantom{0}}

\abstract{%
Abstract
}

\begin{document}
\maketitle

\hypertarget{current}{%
\section{Current}\label{current}}

\hypertarget{methods}{%
\section{Methods}\label{methods}}

\hypertarget{participants}{%
\subsection{Participants}\label{participants}}

\hypertarget{apparatus}{%
\subsection{Apparatus}\label{apparatus}}

\hypertarget{procedure}{%
\subsection{Procedure}\label{procedure}}

\hypertarget{body-position-annotation}{%
\subsection{Body position annotation}\label{body-position-annotation}}

\hypertarget{body-position-classification}{%
\subsection{Body position classification}\label{body-position-classification}}

\hypertarget{goal-1-optimize-and-validate-body-position-classification-model}{%
\section{Goal 1: Optimize and validate body position classification model}\label{goal-1-optimize-and-validate-body-position-classification-model}}

\hypertarget{goal-2-assess-classification-accuracy-across-a-long-delay}{%
\section{Goal 2: Assess classification accuracy across a long delay}\label{goal-2-assess-classification-accuracy-across-a-long-delay}}

\hypertarget{goal-3-examine-wear-time-and-compliance-in-full-day-data-collection}{%
\section{Goal 3: Examine wear time and compliance in full-day data collection}\label{goal-3-examine-wear-time-and-compliance-in-full-day-data-collection}}

\hypertarget{goal-4-compare-classification-estimates-to-prior-literature}{%
\section{Goal 4: Compare classification estimates to prior literature}\label{goal-4-compare-classification-estimates-to-prior-literature}}

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

\newpage

\hypertarget{references}{%
\section{References}\label{references}}

\hypertarget{refs}{}
\begin{CSLReferences}{0}{0}
\end{CSLReferences}

\end{document}

### 2023-02-12 manuscript.Rmd
group split comparison

--- 2023-02-10 manuscript.Rmd
@@
 ```{r setup, include = FALSE}
-library("papaja")
+library(papaja)
 library(knitr)
@@
 library(tidyverse)
+library(rstatix)
 ```
@@
 
-# Current
+# Current Study
 
@@
 
-# Goal 1: Optimize and validate body position classification model
+# Results
 
-# Goal 2: Assess classification accuracy across a long delay
+## Goal 1: Optimize and validate body position classification model
+<!--resampled-metrics, splithalf-metrics, predictions CSVs generated from tune_ml/classification_tidymodels.R -->
+```{r}
+metrics_rs <- read_csv("data/resampled-metrics.csv") 
+metrics_sp <- read_csv("data/splithalf-metrics.csv")
 
-# Goal 3: Examine wear time and compliance in full-day data collection
+```
 
-# Goal 4: Compare classification estimates to prior literature
+## Goal 2: Assess classification accuracy over long recordings
+<!-- compiled_agreement.csv, generated from analysis-pt2/compile_agreement.R -->
+
+## Goal 3: Compare classification estimates to prior literature
+<!-- imu-compiled.csv, generated from analysis-lena-imu/compile_lena_imu.R -->
+
+## Goal 4: Examine wear time and compliance in full-day data collection
+<!-- imu-compiled.csv, generated from analysis-lena-imu/compile_lena_imu.R -->
 

### 2023-02-12 group_split_comparison/group_split.R

@@ -1,0 +1,49 @@
+library(here)
+library(randomForest)
+library(cvms)
+library(caret)
+library(tidyverse)
+library(rstatix)
+
+#LOAD DATA
+load( here("group_split_comparison","compiled_data_lite.RData"))
+ntree = 50
+not_all_na <- function(x) any(!is.na(x))
+
+training <- slide_filt %>% group_by(id, code) %>% slice_head(prop = .75) %>% ungroup 
+testing <- slide_filt %>% group_by(id, code) %>% slice_tail(prop = .25) %>% ungroup
+
+temp_id <- 9908
+
+group_model <- function(temp_id, training, ntree = 50) {
+  temp_group_training <- training %>% 
+    filter(id != temp_id) %>% 
+    select_if(not_all_na) %>% 
+    mutate(code = fct_drop(code)) %>% 
+    select(-id)
+  group <- randomForest(code ~ ., data = temp_group_training, 
+                        localImp = TRUE, proximity = FALSE, ntree = ntree)
+}
+
+split_model <- function(temp_id, training, ntree = 50) {
+  temp_split_training <- training %>% 
+    filter(id == temp_id) %>% 
+    select_if(not_all_na) %>% 
+    mutate(code = fct_drop(code)) %>% 
+    select(-id)
+  split <- randomForest(code ~ ., data = temp_split_training, 
+                        localImp = TRUE, proximity = FALSE, ntree = ntree)
+}
+
+ids <- unique(training$id)
+group_mods <- map(ids, ~split_model(.x, training, ntree = 50)) %>% set_names(ids)
+split_mods <- map(ids, ~split_model(.x, training, ntree = 50)) %>% set_names(ids)
+
+metrics <- function(rfmodel, testing) {
+  predictions <- predict(rfmodel, testing, type = "class")
+  u <- union(predictions, testing$code)
+  res <- confusion_matrix(factor(testing$code, u),factor(predictions, u))
+}
+res_split <- map2(split_mods, ids, ~metrics(.x, filter(testing, id == .y))) %>% set_names(ids)
+res_split %>% map_dbl(~ .x$`Balanced Accuracy`) %>% sort(.)
+

### 2023-02-12 group_split_comparison/group_split.R
Update group_split.R

@@ -14,10 +14,12 @@ training <- slide_filt %>% group_by(id, code) %>% slice_head(prop = .75) %>% ung
 testing <- slide_filt %>% group_by(id, code) %>% slice_tail(prop = .25) %>% ungroup
 
 temp_id <- 9908
+bad_ids <- c(9909, 10204, 10802)
 
-group_model <- function(temp_id, training, ntree = 50) {
+group_model <- function(temp_id, training, ntree = 50, bad_ids = NULL) {
+  print(temp_id)
   temp_group_training <- training %>% 
-    filter(id != temp_id) %>% 
+    filter(id != temp_id, !(id %in% bad_ids)) %>% 
     select_if(not_all_na) %>% 
     mutate(code = fct_drop(code)) %>% 
     select(-id)
@@ -36,7 +38,8 @@ split_model <- function(temp_id, training, ntree = 50) {
 }
 
 ids <- unique(training$id)
-group_mods <- map(ids, ~split_model(.x, training, ntree = 50)) %>% set_names(ids)
+group_mods <- map(ids, ~group_model(.x, training, ntree = 50)) %>% set_names(ids)
+group_select_mods <- map(ids, ~group_model(.x, training, ntree = 50, bad_ids = bad_ids)) %>% set_names(ids)
 split_mods <- map(ids, ~split_model(.x, training, ntree = 50)) %>% set_names(ids)
 
 metrics <- function(rfmodel, testing) {

### 2023-02-13 group_split_comparison/group_split.R
summary tibbles

@@ -42,11 +42,20 @@ group_mods <- map(ids, ~group_model(.x, training, ntree = 50)) %>% set_names(ids
 group_select_mods <- map(ids, ~group_model(.x, training, ntree = 50, bad_ids = bad_ids)) %>% set_names(ids)
 split_mods <- map(ids, ~split_model(.x, training, ntree = 50)) %>% set_names(ids)
 
+save(group_select_mods, split_mods, file = "group_split_comparison/models.RData")
+
 metrics <- function(rfmodel, testing) {
   predictions <- predict(rfmodel, testing, type = "class")
   u <- union(predictions, testing$code)
   res <- confusion_matrix(factor(testing$code, u),factor(predictions, u))
 }
-res_split <- map2(split_mods, ids, ~metrics(.x, filter(testing, id == .y))) %>% set_names(ids)
-res_split %>% map_dbl(~ .x$`Balanced Accuracy`) %>% sort(.)
+res_split <- map2(split_mods, ids, ~metrics(.x, filter(testing, id == .y))) %>% 
+              map_dfr(~ select(.x, `Overall Accuracy`:Kappa)) %>% 
+              add_column(ids) %>% add_column(model = "split")
+res_split %>% get_summary_stats(-ids)
+
+res_select <- map2(group_select_mods, ids, ~metrics(.x, filter(testing, id == .y))) %>% 
+  map_dfr(~ select(.x, `Overall Accuracy`:Kappa)) %>% 
+  add_column(ids) %>% add_column(model = "select")
+res_select %>% get_summary_stats(-ids)
 

### 2023-02-13 group_split_comparison/group_split.R
furrr

@@ -4,6 +4,9 @@ library(cvms)
 library(caret)
 library(tidyverse)
 library(rstatix)
+library(furrr)
+
+plan("multicore", workers = 9)
 
 #LOAD DATA
 load( here("group_split_comparison","compiled_data_lite.RData"))
@@ -13,7 +16,6 @@ not_all_na <- function(x) any(!is.na(x))
 training <- slide_filt %>% group_by(id, code) %>% slice_head(prop = .75) %>% ungroup 
 testing <- slide_filt %>% group_by(id, code) %>% slice_tail(prop = .25) %>% ungroup
 
-temp_id <- 9908
 bad_ids <- c(9909, 10204, 10802)
 
 group_model <- function(temp_id, training, ntree = 50, bad_ids = NULL) {
@@ -38,11 +40,12 @@ split_model <- function(temp_id, training, ntree = 50) {
 }
 
 ids <- unique(training$id)
-group_mods <- map(ids, ~group_model(.x, training, ntree = 50)) %>% set_names(ids)
-group_select_mods <- map(ids, ~group_model(.x, training, ntree = 50, bad_ids = bad_ids)) %>% set_names(ids)
-split_mods <- map(ids, ~split_model(.x, training, ntree = 50)) %>% set_names(ids)
+group_mods <- future_map(ids, ~group_model(.x, training, ntree = 50)) %>% set_names(ids)
+group_select_mods <- future_map(ids, ~group_model(.x, training, ntree = 50, bad_ids = bad_ids)) %>% set_names(ids)
+split_mods <- future_map(ids, ~split_model(.x, training, ntree = 50)) %>% set_names(ids)
 
-save(group_select_mods, split_mods, file = "group_split_comparison/models.RData")
+# save(group_select_mods, split_mods, file = "group_split_comparison/models.RData")
+# load( here("group_split_comparison","models.RData"))
 
 metrics <- function(rfmodel, testing) {
   predictions <- predict(rfmodel, testing, type = "class")
@@ -59,3 +62,4 @@ res_select <- map2(group_select_mods, ids, ~metrics(.x, filter(testing, id == .y
   add_column(ids) %>% add_column(model = "select")
 res_select %>% get_summary_stats(-ids)
 
+ds <- bind_rows(res_split, res_select)

### 2023-02-13 group_split_comparison/group_split.R
write outputs

@@ -6,7 +6,8 @@ library(tidyverse)
 library(rstatix)
 library(furrr)
 
-plan("multicore", workers = 9)
+plan("multisession", workers = 9)
+plan("sequential")
 
 #LOAD DATA
 load( here("group_split_comparison","compiled_data_lite.RData"))
@@ -52,14 +53,24 @@ metrics <- function(rfmodel, testing) {
   u <- union(predictions, testing$code)
   res <- confusion_matrix(factor(testing$code, u),factor(predictions, u))
 }
-res_split <- map2(split_mods, ids, ~metrics(.x, filter(testing, id == .y))) %>% 
-              map_dfr(~ select(.x, `Overall Accuracy`:Kappa)) %>% 
-              add_column(ids) %>% add_column(model = "split")
-res_split %>% get_summary_stats(-ids)
+
+res_group <- map2(group_mods, ids, ~metrics(.x, filter(testing, id == .y))) %>% 
+  map_dfr(~ select(.x, `Overall Accuracy`:Kappa)) %>% 
+  add_column(ids) %>% add_column(model = "group")
+res_group %>% get_summary_stats(-ids)
 
 res_select <- map2(group_select_mods, ids, ~metrics(.x, filter(testing, id == .y))) %>% 
   map_dfr(~ select(.x, `Overall Accuracy`:Kappa)) %>% 
   add_column(ids) %>% add_column(model = "select")
 res_select %>% get_summary_stats(-ids)
 
-ds <- bind_rows(res_split, res_select)
+res_split <- map2(split_mods, ids, ~metrics(.x, filter(testing, id == .y))) %>% 
+  map_dfr(~ select(.x, `Overall Accuracy`:Kappa)) %>% 
+  add_column(ids) %>% add_column(model = "split")
+res_split %>% get_summary_stats(-ids)
+
+ds <- bind_rows(res_split, res_select) %>% bind_rows(res_group)
+write(ds, file = here("data", "group_split_metrics.csv"))
+
+
+

### 2023-02-13 group_split_comparison/group_split.R
updated data files for paper

@@ -11,13 +11,13 @@ plan("sequential")
 
 #LOAD DATA
 load( here("group_split_comparison","compiled_data_lite.RData"))
-ntree = 50
+ntree = 150
 not_all_na <- function(x) any(!is.na(x))
 
 training <- slide_filt %>% group_by(id, code) %>% slice_head(prop = .75) %>% ungroup 
 testing <- slide_filt %>% group_by(id, code) %>% slice_tail(prop = .25) %>% ungroup
 
-bad_ids <- c(9909, 10204, 10802)
+bad_ids <- c(9909, 9911, 10204, 10802, 11104, 11501)
 
 group_model <- function(temp_id, training, ntree = 50, bad_ids = NULL) {
   print(temp_id)
@@ -41,9 +41,9 @@ split_model <- function(temp_id, training, ntree = 50) {
 }
 
 ids <- unique(training$id)
-group_mods <- future_map(ids, ~group_model(.x, training, ntree = 50)) %>% set_names(ids)
-group_select_mods <- future_map(ids, ~group_model(.x, training, ntree = 50, bad_ids = bad_ids)) %>% set_names(ids)
-split_mods <- future_map(ids, ~split_model(.x, training, ntree = 50)) %>% set_names(ids)
+group_mods <- future_map(ids, ~group_model(.x, training)) %>% set_names(ids)
+group_select_mods <- future_map(ids, ~group_model(.x, training, bad_ids = bad_ids)) %>% set_names(ids)
+split_mods <- future_map(ids, ~split_model(.x, training)) %>% set_names(ids)
 
 # save(group_select_mods, split_mods, file = "group_split_comparison/models.RData")
 # load( here("group_split_comparison","models.RData"))
@@ -68,9 +68,12 @@ res_split <- map2(split_mods, ids, ~metrics(.x, filter(testing, id == .y))) %>%
   map_dfr(~ select(.x, `Overall Accuracy`:Kappa)) %>% 
   add_column(ids) %>% add_column(model = "split")
 res_split %>% get_summary_stats(-ids)
+res_split %>% arrange(`Overall Accuracy`)
 
-ds <- bind_rows(res_split, res_select) %>% bind_rows(res_group)
-write(ds, file = here("data", "group_split_metrics.csv"))
+ds <- bind_rows(res_split, res_group) %>% bind_rows(res_select)
[…]

### 2023-02-14 group_split_comparison/group_split.R
updated metrics

@@ -4,10 +4,10 @@ library(cvms)
 library(caret)
 library(tidyverse)
 library(rstatix)
-library(furrr)
-
-plan("multisession", workers = 9)
-plan("sequential")
+# library(furrr)
+# 
+# plan("multisession", workers = 4)
+# plan("sequential")
 
 #LOAD DATA
 load( here("group_split_comparison","compiled_data_lite.RData"))
@@ -17,7 +17,7 @@ not_all_na <- function(x) any(!is.na(x))
 training <- slide_filt %>% group_by(id, code) %>% slice_head(prop = .75) %>% ungroup 
 testing <- slide_filt %>% group_by(id, code) %>% slice_tail(prop = .25) %>% ungroup
 
-bad_ids <- c(9909, 9911, 10204, 10802, 11104, 11501)
+bad_ids <- c(9909, 10204, 12502, 9911, 10701, 10802, 12002, 9910, 11104, 12101, 11501, 11601)
 
 group_model <- function(temp_id, training, ntree = 50, bad_ids = NULL) {
   print(temp_id)
@@ -41,9 +41,9 @@ split_model <- function(temp_id, training, ntree = 50) {
 }
 
 ids <- unique(training$id)
-group_mods <- future_map(ids, ~group_model(.x, training)) %>% set_names(ids)
-group_select_mods <- future_map(ids, ~group_model(.x, training, bad_ids = bad_ids)) %>% set_names(ids)
-split_mods <- future_map(ids, ~split_model(.x, training)) %>% set_names(ids)
+group_mods <-map(ids, ~group_model(.x, training)) %>% set_names(ids)
+group_select_mods <- map(ids, ~group_model(.x, training, bad_ids = bad_ids)) %>% set_names(ids)
+split_mods <- map(ids, ~split_model(.x, training)) %>% set_names(ids)
 
 # save(group_select_mods, split_mods, file = "group_split_comparison/models.RData")
 # load( here("group_split_comparison","models.RData"))

### 2023-02-14 manuscript.Rmd
updated data files for paper
figuring out better select model

@@
 ## Goal 1: Optimize and validate body position classification model
-<!--resampled-metrics, splithalf-metrics, predictions CSVs generated from tune_ml/classification_tidymodels.R -->
+<!-- group_split_metrics, generated from group_split_comparison in this proj, stable copy in data/ -->
 ```{r}
-metrics_rs <- read_csv("data/resampled-metrics.csv") 
-metrics_sp <- read_csv("data/splithalf-metrics.csv")
+metrics <- read_csv("data/group_split_metrics.csv")
 
@@
 ## Goal 2: Assess classification accuracy over long recordings
-<!-- compiled_agreement.csv, generated from analysis-pt2/compile_agreement.R -->
+<!-- compiled_agreement, generated from analysis-part2 -->
+
+```{r}
+
+```
 

### 2023-02-14 group_split_comparison/group_split.R

@@ -77,3 +77,5 @@ ggplot(ds, aes(x = model, y = `Overall Accuracy`)) + geom_boxplot()
 ggplot(ds, aes(x = model, y = `Balanced Accuracy`)) + geom_boxplot()
 ggplot(ds, aes(x = model, y = `Kappa`)) + geom_boxplot()
 
+ds %>% pivot_wider(id_cols = ids, names_from = model, values_from = `Overall Accuracy`) %>% 
+  ggplot(aes(x = split, y = group)) + geom_point() + xlim(0,1) + ylim(0,1) + geom_abline(slope = 1, intercept = 0)

### 2023-02-16 manuscript.Rmd
part 2 graphs

@@
 knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
+pos_levels <- c("Supine", "Prone", "Sitting", "Upright", "Held")
+pal <-  c("#E69F00","#56B4E9", "#009E73","#F0E442", "#0072B2") %>% 
+  set_names(pos_levels)
+
+theme_update(text = element_text(size = 14),
+             axis.text.x = element_text(size = 14, color = "black"), 
+             axis.title.x = element_text(size = 16),
+             axis.text.y = element_text(size = 14,  color = "black"), 
+             axis.title.y = element_text(size = 16), 
+             panel.background = element_blank(),panel.border = element_blank(), 
+             panel.grid.major = element_blank(),
+             panel.grid.minor = element_blank(), axis.line = element_blank(), 
+             legend.key = element_rect(fill = "white"))
+
 ```
@@
 ```{r}
+# GROUP MODEL PART 2 AGREEMENT
+pt2_group <- read_csv("data/compiled_agreement_position.csv") %>% 
+  filter(nap_period == 0 & exclude_period == 0 & !is.na(code) & !is.na(pos)) %>% 
+         mutate(pos = factor(pos, levels = pos_levels))
 
+# Count up position totals and join into an agreement data frame
+model <- pt2_group %>% count(file, pos) %>% add_count(file, wt = n, name = "total") %>% mutate(prop_model = n/total)
+human <- pt2_group %>% count(file, code) %>% add_count(file, wt = n, name = "total") %>% mutate(prop_human = n/total)
+agree_group <- full_join(select(model, file, pos, prop_model), 
+                   select(human, file, code, prop_human), 
+                   by = c("file" = "file", "pos" = "code"))
+agree_group <- agree_group %>% mutate(prop_human = ifelse(is.na(prop_human), 0, prop_human),
+                          prop_model = ifelse(is.na(prop_model), 0, prop_model))
+agree_group <- complete(agree_group, file, pos, fill = list(prop_human = 0, prop_model = 0))
+
+group_agree_cor <- cor_test(agree_group, vars = c("prop_model", "prop_human")) %>% pull(cor)
+
+ggplot(agree_group) +
+  geom_point(aes(x = prop_human, y = prop_model, fill = pos), shape = 21, alpha = .5, size = 4) + 
+  geom_abline(intercept = 0, slope = 1) + 
+  scale_fill_manual(name = "Posture", values = pal, guide = "none") + 
+  scale_x_continuous(breaks = c(0,.25, .5, .75, 1), labels = c(0, 25, 50, 75, 100), limits = c(0, 1.07), name = "Actual Position Prevalence (%)") + 
+  scale_y_continuous(breaks = c(0,.25, .5, .75, 1), labels = c(0, 25, 50, 75, 100), limits = c(0, 1.07), name = "Predicted Position Prevalence (%)") + 
+  facet_wrap("pos", ncol = 2, as.table = F) +
+  theme(panel.spacing.x = unit(1.5, "lines")) +
+  theme(panel.spacing.y = unit(2.5, "lines")) + 
+  theme(strip.background = element_blank(), strip.text.x = element_text(size = 16, face = "bold"))
 ```

### 2023-02-16 manuscript.Rmd
polished graphs

--- 2023-02-14 manuscript.Rmd
@@
 <!-- compiled_agreement, generated from analysis-part2 -->
+```{r}
+# Count up position totals and join into an agreement data frame
+calculate_agreement <- function(ds) {
+  model <- ds %>% count(file, pos) %>% add_count(file, wt = n, name = "total") %>% mutate(prop_model = n/total)
+  human <- ds %>% count(file, code) %>% add_count(file, wt = n, name = "total") %>% mutate(prop_human = n/total)
+  agree_group <- full_join(select(model, file, pos, prop_model), 
+                     select(human, file, code, prop_human), 
+                     by = c("file" = "file", "pos" = "code"))
+  agree_group <- agree_group %>% mutate(prop_human = ifelse(is.na(prop_human), 0, prop_human),
+                            prop_model = ifelse(is.na(prop_model), 0, prop_model))
+  agree_group <- complete(agree_group, file, pos, fill = list(prop_human = 0, prop_model = 0))
+}
 
-```{r}
+agreement_scatter <- function(ds) {
+  ggplot(ds) +
+  geom_point(aes(x = prop_human, y = prop_model, fill = pos), shape = 21, alpha = .5, size = 4) + 
+  geom_abline(intercept = 0, slope = 1) + 
+  scale_fill_manual(name = "Posture", values = pal, guide = "none") + 
+  scale_x_continuous(breaks = c(0,.5,1), labels = c(0, 50, 100), limits = c(-.07, 1.07), name = "Actual Prevalence (%)") + 
+  scale_y_continuous(breaks = c(0,.5,1), labels = c(0, 50, 100), limits = c(-.07, 1.07), name = "Predicted Prevalence (%)") + 
+  facet_wrap("pos", ncol = 2, as.table = F) +
+  theme(panel.spacing.x = unit(1, "lines")) +
+  theme(panel.spacing.y = unit(1, "lines")) + 
+  theme(strip.background = element_blank(), strip.text.x = element_text(size = 16, face = "bold"))
+}
+```
 
+```{r part2overall, warning=FALSE, fig.cap="Overall agreement between human-coded body position and model-predicted body position in the long-delay period.
+Agreement for group models is shown in (A) and agreement for individual models is shown in (B).
+Plots are shown separately for each body position with a reference line that indicates perfect agreement; each point in a plot represent data for a single participant.", fig.align="center", fig.width=7, fig.height=6.7}
+# GROUP MODEL PART 2 AGREEMENT
+pt2_group <- read_csv("data/compiled_agreement_position.csv") %>% 
+  filter(nap_period == 0 & exclude_period == 0 & !is.na(code) & !is.na(pos)) %>% 
+         mutate(pos = factor(pos, levels = pos_levels))
+pt2_split <- read_csv("data/compiled_agreement_split.csv") %>% 
+  filter(nap_period == 0 & exclude_period == 0 & !is.na(code) & !is.na(pos)) %>% 
+         mutate(pos = factor(pos, levels = pos_levels))
+
+agree <- map(list("group" = pt2_group, "split" = pt2_split), calculate_agreement)
+
+group_agree_cor <- cor_test(agree$group, vars = c("prop_model", "prop_human")) %>% pull(cor)
+split_agree_cor <- cor_test(agree$split, vars = c("prop_model", "prop_human")) %>% pull(cor)
+
+agree_plots <- map(agree, agreement_scatter)
+agree_plots <- map2(agree_plots, c("A. Group Models", "B. Individual Models"), ~ .x + ggtitle(.y))
+
+agree_plots[[1]] + agree_plots[[2]]
 ```

### 2023-02-16 manuscript.tex
polished graphs

@@
 
-\hypertarget{current}{%
-\section{Current}\label{current}}
+\hypertarget{current-study}{%
+\section{Current Study}\label{current-study}}
 
@@
 
+\hypertarget{results}{%
+\section{Results}\label{results}}
+
 \hypertarget{goal-1-optimize-and-validate-body-position-classification-model}{%
-\section{Goal 1: Optimize and validate body position classification model}\label{goal-1-optimize-and-validate-body-position-classification-model}}
+\subsection{Goal 1: Optimize and validate body position classification model}\label{goal-1-optimize-and-validate-body-position-classification-model}}
 
-\hypertarget{goal-2-assess-classification-accuracy-across-a-long-delay}{%
-\section{Goal 2: Assess classification accuracy across a long delay}\label{goal-2-assess-classification-accuracy-across-a-long-delay}}
+\hypertarget{goal-2-assess-classification-accuracy-over-long-recordings}{%
+\subsection{Goal 2: Assess classification accuracy over long recordings}\label{goal-2-assess-classification-accuracy-over-long-recordings}}
 
-\hypertarget{goal-3-examine-wear-time-and-compliance-in-full-day-data-collection}{%
-\section{Goal 3: Examine wear time and compliance in full-day data collection}\label{goal-3-examine-wear-time-and-compliance-in-full-day-data-collection}}
+\begin{figure}
 
-\hypertarget{goal-4-compare-classification-estimates-to-prior-literature}{%
-\section{Goal 4: Compare classification estimates to prior literature}\label{goal-4-compare-classification-estimates-to-prior-literature}}
+{\centering \includegraphics{manuscript_files/figure-latex/part2overall-1}
+
+}
+
+\caption{Overall agreement between human-coded body position and model-predicted body position in the long-delay period.
+Agreement for group models is shown in (A) and agreement for individual models is shown in (B).
+Plots are shown separately for each body position with a reference line that indicates perfect agreement; each point in a plot represent data for a single participant.}\label{fig:part2overall}
+\end{figure}
+
+\hypertarget{goal-3-compare-classification-estimates-to-prior-literature}{%
+\subsection{Goal 3: Compare classification estimates to prior literature}\label{goal-3-compare-classification-estimates-to-prior-literature}}
+
+\hypertarget{goal-4-examine-wear-time-and-compliance-in-full-day-data-collection}{%
+\subsection{Goal 4: Examine wear time and compliance in full-day data collection}\label{goal-4-examine-wear-time-and-compliance-in-full-day-data-collection}}
 

### 2023-02-16 manuscript.Rmd
made a difference graph but don't need it

@@
 calculate_agreement <- function(ds) {
-  model <- ds %>% count(file, pos) %>% add_count(file, wt = n, name = "total") %>% mutate(prop_model = n/total)
-  human <- ds %>% count(file, code) %>% add_count(file, wt = n, name = "total") %>% mutate(prop_human = n/total)
+  model <- ds %>% count(file, pos) %>% 
+    add_count(file, wt = n, name = "total") %>% 
+    mutate(prop_model = n/total)
+  human <- ds %>% count(file, code) %>% 
+    add_count(file, wt = n, name = "total") %>% 
+    mutate(prop_human = n/total)
   agree_group <- full_join(select(model, file, pos, prop_model), 
@@
   geom_abline(intercept = 0, slope = 1) + 
-  scale_fill_manual(name = "Posture", values = pal, guide = "none") + 
+  scale_fill_manual(values = pal, guide = "none") + 
   scale_x_continuous(breaks = c(0,.5,1), labels = c(0, 50, 100), limits = c(-.07, 1.07), name = "Actual Prevalence (%)") + 
@@
 }
+
+# difference_plot <- function(ds) {
+#   ds %>% mutate(diff = prop_human - prop_model) %>% 
+#   ggplot(aes(x = pos, y = diff, color = pos)) +
+#   stat_summary() + 
+#   scale_fill_manual(values = pal, guide = "none") + 
+#   ylab("Prediction Difference (%)") + ylim(-.5, .5) + xlab("")
+# }
 ```
@@
 Agreement for group models is shown in (A) and agreement for individual models is shown in (B).
-Plots are shown separately for each body position with a reference line that indicates perfect agreement; each point in a plot represent data for a single participant.", fig.align="center", fig.width=7, fig.height=6.7}
+Plots are shown separately for each body position with a reference line that indicates perfect agreement; each point in a plot represent data for a single participant.", fig.align="center", fig.width=7, fig.height=6.5}
 # GROUP MODEL PART 2 AGREEMENT
@@
 agree_plots <- map2(agree_plots, c("A. Group Models", "B. Individual Models"), ~ .x + ggtitle(.y))
+agree_plots[[1]] + agree_plots[[2]]
+```
 
-agree_plots[[1]] + agree_plots[[2]]
+```{r, pt2overalltable}
+map_dfc(agree, ~ .x %>% group_by(pos) %>% 
+          cor_test(vars = c("prop_model", "prop_human")) %>% 
+          select(pos, cor))
+map_dfc(agree, ~ .x %>% filter(file != "107/3", file != "104/1") %>% group_by(pos) %>% 
+          cor_test(vars = c("prop_model", "prop_human")) %>% 
+          select(pos, cor))
+
+# difference_plots <- map(agree, difference_plot)
+# difference_plots <- map2(difference_plots, c("A. Group Models", "B. Individual Models"), ~ .x + ggtitle(.y))
+# difference_plots[[1]] + difference_plots[[2]]
+
 ```

### 2023-02-17 manuscript.Rmd
pt2 correlation table

@@
 ```{r, pt2overalltable}
-map_dfc(agree, ~ .x %>% group_by(pos) %>% 
-          cor_test(vars = c("prop_model", "prop_human")) %>% 
-          select(pos, cor))
-map_dfc(agree, ~ .x %>% filter(file != "107/3", file != "104/1") %>% group_by(pos) %>% 
-          cor_test(vars = c("prop_model", "prop_human")) %>% 
-          select(pos, cor))
+pt2corr_with_outliers <- map_dfc(agree, ~ .x %>% group_by(pos) %>% cor_test(vars = c("prop_model", "prop_human")) %>% select(pos, cor)) %>% select(-`pos...3`)
+pt2corr_without_outliers <- map_dfc(agree, ~ .x %>% filter(file != "107/3", file != "104/1") %>% group_by(pos) %>% cor_test(vars = c("prop_model", "prop_human")) %>% select(pos, cor)) %>% select(-`pos...1`,-`pos...3`)
+pt2corr <- bind_cols(pt2corr_with_outliers, pt2corr_without_outliers)
+rm(pt2corr_with_outliers)
+rm(pt2corr_without_outliers)
+apa_table(pt2corr, col.names = c("Position", "Group", "Individual", "Group", "Individual"), col_spanners = list(`With Outliers` = c(2, 3), `Without Outliers` = c(4, 5)))
 

### 2023-02-17 manuscript.Rmd
overall row

--- 2023-02-16 manuscript.Rmd
@@
 
-group_agree_cor <- cor_test(agree$group, vars = c("prop_model", "prop_human")) %>% pull(cor)
-split_agree_cor <- cor_test(agree$split, vars = c("prop_model", "prop_human")) %>% pull(cor)
-
 agree_plots <- map(agree, agreement_scatter)
@@
 ```{r, pt2overalltable}
-map_dfc(agree, ~ .x %>% group_by(pos) %>% 
-          cor_test(vars = c("prop_model", "prop_human")) %>% 
-          select(pos, cor))
-map_dfc(agree, ~ .x %>% filter(file != "107/3", file != "104/1") %>% group_by(pos) %>% 
-          cor_test(vars = c("prop_model", "prop_human")) %>% 
-          select(pos, cor))
+pt2corr_with_outliers <- map_dfc(agree, ~ .x %>% group_by(pos) %>% cor_test(vars = c("prop_model", "prop_human")) %>% select(pos, cor)) %>% select(-`pos...3`)
+pt2corr_without_outliers <- map_dfc(agree, ~ .x %>% filter(file != "107/3", file != "104/1") %>% group_by(pos) %>% cor_test(vars = c("prop_model", "prop_human")) %>% select(pos, cor)) %>% select(-`pos...1`,-`pos...3`)
+pt2corr <- bind_cols(pt2corr_with_outliers, pt2corr_without_outliers)
+rm(pt2corr_with_outliers)
+rm(pt2corr_without_outliers)
+
+#Overall
+group_agree_cor <- cor_test(agree$group, vars = c("prop_model", "prop_human")) %>% pull(cor)
+split_agree_cor <- cor_test(agree$split, vars = c("prop_model", "prop_human")) %>% pull(cor)
+group_agree_cor_noout <- agree$group %>% filter(file != "107/3", file != "104/1") %>% 
+  cor_test(vars = c("prop_model", "prop_human")) %>% pull(cor)
+split_agree_cor_noout <- agree$split %>% filter(file != "107/3", file != "104/1") %>%
+  cor_test(vars = c("prop_model", "prop_human")) %>% pull(cor)
+overall_row <- tibble("Overall", group_agree_cor, split_agree_cor, group_agree_cor_noout, split_agree_cor_noout)
+colnames(overall_row) <- colnames(pt2corr)
+pt2corr <- bind_rows(pt2corr, overall_row)
+apa_table(pt2corr, col.names = c("Position", "Group", "Individual", "Group", "Individual"), col_spanners = list(`With Outliers` = c(2, 3), `Without Outliers` = c(4, 5)),
+          midrules = 5)
 

### 2023-02-17 manuscript.Rmd
graphing 10 minute bins

@@
 
+```{r }
+calculate_agreement_bins <- function(ds) {
+  ds <- ds %>% group_by(file) %>% 
+    mutate(sample = row_number(), bin = floor(sample/300)) %>% 
+    add_count(file, bin) %>% ungroup() %>% filter(n > 200) %>% group_by(file) %>% mutate(num_bins = length(unique(bin))) %>% ungroup()
+  model <- ds %>% count(file, bin, num_bins, pos) %>% add_count(file, bin,num_bins,  wt = n, name = "total") %>% mutate(prop_model = n/total)
+  human <- ds %>% count(file, bin, num_bins,  code) %>% add_count(file, bin, num_bins, wt = n, name = "total") %>% mutate(prop_human = n/total)
+  agree <- full_join(select(model, file, pos, prop_model, bin, num_bins), 
+                     select(human, file, code, prop_human, bin, num_bins), 
+                     by = c("file" = "file", "pos" = "code", "bin" = "bin", "num_bins" = "num_bins"))
+  agree <- agree %>% mutate(prop_human = ifelse(is.na(prop_human), 0, prop_human),
+                            prop_model = ifelse(is.na(prop_model), 0, prop_model))
+  agree <- complete(agree, nesting(file, bin, num_bins), pos, fill = list(prop_human = 0, prop_model = 0))
+}
+
+agree_bins <- map(list("group" = pt2_group, "split" = pt2_split), calculate_agreement_bins)
+
+# individual_scatter <- function(ds, pos_select) {
+#   ds %>% filter(pos == pos_select) %>% ggplot() +
+#   geom_abline(intercept = 0, slope = 1) + facet_wrap("file") + 
+#   geom_point(aes(x = prop_human, y = prop_model, fill = pos), shape = 21, alpha = .5, size = 4) + 
+#   scale_fill_manual(values = pal, guide = "none") + 
+#   scale_x_continuous(breaks = c(0,.5,1), labels = c(0, 50, 100), limits = c(-.07, 1.07), name = "Actual Prevalence (%)") + 
+#   scale_y_continuous(breaks = c(0,.5,1), labels = c(0, 50, 100), limits = c(-.07, 1.07), name = "Predicted Prevalence (%)") + 
+#   theme(panel.spacing.x = unit(1.5, "lines")) +
+#   theme(panel.spacing.y = unit(.75, "lines")) + 
+#   theme(strip.background = element_blank(), strip.text.x = element_blank())
+# }
+
+# individual_scatter <- function(ds) {
+#   ds %>% ggplot() +
+#   geom_abline(intercept = 0, slope = 1) + facet_grid(c("file","pos")) + 
+#   geom_point(aes(x = prop_human, y = prop_model, fill = pos), shape = 21, alpha = .5, size = 3) + 
+#   scale_fill_manual(values = pal, guide = "none") + 
+#   scale_x_continuous(breaks = c(0,.5,1), labels = c(0, 50, 100), limits = c(-.07, 1.07), name = "Actual Prevalence (%)") + 
+#   scale_y_continuous(breaks = c(0,.5,1), labels = c(0, 50, 100), limits = c(-.07, 1.07), name = "Predicted Prevalence (%)") + 
+#   theme(panel.spacing.x = unit(.5, "lines")) +
+#   theme(strip.background = element_blank(), strip.text.x = element_blank()) + 
+#    theme(axis.text.x = element_blank(), axis.text.y = element_blank())
+# }
+# individual_scatter(agree_bins$group)
+
+# agree_plots_bins <- map(agree_bins, ~ individual_scatter(.x, "Prone"))
+# agree_plots_bins <- map2(agree_plots_bins, c("A. Group Models", "B. Individual Models"), ~ .x + ggtitle(.y))
+# agree_plots_bins[[1]] / agree_plots_bins[[2]]
+# (individual_scatter(agree_bins$group, "Sitting") + individual_scatter(agree_bins$group, "Upright"))/
+#   (individual_scatter(agree_bins$group, "Prone") + individual_scatter(agree_bins$group, "Supine"))
+#
+
+```
+
 ## Goal 3: Compare classification estimates to prior literature

### 2023-02-17 manuscript.Rmd
table

@@
 
+#Bins
+pt2corr_with_outliers <- map_dfc(agree_bins, ~ .x %>% group_by(pos) %>% cor_test(vars = c("prop_model", "prop_human")) %>% select(pos, cor)) %>% select(-`pos...3`)
+pt2corr_without_outliers <- map_dfc(agree_bins, ~ .x %>% filter(file != "107/3", file != "104/1") %>% group_by(pos) %>% cor_test(vars = c("prop_model", "prop_human")) %>% select(pos, cor)) %>% select(-`pos...1`,-`pos...3`)
+pt2corr <- bind_cols(pt2corr_with_outliers, pt2corr_without_outliers)
+rm(pt2corr_with_outliers)
+rm(pt2corr_without_outliers)
+
+group_agree_cor <- cor_test(agree_bins$group, vars = c("prop_model", "prop_human")) %>% pull(cor)
+split_agree_cor <- cor_test(agree_bins$split, vars = c("prop_model", "prop_human")) %>% pull(cor)
+group_agree_cor_noout <- agree_bins$group %>% filter(file != "107/3", file != "104/1") %>% 
+  cor_test(vars = c("prop_model", "prop_human")) %>% pull(cor)
+split_agree_cor_noout <- agree_bins$split %>% filter(file != "107/3", file != "104/1") %>%
+  cor_test(vars = c("prop_model", "prop_human")) %>% pull(cor)
+overall_row <- tibble("Overall", group_agree_cor, split_agree_cor, group_agree_cor_noout, split_agree_cor_noout)
+colnames(overall_row) <- colnames(pt2corr)
+pt2corr <- bind_rows(pt2corr, overall_row)
+apa_table(pt2corr, col.names = c("Position", "Group", "Individual", "Group", "Individual"), col_spanners = list(`With Outliers` = c(2, 3), `Without Outliers` = c(4, 5)),
+          midrules = 5)
+
 # individual_scatter <- function(ds, pos_select) {

### 2023-02-17 manuscript.tex
pt2 correlation table
overall row
table

@@
 
+\begin{table}[tbp]
+
+\begin{center}
+\begin{threeparttable}
+
+\caption{\label{tab:pt2overalltable}}
+
+\begin{tabular}{lllll}
+\toprule
+ & \multicolumn{2}{c}{With Outliers} & \multicolumn{2}{c}{Without Outliers} \\
+\cmidrule(r){2-3} \cmidrule(r){4-5}
+Position & Group & Individual & Group & Individual\\
+\midrule
+Held & -0.02 & 0.16 & 0.59 & 0.65\\
+Prone & 0.97 & 0.83 & 0.97 & 0.82\\
+Sitting & 0.72 & 0.93 & 0.75 & 0.98\\
+Supine & 0.84 & 0.93 & 0.87 & 0.97\\
+Upright & 0.84 & 0.93 & 0.99 & 0.94\\ \midrule
+Overall & 0.79 & 0.90 & 0.89 & 0.96\\
+\bottomrule
+\end{tabular}
+
+\end{threeparttable}
+\end{center}
+
+\end{table}
+
+\begin{table}[tbp]
+
+\begin{center}
+\begin{threeparttable}
+
+\caption{\label{tab:unnamed-chunk-3}}
+
+\begin{tabular}{lllll}
+\toprule
+ & \multicolumn{2}{c}{With Outliers} & \multicolumn{2}{c}{Without Outliers} \\
+\cmidrule(r){2-3} \cmidrule(r){4-5}
+Position & Group & Individual & Group & Individual\\
+\midrule
+Held & 0.45 & 0.44 & 0.59 & 0.57\\
+Prone & 0.96 & 0.81 & 0.96 & 0.81\\
+Sitting & 0.72 & 0.93 & 0.73 & 0.94\\
+Supine & 0.75 & 0.93 & 0.76 & 0.94\\
+Upright & 0.93 & 0.95 & 0.98 & 0.96\\ \midrule
+Overall & 0.80 & 0.92 & 0.81 & 0.93\\
+\bottomrule
+\end{tabular}
+
+\end{threeparttable}
+\end{center}
+
+\end{table}
+
 \hypertarget{goal-3-compare-classification-estimates-to-prior-literature}{%

### 2023-02-17 manuscript.Rmd
organize code chunks

--- 2023-02-17 manuscript.Rmd
@@
 <!-- compiled_agreement, generated from analysis-part2 -->
-```{r}
+
+```{r part2overall, warning=FALSE, fig.cap="Overall agreement between human-coded body position and model-predicted body position in the long-delay period.
+Agreement for group models is shown in (A) and agreement for individual models is shown in (B).
+Plots are shown separately for each body position with a reference line that indicates perfect agreement; each point in a plot represent data for a single participant.", fig.align="center", fig.width=7, fig.height=6.5}
+
+# PART 2 AGREEMENT
+pt2_group <- read_csv("data/compiled_agreement_position.csv") %>% 
+  filter(nap_period == 0 & exclude_period == 0 & !is.na(code) & !is.na(pos)) %>% 
+         mutate(pos = factor(pos, levels = pos_levels))
+pt2_split <- read_csv("data/compiled_agreement_split.csv") %>% 
+  filter(nap_period == 0 & exclude_period == 0 & !is.na(code) & !is.na(pos)) %>% 
+         mutate(pos = factor(pos, levels = pos_levels))
+
+outlier_files <- c("107/3", "104/1", "106/1")
+filter_outliers <- . %>% filter(!(file %in% outlier_files))
+
 # Count up position totals and join into an agreement data frame
@@
 
+agree <- map(list("group" = pt2_group, "split" = pt2_split), calculate_agreement)
+
 agreement_scatter <- function(ds) {
@@
 
-# difference_plot <- function(ds) {
-#   ds %>% mutate(diff = prop_human - prop_model) %>% 
-#   ggplot(aes(x = pos, y = diff, color = pos)) +
-#   stat_summary() + 
-#   scale_fill_manual(values = pal, guide = "none") + 
-#   ylab("Prediction Difference (%)") + ylim(-.5, .5) + xlab("")
-# }
-```
-
-```{r part2overall, warning=FALSE, fig.cap="Overall agreement between human-coded body position and model-predicted body position in the long-delay period.
-Agreement for group models is shown in (A) and agreement for individual models is shown in (B).
-Plots are shown separately for each body position with a reference line that indicates perfect agreement; each point in a plot represent data for a single participant.", fig.align="center", fig.width=7, fig.height=6.5}
-# GROUP MODEL PART 2 AGREEMENT
-pt2_group <- read_csv("data/compiled_agreement_position.csv") %>% 
-  filter(nap_period == 0 & exclude_period == 0 & !is.na(code) & !is.na(pos)) %>% 
-         mutate(pos = factor(pos, levels = pos_levels))
-pt2_split <- read_csv("data/compiled_agreement_split.csv") %>% 
-  filter(nap_period == 0 & exclude_period == 0 & !is.na(code) & !is.na(pos)) %>% 
-         mutate(pos = factor(pos, levels = pos_levels))
-
-agree <- map(list("group" = pt2_group, "split" = pt2_split), calculate_agreement)
-
 agree_plots <- map(agree, agreement_scatter)
@@
 pt2corr_with_outliers <- map_dfc(agree, ~ .x %>% group_by(pos) %>% cor_test(vars = c("prop_model", "prop_human")) %>% select(pos, cor)) %>% select(-`pos...3`)
-pt2corr_without_outliers <- map_dfc(agree, ~ .x %>% filter(file != "107/3", file != "104/1") %>% group_by(pos) %>% cor_test(vars = c("prop_model", "prop_human")) %>% select(pos, cor)) %>% select(-`pos...1`,-`pos...3`)
+pt2corr_without_outliers <- map_dfc(agree, ~ .x %>% filter_outliers %>% group_by(pos) %>% cor_test(vars = c("prop_model", "prop_human")) %>% select(pos, cor)) %>% select(-`pos...1`,-`pos...3`)
 pt2corr <- bind_cols(pt2corr_with_outliers, pt2corr_without_outliers)
@@
 split_agree_cor <- cor_test(agree$split, vars = c("prop_model", "prop_human")) %>% pull(cor)
-group_agree_cor_noout <- agree$group %>% filter(file != "107/3", file != "104/1") %>% 
+group_agree_cor_noout <- agree$group %>% filter_outliers %>% 
   cor_test(vars = c("prop_model", "prop_human")) %>% pull(cor)
-split_agree_cor_noout <- agree$split %>% filter(file != "107/3", file != "104/1") %>%
+split_agree_cor_noout <- agree$split %>% filter_outliers %>%
   cor_test(vars = c("prop_model", "prop_human")) %>% pull(cor)
@@
           midrules = 5)
-
-# difference_plots <- map(agree, difference_plot)
-# difference_plots <- map2(difference_plots, c("A. Group Models", "B. Individual Models"), ~ .x + ggtitle(.y))
-# difference_plots[[1]] + difference_plots[[2]]
 
@@
 agree_bins <- map(list("group" = pt2_group, "split" = pt2_split), calculate_agreement_bins)
+
+#Bins
+pt2corr_with_outliers <- map_dfc(agree_bins, ~ .x %>% group_by(pos) %>% cor_test(vars = c("prop_model", "prop_human")) %>% select(pos, cor)) %>% select(-`pos...3`)
+pt2corr_without_outliers <- map_dfc(agree_bins, ~ .x %>% filter_outliers %>% group_by(pos) %>% cor_test(vars = c("prop_model", "prop_human")) %>% select(pos, cor)) %>% select(-`pos...1`,-`pos...3`)
+pt2corr <- bind_cols(pt2corr_with_outliers, pt2corr_without_outliers)
+rm(pt2corr_with_outliers)
+rm(pt2corr_without_outliers)
+
+group_agree_cor <- cor_test(agree_bins$group, vars = c("prop_model", "prop_human")) %>% pull(cor)
+split_agree_cor <- cor_test(agree_bins$split, vars = c("prop_model", "prop_human")) %>% pull(cor)
+group_agree_cor_noout <- agree_bins$group %>% filter_outliers %>% 
+  cor_test(vars = c("prop_model", "prop_human")) %>% pull(cor)
+split_agree_cor_noout <- agree_bins$split %>% filter_outliers %>%
+  cor_test(vars = c("prop_model", "prop_human")) %>% pull(cor)
+overall_row <- tibble("Overall", group_agree_cor, split_agree_cor, group_agree_cor_noout, split_agree_cor_noout)
+colnames(overall_row) <- colnames(pt2corr)
+pt2corr <- bind_rows(pt2corr, overall_row)
+apa_table(pt2corr, col.names = c("Position", "Group", "Individual", "Group", "Individual"), col_spanners = list(`With Outliers` = c(2, 3), `Without Outliers` = c(4, 5)),
+          midrules = 5)
+
+difference_plot <- function(ds) {
+  ds <- mutate(ds, diff = prop_human - prop_model)
+  ggplot(filter_outliers(ds), aes(x = pos, y = diff, color = pos, group = file)) +
+  stat_summary(position = position_dodge(.4), alpha = .7) +
+  stat_summary(data = filter(ds, file %in% c("107/3", "104/1", "106/1")), position = position_dodge(.4), alpha = .7, color = "darkgray") +
+  scale_color_manual(values = pal, guide = "none") +
+  coord_cartesian(ylim = c(-1, 1)) + xlab("") + 
+  scale_y_continuous(name = "Prediction Difference (min)", breaks = seq(-1, 1, .2), labels = seq(-10, 10, 2), minor_breaks = waiver())
+}
+difference_plots <- map(agree_bins, ~ difference_plot(.x))
+difference_plots <- map2(difference_plots, c("A. Group Models", "B. Individual Models"), ~ .x + ggtitle(.y))
+difference_plots[[1]] + difference_plots[[2]]
 

### 2023-02-17 manuscript.tex
organize code chunks

--- 2023-02-16 manuscript.tex
@@
 
+\begin{table}[tbp]
+
+\begin{center}
+\begin{threeparttable}
+
+\caption{\label{tab:pt2overalltable}}
+
+\begin{tabular}{lllll}
+\toprule
+ & \multicolumn{2}{c}{With Outliers} & \multicolumn{2}{c}{Without Outliers} \\
+\cmidrule(r){2-3} \cmidrule(r){4-5}
+Position & Group & Individual & Group & Individual\\
+\midrule
+Held & -0.02 & 0.16 & 0.55 & 0.63\\
+Prone & 0.97 & 0.83 & 0.97 & 0.84\\
+Sitting & 0.72 & 0.93 & 0.91 & 0.97\\
+Supine & 0.84 & 0.93 & 0.94 & 0.97\\
+Upright & 0.84 & 0.93 & 0.99 & 0.94\\ \midrule
+Overall & 0.79 & 0.90 & 0.95 & 0.96\\
+\bottomrule
+\end{tabular}
+
+\end{threeparttable}
+\end{center}
+
+\end{table}
+
+\begin{table}[tbp]
+
+\begin{center}
+\begin{threeparttable}
+
+\caption{\label{tab:unnamed-chunk-2}}
+
+\begin{tabular}{lllll}
+\toprule
+ & \multicolumn{2}{c}{With Outliers} & \multicolumn{2}{c}{Without Outliers} \\
+\cmidrule(r){2-3} \cmidrule(r){4-5}
+Position & Group & Individual & Group & Individual\\
+\midrule
+Held & 0.45 & 0.44 & 0.57 & 0.56\\
+Prone & 0.96 & 0.81 & 0.96 & 0.88\\
+Sitting & 0.72 & 0.93 & 0.91 & 0.93\\
+Supine & 0.75 & 0.93 & 0.90 & 0.94\\
+Upright & 0.93 & 0.95 & 0.97 & 0.95\\ \midrule
+Overall & 0.80 & 0.92 & 0.92 & 0.94\\
+\bottomrule
+\end{tabular}
+
+\end{threeparttable}
+\end{center}
+
+\end{table}
+
+\begin{verbatim}
+## Warning: Removed 10 rows containing missing values (`geom_segment()`).
+\end{verbatim}
+
+\begin{verbatim}
+## Warning: Removed 5 rows containing missing values (`geom_segment()`).
+\end{verbatim}
+
+\begin{verbatim}
+## Warning: Removed 10 rows containing missing values (`geom_segment()`).
+\end{verbatim}
+
+\begin{verbatim}
+## Warning: Removed 5 rows containing missing values (`geom_segment()`).
+\end{verbatim}
+
+\includegraphics{manuscript_files/figure-latex/unnamed-chunk-2-1.pdf}
+
 \hypertarget{goal-3-compare-classification-estimates-to-prior-literature}{%

### 2023-02-17 manuscript.Rmd
marking outliers

@@
 agreement_scatter <- function(ds) {
-  ggplot(ds) +
-  geom_point(aes(x = prop_human, y = prop_model, fill = pos), shape = 21, alpha = .5, size = 4) + 
+  ggplot() +
+  geom_point(data = filter(ds, file %in% outlier_files), mapping = aes(x = prop_human, y = prop_model, shape = file, fill = pos), color = "#333333", size = 4, alpha = .5) + 
+  geom_point(data = filter_outliers(ds), mapping = aes(x = prop_human, y = prop_model, fill = pos), shape = 21, alpha = .5, size = 4) + 
   geom_abline(intercept = 0, slope = 1) + 
   scale_fill_manual(values = pal, guide = "none") + 
+  scale_shape_discrete(guide = "none") + 
   scale_x_continuous(breaks = c(0,.5,1), labels = c(0, 50, 100), limits = c(-.07, 1.07), name = "Actual Prevalence (%)") + 
@@
 
-```{r }
+```{r part2bins, warning=FALSE, fig.cap="Prediction performance (difference in minutes between human-coded and model-predicted body position) for 10-minute bins in the long delay period.
+Each point shows the mean and SE for a single participant for each body position, summarizing the prediction difference for each of their 10-minute bins.
+Points falling within the gray shaded region indicate that average prediction errors were less than 1 minute.
+Performance is plotted separately for (A) group models and (B) individual models.
+The three outlier participants are plotted in dark gray, with a different shape marking each individual.", fig.align="center", fig.width=8, fig.height=4.5}
 calculate_agreement_bins <- function(ds) {
@@
   agree <- agree %>% mutate(prop_human = ifelse(is.na(prop_human), 0, prop_human),
-                            prop_model = ifelse(is.na(prop_model), 0, prop_model))
+                            prop_model = ifelse(is.na(prop_model), 0, prop_model),
+                            pos - factor(pos, levels = pos_levels))
   agree <- complete(agree, nesting(file, bin, num_bins), pos, fill = list(prop_human = 0, prop_model = 0))
@@
   ggplot(filter_outliers(ds), aes(x = pos, y = diff, color = pos, group = file)) +
+  geom_rect(aes(xmin = "Held", xmax = "Upright",ymin = -.1, ymax = .1), fill = "#e5e5e5", color = "#e5e5e5") + 
   stat_summary(position = position_dodge(.4), alpha = .7) +
-  stat_summary(data = filter(ds, file %in% c("107/3", "104/1", "106/1")), position = position_dodge(.4), alpha = .7, color = "darkgray") +
+  stat_summary(data = filter(ds, file %in% outlier_files), position = position_dodge(.4), aes(shape = file), alpha = .7, color = "#333333") +
   scale_color_manual(values = pal, guide = "none") +
+  scale_shape_discrete(guide = "none") + 
   coord_cartesian(ylim = c(-1, 1)) + xlab("") + 
-  scale_y_continuous(name = "Prediction Difference (min)", breaks = seq(-1, 1, .2), labels = seq(-10, 10, 2), minor_breaks = waiver())
+  scale_y_continuous(name = "Prediction Difference (min)", breaks = seq(-1, 1, .2), labels = seq(-10, 10, 2), minor_breaks = waiver()) + 
+    theme(axis.text.x = element_text(size = 13, color = "black"))
 }
@@
 difference_plots[[1]] + difference_plots[[2]]
-
-# individual_scatter <- function(ds, pos_select) {
-#   ds %>% filter(pos == pos_select) %>% ggplot() +
-#   geom_abline(intercept = 0, slope = 1) + facet_wrap("file") + 
-#   geom_point(aes(x = prop_human, y = prop_model, fill = pos), shape = 21, alpha = .5, size = 4) + 
-#   scale_fill_manual(values = pal, guide = "none") + 
-#   scale_x_continuous(breaks = c(0,.5,1), labels = c(0, 50, 100), limits = c(-.07, 1.07), name = "Actual Prevalence (%)") + 
-#   scale_y_continuous(breaks = c(0,.5,1), labels = c(0, 50, 100), limits = c(-.07, 1.07), name = "Predicted Prevalence (%)") + 
-#   theme(panel.spacing.x = unit(1.5, "lines")) +
-#   theme(panel.spacing.y = unit(.75, "lines")) + 
-#   theme(strip.background = element_blank(), strip.text.x = element_blank())
-# }
-
-# individual_scatter <- function(ds) {
-#   ds %>% ggplot() +
-#   geom_abline(intercept = 0, slope = 1) + facet_grid(c("file","pos")) + 
-#   geom_point(aes(x = prop_human, y = prop_model, fill = pos), shape = 21, alpha = .5, size = 3) + 
-#   scale_fill_manual(values = pal, guide = "none") + 
-#   scale_x_continuous(breaks = c(0,.5,1), labels = c(0, 50, 100), limits = c(-.07, 1.07), name = "Actual Prevalence (%)") + 
-#   scale_y_continuous(breaks = c(0,.5,1), labels = c(0, 50, 100), limits = c(-.07, 1.07), name = "Predicted Prevalence (%)") + 
-#   theme(panel.spacing.x = unit(.5, "lines")) +
-#   theme(strip.background = element_blank(), strip.text.x = element_blank()) + 
-#    theme(axis.text.x = element_blank(), axis.text.y = element_blank())
-# }
-# individual_scatter(agree_bins$group)
-
-# agree_plots_bins <- map(agree_bins, ~ individual_scatter(.x, "Prone"))
-# agree_plots_bins <- map2(agree_plots_bins, c("A. Group Models", "B. Individual Models"), ~ .x + ggtitle(.y))
-# agree_plots_bins[[1]] / agree_plots_bins[[2]]
-# (individual_scatter(agree_bins$group, "Sitting") + individual_scatter(agree_bins$group, "Upright"))/
-#   (individual_scatter(agree_bins$group, "Prone") + individual_scatter(agree_bins$group, "Supine"))
-#
 

### 2023-02-17 manuscript.tex
marking outliers

@@
 
-\caption{\label{tab:unnamed-chunk-2}}
+\caption{\label{tab:part2bins}}
 
@@
 
-\begin{verbatim}
-## Warning: Removed 10 rows containing missing values (`geom_segment()`).
-\end{verbatim}
+\begin{figure}
 
-\begin{verbatim}
-## Warning: Removed 5 rows containing missing values (`geom_segment()`).
-\end{verbatim}
+{\centering \includegraphics{manuscript_files/figure-latex/part2bins-1}
 
-\begin{verbatim}
-## Warning: Removed 10 rows containing missing values (`geom_segment()`).
-\end{verbatim}
+}
 
-\begin{verbatim}
-## Warning: Removed 5 rows containing missing values (`geom_segment()`).
-\end{verbatim}
-
-\includegraphics{manuscript_files/figure-latex/unnamed-chunk-2-1.pdf}
+\caption{Prediction performance (difference in minutes between human-coded and model-predicted body position) for 10-minute bins in the long delay period.
+Each point shows the mean and SE for a single participant for each body position, summarizing the prediction difference for each of their 10-minute bins.
+Points falling within the gray shaded region indicate that prediction errors were less than 1 minute.
+Performance is plotted separately for (A) group models and (B) individual models.
+The three outlier participants are plotted in dark gray, with a different shape marking each individual.}\label{fig:part2bins}
+\end{figure}
 

### 2023-02-18 manuscript.Rmd
age graphs

@@
 library(patchwork)
+library(scales)
 library(tidyverse)
@@
 
+```{r timelines, warning=FALSE, fig.cap="Timelines", fig.align="center", fig.width=7.75, fig.height=8}
+fullday <- read_csv("data/imu-compiled.csv") %>% 
+  mutate(id_uni = factor(id*100+session),
+         group = factor(ifelse(age <= 8, "Younger", "Older")),
+         nap_time = ifelse(is.na(sit_time), nap_time, NA))
+fullday_long <- fullday %>% pivot_longer(cols = nap_time:upright_time, names_to = "position", values_to = "prop") %>% 
+  mutate(Position = factor(position,
+                           levels = c("nap_time","upright_time", "sit_time", "prone_time", "supine_time", "held_time"),
+                           labels = c("Nap","Upright", "Sitting", "Prone", "Supine", "Held"))) %>% 
+  arrange(Position, age)
+
+timelines <- function(ds) {
+  ggplot(ds, aes(x = clock_time_start, y = prop, fill = Position)) + 
+  geom_bar(stat = "identity") + 
+  facet_wrap("id_uni",ncol = 1, strip.position = "left") + 
+  scale_fill_manual(values = c("Nap" = "gray",pal)) + 
+  scale_x_time(breaks = breaks_width("2 hour"), name = "", minor_breaks = waiver()) + 
+  theme(legend.position = "bottom", 
+        axis.ticks.y = element_blank(),
+        axis.text.y = element_blank(),
+        axis.text.x = element_text(size = 10, color = "black", angle = 315, hjust = -.15),
+        strip.background = element_blank(),
+        strip.text.y = element_blank()) + ylab("")
+}
+
+timeline_plots <- map(c("Younger", "Older"), ~ timelines(filter(fullday_long, group == .x)))
+timeline_plots <- map2(timeline_plots, c("A. 4-7 Months", "B. 11-14 Months"), ~ .x + ggtitle(.y))
+timeline_plots[[1]] + timeline_plots[[2]]
+
+```
+
 ## Goal 2: Assess classification accuracy over long recordings
@@
 Agreement for group models is shown in (A) and agreement for individual models is shown in (B).
-Plots are shown separately for each body position with a reference line that indicates perfect agreement; each point in a plot represent data for a single participant.", fig.align="center", fig.width=7, fig.height=6.5}
+Plots are shown separately for each body position with a reference line that indicates perfect agreement; each point in a plot represent data for a single participant.
+The three outlier participants are plotted in dark gray, with a different shape marking each individual.", fig.align="center", fig.width=7, fig.height=6.5}
 
@@
 ```{r, pt2overalltable}
+#By posture
 pt2corr_with_outliers <- map_dfc(agree, ~ .x %>% group_by(pos) %>% cor_test(vars = c("prop_model", "prop_human")) %>% select(pos, cor)) %>% select(-`pos...3`)
@@
 apa_table(pt2corr, col.names = c("Position", "Group", "Individual", "Group", "Individual"), col_spanners = list(`With Outliers` = c(2, 3), `Without Outliers` = c(4, 5)),
-          midrules = 5)
-
+          midrules = 5, caption = "Correlations between human-coded and model-predicted body position durations across the entire long delay period.
+Correlations are provided within each posture and overall, and computed separately using group and individual models with and without outlier participants.")
 ```
@@
 
-#Bins
-pt2corr_with_outliers <- map_dfc(agree_bins, ~ .x %>% group_by(pos) %>% cor_test(vars = c("prop_model", "prop_human")) %>% select(pos, cor)) %>% select(-`pos...3`)
-pt2corr_without_outliers <- map_dfc(agree_bins, ~ .x %>% filter_outliers %>% group_by(pos) %>% cor_test(vars = c("prop_model", "prop_human")) %>% select(pos, cor)) %>% select(-`pos...1`,-`pos...3`)
-pt2corr <- bind_cols(pt2corr_with_outliers, pt2corr_without_outliers)
-rm(pt2corr_with_outliers)
-rm(pt2corr_without_outliers)
-
-group_agree_cor <- cor_test(agree_bins$group, vars = c("prop_model", "prop_human")) %>% pull(cor)
-split_agree_cor <- cor_test(agree_bins$split, vars = c("prop_model", "prop_human")) %>% pull(cor)
-group_agree_cor_noout <- agree_bins$group %>% filter_outliers %>% 
-  cor_test(vars = c("prop_model", "prop_human")) %>% pull(cor)
-split_agree_cor_noout <- agree_bins$split %>% filter_outliers %>%
-  cor_test(vars = c("prop_model", "prop_human")) %>% pull(cor)
-overall_row <- tibble("Overall", group_agree_cor, split_agree_cor, group_agree_cor_noout, split_agree_cor_noout)
-colnames(overall_row) <- colnames(pt2corr)
-pt2corr <- bind_rows(pt2corr, overall_row)
-apa_table(pt2corr, col.names = c("Position", "Group", "Individual", "Group", "Individual"), col_spanners = list(`With Outliers` = c(2, 3), `Without Outliers` = c(4, 5)),
-          midrules = 5)
-
 difference_plot <- function(ds) {
@@
 difference_plots[[1]] + difference_plots[[2]]
+```
 
+```{r pt2binstable}
+#Bins
+pt2corr_with_outliers <- map_dfc(agree_bins, ~ .x %>% group_by(pos) %>% cor_test(vars = c("prop_model", "prop_human")) %>% select(pos, cor)) %>% select(-`pos...3`)
+pt2corr_without_outliers <- map_dfc(agree_bins, ~ .x %>% filter_outliers %>% group_by(pos) %>% cor_test(vars = c("prop_model", "prop_human")) %>% select(pos, cor)) %>% select(-`pos...1`,-`pos...3`)
+pt2corr_bins <- bind_cols(pt2corr_with_outliers, pt2corr_without_outliers)
+rm(pt2corr_with_outliers)
+rm(pt2corr_without_outliers)
+
+group_agree_cor <- cor_test(agree_bins$group, vars = c("prop_model", "prop_human")) %>% pull(cor)
+split_agree_cor <- cor_test(agree_bins$split, vars = c("prop_model", "prop_human")) %>% pull(cor)
+group_agree_cor_noout <- agree_bins$group %>% filter_outliers %>% 
+  cor_test(vars = c("prop_model", "prop_human")) %>% pull(cor)
+split_agree_cor_noout <- agree_bins$split %>% filter_outliers %>%
+  cor_test(vars = c("prop_model", "prop_human")) %>% pull(cor)
+overall_row <- tibble("Overall", group_agree_cor, split_agree_cor, group_agree_cor_noout, split_agree_cor_noout)
+colnames(overall_row) <- colnames(pt2corr_bins)
+pt2corr_bins <- bind_rows(pt2corr_bins, overall_row)
+rm(overall_row)
+apa_table(pt2corr_bins, col.names = c("Position", "Group", "Individual", "Group", "Individual"), col_spanners = list(`With Outliers` = c(2, 3), `Without Outliers` = c(4, 5)),
+          midrules = 5, caption = "Correlations between human-coded and model-predicted body position durations using 10-minute bins during the long delay period.
+Correlations are provided within each posture and overall, and computed separately using group and individual models with and without outlier participants.")
 ```
@@
 <!-- imu-compiled.csv, generated from analysis-lena-imu/compile_lena_imu.R -->
+
+```{r age, warning=FALSE, fig.cap="Age trends", fig.align="center", fig.width=5, fig.height=8}
+fullday <- read_csv("data/imu-compiled.csv") %>% 
+  mutate(id_uni = factor(id*100+session),
+         group = factor(ifelse(age <= 8, "Younger", "Older")),
+         nap_time = ifelse(is.na(sit_time), nap_time, NA))
+fullday_long <- fullday %>% pivot_longer(cols = sit_time:upright_time, names_to = "position", values_to = "prop") %>% 
+  mutate(position = factor(position,
+                           levels = c("upright_time", "sit_time", "prone_time", "supine_time", "held_time"),
+                           labels = c("Upright", "Sitting", "Prone", "Supine", "Held"))) %>% 
+  arrange(position, age)
+
+outlier_ids <- c(10701, 10401, 10601)
+fullday %>% cor_mat(age, sit_time:upright_time) %>% cor_mark_significant()
+fullday %>% filter(!(id_uni %in% outlier_ids)) %>% cor_mat(age, sit_time:upright_time) %>% cor_mark_significant()
+
+fullday_long %>% drop_na(prop) %>% 
+  ggplot() + facet_wrap("position", as.table = T, ncol = 2) + 
+  geom_smooth(aes(x = age, y = prop), color = "darkgray", method = "lm", se = T) + 
+  stat_summary(fun = "mean", mapping = aes(x = age, y = prop, group = id_uni, fill = position), shape = 21, size = .75) + 
+  scale_fill_manual(values = pal, guide = "none") + 
+  scale_x_continuous(name = "Age (months)", breaks = seq(3, 15, 3), limits = c(3,15)) + 
+  ylim(0,1) + ylab("Proportion of waking day")
+
+fullday_long %>% group_by(position, group) %>% get_summary_stats(prop)
+
+```
 

### 2023-02-18 manuscript.tex
age graphs

@@
 Agreement for group models is shown in (A) and agreement for individual models is shown in (B).
-Plots are shown separately for each body position with a reference line that indicates perfect agreement; each point in a plot represent data for a single participant.}\label{fig:part2overall}
+Plots are shown separately for each body position with a reference line that indicates perfect agreement; each point in a plot represent data for a single participant.
+The three outlier participants are plotted in dark gray, with a different shape marking each individual.}\label{fig:part2overall}
 \end{figure}
@@
 
-\caption{\label{tab:pt2overalltable}}
+\caption{\label{tab:pt2overalltable}Correlations between human-coded and model-predicted body position durations across the entire long delay period.
+Correlations are provided within each posture and overall, and computed separately using group and individual models with and without outlier participants.}
 
@@
 
+\begin{figure}
+
+{\centering \includegraphics{manuscript_files/figure-latex/part2bins-1}
+
+}
+
+\caption{Prediction performance (difference in minutes between human-coded and model-predicted body position) for 10-minute bins in the long delay period.
+Each point shows the mean and SE for a single participant for each body position, summarizing the prediction difference for each of their 10-minute bins.
+Points falling within the gray shaded region indicate that average prediction errors were less than 1 minute.
+Performance is plotted separately for (A) group models and (B) individual models.
+The three outlier participants are plotted in dark gray, with a different shape marking each individual.}\label{fig:part2bins}
+\end{figure}
+
 \begin{table}[tbp]
@@
 
-\caption{\label{tab:part2bins}}
+\caption{\label{tab:pt2binstable}Correlations between human-coded and model-predicted body position durations using 10-minute bins during the long delay period.
+Correlations are provided within each posture and overall, and computed separately using group and individual models with and without outlier participants.}
 
@@
 
+\hypertarget{goal-3-compare-classification-estimates-to-prior-literature}{%
+\subsection{Goal 3: Compare classification estimates to prior literature}\label{goal-3-compare-classification-estimates-to-prior-literature}}
+
 \begin{figure}
 
-{\centering \includegraphics{manuscript_files/figure-latex/part2bins-1}
+{\centering \includegraphics{manuscript_files/figure-latex/timelines-1}
 
@@
 
-\caption{Prediction performance (difference in minutes between human-coded and model-predicted body position) for 10-minute bins in the long delay period.
-Each point shows the mean and SE for a single participant for each body position, summarizing the prediction difference for each of their 10-minute bins.
-Points falling within the gray shaded region indicate that prediction errors were less than 1 minute.
-Performance is plotted separately for (A) group models and (B) individual models.
-The three outlier participants are plotted in dark gray, with a different shape marking each individual.}\label{fig:part2bins}
+\caption{Timelines}\label{fig:timelines}
 \end{figure}
-
-\hypertarget{goal-3-compare-classification-estimates-to-prior-literature}{%
-\subsection{Goal 3: Compare classification estimates to prior literature}\label{goal-3-compare-classification-estimates-to-prior-literature}}
 

### 2023-02-18 manuscript.Rmd
paneled age graph

@@
 
-```{r age, warning=FALSE, fig.cap="Age trends", fig.align="center", fig.width=5, fig.height=8}
-fullday <- read_csv("data/imu-compiled.csv") %>% 
+```{r age, warning=FALSE, fig.cap="Age trends", fig.align="center", fig.width=7.5, fig.height=7}
+fullday_group <- read_csv("data/imu-compiled.csv") %>% 
   mutate(id_uni = factor(id*100+session),
@@
          nap_time = ifelse(is.na(sit_time), nap_time, NA))
-fullday_long <- fullday %>% pivot_longer(cols = sit_time:upright_time, names_to = "position", values_to = "prop") %>% 
+fullday_split <- read_csv("data/imu-compiled-split.csv") %>% 
+  mutate(id_uni = factor(id*100+session),
+         group = factor(ifelse(age <= 8, "Younger", "Older")),
+         nap_time = ifelse(is.na(sit_time), nap_time, NA))
+fullday <- list("group" = fullday_group, "split" = fullday_split)
+
+fullday_long <- function(ds) {
+  ds %>% pivot_longer(cols = sit_time:upright_time, names_to = "position", values_to = "prop") %>% 
   mutate(position = factor(position,
@@
   arrange(position, age)
+}
 
-outlier_ids <- c(10701, 10401, 10601)
-fullday %>% cor_mat(age, sit_time:upright_time) %>% cor_mark_significant()
-fullday %>% filter(!(id_uni %in% outlier_ids)) %>% cor_mat(age, sit_time:upright_time) %>% cor_mark_significant()
-
-fullday_long %>% drop_na(prop) %>% 
+age_scatter <- function(ds) {
+  ds %>% drop_na(prop) %>% 
   ggplot() + facet_wrap("position", as.table = T, ncol = 2) + 
-  geom_smooth(aes(x = age, y = prop), color = "darkgray", method = "lm", se = T) + 
-  stat_summary(fun = "mean", mapping = aes(x = age, y = prop, group = id_uni, fill = position), shape = 21, size = .75) + 
+  geom_smooth(aes(x = age, y = prop), color = "#191919", method = "lm", se = T) + 
+  stat_summary(fun = "mean", mapping = aes(x = age, y = prop, group = id_uni, fill = position), shape = 21, size = .75, alpha = .7) + 
   scale_fill_manual(values = pal, guide = "none") + 
@@
   ylim(0,1) + ylab("Proportion of waking day")
+}
 
-fullday_long %>% group_by(position, group) %>% get_summary_stats(prop)
+age_plots <- map(fullday, fullday_long) %>% map(age_scatter)
+age_plots <- map2(age_plots, c("A. Group Models", "B. Individual Models"), ~ .x + ggtitle(.y))
+age_plots[[1]] + age_plots[[2]]
+
+# outlier_ids <- c(10701, 10401, 10601)
+# fullday %>% cor_mat(age, sit_time:upright_time) %>% cor_mark_significant()
+# fullday %>% filter(!(id_uni %in% outlier_ids)) %>% cor_mat(age, sit_time:upright_time) %>% cor_mark_significant()
+# 
+# fullday_long %>% group_by(position, group) %>% get_summary_stats(prop)
 

### 2023-02-18 manuscript.Rmd
updated facet plots

@@
   scale_x_continuous(name = "Age (months)", breaks = seq(3, 15, 3), limits = c(3,15)) + 
-  ylim(0,1) + ylab("Proportion of waking day")
+  ylim(0,1) + ylab("Proportion of waking day") + 
+  theme(panel.spacing.x = unit(.5, "lines")) +
+  theme(panel.spacing.y = unit(1, "lines")) + 
+  theme(strip.background = element_blank(), strip.text.x = element_text(size = 16, face = "bold"))
 }

### 2023-02-18 manuscript.Rmd
metrics plot

@@
 library(scales)
+library(ggforce)
 library(tidyverse)
@@
 <!-- group_split_metrics, generated from group_split_comparison in this proj, stable copy in data/ -->
-```{r}
-metrics <- read_csv("data/group_split_metrics.csv")
+```{r metrics, warning=FALSE, fig.cap="Metrics", fig.align="center", fig.width=5, fig.height=8}
+metrics <- read_csv("data/group_split_metrics.csv") %>% 
+  filter(model %in% c("group", "split")) %>% 
+  mutate(Model = factor(model, levels = c("group", "split"), labels = c("Group", "Individual")))
+
+metrics_long <- metrics %>% pivot_longer(cols = `Overall Accuracy`:Kappa, 
+                                         names_to = "metric",
+                                         values_to = "value")
+
+metric_plot <- function(ds, met) {
+  ds %>% filter(metric == met) %>% 
+    ggplot(aes(x = Model, y = value, shape = Model)) + 
+    geom_sina(alpha = .7, fill = "orange", size = 4, maxwidth = .5) +
+    stat_summary(fun = "mean", mapping = aes(x = Model, y = value), shape = "-", fill = "black", size = 4) + 
+    scale_shape_manual(guide = "none", values = c(21,23)) + 
+    ylim(0, 1) + ylab(met) + xlab("") + 
+    theme(plot.title.position = "plot",
+          axis.title.y = element_text(size = 13),
+          axis.text.y = element_text(size = 11),
+          axis.text.x = element_text(size = 13))
+}
+metric_labels <- c("Overall Accuracy", "Balanced Accuracy", "F1", "Sensitivity", "Pos Pred Value", "Kappa")
+title_labels <- c("A. Overall Accuracy", "B. Balanced Accuracy", "C. F1 Score", "D. Sensitivity", "E. Pos. Pred. Value", "F. Kappa")
+metric_plots <- map(metric_labels, ~ metric_plot(metrics_long, .x)) %>% map2(., title_labels, ~.x + ggtitle(.y))
+(metric_plots[[1]] + metric_plots[[2]])/(metric_plots[[3]]+metric_plots[[4]])/(metric_plots[[5]] + metric_plots[[6]])
 

### 2023-02-19 manuscript.Rmd
figured out outliers
fixed x-axis on timeline plot

@@
 library(ggforce)
+library(hms)
 library(tidyverse)
@@
 
+#Worst three Kappa individuals are 9, 111-4, and 120-2
+#Worst three Kappa groups are 9, 110-4
+#Worst three Overall Accuracy (grouped) are 104-1, 106-1, 108-2 (first two are pt 2 outliers)
+
 ```
@@
 timelines <- function(ds) {
+  lims <- as_hms(c('09:00:00', '21:59:00'))
+  breaks = as_hms(c('09:00:00', '10:00:00', '11:00:00', '12:00:00', '13:00:00', '14:00:00', '15:00:00', '16:00:00', '17:00:00', '18:00:00', '19:00:00', '20:00:00', '21:00:00'))
   ggplot(ds, aes(x = clock_time_start, y = prop, fill = Position)) + 
@@
   scale_fill_manual(values = c("Nap" = "gray",pal)) + 
-  scale_x_time(breaks = breaks_width("2 hour"), name = "", minor_breaks = waiver()) + 
+  scale_x_time(breaks = breaks_width("4 hour"), name = "", limits = lims) + 
   theme(legend.position = "bottom", 
@@
         axis.text.y = element_blank(),
-        axis.text.x = element_text(size = 10, color = "black", angle = 315, hjust = -.15),
+        axis.text.x = element_text(size = 10, color = "black"),
         strip.background = element_blank(),

### 2023-02-19 manuscript.Rmd
age summaries

@@
   ds %>% filter(metric == met) %>% 
-    ggplot(aes(x = Model, y = value, shape = Model)) + 
-    geom_sina(alpha = .7, fill = "orange", size = 4, maxwidth = .5) +
+    ggplot(aes(x = Model, y = value)) + 
+    geom_sina(alpha = .7, fill = "aquamarine3", size = 4, maxwidth = .5, shape = 21) +
     stat_summary(fun = "mean", mapping = aes(x = Model, y = value), shape = "-", fill = "black", size = 4) + 
-    scale_shape_manual(guide = "none", values = c(21,23)) + 
     ylim(0, 1) + ylab(met) + xlab("") + 
@@
 #Worst three Overall Accuracy (grouped) are 104-1, 106-1, 108-2 (first two are pt 2 outliers)
-
-```
-
-```{r timelines, warning=FALSE, fig.cap="Timelines", fig.align="center", fig.width=7.75, fig.height=8}
-fullday <- read_csv("data/imu-compiled.csv") %>% 
-  mutate(id_uni = factor(id*100+session),
-         group = factor(ifelse(age <= 8, "Younger", "Older")),
-         nap_time = ifelse(is.na(sit_time), nap_time, NA))
-fullday_long <- fullday %>% pivot_longer(cols = nap_time:upright_time, names_to = "position", values_to = "prop") %>% 
-  mutate(Position = factor(position,
-                           levels = c("nap_time","upright_time", "sit_time", "prone_time", "supine_time", "held_time"),
-                           labels = c("Nap","Upright", "Sitting", "Prone", "Supine", "Held"))) %>% 
-  arrange(Position, age)
-
-timelines <- function(ds) {
-  lims <- as_hms(c('09:00:00', '21:59:00'))
-  breaks = as_hms(c('09:00:00', '10:00:00', '11:00:00', '12:00:00', '13:00:00', '14:00:00', '15:00:00', '16:00:00', '17:00:00', '18:00:00', '19:00:00', '20:00:00', '21:00:00'))
-  ggplot(ds, aes(x = clock_time_start, y = prop, fill = Position)) + 
-  geom_bar(stat = "identity") + 
-  facet_wrap("id_uni",ncol = 1, strip.position = "left") + 
-  scale_fill_manual(values = c("Nap" = "gray",pal)) + 
-  scale_x_time(breaks = breaks_width("4 hour"), name = "", limits = lims) + 
-  theme(legend.position = "bottom", 
-        axis.ticks.y = element_blank(),
-        axis.text.y = element_blank(),
-        axis.text.x = element_text(size = 10, color = "black"),
-        strip.background = element_blank(),
-        strip.text.y = element_blank()) + ylab("")
-}
-
-timeline_plots <- map(c("Younger", "Older"), ~ timelines(filter(fullday_long, group == .x)))
-timeline_plots <- map2(timeline_plots, c("A. 4-7 Months", "B. 11-14 Months"), ~ .x + ggtitle(.y))
-timeline_plots[[1]] + timeline_plots[[2]]
 
@@
   ggplot() +
-  geom_point(data = filter(ds, file %in% outlier_files), mapping = aes(x = prop_human, y = prop_model, shape = file, fill = pos), color = "#333333", size = 4, alpha = .5) + 
+  geom_point(data = filter(ds, file %in% outlier_files), mapping = aes(x = prop_human, y = prop_model, shape = file, fill = pos), fill = "#333333", size = 4, alpha = .5) + 
   geom_point(data = filter_outliers(ds), mapping = aes(x = prop_human, y = prop_model, fill = pos), shape = 21, alpha = .5, size = 4) + 
@@
   scale_fill_manual(values = pal, guide = "none") + 
-  scale_shape_discrete(guide = "none") + 
+  scale_shape_manual(guide = "none", values = 22:24) + 
   scale_x_continuous(breaks = c(0,.5,1), labels = c(0, 50, 100), limits = c(-.07, 1.07), name = "Actual Prevalence (%)") + 
@@
   ds <- mutate(ds, diff = prop_human - prop_model)
-  ggplot(filter_outliers(ds), aes(x = pos, y = diff, color = pos, group = file)) +
-  geom_rect(aes(xmin = "Held", xmax = "Upright",ymin = -.1, ymax = .1), fill = "#e5e5e5", color = "#e5e5e5") + 
-  stat_summary(position = position_dodge(.4), alpha = .7) +
-  stat_summary(data = filter(ds, file %in% outlier_files), position = position_dodge(.4), aes(shape = file), alpha = .7, color = "#333333") +
-  scale_color_manual(values = pal, guide = "none") +
-  scale_shape_discrete(guide = "none") + 
+  ggplot(filter_outliers(ds), aes(x = pos, y = diff, fill = pos, group = file)) +
+  annotate("rect", xmin = -Inf, xmax = Inf, ymin = -.1, ymax = .1, fill = "darkgray", color = "darkgray") + 
+  #geom_rect(aes(xmin = "Held", xmax = "Upright",ymin = -.1, ymax = .1), fill = "#e5e5e5", color = "#e5e5e5") + 
+  stat_summary(data = filter(ds, file %in% outlier_files), position = position_dodge(.4), aes(shape = file), fill = "#333333", size = .75, stroke = .25) +
+  stat_summary(position = position_dodge(.4), shape = 21, size = .75, stroke = .25) +
+  scale_fill_manual(values = pal, guide = "none") +
+  scale_shape_manual(guide = "none", values = 22:24) + 
   coord_cartesian(ylim = c(-1, 1)) + xlab("") + 
@@
 
-## Goal 3: Compare classification estimates to prior literature
+## Goal 3: Examine wear time and compliance in full-day data collection
+<!-- imu-compiled.csv, generated from analysis-lena-imu/compile_lena_imu.R -->
+
+```{r timelines, warning=FALSE, fig.cap="Timelines", fig.align="center", fig.width=7.75, fig.height=8}
+timeline_data <- read_csv("data/imu-compiled.csv") %>% 
+  mutate(id_uni = factor(id*100+session),
+         group = factor(ifelse(age <= 8, "Younger", "Older")),
+         nap_time = ifelse(is.na(sit_time), nap_time, NA)) %>% 
+  pivot_longer(cols = nap_time:upright_time, names_to = "position", values_to = "prop") %>% 
+  mutate(Position = factor(position,
+                           levels = c("nap_time","upright_time", "sit_time", "prone_time", "supine_time", "held_time"),
+                           labels = c("Nap","Upright", "Sitting", "Prone", "Supine", "Held"))) %>% 
+  arrange(Position, age)
+
+timelines <- function(ds) {
+  lims <- as_hms(c('09:00:00', '21:59:00'))
+  hour_breaks = as_hms(c('09:00:00', '10:00:00', '11:00:00', '12:00:00', '13:00:00', '14:00:00', '15:00:00', '16:00:00', '17:00:00', '18:00:00', '19:00:00', '20:00:00', '21:00:00'))
+  label_breaks = c("9am","","11am","","1pm","","3pm","","5pm","","7pm","","9pm")
+  ggplot(ds, aes(x = clock_time_start, y = prop, fill = Position)) + 
+  geom_bar(stat = "identity") + 
+  facet_wrap("id_uni",ncol = 1, strip.position = "left") + 
+  scale_fill_manual(values = c("Nap" = "gray",pal), name = "") + 
+  scale_x_time(breaks = hour_breaks, labels = label_breaks, name = "", limits = lims) + 
+  theme(legend.position = "bottom", 
+        axis.ticks.y = element_blank(),
+        axis.text.y = element_blank(),
+        strip.background = element_blank(),
+        axis.text.x = element_text(size = 11),
+        strip.text.y = element_blank()) + ylab("")
+  #https://ggplot2tor.com/scales/scale_x_time for options
+}
+
+timeline_plots <- map(c("Younger", "Older"), ~ timelines(filter(timeline_data, group == .x)))
+timeline_plots <- map2(timeline_plots, c("A. 4-7 Months", "B. 11-14 Months"), ~ .x + ggtitle(.y))
+timeline_plots[[1]] + timeline_plots[[2]]
+```
+
+## Goal 4: Determine sensitivity of classification to measure age differences in body position
 <!-- imu-compiled.csv, generated from analysis-lena-imu/compile_lena_imu.R -->
@@
   geom_smooth(aes(x = age, y = prop), color = "#191919", method = "lm", se = T) + 
-  stat_summary(fun = "mean", mapping = aes(x = age, y = prop, group = id_uni, fill = position), shape = 21, size = .75, alpha = .7) + 
+  stat_summary(fun = "mean", mapping = aes(x = age, y = prop, group = id_uni, fill = position), shape = 21, size = .7) + 
   scale_fill_manual(values = pal, guide = "none") + 
@@
 
-## Goal 4: Examine wear time and compliance in full-day data collection
-<!-- imu-compiled.csv, generated from analysis-lena-imu/compile_lena_imu.R -->
+```{r agetable}
+pos_sum_age <- map(fullday, fullday_long) %>% 
+  map(~ group_by(.x, group, position, id_uni) %>% 
+            mutate(prop = prop * 100) %>% 
+            get_summary_stats(prop) %>% 
+            group_by(group, position) %>% 
+            get_summary_stats(mean) %>% 
+            select(group, position, mean, sd))
+pos_sum_age <- pos_sum_age %>% map2_dfr(c("Group", "Individual"), 
+                         ~ mutate(.x, Model = .y,
+                                  mean_sd = str_glue("{apa_num(mean, digits = 1)}% ({apa_num(sd, digits = 1)})")) %>% 
+                           select(-mean, -sd) %>% 
+                           pivot_wider(names_from = group, values_from = mean_sd)) %>% 
+  pivot_wider(names_from = Model, values_from = c("Younger", "Older")) %>% 
+  relocate(Younger_Individual, .after = Older_Group)
+apa_table(pos_sum_age, col.names = c("Position", "Younger", "Older", "Younger", "Older"), col_spanners = list("Group" = c(2, 3), "Individual" = c(4, 5)),
+          caption = "Summary of age differences in full-day body position for younger (4- to 7-month) and older (11- to 14-month) infants.
+Values shown are the mean percent of time for each body position averaged across infants in each group.
+Standard deviations are shown in parentheses.
+Descriptive statistics are shown separately for group and individual models.")
+
+```
 

### 2023-02-19 manuscript.tex
paneled age graph
metrics plot
age summaries

@@
 
+\begin{figure}
+
+{\centering \includegraphics{manuscript_files/figure-latex/metrics-1}
+
+}
+
+\caption{Metrics}\label{fig:metrics}
+\end{figure}
+
 \hypertarget{goal-2-assess-classification-accuracy-over-long-recordings}{%
@@
 
-\hypertarget{goal-3-compare-classification-estimates-to-prior-literature}{%
-\subsection{Goal 3: Compare classification estimates to prior literature}\label{goal-3-compare-classification-estimates-to-prior-literature}}
+\hypertarget{goal-3-examine-wear-time-and-compliance-in-full-day-data-collection}{%
+\subsection{Goal 3: Examine wear time and compliance in full-day data collection}\label{goal-3-examine-wear-time-and-compliance-in-full-day-data-collection}}
 
@@
 
-\hypertarget{goal-4-examine-wear-time-and-compliance-in-full-day-data-collection}{%
-\subsection{Goal 4: Examine wear time and compliance in full-day data collection}\label{goal-4-examine-wear-time-and-compliance-in-full-day-data-collection}}
+\hypertarget{goal-4-determine-sensitivity-of-classification-to-measure-age-differences-in-body-position}{%
+\subsection{Goal 4: Determine sensitivity of classification to measure age differences in body position}\label{goal-4-determine-sensitivity-of-classification-to-measure-age-differences-in-body-position}}
+
+\begin{figure}
+
+{\centering \includegraphics{manuscript_files/figure-latex/age-1}
+
+}
+
+\caption{Age trends}\label{fig:age}
+\end{figure}
+
+\begin{table}[tbp]
+
+\begin{center}
+\begin{threeparttable}
+
+\caption{\label{tab:agetable}Summary of age differences in full-day body position for younger (4- to 7-month) and older (11- to 14-month) infants.
+Values shown are the mean percent of time for each body position averaged across infants in each group.
+extit\{SD\} is shown in parentheses.
+Estimates are shown separately for group and individual models.}
+
+\begin{tabular}{lllll}
+\toprule
+ & \multicolumn{2}{c}{Group} & \multicolumn{2}{c}{Individual} \\
+\cmidrule(r){2-3} \cmidrule(r){4-5}
+Position & Younger & Older & Younger & Older\\
+\midrule
+Upright & 7.6\% (8.9) & 18.6\% (7.4) & 9.9\% (13.1) & 18.7\% (8.4)\\
+Sitting & 26.3\% (12.1) & 44.4\% (10.1) & 20.2\% (16.3) & 46.9\% (13.3)\\
+Prone & 13.8\% (13.5) & 14.4\% (6.0) & 11.9\% (9.9) & 16.9\% (10.6)\\
+Supine & 37.9\% (23.2) & 14.0\% (8.4) & 39.4\% (30.3) & 10.0\% (9.5)\\
+Held & 12.7\% (6.9) & 8.5\% (5.4) & 17.6\% (19.9) & 7.4\% (7.6)\\
+\bottomrule
+\end{tabular}
+
+\end{threeparttable}
+\end{center}
+
+\end{table}
 

### 2023-02-19 manuscript.Rmd
overall metric table

@@
 
+```
+
+```{r metricstable}
+
+metrics_long %>% group_by(Model, metric) %>% 
+  get_summary_stats(value) %>% 
+  select(-(variable:max), -(q1:mad), -se, -ci) %>% 
+  pivot_wider(names_from = Model, values_from = c("median", "mean", "sd")) %>% 
+  filter(metric != "Neg Pred Value", metric != "Specificity") %>% 
+  relocate(mean_Group, .after = median_Group) %>% 
+  relocate(sd_Group, .after = mean_Group) %>% 
+  mutate(across(-metric, ~ apa_num(.x, digits = 3)),
+         metric = factor(metric, levels = metric_labels)) %>% 
+  arrange(metric) %>% 
+  apa_table(., col.names = c("Metric", "Median", "Mean", "SD", "Median", "Mean", "SD"),    col_spanners = list("Group" = c(2, 4), "Individual" = c(5, 7)),
+          caption = "Summary statistics for model performance metrics shown separately for group and individual models.")
 ```

### 2023-02-19 manuscript.tex
overall metric table

@@
 
+\begin{table}[tbp]
+
+\begin{center}
+\begin{threeparttable}
+
+\caption{\label{tab:metricstable}Summary statistics for model performance metrics shown separately for group and individual models.}
+
+\begin{tabular}{lllllll}
+\toprule
+ & \multicolumn{3}{c}{Group} & \multicolumn{3}{c}{Individual} \\
+\cmidrule(r){2-4} \cmidrule(r){5-7}
+Metric & Median & Mean & SD & Median & Mean & SD\\
+\midrule
+Overall Accuracy & 0.908 & 0.848 & 0.143 & 0.936 & 0.919 & 0.071\\
+Balanced Accuracy & 0.909 & 0.897 & 0.074 & 0.916 & 0.907 & 0.072\\
+F1 & 0.813 & 0.827 & 0.101 & 0.873 & 0.870 & 0.089\\
+Sensitivity & 0.840 & 0.836 & 0.123 & 0.848 & 0.840 & 0.121\\
+Pos Pred Value & 0.822 & 0.815 & 0.125 & 0.915 & 0.899 & 0.108\\
+Kappa & 0.769 & 0.757 & 0.160 & 0.846 & 0.820 & 0.141\\
+\bottomrule
+\end{tabular}
+
+\end{threeparttable}
+\end{center}
+
+\end{table}
+
 \hypertarget{goal-2-assess-classification-accuracy-over-long-recordings}{%
@@
 Values shown are the mean percent of time for each body position averaged across infants in each group.
-extit\{SD\} is shown in parentheses.
-Estimates are shown separately for group and individual models.}
+Standard deviations are shown in parentheses.
+Descriptive statistics are shown separately for group and individual models.}
 

### 2023-02-21 manuscript.Rmd
exemplar timeline

@@
 
+```{r exemplartimeline, out.width="0.9\\linewidth", include=TRUE, fig.align="center", fig.cap="Example Timeline Caption.", echo=FALSE}
+knitr::include_graphics("figures/timeline.png")
+# Generated from analysis-pt2/methods-exemplar-agreement.R
+# Annotated in illustrator based on event times:
+# Exemplar participant is 120-1
+# > pt2_start_time
+# [1] "2022-11-13 11:11:00 PST"
+# > pt2_end_time
+# [1] "2022-11-13 12:41:01 PST"
+# > session_param$start_time_coded
+# [1] "2022-11-13 10:02:40 PST"
+```
+
 ## Body position annotation
@@
     ggplot(aes(x = Model, y = value)) + 
-    geom_sina(alpha = .7, fill = "aquamarine3", size = 4, maxwidth = .5, shape = 21) +
+    geom_sina(alpha = .7, fill = "turquoise3", size = 4, maxwidth = .5, shape = 21) +
     stat_summary(fun = "mean", mapping = aes(x = Model, y = value), shape = "-", fill = "black", size = 4) + 
@@
   ggplot(filter_outliers(ds), aes(x = pos, y = diff, fill = pos, group = file)) +
-  annotate("rect", xmin = -Inf, xmax = Inf, ymin = -.1, ymax = .1, fill = "darkgray", color = "darkgray") + 
+  annotate("rect", xmin = -Inf, xmax = Inf, ymin = -.1, ymax = .1, fill = "gray", color = "gray") + 
   #geom_rect(aes(xmin = "Held", xmax = "Upright",ymin = -.1, ymax = .1), fill = "#e5e5e5", color = "#e5e5e5") + 

### 2023-02-21 group_split_comparison/group_split.R
by class metrics

@@ -42,7 +42,7 @@ split_model <- function(temp_id, training, ntree = 50) {
 
 ids <- unique(training$id)
 group_mods <-map(ids, ~group_model(.x, training)) %>% set_names(ids)
-group_select_mods <- map(ids, ~group_model(.x, training, bad_ids = bad_ids)) %>% set_names(ids)
+# group_select_mods <- map(ids, ~group_model(.x, training, bad_ids = bad_ids)) %>% set_names(ids)
 split_mods <- map(ids, ~split_model(.x, training)) %>% set_names(ids)
 
 # save(group_select_mods, split_mods, file = "group_split_comparison/models.RData")
@@ -59,10 +59,10 @@ res_group <- map2(group_mods, ids, ~metrics(.x, filter(testing, id == .y))) %>%
   add_column(ids) %>% add_column(model = "group")
 res_group %>% get_summary_stats(-ids)
 
-res_select <- map2(group_select_mods, ids, ~metrics(.x, filter(testing, id == .y))) %>% 
-  map_dfr(~ select(.x, `Overall Accuracy`:Kappa)) %>% 
-  add_column(ids) %>% add_column(model = "select")
-res_select %>% get_summary_stats(-ids)
+# res_select <- map2(group_select_mods, ids, ~metrics(.x, filter(testing, id == .y))) %>% 
+#   map_dfr(~ select(.x, `Overall Accuracy`:Kappa)) %>% 
+#   add_column(ids) %>% add_column(model = "select")
+# res_select %>% get_summary_stats(-ids)
 
 res_split <- map2(split_mods, ids, ~metrics(.x, filter(testing, id == .y))) %>% 
   map_dfr(~ select(.x, `Overall Accuracy`:Kappa)) %>% 
@@ -70,12 +70,24 @@ res_split <- map2(split_mods, ids, ~metrics(.x, filter(testing, id == .y))) %>%
 res_split %>% get_summary_stats(-ids)
 res_split %>% arrange(`Overall Accuracy`)
 
-ds <- bind_rows(res_split, res_group) %>% bind_rows(res_select)
+ds <- bind_rows(res_split, res_group) # %>% bind_rows(res_select)
 write_csv(ds, file = "data/group_split_metrics.csv")
 
-ggplot(ds, aes(x = model, y = `Overall Accuracy`)) + geom_boxplot()
-ggplot(ds, aes(x = model, y = `Balanced Accuracy`)) + geom_boxplot()
[…]

### 2023-02-21 manuscript.Rmd
class level table

@@
           caption = "Summary statistics for model performance metrics shown separately for group and individual models.")
+```
+
+\renewcommand{\arraystretch}{.75}
+```{r metricsbyclass}
+metrics_class <- read_csv("data/group_split_metrics_class.csv") %>% 
+  filter(model %in% c("group", "split")) %>% 
+  mutate(Model = factor(model, levels = c("group", "split"), labels = c("Group", "Individual")))
+
+metrics_long_class <- metrics_class %>% pivot_longer(cols = `Balanced Accuracy`:Kappa, 
+                                         names_to = "metric",
+                                         values_to = "value")
+
+metrics_long_summary <- metrics_long_class %>% group_by(Class,Model, metric) %>% 
+  get_summary_stats(value) %>% 
+  select(-(variable:max), -(q1:mad), -se, -ci) %>% 
+  pivot_wider(names_from = Model, values_from = c("median", "mean", "sd")) %>% 
+  filter(metric != "Neg Pred Value", metric != "Specificity") %>% 
+  relocate(mean_Group, .after = median_Group) %>% 
+  relocate(sd_Group, .after = mean_Group) %>% 
+  mutate(across(-(Class:metric), ~ apa_num(.x, digits = 3)),
+         metric = factor(metric, levels = metric_labels),
+         Class = factor(Class, levels = c("Supine", "Prone", "Sitting", "Upright", "Held"))) %>% 
+  relocate(metric, .before = Class) %>% 
+  arrange(metric, Class)
+
+class_table <- metrics_long_summary %>% mutate(metric = as.character(metric))
+class_table$metric[duplicated(class_table$metric)] <- " "
+
+apa_table(class_table, col.names = c("Metric","Position", "Median", "Mean", "SD", "Median", "Mean", "SD"), col_spanners = list("Group" = c(3, 5), "Individual" = c(6, 8)),
+          midrules = c(5, 10, 15, 20),
+          caption = "Model performance metrics for each body position class, shown separately for group and individual models.")
 ```

### 2023-02-21 manuscript.tex
exemplar timeline
class level table

@@
 
+\begin{figure}
+
+{\centering \includegraphics[width=0.9\linewidth]{figures/timeline}
+
+}
+
+\caption{Example Timeline Caption.}\label{fig:exemplartimeline}
+\end{figure}
+
 \hypertarget{body-position-annotation}{%
@@
 Kappa & 0.769 & 0.757 & 0.160 & 0.846 & 0.820 & 0.141\\
+\bottomrule
+\end{tabular}
+
+\end{threeparttable}
+\end{center}
+
+\end{table}
+
+\renewcommand{\arraystretch}{.75}
+
+\begin{table}[tbp]
+
+\begin{center}
+\begin{threeparttable}
+
+\caption{\label{tab:metricsbyclass}Summary statistics for model performance metrics shown separately for group and individual models.}
+
+\begin{tabular}{llllllll}
+\toprule
+ &  & \multicolumn{3}{c}{Group} & \multicolumn{3}{c}{Individual} \\
+\cmidrule(r){3-5} \cmidrule(r){6-8}
+Metric & Position & Median & Mean & SD & Median & Mean & SD\\
+\midrule
+Balanced Accuracy & Supine & 0.977 & 0.935 & 0.099 & 0.995 & 0.957 & 0.090\\
+ & Prone & 0.997 & 0.950 & 0.110 & 0.983 & 0.921 & 0.132\\
+ & Sitting & 0.932 & 0.861 & 0.151 & 0.973 & 0.943 & 0.071\\
+ & Upright & 0.930 & 0.867 & 0.146 & 0.937 & 0.877 & 0.123\\
+ & Held & 0.907 & 0.872 & 0.111 & 0.892 & 0.854 & 0.153\\ \midrule
+F1 & Supine & 0.952 & 0.873 & 0.173 & 0.987 & 0.937 & 0.130\\
+ & Prone & 0.965 & 0.923 & 0.124 & 0.958 & 0.889 & 0.186\\
+ & Sitting & 0.897 & 0.805 & 0.247 & 0.963 & 0.927 & 0.101\\
+ & Upright & 0.781 & 0.748 & 0.235 & 0.868 & 0.769 & 0.227\\
+ & Held & 0.813 & 0.783 & 0.163 & 0.862 & 0.786 & 0.240\\ \midrule
+Sensitivity & Supine & 1.000 & 0.945 & 0.125 & 1.000 & 0.948 & 0.141\\
+ & Prone & 1.000 & 0.907 & 0.216 & 0.988 & 0.856 & 0.266\\
+ & Sitting & 0.901 & 0.787 & 0.291 & 0.977 & 0.917 & 0.136\\
+ & Upright & 0.878 & 0.767 & 0.275 & 0.893 & 0.790 & 0.255\\
+ & Held & 0.867 & 0.774 & 0.229 & 0.817 & 0.722 & 0.309\\ \midrule
+Pos Pred Value & Supine & 0.973 & 0.802 & 0.300 & 1.000 & 0.939 & 0.131\\
+ & Prone & 0.991 & 0.896 & 0.196 & 0.993 & 0.901 & 0.197\\
+ & Sitting & 0.910 & 0.840 & 0.238 & 0.967 & 0.947 & 0.078\\
+ & Upright & 0.848 & 0.743 & 0.285 & 0.875 & 0.817 & 0.205\\
+ & Held & 0.917 & 0.803 & 0.247 & 0.949 & 0.872 & 0.232\\ \midrule
+Kappa & Supine & 0.932 & 0.770 & 0.307 & 0.984 & 0.913 & 0.169\\
+ & Prone & 0.951 & 0.887 & 0.208 & 0.948 & 0.851 & 0.242\\
+ & Sitting & 0.836 & 0.712 & 0.308 & 0.946 & 0.889 & 0.134\\
+ & Upright & 0.729 & 0.700 & 0.277 & 0.829 & 0.744 & 0.232\\
+ & Held & 0.786 & 0.733 & 0.219 & 0.845 & 0.744 & 0.280\\
 \bottomrule

### 2023-02-22 manuscript.tex
updated timeline figure
updated part2 data

@@
 
-\caption{\label{tab:metricsbyclass}Summary statistics for model performance metrics shown separately for group and individual models.}
+\caption{\label{tab:metricsbyclass}Model performance metrics for each body position class, shown separately for group and individual models.}
 
@@
 \midrule
-Held & -0.02 & 0.16 & 0.55 & 0.63\\
-Prone & 0.97 & 0.83 & 0.97 & 0.84\\
-Sitting & 0.72 & 0.93 & 0.91 & 0.97\\
-Supine & 0.84 & 0.93 & 0.94 & 0.97\\
+Held & -0.03 & 0.12 & 0.55 & 0.58\\
+Prone & 0.97 & 0.85 & 0.97 & 0.84\\
+Sitting & 0.73 & 0.93 & 0.91 & 0.97\\
+Supine & 0.84 & 0.94 & 0.94 & 0.97\\
 Upright & 0.84 & 0.93 & 0.99 & 0.94\\ \midrule
-Overall & 0.79 & 0.90 & 0.95 & 0.96\\
+Overall & 0.80 & 0.91 & 0.95 & 0.96\\
 \bottomrule
@@
 \midrule
-Held & 0.45 & 0.44 & 0.57 & 0.56\\
-Prone & 0.96 & 0.81 & 0.96 & 0.88\\
-Sitting & 0.72 & 0.93 & 0.91 & 0.93\\
-Supine & 0.75 & 0.93 & 0.90 & 0.94\\
+Held & 0.45 & 0.43 & 0.57 & 0.55\\
+Prone & 0.96 & 0.89 & 0.96 & 0.88\\
+Sitting & 0.72 & 0.92 & 0.91 & 0.93\\
+Supine & 0.75 & 0.95 & 0.90 & 0.94\\
 Upright & 0.93 & 0.95 & 0.97 & 0.95\\ \midrule
-Overall & 0.80 & 0.92 & 0.92 & 0.94\\
+Overall & 0.80 & 0.94 & 0.93 & 0.94\\
 \bottomrule

### 2023-02-23 manuscript.Rmd
started writing intro

@@
 
-# Current Study
+Infants' movements facilitate and constrain how they interact with their surroundings from moment to moment.
+Changes in *body position*---whether infants are supine on their backs, prone on their bellies, sitting, upright, or held by a  caregiver---have immediate consequences for vision, object exploration, and social interaction.
+When sitting and upright, infants have a better view of faces and distant objects compared to their view while prone [@Freeplay; @LuoFranchak2020; @CWW].
+Infants struggle to manipulate objects while supine and prone, but sitting affords object exploration [@SoskaAdolph2014].
+
+# Goals of the Current Study
 
@@
 
-```{r exemplartimeline, out.width="0.9\\linewidth", include=TRUE, fig.align="center", fig.cap="Example Timeline Caption.", echo=FALSE}
+```{r exemplartimeline, out.width="0.99\\linewidth", include=TRUE, fig.align="center", fig.cap="Example Timeline Caption.", echo=FALSE}
 knitr::include_graphics("figures/timeline.png")
@@
 # Exemplar participant is 120-1
-# > pt2_start_time
-# [1] "2022-11-13 11:11:00 PST"
-# > pt2_end_time
-# [1] "2022-11-13 12:41:01 PST"
-# > session_param$start_time_coded
-# [1] "2022-11-13 10:02:40 PST"
+# > pt2_start_time = "2022-11-13 11:11:00 PST"
+# > pt2_end_time = "2022-11-13 12:41:01 PST"
+# > session_param$start_time_coded = "2022-11-13 10:02:40 PST"
 ```
@@
 
-## Goal 1: Optimize and validate body position classification model
+## Goal 1: Assess the proximal accuracy of body position classification models
+
 <!-- group_split_metrics, generated from group_split_comparison in this proj, stable copy in data/ -->
+
 ```{r metrics, warning=FALSE, fig.cap="Metrics", fig.align="center", fig.width=5, fig.height=8}
@@
 \renewcommand{\arraystretch}{.75}
+
 ```{r metricsbyclass}
@@
           midrules = c(5, 10, 15, 20),
-          caption = "Model performance metrics for each body position class, shown separately for group and individual models.")
+          caption = "Model performance metrics for each body position category, shown separately for group and individual models.")
 ```
 
-## Goal 2: Assess classification accuracy over long recordings
+## Goal 2: Assess the distal accuracy of body position classification models
+
 <!-- compiled_agreement, generated from analysis-part2 -->
 
-```{r part2overall, warning=FALSE, fig.cap="Overall agreement between human-coded body position and model-predicted body position in the long-delay period.
+```{r part2overall, warning=FALSE, fig.cap="Overall agreement between human-coded body position and model-predicted body position in the distal comparison.
 Agreement for group models is shown in (A) and agreement for individual models is shown in (B).
@@
 
-```{r part2bins, warning=FALSE, fig.cap="Prediction performance (difference in minutes between human-coded and model-predicted body position) for 10-minute bins in the long delay period.
+```{r part2bins, warning=FALSE, fig.cap="Prediction performance (difference in minutes between human-coded and model-predicted body position) for 10-minute bins in the distal comparison.
 Each point shows the mean and SE for a single participant for each body position, summarizing the prediction difference for each of their 10-minute bins.
@@
 apa_table(pt2corr_bins, col.names = c("Position", "Group", "Individual", "Group", "Individual"), col_spanners = list(`With Outliers` = c(2, 3), `Without Outliers` = c(4, 5)),
-          midrules = 5, caption = "Correlations between human-coded and model-predicted body position durations using 10-minute bins during the long delay period.
+          midrules = 5, caption = "Correlations between human-coded and model-predicted body position durations using 10-minute bins during the distal comparison.
 Correlations are provided within each posture and overall, and computed separately using group and individual models with and without outlier participants.")
@@
 
-## Goal 3: Examine wear time and compliance in full-day data collection
+## Goal 3: Examine the data quality of full-day home recordings
+
 <!-- imu-compiled.csv, generated from analysis-lena-imu/compile_lena_imu.R -->
@@
 
-## Goal 4: Determine sensitivity of classification to measure age differences in body position
+## Goal 4: Assess the suitability of full-day predictions for capturing age differences in body position
+
 <!-- imu-compiled.csv, generated from analysis-lena-imu/compile_lena_imu.R -->

### 2023-02-23 manuscript.Rmd
intro paragraphs

@@
 Changes in *body position*---whether infants are supine on their backs, prone on their bellies, sitting, upright, or held by a  caregiver---have immediate consequences for vision, object exploration, and social interaction.
-When sitting and upright, infants have a better view of faces and distant objects compared to their view while prone [@Freeplay; @LuoFranchak2020; @CWW].
+When sitting and upright, infants have a better view of faces and distant objects compared to their view while in a prone position [@Freeplay; @LuoFranchak2020; @CWW].
 Infants struggle to manipulate objects while supine and prone, but sitting affords object exploration [@SoskaAdolph2014].
+Upright walking changes how infants and caregivers interact compared with crawling in a prone position; while walking infants move farther away, share toys in different ways, and hear different language from caregivers [@ChenSchneider2022; @Karasik2011; @WestIverson2021; @Karasik2014].
+As infants grow older and acquire new abilities, such as independent sitting and walking, the spend more time sitting and upright and less time held, supine, and prone [@Survey; @Thurman2017; @Freeplay; @AdolphCTL2014].
+Thus, characterizing individual differences in the day-to-day accumulation of body position experiences informs developmental theory by revealing differential opportunities for learning.
+<!-- Other west cites; cite new EMA -->
 
-# Goals of the Current Study
+In this paper, we present an inertial sensing method to classify infant body position from moment-to-moment across an entire day, and validate its accuracy using XX hours of video recorded across XX in-home data collection sessions.
+Our method takes inspiration from a more mature technology: Long-form audio recordings of infants' language experiences using wearable audio recorders such as the LENA device.
+We begin by describing the impact of long-form audio recordings on research in developmental psychology, and identify the key features of the method that should be replicated in long-form recordings of motor behavior.
+Next, we review the current state-of-the-art in measuring infant motor behavior---video and survey data---and their limitations in capturing real-time, full-day behavior.
+Finally, we discuss the advantages of using inertial sensing to classify motor behavior.
+Despite promising past results in brief sessions [CITES], the current investigation takes a needed step forward by testing accuracy over long recordings.
+
+## Inspiration from Long-Form Audio Recordings
+
+## Limitations of Video and Survey Methods
+
+## Promise of Inertial Sensing Methods
+
+## Goals of the Current Study
 

### 2023-02-23 manuscript.tex
intro paragraphs

@@
 
-\hypertarget{current-study}{%
-\section{Current Study}\label{current-study}}
+Infants' movements facilitate and constrain how they interact with their surroundings from moment to moment.
+Changes in \emph{body position}---whether infants are supine on their backs, prone on their bellies, sitting, upright, or held by a caregiver---have immediate consequences for vision, object exploration, and social interaction.
+When sitting and upright, infants have a better view of faces and distant objects compared to their view while in a prone position (Franchak et al., 2018; Kretch et al., 2014; Luo \& Franchak, 2020).
+Infants struggle to manipulate objects while supine and prone, but sitting affords object exploration (Soska \& Adolph, 2014).
+Upright walking changes how infants and caregivers interact compared with crawling in a prone position; while walking infants move farther away, share toys in different ways, and hear different language from caregivers (Chen et al., 2022; Karasik et al., 2011, 2014; West \& Iverson, 2021).
+As infants grow older and acquire new abilities, such as independent sitting and walking, the spend more time sitting and upright and less time held, supine, and prone (Adolph \& Tamis-LeMonda, 2014; Franchak et al., 2018; Franchak, 2019; Thurman \& Corbetta, 2017).
+Thus, characterizing individual differences in the day-to-day accumulation of body position experiences informs developmental theory by revealing differential opportunities for learning.
+
+\hypertarget{goals-of-the-current-study}{%
+\section{Goals of the Current Study}\label{goals-of-the-current-study}}
 
@@
 
-{\centering \includegraphics[width=0.9\linewidth]{figures/timeline}
+{\centering \includegraphics[width=0.99\linewidth]{figures/timeline}
 
@@
 
-\hypertarget{goal-1-optimize-and-validate-body-position-classification-model}{%
-\subsection{Goal 1: Optimize and validate body position classification model}\label{goal-1-optimize-and-validate-body-position-classification-model}}
+\hypertarget{goal-1-assess-the-proximal-accuracy-of-body-position-classification-models}{%
+\subsection{Goal 1: Assess the proximal accuracy of body position classification models}\label{goal-1-assess-the-proximal-accuracy-of-body-position-classification-models}}
 
@@
 
-\caption{\label{tab:metricsbyclass}Model performance metrics for each body position class, shown separately for group and individual models.}
+\caption{\label{tab:metricsbyclass}Model performance metrics for each body position category, shown separately for group and individual models.}
 
@@
 
-\hypertarget{goal-2-assess-classification-accuracy-over-long-recordings}{%
-\subsection{Goal 2: Assess classification accuracy over long recordings}\label{goal-2-assess-classification-accuracy-over-long-recordings}}
+\hypertarget{goal-2-assess-the-distal-accuracy-of-body-position-classification-models}{%
+\subsection{Goal 2: Assess the distal accuracy of body position classification models}\label{goal-2-assess-the-distal-accuracy-of-body-position-classification-models}}
 
@@
 
-\caption{Overall agreement between human-coded body position and model-predicted body position in the long-delay period.
+\caption{Overall agreement between human-coded body position and model-predicted body position in the distal comparison.
 Agreement for group models is shown in (A) and agreement for individual models is shown in (B).
@@
 
-\caption{Prediction performance (difference in minutes between human-coded and model-predicted body position) for 10-minute bins in the long delay period.
+\caption{Prediction performance (difference in minutes between human-coded and model-predicted body position) for 10-minute bins in the distal comparison.
 Each point shows the mean and SE for a single participant for each body position, summarizing the prediction difference for each of their 10-minute bins.
@@
 
-\caption{\label{tab:pt2binstable}Correlations between human-coded and model-predicted body position durations using 10-minute bins during the long delay period.
+\caption{\label{tab:pt2binstable}Correlations between human-coded and model-predicted body position durations using 10-minute bins during the distal comparison.
 Correlations are provided within each posture and overall, and computed separately using group and individual models with and without outlier participants.}
@@
 
-\hypertarget{goal-3-examine-wear-time-and-compliance-in-full-day-data-collection}{%
-\subsection{Goal 3: Examine wear time and compliance in full-day data collection}\label{goal-3-examine-wear-time-and-compliance-in-full-day-data-collection}}
+\hypertarget{goal-3-examine-the-data-quality-of-full-day-home-recordings}{%
+\subsection{Goal 3: Examine the data quality of full-day home recordings}\label{goal-3-examine-the-data-quality-of-full-day-home-recordings}}
 
@@
 
-\hypertarget{goal-4-determine-sensitivity-of-classification-to-measure-age-differences-in-body-position}{%
-\subsection{Goal 4: Determine sensitivity of classification to measure age differences in body position}\label{goal-4-determine-sensitivity-of-classification-to-measure-age-differences-in-body-position}}
+\hypertarget{goal-4-assess-the-suitability-of-full-day-predictions-for-capturing-age-differences-in-body-position}{%
+\subsection{Goal 4: Assess the suitability of full-day predictions for capturing age differences in body position}\label{goal-4-assess-the-suitability-of-full-day-predictions-for-capturing-age-differences-in-body-position}}
 
@@
 \hypertarget{refs}{}
-\begin{CSLReferences}{0}{0}
+\begin{CSLReferences}{1}{0}
+\leavevmode\vadjust pre{\hypertarget{ref-AdolphCTL2014}{}}%
+Adolph, K.
+E., \& Tamis-LeMonda, C.
+S.
+(2014).
+The costs and benefits of development: The transition from crawling to walking.
+\emph{Child Development Perspectives}, \emph{8}, 187--192.
+\url{https://doi.org/10.1111/cdep.12085}
+
+\leavevmode\vadjust pre{\hypertarget{ref-ChenSchneider2022}{}}%
+Chen, Q., Schneider, J.
+L., West, K.
+L., \& Iverson, J.
+M.
+(2022).
+Infant locomotion shapes proximity to adults during everyday play in the u.s.
+\emph{Infancy}.
+\url{https://doi.org/10.1111/infa.12503}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Survey}{}}%
+Franchak, J. M. (2019). Changing opportunities for learning in everyday life: Infant body position over the first year. \emph{Infancy}, \emph{24}, 187--209.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Freeplay}{}}%
+Franchak, J. M., Kretch, K. S., \& Adolph, K. E. (2018). See and be seen: Infant-caregiver social looking during locomotor free play. \emph{Developmental Science}, \emph{21}, e12626.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Karasik2011}{}}%
+Karasik, L.
+B., Tamis-LeMonda, C.
+S., \& Adolph, K.
+E.
+(2011).
+Transition from crawling to walking and infants' actions with objects and people.
+\emph{Child Development}, \emph{82}, 1199--1209.
+\url{https://doi.org/10.1111/j.1467-8624.2011.01595.x}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Karasik2014}{}}%
+Karasik, L.
+B., Tamis-LeMonda, C.
+S., \& Adolph, K.
+E.
+(2014).
+Crawling and walking infants elicit different verbal responses from mothers.
+\emph{Developmental Science}, \emph{17}, 388--395.
+\url{https://doi.org/10.1111/desc.12129}
+
+\leavevmode\vadjust pre{\hypertarget{ref-CWW}{}}%
+Kretch, K.
+S., Franchak, J.
+M., \& Adolph, K.
+E.
+(2014).
+Crawling and walking infants see the world differently.
+\emph{Child Development}, \emph{85}, 1503--1518.
+\url{https://doi.org/10.1111/cdev.12206}
+
+\leavevmode\vadjust pre{\hypertarget{ref-LuoFranchak2020}{}}%
+Luo, C., \& Franchak, J. M. (2020). Head and body structure infants' visual experiences during mobile, naturalistic play. \emph{{PLoS ONE}}, \emph{15}, e0242009.
+
+\leavevmode\vadjust pre{\hypertarget{ref-SoskaAdolph2014}{}}%
+Soska, K. C., \& Adolph, K. E. (2014). Postural position constrains multimodal object exploration in infants. \emph{Infancy}, \emph{19}, 138--161. \url{https://doi.org/10.1111/infa.12039}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Thurman2017}{}}%
+Thurman, S. L., \& Corbetta, D. (2017). Spatial exploration and changes in infant-mother dyads around transitions in infant locomotion. \emph{Developmental Psychology}, \emph{53}, 1207--1221.
+
+\leavevmode\vadjust pre{\hypertarget{ref-WestIverson2021}{}}%
+West, K. L., \& Iverson, J. M. (2021). Communication changes when infants begin to walk. \emph{Developmental Science}, \emph{24}(5). \url{https://doi.org/10.1111/desc.13102}
+
 \end{CSLReferences}

### 2023-02-23 manuscript.Rmd
intro outline

@@
 
-## Inspiration from Long-Form Audio Recordings
+## Inspiration from Long-Form Audio Methods
+
+Impactful findings
+- individual differences in experiences
+- distribution of experiences (burstiness)
+- moderators of experience (RT, Malachowski body position)
+
+Key features
+- unobtrusive (Bergelson; CTL stuff)
+- mobile (no limits to where recording happens)
+- real time
+- long duration (capture entire day and heterogeniety in activities)
+- automatic classification (scaling)
 
@@
 
+Video. Only advantage is real time
+
+Survey. Everything but real time
+
 ## Promise of Inertial Sensing Methods
+
+Wear time is not an issue. Lots of Smith and other studies with long recordings.
+
+Classification accuracy from ML is good. older papers; Airaksinen; Frontiers
+
+What's missing
 

### 2023-02-23 manuscript.tex
intro outline

@@
 
+In this paper, we present an inertial sensing method to classify infant body position from moment-to-moment across an entire day, and validate its accuracy using XX hours of video recorded across XX in-home data collection sessions.
+Our method takes inspiration from a more mature technology: Long-form audio recordings of infants' language experiences using wearable audio recorders such as the LENA device.
+We begin by describing the impact of long-form audio recordings on research in developmental psychology, and identify the key features of the method that should be replicated in long-form recordings of motor behavior.
+Next, we review the current state-of-the-art in measuring infant motor behavior---video and survey data---and their limitations in capturing real-time, full-day behavior.
+Finally, we discuss the advantages of using inertial sensing to classify motor behavior.
+Despite promising past results in brief sessions {[}CITES{]}, the current investigation takes a needed step forward by testing accuracy over long recordings.
+
+\hypertarget{inspiration-from-long-form-audio-recordings}{%
+\subsection{Inspiration from Long-Form Audio Recordings}\label{inspiration-from-long-form-audio-recordings}}
+
+\hypertarget{limitations-of-video-and-survey-methods}{%
+\subsection{Limitations of Video and Survey Methods}\label{limitations-of-video-and-survey-methods}}
+
+\hypertarget{promise-of-inertial-sensing-methods}{%
+\subsection{Promise of Inertial Sensing Methods}\label{promise-of-inertial-sensing-methods}}
+
 \hypertarget{goals-of-the-current-study}{%
-\section{Goals of the Current Study}\label{goals-of-the-current-study}}
+\subsection{Goals of the Current Study}\label{goals-of-the-current-study}}
 

### 2023-02-24 manuscript.Rmd
LENA stuff

--- 2023-02-23 manuscript.Rmd
@@
 Our method takes inspiration from a more mature technology: Long-form audio recordings of infants' language experiences using wearable audio recorders such as the LENA device.
-We begin by describing the impact of long-form audio recordings on research in developmental psychology, and identify the key features of the method that should be replicated in long-form recordings of motor behavior.
+We begin by describing the impact of long-form audio recordings on research in developmental psychology, and identify the key features that should be replicated in long-form recordings of motor behavior.
 Next, we review the current state-of-the-art in measuring infant motor behavior---video and survey data---and their limitations in capturing real-time, full-day behavior.
@@
 
-## Inspiration from Long-Form Audio Recordings
+## Inspiration from Long-Form Audio Methods
+
+The LENA&reg; recorder is a commercial device that is worn by infants in a custom shirt pocket that has sufficient battery life and storage to record for an entire day.
+Closed-source LENA&reg; algorithms analyze the audio recordings to provide counts of useful metrics, such as the number of words spoken by adults in the vicinity of the participant.
+Other long-form audio methods rely on custom-built recorders [@WassPhillips2022], apply alternative classification algorithms to LENA&reg; data [@RasanenSeshadri2020; @MichelettiYao2022], or manually transcribe audio recorded by LENA&reg; devices to improve accuracy or identify behaviors beyond the built-in categories [@Bergelson2019Corpus; @MendozaFausey2021].
+
+Long-form audio recordings have had a transformational impact on language development by allowing researchers to characterize opportunities for learning in daily life.
+Measuring the amount of speech heard by infants in the home [@Weisleder2013] or in a daycare setting [@PerryPrince2018] revealed individual differences input that predict later vocabulary.
+Full-day language recording synchronized with other data sources allows researchers to identify how vocal input and production interacts with other processes.
+Beyond individual differences in aggregated data, long-form recordings can be used to determine the temporal schedule of experiences.
+For example, infants' daily experiences hearing music are clustered in time, with "bursty" episodes of hearing music separated by relatively long periods during which music is absent [@MendozaFausey2022].
+Synchronizing audio recordings with other data sources extends researchers' ability to characterize daily experiences.
+Linking LENA&reg; speech measurements to repeated, time-stamped text-message surveys about infant device placement revealed that infants heard less caregiver speech during moments that they were restrained in devices such as swings, exersaucers, and car seats [@MalachowskiSalo2023].
+A custom-built wearable ECG and audio recorder allowed @WassPhillips2022 to discover that infant arousal increases the likelihood of infant vocalization across the day.
+The ability to record real-time, synchronized data is vital for making inferences about processes that happen on the timescale of minutes or even seconds within in an individual as opposed to comparisons of aggregated data between infants.
+
+Key features
+- unobtrusive (Bergelson; CTL stuff)
+- mobile (no limits to where recording happens)
+- real time
+- long duration (capture entire day and heterogeniety in activities)
+- automatic classification with good-enough accuracy (scaling)
 

### 2023-02-24 manuscript.tex
LENA stuff

@@
 
-\hypertarget{inspiration-from-long-form-audio-recordings}{%
-\subsection{Inspiration from Long-Form Audio Recordings}\label{inspiration-from-long-form-audio-recordings}}
+\hypertarget{inspiration-from-long-form-audio-methods}{%
+\subsection{Inspiration from Long-Form Audio Methods}\label{inspiration-from-long-form-audio-methods}}
+
+Impactful findings (Hailey write this paragraph)
+- individual differences in experiences
+- distribution of experiences (burstiness)
+Full-day language recording synchronized with other data sources allows researchers to identify how vocal input and production interacts with other processes.
+Linking LENA® speech measurements to repeated, time-stamped text-message surveys about infant device use revealed that infants heard less caregiver speech during moments that they were restrained in devices such as swings, exersaucers, and car seats (Malachowski et al., 2023).
+A custom-built wearable ECG and audio recorder allowed Wass et al.
+(2022) to discover that infant arousal increases the likelihood of infant vocalization across the day.
+In both examples, the ability to record real-time, synchronized data is vital for making inferences about processes that happen on the timescale of minutes or even seconds within in an individual as opposed to comparisons of aggregated data between infants.
+
+Key features
+- unobtrusive (Bergelson; CTL stuff)
+- mobile (no limits to where recording happens)
+- real time
+- long duration (capture entire day and heterogeniety in activities)
+- automatic classification with good-enough accuracy (scaling)
 
@@
 
+Video. Only advantage is real time
+
+Survey. Everything but real time
+
 \hypertarget{promise-of-inertial-sensing-methods}{%
 \subsection{Promise of Inertial Sensing Methods}\label{promise-of-inertial-sensing-methods}}
+
+Wear time is not an issue. Lots of Smith and other studies with long recordings.
+
+Classification accuracy from ML is good. older papers; Airaksinen; Frontiers
+
+What's missing
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-MalachowskiSalo2023}{}}%
+Malachowski, L.
+G., Salo, V.
+C., Needham, A.
+W., \& Humphreys, K.
+L.
+(2023).
+Infant placement and language exposure in daily life.
+\emph{Infant and Child Development}.
+\url{https://doi.org/10.1002/icd.2405}
+
 \leavevmode\vadjust pre{\hypertarget{ref-SoskaAdolph2014}{}}%
@@
 Thurman, S. L., \& Corbetta, D. (2017). Spatial exploration and changes in infant-mother dyads around transitions in infant locomotion. \emph{Developmental Psychology}, \emph{53}, 1207--1221.
+
+\leavevmode\vadjust pre{\hypertarget{ref-WassPhillips2022}{}}%
+Wass, S., Phillips, E., Smith, C., Fatimehin, E.
+O., \& Goupil, L.
+(2022).
+Vocal communication is tied to interpersonal arousal coupling in caregiver-infant dyads.
+\emph{{eLife}}, \emph{11}.
+\url{https://doi.org/10.7554/elife.77399}
 

### 2023-02-24 manuscript.Rmd
five things

@@
 In this paper, we present an inertial sensing method to classify infant body position from moment-to-moment across an entire day, and validate its accuracy using XX hours of video recorded across XX in-home data collection sessions.
-Our method takes inspiration from a more mature technology: Long-form audio recordings of infants' language experiences using wearable audio recorders such as the LENA device.
-We begin by describing the impact of long-form audio recordings on research in developmental psychology, and identify the key features that should be replicated in long-form recordings of motor behavior.
+Our method takes inspiration from a more mature technology: Long-form audio recordings of infants' language experiences using wearable audio recorders.
+We begin by describing the impact of long-form audio recordings on research in developmental psychology, and identify the key features that should be replicated in long-form recordings of motor behavior [CITES].
 Next, we review the current state-of-the-art in measuring infant motor behavior---video and survey data---and their limitations in capturing real-time, full-day behavior.
@@
 A custom-built wearable ECG and audio recorder allowed @WassPhillips2022 to discover that infant arousal increases the likelihood of infant vocalization across the day.
-The ability to record real-time, synchronized data is vital for making inferences about processes that happen on the timescale of minutes or even seconds within in an individual as opposed to comparisons of aggregated data between infants.
 
-Key features
-- unobtrusive (Bergelson; CTL stuff)
-- mobile (no limits to where recording happens)
-- real time
-- long duration (capture entire day and heterogeniety in activities)
+We identified five key features of long-form audio methods to replicate in analagous studies of motor behavior.
+First, wearable audio recorders are *mobile*.
+Measurement is not limited to a particular room because the recording device travels with the participant.
+Data are stored on device, so participants do not need to be in range of a receiver.
+Second, wearable audio recording is *unobtrusive*.
+Participants' reactivity to observation, such as from a video camera, may influence behavior.
+For example, caregivers spoke more frequently to infants during a video-recorded portion of a home recording compared with audio-only segments captured by a LENA&reg; device [@Bergelson2019Input].
+Third and fourth, recordings capture *real-time data* over a *full day*.
+The ability to record real-time data is vital for making inferences about processes that happen on the timescale of minutes or even seconds within in an individual as opposed to comparisons of aggregated data between infants.
+Synchronizing real-time data to other data streams helps to reveal sources of variability within an individual.
+Full-day recordings are essential for capturing experiences across the heterogeniety of daily routines that moderate behavior (e.g., play, feeding, errands) [CITE].
+Burstiness of behavior means that long recordings are needed to capture clusters of events amid long periods in which they may be absent [@BarbaroFausey2022; @WarlaumontSobowale2021].
+
 - automatic classification with good-enough accuracy (scaling)

### 2023-02-24 manuscript.tex
five things

@@
 In this paper, we present an inertial sensing method to classify infant body position from moment-to-moment across an entire day, and validate its accuracy using XX hours of video recorded across XX in-home data collection sessions.
-Our method takes inspiration from a more mature technology: Long-form audio recordings of infants' language experiences using wearable audio recorders such as the LENA device.
-We begin by describing the impact of long-form audio recordings on research in developmental psychology, and identify the key features that should be replicated in long-form recordings of motor behavior.
+Our method takes inspiration from a more mature technology: Long-form audio recordings of infants' language experiences using wearable audio recorders.
+We begin by describing the impact of long-form audio recordings on research in developmental psychology, and identify the key features that should be replicated in long-form recordings of motor behavior {[}CITES{]}.
 Next, we review the current state-of-the-art in measuring infant motor behavior---video and survey data---and their limitations in capturing real-time, full-day behavior.
@@
 
-Impactful findings (Hailey write this paragraph)
-- individual differences in experiences
-- distribution of experiences (burstiness)
+The LENA® recorder is a commercial device that is worn by infants in a custom shirt pocket that has sufficient battery life and storage to record for an entire day.
+Closed-source LENA® algorithms analyze the audio recordings to provide counts of useful metrics, such as the number of words spoken by adults in the vicinity of the participant.
+Other long-form audio methods rely on custom-built recorders (Wass et al., 2022), apply alternative classification algorithms to LENA® data (Micheletti et al., 2022; Räsänen et al., 2020), or manually transcribe audio recorded by LENA® devices to improve accuracy or identify behaviors beyond the built-in categories (Bergelson, Casillas, et al., 2019; Mendoza \& Fausey, 2021).
+
+Long-form audio recordings have had a transformational impact on language development by allowing researchers to characterize opportunities for learning in daily life.
+Measuring the amount of speech heard by infants in the home (Weisleder \& Fernald, 2013) or in a daycare setting (Perry et al., 2018) revealed individual differences input that predict later vocabulary.
 Full-day language recording synchronized with other data sources allows researchers to identify how vocal input and production interacts with other processes.
-Linking LENA® speech measurements to repeated, time-stamped text-message surveys about infant device use revealed that infants heard less caregiver speech during moments that they were restrained in devices such as swings, exersaucers, and car seats (Malachowski et al., 2023).
+Beyond individual differences in aggregated data, long-form recordings can be used to determine the temporal schedule of experiences.
+For example, infants' daily experiences hearing music are clustered in time, with ``bursty'' episodes of hearing music separated by relatively long periods during which music is absent (Mendoza \& Fausey, 2022).
+Synchronizing audio recordings with other data sources extends researchers' ability to characterize daily experiences.
+Linking LENA® speech measurements to repeated, time-stamped text-message surveys about infant device placement revealed that infants heard less caregiver speech during moments that they were restrained in devices such as swings, exersaucers, and car seats (Malachowski et al., 2023).
 A custom-built wearable ECG and audio recorder allowed Wass et al.
 (2022) to discover that infant arousal increases the likelihood of infant vocalization across the day.
-In both examples, the ability to record real-time, synchronized data is vital for making inferences about processes that happen on the timescale of minutes or even seconds within in an individual as opposed to comparisons of aggregated data between infants.
 
-Key features
-- unobtrusive (Bergelson; CTL stuff)
-- mobile (no limits to where recording happens)
-- real time
-- long duration (capture entire day and heterogeniety in activities)
-- automatic classification with good-enough accuracy (scaling)
+We identified five key features of long-form audio methods to replicate in analagous studies of motor behavior.
+First, wearable audio recorders are \emph{mobile}.
+Measurement is not limited to a particular room because the recording device travels with the participant.
+Data are stored on device, so participants do not need to be in range of a receiver.
+Second, wearable audio recording is \emph{unobtrusive}.
+Participants' reactivity to observation, such as from a video camera, may influence behavior.
+For example, caregivers spoke more frequently to infants during a video-recorded portion of a home recording compared with audio-only segments captured by a LENA® device (Bergelson, Amatuni, et al., 2019).
+Third and fourth, recordings capture \emph{real-time data} over a \emph{full day}.
+The ability to record real-time data is vital for making inferences about processes that happen on the timescale of minutes or even seconds within in an individual as opposed to comparisons of aggregated data between infants.
+Synchronizing real-time data to other data streams helps to reveal sources of variability within an individual.
+Full-day recordings are essential for capturing experiences across the heterogeniety of daily routines that moderate behavior (e.g., play, feeding, errands) {[}CITE{]}.
+Burstiness of behavior means that long recordings are needed to capture clusters of events amid long periods in which they may be absent (Barbaro \& Fausey, 2022; Warlaumont et al., 2021).
+
+\begin{itemize}
+\tightlist
+\item
+  automatic classification with good-enough accuracy (scaling)
+\end{itemize}
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-BarbaroFausey2022}{}}%
+Barbaro, K.
+de, \& Fausey, C.
+M.
+(2022).
+Ten lessons about infants' everyday experiences.
+\emph{Current Directions in Psychological Science}, \emph{31}(1), 28--33.
+\url{https://doi.org/10.1177/09637214211059536}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Bergelson2019Input}{}}%
+Bergelson, E., Amatuni, A., Dailey, S., Koorathota, S., \& Tor, S. (2019). Day by day, hour by hour: Naturalistic language input to infants. \emph{Developmental Science}, \emph{22}, e12715.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Bergelson2019Corpus}{}}%
+Bergelson, E., Casillas, M., Soderstrom, M., Seidl, A., Warlaumont, A.
+S., \& Amatuni, A.
+(2019).
+What do {N}orth {A}merican babies hear?
+{A} large-scale cross-corpus analysis.
+\emph{Developmental Science}, \emph{22}, e12724.
+
 \leavevmode\vadjust pre{\hypertarget{ref-ChenSchneider2022}{}}%
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-MendozaFausey2021}{}}%
+Mendoza, J. K., \& Fausey, C. M. (2021). Everyday music in infancy. \emph{Developmental Science}, \emph{24}(6). \url{https://doi.org/10.1111/desc.13122}
+
+\leavevmode\vadjust pre{\hypertarget{ref-MendozaFausey2022}{}}%
+Mendoza, J.
+K., \& Fausey, C.
+M.
+(2022).
+Everyday parameters for episode-to-episode dynamics in the daily music of infancy.
+\emph{Cognitive Science}, \emph{46}(8).
+\url{https://doi.org/10.1111/cogs.13178}
+
+\leavevmode\vadjust pre{\hypertarget{ref-MichelettiYao2022}{}}%
+Micheletti, M., Yao, X., Johnson, M., \& Barbaro, K.
+de.
+(2022).
+Validating a model to detect infant crying from naturalistic audio.
+\emph{Behavior Research Methods}.
+\url{https://doi.org/10.3758/s13428-022-01961-x}
+
+\leavevmode\vadjust pre{\hypertarget{ref-PerryPrince2018}{}}%
+Perry, L.
+K., Prince, E.
+B., Valtierra, A.
+M., Rivero-Fernandez, C., Ullery, M.
+A., Katz, L.
+F., Laursen, B., \& Messinger, D.
+S.
+(2018).
+A year in words: The dynamics and consequences of language experiences in an intervention classroom.
+\emph{{PLOS} {ONE}}, \emph{13}(7), e0199893.
+\url{https://doi.org/10.1371/journal.pone.0199893}
+
+\leavevmode\vadjust pre{\hypertarget{ref-RasanenSeshadri2020}{}}%
+Räsänen, O., Seshadri, S., Lavechin, M., Cristia, A., \& Casillas, M.
+(2020).
+{ALICE}: An open-source tool for automatic measurement of phoneme, syllable, and word counts from child-centered daylong recordings.
+\emph{Behavior Research Methods}, \emph{53}(2), 818--835.
+\url{https://doi.org/10.3758/s13428-020-01460-x}
+
 \leavevmode\vadjust pre{\hypertarget{ref-SoskaAdolph2014}{}}%
@@
 Thurman, S. L., \& Corbetta, D. (2017). Spatial exploration and changes in infant-mother dyads around transitions in infant locomotion. \emph{Developmental Psychology}, \emph{53}, 1207--1221.
+
+\leavevmode\vadjust pre{\hypertarget{ref-WarlaumontSobowale2021}{}}%
+Warlaumont, A.
+S., Sobowale, K., \& Fausey, C.
+M.
+(2021).
+Daylong mobile audio recordings reveal multitimescale dynamics in infants' vocal productions and auditory experiences.
+\emph{Current Directions in Psychological Science}, \emph{31}(1), 12--19.
+\url{https://doi.org/10.1177/09637214211058166}
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-Weisleder2013}{}}%
+Weisleder, A., \& Fernald, A. (2013). Talking to children matters: Early language experience strengthens processing and builds vocabulary. \emph{Psychological Science}, \emph{24}, 2143--2152.
+
 \leavevmode\vadjust pre{\hypertarget{ref-WestIverson2021}{}}%

### 2023-02-24 manuscript.Rmd
more about LENA

@@
 Thus, characterizing individual differences in the day-to-day accumulation of body position experiences informs developmental theory by revealing differential opportunities for learning.
+[CLINICAL POSSIBILITIES?]
 <!-- Other west cites; cite new EMA -->
@@
 Burstiness of behavior means that long recordings are needed to capture clusters of events amid long periods in which they may be absent [@BarbaroFausey2022; @WarlaumontSobowale2021].
+Fifth, *automatic classification* means that the approach can scale to analyze large numbers of participants over long recordings without the bottleneck of manual annotation/transcription.
+Automatic classification can only replace human annotation if it is sufficiently accurate and unbiased.
+An independent assessment of the LENA&reg; algorithms found mixed results about the accuracy of different outcomes.
+For example, correlations between human transcribed counts of adult words and child vocalizations against LENA&reg;'s automatic counts were strong, *r* = .698 and *r* = .649, respectively [@CristiaLavechin2020].
+However, poor agreement was found for other metrics, such as the number of "conversational turns" between the child and communicative partners and the identification of male speakers.
 
-- automatic classification with good-enough accuracy (scaling)
+Thus, for some use cases (and for some metrics), long form audio recordings provide a mobile, unobtrusive way to automatically score real-time data over a full day.
+In the remainder of the paper, we turn to the question of how to replicate these qualities in long form recordings of motor behavior.
 
 ## Limitations of Video and Survey Methods
+
+Video and survey methods are the current state-of-the-art in assessing motor behavior.
+Although, each method has  advantages and disadvantages for characterizing infants' everyday motor experiences that complements the other, neither method on its own can provide comparable data to long form audio recordings as reviewed in the previous section.
 

### 2023-02-25 manuscript.Rmd
video and survey limitations

@@
 
-Video. Only advantage is real time
+Video observation is the most common way of measuring infant motor behavior in home recordings.
+Most often, an experimenter with a handheld camera follows infants from room to room to ensure that their movements are visible throughout the recording session [CITE ADOLPH, ANYONE ELSE?], but sometimes a fixed camera may be used [IVERSON?].
+The primary advantage of video recording is that it captures real-time behavior.
+Infants' body position, locomotion, and reaching comprise events that occur on the timescale of seconds, so standard video recording is adequate to score gross motor behavior.
+However, requiring an experimenter to pan a camera is obtrusive, whereas relying on a stationary camera means that infants will be absent from view as they move from place to place.
+Moreover, video observation cannot easily scale to long durations or large numbers of participants.
+Logistically, an experimenter cannot follow behind infants to record their behavior from morning to night (and were they to do so, they would likely alter infants' and caregivers' behavior).
+Typical video recording sessions last XX-XX, short of capturing the variety of activities across the full daily routine.
+Even if full day videos were available, the lack of suitable automatic classification tools means that the human cost of annotation makes it difficult for video methods to scale to testing large numbers of participants.
+Our annotation of 90 minutes of body position takes approximately XX-XX minutes to complete (depending on how often infants switch positions), meaning that a full "waking day" of approximately 11 hours for a 12-month-old [@GallandTaylor2012] would take XX hours to annotate.
 
-Survey. Everything but real time
+In contrast, survey methods such as daily diaries/inventories or ecological momentary assessment (EMA) are mobile, unobtrusive, can be applied across an entire day, and do not need laborious annotation.
+Diary studies provide caregivers with logs or structured interviews to estimate from memory how much time infants spend in particular activities [CITES].
+Ecological momentary assessment uses text-message or app-based notifications to prompt caregivers to make repeated estimates of infants' behavior over the course of a day [CITES].
+And although the responses are valuable in aggregate, the lack the real-time temporal resolution to describe moment-to-moment changes in behavior.
+For diaries and interviews, the memory limits mean that caregivers will report what was most frequent, but cannot remember events that happen on the level of seconds and minutes.
+Repeated EMA surveys might prompt caregivers to make hourly observations; increasing the number of surveys per day would be too burdensome for the respondent.
+Thus, despite being a useful tool for estimating broad developmental changes and/or individual differences in infants' motor experiences, survey methods are not suited for capturing within-participant temporal dynamics.
 

### 2023-02-25 manuscript.tex
video and survey limitations

@@
 Thus, characterizing individual differences in the day-to-day accumulation of body position experiences informs developmental theory by revealing differential opportunities for learning.
+{[}CLINICAL POSSIBILITIES?{]}
 
@@
 Burstiness of behavior means that long recordings are needed to capture clusters of events amid long periods in which they may be absent (Barbaro \& Fausey, 2022; Warlaumont et al., 2021).
+Fifth, \emph{automatic classification} means that the approach can scale to analyze large numbers of participants over long recordings without the bottleneck of manual annotation/transcription.
+Automatic classification can only replace human annotation if it is sufficiently accurate and unbiased.
+An independent assessment of the LENA® algorithms found mixed results about the accuracy of different outcomes.
+For example, correlations between human transcribed counts of adult words and child vocalizations against LENA®'s automatic counts were strong, \emph{r} = .698 and \emph{r} = .649, respectively (Cristia et al., 2020).
+However, poor agreement was found for other metrics, such as the number of ``conversational turns'' between the child and communicative partners and the identification of male speakers.
 
-\begin{itemize}
-\tightlist
-\item
-  automatic classification with good-enough accuracy (scaling)
-\end{itemize}
+Thus, for some use cases (and for some metrics), long form audio recordings provide a mobile, unobtrusive way to automatically score real-time data over a full day.
+In the remainder of the paper, we turn to the question of how to replicate these qualities in long form recordings of motor behavior.
 
@@
 
-Video. Only advantage is real time
+Video and survey methods are the current state-of-the-art in assessing motor behavior.
+Although, each method has advantages and disadvantages for characterizing infants' everyday motor experiences that complements the other, neither method on its own can provide comparable data to long form audio recordings as reviewed in the previous section.
 
-Survey. Everything but real time
+Video observation is the most common way of measuring infant motor behavior in home recordings.
+Most often, an experimenter with a handheld camera follows infants from room to room to ensure that their movements are visible throughout the recording session {[}CITE ADOLPH, ANYONE ELSE?{]}, but sometimes a fixed camera may be used {[}IVERSON?{]}.
+The primary advantage of video recording is that it captures real-time behavior.
+Infants' body position, locomotion, and reaching comprise events that occur on the timescale of seconds, so standard video recording is adequate to score gross motor behavior.
+However, requiring an experimenter to pan a camera is obtrusive, whereas relying on a stationary camera means that infants will be absent from view as they move from place to place.
+Moreover, video observation cannot easily scale to long durations or large numbers of participants.
+Logistically, an experimenter cannot follow behind infants to record their behavior from morning to night (and were they to do so, they would likely alter infants' and caregivers' behavior).
+Typical video recording sessions last XX-XX, short of capturing the variety of activities across the full daily routine.
+Even if full day videos were available, the lack of suitable automatic classification tools means that the human cost of annotation makes it difficult for video methods to scale to testing large numbers of participants.
+Our annotation of 90 minutes of body position takes approximately XX-XX minutes to complete (depending on how often infants switch positions), meaning that a full ``waking day'' of approximately 11 hours for a 12-month-old (Galland et al., 2012) would take XX hours to annotate.
+
+In contrast, survey methods such as daily diaries/inventories or ecological momentary assessment (EMA) are mobile, unobtrusive, can be applied across an entire day, and do not need laborious annotation.
+Diary studies provide caregivers with logs or structured interviews to estimate from memory how much time infants spend in particular activities {[}CITES{]}.
+Ecological momentary assessment uses text-message or app-based notifications to prompt caregivers to make repeated estimates of infants' behavior over the course of a day {[}CITES{]}.
+And although the responses are valuable in aggregate, the lack the real-time temporal resolution to describe moment-to-moment changes in behavior.
+For diaries and interviews, the memory limits mean that caregivers will report what was most frequent, but cannot remember events that happen on the level of seconds and minutes.
+Repeated EMA surveys might prompt caregivers to make hourly observations; increasing the number of surveys per day would be too burdensome for the respondent.
+Thus, despite being a useful tool for estimating broad developmental changes and/or individual differences in infants' motor experiences, survey methods are not suited for capturing within-participant temporal dynamics.
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-CristiaLavechin2020}{}}%
+Cristia, A., Lavechin, M., Scaff, C., Soderstrom, M., Rowland, C., Räsänen, O., Bunce, J., \& Bergelson, E.
+(2020).
+A thorough evaluation of the language environment analysis ({LENA}) system.
+\emph{Behavior Research Methods}, \emph{53}(2), 467--486.
+\url{https://doi.org/10.3758/s13428-020-01393-5}
+
 \leavevmode\vadjust pre{\hypertarget{ref-Survey}{}}%
@@
 Franchak, J. M., Kretch, K. S., \& Adolph, K. E. (2018). See and be seen: Infant-caregiver social looking during locomotor free play. \emph{Developmental Science}, \emph{21}, e12626.
+
+\leavevmode\vadjust pre{\hypertarget{ref-GallandTaylor2012}{}}%
+Galland, B.
+C., Taylor, B.
+J., Elder, D.
+E., \& Herbison, P.
+(2012).
+Normal sleep patterns in infants and children: A systematic review of observational studies.
+\emph{Sleep Medicine Reviews}, \emph{16}(3), 213--222.
+\url{https://doi.org/10.1016/j.smrv.2011.06.001}
 

### 2023-02-25 manuscript.Rmd
more introduction

@@
 
-Wear time is not an issue. Lots of Smith and other studies with long recordings.
+Measuring infant movement with inertial movement units (IMUs) is a promising avenue for long form recordings of motor behavior in the home.
+Lightweight sensors (XX-XXg) can be embedded in garments to make recordings fully *mobile*, and they are *unobtrusive* because they do not require a researcher's presence nor do they record sensitive audio/video that might influence participants' behavior.
+Many commerically-available and inexpensive IMUs have > 12 hour battery life an onboard storage to record *real-time*, *full-day* data of movements at a sampling rate that is higher than typical video (e.g., 50-100 Hz).
+Moreover, past work has successfully recorded the rate of leg kicks [SMITH CITE] and activity intensity [CITES] in infants and children across an entire day.
+
+The open question is whether *automatic classification* is sufficiently accurate to measure movement categories that are relevant to developmental and clinical research, and whether measurement validity is acceptable over long recording periods.
 

### 2023-02-25 manuscript.Rmd
Update manuscript.Rmd

@@
 Even if full day videos were available, the lack of suitable automatic classification tools means that the human cost of annotation makes it difficult for video methods to scale to testing large numbers of participants.
-Our annotation of 90 minutes of body position takes approximately XX-XX minutes to complete (depending on how often infants switch positions), meaning that a full "waking day" of approximately 11 hours for a 12-month-old [@GallandTaylor2012] would take XX hours to annotate.
+For annotation of body position takes approximately 2-5 hours to complete for every hour of video (depending on how often infants switch positions), meaning that a full "waking day" of approximately 11 hours for a 12-month-old [@GallandTaylor2012] could take 22-55 hours of labor to full annotate.
 
@@
 The open question is whether *automatic classification* is sufficiently accurate to measure movement categories that are relevant to developmental and clinical research, and whether measurement validity is acceptable over long recording periods.
+IMUs typically contain accelerometers that measure linear acceleration paired with gyroscopes that measure angular acceleration.
+Unlike motion tracking systems that might be used in a lab, IMUs do not provide data about the position of the body in space.
+Thus, data processing algorithms are needed to classify the raw sensor data (i.e., linear and angular acceleration timeseries) into meaningful categories (e.g., supine, prone, sitting, upright).
+The difficulty of the classification task depends on the categories of interest.
+More basic aspects of movement, such as leg kicks or overall activity intensity, can be identified by taking the magnitude of acceleration (irrespective of direction) and applying thresholds or cut-points to define when a movement has occurred at a particular intensity.
+
+Categorizing body position---supine on the back, prone on the belly, sitting, upright, or held off the ground by a caregiver---is too complex for cut-point definitions to be accurate.
+First, the magnitude of movement can vary greatly *within* a body position.
+An upright infant can be standing still or can be walking briskly across the room.
+A prone infant and be stationary in "tummy time" or crawling in myriad ways [ADOLPH].
+Moreover, the configuration of the arms, legs, and torso within a body position can vary greatly in everyday contexts.
+Infants can sit on the floor in a tripod position with support from an arm, in a "V" position with legs fully extended, in a "W" position with knees bent.
+Sitting on a caregivers' lap without the need to maintain balance means that the legs can dangle and the torso can lean in different directions.
+Sitting in a high chair or car seat reduces the magnitude of postural sway within sitting, and creates still more possibilities for how the arms and legs may move relative to the torso.
+Indeed, creating an all-encompassing set of rules for how to annotate sitting from video is no trivial task because of the various ways that sitting can occur in daily life.
 

### 2023-02-25 manuscript.Rmd
close to the end of the intro

@@
 
-Classification accuracy from ML is good. older papers; Airaksinen; Frontiers
+Thus, modern approaches to human activity recognition have used machine learning to classify activity categories based on features derived from IMU data in adults [CITEs], children [CITEs], and infants [CITEs].
+Synchronized video with ground-truth human annotations creates training data for a machine learning algorithm, such as a random forest models, that can be later used to predict categories from future IMU data that was not annotated.
+Although in the current investigation we focus on classification of body position, this approach can be extended to categorize other aspects of movement, such as locomotion [CITE] and OTHER EXAMPLES?
+Our openly-shared dataset [CITE] contains synchronized video and IMU data, allowing researchers to annotate and train models for behaviors beyond body position.
 
-What's missing
+REVIEW PAST WORK IN BODY POSITION CLASSIFICATION, END WITH WHAT IS MISSING.
 

### 2023-02-25 manuscript.tex
close to the end of the intro

@@
 Even if full day videos were available, the lack of suitable automatic classification tools means that the human cost of annotation makes it difficult for video methods to scale to testing large numbers of participants.
-Our annotation of 90 minutes of body position takes approximately XX-XX minutes to complete (depending on how often infants switch positions), meaning that a full ``waking day'' of approximately 11 hours for a 12-month-old (Galland et al., 2012) would take XX hours to annotate.
+For annotation of body position takes approximately 2-5 hours to complete for every hour of video (depending on how often infants switch positions), meaning that a full ``waking day'' of approximately 11 hours for a 12-month-old (Galland et al., 2012) could take 22-55 hours of labor to full annotate.
 
@@
 
-Wear time is not an issue. Lots of Smith and other studies with long recordings.
+Measuring infant movement with inertial movement units (IMUs) is a promising avenue for long form recordings of motor behavior in the home.
+Lightweight sensors (XX-XXg) can be embedded in garments to make recordings fully \emph{mobile}, and they are \emph{unobtrusive} because they do not require a researcher's presence nor do they record sensitive audio/video that might influence participants' behavior.
+Many commerically-available and inexpensive IMUs have \textgreater{} 12 hour battery life an onboard storage to record \emph{real-time}, \emph{full-day} data of movements at a sampling rate that is higher than typical video (e.g., 50-100 Hz).
+Moreover, past work has successfully recorded the rate of leg kicks {[}SMITH CITE{]} and activity intensity {[}CITES{]} in infants and children across an entire day.
 
-Classification accuracy from ML is good. older papers; Airaksinen; Frontiers
+The open question is whether \emph{automatic classification} is sufficiently accurate to measure movement categories that are relevant to developmental and clinical research, and whether measurement validity is acceptable over long recording periods.
+IMUs typically contain accelerometers that measure linear acceleration paired with gyroscopes that measure angular acceleration.
+Unlike motion tracking systems that might be used in a lab, IMUs do not provide data about the position of the body in space.
+Thus, data processing algorithms are needed to classify the raw sensor data (i.e., linear and angular acceleration timeseries) into meaningful categories (e.g., supine, prone, sitting, upright).
+The difficulty of the classification task depends on the categories of interest.
+More basic aspects of movement, such as leg kicks or overall activity intensity, can be identified by taking the magnitude of acceleration (irrespective of direction) and applying thresholds or cut-points to define when a movement has occurred at a particular intensity.
 
-What's missing
+Categorizing body position---supine on the back, prone on the belly, sitting, upright, or held off the ground by a caregiver---is too complex for cut-point definitions to be accurate.
+First, the magnitude of movement can vary greatly \emph{within} a body position.
+An upright infant can be standing still or can be walking briskly across the room.
+A prone infant and be stationary in ``tummy time'' or crawling in myriad ways {[}ADOLPH{]}.
+Moreover, the configuration of the arms, legs, and torso within a body position can vary greatly in everyday contexts.
+Infants can sit on the floor in a tripod position with support from an arm, in a ``V'' position with legs fully extended, in a ``W'' position with knees bent.
+Sitting on a caregivers' lap without the need to maintain balance means that the legs can dangle and the torso can lean in different directions.
+Sitting in a high chair or car seat reduces the magnitude of postural sway within sitting, and creates still more possibilities for how the arms and legs may move relative to the torso.
+Indeed, creating an all-encompassing set of rules for how to annotate sitting from video is no trivial task because of the various ways that sitting can occur in daily life.
+
+Thus, modern approaches to human activity recognition have used machine learning to classify activity categories based on features derived from IMU data in adults {[}CITEs{]}, children {[}CITEs{]}, and infants {[}CITEs{]}.
+Synchronized video with ground-truth human annotations creates training data for a machine learning algorithm, such as a random forest models, that can be later used to predict categories from future IMU data that was not annotated.
+Although in the current investigation we focus on classification of body position, this approach can be extended to categorize other aspects of movement, such as locomotion {[}CITE{]} and OTHER EXAMPLES?
+Our openly-shared dataset {[}CITE{]} contains synchronized video and IMU data, allowing researchers to annotate and train models for behaviors beyond body position.
+
+REVIEW PAST WORK IN BODY POSITION CLASSIFICATION, END WITH WHAT IS MISSING.
 

### 2023-02-28 manuscript.Rmd
intro

@@
 Measuring infant movement with inertial movement units (IMUs) is a promising avenue for long form recordings of motor behavior in the home.
-Lightweight sensors (XX-XXg) can be embedded in garments to make recordings fully *mobile*, and they are *unobtrusive* because they do not require a researcher's presence nor do they record sensitive audio/video that might influence participants' behavior.
+Lightweight sensors (XX-XXg) can be embedded in garments to make recordings fully *mobile*, and they are *unobtrusive* because they do not require a researcher's presence nor do they record sensitive audio/video that might influence participants' behavior [(Cliff et al., 2009; de Barbaro, 2019; Lobo et al., 2019; Bruijns et al., 2020)].
 Many commerically-available and inexpensive IMUs have > 12 hour battery life an onboard storage to record *real-time*, *full-day* data of movements at a sampling rate that is higher than typical video (e.g., 50-100 Hz).
-Moreover, past work has successfully recorded the rate of leg kicks [SMITH CITE] and activity intensity [CITES] in infants and children across an entire day.
+Moreover, past work has successfully recorded the rate of leg kicks [(Smith et al., 2017; Jiang et al., 2018)] and activity intensity [Hagenbuchner et al., 2015; Trost et al., 2018] in infants and children across an entire day.
+<!-- CHECK THAT CITATIONS ARE FULL DAY -->
 
@@
 The difficulty of the classification task depends on the categories of interest.
-More basic aspects of movement, such as leg kicks or overall activity intensity, can be identified by taking the magnitude of acceleration (irrespective of direction) and applying thresholds or cut-points to define when a movement has occurred at a particular intensity.
+More basic aspects of movement, such as overall activity intensity, can be identified by taking the magnitude of acceleration (irrespective of direction) and applying thresholds or cut-points to define when a movement has occurred at a particular intensity [(Trost et al., 2012; Kuzik et al., 2015; Hager et al., 2017; Armstrong et al., 2019)].
 
@@
 Indeed, creating an all-encompassing set of rules for how to annotate sitting from video is no trivial task because of the various ways that sitting can occur in daily life.
+Finally, caregivers frequently pick up and transport infants, creating motion signals that need to be differentiated from independent activity [Kwon].
+GREENSPAN...
 
-Thus, modern approaches to human activity recognition have used machine learning to classify activity categories based on features derived from IMU data in adults [CITEs], children [CITEs], and infants [CITEs].
+Thus, modern approaches to human activity recognition have used machine learning to classify activity categories based on features derived from IMU data in adults [(Preece et al., 2009; Arif and Kattan, 2015)], children [(Nam and Park, 2013; Zhao et al., 2013; Ren et al., 2016; Stewart et al., 2018)], and infants [Franchak; Airaksinen; Yao].
 Synchronized video with ground-truth human annotations creates training data for a machine learning algorithm, such as a random forest models, that can be later used to predict categories from future IMU data that was not annotated.
 Although in the current investigation we focus on classification of body position, this approach can be extended to categorize other aspects of movement, such as locomotion [CITE] and OTHER EXAMPLES?
-Our openly-shared dataset [CITE] contains synchronized video and IMU data, allowing researchers to annotate and train models for behaviors beyond body position.
+Accompanying our paper, we have created a database [CITE] containing synchronized video and IMU data, allowing researchers to annotate and train models for behaviors beyond body position.
 
-REVIEW PAST WORK IN BODY POSITION CLASSIFICATION, END WITH WHAT IS MISSING.
+Three prior investigations have used machine learning to categorize infant body position from IMUs towards the goal of collecting full-day data, and each has yielded encouraging validation results.
+@AiraksinenRasanen2020 tested 4- to 8-month-olds in a laboratory visit, and found 95% accuracy in distinguishing between body position categories.
+Using a wider age range of XX-XX, [FRANCHAK] found XX% accuracy in a laboratory validation study in categorizing body position.
 

### 2023-02-28 manuscript.Rmd
more introduction

@@
 Three prior investigations have used machine learning to categorize infant body position from IMUs towards the goal of collecting full-day data, and each has yielded encouraging validation results.
-@AiraksinenRasanen2020 tested 4- to 8-month-olds in a laboratory visit, and found 95% accuracy in distinguishing between body position categories.
-Using a wider age range of XX-XX, [FRANCHAK] found XX% accuracy in a laboratory validation study in categorizing body position.
+@AiraksinenRasanen2020 tested 4- to 8-month-olds in a laboratory visit, and found XX% accuracy (kappa = XX%) in distinguishing between body position categories that crawling infants could perform on the floor (excluding times that infants were held by caregivers).
+Using a wider age range of XX-XX, [FRANCHAK] found XX% accuracy (kappa = XX%) in a laboratory validation study in categorizing body position that included infants who could both crawl and walk and also categorized caregiver holding.
+Most recently, [Airaksinen NEW] conducted a validation study of body position classification in either a home or clinic testing in 4- to 19-month-olds, refining their previous method to detect moments that infants were carried by caregivers.
+Classification accuracy did not vary between lab and home settings, and was generally high (95%, kappa = .93).
+However, in all three studies accuracy was assessed within a brief (15-60 minute) session supervised by a researcher, leaving the open question of how well body position classification will scale to testing across an entire day of natural home life.
+YAO?
 
 ## Goals of the Current Study
+
+Accordingly, the overarching goal of the current study is to test the feasibility and validity of long-form body position recording in the home during unsupervised, everyday behavior.
+During a supervised recording, whether in the home or in the lab, the researcher can set up the situation to encourage or restrict certain behaviors and will usually instruct the caregiver to play with the infant.
+Within-day variations in activities (e.g., eating lunch in a high chair) create challenging situations for applying automated classification of body position.
+Will models trained on video recorded observations at the beginning of the day generalize to predict later behavior?
 

### 2023-03-01 manuscript.Rmd
current study draft

@@
 Within-day variations in activities (e.g., eating lunch in a high chair) create challenging situations for applying automated classification of body position.
-Will models trained on video recorded observations at the beginning of the day generalize to predict later behavior?
+Will models trained on video-recorded observations at the beginning of the day generalize to predict behavior at a later time?
+Assessing the validity of temporally *distal* periods is a crucial step to establish whether automatic classification can be used to measure body position across a day.
+
+In the current study, we report the feasibility and validity of body position classification over the full day in the home based on XX testing sessions from XX infants aged 4-14 months.
+Participants received a custom pair of infant leggings embedded with 4 IMUs (one on each ankle and one on each hip) and a video camera to collect ground truth data about infant body position.
+A *proximal comparison* period began when participants received the equipment and completed a guided phone call during which caregivers were asked to elicit different body positions based on prompts from the experimenter.
+Although not directly supervised, this period was most similar to previous recordings because it occurred during a convenient time for the infant and caregiver to play and receive instruction.
+The **first goal** of the current study was to determine the accuracy of body position classification during the proximal comparison period using our novel, semi-supervised procedure in participants' homes.
+
+A second, *distal comparison* period followed the proximal comparison period and captured approximately 90 minutes of home behavior without any input from the experimenter.
+Following a long delay from the start of the study when caregivers were guided in placing the equipment, it is possible that accuracy could become worse if caregivers or infants moved the garment or sensors.
+Moreover, increasing variation in everyday activities creates a challenge for classification.
+Thus, the **second goal** of our study was to assess accuracy during the distal comparison.
+
+After this second video recording period, we asked caregivers to have infants continue to wear IMUs for the rest of the day until they went to bed, creating the first real-time, full-day dataset of infant body position.
+Interpreting such data required caregivers to log when infants napped, when they removed the sensor garment for diaper changes or other reasons, and when infants went to bed at the end of the day.
+The **third goal** of the study was to examine the quality of the full-day data.
+Could body position be successfully recorded over the desired period?
+Finally, the novel full-day dataset affords us a unique chance to ask whether the estimated time infants spent in different body positions align with past results using video and survey methods.
+Thus, the **fourth goal** of the current study was to determine whether our full-day body position estimates yielded expected age-related differences in everyday body position.
+Based on past results, with age infants should spend increasingly more time sitting and upright but less time supine.
 

### 2023-03-01 manuscript.tex
current study draft

@@
 Measuring infant movement with inertial movement units (IMUs) is a promising avenue for long form recordings of motor behavior in the home.
-Lightweight sensors (XX-XXg) can be embedded in garments to make recordings fully \emph{mobile}, and they are \emph{unobtrusive} because they do not require a researcher's presence nor do they record sensitive audio/video that might influence participants' behavior.
+Lightweight sensors (XX-XXg) can be embedded in garments to make recordings fully \emph{mobile}, and they are \emph{unobtrusive} because they do not require a researcher's presence nor do they record sensitive audio/video that might influence participants' behavior {[}(Cliff et al., 2009; de Barbaro, 2019; Lobo et al., 2019; Bruijns et al., 2020){]}.
 Many commerically-available and inexpensive IMUs have \textgreater{} 12 hour battery life an onboard storage to record \emph{real-time}, \emph{full-day} data of movements at a sampling rate that is higher than typical video (e.g., 50-100 Hz).
-Moreover, past work has successfully recorded the rate of leg kicks {[}SMITH CITE{]} and activity intensity {[}CITES{]} in infants and children across an entire day.
+Moreover, past work has successfully recorded the rate of leg kicks {[}(Smith et al., 2017; Jiang et al., 2018){]} and activity intensity {[}Hagenbuchner et al., 2015; Trost et al., 2018{]} in infants and children across an entire day.
 
@@
 The difficulty of the classification task depends on the categories of interest.
-More basic aspects of movement, such as leg kicks or overall activity intensity, can be identified by taking the magnitude of acceleration (irrespective of direction) and applying thresholds or cut-points to define when a movement has occurred at a particular intensity.
+More basic aspects of movement, such as overall activity intensity, can be identified by taking the magnitude of acceleration (irrespective of direction) and applying thresholds or cut-points to define when a movement has occurred at a particular intensity {[}(Trost et al., 2012; Kuzik et al., 2015; Hager et al., 2017; Armstrong et al., 2019){]}.
 
@@
 Indeed, creating an all-encompassing set of rules for how to annotate sitting from video is no trivial task because of the various ways that sitting can occur in daily life.
+Finally, caregivers frequently pick up and transport infants, creating motion signals that need to be differentiated from independent activity {[}Kwon{]}.
+GREENSPAN\ldots{}
 
-Thus, modern approaches to human activity recognition have used machine learning to classify activity categories based on features derived from IMU data in adults {[}CITEs{]}, children {[}CITEs{]}, and infants {[}CITEs{]}.
+Thus, modern approaches to human activity recognition have used machine learning to classify activity categories based on features derived from IMU data in adults {[}(Preece et al., 2009; Arif and Kattan, 2015){]}, children {[}(Nam and Park, 2013; Zhao et al., 2013; Ren et al., 2016; Stewart et al., 2018){]}, and infants {[}Franchak; Airaksinen; Yao{]}.
 Synchronized video with ground-truth human annotations creates training data for a machine learning algorithm, such as a random forest models, that can be later used to predict categories from future IMU data that was not annotated.
 Although in the current investigation we focus on classification of body position, this approach can be extended to categorize other aspects of movement, such as locomotion {[}CITE{]} and OTHER EXAMPLES?
-Our openly-shared dataset {[}CITE{]} contains synchronized video and IMU data, allowing researchers to annotate and train models for behaviors beyond body position.
+Accompanying our paper, we have created a database {[}CITE{]} containing synchronized video and IMU data, allowing researchers to annotate and train models for behaviors beyond body position.
 
-REVIEW PAST WORK IN BODY POSITION CLASSIFICATION, END WITH WHAT IS MISSING.
+Three prior investigations have used machine learning to categorize infant body position from IMUs towards the goal of collecting full-day data, and each has yielded encouraging validation results.
+Airaksinen et al.
+(2020) tested 4- to 8-month-olds in a laboratory visit, and found XX\% accuracy (kappa = XX\%) in distinguishing between body position categories that crawling infants could perform on the floor (excluding times that infants were held by caregivers).
+Using a wider age range of XX-XX, {[}FRANCHAK{]} found XX\% accuracy (kappa = XX\%) in a laboratory validation study in categorizing body position that included infants who could both crawl and walk and also categorized caregiver holding.
+Most recently, {[}Airaksinen NEW{]} conducted a validation study of body position classification in either a home or clinic testing in 4- to 19-month-olds, refining their previous method to detect moments that infants were carried by caregivers.
+Classification accuracy did not vary between lab and home settings, and was generally high (95\%, kappa = .93).
+However, in all three studies accuracy was assessed within a brief (15-60 minute) session supervised by a researcher, leaving the open question of how well body position classification will scale to testing across an entire day of natural home life.
+YAO?
 
@@
 \subsection{Goals of the Current Study}\label{goals-of-the-current-study}}
+
+Accordingly, the overarching goal of the current study is to test the feasibility and validity of long-form body position recording in the home during unsupervised, everyday behavior.
+During a supervised recording, whether in the home or in the lab, the researcher can set up the situation to encourage or restrict certain behaviors and will usually instruct the caregiver to play with the infant.
+Within-day variations in activities (e.g., eating lunch in a high chair) create challenging situations for applying automated classification of body position.
+Will models trained on video-recorded observations at the beginning of the day generalize to predict behavior at a later time?
+Assessing the validity of temporally \emph{distal} periods is a crucial step to establish whether automatic classification can be used to measure body position across a day.
+
+In the current study, we report the feasibility and validity of body position classification over the full day in the home based on XX testing sessions from XX infants aged 4-14 months.
+Participants received a custom pair of infant leggings embedded with 4 IMUs (one on each ankle and one on each hip) and a video camera to collect ground truth data about infant body position.
+A \emph{proximal comparison} period began when participants received the equipment and completed a guided phone call during which caregivers were asked to elicit different body positions based on prompts from the experimenter.
+Although not directly supervised, this period was most similar to previous recordings because it occurred during a convenient time for the infant and caregiver to play and receive instruction.
+The \textbf{first goal} of the current study was to determine the accuracy of body position classification during the proximal comparison period using our novel, semi-supervised procedure in participants' homes.
+
+A second, \emph{distal comparison} period followed the proximal comparison period and captured approximately 90 minutes of home behavior without any input from the experimenter.
+Following a long delay from the start of the study when caregivers were guided in placing the equipment, it is possible that accuracy could become worse if caregivers or infants moved the garment or sensors.
+Moreover, increasing variation in everyday activities creates a challenge for classification.
+Thus, the \textbf{second goal} of our study was to assess accuracy during the distal comparison.
+
+After this second video recording period, we asked caregivers to have infants continue to wear IMUs for the rest of the day until they went to bed, creating the first real-time, full-day dataset of infant body position.
+Interpreting such data required caregivers to log when infants napped, when they removed the sensor garment for diaper changes or other reasons, and when infants went to bed at the end of the day.
+The \textbf{third goal} of the study was to examine the quality of the full-day data.
+Could body position be successfully recorded over the desired period?
+Finally, the novel full-day dataset affords us a unique chance to ask whether the estimated time infants spent in different body positions align with past results using video and survey methods.
+Thus, the \textbf{fourth goal} of the current study was to determine whether our full-day body position estimates yielded expected age-related differences in everyday body position.
+Based on past results, with age infants should spend increasingly more time sitting and upright but less time supine.
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-AiraksinenRasanen2020}{}}%
+Airaksinen, M., Räsänen, O., Ilén, E., Häyrinen, T., Kivi, A., Marchi, V., Gallen, A., Blom, S., Varhe, A., Kaartinen, N., et al.
+(2020).
+Automatic posture and movement tracking of infants with wearable movement sensors.
+\emph{Scientific Reports}, \emph{10}(1), 1--13.
+
 \leavevmode\vadjust pre{\hypertarget{ref-BarbaroFausey2022}{}}%

### 2023-03-02 manuscript.Rmd
line edits

@@
 
-Infants' movements facilitate and constrain how they interact with their surroundings from moment to moment.
+From moment to moment, infants' movements facilitate and constrain how they can interact with their surroundings.
 Changes in *body position*---whether infants are supine on their backs, prone on their bellies, sitting, upright, or held by a  caregiver---have immediate consequences for vision, object exploration, and social interaction.
@@
 
-In this paper, we present an inertial sensing method to classify infant body position from moment-to-moment across an entire day, and validate its accuracy using XX hours of video recorded across XX in-home data collection sessions.
+In this paper, we present an inertial sensing method to classify infant body position from moment-to-moment across an entire day, and validate its accuracy using over 100 hours of video recorded across 35 in-home data collection sessions.
 Our method takes inspiration from a more mature technology: Long-form audio recordings of infants' language experiences using wearable audio recorders.
-We begin by describing the impact of long-form audio recordings on research in developmental psychology, and identify the key features that should be replicated in long-form recordings of motor behavior [CITES].
+We begin by describing the impact of long-form audio recordings on research in developmental psychology [CITES], and identify the key features that should be replicated in long-form recordings of motor behavior.
 Next, we review the current state-of-the-art in measuring infant motor behavior---video and survey data---and their limitations in capturing real-time, full-day behavior.
 Finally, we discuss the advantages of using inertial sensing to classify motor behavior.
-Despite promising past results in brief sessions [CITES], the current investigation takes a needed step forward by testing accuracy over long recordings.
+Despite promising past results in brief, supervised sessions [CITES], the current investigation takes a needed step forward by testing accuracy over long, unsupervised recordings.
 
@@
 The LENA&reg; recorder is a commercial device that is worn by infants in a custom shirt pocket that has sufficient battery life and storage to record for an entire day.
-Closed-source LENA&reg; algorithms analyze the audio recordings to provide counts of useful metrics, such as the number of words spoken by adults in the vicinity of the participant.
+Closed-source LENA&reg; algorithms analyze the audio recordings to provide automatic counts of useful metrics, such as the number of words spoken by adults in the vicinity of the participant.
 Other long-form audio methods rely on custom-built recorders [@WassPhillips2022], apply alternative classification algorithms to LENA&reg; data [@RasanenSeshadri2020; @MichelettiYao2022], or manually transcribe audio recorded by LENA&reg; devices to improve accuracy or identify behaviors beyond the built-in categories [@Bergelson2019Corpus; @MendozaFausey2021].
 
-Long-form audio recordings have had a transformational impact on language development by allowing researchers to characterize opportunities for learning in daily life.
+Long-form audio recordings have had a transformational impact on language development research by allowing researchers to characterize opportunities for learning in daily life.
 Measuring the amount of speech heard by infants in the home [@Weisleder2013] or in a daycare setting [@PerryPrince2018] revealed individual differences input that predict later vocabulary.
-Full-day language recording synchronized with other data sources allows researchers to identify how vocal input and production interacts with other processes.
+Full-day language recording synchronized with other data sources allows researchers to identify how auditory input and vocal production interact with other processes.
 Beyond individual differences in aggregated data, long-form recordings can be used to determine the temporal schedule of experiences.
@@
 
-We identified five key features of long-form audio methods to replicate in analagous studies of motor behavior.
+We identified five key features of long-form audio methods that should be replicated in analogous studies of motor behavior.
 First, wearable audio recorders are *mobile*.
 Measurement is not limited to a particular room because the recording device travels with the participant.
-Data are stored on device, so participants do not need to be in range of a receiver.
+Data are recorded to onboard device memory, so participants do not need to be in range of a receiver.
 Second, wearable audio recording is *unobtrusive*.
@@
 The ability to record real-time data is vital for making inferences about processes that happen on the timescale of minutes or even seconds within in an individual as opposed to comparisons of aggregated data between infants.
-Synchronizing real-time data to other data streams helps to reveal sources of variability within an individual.
+Synchronizing real-time data to other data streams helps to reveal sources of variability within an individual [e.g., @WassPhillips2022].
 Full-day recordings are essential for capturing experiences across the heterogeniety of daily routines that moderate behavior (e.g., play, feeding, errands) [CITE].
-Burstiness of behavior means that long recordings are needed to capture clusters of events amid long periods in which they may be absent [@BarbaroFausey2022; @WarlaumontSobowale2021].
+"Burstiness" of behavior means that long recordings are needed to capture clusters of events amid long periods in which they may be absent [@BarbaroFausey2022; @WarlaumontSobowale2021].
 Fifth, *automatic classification* means that the approach can scale to analyze large numbers of participants over long recordings without the bottleneck of manual annotation/transcription.
@@
 Video and survey methods are the current state-of-the-art in assessing motor behavior.
-Although, each method has  advantages and disadvantages for characterizing infants' everyday motor experiences that complements the other, neither method on its own can provide comparable data to long form audio recordings as reviewed in the previous section.
+Although, each method has  advantages and disadvantages for characterizing infants' everyday motor experiences that complements the other, neither method on its own can provide comparable data to the long form audio recordings reviewed in the previous section.
 
@@
 Infants' body position, locomotion, and reaching comprise events that occur on the timescale of seconds, so standard video recording is adequate to score gross motor behavior.
-However, requiring an experimenter to pan a camera is obtrusive, whereas relying on a stationary camera means that infants will be absent from view as they move from place to place.
+However, requiring an experimenter to operate a camera is obtrusive, whereas relying on a stationary camera means that infants will be absent from view as they move from place to place.
 Moreover, video observation cannot easily scale to long durations or large numbers of participants.
@@
 Even if full day videos were available, the lack of suitable automatic classification tools means that the human cost of annotation makes it difficult for video methods to scale to testing large numbers of participants.
-For annotation of body position takes approximately 2-5 hours to complete for every hour of video (depending on how often infants switch positions), meaning that a full "waking day" of approximately 11 hours for a 12-month-old [@GallandTaylor2012] could take 22-55 hours of labor to full annotate.
+Our annotation of body position takes approximately 2-5 hours to complete for every hour of video (depending on how often infants switch positions), meaning that a full "waking day" of approximately 11 hours for a 12-month-old [@GallandTaylor2012] could take 22-55 hours of labor to full annotate.
 
@@
 Ecological momentary assessment uses text-message or app-based notifications to prompt caregivers to make repeated estimates of infants' behavior over the course of a day [CITES].
-And although the responses are valuable in aggregate, the lack the real-time temporal resolution to describe moment-to-moment changes in behavior.
-For diaries and interviews, the memory limits mean that caregivers will report what was most frequent, but cannot remember events that happen on the level of seconds and minutes.
-Repeated EMA surveys might prompt caregivers to make hourly observations; increasing the number of surveys per day would be too burdensome for the respondent.
-Thus, despite being a useful tool for estimating broad developmental changes and/or individual differences in infants' motor experiences, survey methods are not suited for capturing within-participant temporal dynamics.
+And although the responses are valuable in aggregate, survey methods lack the real-time temporal resolution to describe moment-to-moment changes in behavior.
+For diaries and interviews, limits on caregivers' memory mean that they will report what was most frequent, but cannot remember events that happen on the scale of seconds and minutes.
+At best, EMA surveys prompt caregivers to make hourly observations; increasing the number of surveys per day would be too burdensome for the respondent.
+Thus, despite being a useful tool for estimating broad developmental changes and individual differences in infants' motor experiences, survey methods are not suited for capturing within-participant temporal dynamics.
 

### 2023-03-02 manuscript.Rmd
intro revisions

@@
 Despite promising past results in brief, supervised sessions [CITES], the current investigation takes a needed step forward by testing accuracy over long, unsupervised recordings.
+Although in the current investigation we focus on classification of body position, this approach can be extended to categorize other aspects of movement, such as locomotion [CITE] and OTHER EXAMPLES?
+Accompanying our paper, we have created a database [CITE] containing synchronized video and IMU data, allowing researchers to annotate and train models for behaviors beyond body position.
 
@@
 Lightweight sensors (XX-XXg) can be embedded in garments to make recordings fully *mobile*, and they are *unobtrusive* because they do not require a researcher's presence nor do they record sensitive audio/video that might influence participants' behavior [(Cliff et al., 2009; de Barbaro, 2019; Lobo et al., 2019; Bruijns et al., 2020)].
-Many commerically-available and inexpensive IMUs have > 12 hour battery life an onboard storage to record *real-time*, *full-day* data of movements at a sampling rate that is higher than typical video (e.g., 50-100 Hz).
+Many commercially-available and inexpensive IMUs have > 12 hour battery life with onboard storage to record *real-time*, *full-day* motion data at a sampling rate that is higher than typical video (e.g., 50-100 Hz).
 Moreover, past work has successfully recorded the rate of leg kicks [(Smith et al., 2017; Jiang et al., 2018)] and activity intensity [Hagenbuchner et al., 2015; Trost et al., 2018] in infants and children across an entire day.
@@
 Sitting on a caregivers' lap without the need to maintain balance means that the legs can dangle and the torso can lean in different directions.
-Sitting in a high chair or car seat reduces the magnitude of postural sway within sitting, and creates still more possibilities for how the arms and legs may move relative to the torso.
+Sitting in a high chair or car seat reduces the magnitude of postural sway within sitting, and creates even more possibilities for how the arms and legs may move relative to the torso.
 Indeed, creating an all-encompassing set of rules for how to annotate sitting from video is no trivial task because of the various ways that sitting can occur in daily life.
 Finally, caregivers frequently pick up and transport infants, creating motion signals that need to be differentiated from independent activity [Kwon].
+Thus, modern approaches to human activity recognition have used machine learning to classify activity categories based on features derived from IMU data in adults [(Preece et al., 2009; Arif and Kattan, 2015)], children [(Nam and Park, 2013; Zhao et al., 2013; Ren et al., 2016; Stewart et al., 2018)], and infants [Franchak; Airaksinen; Yao].
+Synchronized video with ground-truth human annotations creates training data for a machine learning algorithm, such as a random forest model, that can be later used to predict categories from IMU data that was not annotated.
+Crucially, this allows automatic classification to scale to full day recording by relying on a relatively smaller set of video annotation.
 GREENSPAN...
 
-Thus, modern approaches to human activity recognition have used machine learning to classify activity categories based on features derived from IMU data in adults [(Preece et al., 2009; Arif and Kattan, 2015)], children [(Nam and Park, 2013; Zhao et al., 2013; Ren et al., 2016; Stewart et al., 2018)], and infants [Franchak; Airaksinen; Yao].
-Synchronized video with ground-truth human annotations creates training data for a machine learning algorithm, such as a random forest models, that can be later used to predict categories from future IMU data that was not annotated.
-Although in the current investigation we focus on classification of body position, this approach can be extended to categorize other aspects of movement, such as locomotion [CITE] and OTHER EXAMPLES?
-Accompanying our paper, we have created a database [CITE] containing synchronized video and IMU data, allowing researchers to annotate and train models for behaviors beyond body position.
-
-Three prior investigations have used machine learning to categorize infant body position from IMUs towards the goal of collecting full-day data, and each has yielded encouraging validation results.
+Three prior investigations have used machine learning to categorize infant body position from IMUs towards the goal of collecting full-day data.
 @AiraksinenRasanen2020 tested 4- to 8-month-olds in a laboratory visit, and found XX% accuracy (kappa = XX%) in distinguishing between body position categories that crawling infants could perform on the floor (excluding times that infants were held by caregivers).
@@
 Classification accuracy did not vary between lab and home settings, and was generally high (95%, kappa = .93).
-However, in all three studies accuracy was assessed within a brief (15-60 minute) session supervised by a researcher, leaving the open question of how well body position classification will scale to testing across an entire day of natural home life.
+Although all three studies yielded promising classification accuracy, accuracy was assessed in brief (15-60 minute) sessions supervised by a researcher, leaving the open question of how well body position classification will scale to testing across an entire day of natural home life.
 YAO?
@@
 Accordingly, the overarching goal of the current study is to test the feasibility and validity of long-form body position recording in the home during unsupervised, everyday behavior.
-During a supervised recording, whether in the home or in the lab, the researcher can set up the situation to encourage or restrict certain behaviors and will usually instruct the caregiver to play with the infant.
-Within-day variations in activities (e.g., eating lunch in a high chair) create challenging situations for applying automated classification of body position.
+Supervised recordings from past work [CITES], whether in the home or in the lab, let researchers set up the situation to encourage or restrict certain behaviors.
+Usually, caregivers are asked to play with the infant.
+However, across a real day, non-play activities (e.g., eating lunch in a high chair) create challenging situations for applying automated classification of body position.
 Will models trained on video-recorded observations at the beginning of the day generalize to predict behavior at a later time?
@@
 
-In the current study, we report the feasibility and validity of body position classification over the full day in the home based on XX testing sessions from XX infants aged 4-14 months.
+In the current study, we report the feasibility and validity of body position classification over the full day in the home based on 35 testing sessions from XX infants aged 4-14 months.
 Participants received a custom pair of infant leggings embedded with 4 IMUs (one on each ankle and one on each hip) and a video camera to collect ground truth data about infant body position.
 A *proximal comparison* period began when participants received the equipment and completed a guided phone call during which caregivers were asked to elicit different body positions based on prompts from the experimenter.
-Although not directly supervised, this period was most similar to previous recordings because it occurred during a convenient time for the infant and caregiver to play and receive instruction.
-The **first goal** of the current study was to determine the accuracy of body position classification during the proximal comparison period using our novel, semi-supervised procedure in participants' homes.
+Although not directly supervised, this period was most similar to previous recordings because it occurred during a convenient time for the infant and caregiver to play while they received instructions from the experimenter.
+The **first goal** of the current study was to determine the accuracy of body position classification during the proximal comparison period using this novel, semi-supervised procedure in participants' homes.
 
-A second, *distal comparison* period followed the proximal comparison period and captured approximately 90 minutes of home behavior without any input from the experimenter.
-Following a long delay from the start of the study when caregivers were guided in placing the equipment, it is possible that accuracy could become worse if caregivers or infants moved the garment or sensors.
-Moreover, increasing variation in everyday activities creates a challenge for classification.
+A second, *distal comparison* period followed the proximal comparison period and captured approximately 90 minutes of home behavior that was completely unsupervised.
+Caregivers and infants could (and did) do whatever they wished.
+Because this recording happened a considerable amount of time after the initial setup and instructions from the experimenter, accuracy could decline if caregivers or infants moved the garment or sensors.
+Moreover, increasing variation in everyday activities during the distal comparison creates a greater challenge for classification, testing whether machine learning models can generalize to novel test cases.
 Thus, the **second goal** of our study was to assess accuracy during the distal comparison.
 
-After this second video recording period, we asked caregivers to have infants continue to wear IMUs for the rest of the day until they went to bed, creating the first real-time, full-day dataset of infant body position.
+After this second video recording period, we asked caregivers to keep infants wearing IMUs for the rest of the day until their regular bedtime, creating the **first real-time, full-day dataset of infant body position**.
 Interpreting such data required caregivers to log when infants napped, when they removed the sensor garment for diaper changes or other reasons, and when infants went to bed at the end of the day.
@@
 Finally, the novel full-day dataset affords us a unique chance to ask whether the estimated time infants spent in different body positions align with past results using video and survey methods.
-Thus, the **fourth goal** of the current study was to determine whether our full-day body position estimates yielded expected age-related differences in everyday body position.
-Based on past results, with age infants should spend increasingly more time sitting and upright but less time supine.
+Thus, the **fourth goal** of the current study was to determine whether full-day body position measurements conformed to expectations about age differences in body position.
+Based on past results [CITE], infants should spend increasingly more time sitting and upright but less time supine over the age range tested (4 to 14 months).
 

### 2023-03-02 manuscript.tex
intro revisions

@@
 
-Infants' movements facilitate and constrain how they interact with their surroundings from moment to moment.
+From moment to moment, infants' movements facilitate and constrain how they can interact with their surroundings.
 Changes in \emph{body position}---whether infants are supine on their backs, prone on their bellies, sitting, upright, or held by a caregiver---have immediate consequences for vision, object exploration, and social interaction.
@@
 
-In this paper, we present an inertial sensing method to classify infant body position from moment-to-moment across an entire day, and validate its accuracy using XX hours of video recorded across XX in-home data collection sessions.
+In this paper, we present an inertial sensing method to classify infant body position from moment-to-moment across an entire day, and validate its accuracy using over 100 hours of video recorded across 35 in-home data collection sessions.
 Our method takes inspiration from a more mature technology: Long-form audio recordings of infants' language experiences using wearable audio recorders.
-We begin by describing the impact of long-form audio recordings on research in developmental psychology, and identify the key features that should be replicated in long-form recordings of motor behavior {[}CITES{]}.
+We begin by describing the impact of long-form audio recordings on research in developmental psychology {[}CITES{]}, and identify the key features that should be replicated in long-form recordings of motor behavior.
 Next, we review the current state-of-the-art in measuring infant motor behavior---video and survey data---and their limitations in capturing real-time, full-day behavior.
 Finally, we discuss the advantages of using inertial sensing to classify motor behavior.
-Despite promising past results in brief sessions {[}CITES{]}, the current investigation takes a needed step forward by testing accuracy over long recordings.
+Despite promising past results in brief, supervised sessions {[}CITES{]}, the current investigation takes a needed step forward by testing accuracy over long, unsupervised recordings.
+Although in the current investigation we focus on classification of body position, this approach can be extended to categorize other aspects of movement, such as locomotion {[}CITE{]} and OTHER EXAMPLES?
+Accompanying our paper, we have created a database {[}CITE{]} containing synchronized video and IMU data, allowing researchers to annotate and train models for behaviors beyond body position.
 
@@
 The LENA® recorder is a commercial device that is worn by infants in a custom shirt pocket that has sufficient battery life and storage to record for an entire day.
-Closed-source LENA® algorithms analyze the audio recordings to provide counts of useful metrics, such as the number of words spoken by adults in the vicinity of the participant.
+Closed-source LENA® algorithms analyze the audio recordings to provide automatic counts of useful metrics, such as the number of words spoken by adults in the vicinity of the participant.
 Other long-form audio methods rely on custom-built recorders (Wass et al., 2022), apply alternative classification algorithms to LENA® data (Micheletti et al., 2022; Räsänen et al., 2020), or manually transcribe audio recorded by LENA® devices to improve accuracy or identify behaviors beyond the built-in categories (Bergelson, Casillas, et al., 2019; Mendoza \& Fausey, 2021).
 
-Long-form audio recordings have had a transformational impact on language development by allowing researchers to characterize opportunities for learning in daily life.
+Long-form audio recordings have had a transformational impact on language development research by allowing researchers to characterize opportunities for learning in daily life.
 Measuring the amount of speech heard by infants in the home (Weisleder \& Fernald, 2013) or in a daycare setting (Perry et al., 2018) revealed individual differences input that predict later vocabulary.
-Full-day language recording synchronized with other data sources allows researchers to identify how vocal input and production interacts with other processes.
+Full-day language recording synchronized with other data sources allows researchers to identify how auditory input and vocal production interact with other processes.
 Beyond individual differences in aggregated data, long-form recordings can be used to determine the temporal schedule of experiences.
@@
 
-We identified five key features of long-form audio methods to replicate in analagous studies of motor behavior.
+We identified five key features of long-form audio methods that should be replicated in analogous studies of motor behavior.
 First, wearable audio recorders are \emph{mobile}.
 Measurement is not limited to a particular room because the recording device travels with the participant.
-Data are stored on device, so participants do not need to be in range of a receiver.
+Data are recorded to onboard device memory, so participants do not need to be in range of a receiver.
 Second, wearable audio recording is \emph{unobtrusive}.
@@
 The ability to record real-time data is vital for making inferences about processes that happen on the timescale of minutes or even seconds within in an individual as opposed to comparisons of aggregated data between infants.
-Synchronizing real-time data to other data streams helps to reveal sources of variability within an individual.
+Synchronizing real-time data to other data streams helps to reveal sources of variability within an individual (e.g., Wass et al., 2022).
 Full-day recordings are essential for capturing experiences across the heterogeniety of daily routines that moderate behavior (e.g., play, feeding, errands) {[}CITE{]}.
-Burstiness of behavior means that long recordings are needed to capture clusters of events amid long periods in which they may be absent (Barbaro \& Fausey, 2022; Warlaumont et al., 2021).
+``Burstiness'' of behavior means that long recordings are needed to capture clusters of events amid long periods in which they may be absent (Barbaro \& Fausey, 2022; Warlaumont et al., 2021).
 Fifth, \emph{automatic classification} means that the approach can scale to analyze large numbers of participants over long recordings without the bottleneck of manual annotation/transcription.
@@
 Video and survey methods are the current state-of-the-art in assessing motor behavior.
-Although, each method has advantages and disadvantages for characterizing infants' everyday motor experiences that complements the other, neither method on its own can provide comparable data to long form audio recordings as reviewed in the previous section.
+Although, each method has advantages and disadvantages for characterizing infants' everyday motor experiences that complements the other, neither method on its own can provide comparable data to the long form audio recordings reviewed in the previous section.
 
@@
 Even if full day videos were available, the lack of suitable automatic classification tools means that the human cost of annotation makes it difficult for video methods to scale to testing large numbers of participants.
-For annotation of body position takes approximately 2-5 hours to complete for every hour of video (depending on how often infants switch positions), meaning that a full ``waking day'' of approximately 11 hours for a 12-month-old (Galland et al., 2012) could take 22-55 hours of labor to full annotate.
+Our annotation of body position takes approximately 2-5 hours to complete for every hour of video (depending on how often infants switch positions), meaning that a full ``waking day'' of approximately 11 hours for a 12-month-old (Galland et al., 2012) could take 22-55 hours of labor to full annotate.
 
@@
 Ecological momentary assessment uses text-message or app-based notifications to prompt caregivers to make repeated estimates of infants' behavior over the course of a day {[}CITES{]}.
-And although the responses are valuable in aggregate, the lack the real-time temporal resolution to describe moment-to-moment changes in behavior.
-For diaries and interviews, the memory limits mean that caregivers will report what was most frequent, but cannot remember events that happen on the level of seconds and minutes.
-Repeated EMA surveys might prompt caregivers to make hourly observations; increasing the number of surveys per day would be too burdensome for the respondent.
-Thus, despite being a useful tool for estimating broad developmental changes and/or individual differences in infants' motor experiences, survey methods are not suited for capturing within-participant temporal dynamics.
+And although the responses are valuable in aggregate, survey methods lack the real-time temporal resolution to describe moment-to-moment changes in behavior.
+For diaries and interviews, limits on caregivers' memory mean that they will report what was most frequent, but cannot remember events that happen on the scale of seconds and minutes.
+At best, EMA surveys prompt caregivers to make hourly observations; increasing the number of surveys per day would be too burdensome for the respondent.
+Thus, despite being a useful tool for estimating broad developmental changes and individual differences in infants' motor experiences, survey methods are not suited for capturing within-participant temporal dynamics.
 
@@
 Lightweight sensors (XX-XXg) can be embedded in garments to make recordings fully \emph{mobile}, and they are \emph{unobtrusive} because they do not require a researcher's presence nor do they record sensitive audio/video that might influence participants' behavior {[}(Cliff et al., 2009; de Barbaro, 2019; Lobo et al., 2019; Bruijns et al., 2020){]}.
-Many commerically-available and inexpensive IMUs have \textgreater{} 12 hour battery life an onboard storage to record \emph{real-time}, \emph{full-day} data of movements at a sampling rate that is higher than typical video (e.g., 50-100 Hz).
+Many commercially-available and inexpensive IMUs have \textgreater{} 12 hour battery life with onboard storage to record \emph{real-time}, \emph{full-day} motion data at a sampling rate that is higher than typical video (e.g., 50-100 Hz).
 Moreover, past work has successfully recorded the rate of leg kicks {[}(Smith et al., 2017; Jiang et al., 2018){]} and activity intensity {[}Hagenbuchner et al., 2015; Trost et al., 2018{]} in infants and children across an entire day.
@@
 Sitting on a caregivers' lap without the need to maintain balance means that the legs can dangle and the torso can lean in different directions.
-Sitting in a high chair or car seat reduces the magnitude of postural sway within sitting, and creates still more possibilities for how the arms and legs may move relative to the torso.
+Sitting in a high chair or car seat reduces the magnitude of postural sway within sitting, and creates even more possibilities for how the arms and legs may move relative to the torso.
 Indeed, creating an all-encompassing set of rules for how to annotate sitting from video is no trivial task because of the various ways that sitting can occur in daily life.
 Finally, caregivers frequently pick up and transport infants, creating motion signals that need to be differentiated from independent activity {[}Kwon{]}.
+Thus, modern approaches to human activity recognition have used machine learning to classify activity categories based on features derived from IMU data in adults {[}(Preece et al., 2009; Arif and Kattan, 2015){]}, children {[}(Nam and Park, 2013; Zhao et al., 2013; Ren et al., 2016; Stewart et al., 2018){]}, and infants {[}Franchak; Airaksinen; Yao{]}.
+Synchronized video with ground-truth human annotations creates training data for a machine learning algorithm, such as a random forest model, that can be later used to predict categories from IMU data that was not annotated.
+Crucially, this allows automatic classification to scale to full day recording by relying on a relatively smaller set of video annotation.
 GREENSPAN\ldots{}
 
-Thus, modern approaches to human activity recognition have used machine learning to classify activity categories based on features derived from IMU data in adults {[}(Preece et al., 2009; Arif and Kattan, 2015){]}, children {[}(Nam and Park, 2013; Zhao et al., 2013; Ren et al., 2016; Stewart et al., 2018){]}, and infants {[}Franchak; Airaksinen; Yao{]}.
-Synchronized video with ground-truth human annotations creates training data for a machine learning algorithm, such as a random forest models, that can be later used to predict categories from future IMU data that was not annotated.
-Although in the current investigation we focus on classification of body position, this approach can be extended to categorize other aspects of movement, such as locomotion {[}CITE{]} and OTHER EXAMPLES?
-Accompanying our paper, we have created a database {[}CITE{]} containing synchronized video and IMU data, allowing researchers to annotate and train models for behaviors beyond body position.
-
-Three prior investigations have used machine learning to categorize infant body position from IMUs towards the goal of collecting full-day data, and each has yielded encouraging validation results.
+Three prior investigations have used machine learning to categorize infant body position from IMUs towards the goal of collecting full-day data.
 Airaksinen et al.
@@
 Classification accuracy did not vary between lab and home settings, and was generally high (95\%, kappa = .93).
-However, in all three studies accuracy was assessed within a brief (15-60 minute) session supervised by a researcher, leaving the open question of how well body position classification will scale to testing across an entire day of natural home life.
+Although all three studies yielded promising classification accuracy, accuracy was assessed in brief (15-60 minute) sessions supervised by a researcher, leaving the open question of how well body position classification will scale to testing across an entire day of natural home life.
 YAO?
@@
 Accordingly, the overarching goal of the current study is to test the feasibility and validity of long-form body position recording in the home during unsupervised, everyday behavior.
-During a supervised recording, whether in the home or in the lab, the researcher can set up the situation to encourage or restrict certain behaviors and will usually instruct the caregiver to play with the infant.
-Within-day variations in activities (e.g., eating lunch in a high chair) create challenging situations for applying automated classification of body position.
+Supervised recordings from past work {[}CITES{]}, whether in the home or in the lab, let researchers set up the situation to encourage or restrict certain behaviors.
+Usually, caregivers are asked to play with the infant.
+However, across a real day, non-play activities (e.g., eating lunch in a high chair) create challenging situations for applying automated classification of body position.
 Will models trained on video-recorded observations at the beginning of the day generalize to predict behavior at a later time?
@@
 
-In the current study, we report the feasibility and validity of body position classification over the full day in the home based on XX testing sessions from XX infants aged 4-14 months.
+In the current study, we report the feasibility and validity of body position classification over the full day in the home based on 35 testing sessions from XX infants aged 4-14 months.
 Participants received a custom pair of infant leggings embedded with 4 IMUs (one on each ankle and one on each hip) and a video camera to collect ground truth data about infant body position.
 A \emph{proximal comparison} period began when participants received the equipment and completed a guided phone call during which caregivers were asked to elicit different body positions based on prompts from the experimenter.
-Although not directly supervised, this period was most similar to previous recordings because it occurred during a convenient time for the infant and caregiver to play and receive instruction.
-The \textbf{first goal} of the current study was to determine the accuracy of body position classification during the proximal comparison period using our novel, semi-supervised procedure in participants' homes.
+Although not directly supervised, this period was most similar to previous recordings because it occurred during a convenient time for the infant and caregiver to play while they received instructions from the experimenter.
+The \textbf{first goal} of the current study was to determine the accuracy of body position classification during the proximal comparison period using this novel, semi-supervised procedure in participants' homes.
 
-A second, \emph{distal comparison} period followed the proximal comparison period and captured approximately 90 minutes of home behavior without any input from the experimenter.
-Following a long delay from the start of the study when caregivers were guided in placing the equipment, it is possible that accuracy could become worse if caregivers or infants moved the garment or sensors.
-Moreover, increasing variation in everyday activities creates a challenge for classification.
+A second, \emph{distal comparison} period followed the proximal comparison period and captured approximately 90 minutes of home behavior that was completely unsupervised.
+Caregivers and infants could (and did) do whatever they wished.
+Because this recording happened a considerable amount of time after the initial setup and instructions from the experimenter, accuracy could decline if caregivers or infants moved the garment or sensors.
+Moreover, increasing variation in everyday activities during the distal comparison creates a greater challenge for classification, testing whether machine learning models can generalize to novel test cases.
 Thus, the \textbf{second goal} of our study was to assess accuracy during the distal comparison.
 
-After this second video recording period, we asked caregivers to have infants continue to wear IMUs for the rest of the day until they went to bed, creating the first real-time, full-day dataset of infant body position.
+After this second video recording period, we asked caregivers to keep infants wearing IMUs for the rest of the day until their regular bedtime, creating the \textbf{first real-time, full-day dataset of infant body position}.
 Interpreting such data required caregivers to log when infants napped, when they removed the sensor garment for diaper changes or other reasons, and when infants went to bed at the end of the day.
@@
 Finally, the novel full-day dataset affords us a unique chance to ask whether the estimated time infants spent in different body positions align with past results using video and survey methods.
-Thus, the \textbf{fourth goal} of the current study was to determine whether our full-day body position estimates yielded expected age-related differences in everyday body position.
-Based on past results, with age infants should spend increasingly more time sitting and upright but less time supine.
+Thus, the \textbf{fourth goal} of the current study was to determine whether full-day body position measurements conformed to expectations about age differences in body position.
+Based on past results {[}CITE{]}, infants should spend increasingly more time sitting and upright but less time supine over the age range tested (4 to 14 months).
 

### 2023-03-02 group_split_comparison/group_split.R
updated data

@@ -11,7 +11,8 @@ library(rstatix)
 
 #LOAD DATA
 load( here("group_split_comparison","compiled_data_lite.RData"))
-ntree = 150
+ntree = 550
+mtry = 44
 not_all_na <- function(x) any(!is.na(x))
 
 training <- slide_filt %>% group_by(id, code) %>% slice_head(prop = .75) %>% ungroup 
@@ -27,7 +28,7 @@ group_model <- function(temp_id, training, ntree = 50, bad_ids = NULL) {
     mutate(code = fct_drop(code)) %>% 
     select(-id)
   group <- randomForest(code ~ ., data = temp_group_training, 
-                        localImp = TRUE, proximity = FALSE, ntree = ntree)
+                        localImp = TRUE, proximity = FALSE, ntree = ntree, mtry = mtry)
 }
 
 split_model <- function(temp_id, training, ntree = 50) {
@@ -37,7 +38,7 @@ split_model <- function(temp_id, training, ntree = 50) {
     mutate(code = fct_drop(code)) %>% 
     select(-id)
   split <- randomForest(code ~ ., data = temp_split_training, 
-                        localImp = TRUE, proximity = FALSE, ntree = ntree)
+                        localImp = TRUE, proximity = FALSE, ntree = ntree, mtry = mtry)
 }
 
 ids <- unique(training$id)

### 2023-03-02 manuscript.Rmd
adding refs

@@
 As infants grow older and acquire new abilities, such as independent sitting and walking, the spend more time sitting and upright and less time held, supine, and prone [@Survey; @Thurman2017; @Freeplay; @AdolphCTL2014].
-Thus, characterizing individual differences in the day-to-day accumulation of body position experiences informs developmental theory by revealing differential opportunities for learning.
-[CLINICAL POSSIBILITIES?]
-<!-- Other west cites; cite new EMA -->
+Thus, characterizing individual differences in the day-to-day accumulation of body position experiences informs developmental theory by revealing differential opportunities for learning [@CurrentOpinion].
+<!-- Other west cites; cite new EMA; KretchKoziol2022 -->
 
@@
 Our method takes inspiration from a more mature technology: Long-form audio recordings of infants' language experiences using wearable audio recorders.
-We begin by describing the impact of long-form audio recordings on research in developmental psychology [CITES], and identify the key features that should be replicated in long-form recordings of motor behavior.
+We begin by describing the impact of long-form audio recordings on research in developmental psychology, and identify the key features that should be replicated in long-form recordings of motor behavior.
 Next, we review the current state-of-the-art in measuring infant motor behavior---video and survey data---and their limitations in capturing real-time, full-day behavior.
 Finally, we discuss the advantages of using inertial sensing to classify motor behavior.
-Despite promising past results in brief, supervised sessions [CITES], the current investigation takes a needed step forward by testing accuracy over long, unsupervised recordings.
-Although in the current investigation we focus on classification of body position, this approach can be extended to categorize other aspects of movement, such as locomotion [CITE] and OTHER EXAMPLES?
+Despite promising past results in brief, supervised sessions [@FranchakScott2021; @AiraksinenRasanen2020; @AiraksinenGallen2022], the current investigation takes a needed step forward by testing accuracy over long, unsupervised recordings.
+Although in the current investigation we focus on classification of body position, this approach can be extended to categorize other aspects of movement, such as locomotion [@AiraksinenRasanen2020] and infant-caregiver contact [@YaoPlotz2019].
 Accompanying our paper, we have created a database [CITE] containing synchronized video and IMU data, allowing researchers to annotate and train models for behaviors beyond body position.
@@
 Synchronizing real-time data to other data streams helps to reveal sources of variability within an individual [e.g., @WassPhillips2022].
-Full-day recordings are essential for capturing experiences across the heterogeniety of daily routines that moderate behavior (e.g., play, feeding, errands) [CITE].
+Full-day recordings are essential for capturing experiences across the heterogeniety of daily routines that moderate behavior (e.g., play, feeding, errands) [@KadookaCaufield2021].
 "Burstiness" of behavior means that long recordings are needed to capture clusters of events amid long periods in which they may be absent [@BarbaroFausey2022; @WarlaumontSobowale2021].
@@
 However, poor agreement was found for other metrics, such as the number of "conversational turns" between the child and communicative partners and the identification of male speakers.
+<!-- EVENTUALLY CITE EMA HERE -->
 
@@
 Video observation is the most common way of measuring infant motor behavior in home recordings.
-Most often, an experimenter with a handheld camera follows infants from room to room to ensure that their movements are visible throughout the recording session [CITE ADOLPH, ANYONE ELSE?], but sometimes a fixed camera may be used [IVERSON?].
+Most often, an experimenter with a handheld camera follows infants from room to room to ensure that their movements are visible throughout the recording session [@Karasik2011; @ChenSchneider2022; @HerzbergFletcher2021].
 The primary advantage of video recording is that it captures real-time behavior.
@@
 Logistically, an experimenter cannot follow behind infants to record their behavior from morning to night (and were they to do so, they would likely alter infants' and caregivers' behavior).
-Typical video recording sessions last XX-XX, short of capturing the variety of activities across the full daily routine.
+Typical video recording sessions last 45-120 minutes [@Karasik2011; @ChenSchneider2022; @HerzbergFletcher2021], short of capturing the variety of activities across the full daily routine.
 Even if full day videos were available, the lack of suitable automatic classification tools means that the human cost of annotation makes it difficult for video methods to scale to testing large numbers of participants.
@@
 In contrast, survey methods such as daily diaries/inventories or ecological momentary assessment (EMA) are mobile, unobtrusive, can be applied across an entire day, and do not need laborious annotation.
-Diary studies provide caregivers with logs or structured interviews to estimate from memory how much time infants spend in particular activities [CITES].
-Ecological momentary assessment uses text-message or app-based notifications to prompt caregivers to make repeated estimates of infants' behavior over the course of a day [CITES].
+Diary studies provide caregivers with logs or structured interviews to estimate from memory how much time infants spend in particular activities [@MajnemerBarr2005; @KarasikKuchirko2022].
+Ecological momentary assessment uses text-message or app-based notifications to prompt caregivers to make repeated estimates of infants' behavior over the course of a day [@Survey; @KadookaCaufield2021].
 And although the responses are valuable in aggregate, survey methods lack the real-time temporal resolution to describe moment-to-moment changes in behavior.
@@
 Measuring infant movement with inertial movement units (IMUs) is a promising avenue for long form recordings of motor behavior in the home.
-Lightweight sensors (XX-XXg) can be embedded in garments to make recordings fully *mobile*, and they are *unobtrusive* because they do not require a researcher's presence nor do they record sensitive audio/video that might influence participants' behavior [(Cliff et al., 2009; de Barbaro, 2019; Lobo et al., 2019; Bruijns et al., 2020)].
+Lightweight sensors (10-30 g) can be embedded in garments to make recordings fully *mobile*, and they are *unobtrusive* because they do not require a researcher's presence nor do they record sensitive audio/video that might influence participants' behavior [(Cliff et al., 2009; de Barbaro, 2019; Lobo et al., 2019; Bruijns et al., 2020)].
 Many commercially-available and inexpensive IMUs have > 12 hour battery life with onboard storage to record *real-time*, *full-day* motion data at a sampling rate that is higher than typical video (e.g., 50-100 Hz).
-Moreover, past work has successfully recorded the rate of leg kicks [(Smith et al., 2017; Jiang et al., 2018)] and activity intensity [Hagenbuchner et al., 2015; Trost et al., 2018] in infants and children across an entire day.
+Moreover, past work has successfully recorded the rate of leg kicks [e.g., @DengTrujillo-Priego2019] and activity intensity [@SchnellerBentsen2017] in infants and children across multiple days.
 <!-- CHECK THAT CITATIONS ARE FULL DAY -->
@@
 Indeed, creating an all-encompassing set of rules for how to annotate sitting from video is no trivial task because of the various ways that sitting can occur in daily life.
-Finally, caregivers frequently pick up and transport infants, creating motion signals that need to be differentiated from independent activity [Kwon].
+Finally, caregivers frequently pick up and transport infants, creating motion signals that need to be differentiated from independent activity [@KwonZavos2019; @PatelShi2019].
 Thus, modern approaches to human activity recognition have used machine learning to classify activity categories based on features derived from IMU data in adults [(Preece et al., 2009; Arif and Kattan, 2015)], children [(Nam and Park, 2013; Zhao et al., 2013; Ren et al., 2016; Stewart et al., 2018)], and infants [Franchak; Airaksinen; Yao].

### 2023-03-02 manuscript.tex
adding refs

@@
 As infants grow older and acquire new abilities, such as independent sitting and walking, the spend more time sitting and upright and less time held, supine, and prone (Adolph \& Tamis-LeMonda, 2014; Franchak et al., 2018; Franchak, 2019; Thurman \& Corbetta, 2017).
-Thus, characterizing individual differences in the day-to-day accumulation of body position experiences informs developmental theory by revealing differential opportunities for learning.
-{[}CLINICAL POSSIBILITIES?{]}
+Thus, characterizing individual differences in the day-to-day accumulation of body position experiences informs developmental theory by revealing differential opportunities for learning (Franchak, 2020).
 
@@
 Our method takes inspiration from a more mature technology: Long-form audio recordings of infants' language experiences using wearable audio recorders.
-We begin by describing the impact of long-form audio recordings on research in developmental psychology {[}CITES{]}, and identify the key features that should be replicated in long-form recordings of motor behavior.
+We begin by describing the impact of long-form audio recordings on research in developmental psychology, and identify the key features that should be replicated in long-form recordings of motor behavior.
 Next, we review the current state-of-the-art in measuring infant motor behavior---video and survey data---and their limitations in capturing real-time, full-day behavior.
 Finally, we discuss the advantages of using inertial sensing to classify motor behavior.
-Despite promising past results in brief, supervised sessions {[}CITES{]}, the current investigation takes a needed step forward by testing accuracy over long, unsupervised recordings.
-Although in the current investigation we focus on classification of body position, this approach can be extended to categorize other aspects of movement, such as locomotion {[}CITE{]} and OTHER EXAMPLES?
+Despite promising past results in brief, supervised sessions (Airaksinen et al., 2022, 2020; Franchak et al., 2021), the current investigation takes a needed step forward by testing accuracy over long, unsupervised recordings.
+Although in the current investigation we focus on classification of body position, this approach can be extended to categorize other aspects of movement, such as locomotion (Airaksinen et al., 2020) and infant-caregiver contact (Yao et al., 2019).
 Accompanying our paper, we have created a database {[}CITE{]} containing synchronized video and IMU data, allowing researchers to annotate and train models for behaviors beyond body position.
@@
 Synchronizing real-time data to other data streams helps to reveal sources of variability within an individual (e.g., Wass et al., 2022).
-Full-day recordings are essential for capturing experiences across the heterogeniety of daily routines that moderate behavior (e.g., play, feeding, errands) {[}CITE{]}.
+Full-day recordings are essential for capturing experiences across the heterogeniety of daily routines that moderate behavior (e.g., play, feeding, errands) (Kadooka et al., 2021, April).
 ``Burstiness'' of behavior means that long recordings are needed to capture clusters of events amid long periods in which they may be absent (Barbaro \& Fausey, 2022; Warlaumont et al., 2021).
@@
 Video observation is the most common way of measuring infant motor behavior in home recordings.
-Most often, an experimenter with a handheld camera follows infants from room to room to ensure that their movements are visible throughout the recording session {[}CITE ADOLPH, ANYONE ELSE?{]}, but sometimes a fixed camera may be used {[}IVERSON?{]}.
+Most often, an experimenter with a handheld camera follows infants from room to room to ensure that their movements are visible throughout the recording session (Chen et al., 2022; Herzberg et al., 2021; Karasik et al., 2011).
 The primary advantage of video recording is that it captures real-time behavior.
@@
 Logistically, an experimenter cannot follow behind infants to record their behavior from morning to night (and were they to do so, they would likely alter infants' and caregivers' behavior).
-Typical video recording sessions last XX-XX, short of capturing the variety of activities across the full daily routine.
+Typical video recording sessions last 45-120 minutes (Chen et al., 2022; Herzberg et al., 2021; Karasik et al., 2011), short of capturing the variety of activities across the full daily routine.
 Even if full day videos were available, the lack of suitable automatic classification tools means that the human cost of annotation makes it difficult for video methods to scale to testing large numbers of participants.
@@
 In contrast, survey methods such as daily diaries/inventories or ecological momentary assessment (EMA) are mobile, unobtrusive, can be applied across an entire day, and do not need laborious annotation.
-Diary studies provide caregivers with logs or structured interviews to estimate from memory how much time infants spend in particular activities {[}CITES{]}.
-Ecological momentary assessment uses text-message or app-based notifications to prompt caregivers to make repeated estimates of infants' behavior over the course of a day {[}CITES{]}.
+Diary studies provide caregivers with logs or structured interviews to estimate from memory how much time infants spend in particular activities (Karasik et al., 2022; Majnemer \& Barr, 2005).
+Ecological momentary assessment uses text-message or app-based notifications to prompt caregivers to make repeated estimates of infants' behavior over the course of a day (Franchak, 2019; Kadooka et al., 2021, April).
 And although the responses are valuable in aggregate, survey methods lack the real-time temporal resolution to describe moment-to-moment changes in behavior.
@@
 Measuring infant movement with inertial movement units (IMUs) is a promising avenue for long form recordings of motor behavior in the home.
-Lightweight sensors (XX-XXg) can be embedded in garments to make recordings fully \emph{mobile}, and they are \emph{unobtrusive} because they do not require a researcher's presence nor do they record sensitive audio/video that might influence participants' behavior {[}(Cliff et al., 2009; de Barbaro, 2019; Lobo et al., 2019; Bruijns et al., 2020){]}.
+Lightweight sensors (10-30 g) can be embedded in garments to make recordings fully \emph{mobile}, and they are \emph{unobtrusive} because they do not require a researcher's presence nor do they record sensitive audio/video that might influence participants' behavior {[}(Cliff et al., 2009; de Barbaro, 2019; Lobo et al., 2019; Bruijns et al., 2020){]}.
 Many commercially-available and inexpensive IMUs have \textgreater{} 12 hour battery life with onboard storage to record \emph{real-time}, \emph{full-day} motion data at a sampling rate that is higher than typical video (e.g., 50-100 Hz).
-Moreover, past work has successfully recorded the rate of leg kicks {[}(Smith et al., 2017; Jiang et al., 2018){]} and activity intensity {[}Hagenbuchner et al., 2015; Trost et al., 2018{]} in infants and children across an entire day.
+Moreover, past work has successfully recorded the rate of leg kicks (e.g., Deng et al., 2019) and activity intensity (SCHNELLER et al., 2017) in infants and children across multiple days.
 
@@
 Indeed, creating an all-encompassing set of rules for how to annotate sitting from video is no trivial task because of the various ways that sitting can occur in daily life.
-Finally, caregivers frequently pick up and transport infants, creating motion signals that need to be differentiated from independent activity {[}Kwon{]}.
+Finally, caregivers frequently pick up and transport infants, creating motion signals that need to be differentiated from independent activity (Kwon et al., 2019; Patel et al., 2019).
 Thus, modern approaches to human activity recognition have used machine learning to classify activity categories based on features derived from IMU data in adults {[}(Preece et al., 2009; Arif and Kattan, 2015){]}, children {[}(Nam and Park, 2013; Zhao et al., 2013; Ren et al., 2016; Stewart et al., 2018){]}, and infants {[}Franchak; Airaksinen; Yao{]}.
@@
 \midrule
-Held & -0.03 & 0.12 & 0.55 & 0.58\\
-Prone & 0.97 & 0.85 & 0.97 & 0.84\\
-Sitting & 0.73 & 0.93 & 0.91 & 0.97\\
-Supine & 0.84 & 0.94 & 0.94 & 0.97\\
-Upright & 0.84 & 0.93 & 0.99 & 0.94\\ \midrule
-Overall & 0.80 & 0.91 & 0.95 & 0.96\\
+Held & 0.02 & 0.05 & 0.75 & 0.60\\
+Prone & 0.97 & 0.86 & 0.97 & 0.84\\
+Sitting & 0.70 & 0.88 & 0.91 & 0.95\\
+Supine & 0.84 & 0.92 & 0.94 & 0.97\\
+Upright & 0.62 & 0.84 & 0.99 & 0.94\\ \midrule
+Overall & 0.76 & 0.87 & 0.95 & 0.96\\
 \bottomrule
@@
 \midrule
-Held & 0.45 & 0.43 & 0.57 & 0.55\\
-Prone & 0.96 & 0.89 & 0.96 & 0.88\\
-Sitting & 0.72 & 0.92 & 0.91 & 0.93\\
-Supine & 0.75 & 0.95 & 0.90 & 0.94\\
-Upright & 0.93 & 0.95 & 0.97 & 0.95\\ \midrule
-Overall & 0.80 & 0.94 & 0.93 & 0.94\\
+Held & 0.51 & 0.45 & 0.67 & 0.64\\
+Prone & 0.96 & 0.90 & 0.96 & 0.89\\
+Sitting & 0.71 & 0.91 & 0.89 & 0.92\\
+Supine & 0.75 & 0.94 & 0.88 & 0.94\\
+Upright & 0.90 & 0.93 & 0.98 & 0.96\\ \midrule
+Overall & 0.79 & 0.93 & 0.92 & 0.94\\
 \bottomrule
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-AiraksinenGallen2022}{}}%
+Airaksinen, M., Gallen, A., Kivi, A., Vijayakrishnan, P., Häyrinen, T., Ilén, E., Räsänen, O., Haataja, L.
+M., \& Vanhatalo, S.
+(2022).
+Intelligent wearable allows out-of-the-lab tracking of developing motor abilities in infants.
+\emph{Communications Medicine}, \emph{2}(1).
+\url{https://doi.org/10.1038/s43856-022-00131-6}
+
 \leavevmode\vadjust pre{\hypertarget{ref-AiraksinenRasanen2020}{}}%
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-DengTrujillo-Priego2019}{}}%
+Deng, W., Trujillo-Priego, I.
+A., \& Smith, B.
+A.
+(2019).
+How many days are necessary to represent an infant{\textquotesingle}s typical daily leg movement behavior using wearable sensors?
+\emph{Physical Therapy}, \emph{99}(6), 730--738.
+\url{https://doi.org/10.1093/ptj/pzz036}
+
 \leavevmode\vadjust pre{\hypertarget{ref-Survey}{}}%
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-CurrentOpinion}{}}%
+Franchak, J. M. (2020). The ecology of infants' perceptual-motor exploration. \emph{Current Opinion in Psychology}, \emph{32}, 110--114.
+
 \leavevmode\vadjust pre{\hypertarget{ref-Freeplay}{}}%
 Franchak, J. M., Kretch, K. S., \& Adolph, K. E. (2018). See and be seen: Infant-caregiver social looking during locomotor free play. \emph{Developmental Science}, \emph{21}, e12626.
+
+\leavevmode\vadjust pre{\hypertarget{ref-FranchakScott2021}{}}%
+Franchak, J.
+M., Scott, V., \& Luo, C.
+(2021).
+A contactless method for measuring full-day, naturalistic motor behavior using wearable inertial sensors.
+\emph{Frontiers in Psychology}, \emph{12}.
+\url{https://doi.org/10.3389/fpsyg.2021.701343}
 
@@
 \url{https://doi.org/10.1016/j.smrv.2011.06.001}
+
+\leavevmode\vadjust pre{\hypertarget{ref-HerzbergFletcher2021}{}}%
+Herzberg, O., Fletcher, K.
+K., Schatz, J.
+L., Adolph, K.
+E., \& Tamis-LeMonda, C.
+S.
+(2021).
+Infant exuberant object play at home: Immense amounts of time-distributed, variable practice.
+\emph{Child Development}, \emph{93}(1), 150--164.
+
+\leavevmode\vadjust pre{\hypertarget{ref-KadookaCaufield2021}{}}%
+Kadooka, K., Caufield, M., Fausey, C.
+M., \& Franchak, J.
+M.
+(2021, April).
+Visuomotor learning opportunities are nested within everyday activities.
+\emph{Paper Presented at the Biennial Meeting of the Society for Research in Child Development}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-KarasikKuchirko2022}{}}%
+Karasik, L.
+B., Kuchirko, Y., Dodojonova, R.
+M., \& Elison, J.
+T.
+(2022).
+Comparison of {U.S.} And {Tajik} infants' time in containment devices.
+\emph{Infant and Child Development}, \emph{31}(4).
+\url{https://doi.org/10.1002/icd.2340}
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-KwonZavos2019}{}}%
+Kwon, S., Zavos, P., Nickele, K., Sugianto, A., \& Albert, M.
+V.
+(2019).
+Hip and wrist-worn accelerometer data analysis for toddler activities.
+\emph{International Journal of Environmental Research and Public Health}, \emph{16}(14), 2598.
+
 \leavevmode\vadjust pre{\hypertarget{ref-LuoFranchak2020}{}}%
 Luo, C., \& Franchak, J. M. (2020). Head and body structure infants' visual experiences during mobile, naturalistic play. \emph{{PLoS ONE}}, \emph{15}, e0242009.
+
+\leavevmode\vadjust pre{\hypertarget{ref-MajnemerBarr2005}{}}%
+Majnemer, A., \& Barr, R. G. (2005). Influence of supine sleep positioning on early motor milestone acquisition. \emph{Developmental Medicine and Child Neurology}, \emph{47}, 370--376.
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-PatelShi2019}{}}%
+Patel, P., Shi, Y., Hajiaghajani, F., Biswas, S., \& Lee, M.-H.
+(2019).
+A novel two-body sensor system to study spontaneous movements in infants during caregiver physical contact.
+\emph{Infant Behavior and Development}, \emph{57}, 101383.
+\url{https://doi.org/10.1016/j.infbeh.2019.101383}
+
 \leavevmode\vadjust pre{\hypertarget{ref-PerryPrince2018}{}}%
@@
 \url{https://doi.org/10.3758/s13428-020-01460-x}
+
+\leavevmode\vadjust pre{\hypertarget{ref-SchnellerBentsen2017}{}}%
+SCHNELLER, M.
+B., BENTSEN, P., NIELSEN, G., BRØND, J.
+C., RIED-LARSEN, M., MYGIND, E., \& SCHIPPERIJN, J.
+(2017).
+Measuring children's physical activity.
+\emph{Medicine \& Science in Sports \& Exercise}, \emph{49}(6), 1261--1269.
+\url{https://doi.org/10.1249/mss.0000000000001222}
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-YaoPlotz2019}{}}%
+Yao, X., Plötz, T., Johnson, M., \& Barbaro, K.
+de.
+(2019).
+Automated detection of infant holding using wearable sensing: Implications for developmental science and intervention.
+\emph{Proceedings of the ACM on Interactive, Mobile, Wearable and Ubiquitous Technologies}, \emph{3}(2), 1--17.
+
 \end{CSLReferences}

### 2023-03-03 manuscript.Rmd
more intro citations

@@
 
-Measuring infant movement with inertial movement units (IMUs) is a promising avenue for long form recordings of motor behavior in the home.
-Lightweight sensors (10-30 g) can be embedded in garments to make recordings fully *mobile*, and they are *unobtrusive* because they do not require a researcher's presence nor do they record sensitive audio/video that might influence participants' behavior [(Cliff et al., 2009; de Barbaro, 2019; Lobo et al., 2019; Bruijns et al., 2020)].
+Measuring infant movement with inertial movement units (IMUs) is a promising avenue for long form recordings of motor behavior in the home [@CliffReilly2009; @Barbaro2019; @BruijnsTruelove2020; @LoboHall2019].
+Lightweight sensors (10-30 g) can be embedded in garments to make recordings fully *mobile*, and they are *unobtrusive* because they do not require a researcher's presence nor do they record sensitive audio/video that might influence participants' behavior.
 Many commercially-available and inexpensive IMUs have > 12 hour battery life with onboard storage to record *real-time*, *full-day* motion data at a sampling rate that is higher than typical video (e.g., 50-100 Hz).
 Moreover, past work has successfully recorded the rate of leg kicks [e.g., @DengTrujillo-Priego2019] and activity intensity [@SchnellerBentsen2017] in infants and children across multiple days.
-<!-- CHECK THAT CITATIONS ARE FULL DAY -->
 
@@
 The difficulty of the classification task depends on the categories of interest.
-More basic aspects of movement, such as overall activity intensity, can be identified by taking the magnitude of acceleration (irrespective of direction) and applying thresholds or cut-points to define when a movement has occurred at a particular intensity [(Trost et al., 2012; Kuzik et al., 2015; Hager et al., 2017; Armstrong et al., 2019)].
+More basic aspects of movement, such as overall activity intensity, can be identified by taking the magnitude of acceleration (irrespective of direction) and applying thresholds or cut-points to define when a movement has occurred at a particular intensity [e.g., @HagerTilton2017; @ArmstrongCovington2019].
 
@@
 An upright infant can be standing still or can be walking briskly across the room.
-A prone infant and be stationary in "tummy time" or crawling in myriad ways [ADOLPH].
+A prone infant and be stationary in "tummy time", or they can crawl in myriad ways [@AdolphVereijken1998].
 Moreover, the configuration of the arms, legs, and torso within a body position can vary greatly in everyday contexts.
@@
 Finally, caregivers frequently pick up and transport infants, creating motion signals that need to be differentiated from independent activity [@KwonZavos2019; @PatelShi2019].
-Thus, modern approaches to human activity recognition have used machine learning to classify activity categories based on features derived from IMU data in adults [(Preece et al., 2009; Arif and Kattan, 2015)], children [(Nam and Park, 2013; Zhao et al., 2013; Ren et al., 2016; Stewart et al., 2018)], and infants [Franchak; Airaksinen; Yao].
+Thus, modern approaches to human activity recognition have used machine learning to classify activity categories based on features derived from IMU data in adults [@Preece2009; @Arif2015], children [@NamPark2013; @RenDing2016; @StewartNarayanan2018], and infants [@YaoPlotz2019; @FranchakScott2021; @AiraksinenRasanen2020].
 Synchronized video with ground-truth human annotations creates training data for a machine learning algorithm, such as a random forest model, that can be later used to predict categories from IMU data that was not annotated.
 Crucially, this allows automatic classification to scale to full day recording by relying on a relatively smaller set of video annotation.
-GREENSPAN...
+
+<!-- NEED TO FIT IN GREENSPAN posture from cut points, Hewitt Tummy time, Yao holding-->
 
 Three prior investigations have used machine learning to categorize infant body position from IMUs towards the goal of collecting full-day data.
-@AiraksinenRasanen2020 tested 4- to 8-month-olds in a laboratory visit, and found XX% accuracy (kappa = XX%) in distinguishing between body position categories that crawling infants could perform on the floor (excluding times that infants were held by caregivers).
-Using a wider age range of XX-XX, [FRANCHAK] found XX% accuracy (kappa = XX%) in a laboratory validation study in categorizing body position that included infants who could both crawl and walk and also categorized caregiver holding.
-Most recently, [Airaksinen NEW] conducted a validation study of body position classification in either a home or clinic testing in 4- to 19-month-olds, refining their previous method to detect moments that infants were carried by caregivers.
-Classification accuracy did not vary between lab and home settings, and was generally high (95%, kappa = .93).
+@AiraksinenRasanen2020 tested 4- to 8-month-olds in a laboratory visit, and found 95% accuracy in distinguishing between body position categories that crawling infants could perform on the floor (excluding times that infants were held by caregivers).
+Using a wider age range of 6-18 months, @FranchakScott2021 found 98% accuracy (*kappa* = 95%) in a laboratory validation study in categorizing body position that included infants who could both crawl and walk and also included a category for caregiver holding.
+Most recently, @AiraksinenGallen2022 conducted a validation study of body position classification in either a home or clinic testing in 4- to 19-month-olds, refining their previous method to detect moments that infants were carried by caregivers.
+Classification accuracy did not vary between lab and home settings, and was generally high (95%, *kappa* = .93).
 Although all three studies yielded promising classification accuracy, accuracy was assessed in brief (15-60 minute) sessions supervised by a researcher, leaving the open question of how well body position classification will scale to testing across an entire day of natural home life.
-YAO?
+
+<!-- Could have a bit more detail in this last paragraph -->
 
@@
 Accordingly, the overarching goal of the current study is to test the feasibility and validity of long-form body position recording in the home during unsupervised, everyday behavior.
-Supervised recordings from past work [CITES], whether in the home or in the lab, let researchers set up the situation to encourage or restrict certain behaviors.
+Supervised recordings from past work [@FranchakScott2021; @AiraksinenRasanen2020; @AiraksinenGallen2022], whether in the home or in the lab, let researchers set up the situation to encourage or restrict certain behaviors.
 Usually, caregivers are asked to play with the infant.
@@
 
-In the current study, we report the feasibility and validity of body position classification over the full day in the home based on 35 testing sessions from XX infants aged 4-14 months.
+In the current study, we report the feasibility and validity of body position classification over the full day in the home based on 35 testing sessions from 23 infants aged 4-14 months.
 Participants received a custom pair of infant leggings embedded with 4 IMUs (one on each ankle and one on each hip) and a video camera to collect ground truth data about infant body position.
@@
 Thus, the **fourth goal** of the current study was to determine whether full-day body position measurements conformed to expectations about age differences in body position.
-Based on past results [CITE], infants should spend increasingly more time sitting and upright but less time supine over the age range tested (4 to 14 months).
+Based on past results [@Survey], infants should spend increasingly more time sitting and upright but less time supine over the age range tested (4 to 14 months).
 

### 2023-03-03 manuscript.tex
more intro citations

@@
 
-Measuring infant movement with inertial movement units (IMUs) is a promising avenue for long form recordings of motor behavior in the home.
-Lightweight sensors (10-30 g) can be embedded in garments to make recordings fully \emph{mobile}, and they are \emph{unobtrusive} because they do not require a researcher's presence nor do they record sensitive audio/video that might influence participants' behavior {[}(Cliff et al., 2009; de Barbaro, 2019; Lobo et al., 2019; Bruijns et al., 2020){]}.
+Measuring infant movement with inertial movement units (IMUs) is a promising avenue for long form recordings of motor behavior in the home (Barbaro, 2019; Bruijns et al., 2020; Cliff et al., 2009; Lobo et al., 2019).
+Lightweight sensors (10-30 g) can be embedded in garments to make recordings fully \emph{mobile}, and they are \emph{unobtrusive} because they do not require a researcher's presence nor do they record sensitive audio/video that might influence participants' behavior.
 Many commercially-available and inexpensive IMUs have \textgreater{} 12 hour battery life with onboard storage to record \emph{real-time}, \emph{full-day} motion data at a sampling rate that is higher than typical video (e.g., 50-100 Hz).
-Moreover, past work has successfully recorded the rate of leg kicks (e.g., Deng et al., 2019) and activity intensity (SCHNELLER et al., 2017) in infants and children across multiple days.
+Moreover, past work has successfully recorded the rate of leg kicks (e.g., Deng et al., 2019) and activity intensity (Schneller et al., 2017) in infants and children across multiple days.
 
@@
 The difficulty of the classification task depends on the categories of interest.
-More basic aspects of movement, such as overall activity intensity, can be identified by taking the magnitude of acceleration (irrespective of direction) and applying thresholds or cut-points to define when a movement has occurred at a particular intensity {[}(Trost et al., 2012; Kuzik et al., 2015; Hager et al., 2017; Armstrong et al., 2019){]}.
+More basic aspects of movement, such as overall activity intensity, can be identified by taking the magnitude of acceleration (irrespective of direction) and applying thresholds or cut-points to define when a movement has occurred at a particular intensity (Armstrong et al., 2019; e.g., Hager et al., 2017).
 
@@
 An upright infant can be standing still or can be walking briskly across the room.
-A prone infant and be stationary in ``tummy time'' or crawling in myriad ways {[}ADOLPH{]}.
+A prone infant and be stationary in ``tummy time'', or they can crawl in myriad ways (Adolph et al., 1998).
 Moreover, the configuration of the arms, legs, and torso within a body position can vary greatly in everyday contexts.
@@
 Finally, caregivers frequently pick up and transport infants, creating motion signals that need to be differentiated from independent activity (Kwon et al., 2019; Patel et al., 2019).
-Thus, modern approaches to human activity recognition have used machine learning to classify activity categories based on features derived from IMU data in adults {[}(Preece et al., 2009; Arif and Kattan, 2015){]}, children {[}(Nam and Park, 2013; Zhao et al., 2013; Ren et al., 2016; Stewart et al., 2018){]}, and infants {[}Franchak; Airaksinen; Yao{]}.
+Thus, modern approaches to human activity recognition have used machine learning to classify activity categories based on features derived from IMU data in adults (Arif \& Kattan, 2015; Preece et al., 2009), children (Nam \& Park, 2013; Ren et al., 2016; Stewart et al., 2018), and infants (Airaksinen et al., 2020; Franchak et al., 2021; Yao et al., 2019).
 Synchronized video with ground-truth human annotations creates training data for a machine learning algorithm, such as a random forest model, that can be later used to predict categories from IMU data that was not annotated.
 Crucially, this allows automatic classification to scale to full day recording by relying on a relatively smaller set of video annotation.
-GREENSPAN\ldots{}
 
@@
 Airaksinen et al.
-(2020) tested 4- to 8-month-olds in a laboratory visit, and found XX\% accuracy (kappa = XX\%) in distinguishing between body position categories that crawling infants could perform on the floor (excluding times that infants were held by caregivers).
-Using a wider age range of XX-XX, {[}FRANCHAK{]} found XX\% accuracy (kappa = XX\%) in a laboratory validation study in categorizing body position that included infants who could both crawl and walk and also categorized caregiver holding.
-Most recently, {[}Airaksinen NEW{]} conducted a validation study of body position classification in either a home or clinic testing in 4- to 19-month-olds, refining their previous method to detect moments that infants were carried by caregivers.
-Classification accuracy did not vary between lab and home settings, and was generally high (95\%, kappa = .93).
+(2020) tested 4- to 8-month-olds in a laboratory visit, and found 95\% accuracy in distinguishing between body position categories that crawling infants could perform on the floor (excluding times that infants were held by caregivers).
+Using a wider age range of 6-18 months, Franchak et al.
+(2021) found 98\% accuracy (\emph{kappa} = 95\%) in a laboratory validation study in categorizing body position that included infants who could both crawl and walk and also included a category for caregiver holding.
+Most recently, Airaksinen et al.
+(2022) conducted a validation study of body position classification in either a home or clinic testing in 4- to 19-month-olds, refining their previous method to detect moments that infants were carried by caregivers.
+Classification accuracy did not vary between lab and home settings, and was generally high (95\%, \emph{kappa} = .93).
 Although all three studies yielded promising classification accuracy, accuracy was assessed in brief (15-60 minute) sessions supervised by a researcher, leaving the open question of how well body position classification will scale to testing across an entire day of natural home life.
-YAO?
 
@@
 Accordingly, the overarching goal of the current study is to test the feasibility and validity of long-form body position recording in the home during unsupervised, everyday behavior.
-Supervised recordings from past work {[}CITES{]}, whether in the home or in the lab, let researchers set up the situation to encourage or restrict certain behaviors.
+Supervised recordings from past work (Airaksinen et al., 2022, 2020; Franchak et al., 2021), whether in the home or in the lab, let researchers set up the situation to encourage or restrict certain behaviors.
 Usually, caregivers are asked to play with the infant.
@@
 Thus, the \textbf{fourth goal} of the current study was to determine whether full-day body position measurements conformed to expectations about age differences in body position.
-Based on past results {[}CITE{]}, infants should spend increasingly more time sitting and upright but less time supine over the age range tested (4 to 14 months).
+Based on past results (Franchak, 2019), infants should spend increasingly more time sitting and upright but less time supine over the age range tested (4 to 14 months).
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-AdolphVereijken1998}{}}%
+Adolph, K. E., Vereijken, B., \& Denny, M. A. (1998). Learning to crawl. \emph{Child Development}, \emph{69}, 1299--1312. \url{https://doi.org/10.1111/j.1467-8624.1998.tb06213.x}
+
 \leavevmode\vadjust pre{\hypertarget{ref-AiraksinenGallen2022}{}}%
@@
 \emph{Scientific Reports}, \emph{10}(1), 1--13.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Arif2015}{}}%
+Arif, M., \& Kattan, A. (2015). Physical activities monitoring using wearable acceleration sensors attached to the body. \emph{{PLoS ONE}}, \emph{10}, e0130851.
+
+\leavevmode\vadjust pre{\hypertarget{ref-ArmstrongCovington2019}{}}%
+Armstrong, B., Covington, L.
+B., Hager, E.
+R., \& Black, M.
+M.
+(2019).
+Objective sleep and physical activity using 24-hour ankle-worn accelerometry among toddlers from low-income families.
+\emph{Sleep Health}, \emph{5}(5), 459--465.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Barbaro2019}{}}%
+Barbaro, K. de. (2019). Automated sensing of daily activity: A new lens into development. \emph{Developmental Psychobiology}, \emph{61}(3), 444--464.
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-BruijnsTruelove2020}{}}%
+Bruijns, B.
+A., Truelove, S., Johnson, A.
+M., Gilliland, J., \& Tucker, P.
+(2020).
+Infants' and toddlers' physical activity and sedentary time as measured by accelerometry: A systematic review and meta-analysis.
+\emph{International Journal of Behavioral Nutrition and Physical Activity}, \emph{17}(1), 14.
+
 \leavevmode\vadjust pre{\hypertarget{ref-ChenSchneider2022}{}}%
@@
 \url{https://doi.org/10.1111/infa.12503}
+
+\leavevmode\vadjust pre{\hypertarget{ref-CliffReilly2009}{}}%
+Cliff, D.
+P., Reilly, J.
+J., \& Okely, A.
+D.
+(2009).
+Methodological considerations in using accelerometers to assess habitual physical activity in children aged 0--5 years.
+\emph{Journal of Science and Medicine in Sport}, \emph{12}(5), 557--567.
 
@@
 \url{https://doi.org/10.1016/j.smrv.2011.06.001}
+
+\leavevmode\vadjust pre{\hypertarget{ref-HagerTilton2017}{}}%
+Hager, E., Tilton, N., Wang, Y., Kapur, N., Arbaiza, R., Merry, B., \& Black, M.
+(2017).
+The home environment and toddler physical activity: An ecological momentary assessment study.
+\emph{Pediatric Obesity}, \emph{12}(1), 1--9.
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-LoboHall2019}{}}%
+Lobo, M.
+A., Hall, M.
+L., Greenspan, B., Rohloff, P., Prosser, L.
+A., \& Smith, B.
+A.
+(2019).
+Wearables for pediatric rehabilitation: How to optimally design and use products to meet the needs of users.
+\emph{Physical Therapy}, \emph{99}(6), 647--657.
+
 \leavevmode\vadjust pre{\hypertarget{ref-LuoFranchak2020}{}}%
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-NamPark2013}{}}%
+Nam, Y., \& Park, J.
+W.
+(2013).
+Child activity recognition based on cooperative fusion model of a triaxial accelerometer and a barometric pressure sensor.
+\emph{IEEE Journal of Biomedical and Health Informatics}, \emph{17}, 420--426.
+
 \leavevmode\vadjust pre{\hypertarget{ref-PatelShi2019}{}}%
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-Preece2009}{}}%
+Preece, S.
+J., Goulermas, J.
+Y., Kenney, L.
+P.
+J., \& Howard, D.
+(2009).
+A comparison of feature extraction methods for the classification of dynamic activities from accelerometer data.
+\emph{IEEE Transactions on Biomedical Engineering}, \emph{56}, 871--879.
+
 \leavevmode\vadjust pre{\hypertarget{ref-RasanenSeshadri2020}{}}%
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-RenDing2016}{}}%
+Ren, X., Ding, W., Crouter, S.
+E., Mu, Y., \& Xie, R.
+(2016).
+Activity recognition and intensity estimation in youth from accelerometer data aided by machine learning.
+\emph{Applied Intelligence}, \emph{45}(2), 512--529.
+
 \leavevmode\vadjust pre{\hypertarget{ref-SchnellerBentsen2017}{}}%
-SCHNELLER, M.
-B., BENTSEN, P., NIELSEN, G., BRØND, J.
-C., RIED-LARSEN, M., MYGIND, E., \& SCHIPPERIJN, J.
+Schneller, M.
+B., Bentsen, P., Nielsen, G., Brønd, J.
+C., Ried-Larsen, M., Mygind, E., \& Schipperijn, J.
 (2017).
@@
 Soska, K. C., \& Adolph, K. E. (2014). Postural position constrains multimodal object exploration in infants. \emph{Infancy}, \emph{19}, 138--161. \url{https://doi.org/10.1111/infa.12039}
+
+\leavevmode\vadjust pre{\hypertarget{ref-StewartNarayanan2018}{}}%
+Stewart, T., Narayanan, A., Hedayatrad, L., Neville, J., Mackay, L., \& Duncan, S.
+(2018).
+A dual-accelerometer system for classifying physical activity in children and adults.
+\emph{Medicine and Science in Sports and Exercise}, \emph{50}(12), 2595--2602.
 

### 2023-03-03 group_split_comparison/group_split.R
final data processed

@@ -46,7 +46,7 @@ group_mods <-map(ids, ~group_model(.x, training)) %>% set_names(ids)
 # group_select_mods <- map(ids, ~group_model(.x, training, bad_ids = bad_ids)) %>% set_names(ids)
 split_mods <- map(ids, ~split_model(.x, training)) %>% set_names(ids)
 
-# save(group_select_mods, split_mods, file = "group_split_comparison/models.RData")
+# save(group_mods, split_mods, file = "group_split_comparison/models.RData")
 # load( here("group_split_comparison","models.RData"))
 
 metrics <- function(rfmodel, testing) {

### 2023-03-03 manuscript.tex
knit

@@
 \midrule
-Overall Accuracy & 0.908 & 0.848 & 0.143 & 0.936 & 0.919 & 0.071\\
-Balanced Accuracy & 0.909 & 0.897 & 0.074 & 0.916 & 0.907 & 0.072\\
-F1 & 0.813 & 0.827 & 0.101 & 0.873 & 0.870 & 0.089\\
-Sensitivity & 0.840 & 0.836 & 0.123 & 0.848 & 0.840 & 0.121\\
-Pos Pred Value & 0.822 & 0.815 & 0.125 & 0.915 & 0.899 & 0.108\\
-Kappa & 0.769 & 0.757 & 0.160 & 0.846 & 0.820 & 0.141\\
+Overall Accuracy & 0.913 & 0.852 & 0.145 & 0.926 & 0.917 & 0.074\\
+Balanced Accuracy & 0.918 & 0.899 & 0.074 & 0.918 & 0.910 & 0.074\\
+F1 & 0.838 & 0.830 & 0.102 & 0.879 & 0.870 & 0.095\\
+Sensitivity & 0.859 & 0.840 & 0.122 & 0.856 & 0.846 & 0.125\\
+Pos Pred Value & 0.841 & 0.826 & 0.115 & 0.925 & 0.892 & 0.118\\
+Kappa & 0.783 & 0.764 & 0.156 & 0.862 & 0.826 & 0.148\\
 \bottomrule
@@
 \midrule
-Balanced Accuracy & Supine & 0.977 & 0.935 & 0.099 & 0.995 & 0.957 & 0.090\\
- & Prone & 0.997 & 0.950 & 0.110 & 0.983 & 0.921 & 0.132\\
- & Sitting & 0.932 & 0.861 & 0.151 & 0.973 & 0.943 & 0.071\\
- & Upright & 0.930 & 0.867 & 0.146 & 0.937 & 0.877 & 0.123\\
- & Held & 0.907 & 0.872 & 0.111 & 0.892 & 0.854 & 0.153\\ \midrule
-F1 & Supine & 0.952 & 0.873 & 0.173 & 0.987 & 0.937 & 0.130\\
- & Prone & 0.965 & 0.923 & 0.124 & 0.958 & 0.889 & 0.186\\
- & Sitting & 0.897 & 0.805 & 0.247 & 0.963 & 0.927 & 0.101\\
- & Upright & 0.781 & 0.748 & 0.235 & 0.868 & 0.769 & 0.227\\
- & Held & 0.813 & 0.783 & 0.163 & 0.862 & 0.786 & 0.240\\ \midrule
-Sensitivity & Supine & 1.000 & 0.945 & 0.125 & 1.000 & 0.948 & 0.141\\
- & Prone & 1.000 & 0.907 & 0.216 & 0.988 & 0.856 & 0.266\\
- & Sitting & 0.901 & 0.787 & 0.291 & 0.977 & 0.917 & 0.136\\
- & Upright & 0.878 & 0.767 & 0.275 & 0.893 & 0.790 & 0.255\\
- & Held & 0.867 & 0.774 & 0.229 & 0.817 & 0.722 & 0.309\\ \midrule
-Pos Pred Value & Supine & 0.973 & 0.802 & 0.300 & 1.000 & 0.939 & 0.131\\
- & Prone & 0.991 & 0.896 & 0.196 & 0.993 & 0.901 & 0.197\\
- & Sitting & 0.910 & 0.840 & 0.238 & 0.967 & 0.947 & 0.078\\
- & Upright & 0.848 & 0.743 & 0.285 & 0.875 & 0.817 & 0.205\\
- & Held & 0.917 & 0.803 & 0.247 & 0.949 & 0.872 & 0.232\\ \midrule
-Kappa & Supine & 0.932 & 0.770 & 0.307 & 0.984 & 0.913 & 0.169\\
- & Prone & 0.951 & 0.887 & 0.208 & 0.948 & 0.851 & 0.242\\
- & Sitting & 0.836 & 0.712 & 0.308 & 0.946 & 0.889 & 0.134\\
- & Upright & 0.729 & 0.700 & 0.277 & 0.829 & 0.744 & 0.232\\
- & Held & 0.786 & 0.733 & 0.219 & 0.845 & 0.744 & 0.280\\
+Balanced Accuracy & Supine & 0.974 & 0.923 & 0.109 & 0.998 & 0.963 & 0.082\\
+ & Prone & 0.997 & 0.953 & 0.105 & 0.965 & 0.924 & 0.124\\
+ & Sitting & 0.918 & 0.864 & 0.155 & 0.963 & 0.944 & 0.069\\
+ & Upright & 0.929 & 0.861 & 0.154 & 0.935 & 0.872 & 0.134\\
+ & Held & 0.915 & 0.884 & 0.107 & 0.861 & 0.851 & 0.147\\ \midrule
+F1 & Supine & 0.949 & 0.863 & 0.185 & 0.994 & 0.945 & 0.120\\
+ & Prone & 0.971 & 0.921 & 0.139 & 0.943 & 0.898 & 0.160\\
+ & Sitting & 0.912 & 0.810 & 0.250 & 0.955 & 0.926 & 0.098\\
+ & Upright & 0.782 & 0.758 & 0.214 & 0.850 & 0.756 & 0.259\\
+ & Held & 0.832 & 0.782 & 0.196 & 0.845 & 0.782 & 0.227\\ \midrule
+Sensitivity & Supine & 1.000 & 0.921 & 0.166 & 1.000 & 0.959 & 0.123\\
+ & Prone & 1.000 & 0.915 & 0.205 & 0.985 & 0.864 & 0.249\\
+ & Sitting & 0.911 & 0.795 & 0.288 & 0.985 & 0.916 & 0.136\\
+ & Upright & 0.870 & 0.755 & 0.293 & 0.907 & 0.783 & 0.274\\
+ & Held & 0.875 & 0.793 & 0.218 & 0.793 & 0.715 & 0.297\\ \midrule
+Pos Pred Value & Supine & 0.993 & 0.814 & 0.300 & 1.000 & 0.941 & 0.127\\
+ & Prone & 1.000 & 0.892 & 0.217 & 0.986 & 0.903 & 0.191\\
+ & Sitting & 0.926 & 0.841 & 0.242 & 0.974 & 0.947 & 0.071\\
+ & Upright & 0.842 & 0.738 & 0.292 & 0.888 & 0.803 & 0.235\\
+ & Held & 0.935 & 0.839 & 0.195 & 0.941 & 0.868 & 0.218\\ \midrule
+Kappa & Supine & 0.934 & 0.762 & 0.311 & 0.989 & 0.921 & 0.158\\
+ & Prone & 0.963 & 0.885 & 0.216 & 0.929 & 0.860 & 0.224\\
+ & Sitting & 0.838 & 0.720 & 0.313 & 0.930 & 0.890 & 0.129\\
+ & Upright & 0.735 & 0.688 & 0.284 & 0.820 & 0.730 & 0.264\\
+ & Held & 0.808 & 0.757 & 0.205 & 0.827 & 0.740 & 0.268\\
 \bottomrule

### 2023-03-03 manuscript.Rmd
final data processed
expalined outliers, got data about samples

@@
              legend.key = element_rect(fill = "white"))
+
+# SUBMITTED DATA IS BASED ON THE FOLLOWING IDS
+#9908  9909  9910  9911 10201 10203 10204 10301 10501 10601 10701 10702 10703 10704 10802 10901 11001 11002 11004 11102 11103 11104 11202 11501 11601 11901 12001 12002 12101 12302 12303 12501 12502 13001
 
@@
 
-In the current study, we report the feasibility and validity of body position classification over the full day in the home based on 35 testing sessions from 23 infants aged 4-14 months.
+In the current study, we report the feasibility and validity of body position classification over the full day in the home based on 35 testing sessions from 22 infants aged 4-14 months.
 Participants received a custom pair of infant leggings embedded with 4 IMUs (one on each ankle and one on each hip) and a video camera to collect ground truth data about infant body position.
@@
 
+What can we learn about the two outliers? What do they tell us about how this will work in the future.
+
+One outlier (106-4) had significant confusion between sitting and supine.
+Both their proximal and distal performance was bad, but only for sitting and supine.
+Watching the videos showed that this participant spent a long period of time in a seating device that was reclined between supine and sitting.
+Also spent long time in mother's arms in a similar supine/sitting position that was hard to classify.
+
+The other outlier (107-3) had good proximal accuracy but poor distal accuracy.
+Confusion between upright and suspended.
+Infant was in a baby walker for most of part 2.
+Moved more like a baby held by a CG than a baby walking upright.
+
+For both, other class performance was good, suggesting that it's about the particular events, not about the quality of the algorithm in applying to the participant overall.
+
 ## Goal 3: Examine the data quality of full-day home recordings

### 2023-03-03 manuscript.Rmd

@@
   filter(nap_period == 0 & exclude_period == 0 & !is.na(code) & !is.na(pos)) %>% 
-         mutate(pos = factor(pos, levels = pos_levels))
+         mutate(pos = factor(pos, levels = pos_levels)) %>% 
+  filter(file != "104/1")
 pt2_split <- read_csv("data/compiled_agreement_split.csv") %>% 
   filter(nap_period == 0 & exclude_period == 0 & !is.na(code) & !is.na(pos)) %>% 
-         mutate(pos = factor(pos, levels = pos_levels))
+         mutate(pos = factor(pos, levels = pos_levels)) %>% 
+  filter(file != "104/1")
 
-outlier_files <- c("107/3", "104/1", "106/1")
+outlier_files <- c("107/3", "106/1")
 filter_outliers <- . %>% filter(!(file %in% outlier_files))
@@
   scale_fill_manual(values = pal, guide = "none") + 
-  scale_shape_manual(guide = "none", values = 22:24) + 
+  scale_shape_manual(guide = "none", values = 22:23) + 
   scale_x_continuous(breaks = c(0,.5,1), labels = c(0, 50, 100), limits = c(-.07, 1.07), name = "Actual Prevalence (%)") + 
@@
   scale_fill_manual(values = pal, guide = "none") +
-  scale_shape_manual(guide = "none", values = 22:24) + 
+  scale_shape_manual(guide = "none", values = 22:23) + 
   coord_cartesian(ylim = c(-1, 1)) + xlab("") + 
@@
          nap_time = ifelse(is.na(sit_time), nap_time, NA)) %>% 
+  filter(id_uni != 10401) %>% 
   pivot_longer(cols = nap_time:upright_time, names_to = "position", values_to = "prop") %>% 
@@
          group = factor(ifelse(age <= 8, "Younger", "Older")),
-         nap_time = ifelse(is.na(sit_time), nap_time, NA))
+         nap_time = ifelse(is.na(sit_time), nap_time, NA)) %>% 
+  filter(id_uni != 10401)
 fullday_split <- read_csv("data/imu-compiled-split.csv") %>% 
@@
          group = factor(ifelse(age <= 8, "Younger", "Older")),
-         nap_time = ifelse(is.na(sit_time), nap_time, NA))
+         nap_time = ifelse(is.na(sit_time), nap_time, NA)) %>% 
+  filter(id_uni != 10401)
 fullday <- list("group" = fullday_group, "split" = fullday_split)

### 2023-03-03 manuscript.tex

@@
 \midrule
-Held & 0.02 & 0.05 & 0.75 & 0.60\\
+Held & 0.02 & 0.04 & 0.73 & 0.60\\
 Prone & 0.97 & 0.86 & 0.97 & 0.84\\
-Sitting & 0.70 & 0.88 & 0.91 & 0.95\\
-Supine & 0.84 & 0.92 & 0.94 & 0.97\\
-Upright & 0.62 & 0.84 & 0.99 & 0.94\\ \midrule
-Overall & 0.76 & 0.87 & 0.95 & 0.96\\
+Sitting & 0.79 & 0.97 & 0.91 & 0.95\\
+Supine & 0.88 & 0.98 & 0.94 & 0.97\\
+Upright & 0.63 & 0.83 & 0.99 & 0.95\\ \midrule
+Overall & 0.80 & 0.91 & 0.95 & 0.96\\
 \bottomrule
@@
 \midrule
-Held & 0.51 & 0.45 & 0.67 & 0.64\\
+Held & 0.51 & 0.46 & 0.67 & 0.63\\
 Prone & 0.96 & 0.90 & 0.96 & 0.89\\
-Sitting & 0.71 & 0.91 & 0.89 & 0.92\\
-Supine & 0.75 & 0.94 & 0.88 & 0.94\\
-Upright & 0.90 & 0.93 & 0.98 & 0.96\\ \midrule
-Overall & 0.79 & 0.93 & 0.92 & 0.94\\
+Sitting & 0.72 & 0.93 & 0.89 & 0.92\\
+Supine & 0.76 & 0.96 & 0.88 & 0.93\\
+Upright & 0.91 & 0.93 & 0.98 & 0.96\\ \midrule
+Overall & 0.80 & 0.94 & 0.92 & 0.94\\
 \bottomrule
@@
 \end{table}
+
+What can we learn about the two outliers? What do they tell us about how this will work in the future.
+
+One outlier (106-4) had significant confusion between sitting and supine.
+Both their proximal and distal performance was bad, but only for sitting and supine.
+Watching the videos showed that this participant spent a long period of time in a seating device that was reclined between supine and sitting.
+Also spent long time in mother's arms in a similar supine/sitting position that was hard to classify.
+
+The other outlier (107-3) had good proximal accuracy but poor distal accuracy.
+Confusion between upright and suspended.
+Infant was in a baby walker for most of part 2.
+Moved more like a baby held by a CG than a baby walking upright.
+
+For both, other class performance was good, suggesting that it's about the particular events, not about the quality of the algorithm in applying to the participant overall.
 
@@
 \midrule
-Upright & 7.6\% (8.9) & 18.6\% (7.4) & 9.9\% (13.1) & 18.7\% (8.4)\\
-Sitting & 26.3\% (12.1) & 44.4\% (10.1) & 20.2\% (16.3) & 46.9\% (13.3)\\
-Prone & 13.8\% (13.5) & 14.4\% (6.0) & 11.9\% (9.9) & 16.9\% (10.6)\\
-Supine & 37.9\% (23.2) & 14.0\% (8.4) & 39.4\% (30.3) & 10.0\% (9.5)\\
-Held & 12.7\% (6.9) & 8.5\% (5.4) & 17.6\% (19.9) & 7.4\% (7.6)\\
+Upright & 7.7\% (9.3) & 18.6\% (7.4) & 10.4\% (13.6) & 18.7\% (8.4)\\
+Sitting & 24.9\% (11.5) & 44.4\% (10.1) & 18.8\% (16.3) & 46.9\% (13.3)\\
+Prone & 14.4\% (13.8) & 14.4\% (6.0) & 12.5\% (10.2) & 16.9\% (10.6)\\
+Supine & 38.5\% (24.0) & 14.0\% (8.4) & 40.2\% (31.6) & 10.0\% (9.5)\\
+Held & 12.7\% (7.1) & 8.5\% (5.4) & 17.1\% (20.8) & 7.4\% (7.6)\\
 \bottomrule

### 2023-03-03 manuscript.tex

@@
 \midrule
-Overall Accuracy & 0.913 & 0.852 & 0.145 & 0.926 & 0.917 & 0.074\\
-Balanced Accuracy & 0.918 & 0.899 & 0.074 & 0.918 & 0.910 & 0.074\\
-F1 & 0.838 & 0.830 & 0.102 & 0.879 & 0.870 & 0.095\\
-Sensitivity & 0.859 & 0.840 & 0.122 & 0.856 & 0.846 & 0.125\\
-Pos Pred Value & 0.841 & 0.826 & 0.115 & 0.925 & 0.892 & 0.118\\
-Kappa & 0.783 & 0.764 & 0.156 & 0.862 & 0.826 & 0.148\\
+Overall Accuracy & 0.904 & 0.843 & 0.165 & 0.929 & 0.915 & 0.076\\
+Balanced Accuracy & 0.923 & 0.898 & 0.077 & 0.918 & 0.909 & 0.072\\
+F1 & 0.834 & 0.826 & 0.115 & 0.877 & 0.869 & 0.091\\
+Sensitivity & 0.870 & 0.838 & 0.127 & 0.856 & 0.845 & 0.119\\
+Pos Pred Value & 0.828 & 0.816 & 0.127 & 0.932 & 0.900 & 0.109\\
+Kappa & 0.766 & 0.759 & 0.166 & 0.836 & 0.825 & 0.143\\
 \bottomrule
@@
 \midrule
-Balanced Accuracy & Supine & 0.974 & 0.923 & 0.109 & 0.998 & 0.963 & 0.082\\
- & Prone & 0.997 & 0.953 & 0.105 & 0.965 & 0.924 & 0.124\\
- & Sitting & 0.918 & 0.864 & 0.155 & 0.963 & 0.944 & 0.069\\
- & Upright & 0.929 & 0.861 & 0.154 & 0.935 & 0.872 & 0.134\\
- & Held & 0.915 & 0.884 & 0.107 & 0.861 & 0.851 & 0.147\\ \midrule
-F1 & Supine & 0.949 & 0.863 & 0.185 & 0.994 & 0.945 & 0.120\\
- & Prone & 0.971 & 0.921 & 0.139 & 0.943 & 0.898 & 0.160\\
- & Sitting & 0.912 & 0.810 & 0.250 & 0.955 & 0.926 & 0.098\\
- & Upright & 0.782 & 0.758 & 0.214 & 0.850 & 0.756 & 0.259\\
- & Held & 0.832 & 0.782 & 0.196 & 0.845 & 0.782 & 0.227\\ \midrule
-Sensitivity & Supine & 1.000 & 0.921 & 0.166 & 1.000 & 0.959 & 0.123\\
- & Prone & 1.000 & 0.915 & 0.205 & 0.985 & 0.864 & 0.249\\
- & Sitting & 0.911 & 0.795 & 0.288 & 0.985 & 0.916 & 0.136\\
- & Upright & 0.870 & 0.755 & 0.293 & 0.907 & 0.783 & 0.274\\
- & Held & 0.875 & 0.793 & 0.218 & 0.793 & 0.715 & 0.297\\ \midrule
-Pos Pred Value & Supine & 0.993 & 0.814 & 0.300 & 1.000 & 0.941 & 0.127\\
- & Prone & 1.000 & 0.892 & 0.217 & 0.986 & 0.903 & 0.191\\
- & Sitting & 0.926 & 0.841 & 0.242 & 0.974 & 0.947 & 0.071\\
- & Upright & 0.842 & 0.738 & 0.292 & 0.888 & 0.803 & 0.235\\
- & Held & 0.935 & 0.839 & 0.195 & 0.941 & 0.868 & 0.218\\ \midrule
-Kappa & Supine & 0.934 & 0.762 & 0.311 & 0.989 & 0.921 & 0.158\\
- & Prone & 0.963 & 0.885 & 0.216 & 0.929 & 0.860 & 0.224\\
- & Sitting & 0.838 & 0.720 & 0.313 & 0.930 & 0.890 & 0.129\\
- & Upright & 0.735 & 0.688 & 0.284 & 0.820 & 0.730 & 0.264\\
- & Held & 0.808 & 0.757 & 0.205 & 0.827 & 0.740 & 0.268\\
+Balanced Accuracy & Supine & 0.973 & 0.927 & 0.121 & 0.994 & 0.959 & 0.083\\
+ & Prone & 0.997 & 0.949 & 0.112 & 0.978 & 0.924 & 0.126\\
+ & Sitting & 0.915 & 0.862 & 0.147 & 0.967 & 0.941 & 0.074\\
+ & Upright & 0.934 & 0.867 & 0.153 & 0.933 & 0.879 & 0.120\\
+ & Held & 0.903 & 0.876 & 0.114 & 0.864 & 0.842 & 0.155\\ \midrule
+F1 & Supine & 0.962 & 0.859 & 0.225 & 0.987 & 0.941 & 0.120\\
+ & Prone & 0.972 & 0.911 & 0.164 & 0.943 & 0.894 & 0.171\\
+ & Sitting & 0.907 & 0.804 & 0.237 & 0.957 & 0.925 & 0.106\\
+ & Upright & 0.818 & 0.772 & 0.217 & 0.864 & 0.771 & 0.234\\
+ & Held & 0.841 & 0.793 & 0.164 & 0.839 & 0.764 & 0.241\\ \midrule
+Sensitivity & Supine & 1.000 & 0.904 & 0.224 & 1.000 & 0.952 & 0.124\\
+ & Prone & 1.000 & 0.906 & 0.219 & 0.985 & 0.865 & 0.254\\
+ & Sitting & 0.913 & 0.819 & 0.257 & 0.950 & 0.911 & 0.141\\
+ & Upright & 0.880 & 0.765 & 0.291 & 0.883 & 0.798 & 0.243\\
+ & Held & 0.864 & 0.780 & 0.235 & 0.789 & 0.698 & 0.315\\ \midrule
+Pos Pred Value & Supine & 0.998 & 0.832 & 0.291 & 1.000 & 0.939 & 0.128\\
+ & Prone & 1.000 & 0.889 & 0.228 & 0.972 & 0.894 & 0.193\\
+ & Sitting & 0.924 & 0.805 & 0.277 & 0.974 & 0.951 & 0.077\\
+ & Upright & 0.882 & 0.742 & 0.294 & 0.939 & 0.820 & 0.234\\
+ & Held & 0.900 & 0.812 & 0.238 & 0.936 & 0.884 & 0.171\\ \midrule
+Kappa & Supine & 0.942 & 0.766 & 0.326 & 0.984 & 0.916 & 0.159\\
+ & Prone & 0.962 & 0.875 & 0.236 & 0.933 & 0.855 & 0.232\\
+ & Sitting & 0.820 & 0.708 & 0.313 & 0.929 & 0.889 & 0.139\\
+ & Upright & 0.773 & 0.701 & 0.291 & 0.804 & 0.743 & 0.242\\
+ & Held & 0.809 & 0.743 & 0.220 & 0.826 & 0.721 & 0.279\\
 \bottomrule

### 2023-03-04 group_split_comparison/group_split.R
updated metrics

@@ -11,7 +11,7 @@ library(rstatix)
 
 #LOAD DATA
 load( here("group_split_comparison","compiled_data_lite.RData"))
-ntree = 550
+ntree = 1100
 mtry = 44
 not_all_na <- function(x) any(!is.na(x))
 

### 2023-03-04 manuscript.tex
render ms

@@
 \midrule
-Overall Accuracy & 0.904 & 0.843 & 0.165 & 0.929 & 0.915 & 0.076\\
-Balanced Accuracy & 0.923 & 0.898 & 0.077 & 0.918 & 0.909 & 0.072\\
-F1 & 0.834 & 0.826 & 0.115 & 0.877 & 0.869 & 0.091\\
-Sensitivity & 0.870 & 0.838 & 0.127 & 0.856 & 0.845 & 0.119\\
-Pos Pred Value & 0.828 & 0.816 & 0.127 & 0.932 & 0.900 & 0.109\\
-Kappa & 0.766 & 0.759 & 0.166 & 0.836 & 0.825 & 0.143\\
+Overall Accuracy & 0.894 & 0.846 & 0.131 & 0.933 & 0.916 & 0.072\\
+Balanced Accuracy & 0.913 & 0.891 & 0.077 & 0.911 & 0.907 & 0.071\\
+F1 & 0.822 & 0.816 & 0.106 & 0.882 & 0.865 & 0.091\\
+Sensitivity & 0.856 & 0.825 & 0.127 & 0.847 & 0.841 & 0.119\\
+Pos Pred Value & 0.826 & 0.810 & 0.125 & 0.928 & 0.899 & 0.107\\
+Kappa & 0.768 & 0.746 & 0.161 & 0.849 & 0.821 & 0.143\\
 \bottomrule
@@
 \midrule
-Balanced Accuracy & Supine & 0.973 & 0.927 & 0.121 & 0.994 & 0.959 & 0.083\\
- & Prone & 0.997 & 0.949 & 0.112 & 0.978 & 0.924 & 0.126\\
- & Sitting & 0.915 & 0.862 & 0.147 & 0.967 & 0.941 & 0.074\\
- & Upright & 0.934 & 0.867 & 0.153 & 0.933 & 0.879 & 0.120\\
- & Held & 0.903 & 0.876 & 0.114 & 0.864 & 0.842 & 0.155\\ \midrule
-F1 & Supine & 0.962 & 0.859 & 0.225 & 0.987 & 0.941 & 0.120\\
- & Prone & 0.972 & 0.911 & 0.164 & 0.943 & 0.894 & 0.171\\
- & Sitting & 0.907 & 0.804 & 0.237 & 0.957 & 0.925 & 0.106\\
- & Upright & 0.818 & 0.772 & 0.217 & 0.864 & 0.771 & 0.234\\
- & Held & 0.841 & 0.793 & 0.164 & 0.839 & 0.764 & 0.241\\ \midrule
-Sensitivity & Supine & 1.000 & 0.904 & 0.224 & 1.000 & 0.952 & 0.124\\
- & Prone & 1.000 & 0.906 & 0.219 & 0.985 & 0.865 & 0.254\\
- & Sitting & 0.913 & 0.819 & 0.257 & 0.950 & 0.911 & 0.141\\
- & Upright & 0.880 & 0.765 & 0.291 & 0.883 & 0.798 & 0.243\\
- & Held & 0.864 & 0.780 & 0.235 & 0.789 & 0.698 & 0.315\\ \midrule
-Pos Pred Value & Supine & 0.998 & 0.832 & 0.291 & 1.000 & 0.939 & 0.128\\
- & Prone & 1.000 & 0.889 & 0.228 & 0.972 & 0.894 & 0.193\\
- & Sitting & 0.924 & 0.805 & 0.277 & 0.974 & 0.951 & 0.077\\
- & Upright & 0.882 & 0.742 & 0.294 & 0.939 & 0.820 & 0.234\\
- & Held & 0.900 & 0.812 & 0.238 & 0.936 & 0.884 & 0.171\\ \midrule
-Kappa & Supine & 0.942 & 0.766 & 0.326 & 0.984 & 0.916 & 0.159\\
- & Prone & 0.962 & 0.875 & 0.236 & 0.933 & 0.855 & 0.232\\
- & Sitting & 0.820 & 0.708 & 0.313 & 0.929 & 0.889 & 0.139\\
- & Upright & 0.773 & 0.701 & 0.291 & 0.804 & 0.743 & 0.242\\
- & Held & 0.809 & 0.743 & 0.220 & 0.826 & 0.721 & 0.279\\
+Balanced Accuracy & Supine & 0.964 & 0.925 & 0.101 & 0.994 & 0.959 & 0.086\\
+ & Prone & 0.996 & 0.941 & 0.121 & 0.968 & 0.915 & 0.135\\
+ & Sitting & 0.909 & 0.862 & 0.141 & 0.963 & 0.942 & 0.070\\
+ & Upright & 0.907 & 0.849 & 0.155 & 0.938 & 0.874 & 0.123\\
+ & Held & 0.901 & 0.871 & 0.110 & 0.868 & 0.844 & 0.154\\ \midrule
+F1 & Supine & 0.943 & 0.867 & 0.176 & 0.986 & 0.938 & 0.127\\
+ & Prone & 0.979 & 0.897 & 0.190 & 0.957 & 0.880 & 0.191\\
+ & Sitting & 0.879 & 0.803 & 0.217 & 0.951 & 0.924 & 0.099\\
+ & Upright & 0.738 & 0.722 & 0.246 & 0.844 & 0.766 & 0.231\\
+ & Held & 0.771 & 0.779 & 0.151 & 0.839 & 0.768 & 0.238\\ \midrule
+Sensitivity & Supine & 1.000 & 0.905 & 0.180 & 1.000 & 0.954 & 0.129\\
+ & Prone & 1.000 & 0.894 & 0.234 & 0.974 & 0.849 & 0.272\\
+ & Sitting & 0.910 & 0.811 & 0.257 & 0.964 & 0.915 & 0.136\\
+ & Upright & 0.837 & 0.730 & 0.297 & 0.891 & 0.786 & 0.253\\
+ & Held & 0.852 & 0.772 & 0.228 & 0.773 & 0.702 & 0.312\\ \midrule
+Pos Pred Value & Supine & 0.995 & 0.828 & 0.292 & 1.000 & 0.932 & 0.134\\
+ & Prone & 0.987 & 0.877 & 0.236 & 1.000 & 0.892 & 0.210\\
+ & Sitting & 0.896 & 0.802 & 0.261 & 0.972 & 0.945 & 0.079\\
+ & Upright & 0.839 & 0.739 & 0.283 & 0.923 & 0.825 & 0.228\\
+ & Held & 0.852 & 0.794 & 0.240 & 0.943 & 0.899 & 0.134\\ \midrule
+Kappa & Supine & 0.907 & 0.764 & 0.295 & 0.983 & 0.912 & 0.166\\
+ & Prone & 0.968 & 0.860 & 0.259 & 0.942 & 0.841 & 0.246\\
+ & Sitting & 0.816 & 0.702 & 0.297 & 0.915 & 0.887 & 0.127\\
+ & Upright & 0.707 & 0.673 & 0.281 & 0.822 & 0.741 & 0.236\\
+ & Held & 0.732 & 0.726 & 0.209 & 0.826 & 0.727 & 0.277\\
 \bottomrule

### 2023-03-04 group_split_comparison/group_split.R
Revert "updated metrics"
This reverts commit c2036f40614ba5d0c5cf475c25c0db9b83734f5b.

@@ -11,7 +11,7 @@ library(rstatix)
 
 #LOAD DATA
 load( here("group_split_comparison","compiled_data_lite.RData"))
-ntree = 1100
+ntree = 550
 mtry = 44
 not_all_na <- function(x) any(!is.na(x))
 

### 2023-03-07 manuscript.Rmd
ppt stats

@@
 
-## Participants
+## Participants and Design
+
+```{r participants}
+ppts <- read_csv("data/imu-compiled.csv") %>% 
+  mutate(id_uni = factor(id*100+session),
+         group = factor(ifelse(age <= 8, "Younger", "Older")))
+num_younger <- ppts %>% filter(group == "Younger") %>% .$id %>% unique(.) %>% length(.)
+num_older <-  ppts %>% filter(group == "Older") %>% .$id %>% unique(.) %>% length(.)
+num_sessions <- ppts %>% group_by(group, id, session) %>% slice_head(n = 1) %>% count %>% group_by(group, id) %>% count
+age <- ppts %>% group_by(group) %>% get_summary_stats(age) %>% column_to_rownames("group")
+```
+
+Infants were recruited in one of two age groups: Younger infants were 4-7 months and older infants were 11-14 months.
+Families were recruited through social media advertisements and from community events in Southern California.
+There were `r num_younger` infants in the 4-7 month group (X female) and `r num_older` in the 11-14 month group (X female).
+Only 1 of session was excluded due to a technical error---one of the four IMU sensors failed to record, resulting in an unusable recording.
+The ethnicity for infants was reported as X Hispanix/Latinx and X for not Hispanix/Latinx.
+Race was reported as X Caucasian, X Asian, X ~~~ and X others.
+Parents were compensated \$30 in cash for each visit they completed.
+The BLINDED Institutional Review Board reviewed and approved all procedures associated with the study.
+All participants gave their informed consent to participate.
+<!-- TODO RACE ETHNICITY SEX -->
+
+Most participants were tested in a single session (`r nrow(num_sessions %>% filter(n == 1))`), but `r nrow(num_sessions %>% filter(n > 1))` participants contributed between 2-4 sessions as part of an ongoing longitudinal study.
+Across the two age groups, we report data on a total of `r length(unique(ppts$id_uni))` sessions, with `r num_sessions %>% filter(group == "Younger") %>% pull(n) %>% sum` from younger infants and `r num_sessions %>% filter(group == "Older") %>% pull(n) %>% sum` from older infants.
+Across sessions, younger infants' age ranged from `r age["Younger","min"]` to `r age["Younger","max"]` (*M* = `r age["Younger","mean"]`) and older infants' age ranged from `r age["Older","min"]` to `r age["Older","max"]` (*M* = `r age["Older","mean"]`).
+
+<!-- with X including permission for data to be shared on Databrary.  -->
 
@@
 (metric_plots[[1]] + metric_plots[[2]])/(metric_plots[[3]]+metric_plots[[4]])/(metric_plots[[5]] + metric_plots[[6]])
-
-#Worst three Kappa individuals are 9, 111-4, and 120-2
-#Worst three Kappa groups are 9, 110-4
-#Worst three Overall Accuracy (grouped) are 104-1, 106-1, 108-2 (first two are pt 2 outliers)
 

### 2023-03-07 manuscript.tex
ppt stats

@@
 
-\hypertarget{participants}{%
-\subsection{Participants}\label{participants}}
+\hypertarget{participants-and-design}{%
+\subsection{Participants and Design}\label{participants-and-design}}
+
+Infants were recruited in one of two age groups: Younger infants were 4-7 months and older infants were 11-14 months.
+Families were recruited through social media advertisements and from community events in Southern California.
+There were 9 infants in the 4-7 month group (X female) and 14 in the 11-14 month group (X female).
+Only 1 of session was excluded due to a technical error---one of the four IMU sensors failed to record, resulting in an unusable recording.
+The ethnicity for infants was reported as X Hispanix/Latinx and X for not Hispanix/Latinx.
+Race was reported as X Caucasian, X Asian, X \textasciitilde\textasciitilde\textasciitilde{} and X others.
+Parents were compensated \$30 in cash for each visit they completed.
+The BLINDED Institutional Review Board reviewed and approved all procedures associated with the study.
+All participants gave their informed consent to participate.
+
+Most participants were tested in a single session (17), but 6 participants contributed between 2-4 sessions as part of an ongoing longitudinal study.
+Across the two age groups, we report data on a total of 34 sessions, with 14 from younger infants and 20 from older infants.
+Across sessions, younger infants' age ranged from 3.84 to 7.16 (\emph{M} = 4.99) and older infants' age ranged from 10.74 to 14.23 (\emph{M} = 11.75).
 

### 2023-03-07 manuscript.Rmd
materials section

@@
 
-Infants were recruited in one of two age groups: Younger infants were 4-7 months and older infants were 11-14 months.
+Infants were recruited in one of two age groups: *Younger* infants were between 4 and 7 months and *Older* infants were between 11 and 14 months.
+There were `r num_younger` infants in the 4-7 month group (X female) and `r num_older` in the 11-14 month group (X female).
 Families were recruited through social media advertisements and from community events in Southern California.
-There were `r num_younger` infants in the 4-7 month group (X female) and `r num_older` in the 11-14 month group (X female).
-Only 1 of session was excluded due to a technical error---one of the four IMU sensors failed to record, resulting in an unusable recording.
 The ethnicity for infants was reported as X Hispanix/Latinx and X for not Hispanix/Latinx.
@@
 All participants gave their informed consent to participate.
-<!-- TODO RACE ETHNICITY SEX -->
+<!-- TODO RACE ETHNICITY SEX, COUNT BY HAND -->
 
-Most participants were tested in a single session (`r nrow(num_sessions %>% filter(n == 1))`), but `r nrow(num_sessions %>% filter(n > 1))` participants contributed between 2-4 sessions as part of an ongoing longitudinal study.
-Across the two age groups, we report data on a total of `r length(unique(ppts$id_uni))` sessions, with `r num_sessions %>% filter(group == "Younger") %>% pull(n) %>% sum` from younger infants and `r num_sessions %>% filter(group == "Older") %>% pull(n) %>% sum` from older infants.
+Most participants were tested in a single session (*n* = `r nrow(num_sessions %>% filter(n == 1))`), but `r nrow(num_sessions %>% filter(n > 1))` participants contributed between 2-4 sessions as part of an ongoing longitudinal study.
+Only 1 session was excluded due to a technical error---one of the four IMU sensors failed to record, resulting in an unusable set of data for classification.
+Across the two age groups, we report data on a total of `r length(unique(ppts$id_uni))` sessions, with `r num_sessions %>% filter(group == "Younger") %>% pull(n) %>% sum` sessions from younger infants and `r num_sessions %>% filter(group == "Older") %>% pull(n) %>% sum` sessions from older infants.
 Across sessions, younger infants' age ranged from `r age["Younger","min"]` to `r age["Younger","max"]` (*M* = `r age["Younger","mean"]`) and older infants' age ranged from `r age["Older","min"]` to `r age["Older","max"]` (*M* = `r age["Older","mean"]`).
@@
 ## Apparatus
+
+Four inertial movement units (IMUs) were used to record infant movement across the day (MC10 Biostamp).
+A custom garment was made to hold the IMUs: Internal pockets were sewn into a snug-fitting pair of infant leggings so that IMUs would stay close to the body and so that infants could not pull out the sensors.
+On each side of the body there was a pocket over the hip and a pocket just above the ankle.
+Each sensor recorded accelerometer and gyroscope data at 62.5 Hz throughout the day, with sufficient battery and on-device storage to record for approximately 12 hours.
+Infants also wore a LENA&reg; recorder throughout the day in the front pocket of a LENA&reg; shirt, located near the infant’s chest.
+
+Video recordings were captured using an action camera on a miniature tripod (Insta360 ONE R) that caregivers placed in the room that the infant was in.
+The proprietary "Boosted Battery Base" attached to the action camera to allow for a total of 3 hours of recording.
+However, as described below, this created two video files separated by a variable gap of approximately 40 s.
+Caregivers also received a log sheet to record times that infants napped as well as times that the sensor garment was removed from the infant (e.g., baths, diaper changes).
 

### 2023-03-07 manuscript.tex
materials section

@@
 
-Infants were recruited in one of two age groups: Younger infants were 4-7 months and older infants were 11-14 months.
+Infants were recruited in one of two age groups: \emph{Younger} infants were between 4 and 7 months and \emph{Older} infants were between 11 and 14 months.
+There were 9 infants in the 4-7 month group (X female) and 14 in the 11-14 month group (X female).
 Families were recruited through social media advertisements and from community events in Southern California.
-There were 9 infants in the 4-7 month group (X female) and 14 in the 11-14 month group (X female).
-Only 1 of session was excluded due to a technical error---one of the four IMU sensors failed to record, resulting in an unusable recording.
 The ethnicity for infants was reported as X Hispanix/Latinx and X for not Hispanix/Latinx.
@@
 
-Most participants were tested in a single session (17), but 6 participants contributed between 2-4 sessions as part of an ongoing longitudinal study.
-Across the two age groups, we report data on a total of 34 sessions, with 14 from younger infants and 20 from older infants.
+Most participants were tested in a single session (\emph{n} = 17), but 6 participants contributed between 2-4 sessions as part of an ongoing longitudinal study.
+Only 1 session was excluded due to a technical error---one of the four IMU sensors failed to record, resulting in an unusable set of data for classification.
+Across the two age groups, we report data on a total of 34 sessions, with 14 sessions from younger infants and 20 sessions from older infants.
 Across sessions, younger infants' age ranged from 3.84 to 7.16 (\emph{M} = 4.99) and older infants' age ranged from 10.74 to 14.23 (\emph{M} = 11.75).
@@
 \subsection{Apparatus}\label{apparatus}}
+
+Four inertial movement units (IMUs) were used to record infant movement across the day (MC10 Biostamp).
+A custom garment was made to hold the IMUs: Internal pockets were sewn into a snug-fitting pair of infant leggings so that IMUs would stay close to the body and so that infants could not pull out the sensors.
+On each side of the body there was a pocket over the hip and a pocket just above the ankle.
+Each sensor recorded accelerometer and gyroscope data at 62.5 Hz throughout the day, with sufficient battery and on-device storage to record for approximately 12 hours.
+Infants also wore a LENA® recorder throughout the day in the front pocket of a LENA® shirt, located near the infant's chest.
+
+Video recordings were captured using an action camera on a miniature tripod (Insta360 ONE R) that caregivers placed in the room that the infant was in.
+The proprietary ``Boosted Battery Base'' attached to the action camera to allow for a total of 3 hours of recording.
+However, as described below, this created two video files separated by a variable gap of approximately 40 s.
+Caregivers also received a log sheet to record times that infants napped as well as times that the sensor garment was removed from the infant (e.g., baths, diaper changes).
 

### 2023-03-07 manuscript.Rmd
start of procedure

@@
 
-## Body position annotation
+On the day of the visit, a researcher arrived at the participant’s home in the morning between and prepared all the equipment at the doorstep.
+In order to synchronize all three recording devices, the researcher began by first turning on the video camera and the LENA audio device (the IMUs were already configured and placed at arrival because they required a proprietary sensor dock to begin recording in the laboratory).
+To create an easily recognizable synchronization point between the video recording and IMU data, the researcher dropped or slammed the leggings (containing the IMUs) on a surface in view of the camera, as in [@FranchakScott2021].
+All the equipment---now recording and synchronized---was placed inside a large bucket and left outside the family's front door.
+
+Returning to their vehicle, the researcher then called the caregiver and walked them through all of the paperwork, having them sign in the designated areas.
+The caregiver was then instructed to place the camera in an area that captured the majority of the room.
+They then were asked to put the pair of leggings and shirt on their infant.
+
+The researcher then had the caregiver complete a number of guided activities with their infant.
+Within view of the camera the caregiver was asked to place their infant in several different positions.
+These positions were as follows: lying supine, lying prone, sitting on the floor, standing upright, held by the caregiver while they walked back and forth, crawling or walking independently, and sitting restrained.
+The positions were assisted or unassisted and were timed by the researcher.
+Except for lying supine and held by a caregiver which lasted for a duration of 3-4 minutes, all other positions lasted for 1 minute.The caregiver was then instructed to remove the leggings from the child, raise it above their head and drop it to the floor standing approximately six feet away from the camera in order to capture another identifiable synchronization point.
+Once completed, the caregiver was then instructed to play with their infant for 10 minutes within view of the camera.
+Afterwards they were to go about their day as usual.
+The sensors were to be worn up until the child’s bedtime.
+The caregiver logged the times the sensors had to be removed or the child took a nap.
+
+The Camera (Insta360) only had the battery life to record for ~three hours. Afterwards, the participant was instructed at the 3.5 hour mark to put the camera back in the bucket.
+
+The following day a researcher picked up the equipment, verified the paperwork was signed, and compensated the participant.
+
+## Body Position Annotation
+
+The insta360 recordings were used by trained human coders to identify infant body position: supine, prone, sitting, sitting restrained, upright, or held. All coding was done using Datavyu software.
+
+When the infant was lying on their back or was reclined 45 degrees or more, supine was coded.
+Supine was also coded when the infant was laying on their side.
+Prone was coded when the infant was lying on their stomach, was on all fours in a downward dog position, or was crawling.
+When an infant was sitting with their butt on a surface, such as on the floor or a caregiver’s lap, the time frame was coded sitting on a surface by researchers.
+Sitting was also coded when the infant was in a kneeling-sit position, in which their knees were on the ground with their legging tucked underneath.
+When an infant was in a seating device that restricted movement, such as a high chair, sitting restrained was coded.
+Upright was coded when the infant was standing on the ground with two feet, walking, or kneeling with their thighs and torso in a straight line.
+When an infant was carried by a caregiver, held was coded.
+However, when the caregiver was sitting with the infant in their lap the infant’s body position was coded as if the caregiver was a surface (e.g., if the infant was sitting on the caregiver’s lap this was coded as sitting).
+The positions that did not fit in one of the six categories or when the infant was out of view were excluded.
+Periods when the leggings were being adjusted or taken off the infant were also excluded.
+
+Primary coding was completed for the full length of the video, while reliability coding was completed for the first thirty minutes.
+Interrater reliability between coders … (overall agreement = XX.X%, kappa = X.XX).
 

### 2023-03-07 manuscript.Rmd
procedure

--- 2023-03-07 manuscript.Rmd
@@
 Synchronizing real-time data to other data streams helps to reveal sources of variability within an individual [e.g., @WassPhillips2022].
-Full-day recordings are essential for capturing experiences across the heterogeniety of daily routines that moderate behavior (e.g., play, feeding, errands) [@KadookaCaufield2021].
+Full-day recordings are essential for capturing experiences across the heterogeniety of daily routines that moderate behavior (e.g., play, feeding, errands) [@KadookaCaufield2021; @Tamis-LeMondaCustode2018].
 "Burstiness" of behavior means that long recordings are needed to capture clusters of events amid long periods in which they may be absent [@BarbaroFausey2022; @WarlaumontSobowale2021].
@@
 
-```{r exemplartimeline, out.width="0.99\\linewidth", include=TRUE, fig.align="center", fig.cap="Example Timeline Caption.", echo=FALSE}
+```{r exemplar-timeline, out.width="0.99\\linewidth", include=TRUE, fig.align="center", fig.cap="Example Timeline Caption.", echo=FALSE}
 knitr::include_graphics("figures/timeline.png")
@@
 
-## Body position annotation
+<!-- AVERAGE START TIME -->
+Figure \@ref(fig:exemplar-timeline)) shows an exemplar timeline of the entire procedure and recording periods for a single participant.
+On the day of the visit, a researcher arrived at the participant’s home in the morning between and prepared all the equipment at the doorstep.
+In order to synchronize all three recording devices, the researcher began by first turning on the video camera and the LENA audio device (the IMUs were already configured and placed at arrival because they required a proprietary sensor dock to begin recording in the laboratory).
+To create an easily recognizable synchronization point between the video recording and IMU data, the researcher dropped or struck the leggings (containing the IMUs) on a surface in view of the camera, as in [@FranchakScott2021].
+All the equipment---once recording and with synchronization information recorded---was placed inside a large bucket and left outside the family's front door.
+
+The researcher then called the caregiver on the phone and walked them through a set of procedures needed to properly set up the equipment and record video for classifier training and testing.
+At the start of this "guided call", the caregiver was instructed to place the camera in an area that captured the majority of the room.
+Next, they were asked to put the pair of leggings and shirt on their infant, with the researcher providing guidance about how to correctly orient the garments.
+
+Afterwards, the researcher asked the caregiver to complete a number of guided activities with their infant.
+Within view of the camera, the caregiver was asked to place their infant in several different positions: lying supine, lying prone, sitting on the floor, standing upright, held by the caregiver while the caregiver walked back and forth, crawling, walking, and sitting in a restrained seat (e.g., high chair).
+Depending on the infants' age and motor skill level, the positions could be done independently or were completed with assistance from the caregiver.
+The researcher kept time to ensure at least 1 minute of behavior for each activity.
+Once completed, the caregiver was then instructed to play with their infant for 10 minutes within view of the camera to collect additional training data with the infant in positions that would be typical of play.
+
+Afterwards, they were to go about their day as usual with the infant wearing the sensor garment until their bedtime, only taking off the sensor for naps, baths, and diaper changes.
+The caregiver logged the times the sensors were removed (blank areas in the timeline in Figure \@ref(fig:exemplar-timeline))) or the child took a nap (gray areas in the timeline in Figure \@ref(fig:exemplar-timeline))) so that those times could be excluded from analysis.
+The following day a researcher picked up the equipment, verified the paperwork was signed, and compensated the participant.
+
+Because the camera only had the battery life to record for ~3 hours (divided into two 90-minute video files), this divided the day into different periods for analysis.
+As seen in the bottom of Figure \@ref(fig:exemplar-timeline)), the *video period* comprised the first three hours of recording starting from the researcher's arrival when they turned on the camera.
+The first 90-minute video file, termed the *proximal comparison*, contained the activities during the guided call followed by a period of infants and caregivers resuming their normal activities.
+Because this video contained the synchronization point, the data in this period had temporal synchrony between IMU and video data that contained errors of no more than 30-60 ms (1-2 video frames).
+When the first video file ended, a second video file was recorded, termed the *distal comparison*.
+This video recorded the next 90 minutes of natural activity.
+However, because there was a variable gap of ~40 s between the two videos, temporal synchrony in the distal comparison video was approximate containing offsets of ~ 5-10 s.
+<!-- HOW MANY PPTS DON"T HAVE DISTAL COMPARISON VIDEOS? -->
+
+## Body Position Annotation
+
+The proximal and distal comparison videos for each participant were annotated by trained human coders to identify infant body position: supine, prone, sitting, sitting restrained, upright, or held.
+All coding was done using Datavyu software (datavyu.org).
+
+When the infant was lying on their back or was reclined 45 degrees or more, supine was coded.
+Supine was also coded when the infant was laying on their side.
+Prone was coded when the infant was lying on their stomach, was on all fours in a downward dog position, or was crawling.
+When an infant was sitting with their butt on a surface, such as on the floor or a caregiver’s lap, the time frame was coded sitting on a surface by researchers.
+Sitting was also coded when the infant was in a kneeling-sit position, in which their knees were on the ground with their legging tucked underneath.
+When an infant was in a seating device that restricted movement, such as a high chair, sitting restrained was coded.
+Upright was coded when the infant was standing on the ground with two feet, walking, or kneeling with their thighs and torso in a straight line.
+When an infant was carried by a caregiver, held was coded.
+However, when the caregiver was sitting with the infant in their lap the infant’s body position was coded as if the caregiver was a surface (e.g., if the infant was sitting on the caregiver’s lap this was coded as sitting).
+The positions that did not fit in one of the six categories or when the infant was out of view were excluded.
+Periods when the leggings were being adjusted or taken off the infant were also excluded.
+
+Primary coding was completed for the full length of the video, while reliability coding was completed for the first thirty minutes.
+Interrater reliability between coders … (overall agreement = XX.X%, kappa = X.XX).
 

### 2023-03-07 manuscript.Rmd
coding

@@
 
-The proximal and distal comparison videos for each participant were annotated by trained human coders to identify infant body position: supine, prone, sitting, sitting restrained, upright, or held.
+The proximal and distal comparison videos for each participant were annotated by trained human coders to identify infant *body position* into one of 5 mutually-exclusive categories: supine, prone, sitting, upright, or held by caregiver.
 All coding was done using Datavyu software (datavyu.org).
 
-When the infant was lying on their back or was reclined 45 degrees or more, supine was coded.
-Supine was also coded when the infant was laying on their side.
+Supine was coded when the infant was lying on their back or was reclined up to a 45 degrees angle.
+Supine was also coded in the rare cases when the infant was laying on their side.
 Prone was coded when the infant was lying on their stomach, was on all fours in a downward dog position, or was crawling.
-When an infant was sitting with their butt on a surface, such as on the floor or a caregiver’s lap, the time frame was coded sitting on a surface by researchers.
-Sitting was also coded when the infant was in a kneeling-sit position, in which their knees were on the ground with their legging tucked underneath.
-When an infant was in a seating device that restricted movement, such as a high chair, sitting restrained was coded.
-Upright was coded when the infant was standing on the ground with two feet, walking, or kneeling with their thighs and torso in a straight line.
+We scored sitting to include any form of the following seated positions: 1) infants sat with their buttocks on a surface, such as on the floor or a caregiver’s lap, 2) infant was in a kneeling-sit position, in which their knees were on the ground with their legs tucked underneath the buttocks, and 3) infant was in a seating device, such as a high chair, that kept the torso oriented perpendicular to the ground (a reclined position, such as in a young infant's car seat, would be counted as supine).
+Upright was coded when the infant was standing on the ground with two feet or walking (regardless of whether infants' balance was assisted by a caregiver or with their hands holding onto something for support).
 When an infant was carried by a caregiver, held was coded.
 However, when the caregiver was sitting with the infant in their lap the infant’s body position was coded as if the caregiver was a surface (e.g., if the infant was sitting on the caregiver’s lap this was coded as sitting).
-The positions that did not fit in one of the six categories or when the infant was out of view were excluded.
-Periods when the leggings were being adjusted or taken off the infant were also excluded.
+Times during the video when the infant was out of view were excluded.
+Periods when the leggings were being adjusted or taken off the infant were also excluded, as were transitions between body positions.
 
-Primary coding was completed for the full length of the video, while reliability coding was completed for the first thirty minutes.
-Interrater reliability between coders … (overall agreement = XX.X%, kappa = X.XX).
+A primary coder completed annotation for the full length of the video, while an independent reliability coder completed annotation for the first thirty minutes of each video.
+Interrater reliability was based on the proportion of video frames that the two coders chose the same body position code.
+Overall agreement averaged XX.X% across video files, ranging from XX-XX.
+Cohen's *kappa* ranged from XX-XX across files, with a mean of XX.
 
 ## Body position classification
+
+## Data Sharing and Transparency
+
+<!-- WHAT CAN WE SHARE AND HOW -->
 

### 2023-03-07 manuscript.tex
knit

@@
 Synchronizing real-time data to other data streams helps to reveal sources of variability within an individual (e.g., Wass et al., 2022).
-Full-day recordings are essential for capturing experiences across the heterogeniety of daily routines that moderate behavior (e.g., play, feeding, errands) (Kadooka et al., 2021, April).
+Full-day recordings are essential for capturing experiences across the heterogeniety of daily routines that moderate behavior (e.g., play, feeding, errands) (Kadooka et al., 2021, April; Tamis-LeMonda et al., 2018).
 ``Burstiness'' of behavior means that long recordings are needed to capture clusters of events amid long periods in which they may be absent (Barbaro \& Fausey, 2022; Warlaumont et al., 2021).
@@
 
-\caption{Example Timeline Caption.}\label{fig:exemplartimeline}
+\caption{Example Timeline Caption.}\label{fig:exemplar-timeline}
 \end{figure}
 
+Figure \ref{fig:exemplar-timeline}) shows an exemplar timeline of the entire procedure and recording periods for a single participant.
+On the day of the visit, a researcher arrived at the participant's home in the morning between and prepared all the equipment at the doorstep.
+In order to synchronize all three recording devices, the researcher began by first turning on the video camera and the LENA audio device (the IMUs were already configured and placed at arrival because they required a proprietary sensor dock to begin recording in the laboratory).
+To create an easily recognizable synchronization point between the video recording and IMU data, the researcher dropped or struck the leggings (containing the IMUs) on a surface in view of the camera, as in (Franchak et al., 2021).
+All the equipment---once recording and with synchronization information recorded---was placed inside a large bucket and left outside the family's front door.
+
+The researcher then called the caregiver on the phone and walked them through a set of procedures needed to properly set up the equipment and record video for classifier training and testing.
+At the start of this ``guided call'', the caregiver was instructed to place the camera in an area that captured the majority of the room.
+Next, they were asked to put the pair of leggings and shirt on their infant, with the researcher providing guidance about how to correctly orient the garments.
+
+Afterwards, the researcher asked the caregiver to complete a number of guided activities with their infant.
+Within view of the camera, the caregiver was asked to place their infant in several different positions: lying supine, lying prone, sitting on the floor, standing upright, held by the caregiver while the caregiver walked back and forth, crawling, walking, and sitting in a restrained seat (e.g., high chair).
+Depending on the infants' age and motor skill level, the positions could be done independently or were completed with assistance from the caregiver.
+The researcher kept time to ensure at least 1 minute of behavior for each activity.
+Once completed, the caregiver was then instructed to play with their infant for 10 minutes within view of the camera to collect additional training data with the infant in positions that would be typical of play.
+
+Afterwards, they were to go about their day as usual with the infant wearing the sensor garment until their bedtime, only taking off the sensor for naps, baths, and diaper changes.
+The caregiver logged the times the sensors were removed (blank areas in the timeline in Figure \ref{fig:exemplar-timeline})) or the child took a nap (gray areas in the timeline in Figure \ref{fig:exemplar-timeline})) so that those times could be excluded from analysis.
+The following day a researcher picked up the equipment, verified the paperwork was signed, and compensated the participant.
+
+Because the camera only had the battery life to record for \textasciitilde3 hours (divided into two 90-minute video files), this divided the day into different periods for analysis.
+As seen in the bottom of Figure \ref{fig:exemplar-timeline}), the \emph{video period} comprised the first three hours of recording starting from the researcher's arrival when they turned on the camera.
+The first 90-minute video file, termed the \emph{proximal comparison}, contained the activities during the guided call followed by a period of infants and caregivers resuming their normal activities.
+Because this video contained the synchronization point, the data in this period had temporal synchrony between IMU and video data that contained errors of no more than 30-60 ms (1-2 video frames).
+When the first video file ended, a second video file was recorded, termed the \emph{distal comparison}.
+This video recorded the next 90 minutes of natural activity.
+However, because there was a variable gap of \textasciitilde40 s between the two videos, temporal synchrony in the distal comparison video was approximate containing offsets of \textasciitilde{} 5-10 s.
+
 \hypertarget{body-position-annotation}{%
-\subsection{Body position annotation}\label{body-position-annotation}}
+\subsection{Body Position Annotation}\label{body-position-annotation}}
+
+The proximal and distal comparison videos for each participant were annotated by trained human coders to identify infant \emph{body position} into one of 5 mutually-exclusive categories: supine, prone, sitting, upright, or held by caregiver.
+All coding was done using Datavyu software (datavyu.org).
+
+Supine was coded when the infant was lying on their back or was reclined up to a 45 degrees angle.
+Supine was also coded in the rare cases when the infant was laying on their side.
+Prone was coded when the infant was lying on their stomach, was on all fours in a downward dog position, or was crawling.
+We scored sitting to include any form of the following seated positions: 1) infants sat with their buttocks on a surface, such as on the floor or a caregiver's lap, 2) infant was in a kneeling-sit position, in which their knees were on the ground with their legs tucked underneath the buttocks, and 3) infant was in a seating device, such as a high chair, that kept the torso oriented perpendicular to the ground (a reclined position, such as in a young infant's car seat, would be counted as supine).
+Upright was coded when the infant was standing on the ground with two feet or walking (regardless of whether infants' balance was assisted by a caregiver or with their hands holding onto something for support).
+When an infant was carried by a caregiver, held was coded.
+However, when the caregiver was sitting with the infant in their lap the infant's body position was coded as if the caregiver was a surface (e.g., if the infant was sitting on the caregiver's lap this was coded as sitting).
+Times during the video when the infant was out of view were excluded.
+Periods when the leggings were being adjusted or taken off the infant were also excluded, as were transitions between body positions.
+
+A primary coder completed annotation for the full length of the video, while an independent reliability coder completed annotation for the first thirty minutes of each video.
+Interrater reliability was based on the proportion of video frames that the two coders chose the same body position code.
+Overall agreement averaged XX.X\% across video files, ranging from XX-XX.
+Cohen's \emph{kappa} ranged from XX-XX across files, with a mean of XX.
 
@@
 \subsection{Body position classification}\label{body-position-classification}}
+
+\hypertarget{data-sharing-and-transparency}{%
+\subsection{Data Sharing and Transparency}\label{data-sharing-and-transparency}}
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-Tamis-LeMondaCustode2018}{}}%
+Tamis-LeMonda, C.
+S., Custode, S., Kuchirko, Y., Escobar, K., \& Lo, T.
+(2018).
+Routine language: Speech directed to infants during home activities.
+\emph{Child Development}, \emph{90}(6), 2135--2152.
+\url{https://doi.org/10.1111/cdev.13089}
+
 \leavevmode\vadjust pre{\hypertarget{ref-Thurman2017}{}}%

### 2023-03-07 manuscript.Rmd
knit
cache results

@@
 
-```{r participants}
+```{r participants, cache=TRUE, cache.extra=tools::md5sum("data/imu-compiled.csv")}
 ppts <- read_csv("data/imu-compiled.csv") %>% 
@@
 Overall agreement averaged XX.X% across video files, ranging from XX-XX.
-Cohen's *kappa* ranged from XX-XX across files, with a mean of XX.
+Cohen's *kappa* averaged XX.X% across video files, ranging from XX-XX.
 
-## Body position classification
+## Body Position Classification
 
@@
 
-```{r metrics, warning=FALSE, fig.cap="Metrics", fig.align="center", fig.width=5, fig.height=8}
+```{r metrics, warning=FALSE, fig.cap="Metrics", fig.align="center", fig.width=5, fig.height=8, cache=TRUE, cache.extra=tools::md5sum("data/group_split_metrics.csv")}
 metrics <- read_csv("data/group_split_metrics.csv") %>% 
@@
 
-```{r metricstable}
+```{r metricstable, cache=TRUE, cache.extra=tools::md5sum("data/group_split_metrics.csv")}
 
@@
 
-```{r metricsbyclass}
+```{r metricsbyclass, cache=TRUE, cache.extra=tools::md5sum("data/group_split_metrics_class.csv")}
 metrics_class <- read_csv("data/group_split_metrics_class.csv") %>% 
@@
 Plots are shown separately for each body position with a reference line that indicates perfect agreement; each point in a plot represent data for a single participant.
-The three outlier participants are plotted in dark gray, with a different shape marking each individual.", fig.align="center", fig.width=7, fig.height=6.5}
+The three outlier participants are plotted in dark gray, with a different shape marking each individual.", fig.align="center", fig.width=7, fig.height=6.5, cache=TRUE, cache.extra=tools::md5sum("data/compiled_agreement_position.csv"), cache.extra=tools::md5sum("data/compiled_agreement_split.csv")}
 
@@
 
-```{r, pt2overalltable}
+```{r, pt2overalltable, cache=TRUE, cache.extra=tools::md5sum("data/compiled_agreement_position.csv"), cache.extra=tools::md5sum("data/compiled_agreement_split.csv")}
 #By posture
@@
 Performance is plotted separately for (A) group models and (B) individual models.
-The three outlier participants are plotted in dark gray, with a different shape marking each individual.", fig.align="center", fig.width=8, fig.height=4.5}
+The three outlier participants are plotted in dark gray, with a different shape marking each individual.", fig.align="center", fig.width=8, fig.height=4.5, cache=TRUE, cache.extra=tools::md5sum("data/compiled_agreement_position.csv"), cache.extra=tools::md5sum("data/compiled_agreement_split.csv")}
 calculate_agreement_bins <- function(ds) {
@@
 
-```{r pt2binstable}
+```{r pt2binstable, cache=TRUE, cache.extra=tools::md5sum("data/compiled_agreement_position.csv"), cache.extra=tools::md5sum("data/compiled_agreement_split.csv")}
 #Bins
@@
 
-```{r timelines, warning=FALSE, fig.cap="Timelines", fig.align="center", fig.width=7.75, fig.height=8}
+```{r timelines, warning=FALSE, fig.cap="Timelines", fig.align="center", fig.width=7.75, fig.height=8, cache=TRUE, cache.extra=tools::md5sum("data/imu-compiled.csv")}
 timeline_data <- read_csv("data/imu-compiled.csv") %>% 
@@
 
-```{r age, warning=FALSE, fig.cap="Age trends", fig.align="center", fig.width=7.5, fig.height=7}
+```{r age, warning=FALSE, fig.cap="Age trends", fig.align="center", fig.width=7.5, fig.height=7, cache=TRUE, cache.extra=tools::md5sum("data/imu-compiled.csv")}
 fullday_group <- read_csv("data/imu-compiled.csv") %>% 
@@
 
-```{r agetable}
+```{r agetable, cache=TRUE, cache.extra=tools::md5sum("data/imu-compiled.csv")}
 pos_sum_age <- map(fullday, fullday_long) %>% 

### 2023-03-07 manuscript.tex
cache results

@@
 Overall agreement averaged XX.X\% across video files, ranging from XX-XX.
-Cohen's \emph{kappa} ranged from XX-XX across files, with a mean of XX.
+Cohen's \emph{kappa} averaged XX.X\% across video files, ranging from XX-XX.
 
 \hypertarget{body-position-classification}{%
-\subsection{Body position classification}\label{body-position-classification}}
+\subsection{Body Position Classification}\label{body-position-classification}}
 

### 2023-03-07 manuscript.Rmd
classifiers

@@
 In order to synchronize all three recording devices, the researcher began by first turning on the video camera and the LENA audio device (the IMUs were already configured and placed at arrival because they required a proprietary sensor dock to begin recording in the laboratory).
-To create an easily recognizable synchronization point between the video recording and IMU data, the researcher dropped or struck the leggings (containing the IMUs) on a surface in view of the camera, as in [@FranchakScott2021].
+To create an easily recognizable synchronization point between the video recording and IMU data, the researcher dropped or struck the leggings (containing the IMUs) on a surface in view of the camera, as in @FranchakScott2021.
 All the equipment---once recording and with synchronization information recorded---was placed inside a large bucket and left outside the family's front door.
@@
 ## Body Position Classification
+
+<!-- R packages? Shared code -->
+The same machine learning classification process was used as in prior work [@FranchakScott2021].
+Using the synchronization point, human-coded body annotations from video were linked to the corresponding times in the IMU timeseries data.
+A single, merged dataset was created with synchronized accelerometer signals (in three orientations: X, Y, and Z) and gyroscope signals (in three orientations: roll, pitch, and yaw) for each of the four sensors (left hip, right hip, left ankle, right ankle) with the corresponding timestamp and body position code.
+
+Classification training and prediction was conducted on a windowed dataset that summarized the raw, 62.5 Hz motion signals within 4-s windows.
+Overlapping moving windows captured 4-s (250 samples) of data starting each second, which is a common unit of analysis in prior studies of human activity classification [@FranchakScott2021; @NamPark2013, @AiraksinenRasanen2020].
+The 250 samples of data in each window were aggregated into a variety of motion features---summary statistics that could be fed into the machine learning model.
+The minimum, maximum, 25th percentile, 75th percentile, mean, median, skew, kurtosis, standard deviation, and sum were computed for each signal (e.g., right hip X linear acceleration, left ankle pitch angular acceleration).
+The 10 summary statistics and 24 sensor signals generated 240 columns of motion features that described movement within each window.
+Furthermore, a series of cross-sensor and cross-orientation summaries (such as the correlation, magnitude, and difference between pairs of sensors) added an additional 196 columns of motion features.
+The 436 total motion features corresponded to a single body annotation code for each 4-s window.
+Windows were only used for training/testing if they contained a single body position for > 75% (3 s) of time within the window to ensure that motion signals could be tied to a clear example of each behavior.
 

### 2023-03-07 manuscript.Rmd
classification description

@@
 
-## Data Sharing and Transparency
+Windowed datasets were used for machine learning classification and validation.
+For each analysis reported in the results, a subset of data were defined as a "training" set and another, independent portion of the data were defined as a "testing" set.
+Different training/testing set combinations were used to address different goals; here we describe the general procedure of how a classifier was derived from a training set.
+Random forest models [@Breiman2001] were trained to predict the human coded body position label for each window from the set of 436 motion features using the *randomForest* package in R [@LiawWiener2002].
+Random forest models use random subsets of features and random subsets of data across many iterations; this random subsetting creates models that are less likely to be overfit.
+The *tidymodels* set of packages in R was used to tune the "mtry" and "ntrees" hyperparameters, which control the number of columns used and the total number of trees created when modeling.
+We use an mtry parameter of 44 and and ntree parameter of 550, however, changing hyperparameters had little impact on model performance.
+Just like a linear regression model, the resulting random forest model could later be applied to a set of testing data with the *predict* function.
+Validation analyses below will compare the human-coded annotations and random forest model predictions for testing data that are independent from training data, providing a test of whether models can successfully generalize.
+<!-- cite tidymodels -->
 
+<!-- ## Data Sharing and Transparency -->
 <!-- WHAT CAN WE SHARE AND HOW -->

### 2023-03-07 manuscript.tex
reliability numbers

@@
 In order to synchronize all three recording devices, the researcher began by first turning on the video camera and the LENA audio device (the IMUs were already configured and placed at arrival because they required a proprietary sensor dock to begin recording in the laboratory).
-To create an easily recognizable synchronization point between the video recording and IMU data, the researcher dropped or struck the leggings (containing the IMUs) on a surface in view of the camera, as in (Franchak et al., 2021).
+To create an easily recognizable synchronization point between the video recording and IMU data, the researcher dropped or struck the leggings (containing the IMUs) on a surface in view of the camera, as in Franchak et al.
+(2021).
 All the equipment---once recording and with synchronization information recorded---was placed inside a large bucket and left outside the family's front door.
@@
 Interrater reliability was based on the proportion of video frames that the two coders chose the same body position code.
-Overall agreement averaged XX.X\% across video files, ranging from XX-XX.
-Cohen's \emph{kappa} averaged XX.X\% across video files, ranging from XX-XX.
+Overall agreement averaged 90.5\% across video files, ranging from 68.4\%-100\%.
+Cohen's \emph{kappa} averaged 85.5\% across video files, ranging from 31.0\%-100\%.
 
@@
 
-\hypertarget{data-sharing-and-transparency}{%
-\subsection{Data Sharing and Transparency}\label{data-sharing-and-transparency}}
+The same machine learning classification process was used as in prior work (Franchak et al., 2021).
+Using the synchronization point, human-coded body annotations from video were linked to the corresponding times in the IMU timeseries data.
+A single, merged dataset was created with synchronized accelerometer signals (in three orientations: X, Y, and Z) and gyroscope signals (in three orientations: roll, pitch, and yaw) for each of the four sensors (left hip, right hip, left ankle, right ankle) with the corresponding timestamp and body position code.
+
+Classification training and prediction was conducted on a windowed dataset that summarized the raw, 62.5 Hz motion signals within 4-s windows.
+Overlapping moving windows captured 4-s (250 samples) of data starting each second, which is a common unit of analysis in prior studies of human activity classification Airaksinen et al.
+(2020).
+The 250 samples of data in each window were aggregated into a variety of motion features---summary statistics that could be fed into the machine learning model.
+The minimum, maximum, 25th percentile, 75th percentile, mean, median, skew, kurtosis, standard deviation, and sum were computed for each signal (e.g., right hip X linear acceleration, left ankle pitch angular acceleration).
+The 10 summary statistics and 24 sensor signals generated 240 columns of motion features that described movement within each window.
+Furthermore, a series of cross-sensor and cross-orientation summaries (such as the correlation, magnitude, and difference between pairs of sensors) added an additional 196 columns of motion features.
+The 436 total motion features corresponded to a single body annotation code for each 4-s window.
+Windows were only used for training/testing if they contained a single body position for \textgreater{} 75\% (3 s) of time within the window to ensure that motion signals could be tied to a clear example of each behavior.
+
+Windowed datasets were used for machine learning classification and validation.
+For each analysis reported in the results, a subset of data were defined as a ``training'' set and another, independent portion of the data were defined as a ``testing'' set.
+Different training/testing set combinations were used to address different goals; here we describe the general procedure of how a classifier was derived from a training set.
+Random forest models (Breiman, 2001) were trained to predict the human coded body position label for each window from the set of 436 motion features using the \emph{randomForest} package in R (Liaw et al., 2002).
+Random forest models use random subsets of features and random subsets of data across many iterations; this random subsetting creates models that are less likely to be overfit.
+The \emph{tidymodels} set of packages in R was used to tune the ``mtry'' and ``ntrees'' hyperparameters, which control the number of columns used and the total number of trees created when modeling.
+We use an mtry parameter of 44 and and ntree parameter of 550, however, changing hyperparameters had little impact on model performance.
+Just like a linear regression model, the resulting random forest model could later be applied to a set of testing data with the \emph{predict} function.
+Validation analyses below will compare the human-coded annotations and random forest model predictions for testing data that are independent from training data, providing a test of whether models can successfully generalize.
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-Breiman2001}{}}%
+Breiman, L. (2001). Random forests. \emph{Machine Learning}, \emph{45}, 5--32.
+
 \leavevmode\vadjust pre{\hypertarget{ref-BruijnsTruelove2020}{}}%
@@
 \emph{International Journal of Environmental Research and Public Health}, \emph{16}(14), 2598.
+
+\leavevmode\vadjust pre{\hypertarget{ref-LiawWiener2002}{}}%
+Liaw, A., Wiener, M., et al. (2002). Classification and regression by randomForest. \emph{R News}, \emph{2}(3), 18--22.
 

### 2023-03-07 manuscript.Rmd
reliability numbers
updated refs

@@
 Interrater reliability was based on the proportion of video frames that the two coders chose the same body position code.
-Overall agreement averaged XX.X% across video files, ranging from XX-XX.
-Cohen's *kappa* averaged XX.X% across video files, ranging from XX-XX.
+Overall agreement averaged 90.5% across video files, ranging from 68.4%-100%.
+Cohen's *kappa* averaged 85.5% across video files, ranging from 31.0%-100%.
+<!-- UPDATE THESE BEFORE SUBMITTING -->
 
@@
 Classification training and prediction was conducted on a windowed dataset that summarized the raw, 62.5 Hz motion signals within 4-s windows.
-Overlapping moving windows captured 4-s (250 samples) of data starting each second, which is a common unit of analysis in prior studies of human activity classification [@FranchakScott2021; @NamPark2013, @AiraksinenRasanen2020].
+Overlapping moving windows captured 4-s (250 samples) of data starting each second, which is a common unit of analysis in prior studies of human activity classification [@FranchakScott2021; @NamPark2013; @AiraksinenRasanen2020].
 The 250 samples of data in each window were aggregated into a variety of motion features---summary statistics that could be fed into the machine learning model.
@@
 Random forest models use random subsets of features and random subsets of data across many iterations; this random subsetting creates models that are less likely to be overfit.
-The *tidymodels* set of packages in R was used to tune the "mtry" and "ntrees" hyperparameters, which control the number of columns used and the total number of trees created when modeling.
+The *tidymodels* set of packages in R [@tidymodels] was used to tune the "mtry" and "ntrees" hyperparameters, which control the number of columns sampled and the total number of trees created when modeling.
 We use an mtry parameter of 44 and and ntree parameter of 550, however, changing hyperparameters had little impact on model performance.
@@
 Validation analyses below will compare the human-coded annotations and random forest model predictions for testing data that are independent from training data, providing a test of whether models can successfully generalize.
-<!-- cite tidymodels -->
 

### 2023-03-07 manuscript.tex
updated refs

@@
 Classification training and prediction was conducted on a windowed dataset that summarized the raw, 62.5 Hz motion signals within 4-s windows.
-Overlapping moving windows captured 4-s (250 samples) of data starting each second, which is a common unit of analysis in prior studies of human activity classification Airaksinen et al.
-(2020).
+Overlapping moving windows captured 4-s (250 samples) of data starting each second, which is a common unit of analysis in prior studies of human activity classification (Airaksinen et al., 2020; Franchak et al., 2021; Nam \& Park, 2013).
 The 250 samples of data in each window were aggregated into a variety of motion features---summary statistics that could be fed into the machine learning model.
@@
 Random forest models use random subsets of features and random subsets of data across many iterations; this random subsetting creates models that are less likely to be overfit.
-The \emph{tidymodels} set of packages in R was used to tune the ``mtry'' and ``ntrees'' hyperparameters, which control the number of columns used and the total number of trees created when modeling.
+The \emph{tidymodels} set of packages in R (Kuhn \& Wickham, 2020) was used to tune the ``mtry'' and ``ntrees'' hyperparameters, which control the number of columns sampled and the total number of trees created when modeling.
 We use an mtry parameter of 44 and and ntree parameter of 550, however, changing hyperparameters had little impact on model performance.
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-tidymodels}{}}%
+Kuhn, M., \& Wickham, H. (2020). Tidymodels: A collection of packages for modeling and machine learning using tidyverse principles. \emph{Boston, MA, USA}.
+
 \leavevmode\vadjust pre{\hypertarget{ref-KwonZavos2019}{}}%

### 2023-03-08 manuscript.Rmd
duration in hours

@@
 library(tidyverse)
+library(lubridate)
 library(rstatix)
@@
 Although in the current investigation we focus on classification of body position, this approach can be extended to categorize other aspects of movement, such as locomotion [@AiraksinenRasanen2020] and infant-caregiver contact [@YaoPlotz2019].
-Accompanying our paper, we have created a database [CITE] containing synchronized video and IMU data, allowing researchers to annotate and train models for behaviors beyond body position.
+<!-- Accompanying our paper, we have created a database [CITE] containing synchronized video and IMU data, allowing researchers to annotate and train models for behaviors beyond body position.  -->
 
@@
 The **first goal** of the current study was to determine the accuracy of body position classification during the proximal comparison period using this novel, semi-supervised procedure in participants' homes.
+Past work found varying better performance using "individual models"---models that were trained on one participant's data to predict their later behavior---compared with a "group model" that aggregated data from all infants to create a single body position classifier [@FranchakScott2021], so we compared both modeling approaches in the current investigation.
 
@@
 The proprietary "Boosted Battery Base" attached to the action camera to allow for a total of 3 hours of recording.
-However, as described below, this created two video files separated by a variable gap of approximately 40 s.
+However, as described below, this created two video files temporally separated by a gap of approximately 40-45 s.
 Caregivers also received a log sheet to record times that infants napped as well as times that the sensor garment was removed from the infant (e.g., baths, diaper changes).
@@
 
-<!-- AVERAGE START TIME -->
 Figure \@ref(fig:exemplar-timeline)) shows an exemplar timeline of the entire procedure and recording periods for a single participant.
@@
 
-## Goal 1: Assess the proximal accuracy of body position classification models
+```{r total-duration}
+duration <- ppts %>% group_by(id_uni) %>% 
+  summarize(start_time = min(clock_time_start), end_time= max(clock_time_end)) %>% 
+  mutate(duration = as.duration(end_time - start_time), duration_hours = as.numeric(duration, "hours")) %>% arrange(duration)
+```
+
+We report four sets of results based on `r length(unique(ppts$id_uni))` full-day testing sessions resulting in a total of `r sum(duration$duration_hours)` hours of movement recording.
+
+## Assess the Proximal Accuracy of Body Position Classification Models
 
@@
 
+```{r start-time}
+start_times <- ppts %>% group_by(id_uni) %>% slice_min(clock_time_start) %>% pull(clock_time_start) %>% sort
+start_times <- str_sub(as.character(start_times), start = 1, end = -4)
+```
+The start time ranged from `r start_times[1]` to `r start_times[length(start_times)]` with a median of `r start_times[floor(length(start_times)/2)]`.
+Although the equipment was always dropped off in the morning,
+
 ## Goal 4: Assess the suitability of full-day predictions for capturing age differences in body position

### 2023-03-08 manuscript.Rmd
got intro to results

@@
 We report four sets of results based on `r length(unique(ppts$id_uni))` full-day testing sessions resulting in a total of `r sum(duration$duration_hours)` hours of movement recording.
+First, we focus on the accuracy of group and individual models trained and validated on the proximal comparison period.
+This period is unique in having the highest degree of temporal synchrony between video and motion data, allowing us to assess the accuracy of individual body position events.
+Second, we applied models trained from data during the proximal comparison to predict infant behavior during the distal comparison, when coarser synchrony between video and motion data was available.
+This distal comparison is crucial, because it provides the first ever test of how well body classification models predict behavior beyond the initial, supervised period.
+Data from the first two sets of analyses confirm that body position classification models are accurate across a long period of time, suggesting that data from the entire day should be valid.
+In the third set of results we examine the data quality of full-day recordings, and in the fourth set of results we test whether full-day recordings capture well-established age differences in body position experiences.
 

### 2023-03-08 manuscript.Rmd
proximal metrics

@@
 
-```{r metricstable, cache=TRUE, cache.extra=tools::md5sum("data/group_split_metrics.csv")}
+```{r metric-stable, cache=TRUE, cache.extra=tools::md5sum("data/group_split_metrics.csv")}
 
-metrics_long %>% group_by(Model, metric) %>% 
+metrics_summary <- metrics_long %>% group_by(Model, metric) %>% 
   get_summary_stats(value) %>% 
@@
          metric = factor(metric, levels = metric_labels)) %>% 
-  arrange(metric) %>% 
-  apa_table(., col.names = c("Metric", "Median", "Mean", "SD", "Median", "Mean", "SD"),    col_spanners = list("Group" = c(2, 4), "Individual" = c(5, 7)),
+  arrange(metric)
+
+  apa_table(metrics_summary, col.names = c("Metric", "Median", "Mean", "SD", "Median", "Mean", "SD"),    col_spanners = list("Group" = c(2, 4), "Individual" = c(5, 7)),
           caption = "Summary statistics for model performance metrics shown separately for group and individual models.")
+
+metrics_lookup <- metrics_long %>% group_by(Model, metric) %>% 
+  get_summary_stats(value) %>% 
+  select(-(variable:n), -(q1:mad), -se, -ci) %>%
+  pivot_wider(names_from = Model, values_from = c("min","max","median", "mean", "sd"))
+    column_to_rownames("metric")
 ```
@@
 ```
+
+The first set of analyses use data from the proximal comparison, in which video annotations and motion data were tightly synchronized (within 30-60 ms).
+This high degree of synchronization makes it possible to link human-coded body position annotations to each 4-s window of motion data, providing ground truth data for model training and testing.
+As in past work [@FranchakScott2021], we compared two types of models.
+To assess the accuracy of body position classification for each recording session, we held back the last 25% of a session's proximal comparison data as the testing set.
+The testing set was never used as training data.
+A *group model* was trained for each session using an aggregated dataset containing the first 75% of all **other** sessions' proximal comparison data.
+This leave-one-out cross-validation tested the generalization of the model to a recording session that was not used at all in training.
+In contrast, *individual models* were trained using the first 75% of a recording session, then tested against the last 25%.
+The individual model tests whether earlier training data generalize to later testing data within an individual participant's recording.
+In both cases, the testing set was the same, allowing for direct comparison between the two modeling approaches.
+
+Figure \@ref(fig:metrics)) and Table \@ref(tab:metrics-table)) summarize the performance of group and individual models using standard metrics for classification.
 

### 2023-03-08 manuscript.Rmd
kappa

@@
 
-```{r metric-stable, cache=TRUE, cache.extra=tools::md5sum("data/group_split_metrics.csv")}
+```{r metrics-table, cache=TRUE, cache.extra=tools::md5sum("data/group_split_metrics.csv")}
 
@@
   select(-(variable:n), -(q1:mad), -se, -ci) %>%
-  pivot_wider(names_from = Model, values_from = c("min","max","median", "mean", "sd"))
+  pivot_wider(names_from = Model, values_from = c("min","max","median", "mean", "sd")) %>% 
     column_to_rownames("metric")
@@
 
-```{r metricsbyclass, cache=TRUE, cache.extra=tools::md5sum("data/group_split_metrics_class.csv")}
+```{r metrics-by-class, cache=TRUE, cache.extra=tools::md5sum("data/group_split_metrics_class.csv")}
 metrics_class <- read_csv("data/group_split_metrics_class.csv") %>% 
@@
 
-Figure \@ref(fig:metrics)) and Table \@ref(tab:metrics-table)) summarize the performance of group and individual models using standard metrics for classification.
+<!-- GET RID OF BALANCED ACCURACY AND F1, MOVE KAPPA UP, THEN DO SENS/PPV -->
+Figure \@ref(fig:metrics) and Table \@ref(tab:metrics-table) summarize the performance of group and individual models using standard metrics for classification.
+Overall accuracy (Figure \@ref(fig:metrics)A) represents the proportion of 4-s windows in the testing set in which the model prediction matched the human annotation of body position.
+Median overall accuracy was `r metrics_lookup["Overall Accuracy", "median_Group"]` for group models and `r metrics_lookup["Overall Accuracy", "median_Individual"]` for individual models.
+Although individual model accuracy was greater than group model accuracy, group model accuracy approached the level of agreement found between two human coders (median = .92) indicating that group models could adequately replace human coders.
+Still, it is notable that the worst-case accuracy from group models (minimum = `r metrics_lookup["Overall Accuracy", "min_Group"]`) was considerably lower than in individual models (minimum = `r metrics_lookup["Overall Accuracy", "min_Individual"]`).
+Possibly, individual models accounted for idiosyncrasies in behavior or inconsistencies in sensor placement for those sessions that group models could not generalize to.
+
+Strong overall accuracy can be misleading when class prevalence is unbalanced.
+For example, most infants spent longer sitting than prone, so good performance for sitting classification could mask poorer performance in prone classification.
+Accordingly, we report Cohen's kappa, a commonly-used metric that penalizes missing rare events (Figure \@ref(fig:metrics)), and provide classification metrics for each individual body position (Table \@ref(tab:metrics-by-class)) to account for imbalance in body position rates within and between individuals.
+Similar to overall accuracy, kappa values were strong for both model types with median group kappas (`r metrics_lookup["Kappa", "median_Group"]`) somewhat worse compared with median individual kappas (`r metrics_lookup["Kappa", "median_Individual"]`).
+Guidelines for interpreting kappa statistics [@LandisKoch1977] consider 0.81–1.00 "Almost Perfect,” 0.61–0.80 “Substantial,” 0.41–0.60 “Moderate,” 0.21– 0.40 “Fair,” and 0–0.20 “Slight to Poor”, indicating that agreement for most group and individual model predictions fell in the Substantial to Almost Perfect range.
 

### 2023-03-08 manuscript.tex
kappa

@@
 Although in the current investigation we focus on classification of body position, this approach can be extended to categorize other aspects of movement, such as locomotion (Airaksinen et al., 2020) and infant-caregiver contact (Yao et al., 2019).
-Accompanying our paper, we have created a database {[}CITE{]} containing synchronized video and IMU data, allowing researchers to annotate and train models for behaviors beyond body position.
 
@@
 The \textbf{first goal} of the current study was to determine the accuracy of body position classification during the proximal comparison period using this novel, semi-supervised procedure in participants' homes.
+Past work found varying better performance using ``individual models''---models that were trained on one participant's data to predict their later behavior---compared with a ``group model'' that aggregated data from all infants to create a single body position classifier (Franchak et al., 2021), so we compared both modeling approaches in the current investigation.
 
@@
 The proprietary ``Boosted Battery Base'' attached to the action camera to allow for a total of 3 hours of recording.
-However, as described below, this created two video files separated by a variable gap of approximately 40 s.
+However, as described below, this created two video files temporally separated by a gap of approximately 40-45 s.
 Caregivers also received a log sheet to record times that infants napped as well as times that the sensor garment was removed from the infant (e.g., baths, diaper changes).
@@
 
-\hypertarget{goal-1-assess-the-proximal-accuracy-of-body-position-classification-models}{%
-\subsection{Goal 1: Assess the proximal accuracy of body position classification models}\label{goal-1-assess-the-proximal-accuracy-of-body-position-classification-models}}
+We report four sets of results based on 34 full-day testing sessions resulting in a total of 301.74 hours of movement recording.
+First, we focus on the accuracy of group and individual models trained and validated on the proximal comparison period.
+This period is unique in having the highest degree of temporal synchrony between video and motion data, allowing us to assess the accuracy of individual body position events.
+Second, we applied models trained from data during the proximal comparison to predict infant behavior during the distal comparison, when coarser synchrony between video and motion data was available.
+This distal comparison is crucial, because it provides the first ever test of how well body classification models predict behavior beyond the initial, supervised period.
+Data from the first two sets of analyses confirm that body position classification models are accurate across a long period of time, suggesting that data from the entire day should be valid.
+In the third set of results we examine the data quality of full-day recordings, and in the fourth set of results we test whether full-day recordings capture well-established age differences in body position experiences.
+
+\hypertarget{assess-the-proximal-accuracy-of-body-position-classification-models}{%
+\subsection{Assess the Proximal Accuracy of Body Position Classification Models}\label{assess-the-proximal-accuracy-of-body-position-classification-models}}
 
@@
 
-\caption{\label{tab:metricstable}Summary statistics for model performance metrics shown separately for group and individual models.}
+\caption{\label{tab:metrics-table}Summary statistics for model performance metrics shown separately for group and individual models.}
 
@@
 
-\caption{\label{tab:metricsbyclass}Model performance metrics for each body position category, shown separately for group and individual models.}
+\caption{\label{tab:metrics-by-class}Model performance metrics for each body position category, shown separately for group and individual models.}
 
@@
 
+The first set of analyses use data from the proximal comparison, in which video annotations and motion data were tightly synchronized (within 30-60 ms).
+This high degree of synchronization makes it possible to link human-coded body position annotations to each 4-s window of motion data, providing ground truth data for model training and testing.
+As in past work (Franchak et al., 2021), we compared two types of models.
+To assess the accuracy of body position classification for each recording session, we held back the last 25\% of a session's proximal comparison data as the testing set.
+The testing set was never used as training data.
+A \emph{group model} was trained for each session using an aggregated dataset containing the first 75\% of all \textbf{other} sessions' proximal comparison data.
+This leave-one-out cross-validation tested the generalization of the model to a recording session that was not used at all in training.
+In contrast, \emph{individual models} were trained using the first 75\% of a recording session, then tested against the last 25\%.
+The individual model tests whether earlier training data generalize to later testing data within an individual participant's recording.
+In both cases, the testing set was the same, allowing for direct comparison between the two modeling approaches.
+
+Figure \ref{fig:metrics} and Table \ref{tab:metrics-table} summarize the performance of group and individual models using standard metrics for classification.
+Overall accuracy (Figure \ref{fig:metrics}A) represents the proportion of 4-s windows in the testing set in which the model prediction matched the human annotation of body position.
+Median overall accuracy was 0.89 for group models and 0.93 for individual models.
+Although individual model accuracy was greater than group model accuracy, group model accuracy approached the level of agreement found between two human coders (median = .92) indicating that group models could adequately replace human coders.
+Still, it is notable that the worst-case accuracy from group models (minimum = 0.50) was considerably lower than in individual models (minimum = 0.73).
+Possibly, individual models accounted for idiosyncrasies in behavior or inconsistencies in sensor placement for those sessions that group models could not generalize to.
+
+Strong overall accuracy can be misleading when class prevalence is unbalanced.
+For example, most infants spent longer sitting than prone, so good performance for sitting classification could mask poorer performance in prone classification.
+Accordingly, we report Cohen's kappa, a commonly-used metric that penalizes missing rare events (Figure \ref{fig:metrics}), and provide classification metrics for each individual body position (Table \ref{tab:metrics-by-class}) to account for imbalance in body position rates within and between individuals.
+Similar to overall accuracy, kappa values were strong for both model types with median group kappas (0.77) somewhat worse compared with median individual kappas (0.85).
+Guidelines for interpreting kappa statistics (Landis \& Koch, 1977) consider 0.81--1.00 ``Almost Perfect,'' 0.61--0.80 ``Substantial,'' 0.41--0.60 ``Moderate,'' 0.21-- 0.40 ``Fair,'' and 0--0.20 ``Slight to Poor'', indicating that agreement for most group and individual model predictions fell in the Substantial to Almost Perfect range.
+
 \hypertarget{goal-2-assess-the-distal-accuracy-of-body-position-classification-models}{%
@@
 
+The start time ranged from 08:55 to 13:20 with a median of 10:20. Although the equipment was always dropped off in the morning,
+
 \hypertarget{goal-4-assess-the-suitability-of-full-day-predictions-for-capturing-age-differences-in-body-position}{%
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-LandisKoch1977}{}}%
+Landis, J. R., \& Koch, G. G. (1977). The measurement of observer agreement for categorical data. \emph{Biometrics}, \emph{33}(1), 159. \url{https://doi.org/10.2307/2529310}
+
 \leavevmode\vadjust pre{\hypertarget{ref-LiawWiener2002}{}}%

### 2023-03-08 manuscript.Rmd
fewer metrics

@@
 
-```{r metrics, warning=FALSE, fig.cap="Metrics", fig.align="center", fig.width=5, fig.height=8, cache=TRUE, cache.extra=tools::md5sum("data/group_split_metrics.csv")}
+```{r metrics, warning=FALSE, fig.cap="Metrics", fig.align="center", fig.width=5.5, fig.height=5.5, cache=TRUE, cache.extra=tools::md5sum("data/group_split_metrics.csv")}
 metrics <- read_csv("data/group_split_metrics.csv") %>% 
@@
 }
-metric_labels <- c("Overall Accuracy", "Balanced Accuracy", "F1", "Sensitivity", "Pos Pred Value", "Kappa")
-title_labels <- c("A. Overall Accuracy", "B. Balanced Accuracy", "C. F1 Score", "D. Sensitivity", "E. Pos. Pred. Value", "F. Kappa")
+# metric_labels <- c("Overall Accuracy", "Balanced Accuracy", "F1", "Sensitivity", "Pos Pred Value", "Kappa")
+# title_labels <- c("A. Overall Accuracy", "B. Balanced Accuracy", "C. F1 Score", "D. Sensitivity", "E. Pos. Pred. Value", "F. Kappa")
+metric_labels <- c("Overall Accuracy", "Kappa", "Sensitivity", "Pos Pred Value")
+title_labels <- c("A. Overall Accuracy", "B. Kappa", "C. Sensitivity", "D. Pos. Pred. Value")
 metric_plots <- map(metric_labels, ~ metric_plot(metrics_long, .x)) %>% map2(., title_labels, ~.x + ggtitle(.y))
-(metric_plots[[1]] + metric_plots[[2]])/(metric_plots[[3]]+metric_plots[[4]])/(metric_plots[[5]] + metric_plots[[6]])
+(metric_plots[[1]] + metric_plots[[2]])/(metric_plots[[3]]+metric_plots[[4]])
+# (metric_plots[[1]] + metric_plots[[2]])/(metric_plots[[3]]+metric_plots[[4]])/(metric_plots[[5]] + metric_plots[[6]])
 
@@
 
-<!-- GET RID OF BALANCED ACCURACY AND F1, MOVE KAPPA UP, THEN DO SENS/PPV -->
 Figure \@ref(fig:metrics) and Table \@ref(tab:metrics-table) summarize the performance of group and individual models using standard metrics for classification.

### 2023-03-08 manuscript.Rmd
more kappa

@@
   pivot_wider(names_from = Model, values_from = c("median", "mean", "sd")) %>% 
-  filter(metric != "Neg Pred Value", metric != "Specificity") %>% 
+  filter(metric != "Neg Pred Value", metric != "Specificity", metric != "Balanced Accuracy", metric != "F1") %>% 
   relocate(mean_Group, .after = median_Group) %>% 
@@
 apa_table(class_table, col.names = c("Metric","Position", "Median", "Mean", "SD", "Median", "Mean", "SD"), col_spanners = list("Group" = c(3, 5), "Individual" = c(6, 8)),
-          midrules = c(5, 10, 15, 20),
+          midrules = c(5, 10),
           caption = "Model performance metrics for each body position category, shown separately for group and individual models.")
+metrics_long_summary <- metrics_long_summary %>% 
+  mutate(rowname = paste(Class, metric)) %>% 
+  column_to_rownames("rowname")
 ```
@@
 
+### Overall Accuracy
+
 Figure \@ref(fig:metrics) and Table \@ref(tab:metrics-table) summarize the performance of group and individual models using standard metrics for classification.
@@
 
+### Cohen's Kappa 
 Strong overall accuracy can be misleading when class prevalence is unbalanced.
 For example, most infants spent longer sitting than prone, so good performance for sitting classification could mask poorer performance in prone classification.
-Accordingly, we report Cohen's kappa, a commonly-used metric that penalizes missing rare events (Figure \@ref(fig:metrics)), and provide classification metrics for each individual body position (Table \@ref(tab:metrics-by-class)) to account for imbalance in body position rates within and between individuals.
+Accordingly, we report Cohen's kappa, a commonly-used metric that penalizes missing rare events (Figure \@ref(fig:metrics)B), and provide classification metrics for each individual body position (Table \@ref(tab:metrics-by-class)) to account for imbalance in body position rates within and between individuals.
 Similar to overall accuracy, kappa values were strong for both model types with median group kappas (`r metrics_lookup["Kappa", "median_Group"]`) somewhat worse compared with median individual kappas (`r metrics_lookup["Kappa", "median_Individual"]`).
 Guidelines for interpreting kappa statistics [@LandisKoch1977] consider 0.81–1.00 "Almost Perfect,” 0.61–0.80 “Substantial,” 0.41–0.60 “Moderate,” 0.21– 0.40 “Fair,” and 0–0.20 “Slight to Poor”, indicating that agreement for most group and individual model predictions fell in the Substantial to Almost Perfect range.
+
+As in past work, all body positions were accurately classified even though performance varied somewhat between positions.
+As Table \@ref(tab:metrics-by-class) shows, median kappa statistics were strongest for prone (group = `r metrics_long_summary["Prone Kappa", "median_Group"]`, individual = `r metrics_long_summary["Prone Kappa", "median_Individual"]`) and supine (group = `r metrics_long_summary["Supine Kappa", "median_Group"]`, individual = `r metrics_long_summary["Supine Kappa", "median_Individual"]`).
+Sitting performance fell in the middle, and was considerably worse for group models than individual models (group = `r metrics_long_summary["Sitting Kappa", "median_Group"]`, individual = `r metrics_long_summary["Sitting Kappa", "median_Individual"]`).
+Held (group = `r metrics_long_summary["Held Kappa", "median_Group"]`, individual = `r metrics_long_summary["Held Kappa", "median_Individual"]`) and upright (group = `r metrics_long_summary["Upright Kappa", "median_Group"]`, individual = `r metrics_long_summary["Upright Kappa", "median_Individual"]`) performance was the least accurate, however, median performance was still within the "Substantial" range.
+
+### Sensitivity and Positive Predictive Value
 

### 2023-03-08 manuscript.tex
more kappa

@@
 Overall Accuracy & 0.894 & 0.846 & 0.131 & 0.933 & 0.916 & 0.072\\
-Balanced Accuracy & 0.913 & 0.891 & 0.077 & 0.911 & 0.907 & 0.071\\
-F1 & 0.822 & 0.816 & 0.106 & 0.882 & 0.865 & 0.091\\
+Kappa & 0.768 & 0.746 & 0.161 & 0.849 & 0.821 & 0.143\\
 Sensitivity & 0.856 & 0.825 & 0.127 & 0.847 & 0.841 & 0.119\\
 Pos Pred Value & 0.826 & 0.810 & 0.125 & 0.928 & 0.899 & 0.107\\
-Kappa & 0.768 & 0.746 & 0.161 & 0.849 & 0.821 & 0.143\\
 \bottomrule
@@
 \midrule
-Balanced Accuracy & Supine & 0.964 & 0.925 & 0.101 & 0.994 & 0.959 & 0.086\\
- & Prone & 0.996 & 0.941 & 0.121 & 0.968 & 0.915 & 0.135\\
- & Sitting & 0.909 & 0.862 & 0.141 & 0.963 & 0.942 & 0.070\\
- & Upright & 0.907 & 0.849 & 0.155 & 0.938 & 0.874 & 0.123\\
- & Held & 0.901 & 0.871 & 0.110 & 0.868 & 0.844 & 0.154\\ \midrule
-F1 & Supine & 0.943 & 0.867 & 0.176 & 0.986 & 0.938 & 0.127\\
- & Prone & 0.979 & 0.897 & 0.190 & 0.957 & 0.880 & 0.191\\
- & Sitting & 0.879 & 0.803 & 0.217 & 0.951 & 0.924 & 0.099\\
- & Upright & 0.738 & 0.722 & 0.246 & 0.844 & 0.766 & 0.231\\
- & Held & 0.771 & 0.779 & 0.151 & 0.839 & 0.768 & 0.238\\ \midrule
+Kappa & Supine & 0.907 & 0.764 & 0.295 & 0.983 & 0.912 & 0.166\\
+ & Prone & 0.968 & 0.860 & 0.259 & 0.942 & 0.841 & 0.246\\
+ & Sitting & 0.816 & 0.702 & 0.297 & 0.915 & 0.887 & 0.127\\
+ & Upright & 0.707 & 0.673 & 0.281 & 0.822 & 0.741 & 0.236\\
+ & Held & 0.732 & 0.726 & 0.209 & 0.826 & 0.727 & 0.277\\ \midrule
 Sensitivity & Supine & 1.000 & 0.905 & 0.180 & 1.000 & 0.954 & 0.129\\
@@
  & Upright & 0.839 & 0.739 & 0.283 & 0.923 & 0.825 & 0.228\\
- & Held & 0.852 & 0.794 & 0.240 & 0.943 & 0.899 & 0.134\\ \midrule
-Kappa & Supine & 0.907 & 0.764 & 0.295 & 0.983 & 0.912 & 0.166\\
- & Prone & 0.968 & 0.860 & 0.259 & 0.942 & 0.841 & 0.246\\
- & Sitting & 0.816 & 0.702 & 0.297 & 0.915 & 0.887 & 0.127\\
- & Upright & 0.707 & 0.673 & 0.281 & 0.822 & 0.741 & 0.236\\
- & Held & 0.732 & 0.726 & 0.209 & 0.826 & 0.727 & 0.277\\
+ & Held & 0.852 & 0.794 & 0.240 & 0.943 & 0.899 & 0.134\\
 \bottomrule
@@
 For example, most infants spent longer sitting than prone, so good performance for sitting classification could mask poorer performance in prone classification.
-Accordingly, we report Cohen's kappa, a commonly-used metric that penalizes missing rare events (Figure \ref{fig:metrics}), and provide classification metrics for each individual body position (Table \ref{tab:metrics-by-class}) to account for imbalance in body position rates within and between individuals.
+Accordingly, we report Cohen's kappa, a commonly-used metric that penalizes missing rare events (Figure \ref{fig:metrics}B), and provide classification metrics for each individual body position (Table \ref{tab:metrics-by-class}) to account for imbalance in body position rates within and between individuals.
 Similar to overall accuracy, kappa values were strong for both model types with median group kappas (0.77) somewhat worse compared with median individual kappas (0.85).
 Guidelines for interpreting kappa statistics (Landis \& Koch, 1977) consider 0.81--1.00 ``Almost Perfect,'' 0.61--0.80 ``Substantial,'' 0.41--0.60 ``Moderate,'' 0.21-- 0.40 ``Fair,'' and 0--0.20 ``Slight to Poor'', indicating that agreement for most group and individual model predictions fell in the Substantial to Almost Perfect range.
+
+As in past work, all body positions were accurately classified even though performance varied somewhat between positions.
+As Table \ref{tab:metrics-by-class} shows, median kappa statistics were strongest for prone (group = 0.968, individual = (0.942)) and supine (group = 0.907, individual = (0.983)).
+Sitting performance fell in the middle, and was considerably worse for group models than individual models (group = 0.816, individual = (0.915)).
+Held (group = 0.732, individual = (0.826)) and upright (group = 0.707, individual = (0.822)) performance was the least accurate, however, median performance was still within the ``Substantial'' range.
 

### 2023-03-09 manuscript.Rmd
finished results 1

@@
 
-```{r metrics, warning=FALSE, fig.cap="Metrics", fig.align="center", fig.width=5.5, fig.height=5.5, cache=TRUE, cache.extra=tools::md5sum("data/group_split_metrics.csv")}
+```{r metrics, warning=FALSE, fig.cap="Metrics", fig.align="center", fig.width=6, fig.height=3, cache=TRUE, cache.extra=tools::md5sum("data/group_split_metrics.csv")}
 metrics <- read_csv("data/group_split_metrics.csv") %>% 
@@
 metric_plots <- map(metric_labels, ~ metric_plot(metrics_long, .x)) %>% map2(., title_labels, ~.x + ggtitle(.y))
-(metric_plots[[1]] + metric_plots[[2]])/(metric_plots[[3]]+metric_plots[[4]])
+(metric_plots[[1]] + metric_plots[[2]])
 # (metric_plots[[1]] + metric_plots[[2]])/(metric_plots[[3]]+metric_plots[[4]])/(metric_plots[[5]] + metric_plots[[6]])
@@
 In both cases, the testing set was the same, allowing for direct comparison between the two modeling approaches.
+Figure \@ref(fig:metrics), Table \@ref(tab:metrics-table), and Table \@ref(tab:metrics-by-class) summarize the performance of group and individual models using standard metrics for classification.
 
@@
 
-Figure \@ref(fig:metrics) and Table \@ref(tab:metrics-table) summarize the performance of group and individual models using standard metrics for classification.
 Overall accuracy (Figure \@ref(fig:metrics)A) represents the proportion of 4-s windows in the testing set in which the model prediction matched the human annotation of body position.
@@
 ### Sensitivity and Positive Predictive Value
+
+Beyond accuracy, it is important to establish that classification is neither too sensitive nor too conservative in choosing one position versus another.
+Sensitivity refers to the proportion of events of a given position that were correctly identified (e.g., out of 100 human-coded sitting windows, how many of those windows did the model correctly classify as sitting?).
+High sensitivity means that events are unlikely to be missed.
+In contrast, positive predictive value (PPV) refers to the proportion of events classified for a given position actually belonged to that position (e.g., if the model said a baby was upright during 100 windows, how many of those windows were indeed human-coded sitting events?).
+High PPV means that we can be confident in the event label.
+Table \@ref(tab:metrics-by-class) shows the sensitivity and PPV by body position class for group and individual models.
+For group models, sensitivity and PPV were similar: They were highest for supine, prone, and sitting (the most accurately identified class) and lowest for upright and held.
+All values were sufficient (> .8).
+Results for individual models were similar to group models, with the exception of a somewhat lower sensitivity score for held.
+Overall, the results suggest a reasonable balance between sensitivity and PPV across classes for both model types.
 

### 2023-03-09 manuscript.tex
finished results 1

@@
 In both cases, the testing set was the same, allowing for direct comparison between the two modeling approaches.
-
-Figure \ref{fig:metrics} and Table \ref{tab:metrics-table} summarize the performance of group and individual models using standard metrics for classification.
+Figure \ref{fig:metrics}, Table \ref{tab:metrics-table}, and Table \ref{tab:metrics-by-class} summarize the performance of group and individual models using standard metrics for classification.
+
+\hypertarget{overall-accuracy}{%
+\subsubsection{Overall Accuracy}\label{overall-accuracy}}
+
 Overall accuracy (Figure \ref{fig:metrics}A) represents the proportion of 4-s windows in the testing set in which the model prediction matched the human annotation of body position.
@@
 
+\hypertarget{cohens-kappa}{%
+\subsubsection{Cohen's Kappa}\label{cohens-kappa}}
+
 Strong overall accuracy can be misleading when class prevalence is unbalanced.
@@
 As in past work, all body positions were accurately classified even though performance varied somewhat between positions.
-As Table \ref{tab:metrics-by-class} shows, median kappa statistics were strongest for prone (group = 0.968, individual = (0.942)) and supine (group = 0.907, individual = (0.983)).
-Sitting performance fell in the middle, and was considerably worse for group models than individual models (group = 0.816, individual = (0.915)).
-Held (group = 0.732, individual = (0.826)) and upright (group = 0.707, individual = (0.822)) performance was the least accurate, however, median performance was still within the ``Substantial'' range.
+As Table \ref{tab:metrics-by-class} shows, median kappa statistics were strongest for prone (group = 0.968, individual = 0.942) and supine (group = 0.907, individual = 0.983).
+Sitting performance fell in the middle, and was considerably worse for group models than individual models (group = 0.816, individual = 0.915).
+Held (group = 0.732, individual = 0.826) and upright (group = 0.707, individual = 0.822) performance was the least accurate, however, median performance was still within the ``Substantial'' range.
+
+\hypertarget{sensitivity-and-positive-predictive-value}{%
+\subsubsection{Sensitivity and Positive Predictive Value}\label{sensitivity-and-positive-predictive-value}}
+
+Beyond accuracy, it is important to establish that classification is neither too sensitive nor too conservative in choosing one position versus another.
+Sensitivity refers to the proportion of events of a given position that were correctly identified (e.g., out of 100 human-coded sitting windows, how many of those windows did the model correctly classify as sitting?).
+High sensitivity means that events are unlikely to be missed.
+In contrast, positive predictive value (PPV) refers to the proportion of events classified for a given position actually belonged to that position (e.g., if the model said a baby was upright during 100 windows, how many of those windows were indeed human-coded sitting events?).
+High PPV means that we can be confident in the event label.
+Table \ref{tab:metrics-by-class} shows the sensitivity and PPV by body position class for group and individual models.
+For group models, sensitivity and PPV were similar: They were highest for supine, prone, and sitting (the most accurately identified class) and lowest for upright and held.
+All values were sufficient (\textgreater{} .8).
+Results for individual models were similar to group models, with the exception of a somewhat lower sensitivity score for held.
+Overall, the results suggest a reasonable balance between sensitivity and PPV across classes for both model types.
 

### 2023-03-09 manuscript.Rmd
results edits

@@
 This high degree of synchronization makes it possible to link human-coded body position annotations to each 4-s window of motion data, providing ground truth data for model training and testing.
-As in past work [@FranchakScott2021], we compared two types of models.
+As in past work [@FranchakScott2021], we compared two types of models: *group models* and *individual models*.
 To assess the accuracy of body position classification for each recording session, we held back the last 25% of a session's proximal comparison data as the testing set.
 The testing set was never used as training data.
-A *group model* was trained for each session using an aggregated dataset containing the first 75% of all **other** sessions' proximal comparison data.
-This leave-one-out cross-validation tested the generalization of the model to a recording session that was not used at all in training.
-In contrast, *individual models* were trained using the first 75% of a recording session, then tested against the last 25%.
+A group model was trained for each session using an aggregated dataset containing the first 75% of all **other** sessions' proximal comparison data.
+This leave-one-out cross-validation tested the generalization of the model to a recording session that was not used at all in the training set.
+In contrast, individual models were trained using the first 75% of a recording session, then tested against the last 25%.
 The individual model tests whether earlier training data generalize to later testing data within an individual participant's recording.
@@
 Figure \@ref(fig:metrics), Table \@ref(tab:metrics-table), and Table \@ref(tab:metrics-by-class) summarize the performance of group and individual models using standard metrics for classification.
+<!-- GRAPHIC FOR TWO KINDS OF MODELS? -->
 
@@
 Overall accuracy (Figure \@ref(fig:metrics)A) represents the proportion of 4-s windows in the testing set in which the model prediction matched the human annotation of body position.
-Median overall accuracy was `r metrics_lookup["Overall Accuracy", "median_Group"]` for group models and `r metrics_lookup["Overall Accuracy", "median_Individual"]` for individual models.
-Although individual model accuracy was greater than group model accuracy, group model accuracy approached the level of agreement found between two human coders (median = .92) indicating that group models could adequately replace human coders.
-Still, it is notable that the worst-case accuracy from group models (minimum = `r metrics_lookup["Overall Accuracy", "min_Group"]`) was considerably lower than in individual models (minimum = `r metrics_lookup["Overall Accuracy", "min_Individual"]`).
+Overall accuracy for group models  (*M* = `r metrics_lookup["Overall Accuracy", "mean_Group"]`) was slightly lower than accuracy for individual models (*M* = `r metrics_lookup["Overall Accuracy", "mean_Individual"]`).
+Although overall accuracy from our unsupervised, in-home data collection did not match the near-perfect accuracy  (.95-.98) found in prior in-lab studies [@FranchakScott2021; @AiraksinenRasanen2020], both models approached the level of agreement found between two human coders (*M* = .905).
+Most likely, lower accuracy in the current study results from the more variable and complex behavior observed in an unsupervised setting rather than from a difference in the quality of the classification model.
+Visual inspection of Figure \@ref(fig:metrics)A shows that accuracy values were heavily skewed, with most approaching perfect accuracy but a few participants with very poor accuracy.
+Looking at the median performance suggests that the difference between models was not as wide for the typical participant (group median accuracy = `r metrics_lookup["Overall Accuracy", "median_Group"]`; individual median accuracy = `r metrics_lookup["Overall Accuracy", "median_Individual"]`.
+Indeed, it is notable that the worst-case accuracy from group models (minimum = `r metrics_lookup["Overall Accuracy", "min_Group"]`) was considerably lower than in individual models (minimum = `r metrics_lookup["Overall Accuracy", "min_Individual"]`).
 Possibly, individual models accounted for idiosyncrasies in behavior or inconsistencies in sensor placement for those sessions that group models could not generalize to.
@@
 Accordingly, we report Cohen's kappa, a commonly-used metric that penalizes missing rare events (Figure \@ref(fig:metrics)B), and provide classification metrics for each individual body position (Table \@ref(tab:metrics-by-class)) to account for imbalance in body position rates within and between individuals.
-Similar to overall accuracy, kappa values were strong for both model types with median group kappas (`r metrics_lookup["Kappa", "median_Group"]`) somewhat worse compared with median individual kappas (`r metrics_lookup["Kappa", "median_Individual"]`).
+Similar to overall accuracy, kappa values were strong for both model types with group kappas (*M* = `r metrics_lookup["Kappa", "mean_Group"]`) somewhat worse compared with individual kappas (*M* = `r metrics_lookup["Kappa", "mean_Individual"]`).
 Guidelines for interpreting kappa statistics [@LandisKoch1977] consider 0.81–1.00 "Almost Perfect,” 0.61–0.80 “Substantial,” 0.41–0.60 “Moderate,” 0.21– 0.40 “Fair,” and 0–0.20 “Slight to Poor”, indicating that agreement for most group and individual model predictions fell in the Substantial to Almost Perfect range.
 
-As in past work, all body positions were accurately classified even though performance varied somewhat between positions.
-As Table \@ref(tab:metrics-by-class) shows, median kappa statistics were strongest for prone (group = `r metrics_long_summary["Prone Kappa", "median_Group"]`, individual = `r metrics_long_summary["Prone Kappa", "median_Individual"]`) and supine (group = `r metrics_long_summary["Supine Kappa", "median_Group"]`, individual = `r metrics_long_summary["Supine Kappa", "median_Individual"]`).
-Sitting performance fell in the middle, and was considerably worse for group models than individual models (group = `r metrics_long_summary["Sitting Kappa", "median_Group"]`, individual = `r metrics_long_summary["Sitting Kappa", "median_Individual"]`).
-Held (group = `r metrics_long_summary["Held Kappa", "median_Group"]`, individual = `r metrics_long_summary["Held Kappa", "median_Individual"]`) and upright (group = `r metrics_long_summary["Upright Kappa", "median_Group"]`, individual = `r metrics_long_summary["Upright Kappa", "median_Individual"]`) performance was the least accurate, however, median performance was still within the "Substantial" range.
+As in past work [@FranchakScott2021; @AiraksinenRasanen2020], all body positions were accurately classified even though performance varied somewhat between positions.
+As Table \@ref(tab:metrics-by-class) shows, mean kappa statistics were strongest for prone (group *M* =  `r metrics_long_summary["Prone Kappa", "mean_Group"]`, individual *M* =  `r metrics_long_summary["Prone Kappa", "mean_Individual"]`) and supine (group *M* =  `r metrics_long_summary["Supine Kappa", "mean_Group"]`, individual *M* =  `r metrics_long_summary["Supine Kappa", "mean_Individual"]`).
+Sitting performance fell in the middle, and was considerably worse for group models than individual models (group *M* =  `r metrics_long_summary["Sitting Kappa", "mean_Group"]`, individual *M* =  `r metrics_long_summary["Sitting Kappa", "mean_Individual"]`).
+Held (group *M* =  `r metrics_long_summary["Held Kappa", "mean_Group"]`, individual *M* =  `r metrics_long_summary["Held Kappa", "mean_Individual"]`) and upright (group *M* =  `r metrics_long_summary["Upright Kappa", "mean_Group"]`, individual *M* =  `r metrics_long_summary["Upright Kappa", "mean_Individual"]`) performance was the least accurate, however, average performance was still within the "Substantial" range.
 
@@
 Results for individual models were similar to group models, with the exception of a somewhat lower sensitivity score for held.
-Overall, the results suggest a reasonable balance between sensitivity and PPV across classes for both model types.
+Overall, the results suggest a reasonable balance between sensitivity and PPV among different body position classes for both model types.
+Table \@ref(tab:metrics-table) shows the overall sensitivity and PPV across classes.
 

### 2023-03-09 manuscript.tex
results edits

@@
 This high degree of synchronization makes it possible to link human-coded body position annotations to each 4-s window of motion data, providing ground truth data for model training and testing.
-As in past work (Franchak et al., 2021), we compared two types of models.
+As in past work (Franchak et al., 2021), we compared two types of models: \emph{group models} and \emph{individual models}.
 To assess the accuracy of body position classification for each recording session, we held back the last 25\% of a session's proximal comparison data as the testing set.
 The testing set was never used as training data.
-A \emph{group model} was trained for each session using an aggregated dataset containing the first 75\% of all \textbf{other} sessions' proximal comparison data.
-This leave-one-out cross-validation tested the generalization of the model to a recording session that was not used at all in training.
-In contrast, \emph{individual models} were trained using the first 75\% of a recording session, then tested against the last 25\%.
+A group model was trained for each session using an aggregated dataset containing the first 75\% of all \textbf{other} sessions' proximal comparison data.
+This leave-one-out cross-validation tested the generalization of the model to a recording session that was not used at all in the training set.
+In contrast, individual models were trained using the first 75\% of a recording session, then tested against the last 25\%.
 The individual model tests whether earlier training data generalize to later testing data within an individual participant's recording.
@@
 Overall accuracy (Figure \ref{fig:metrics}A) represents the proportion of 4-s windows in the testing set in which the model prediction matched the human annotation of body position.
-Median overall accuracy was 0.89 for group models and 0.93 for individual models.
-Although individual model accuracy was greater than group model accuracy, group model accuracy approached the level of agreement found between two human coders (median = .92) indicating that group models could adequately replace human coders.
-Still, it is notable that the worst-case accuracy from group models (minimum = 0.50) was considerably lower than in individual models (minimum = 0.73).
+Overall accuracy for group models (\emph{M} = 0.85) was slightly lower than accuracy for individual models (\emph{M} = 0.92).
+Although overall accuracy from our unsupervised, in-home data collection did not match the near-perfect accuracy (.95-.98) found in prior in-lab studies (Airaksinen et al., 2020; Franchak et al., 2021), both models approached the level of agreement found between two human coders (\emph{M} = .905).
+Most likely, lower accuracy in the current study results from the more variable and complex behavior observed in an unsupervised setting rather than from a difference in the quality of the classification model.
+Visual inspection of Figure \ref{fig:metrics}A shows that accuracy values were heavily skewed, with most approaching perfect accuracy but a few participants with very poor accuracy.
+Looking at the median performance suggests that the difference between models was not as wide for the typical participant (group median accuracy = 0.89; individual median accuracy = 0.93.
+Indeed, it is notable that the worst-case accuracy from group models (minimum = 0.50) was considerably lower than in individual models (minimum = 0.73).
 Possibly, individual models accounted for idiosyncrasies in behavior or inconsistencies in sensor placement for those sessions that group models could not generalize to.
@@
 Accordingly, we report Cohen's kappa, a commonly-used metric that penalizes missing rare events (Figure \ref{fig:metrics}B), and provide classification metrics for each individual body position (Table \ref{tab:metrics-by-class}) to account for imbalance in body position rates within and between individuals.
-Similar to overall accuracy, kappa values were strong for both model types with median group kappas (0.77) somewhat worse compared with median individual kappas (0.85).
+Similar to overall accuracy, kappa values were strong for both model types with group kappas (\emph{M} = 0.75) somewhat worse compared with individual kappas (\emph{M} = 0.82).
 Guidelines for interpreting kappa statistics (Landis \& Koch, 1977) consider 0.81--1.00 ``Almost Perfect,'' 0.61--0.80 ``Substantial,'' 0.41--0.60 ``Moderate,'' 0.21-- 0.40 ``Fair,'' and 0--0.20 ``Slight to Poor'', indicating that agreement for most group and individual model predictions fell in the Substantial to Almost Perfect range.
 
-As in past work, all body positions were accurately classified even though performance varied somewhat between positions.
-As Table \ref{tab:metrics-by-class} shows, median kappa statistics were strongest for prone (group = 0.968, individual = 0.942) and supine (group = 0.907, individual = 0.983).
-Sitting performance fell in the middle, and was considerably worse for group models than individual models (group = 0.816, individual = 0.915).
-Held (group = 0.732, individual = 0.826) and upright (group = 0.707, individual = 0.822) performance was the least accurate, however, median performance was still within the ``Substantial'' range.
+As in past work (Airaksinen et al., 2020; Franchak et al., 2021), all body positions were accurately classified even though performance varied somewhat between positions.
+As Table \ref{tab:metrics-by-class} shows, median kappa statistics were strongest for prone (group \emph{M} = 0.860, individual \emph{M} = 0.841) and supine (group \emph{M} = 0.764, individual \emph{M} = 0.912).
+Sitting performance fell in the middle, and was considerably worse for group models than individual models (group \emph{M} = 0.702, individual \emph{M} = 0.887).
+Held (group \emph{M} = 0.726, individual \emph{M} = 0.727) and upright (group \emph{M} = 0.673, individual \emph{M} = 0.741) performance was the least accurate, however, median performance was still within the ``Substantial'' range.
 
@@
 Results for individual models were similar to group models, with the exception of a somewhat lower sensitivity score for held.
-Overall, the results suggest a reasonable balance between sensitivity and PPV across classes for both model types.
+Overall, the results suggest a reasonable balance between sensitivity and PPV among different body position classes for both model types.
+Table \ref{tab:metrics-table} shows the overall sensitivity and PPV across classes.
 

### 2023-03-09 manuscript.Rmd
started distal

@@
 This video recorded the next 90 minutes of natural activity.
-However, because there was a variable gap of ~40 s between the two videos, temporal synchrony in the distal comparison video was approximate containing offsets of ~ 5-10 s.
+However, because there was a variable gap of ~40 s between the two videos, temporal synchrony in the distal comparison video was approximate containing offsets of ~ 5 s in either direction.
 <!-- HOW MANY PPTS DON"T HAVE DISTAL COMPARISON VIDEOS? -->
@@
 Figure \@ref(fig:metrics), Table \@ref(tab:metrics-table), and Table \@ref(tab:metrics-by-class) summarize the performance of group and individual models using standard metrics for classification.
+
 <!-- GRAPHIC FOR TWO KINDS OF MODELS? -->
@@
 
-## Goal 2: Assess the distal accuracy of body position classification models
+## Measure the Distal Accuracy of Body Position Classification Models
 
@@
 
+In the first set of results, we showed that group and individual models trained from data during the proximal comparison period were accurate at classifying body position in the same recording.
+In the next analysis, we examine long-term performance by testing how accurately models trained from the proximal period could predict body position during the distal period---the 90-minute long recordings that followed the proximal recordings (Figure \@ref(fig:exemplar-timeline)).
+As previously mentioned, distal videos had only coarse temporal synchrony with motion recordings (misalignment of up to ~ 5 s in either direction).
+This misalignment precluded calculating accuracy based on the proportion of matching events, so we used another technique following past work [@YaoPlotz2019; @FranchakScott2021].
+For each distal comparison, we summed the amount of time infants were predicted to be in each of the 5 body position categories from the model and compared that to the summed time for the body positions based on human coding.
+Below, we report correlations between the model-predicted and human-coded aggregated body position time across the entire distal period as well as within finer, 10-minute bins.
+
+Whereas the first set of analyses used all `r length(unique(ppts$id_uni))` sessions, this was not possible in the distal comparison.
+Because the start of the visit was scheduled during a time the infant was awake, it was common for a nap to be needed following the proximal period.
+Nine sessions were excluded because the infant was either napping or otherwise not on camera during the entire 90-minute distal recording.
+Three additional sessions were excluded because a caregiver accidentally turned off the video camera (*n* = 1) or left the house (*n* = 2).
+This left `r length(unique(agree$group$file))` sessions with usable distal comparison data.
+
+### Overall Agreement During the Distal Comparison
+
 ```{r part2bins, warning=FALSE, fig.cap="Prediction performance (difference in minutes between human-coded and model-predicted body position) for 10-minute bins in the distal comparison.
@@
 ```
+
+### Distal Comparison Agreement at a Finer Timescale
 

### 2023-03-09 manuscript.tex
started distal

@@
 This video recorded the next 90 minutes of natural activity.
-However, because there was a variable gap of \textasciitilde40 s between the two videos, temporal synchrony in the distal comparison video was approximate containing offsets of \textasciitilde{} 5-10 s.
+However, because there was a variable gap of \textasciitilde40 s between the two videos, temporal synchrony in the distal comparison video was approximate containing offsets of \textasciitilde{} 5 s in either direction.
 
@@
 As in past work (Airaksinen et al., 2020; Franchak et al., 2021), all body positions were accurately classified even though performance varied somewhat between positions.
-As Table \ref{tab:metrics-by-class} shows, median kappa statistics were strongest for prone (group \emph{M} = 0.860, individual \emph{M} = 0.841) and supine (group \emph{M} = 0.764, individual \emph{M} = 0.912).
+As Table \ref{tab:metrics-by-class} shows, mean kappa statistics were strongest for prone (group \emph{M} = 0.860, individual \emph{M} = 0.841) and supine (group \emph{M} = 0.764, individual \emph{M} = 0.912).
 Sitting performance fell in the middle, and was considerably worse for group models than individual models (group \emph{M} = 0.702, individual \emph{M} = 0.887).
-Held (group \emph{M} = 0.726, individual \emph{M} = 0.727) and upright (group \emph{M} = 0.673, individual \emph{M} = 0.741) performance was the least accurate, however, median performance was still within the ``Substantial'' range.
+Held (group \emph{M} = 0.726, individual \emph{M} = 0.727) and upright (group \emph{M} = 0.673, individual \emph{M} = 0.741) performance was the least accurate, however, average performance was still within the ``Substantial'' range.
 
@@
 
-\hypertarget{goal-2-assess-the-distal-accuracy-of-body-position-classification-models}{%
-\subsection{Goal 2: Assess the distal accuracy of body position classification models}\label{goal-2-assess-the-distal-accuracy-of-body-position-classification-models}}
+\hypertarget{measure-the-distal-accuracy-of-body-position-classification-models}{%
+\subsection{Measure the Distal Accuracy of Body Position Classification Models}\label{measure-the-distal-accuracy-of-body-position-classification-models}}
 
@@
 
+In the first set of results, we showed that group and individual models trained from data during the proximal comparison period were accurate at classifying body position in the same recording.
+In the next analysis, we examine long-term performance by testing how accurately models trained from the proximal period could predict body position during the distal period---the 90-minute long recordings that followed the proximal recordings (Figure \ref{fig:exemplar-timeline}).
+As previously mentioned, distal videos had only coarse temporal synchrony with motion recordings (misalignment of up to \textasciitilde{} 5 s in either direction).
+This misalignment precluded calculating accuracy based on the proportion of matching events, so we used another technique following past work (Franchak et al., 2021; Yao et al., 2019).
+For each distal comparison, we summed the amount of time infants were predicted to be in each of the 5 body position categories from the model and compared that to the summed time for the body positions based on human coding.
+Below, we report correlations between the model-predicted and human-coded aggregated body position time across the entire distal period as well as within finer, 10-minute bins.
+
+Whereas the first set of analyses used all 34 sessions, this was not possible in the distal comparison.
+Because the start of the visit was scheduled during a time the infant was awake, it was common for a nap to be needed following the proximal period.
+Nine sessions were excluded because the infant was either napping or otherwise not on camera during the entire 90-minute distal recording.
+Three additional sessions were excluded because a caregiver accidentally turned off the video camera (\emph{n} = 1) or left the house (\emph{n} = 2).
+This left 22 sessions with usable distal comparison data.
+
+\hypertarget{overall-agreement-during-the-distal-comparison}{%
+\subsubsection{Overall Agreement During the Distal Comparison}\label{overall-agreement-during-the-distal-comparison}}
+
 \begin{figure}
@@
 \end{figure}
+
+\hypertarget{distal-comparison-agreement-at-a-finer-timescale}{%
+\subsubsection{Distal Comparison Agreement at a Finer Timescale}\label{distal-comparison-agreement-at-a-finer-timescale}}
 

### 2023-03-09 manuscript.Rmd
distal accuracy

@@
 In the next analysis, we examine long-term performance by testing how accurately models trained from the proximal period could predict body position during the distal period---the 90-minute long recordings that followed the proximal recordings (Figure \@ref(fig:exemplar-timeline)).
+A single group model was created using all sessions' proximal period training data (rather than group models leaving out a single session); the same individual models were used.
 As previously mentioned, distal videos had only coarse temporal synchrony with motion recordings (misalignment of up to ~ 5 s in either direction).
@@
 ### Overall Agreement During the Distal Comparison
+
+Figure \@ref(fig:part2overall) and Table \@ref(tab:part2overalltable) summarize the overall agreement during the distal comparison period.
+For each session, we calculated the actual prevalence of each body position as a percentage of the time that the infant was on video and awake using human annotated body position (x-axis on Figure \@ref(fig:part2overall)).
+Predicted prevalence was calculated the same way for group and individual model predictions, omitting the off-camera and nap periods to make a direct comparison.
+Overall---across participants and across body position classes---agreement was strong: The correlation between group model predictions and human-coded prevalence was *r* = `r group_agree_cor`, and the correlation between individual model predictions and human-coded prevalence was *r* = `r split_agree_cor`.
+As in the proximal comparison, agreement varied somewhat between body positions; in particular, agreement for held was poor.
+Unlike in the proximal period, some body positions were better predicted by group models and others by individual models.
+
+Visual inspection of \@ref(fig:part2overall) indicated two extreme outliers, marked by a gray square and a gray diamond.
+We closely investigated each of the two outliers to understand why their agreement was so much worse compared with the other sessions.
+The "gray square" outlier had significant confusion between sitting and supine classification.
+Reviewing the video indicated that this participant spent a long period of time in a seating device that was reclined almost exactly at 45 degrees, making it difficult to determine if the infant was sitting or supine.
+The infant also spent a long time in mother's arms in an ambiguous supine/sitting position that was hard to classify.
+This participant's proximal accuracy was also poor because similar ambiguous appeared during the initial recording.
+In contrast, the "gray diamond outlier" had strong proximal accuracy, with confusion only arising in the distal period between upright and held categories.
+Infant was in a baby walker for most of part 2.
+Moved more like a baby held by a CG than a baby walking upright.
+
+For both, other class performance was good, suggesting that it's about the particular events, not about the quality of the algorithm in applying to the participant overall.
 
@@
 
-### Distal Comparison Agreement at a Finer Timescale
-
 ```{r pt2binstable, cache=TRUE, cache.extra=tools::md5sum("data/compiled_agreement_position.csv"), cache.extra=tools::md5sum("data/compiled_agreement_split.csv")}
@@
 
-What can we learn about the two outliers? What do they tell us about how this will work in the future.
-
-One outlier (106-4) had significant confusion between sitting and supine.
-Both their proximal and distal performance was bad, but only for sitting and supine.
-Watching the videos showed that this participant spent a long period of time in a seating device that was reclined between supine and sitting.
-Also spent long time in mother's arms in a similar supine/sitting position that was hard to classify.
-
-The other outlier (107-3) had good proximal accuracy but poor distal accuracy.
-Confusion between upright and suspended.
-Infant was in a baby walker for most of part 2.
-Moved more like a baby held by a CG than a baby walking upright.
-
-For both, other class performance was good, suggesting that it's about the particular events, not about the quality of the algorithm in applying to the participant overall.
+### Distal Comparison Agreement at a Finer Timescale
 

### 2023-03-09 manuscript.Rmd
first half of distal is done

@@
 
-Figure \@ref(fig:part2overall) and Table \@ref(tab:part2overalltable) summarize the overall agreement during the distal comparison period.
+Figure \@ref(fig:part2overall) and Table \@ref(tab:pt2overalltable) summarize the overall agreement during the distal comparison period.
 For each session, we calculated the actual prevalence of each body position as a percentage of the time that the infant was on video and awake using human annotated body position (x-axis on Figure \@ref(fig:part2overall)).
@@
 In contrast, the "gray diamond outlier" had strong proximal accuracy, with confusion only arising in the distal period between upright and held categories.
-Infant was in a baby walker for most of part 2.
-Moved more like a baby held by a CG than a baby walking upright.
+Reviewing the video showed that all disagreements occurred when the infant was in a baby walker; human coders scored this as "upright" but the models predicted it as "held".
+Most likely, the infant's movements in the baby walker were more similar to how a baby moved while carried, and unlike how most infant's moved while walking upright.
 
-For both, other class performance was good, suggesting that it's about the particular events, not about the quality of the algorithm in applying to the participant overall.
+What is notable about both outliers is that disagreements were restricted to a particular border case (supine vs.
+sitting; upright vs.
+held) and their accuracy for other classes was strong.
+This suggests that their poor performance came as a result of spending a long time in an ambiguous position, not the result of the entire model failing to generalize to the later time period (or an error in sensor placement, such as if the parent removed the leggings and put them on backwards after a diaper change).
+To better capture the typical level of agreement, we report all correlations in Table \@ref(tab:pt2overalltable) excluding the two outliers.
+Overall agreement among the non-outlier sessions was excellent for both group models (*r* = `r group_agree_cor_noout`) and individual models (*r* = `r split_agree_cor_noout`).
+With the exception of held, within-class agreement was strong for the other four body positions.
 

### 2023-03-09 manuscript.tex
first half of distal is done

@@
 In the next analysis, we examine long-term performance by testing how accurately models trained from the proximal period could predict body position during the distal period---the 90-minute long recordings that followed the proximal recordings (Figure \ref{fig:exemplar-timeline}).
+A single group model was created using all sessions' proximal period training data (rather than group models leaving out a single session); the same individual models were used.
 As previously mentioned, distal videos had only coarse temporal synchrony with motion recordings (misalignment of up to \textasciitilde{} 5 s in either direction).
@@
 \subsubsection{Overall Agreement During the Distal Comparison}\label{overall-agreement-during-the-distal-comparison}}
+
+Figure \ref{fig:part2overall} and Table \ref{tab:part2overalltable} summarize the overall agreement during the distal comparison period.
+For each session, we calculated the actual prevalence of each body position as a percentage of the time that the infant was on video and awake using human annotated body position (x-axis on Figure \ref{fig:part2overall}).
+Predicted prevalence was calculated the same way for group and individual model predictions, omitting the off-camera and nap periods to make a direct comparison.
+Overall---across participants and across body position classes---agreement was strong: The correlation between group model predictions and human-coded prevalence was \emph{r} = 0.80, and the correlation between individual model predictions and human-coded prevalence was \emph{r} = 0.91.
+As in the proximal comparison, agreement varied somewhat between body positions; in particular, agreement for held was poor.
+Unlike in the proximal period, some body positions were better predicted by group models and others by individual models.
+
+Visual inspection of \ref{fig:part2overall} indicated two extreme outliers, marked by a gray square and a gray diamond.
+We closely investigated each of the two outliers to understand why their agreement was so much worse compared with the other sessions.
+The ``gray square'' outlier had significant confusion between sitting and supine classification.
+Reviewing the video indicated that this participant spent a long period of time in a seating device that was reclined almost exactly at 45 degrees, making it difficult to determine if the infant was sitting or supine.
+The infant also spent a long time in mother's arms in an ambiguous supine/sitting position that was hard to classify.
+This participant's proximal accuracy was also poor because similar ambiguous appeared during the initial recording.
+In contrast, the ``gray diamond outlier'' had strong proximal accuracy, with confusion only arising in the distal period between upright and held categories.
+Reviewing the video showed that all disagreements occurred when the infant was in a baby walker; human coders scored this as ``upright'' but the models predicted it as ``held''.
+Most likely, the infant's movements in the baby walker were more similar to how a baby moved while carried, and unlike how most infant's moved while walking upright.
+
+What is notable about both outliers is that disagreements were restricted to a particular border case (supine vs.~sitting; upright vs.~held) and their accuracy for other classes was strong.
+This suggests that their poor performance came as a result of spending a long time in an ambiguous position, not the result of the entire model failing to generalize to the later time period (or an error in sensor placement, such as if the parent removed the leggings and put them on backwards after a diaper change).
+To better capture the typical level of agreement, we report all correlations in Table \ref{tab:part2overalltable} excluding the two outliers.
+Overall agreement among the non-outlier sessions was excellent for both group models (\emph{r} = 0.95) and individual models (\emph{r} = 0.96).
+With the exception of held, within-class agreement was strong for the other four body positions.
 
@@
 \end{figure}
-
-\hypertarget{distal-comparison-agreement-at-a-finer-timescale}{%
-\subsubsection{Distal Comparison Agreement at a Finer Timescale}\label{distal-comparison-agreement-at-a-finer-timescale}}
 
@@
 
-What can we learn about the two outliers? What do they tell us about how this will work in the future.
-
-One outlier (106-4) had significant confusion between sitting and supine.
-Both their proximal and distal performance was bad, but only for sitting and supine.
-Watching the videos showed that this participant spent a long period of time in a seating device that was reclined between supine and sitting.
-Also spent long time in mother's arms in a similar supine/sitting position that was hard to classify.
-
-The other outlier (107-3) had good proximal accuracy but poor distal accuracy.
-Confusion between upright and suspended.
-Infant was in a baby walker for most of part 2.
-Moved more like a baby held by a CG than a baby walking upright.
-
-For both, other class performance was good, suggesting that it's about the particular events, not about the quality of the algorithm in applying to the participant overall.
+\hypertarget{distal-comparison-agreement-at-a-finer-timescale}{%
+\subsubsection{Distal Comparison Agreement at a Finer Timescale}\label{distal-comparison-agreement-at-a-finer-timescale}}
 

### 2023-03-09 manuscript.Rmd
10 min bins

@@
 
-Visual inspection of \@ref(fig:part2overall) indicated two extreme outliers, marked by a gray square and a gray diamond.
-We closely investigated each of the two outliers to understand why their agreement was so much worse compared with the other sessions.
+Visual inspection of Figure \@ref(fig:part2overall) indicated two extreme outliers, which we marked by a gray square and a gray diamond.
+We closely investigated each of the two outliers to understand why their agreement was below the standard of the other sessions.
 The "gray square" outlier had significant confusion between sitting and supine classification.
@@
 sitting; upright vs.
-held) and their accuracy for other classes was strong.
+held); accuracy for other classes remained strong.
 This suggests that their poor performance came as a result of spending a long time in an ambiguous position, not the result of the entire model failing to generalize to the later time period (or an error in sensor placement, such as if the parent removed the leggings and put them on backwards after a diaper change).
@@
 
-### Distal Comparison Agreement at a Finer Timescale
+### Short-Timescale Agreement during the Distal Comparison
+
+Collecting real-time data that can be synchronized to other sources is a key goal in our development of long-form body position classification.
+Although overall aggregate agreement in the distal comparison was strong, it is important to show that similarly strong agreement is found within a shorter timescale.
+We repeated agreement analysis, correlating human-coded body position prevalence with model-predicted prevalence, after dividing the distal comparison period into nine 10-min bins (marked by vertical dashed lines in Figure \@ref(fig:exemplar-timeline)).
 

### 2023-03-09 manuscript.Rmd
finished distal

@@
 
-```{r part2bins, warning=FALSE, fig.cap="Prediction performance (difference in minutes between human-coded and model-predicted body position) for 10-minute bins in the distal comparison.
+```{r part2bins, warning=FALSE, fig.cap="Prediction difference (difference in minutes between human-coded and model-predicted body position) for 10-minute bins in the distal comparison.
 Each point shows the mean and SE for a single participant for each body position, summarizing the prediction difference for each of their 10-minute bins.
@@
 }
+#Difference variable is scaled as proportion, but plotted against min/max of 10 minutes
+
 difference_plots <- map(agree_bins, ~ difference_plot(.x))
@@
 
+```{r prediction-difference}
+group_diff <- agree_bins$group %>% 
+  mutate(diff = (prop_human - prop_model)*10,
+         less_than_1 = ifelse(diff >= -1 & diff <= 1, 1, 0))
+group_diff <- group_diff %>% group_by(pos) %>% mutate(n = n()) %>% summarize(prop = sum(less_than_1), total = mean(n)) %>% mutate(prop = apa_num(prop/total*100)) %>% column_to_rownames("pos")
+
+ind_diff <- agree_bins$split %>% 
+  mutate(diff = (prop_human - prop_model)*10,
+         less_than_1 = ifelse(diff >= -1 & diff <= 1, 1, 0))
+ind_diff <- ind_diff %>% group_by(pos) %>% mutate(n = n()) %>% summarize(prop = sum(less_than_1), total = mean(n)) %>% mutate(prop = apa_num(prop/total*100)) %>% column_to_rownames("pos")
+```
+
 ### Short-Timescale Agreement during the Distal Comparison
@@
 Although overall aggregate agreement in the distal comparison was strong, it is important to show that similarly strong agreement is found within a shorter timescale.
-We repeated agreement analysis, correlating human-coded body position prevalence with model-predicted prevalence, after dividing the distal comparison period into nine 10-min bins (marked by vertical dashed lines in Figure \@ref(fig:exemplar-timeline)).
+We repeated agreement analysis, correlating human-coded body position prevalence with model-predicted prevalence, after dividing the distal comparison period into nine 10-minute bins (marked by vertical dashed lines in Figure \@ref(fig:exemplar-timeline)).
+Infants had varying numbers of 10-minute bins depending on much time they were awake and on camera.
+Bins were included only if there was > X minutes of usable data.
+<!-- RANGE OF BINSm USABLE DATA -->
+
+Table \@ref(tab:pt2binstable) shows the agreement correlation coefficients for group and individual models, including and excluding the outliers identified in the previous section.
+Performance at a short timescale was similar to performance overall: Overall agreement after excluding outliers was excellent for both group (*r* = `r group_agree_cor_noout`) and individual models (*r* = `r split_agree_cor_noout`).
+Within-position correlations were weakest for held and strongest for upright regardless of the model type.
+Agreement for prone was better for group models, whereas sitting and supine were better predicted by individual models.
+
+To describe the observed amount of prediction error in 10-minute bins, we subtracted the predicted duration (in minutes) for each body position in each bin from the human coded duration in that bin to create a prediction difference score.
+A score of 0 would indicate no error; positive differences indicate that the model overestimated the amount of time in a position, whereas negative differences indicate underestimation.
+Figure \@ref(fig:part2bins) plots the mean prediction difference for each session for each body position.
+The gray shaded area marks ± 1 minute of prediction error.
+The vast majority of session-averaged predictions fall within 1 minute of error without a clear bias towards overestimation or underestimation (with the previously-identified outliers as exceptions).
+For group models, we calculated the percentage of 10-min bins across participants that had errors < 1 minute: `r group_diff["Held","prop"]`% for held, `r group_diff["Supine","prop"]`% for supine, `r group_diff["Prone","prop"]`% for prone, `r group_diff["Sitting","prop"]`% for sitting, and `r group_diff["Upright","prop"]`% for upright.
+For individual models, the percent of 10-min bins with < 1 minute of error was: `r ind_diff["Held","prop"]`% for held, `r ind_diff["Supine","prop"]`% for supine, `r ind_diff["Prone","prop"]`% for prone, `r ind_diff["Sitting","prop"]`% for sitting, and `r ind_diff["Upright","prop"]`% for upright.
+Taken together, these results suggest that most estimates across models, sessions, and body positions had small amounts of prediction error even after a long delay separating the period that models were trained and tested.
 

### 2023-03-09 manuscript.tex
finished distal

@@
 
-Figure \ref{fig:part2overall} and Table \ref{tab:part2overalltable} summarize the overall agreement during the distal comparison period.
+Figure \ref{fig:part2overall} and Table \ref{tab:pt2overalltable} summarize the overall agreement during the distal comparison period.
 For each session, we calculated the actual prevalence of each body position as a percentage of the time that the infant was on video and awake using human annotated body position (x-axis on Figure \ref{fig:part2overall}).
@@
 
-Visual inspection of \ref{fig:part2overall} indicated two extreme outliers, marked by a gray square and a gray diamond.
-We closely investigated each of the two outliers to understand why their agreement was so much worse compared with the other sessions.
+Visual inspection of Figure \ref{fig:part2overall} indicated two extreme outliers, which we marked by a gray square and a gray diamond.
+We closely investigated each of the two outliers to understand why their agreement was below the standard of the other sessions.
 The ``gray square'' outlier had significant confusion between sitting and supine classification.
@@
 
-What is notable about both outliers is that disagreements were restricted to a particular border case (supine vs.~sitting; upright vs.~held) and their accuracy for other classes was strong.
+What is notable about both outliers is that disagreements were restricted to a particular border case (supine vs.~sitting; upright vs.~held); accuracy for other classes remained strong.
 This suggests that their poor performance came as a result of spending a long time in an ambiguous position, not the result of the entire model failing to generalize to the later time period (or an error in sensor placement, such as if the parent removed the leggings and put them on backwards after a diaper change).
-To better capture the typical level of agreement, we report all correlations in Table \ref{tab:part2overalltable} excluding the two outliers.
+To better capture the typical level of agreement, we report all correlations in Table \ref{tab:pt2overalltable} excluding the two outliers.
 Overall agreement among the non-outlier sessions was excellent for both group models (\emph{r} = 0.95) and individual models (\emph{r} = 0.96).
@@
 
-\caption{Prediction performance (difference in minutes between human-coded and model-predicted body position) for 10-minute bins in the distal comparison.
+\caption{Prediction difference (difference in minutes between human-coded and model-predicted body position) for 10-minute bins in the distal comparison.
 Each point shows the mean and SE for a single participant for each body position, summarizing the prediction difference for each of their 10-minute bins.
@@
 
-\hypertarget{distal-comparison-agreement-at-a-finer-timescale}{%
-\subsubsection{Distal Comparison Agreement at a Finer Timescale}\label{distal-comparison-agreement-at-a-finer-timescale}}
+\hypertarget{short-timescale-agreement-during-the-distal-comparison}{%
+\subsubsection{Short-Timescale Agreement during the Distal Comparison}\label{short-timescale-agreement-during-the-distal-comparison}}
+
+Collecting real-time data that can be synchronized to other sources is a key goal in our development of long-form body position classification.
+Although overall aggregate agreement in the distal comparison was strong, it is important to show that similarly strong agreement is found within a shorter timescale.
+We repeated agreement analysis, correlating human-coded body position prevalence with model-predicted prevalence, after dividing the distal comparison period into nine 10-minute bins (marked by vertical dashed lines in Figure \ref{fig:exemplar-timeline}).
+Infants had varying numbers of 10-minute bins depending on much time they were awake and on camera.
+Bins were included only if there was \textgreater{} X minutes of usable data.
+
+Table \ref{tab:pt2binstable} shows the agreement correlation coefficients for group and individual models, including and excluding the outliers identified in the previous section.
+Performance at a short timescale was similar to performance overall: Overall agreement after excluding outliers was excellent for both group (\emph{r} = 0.92) and individual models (\emph{r} = 0.94).
+Within-position correlations were weakest for held and strongest for upright regardless of the model type.
+Agreement for prone was better for group models, whereas sitting and supine were better predicted by individual models.
+
+To describe the observed amount of prediction error in 10-minute bins, we subtracted the predicted duration (in minutes) for each body position in each bin from the human coded duration in that bin to create a prediction difference score.
+A score of 0 would indicate no error; positive differences indicate that the model overestimated the amount of time in a position, whereas negative differences indicate underestimation.
+Figure \ref{fig:part2bins} plots the mean prediction difference for each session for each body position.
+The gray shaded area marks ± 1 minute of prediction error.
+The vast majority of session-averaged predictions fall within 1 minute of error without a clear bias towards overestimation or underestimation (with the previously-identified outliers as exceptions).
+For group models, we calculated the percentage of 10-min bins across participants that had errors \textless{} 1 minute: 94.68\% for held, 80.85\% for supine, 94.68\% for prone, 77.66\% for sitting, and 94.68\% for upright.
+For individual models, the percent of 10-min bins with \textless{} 1 minute of error was: 92.94\% for held, 88.24\% for supine, 88.24\% for prone, 83.53\% for sitting, and 96.47\% for upright.
+Taken together, these results suggest that most estimates across models, sessions, and body positions had small amounts of prediction error even after a long delay from when models were trained.
 

### 2023-03-10 manuscript.Rmd
fixed timelines

@@
 Interpreting such data required caregivers to log when infants napped, when they removed the sensor garment for diaper changes or other reasons, and when infants went to bed at the end of the day.
-The **third goal** of the study was to examine the quality of the full-day data.
+Finally, the **third goal** of the study was to examine the quality of the full-day data.
 Could body position be successfully recorded over the desired period?
-Finally, the novel full-day dataset affords us a unique chance to ask whether the estimated time infants spent in different body positions align with past results using video and survey methods.
-Thus, the **fourth goal** of the current study was to determine whether full-day body position measurements conformed to expectations about age differences in body position.
+The novel full-day dataset affords us a unique chance to ask whether the estimated time infants spent in different body positions align with past results using video and survey methods.
+Thus, we determined whether full-day body position measurements conformed to expected age differences in body position.
 Based on past results [@Survey], infants should spend increasingly more time sitting and upright but less time supine over the age range tested (4 to 14 months).
@@
 
-## Goal 3: Examine the data quality of full-day home recordings
+## Examine the Data Quality of Full-Day Home Recordings
 
@@
 timelines <- function(ds) {
+  ds$id_uni <- as.numeric(factor(ds$id_uni))
   lims <- as_hms(c('09:00:00', '21:59:00'))
@@
   facet_wrap("id_uni",ncol = 1, strip.position = "left") + 
-  scale_fill_manual(values = c("Nap" = "gray",pal), name = "") + 
+  scale_fill_manual(values = c("Nap" = "gray",pal), name = "") + ylab("") + 
   scale_x_time(breaks = hour_breaks, labels = label_breaks, name = "", limits = lims) + 
@@
         axis.text.y = element_blank(),
-        strip.background = element_blank(),
         axis.text.x = element_text(size = 11),
-        strip.text.y = element_blank()) + ylab("")
+        strip.text.y.left = element_text(angle = 360))
   #https://ggplot2tor.com/scales/scale_x_time for options
+  # strip.text.y = element_blank()
+  # strip.background = element_blank(),
 }
@@
 timeline_plots <- map(c("Younger", "Older"), ~ timelines(filter(timeline_data, group == .x)))
-timeline_plots <- map2(timeline_plots, c("A. 4-7 Months", "B. 11-14 Months"), ~ .x + ggtitle(.y))
+timeline_plots <- map2(timeline_plots, c("A. Younger Infants (4-7 mo)", "B. Older Infants (11-14 mo)"), ~ .x + ggtitle(.y))
 timeline_plots[[1]] + timeline_plots[[2]]
@@
 start_times <- str_sub(as.character(start_times), start = 1, end = -4)
+
+end_times <- ppts %>% group_by(id_uni) %>% slice_max(clock_time_start) %>% pull(clock_time_start) %>% sort
+end_times <- str_sub(as.character(end_times), start = 1, end = -4)
+
+coverage <- ppts %>% mutate(missing = ifelse(is.na(nap_time), 1, 0)) %>%
+  group_by(id_uni) %>% 
+  summarize(blocks = n(), missing = sum(missing, na.rm = T), naps = sum(nap_time, na.rm = T)) %>% 
+  mutate(total = blocks * 5/60, prop_accounted = 1-(missing/blocks), naptime = naps * 5)
 ```
+
+After the distal comparison video ended, caregivers were instructed to keep the sensors on their infants for the remainder of the day until infants went to bed, removing the sensors for naps, diaper changes, and trips out of the house.
+The first two sets of results show that accuracy was consistently high across the proximal and distal recordings, making us confident that predictions over the remainder of the day would continue to be accurate after the video recordings stopped.
+
+This leads to our last questions---how successfully did recordings capture infants' entire day, and how suitable were daylong recordings for capturing individual differences in body position?
+Figure \@ref(fig:timelines) depicts body position timelines across the day for each session, divided into younger and older infants, predicted from the group model.
 The start time ranged from `r start_times[1]` to `r start_times[length(start_times)]` with a median of `r start_times[floor(length(start_times)/2)]`.
-Although the equipment was always dropped off in the morning,
-
-## Goal 4: Assess the suitability of full-day predictions for capturing age differences in body position
-
-<!-- imu-compiled.csv, generated from analysis-lena-imu/compile_lena_imu.R -->
+Although the equipment was always dropped off in the morning, sometimes infants were unexpectedly asleep, leading to a few sessions in which infants began wearing the sensors later than intended.
 

### 2023-03-10 manuscript.tex
fixed timelines

@@
 Interpreting such data required caregivers to log when infants napped, when they removed the sensor garment for diaper changes or other reasons, and when infants went to bed at the end of the day.
-The \textbf{third goal} of the study was to examine the quality of the full-day data.
+Finally, the \textbf{third goal} of the study was to examine the quality of the full-day data.
 Could body position be successfully recorded over the desired period?
-Finally, the novel full-day dataset affords us a unique chance to ask whether the estimated time infants spent in different body positions align with past results using video and survey methods.
-Thus, the \textbf{fourth goal} of the current study was to determine whether full-day body position measurements conformed to expectations about age differences in body position.
+The novel full-day dataset affords us a unique chance to ask whether the estimated time infants spent in different body positions align with past results using video and survey methods.
+Thus, we determined whether full-day body position measurements conformed to expected age differences in body position.
 Based on past results (Franchak, 2019), infants should spend increasingly more time sitting and upright but less time supine over the age range tested (4 to 14 months).
@@
 For individual models, the percent of 10-min bins with \textless{} 1 minute of error was: 92.94\% for held, 88.24\% for supine, 88.24\% for prone, 83.53\% for sitting, and 96.47\% for upright.
-Taken together, these results suggest that most estimates across models, sessions, and body positions had small amounts of prediction error even after a long delay from when models were trained.
-
-\hypertarget{goal-3-examine-the-data-quality-of-full-day-home-recordings}{%
-\subsection{Goal 3: Examine the data quality of full-day home recordings}\label{goal-3-examine-the-data-quality-of-full-day-home-recordings}}
+Taken together, these results suggest that most estimates across models, sessions, and body positions had small amounts of prediction error even after a long delay separating the period that models were trained and tested.
+
+\hypertarget{examine-the-data-quality-of-full-day-home-recordings}{%
+\subsection{Examine the Data Quality of Full-Day Home Recordings}\label{examine-the-data-quality-of-full-day-home-recordings}}
 
@@
 
-The start time ranged from 08:55 to 13:20 with a median of 10:20. Although the equipment was always dropped off in the morning,
-
-\hypertarget{goal-4-assess-the-suitability-of-full-day-predictions-for-capturing-age-differences-in-body-position}{%
-\subsection{Goal 4: Assess the suitability of full-day predictions for capturing age differences in body position}\label{goal-4-assess-the-suitability-of-full-day-predictions-for-capturing-age-differences-in-body-position}}
+After the distal comparison video ended, caregivers were instructed to keep the sensors on their infants for the remainder of the day until infants went to bed, removing the sensors for naps, diaper changes, and trips out of the house.
+The first two sets of results show that accuracy was consistently high across the proximal and distal recordings, making us confident that predictions over the remainder of the day would continue to be accurate after the video recordings stopped.
+
+This leads to our last questions---how successfully did recordings capture infants' entire day, and how suitable were daylong recordings for capturing individual differences in body position?
+Figure \ref{fig:timelines} depicts body position timelines across the day for each session, divided into younger and older infants, predicted from the group model.
+The start time ranged from 08:55 to 13:20 with a median of 10:20.
+Although the equipment was always dropped off in the morning, sometimes infants were unexpectedly asleep, leading to a few sessions in which infants began wearing the sensors later than intended.
 

### 2023-03-10 manuscript.Rmd
recording stats

@@
   mutate(id_uni = factor(id*100+session),
-         group = factor(ifelse(age <= 8, "Younger", "Older")))
+         group = factor(ifelse(age <= 8, "Younger", "Older"))) %>% 
+  filter(id_uni != 10401)
 num_younger <- ppts %>% filter(group == "Younger") %>% .$id %>% unique(.) %>% length(.)
@@
 
-end_times <- ppts %>% group_by(id_uni) %>% slice_max(clock_time_start) %>% pull(clock_time_start) %>% sort
+end_times <- ppts %>% filter(id_uni != 11104, id_uni != 801) %>%
+  group_by(id_uni) %>% slice_max(clock_time_start) %>% pull(clock_time_start) %>% sort
 end_times <- str_sub(as.character(end_times), start = 1, end = -4)
@@
 coverage <- ppts %>% mutate(missing = ifelse(is.na(nap_time), 1, 0)) %>%
+  filter(id_uni != 11104, id_uni != 801) %>% 
   group_by(id_uni) %>% 
   summarize(blocks = n(), missing = sum(missing, na.rm = T), naps = sum(nap_time, na.rm = T)) %>% 
-  mutate(total = blocks * 5/60, prop_accounted = 1-(missing/blocks), naptime = naps * 5)
+  mutate(total = blocks * 5/60, 
+         prop_accounted = 1-(missing/(blocks-missing)), 
+         hours_recorded = (blocks-missing-naps)*5/60,
+         naptime = naps * 5/60)
 ```
@@
 After the distal comparison video ended, caregivers were instructed to keep the sensors on their infants for the remainder of the day until infants went to bed, removing the sensors for naps, diaper changes, and trips out of the house.
-The first two sets of results show that accuracy was consistently high across the proximal and distal recordings, making us confident that predictions over the remainder of the day would continue to be accurate after the video recordings stopped.
+The first two sets of results show that accuracy was consistently high across the proximal and distal recordings, providing confidence that predictions over the remainder of the day would continue to be accurate after the video recordings stopped.
 
-This leads to our last questions---how successfully did recordings capture infants' entire day, and how suitable were daylong recordings for capturing individual differences in body position?
-Figure \@ref(fig:timelines) depicts body position timelines across the day for each session, divided into younger and older infants, predicted from the group model.
-The start time ranged from `r start_times[1]` to `r start_times[length(start_times)]` with a median of `r start_times[floor(length(start_times)/2)]`.
-Although the equipment was always dropped off in the morning, sometimes infants were unexpectedly asleep, leading to a few sessions in which infants began wearing the sensors later than intended.
+This leads to two final questions---how successfully did recordings capture infants' entire day, and how suitable were day-long recordings for capturing individual differences in body position?
+Figure \@ref(fig:timelines) depicts body position timelines across the day for each session, divided into younger (A) and older infants (B).
+Predictions from the group model were used to ensure that motion data were classified consistently across all sessions.
+Session start times ranged from `r start_times[1]` to `r start_times[length(start_times)]` with a median of `r start_times[floor(length(start_times)/2)]`.
+Although the equipment was always dropped off in the morning, sometimes infants were unexpectedly asleep, leading to a few sessions in which infants began wearing the sensors later than intended (such as \#11 in the younger group).
+With two exceptions, recordings lasted until the infants' bed time.
+Older participant \#1 had the equipment picked up on the same day rather than the next day, so the recording ended at 17:00.
+Older participant \#11 wore the equipment in the morning, but the family left the house at 10:15 and remained out for the rest of the day, choosing not to put the equipment back on when they returned home in the late afternoon.
+Among those participants who wore the equipment until bedtime, recordings ended between `r end_times[1]` and `r end_times[length(end_times)]` with a median of `r end_times[floor(length(end_times)/2)]`.
+
+Among participants who wore the sensors from morning to bedtime, the majority of the time during the day either resulted in usable body position data or was excluded due to caregiver-reported naps (gray periods in Figure \@ref(fig:timelines)).
+The total length of the recording period ranged from `r apa_num(min(coverage$total))`-`r apa_num(max(coverage$total))` hours with a median of `r apa_num(median(coverage$total))` hours.
+Nap times reported during the recording period ranged from `r apa_num(min(coverage$naptime))`-`r apa_num(max(coverage$naptime))` hours with a median of `r apa_num(median(coverage$naptime))` hours.
+Of the remaining awake time during the recording period, body position data was collecting during proportions that ranged from `r apa_num(min(coverage$prop_accounted))`-`r apa_num(max(coverage$prop_accounted))` hours with a median of `r apa_num(median(coverage$prop_accounted))`.
+Younger infant \#13 had the least portion of the waking day accounted for; the family left the house to run errands during the majority of the recording period, and the infant napped during much of the remaining time at home.
+Overall, the recordings produced a median of `r apa_num(median(coverage$hours_recorded))` hours of motion data, with the entire dataset totaling `r apa_num(sum(coverage$hours_recorded))` hours.
 

### 2023-03-10 manuscript.tex
recording stats

@@
 Infants were recruited in one of two age groups: \emph{Younger} infants were between 4 and 7 months and \emph{Older} infants were between 11 and 14 months.
-There were 9 infants in the 4-7 month group (X female) and 14 in the 11-14 month group (X female).
+There were 8 infants in the 4-7 month group (X female) and 14 in the 11-14 month group (X female).
 Families were recruited through social media advertisements and from community events in Southern California.
@@
 
-Most participants were tested in a single session (\emph{n} = 17), but 6 participants contributed between 2-4 sessions as part of an ongoing longitudinal study.
+Most participants were tested in a single session (\emph{n} = 16), but 6 participants contributed between 2-4 sessions as part of an ongoing longitudinal study.
 Only 1 session was excluded due to a technical error---one of the four IMU sensors failed to record, resulting in an unusable set of data for classification.
-Across the two age groups, we report data on a total of 34 sessions, with 14 sessions from younger infants and 20 sessions from older infants.
-Across sessions, younger infants' age ranged from 3.84 to 7.16 (\emph{M} = 4.99) and older infants' age ranged from 10.74 to 14.23 (\emph{M} = 11.75).
+Across the two age groups, we report data on a total of 33 sessions, with 13 sessions from younger infants and 20 sessions from older infants.
+Across sessions, younger infants' age ranged from 3.84 to 7.16 (\emph{M} = 5.03) and older infants' age ranged from 10.74 to 14.23 (\emph{M} = 11.75).
 
@@
 
-We report four sets of results based on 34 full-day testing sessions resulting in a total of 301.74 hours of movement recording.
+We report four sets of results based on 33 full-day testing sessions resulting in a total of 294.07 hours of movement recording.
 First, we focus on the accuracy of group and individual models trained and validated on the proximal comparison period.
@@
 
-Whereas the first set of analyses used all 34 sessions, this was not possible in the distal comparison.
+Whereas the first set of analyses used all 33 sessions, this was not possible in the distal comparison.
 Because the start of the visit was scheduled during a time the infant was awake, it was common for a nap to be needed following the proximal period.
@@
 After the distal comparison video ended, caregivers were instructed to keep the sensors on their infants for the remainder of the day until infants went to bed, removing the sensors for naps, diaper changes, and trips out of the house.
-The first two sets of results show that accuracy was consistently high across the proximal and distal recordings, making us confident that predictions over the remainder of the day would continue to be accurate after the video recordings stopped.
-
-This leads to our last questions---how successfully did recordings capture infants' entire day, and how suitable were daylong recordings for capturing individual differences in body position?
-Figure \ref{fig:timelines} depicts body position timelines across the day for each session, divided into younger and older infants, predicted from the group model.
-The start time ranged from 08:55 to 13:20 with a median of 10:20.
-Although the equipment was always dropped off in the morning, sometimes infants were unexpectedly asleep, leading to a few sessions in which infants began wearing the sensors later than intended.
+The first two sets of results show that accuracy was consistently high across the proximal and distal recordings, providing confidence that predictions over the remainder of the day would continue to be accurate after the video recordings stopped.
+
+This leads to two final questions---how successfully did recordings capture infants' entire day, and how suitable were day-long recordings for capturing individual differences in body position?
+Figure \ref{fig:timelines} depicts body position timelines across the day for each session, divided into younger (A) and older infants (B).
+Predictions from the group model were used to ensure that motion data were classified consistently across all sessions.
+Session start times ranged from 08:55 to 13:20 with a median of 10:20.
+Although the equipment was always dropped off in the morning, sometimes infants were unexpectedly asleep, leading to a few sessions in which infants began wearing the sensors later than intended (such as \#11 in the younger group).
+With two exceptions, recordings lasted until the infants' bed time.
+Older participant \#1 had the equipment picked up on the same day rather than the next day, so the recording ended at 17:00.
+Older participant \#11 wore the equipment in the morning, but the family left the house at 10:15 and remained out for the rest of the day, choosing not to put the equipment back on when they returned home in the late afternoon.
+Among those participants who wore the equipment until bedtime, recordings ended between 17:25 and 22:15 with a median of 19:20.
+
+Among participants who wore the sensors from morning to bedtime, the majority of the time during the day either resulted in usable body position data or was excluded due to caregiver-reported naps (gray periods in Figure \ref{fig:timelines}).
+The total length of the recording period ranged from 6.92-11.75 hours with a median of 9.08 hours.
+Nap times reported during the recording period ranged from 0.00-4.99 hours with a median of 2.22 hours.
+Of the remaining awake time during the recording period, body position data was collecting during proportions that ranged from 0.18-1.00 hours with a median of 1.00.
+Younger infant \#13 had the least portion of the waking day accounted for; the family left the house to run errands during the majority of the recording period, and the infant napped during much of the remaining time at home.
+Overall, the recordings produced a median of 6.15 hours of motion data, with the entire dataset totaling 201.06 hours.
 

### 2023-03-10 manuscript.Rmd
more results

@@
 The first two sets of results show that accuracy was consistently high across the proximal and distal recordings, providing confidence that predictions over the remainder of the day would continue to be accurate after the video recordings stopped.
+This leads to two final questions---how successfully did recordings capture infants' entire day, and how suitable were day-long recordings for capturing individual differences in body position?
 
-This leads to two final questions---how successfully did recordings capture infants' entire day, and how suitable were day-long recordings for capturing individual differences in body position?
+### How Well did Recordings Capture Infants' Entire Day?
+
 Figure \@ref(fig:timelines) depicts body position timelines across the day for each session, divided into younger (A) and older infants (B).
@@
 Nap times reported during the recording period ranged from `r apa_num(min(coverage$naptime))`-`r apa_num(max(coverage$naptime))` hours with a median of `r apa_num(median(coverage$naptime))` hours.
-Of the remaining awake time during the recording period, body position data was collecting during proportions that ranged from `r apa_num(min(coverage$prop_accounted))`-`r apa_num(max(coverage$prop_accounted))` hours with a median of `r apa_num(median(coverage$prop_accounted))`.
+Body position data was available to describe `r apa_num(min(coverage$prop_accounted))`-`r apa_num(max(coverage$prop_accounted))`% of the awake portion of the recording period (median = `r apa_num(median(coverage$prop_accounted))`%).
 Younger infant \#13 had the least portion of the waking day accounted for; the family left the house to run errands during the majority of the recording period, and the infant napped during much of the remaining time at home.
 Overall, the recordings produced a median of `r apa_num(median(coverage$hours_recorded))` hours of motion data, with the entire dataset totaling `r apa_num(sum(coverage$hours_recorded))` hours.
+Even the shortest full recording (`r apa_num(min(coverage$hours_recorded))` hours) exceeded the longest observational sessions in past work in which an experimenter operated a camera.
+
+Considering that body position annotation takes 2-5 hours of labor per 1 hour of behavior, the `r round(sum(coverage$hours_recorded))` hours of data we recorded would have taken `r round(sum(coverage$hours_recorded)*2)`-`r round(sum(coverage$hours_recorded)*5)` hours of labor to annotate from video.
+However, only an estimated `r round(nrow(coverage)*2)`-`r round(nrow(coverage)*5)` hours of labor was needed to annotate the initial hour of each session's video that was used to train the machine learning models---an immense savings in the human cost of annotation that allows the method to scale to larger sample sizes and recording durations.
+<!-- CITES -->
 
@@
 
+### Did Full-Day Estimates of Body Position Reveal Age Differences?
+
 # Discussion

### 2023-03-10 manuscript.tex
more results

@@
 The first two sets of results show that accuracy was consistently high across the proximal and distal recordings, providing confidence that predictions over the remainder of the day would continue to be accurate after the video recordings stopped.
-
 This leads to two final questions---how successfully did recordings capture infants' entire day, and how suitable were day-long recordings for capturing individual differences in body position?
+
+\hypertarget{how-well-did-recordings-capture-infants-entire-day}{%
+\subsubsection{How Well did Recordings Capture Infants' Entire Day?}\label{how-well-did-recordings-capture-infants-entire-day}}
+
 Figure \ref{fig:timelines} depicts body position timelines across the day for each session, divided into younger (A) and older infants (B).
@@
 Nap times reported during the recording period ranged from 0.00-4.99 hours with a median of 2.22 hours.
-Of the remaining awake time during the recording period, body position data was collecting during proportions that ranged from 0.18-1.00 hours with a median of 1.00.
+Body position data was available to describe 0.18-1.00\% of the awake portion of the recording period (median = 1.00\%).
 Younger infant \#13 had the least portion of the waking day accounted for; the family left the house to run errands during the majority of the recording period, and the infant napped during much of the remaining time at home.
 Overall, the recordings produced a median of 6.15 hours of motion data, with the entire dataset totaling 201.06 hours.
+Even the shortest full recording (3.06 hours) exceeded the longest observational sessions in past work in which an experimenter operated a camera.
+
+Considering that body position annotation takes 2-5 hours of labor per 1 hour of behavior, the 201 hours of data we recorded would have taken 402-1005 hours of labor to annotate from video.
+However, only an estimated 62-155 hours of labor was needed to annotate the initial hour of each session's video that was used to train the machine learning models---an immense savings in the human cost of annotation that allows the method to scale to larger sample sizes and recording durations.
 
@@
 
+\hypertarget{did-full-day-estimates-of-body-position-reveal-age-differences}{%
+\subsubsection{Did Full-Day Estimates of Body Position Reveal Age Differences?}\label{did-full-day-estimates-of-body-position-reveal-age-differences}}
+
 \hypertarget{discussion}{%

### 2023-03-10 manuscript.Rmd
correlations

@@
             select(group, position, mean, sd))
+
+pos_sum_lookup <- pos_sum_age %>% 
+  map2_dfr(c("Group", "Individual"), 
+                         ~ mutate(.x, Model = .y) %>%
+                           select(-sd) %>% 
+                           pivot_wider(names_from = group, values_from = mean)) %>% 
+  pivot_wider(names_from = Model, values_from = c("Younger", "Older")) %>% 
+  column_to_rownames("position")
+
 pos_sum_age <- pos_sum_age %>% map2_dfr(c("Group", "Individual"), 
@@
   relocate(Younger_Individual, .after = Older_Group)
+
 apa_table(pos_sum_age, col.names = c("Position", "Younger", "Older", "Younger", "Older"), col_spanners = list("Group" = c(2, 3), "Individual" = c(4, 5)),
@@
 
-### Did Full-Day Estimates of Body Position Reveal Age Differences?
+```{r pos-correlations}
+pos_sum_ppt <- map(fullday, 
+                   ~ group_by(.x, age, id_uni) %>% 
+                     filter(is.na(nap_time)) %>% 
+                     summarize(across(sit_time:upright_time, \(x) mean(x, na.rm = TRUE))) %>% 
+                     ungroup %>% select(-id_uni))
+
+age_cor_group <- cor_mat(pos_sum_ppt$group)
+age_cor_group_pval <- age_cor_group %>% cor_get_pval() %>% column_to_rownames("rowname")
+age_cor_group <- age_cor_group %>% column_to_rownames("rowname")
+age_cor_ind <- cor_mat(pos_sum_ppt$split)
+age_cor_ind_pval <- age_cor_ind %>% cor_get_pval() %>% column_to_rownames("rowname")
+age_cor_ind <- age_cor_ind %>% column_to_rownames("rowname")
+```
+
+### Can Full-Day Estimates of Body Position Reveal Age Differences?
+
+In the last analysis, we demonstrate how the full-day recordings can reveal individual differences in body position according to age.
+Past work using sparse EMA surveys across the day [@Survey] found that from 3-12 months, infants spent less time held, reclined, and supine but spent increasingly more time sitting and upright.
+Prone time changed little.
+Table \@ref(tab:agetable) summarizes the percent of time that younger and older infants spent in each body position out of their awake samples, as predicted by group and individual models.
+
+The most straightforward comparisons between the two investigations are upright and prone, because they were defined identically.
+In @Survey, upright time increased from 0.6-5.5% of the time for infants 3-6 months, increasing to 22.0% at 12 months.
+Similarly, younger infants' sitting time estimated from full-day group models was `r pos_sum_lookup["Upright", "Younger_Group"]`% compared with `r pos_sum_lookup["Upright", "Older_Group"]`% for the older group.
+In @Survey, prone time increased ranged from 2.9-9.2% of the time for infants 3-6 months, and was 7.2% at 12 months.
+In the current study, younger infants' prone time was was `r pos_sum_lookup["Prone", "Younger_Group"]`% and `r pos_sum_lookup["Prone", "Older_Group"]`% for older infants.
+Held, sitting, and supine are more difficult to compare because prior work [@Survey] considered held any time that infants were in caregivers' arms or on their laps, however, we only counted laying or sitting on caregivers' laps as supine and sitting, respectively.
+Further complicating the comparison, @Survey included a reclined category that we would classify as either supine or sitting depending on the specifics.
+Although the numbers are less directly comparable, our results are similar in showing that held and supine time was less for older infants than younger infants.
+Sitting was greater for older infants (`r pos_sum_lookup["Sitting", "Older_Group"]`%) compared with younger infants (`r pos_sum_lookup["Sitting", "Younger_Group"]`%).
+Moreover, sitting was the most frequent body position followed by upright in 12-month-olds [@Survey], just as we observed in the current study.
+Overall, the estimates from the current investigation are consistent with age differences found using a different methodology.
+
+Figure \@ref(fig:age) shows that both group and individual model predictions captured age-related differences in body position.
+Age in months and upright time were significantly, positively correlated using group (*r* = `r age_cor_group["age","upright_time"]`, *p* `r apa_p(age_cor_group_pval["age","upright_time"])`)
 

### 2023-03-10 manuscript.tex
correlations

@@
 
-\hypertarget{did-full-day-estimates-of-body-position-reveal-age-differences}{%
-\subsubsection{Did Full-Day Estimates of Body Position Reveal Age Differences?}\label{did-full-day-estimates-of-body-position-reveal-age-differences}}
+\hypertarget{can-full-day-estimates-of-body-position-reveal-age-differences}{%
+\subsubsection{Can Full-Day Estimates of Body Position Reveal Age Differences?}\label{can-full-day-estimates-of-body-position-reveal-age-differences}}
+
+In the last analysis, we demonstrate how the full-day recordings can reveal individual differences in body position according to age.
+Past work using sparse EMA surveys across the day (Franchak, 2019) found that from 3-12 months, infants spent less time held, reclined, and supine but spent increasingly more time sitting and upright.
+Prone time changed little.
+Table \ref{tab:agetable} summarizes the percent of time that younger and older infants spent in each body position out of their awake samples, as predicted by group and individual models.
+
+The most straightforward comparisons between the two investigations are upright and prone, because they were defined identically.
+In Franchak (2019), upright time increased from 0.6-5.5\% of the time for infants 3-6 months, increasing to 22.0\% at 12 months.
+Similarly, younger infants' sitting time estimated from full-day group models was 7.68\% compared with 18.57\% for the older group.
+In Franchak (2019), prone time increased ranged from 2.9-9.2\% of the time for infants 3-6 months, and was 7.2\% at 12 months.
+In the current study, younger infants' prone time was was 14.38\% and 14.45\% for older infants.
+Held, sitting, and supine are more difficult to compare because prior work (Franchak, 2019) considered held any time that infants were in caregivers' arms or on their laps, however, we only counted laying or sitting on caregivers' laps as supine and sitting, respectively.
+Further complicating the comparison, Franchak (2019) included a reclined category that we would classify as either supine or sitting depending on the specifics.
+Although the numbers are less directly comparable, our results are similar in showing that held and supine time was less for older infants than younger infants.
+Sitting was greater for older infants (44.42\%) compared with younger infants (24.92\%).
+Moreover, sitting was the most frequent body position followed by upright in 12-month-olds (Franchak, 2019), just as we observed in the current study.
+Overall, the estimates from the current investigation are consistent with age differences found using a different methodology.
+
+Figure \ref{fig:age} shows that both group and individual model predictions captured age-related differences in body position.
+Age in months and upright time were significantly, positively correlated using group (\emph{r} = 0.59, \emph{p} \textless{} .001)
 

### 2023-03-10 manuscript.Rmd
finished results

@@
 
-```{r pos-correlations}
+```{r pos-correlations, cache=TRUE, cache.extra=tools::md5sum("data/imu-compiled.csv")}
 pos_sum_ppt <- map(fullday, 
@@
 Figure \@ref(fig:age) shows that both group and individual model predictions captured age-related differences in body position.
-Age in months and upright time were significantly, positively correlated using group (*r* = `r age_cor_group["age","upright_time"]`, *p* `r apa_p(age_cor_group_pval["age","upright_time"])`)
+Age in months and upright time were significantly, positively correlated using group (*r* = `r age_cor_group["age","upright_time"]`, *p* `r apa_p(age_cor_group_pval["age","upright_time"], add_equals = T)`) and individual models (*r* = `r age_cor_ind["age","upright_time"]`, *p* `r apa_p(age_cor_ind_pval["age","upright_time"], add_equals = T)`).
+Similarly, age predicted increases in sitting time using group (*r* = `r age_cor_group["age","sit_time"]`, *p* `r apa_p(age_cor_group_pval["age","sit_time"], add_equals = T)`) and individual models (*r* = `r age_cor_ind["age","sit_time"]`, *p* `r apa_p(age_cor_ind_pval["age","sit_time"], add_equals = T)`).
+In contrast, age predicted decreases in supine time using group (*r* = `r age_cor_group["age","supine_time"]`, *p* `r apa_p(age_cor_group_pval["age","supine_time"], add_equals = T)`) and individual models (*r* = `r age_cor_ind["age","supine_time"]`, *p* `r apa_p(age_cor_ind_pval["age","supine_time"], add_equals = T)`).
+No clear age trends were found for prone time.
+Age was not significantly related to held time for group models (*r* = `r age_cor_group["age","held_time"]`, *p* `r apa_p(age_cor_group_pval["age","held_time"], add_equals = T)`), but individual models showed a significant negative correlation (*r* = `r age_cor_ind["age","held_time"]`, *p* `r apa_p(age_cor_ind_pval["age","held_time"], add_equals = T)`) resulting from an outlier in the younger group.
 

### 2023-03-10 manuscript.tex
finished results

@@
 Figure \ref{fig:age} shows that both group and individual model predictions captured age-related differences in body position.
-Age in months and upright time were significantly, positively correlated using group (\emph{r} = 0.59, \emph{p} \textless{} .001)
+Age in months and upright time were significantly, positively correlated using group (\emph{r} = 0.59, \emph{p} \textless{} .001) and individual models (\emph{r} = 0.50, \emph{p} = .005).
+Similarly, age predicted increases in sitting time using group (\emph{r} = 0.70, \emph{p} \textless{} .001) and individual models (\emph{r} = 0.72, \emph{p} \textless{} .001).
+In contrast, age predicted decreases in supine time using group (\emph{r} = -0.66, \emph{p} \textless{} .001) and individual models (\emph{r} = -0.61, \emph{p} \textless{} .001).
+No clear age trends were found for prone time.
+Age was not significantly related to held time for group models (\emph{r} = -0.22, \emph{p} = .212), but individual models showed a significant negative correlation (\emph{r} = -0.42, \emph{p} = .021) resulting from an outlier in the younger group.
 

### 2023-03-14 manuscript.Rmd
even more NA fixes

@@
   ds$id_uni <- as.numeric(factor(ds$id_uni))
+  ds <- ds %>% mutate(prop = na_if(prop, 0))
   lims <- as_hms(c('09:00:00', '21:59:00'))

### 2023-03-14 manuscript.tex
even more NA fixes

@@
 Upright & 7.7\% (9.3) & 18.6\% (7.4) & 10.4\% (13.6) & 18.7\% (8.4)\\
-Sitting & 24.9\% (11.5) & 44.4\% (10.1) & 18.8\% (16.3) & 46.9\% (13.3)\\
-Prone & 14.4\% (13.8) & 14.4\% (6.0) & 12.5\% (10.2) & 16.9\% (10.6)\\
+Sitting & 25.4\% (12.4) & 44.4\% (10.1) & 18.8\% (16.3) & 46.9\% (13.3)\\
+Prone & 14.5\% (13.8) & 14.4\% (6.0) & 12.5\% (10.2) & 16.9\% (10.6)\\
 Supine & 38.5\% (24.0) & 14.0\% (8.4) & 40.2\% (31.6) & 10.0\% (9.5)\\
-Held & 12.7\% (7.1) & 8.5\% (5.4) & 17.1\% (20.8) & 7.4\% (7.6)\\
+Held & 13.1\% (7.4) & 8.5\% (5.4) & 17.1\% (20.8) & 7.4\% (7.6)\\
 \bottomrule
@@
 In Franchak (2019), upright time increased from 0.6-5.5\% of the time for infants 3-6 months, increasing to 22.0\% at 12 months.
-Similarly, younger infants' sitting time estimated from full-day group models was 7.68\% compared with 18.57\% for the older group.
+Similarly, younger infants' sitting time estimated from full-day group models was 7.72\% compared with 18.57\% for the older group.
 In Franchak (2019), prone time increased ranged from 2.9-9.2\% of the time for infants 3-6 months, and was 7.2\% at 12 months.
-In the current study, younger infants' prone time was was 14.38\% and 14.45\% for older infants.
+In the current study, younger infants' prone time was was 14.46\% and 14.45\% for older infants.
 Held, sitting, and supine are more difficult to compare because prior work (Franchak, 2019) considered held any time that infants were in caregivers' arms or on their laps, however, we only counted laying or sitting on caregivers' laps as supine and sitting, respectively.
@@
 Although the numbers are less directly comparable, our results are similar in showing that held and supine time was less for older infants than younger infants.
-Sitting was greater for older infants (44.42\%) compared with younger infants (24.92\%).
+Sitting was greater for older infants (44.42\%) compared with younger infants (25.37\%).
 Moreover, sitting was the most frequent body position followed by upright in 12-month-olds (Franchak, 2019), just as we observed in the current study.
@@
 Age in months and upright time were significantly, positively correlated using group (\emph{r} = 0.59, \emph{p} \textless{} .001) and individual models (\emph{r} = 0.50, \emph{p} = .005).
-Similarly, age predicted increases in sitting time using group (\emph{r} = 0.70, \emph{p} \textless{} .001) and individual models (\emph{r} = 0.72, \emph{p} \textless{} .001).
+Similarly, age predicted increases in sitting time using group (\emph{r} = 0.68, \emph{p} \textless{} .001) and individual models (\emph{r} = 0.72, \emph{p} \textless{} .001).
 In contrast, age predicted decreases in supine time using group (\emph{r} = -0.66, \emph{p} \textless{} .001) and individual models (\emph{r} = -0.61, \emph{p} \textless{} .001).
 No clear age trends were found for prone time.
-Age was not significantly related to held time for group models (\emph{r} = -0.22, \emph{p} = .212), but individual models showed a significant negative correlation (\emph{r} = -0.42, \emph{p} = .021) resulting from an outlier in the younger group.
+Age was not significantly related to held time for group models (\emph{r} = -0.23, \emph{p} = .200), but individual models showed a significant negative correlation (\emph{r} = -0.42, \emph{p} = .021) resulting from an outlier in the younger group.
 

### 2023-03-14 manuscript.Rmd
consistent factor orders

@@
   filter(nap_period == 0 & exclude_period == 0 & !is.na(code) & !is.na(pos)) %>% 
-         mutate(pos = factor(pos, levels = pos_levels)) %>% 
+         mutate(pos = factor(pos, levels = pos_levels),
+                code = factor(code, levels = pos_levels)) %>% 
   filter(file != "104/1")
@@
   agree_group <- complete(agree_group, file, pos, fill = list(prop_human = 0, prop_model = 0))
+
 }
@@
   scale_y_continuous(breaks = c(0,.5,1), labels = c(0, 50, 100), limits = c(-.07, 1.07), name = "Predicted Prevalence (%)") + 
-  facet_wrap("pos", ncol = 2, as.table = F) +
+  facet_wrap("pos", ncol = 2, as.table = T) +
   theme(panel.spacing.x = unit(1, "lines")) +
@@
   ds %>% pivot_longer(cols = sit_time:upright_time, names_to = "position", values_to = "prop") %>% 
-  mutate(position = factor(position,
-                           levels = c("upright_time", "sit_time", "prone_time", "supine_time", "held_time"),
-                           labels = c("Upright", "Sitting", "Prone", "Supine", "Held"))) %>% 
+  mutate(position = factor(position, levels = c("supine_time", "prone_time", "sit_time", "upright_time", "held_time"), labels = c("Supine", "Prone", "Sitting", "Upright", "Held"))) %>% 
   arrange(position, age)

### 2023-03-14 manuscript.tex
consistent factor orders

@@
 \midrule
-Held & 0.02 & 0.04 & 0.73 & 0.60\\
+Supine & 0.88 & 0.98 & 0.94 & 0.97\\
 Prone & 0.97 & 0.86 & 0.97 & 0.84\\
 Sitting & 0.79 & 0.97 & 0.91 & 0.95\\
-Supine & 0.88 & 0.98 & 0.94 & 0.97\\
-Upright & 0.63 & 0.83 & 0.99 & 0.95\\ \midrule
+Upright & 0.63 & 0.83 & 0.99 & 0.95\\
+Held & 0.02 & 0.04 & 0.73 & 0.60\\ \midrule
 Overall & 0.80 & 0.91 & 0.95 & 0.96\\
@@
 \midrule
-Held & 0.51 & 0.46 & 0.67 & 0.63\\
+Supine & 0.76 & 0.96 & 0.88 & 0.93\\
 Prone & 0.96 & 0.90 & 0.96 & 0.89\\
 Sitting & 0.72 & 0.93 & 0.89 & 0.92\\
-Supine & 0.76 & 0.96 & 0.88 & 0.93\\
-Upright & 0.91 & 0.93 & 0.98 & 0.96\\ \midrule
+Upright & 0.91 & 0.93 & 0.98 & 0.96\\
+Held & 0.51 & 0.46 & 0.67 & 0.63\\ \midrule
 Overall & 0.80 & 0.94 & 0.92 & 0.94\\
@@
 \midrule
+Supine & 38.5\% (24.0) & 14.0\% (8.4) & 40.2\% (31.6) & 10.0\% (9.5)\\
+Prone & 14.5\% (13.8) & 14.4\% (6.0) & 12.5\% (10.2) & 16.9\% (10.6)\\
+Sitting & 25.4\% (12.4) & 44.4\% (10.1) & 18.8\% (16.3) & 46.9\% (13.3)\\
 Upright & 7.7\% (9.3) & 18.6\% (7.4) & 10.4\% (13.6) & 18.7\% (8.4)\\
-Sitting & 25.4\% (12.4) & 44.4\% (10.1) & 18.8\% (16.3) & 46.9\% (13.3)\\
-Prone & 14.5\% (13.8) & 14.4\% (6.0) & 12.5\% (10.2) & 16.9\% (10.6)\\
-Supine & 38.5\% (24.0) & 14.0\% (8.4) & 40.2\% (31.6) & 10.0\% (9.5)\\
 Held & 13.1\% (7.4) & 8.5\% (5.4) & 17.1\% (20.8) & 7.4\% (7.6)\\

### 2023-03-14 manuscript.Rmd
edited intro

@@
 Infants struggle to manipulate objects while supine and prone, but sitting affords object exploration [@SoskaAdolph2014].
-Upright walking changes how infants and caregivers interact compared with crawling in a prone position; while walking infants move farther away, share toys in different ways, and hear different language from caregivers [@ChenSchneider2022; @Karasik2011; @WestIverson2021; @Karasik2014].
-As infants grow older and acquire new abilities, such as independent sitting and walking, the spend more time sitting and upright and less time held, supine, and prone [@Survey; @Thurman2017; @Freeplay; @AdolphCTL2014].
+Upright walking changes how infants and caregivers interact; while walking infants move farther away from caregivers and share toys in different ways compared to when crawling in a prone position [@ChenSchneider2022; @Karasik2011].
+As infants grow older and acquire new abilities, such as independent sitting and walking, they spend more time sitting and upright and less time held, supine, and prone [@Survey; @Thurman2017; @Freeplay; @AdolphCTL2014].
+Caregivers provide different learning input to infants depending on infants' body position [@KretchKoziol2022; @WestFletcher2022; @WestIverson2021; @Karasik2014].
 Thus, characterizing individual differences in the day-to-day accumulation of body position experiences informs developmental theory by revealing differential opportunities for learning [@CurrentOpinion].
-<!-- Other west cites; cite new EMA; KretchKoziol2022 -->
 
-In this paper, we present an inertial sensing method to classify infant body position from moment-to-moment across an entire day, and validate its accuracy using over 100 hours of video recorded across 35 in-home data collection sessions.
+In this paper, we present an inertial sensing method to classify infant body position from moment-to-moment across an entire day, and validate its accuracy using over 100 hours of video recorded across 34 in-home data collection sessions.
 Our method takes inspiration from a more mature technology: Long-form audio recordings of infants' language experiences using wearable audio recorders.
@@
 
-The LENA&reg; recorder is a commercial device that is worn by infants in a custom shirt pocket that has sufficient battery life and storage to record for an entire day.
-Closed-source LENA&reg; algorithms analyze the audio recordings to provide automatic counts of useful metrics, such as the number of words spoken by adults in the vicinity of the participant.
+The LENA&reg; recorder is a commercial device that is worn by infants in a custom shirt pocket; the recorder has sufficient battery life and storage to record for an entire day.
+Closed-source LENA&reg; algorithms analyze the audio recordings to provide automatic counts of useful metrics, such as the number of words spoken by adults in the vicinity of the infant.
 Other long-form audio methods rely on custom-built recorders [@WassPhillips2022], apply alternative classification algorithms to LENA&reg; data [@RasanenSeshadri2020; @MichelettiYao2022], or manually transcribe audio recorded by LENA&reg; devices to improve accuracy or identify behaviors beyond the built-in categories [@Bergelson2019Corpus; @MendozaFausey2021].
@@
 Long-form audio recordings have had a transformational impact on language development research by allowing researchers to characterize opportunities for learning in daily life.
-Measuring the amount of speech heard by infants in the home [@Weisleder2013] or in a daycare setting [@PerryPrince2018] revealed individual differences input that predict later vocabulary.
+For example, measuring the amount of speech heard by infants in the home [@Weisleder2013] or in a daycare setting [@PerryPrince2018] revealed how individual differences in speech input predict later vocabulary.
 Full-day language recording synchronized with other data sources allows researchers to identify how auditory input and vocal production interact with other processes.
 Beyond individual differences in aggregated data, long-form recordings can be used to determine the temporal schedule of experiences.
-For example, infants' daily experiences hearing music are clustered in time, with "bursty" episodes of hearing music separated by relatively long periods during which music is absent [@MendozaFausey2022].
+For example, infants' daily experiences hearing music are clustered in time, with "bursty" episodes of music presence separated by relatively long periods during which music is absent [@MendozaFausey2022].
 Synchronizing audio recordings with other data sources extends researchers' ability to characterize daily experiences.
-Linking LENA&reg; speech measurements to repeated, time-stamped text-message surveys about infant device placement revealed that infants heard less caregiver speech during moments that they were restrained in devices such as swings, exersaucers, and car seats [@MalachowskiSalo2023].
+Linking LENA&reg; speech measurements to repeated, time-stamped text-message surveys about infant device placement revealed that infants heard less caregiver speech during periods that they were restrained in devices such as swings, exersaucers, and car seats [@MalachowskiSalo2023].
 A custom-built wearable ECG and audio recorder allowed @WassPhillips2022 to discover that infant arousal increases the likelihood of infant vocalization across the day.
@@
 Synchronizing real-time data to other data streams helps to reveal sources of variability within an individual [e.g., @WassPhillips2022].
-Full-day recordings are essential for capturing experiences across the heterogeniety of daily routines that moderate behavior (e.g., play, feeding, errands) [@KadookaCaufield2021; @Tamis-LeMondaCustode2018].
+Full-day recordings are essential for capturing experiences across the heterogeneity of daily routines that moderate behavior (e.g., play, feeding, errands) [@KadookaCaufield2021; @Tamis-LeMondaCustode2018].
 "Burstiness" of behavior means that long recordings are needed to capture clusters of events amid long periods in which they may be absent [@BarbaroFausey2022; @WarlaumontSobowale2021].
 Fifth, *automatic classification* means that the approach can scale to analyze large numbers of participants over long recordings without the bottleneck of manual annotation/transcription.
-Automatic classification can only replace human annotation if it is sufficiently accurate and unbiased.
+But automatic classification can only replace human annotation if it is sufficiently accurate and unbiased.
 An independent assessment of the LENA&reg; algorithms found mixed results about the accuracy of different outcomes.
@@
 However, poor agreement was found for other metrics, such as the number of "conversational turns" between the child and communicative partners and the identification of male speakers.
-<!-- EVENTUALLY CITE EMA HERE -->
+<!-- EVENTUALLY CITE NEW EMA HERE -->
 
 Thus, for some use cases (and for some metrics), long form audio recordings provide a mobile, unobtrusive way to automatically score real-time data over a full day.
-In the remainder of the paper, we turn to the question of how to replicate these qualities in long form recordings of motor behavior.
+In the remainder of the paper, we turn to the question of how to replicate those key features in long form recordings of motor behavior.
 
@@
 Video and survey methods are the current state-of-the-art in assessing motor behavior.
-Although, each method has  advantages and disadvantages for characterizing infants' everyday motor experiences that complements the other, neither method on its own can provide comparable data to the long form audio recordings reviewed in the previous section.
+Although each method has complementary advantages and disadvantages for characterizing infants' everyday motor experiences, neither method on its own can provide comparable data to the long form audio recordings reviewed in the previous section.
 
@@
 Logistically, an experimenter cannot follow behind infants to record their behavior from morning to night (and were they to do so, they would likely alter infants' and caregivers' behavior).
-Typical video recording sessions last 45-120 minutes [@Karasik2011; @ChenSchneider2022; @HerzbergFletcher2021], short of capturing the variety of activities across the full daily routine.
+Typical video recording sessions last 45-120 minutes [@Karasik2011; @ChenSchneider2022; @HerzbergFletcher2021], far short of capturing the variety of activities across the full daily routine.
 Even if full day videos were available, the lack of suitable automatic classification tools means that the human cost of annotation makes it difficult for video methods to scale to testing large numbers of participants.
@@
 Many commercially-available and inexpensive IMUs have > 12 hour battery life with onboard storage to record *real-time*, *full-day* motion data at a sampling rate that is higher than typical video (e.g., 50-100 Hz).
-Moreover, past work has successfully recorded the rate of leg kicks [e.g., @DengTrujillo-Priego2019] and activity intensity [@SchnellerBentsen2017] in infants and children across multiple days.
+Indeed, past work has successfully recorded the rate of leg kicks [e.g., @DengTrujillo-Priego2019] and activity intensity [@SchnellerBentsen2017] in infants and children across multiple days.
 
-The open question is whether *automatic classification* is sufficiently accurate to measure movement categories that are relevant to developmental and clinical research, and whether measurement validity is acceptable over long recording periods.
+The open question is whether *automatic classification* is sufficiently accurate to measure movement categories that are relevant to developmental and clinical research, and whether measurements continue to be accurate over long recording periods.
 IMUs typically contain accelerometers that measure linear acceleration paired with gyroscopes that measure angular acceleration.
@@
 Finally, caregivers frequently pick up and transport infants, creating motion signals that need to be differentiated from independent activity [@KwonZavos2019; @PatelShi2019].
+Although @GreenspanCunha2021 were able to use cut-point definitions about infant trunk angle to categorize body positions, this method was not able to differentiate times that infants were held by caregivers off the ground.
 Thus, modern approaches to human activity recognition have used machine learning to classify activity categories based on features derived from IMU data in adults [@Preece2009; @Arif2015], children [@NamPark2013; @RenDing2016; @StewartNarayanan2018], and infants [@YaoPlotz2019; @FranchakScott2021; @AiraksinenRasanen2020].
@@
 Crucially, this allows automatic classification to scale to full day recording by relying on a relatively smaller set of video annotation.
-
-<!-- NEED TO FIT IN GREENSPAN posture from cut points, Hewitt Tummy time, Yao holding-->
 
@@
 Although all three studies yielded promising classification accuracy, accuracy was assessed in brief (15-60 minute) sessions supervised by a researcher, leaving the open question of how well body position classification will scale to testing across an entire day of natural home life.
-
-<!-- Could have a bit more detail in this last paragraph -->
 

### 2023-03-14 manuscript.Rmd
current study edits

@@
 
-Accordingly, the overarching goal of the current study is to test the feasibility and validity of long-form body position recording in the home during unsupervised, everyday behavior.
+Accordingly, the overarching goal of the current study is to test the validity of long-form body position recording in the home during unsupervised, everyday behavior.
 Supervised recordings from past work [@FranchakScott2021; @AiraksinenRasanen2020; @AiraksinenGallen2022], whether in the home or in the lab, let researchers set up the situation to encourage or restrict certain behaviors.
-Usually, caregivers are asked to play with the infant.
-However, across a real day, non-play activities (e.g., eating lunch in a high chair) create challenging situations for applying automated classification of body position.
+Usually, caregivers were asked to play with the infant.
+However, in a real day, non-play activities (e.g., eating lunch in a high chair) create challenging situations for applying automated classification of body position.
 Will models trained on video-recorded observations at the beginning of the day generalize to predict behavior at a later time?
@@
 
-In the current study, we report the feasibility and validity of body position classification over the full day in the home based on 35 testing sessions from 22 infants aged 4-14 months.
+In the current study, we report the feasibility and validity of body position classification over the full day in the home based on 34 testing sessions from 22 infants aged 4-14 months.
 Participants received a custom pair of infant leggings embedded with 4 IMUs (one on each ankle and one on each hip) and a video camera to collect ground truth data about infant body position.
@@
 The **first goal** of the current study was to determine the accuracy of body position classification during the proximal comparison period using this novel, semi-supervised procedure in participants' homes.
-Past work found varying better performance using "individual models"---models that were trained on one participant's data to predict their later behavior---compared with a "group model" that aggregated data from all infants to create a single body position classifier [@FranchakScott2021], so we compared both modeling approaches in the current investigation.
+Past work found better performance using "individual models"---models that were trained on one participant's data to predict their later behavior---compared with a "group model" that aggregated data from all infants to create a single body position classifier [@FranchakScott2021], so we compared both modeling approaches in the current investigation.
 
+The crucial test was how well models predicted later behavior over longer recordings of everyday activities.
 A second, *distal comparison* period followed the proximal comparison period and captured approximately 90 minutes of home behavior that was completely unsupervised.
@@
 Because this recording happened a considerable amount of time after the initial setup and instructions from the experimenter, accuracy could decline if caregivers or infants moved the garment or sensors.
-Moreover, increasing variation in everyday activities during the distal comparison creates a greater challenge for classification, testing whether machine learning models can generalize to novel test cases.
+Moreover, increasing variation in everyday activities during the distal comparison creates a greater challenge, testing whether body position classification models can generalize to novel test cases.
 Thus, the **second goal** of our study was to assess accuracy during the distal comparison.
 
-After this second video recording period, we asked caregivers to keep infants wearing IMUs for the rest of the day until their regular bedtime, creating the **first real-time, full-day dataset of infant body position**.
+After the distal comparison period when video recording ceased, we asked caregivers to have infants wear the IMUs for the rest of the day until their regular bedtime, creating the **first real-time, full-day dataset of infant body position**.
 Interpreting such data required caregivers to log when infants napped, when they removed the sensor garment for diaper changes or other reasons, and when infants went to bed at the end of the day.
-Finally, the **third goal** of the study was to examine the quality of the full-day data.
+Thus, the **third goal** of the study was to examine the quality of the full-day data.
 Could body position be successfully recorded over the desired period?

### 2023-03-14 manuscript.tex
current study edits

@@
 Infants struggle to manipulate objects while supine and prone, but sitting affords object exploration (Soska \& Adolph, 2014).
-Upright walking changes how infants and caregivers interact compared with crawling in a prone position; while walking infants move farther away, share toys in different ways, and hear different language from caregivers (Chen et al., 2022; Karasik et al., 2011, 2014; West \& Iverson, 2021).
-As infants grow older and acquire new abilities, such as independent sitting and walking, the spend more time sitting and upright and less time held, supine, and prone (Adolph \& Tamis-LeMonda, 2014; Franchak et al., 2018; Franchak, 2019; Thurman \& Corbetta, 2017).
+Upright walking changes how infants and caregivers interact; while walking infants move farther away from caregivers and share toys in different ways compared to when crawling in a prone position (Chen et al., 2022; Karasik et al., 2011).
+As infants grow older and acquire new abilities, such as independent sitting and walking, they spend more time sitting and upright and less time held, supine, and prone (Adolph \& Tamis-LeMonda, 2014; Franchak et al., 2018; Franchak, 2019; Thurman \& Corbetta, 2017).
+Caregivers provide different learning input to infants depending on infants' body position (Karasik et al., 2014; Kretch et al., 2022; West et al., 2022; West \& Iverson, 2021).
 Thus, characterizing individual differences in the day-to-day accumulation of body position experiences informs developmental theory by revealing differential opportunities for learning (Franchak, 2020).
 
-In this paper, we present an inertial sensing method to classify infant body position from moment-to-moment across an entire day, and validate its accuracy using over 100 hours of video recorded across 35 in-home data collection sessions.
+In this paper, we present an inertial sensing method to classify infant body position from moment-to-moment across an entire day, and validate its accuracy using over 100 hours of video recorded across 34 in-home data collection sessions.
 Our method takes inspiration from a more mature technology: Long-form audio recordings of infants' language experiences using wearable audio recorders.
@@
 
-The LENA® recorder is a commercial device that is worn by infants in a custom shirt pocket that has sufficient battery life and storage to record for an entire day.
-Closed-source LENA® algorithms analyze the audio recordings to provide automatic counts of useful metrics, such as the number of words spoken by adults in the vicinity of the participant.
+The LENA® recorder is a commercial device that is worn by infants in a custom shirt pocket; the recorder has sufficient battery life and storage to record for an entire day.
+Closed-source LENA® algorithms analyze the audio recordings to provide automatic counts of useful metrics, such as the number of words spoken by adults in the vicinity of the infant.
 Other long-form audio methods rely on custom-built recorders (Wass et al., 2022), apply alternative classification algorithms to LENA® data (Micheletti et al., 2022; Räsänen et al., 2020), or manually transcribe audio recorded by LENA® devices to improve accuracy or identify behaviors beyond the built-in categories (Bergelson, Casillas, et al., 2019; Mendoza \& Fausey, 2021).
@@
 Long-form audio recordings have had a transformational impact on language development research by allowing researchers to characterize opportunities for learning in daily life.
-Measuring the amount of speech heard by infants in the home (Weisleder \& Fernald, 2013) or in a daycare setting (Perry et al., 2018) revealed individual differences input that predict later vocabulary.
+For example, measuring the amount of speech heard by infants in the home (Weisleder \& Fernald, 2013) or in a daycare setting (Perry et al., 2018) revealed how individual differences in speech input predict later vocabulary.
 Full-day language recording synchronized with other data sources allows researchers to identify how auditory input and vocal production interact with other processes.
 Beyond individual differences in aggregated data, long-form recordings can be used to determine the temporal schedule of experiences.
-For example, infants' daily experiences hearing music are clustered in time, with ``bursty'' episodes of hearing music separated by relatively long periods during which music is absent (Mendoza \& Fausey, 2022).
+For example, infants' daily experiences hearing music are clustered in time, with ``bursty'' episodes of music presence separated by relatively long periods during which music is absent (Mendoza \& Fausey, 2022).
 Synchronizing audio recordings with other data sources extends researchers' ability to characterize daily experiences.
-Linking LENA® speech measurements to repeated, time-stamped text-message surveys about infant device placement revealed that infants heard less caregiver speech during moments that they were restrained in devices such as swings, exersaucers, and car seats (Malachowski et al., 2023).
+Linking LENA® speech measurements to repeated, time-stamped text-message surveys about infant device placement revealed that infants heard less caregiver speech during periods that they were restrained in devices such as swings, exersaucers, and car seats (Malachowski et al., 2023).
 A custom-built wearable ECG and audio recorder allowed Wass et al.
@@
 Synchronizing real-time data to other data streams helps to reveal sources of variability within an individual (e.g., Wass et al., 2022).
-Full-day recordings are essential for capturing experiences across the heterogeniety of daily routines that moderate behavior (e.g., play, feeding, errands) (Kadooka et al., 2021, April; Tamis-LeMonda et al., 2018).
+Full-day recordings are essential for capturing experiences across the heterogeneity of daily routines that moderate behavior (e.g., play, feeding, errands) (Kadooka et al., 2021, April; Tamis-LeMonda et al., 2018).
 ``Burstiness'' of behavior means that long recordings are needed to capture clusters of events amid long periods in which they may be absent (Barbaro \& Fausey, 2022; Warlaumont et al., 2021).
 Fifth, \emph{automatic classification} means that the approach can scale to analyze large numbers of participants over long recordings without the bottleneck of manual annotation/transcription.
-Automatic classification can only replace human annotation if it is sufficiently accurate and unbiased.
+But automatic classification can only replace human annotation if it is sufficiently accurate and unbiased.
 An independent assessment of the LENA® algorithms found mixed results about the accuracy of different outcomes.
@@
 Thus, for some use cases (and for some metrics), long form audio recordings provide a mobile, unobtrusive way to automatically score real-time data over a full day.
-In the remainder of the paper, we turn to the question of how to replicate these qualities in long form recordings of motor behavior.
+In the remainder of the paper, we turn to the question of how to replicate those key features in long form recordings of motor behavior.
 
@@
 Video and survey methods are the current state-of-the-art in assessing motor behavior.
-Although, each method has advantages and disadvantages for characterizing infants' everyday motor experiences that complements the other, neither method on its own can provide comparable data to the long form audio recordings reviewed in the previous section.
+Although each method has complementary advantages and disadvantages for characterizing infants' everyday motor experiences, neither method on its own can provide comparable data to the long form audio recordings reviewed in the previous section.
 
@@
 Logistically, an experimenter cannot follow behind infants to record their behavior from morning to night (and were they to do so, they would likely alter infants' and caregivers' behavior).
-Typical video recording sessions last 45-120 minutes (Chen et al., 2022; Herzberg et al., 2021; Karasik et al., 2011), short of capturing the variety of activities across the full daily routine.
+Typical video recording sessions last 45-120 minutes (Chen et al., 2022; Herzberg et al., 2021; Karasik et al., 2011), far short of capturing the variety of activities across the full daily routine.
 Even if full day videos were available, the lack of suitable automatic classification tools means that the human cost of annotation makes it difficult for video methods to scale to testing large numbers of participants.
@@
 Many commercially-available and inexpensive IMUs have \textgreater{} 12 hour battery life with onboard storage to record \emph{real-time}, \emph{full-day} motion data at a sampling rate that is higher than typical video (e.g., 50-100 Hz).
-Moreover, past work has successfully recorded the rate of leg kicks (e.g., Deng et al., 2019) and activity intensity (Schneller et al., 2017) in infants and children across multiple days.
-
-The open question is whether \emph{automatic classification} is sufficiently accurate to measure movement categories that are relevant to developmental and clinical research, and whether measurement validity is acceptable over long recording periods.
+Indeed, past work has successfully recorded the rate of leg kicks (e.g., Deng et al., 2019) and activity intensity (Schneller et al., 2017) in infants and children across multiple days.
+
+The open question is whether \emph{automatic classification} is sufficiently accurate to measure movement categories that are relevant to developmental and clinical research, and whether measurements continue to be accurate over long recording periods.
 IMUs typically contain accelerometers that measure linear acceleration paired with gyroscopes that measure angular acceleration.
@@
 Finally, caregivers frequently pick up and transport infants, creating motion signals that need to be differentiated from independent activity (Kwon et al., 2019; Patel et al., 2019).
+Although Greenspan et al.
+(2021) were able to use cut-point definitions about infant trunk angle to categorize body positions, this method was not able to differentiate times that infants were held by caregivers off the ground.
 Thus, modern approaches to human activity recognition have used machine learning to classify activity categories based on features derived from IMU data in adults (Arif \& Kattan, 2015; Preece et al., 2009), children (Nam \& Park, 2013; Ren et al., 2016; Stewart et al., 2018), and infants (Airaksinen et al., 2020; Franchak et al., 2021; Yao et al., 2019).
@@
 
-Accordingly, the overarching goal of the current study is to test the feasibility and validity of long-form body position recording in the home during unsupervised, everyday behavior.
+Accordingly, the overarching goal of the current study is to test the validity of long-form body position recording in the home during unsupervised, everyday behavior.
 Supervised recordings from past work (Airaksinen et al., 2022, 2020; Franchak et al., 2021), whether in the home or in the lab, let researchers set up the situation to encourage or restrict certain behaviors.
-Usually, caregivers are asked to play with the infant.
-However, across a real day, non-play activities (e.g., eating lunch in a high chair) create challenging situations for applying automated classification of body position.
+Usually, caregivers were asked to play with the infant.
+However, in a real day, non-play activities (e.g., eating lunch in a high chair) create challenging situations for applying automated classification of body position.
 Will models trained on video-recorded observations at the beginning of the day generalize to predict behavior at a later time?
@@
 The \textbf{first goal} of the current study was to determine the accuracy of body position classification during the proximal comparison period using this novel, semi-supervised procedure in participants' homes.
-Past work found varying better performance using ``individual models''---models that were trained on one participant's data to predict their later behavior---compared with a ``group model'' that aggregated data from all infants to create a single body position classifier (Franchak et al., 2021), so we compared both modeling approaches in the current investigation.
-
+Past work found better performance using ``individual models''---models that were trained on one participant's data to predict their later behavior---compared with a ``group model'' that aggregated data from all infants to create a single body position classifier (Franchak et al., 2021), so we compared both modeling approaches in the current investigation.
+
+The crucial test was how well models predicted later behavior over longer recordings of everyday activities.
 A second, \emph{distal comparison} period followed the proximal comparison period and captured approximately 90 minutes of home behavior that was completely unsupervised.
@@
 Because this recording happened a considerable amount of time after the initial setup and instructions from the experimenter, accuracy could decline if caregivers or infants moved the garment or sensors.
-Moreover, increasing variation in everyday activities during the distal comparison creates a greater challenge for classification, testing whether machine learning models can generalize to novel test cases.
+Moreover, increasing variation in everyday activities during the distal comparison creates a greater challenge, testing whether body position classification models can generalize to novel test cases.
 Thus, the \textbf{second goal} of our study was to assess accuracy during the distal comparison.
 
-After this second video recording period, we asked caregivers to keep infants wearing IMUs for the rest of the day until their regular bedtime, creating the \textbf{first real-time, full-day dataset of infant body position}.
+After the distal comparison period when video recording ceased, we asked caregivers to have infants wear the IMUs for the rest of the day until their regular bedtime, creating the \textbf{first real-time, full-day dataset of infant body position}.
 Interpreting such data required caregivers to log when infants napped, when they removed the sensor garment for diaper changes or other reasons, and when infants went to bed at the end of the day.
-Finally, the \textbf{third goal} of the study was to examine the quality of the full-day data.
+Thus, the \textbf{third goal} of the study was to examine the quality of the full-day data.
 Could body position be successfully recorded over the desired period?
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-GreenspanCunha2021}{}}%
+Greenspan, B., Cunha, A.
+B., \& Lobo, M.
+A.
+(2021).
+Design and validation of a smart garment to measure positioning practices of parents with young infants.
+\emph{Infant Behavior and Development}, \emph{62}, 101530.
+
 \leavevmode\vadjust pre{\hypertarget{ref-HagerTilton2017}{}}%
@@
 \url{https://doi.org/10.1111/cdev.12206}
+
+\leavevmode\vadjust pre{\hypertarget{ref-KretchKoziol2022}{}}%
+Kretch, K.
+S., Koziol, N.
+A., Marcinowski, E.
+C., Kane, A.
+E., Inamdar, K., Brown, E.
+D., Bovaird, J.
+A., Harbourne, R.
+T., Hsu, L.-Y., Lobo, M.
+A., \& Dusing, S.
+C.
+(2022).
+Infant posture and caregiver-provided cognitive opportunities in typically developing infants and infants with motor delay.
+\emph{Developmental Psychobiology}, \emph{64}(1).
+\url{https://doi.org/10.1002/dev.22233}
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-WestFletcher2022}{}}%
+West, K.
+L., Fletcher, K.
+K., Adolph, K.
+E., \& Tamis-LeMonda, C.
+S.
+(2022).
+Mothers talk about infants' actions: How verbs correspond to infants' real-time behavior.
+\emph{Developmental Psychology}, \emph{58}(3), 405--416.
+\url{https://doi.org/10.1037/dev0001285}
+
 \leavevmode\vadjust pre{\hypertarget{ref-WestIverson2021}{}}%

### 2023-03-14 manuscript.Rmd
R citations

@@
 
-Infants were recruited in one of two age groups: *Younger* infants were between 4 and 7 months and *Older* infants were between 11 and 14 months.
+Infants were recruited in one of two age groups: *Younger* infants were between 4 and 7 months and *older* infants were between 11 and 14 months.
 There were `r num_younger` infants in the 4-7 month group (X female) and `r num_older` in the 11-14 month group (X female).
 Families were recruited through social media advertisements and from community events in Southern California.
-The ethnicity for infants was reported as X Hispanix/Latinx and X for not Hispanix/Latinx.
+Caregivers reported the ethnicity of infants as X Hispanix/Latinx and X for not Hispanix/Latinx.
 Race was reported as X Caucasian, X Asian, X ~~~ and X others.
-Parents were compensated \$30 in cash for each visit they completed.
+Families were compensated \$30 for every home recording session they completed.
 The BLINDED Institutional Review Board reviewed and approved all procedures associated with the study.
-All participants gave their informed consent to participate.
+All caregivers gave their informed consent before the start of the study.
 <!-- TODO RACE ETHNICITY SEX, COUNT BY HAND -->
@@
 Across the two age groups, we report data on a total of `r length(unique(ppts$id_uni))` sessions, with `r num_sessions %>% filter(group == "Younger") %>% pull(n) %>% sum` sessions from younger infants and `r num_sessions %>% filter(group == "Older") %>% pull(n) %>% sum` sessions from older infants.
-Across sessions, younger infants' age ranged from `r age["Younger","min"]` to `r age["Younger","max"]` (*M* = `r age["Younger","mean"]`) and older infants' age ranged from `r age["Older","min"]` to `r age["Older","max"]` (*M* = `r age["Older","mean"]`).
-
-<!-- with X including permission for data to be shared on Databrary.  -->
+Across sessions, younger infants' age ranged from `r apa_num(age["Younger","min"], digits = 1)` to `r apa_num(age["Younger","max"], digits = 1)` months (*M* = `r apa_num(age["Younger","mean"], digits = 1)`) and older infants' age ranged from `r apa_num(age["Older","min"], digits = 1)` to `r apa_num(age["Older","max"], digits = 1)` months (*M* = `r apa_num(age["Older","mean"], digits = 1)`) .
 
@@
 Four inertial movement units (IMUs) were used to record infant movement across the day (MC10 Biostamp).
-A custom garment was made to hold the IMUs: Internal pockets were sewn into a snug-fitting pair of infant leggings so that IMUs would stay close to the body and so that infants could not pull out the sensors.
-On each side of the body there was a pocket over the hip and a pocket just above the ankle.
-Each sensor recorded accelerometer and gyroscope data at 62.5 Hz throughout the day, with sufficient battery and on-device storage to record for approximately 12 hours.
+A custom garment was designed to hold the IMUs: Internal pockets were sewn into a snug-fitting pair of infant leggings so that IMUs would stay close to the body (reducing vibration), which also prevented infants from pulling out the sensors.
+A pocket over the hip and a pocket just above the ankle were sewn on the outside of the right and left legs of the garment.
+Each sensor recorded accelerometer and gyroscope data at 62.5 Hz throughout the day, with sufficient battery and onboard storage to record for approximately 12 hours.
 Infants also wore a LENA&reg; recorder throughout the day in the front pocket of a LENA&reg; shirt, located near the infant’s chest.
@@
 The proprietary "Boosted Battery Base" attached to the action camera to allow for a total of 3 hours of recording.
-However, as described below, this created two video files temporally separated by a gap of approximately 40-45 s.
+As described below, the video recording was not continuous.
+The camera created two video files temporally separated by a gap of approximately 40-45 s.
 Caregivers also received a log sheet to record times that infants napped as well as times that the sensor garment was removed from the infant (e.g., baths, diaper changes).
@@
 
-```{r exemplar-timeline, out.width="0.99\\linewidth", include=TRUE, fig.align="center", fig.cap="Example Timeline Caption.", echo=FALSE}
+```{r exemplar-timeline, out.width="0.99\\linewidth", include=TRUE, fig.align="center", fig.cap="Timeline from an exemplar participant (older infant 15).
+The top row shows the model-predicted body position across the entire recording period.
+Annotations indicate when the video camera was turned on by the experimenter when arriving at the house, when the sensors were first placed on the infant, when the guided call took place, and when the video files were recorded.
+Gray areas on the timeline indicate naps, and white areas indicate times when the sensors were removed.
+The bottom row shows a zoomed-in view of the video period during which ground truth data were available.
+The top timeline shows human-coded body position and the bottom row shows model-predicted body position; these were the data used for validation.
+The first part of the video period was the proximal comparison, when video and motion data were highly synchronized.
+The second part of the video recording was the distal comparison that had coarser synchronization.
+Accuracy data for the distal comparison are provided overall and during 10-minute bins, marked by vertical dashed lines.", echo=FALSE}
 knitr::include_graphics("figures/timeline.png")
@@
 On the day of the visit, a researcher arrived at the participant’s home in the morning between and prepared all the equipment at the doorstep.
-In order to synchronize all three recording devices, the researcher began by first turning on the video camera and the LENA audio device (the IMUs were already configured and placed at arrival because they required a proprietary sensor dock to begin recording in the laboratory).
+In order to synchronize all three recording devices, the researcher began by first turning on the video camera and the LENA audio device (the IMUs were already configured and placed before arrival because they required a proprietary sensor dock to begin recording in the laboratory).
 To create an easily recognizable synchronization point between the video recording and IMU data, the researcher dropped or struck the leggings (containing the IMUs) on a surface in view of the camera, as in @FranchakScott2021.
 All the equipment---once recording and with synchronization information recorded---was placed inside a large bucket and left outside the family's front door.
-
-The researcher then called the caregiver on the phone and walked them through a set of procedures needed to properly set up the equipment and record video for classifier training and testing.
+The researcher then called the caregiver on the phone and walked them through a set of procedures needed to properly set up the equipment and record video for ground truth human annotation of body position.
 At the start of this "guided call", the caregiver was instructed to place the camera in an area that captured the majority of the room.
@@
 
-Afterwards, they were to go about their day as usual with the infant wearing the sensor garment until their bedtime, only taking off the sensor for naps, baths, and diaper changes.
+Afterwards, they were asked to go about their day as usual with the infant wearing the sensor garment until their bedtime, only taking off the sensor for naps, baths, and diaper changes.
 The caregiver logged the times the sensors were removed (blank areas in the timeline in Figure \@ref(fig:exemplar-timeline))) or the child took a nap (gray areas in the timeline in Figure \@ref(fig:exemplar-timeline))) so that those times could be excluded from analysis.
-The following day a researcher picked up the equipment, verified the paperwork was signed, and compensated the participant.
+The following day a researcher picked up the equipment, verified that the paperwork was signed, and compensated the participant.
 
-Because the camera only had the battery life to record for ~3 hours (divided into two 90-minute video files), this divided the day into different periods for analysis.
+Because the camera only had the battery life to record for ~3 hours (split into two 90-minute video files), this divided the day into different periods for analysis.
 As seen in the bottom of Figure \@ref(fig:exemplar-timeline)), the *video period* comprised the first three hours of recording starting from the researcher's arrival when they turned on the camera.
 The first 90-minute video file, termed the *proximal comparison*, contained the activities during the guided call followed by a period of infants and caregivers resuming their normal activities.
-Because this video contained the synchronization point, the data in this period had temporal synchrony between IMU and video data that contained errors of no more than 30-60 ms (1-2 video frames).
+Because this video contained the synchronization point, the data in this period had high temporal synchrony between IMU and video data.
+Synchrony errors were less than 30-60 ms (1-2 video frames).
 When the first video file ended, a second video file was recorded, termed the *distal comparison*.
@@
 However, because there was a variable gap of ~40 s between the two videos, temporal synchrony in the distal comparison video was approximate containing offsets of ~ 5 s in either direction.
-<!-- HOW MANY PPTS DON"T HAVE DISTAL COMPARISON VIDEOS? -->
 
@@
 
-The proximal and distal comparison videos for each participant were annotated by trained human coders to identify infant *body position* into one of 5 mutually-exclusive categories: supine, prone, sitting, upright, or held by caregiver.
+The proximal and distal comparison videos for each participant were annotated by trained human coders to identify infant *body position* into one of 5 mutually-exclusive categories following the definitions in prior work [@FranchakScott2021]: supine, prone, sitting, upright, or held by caregiver.
 All coding was done using Datavyu software (datavyu.org).
 
-Supine was coded when the infant was lying on their back or was reclined up to a 45 degrees angle.
+Supine was coded when the infant was lying on their back or was reclined up to a 45 degree angle.
 Supine was also coded in the rare cases when the infant was laying on their side.
-Prone was coded when the infant was lying on their stomach, was on all fours in a downward dog position, or was crawling.
+Prone was coded when the infant was lying on their stomach, was stationary supported by the hands/knees or the hands/feet, or was crawling.
 We scored sitting to include any form of the following seated positions: 1) infants sat with their buttocks on a surface, such as on the floor or a caregiver’s lap, 2) infant was in a kneeling-sit position, in which their knees were on the ground with their legs tucked underneath the buttocks, and 3) infant was in a seating device, such as a high chair, that kept the torso oriented perpendicular to the ground (a reclined position, such as in a young infant's car seat, would be counted as supine).
@@
 
-<!-- R packages? Shared code -->
 The same machine learning classification process was used as in prior work [@FranchakScott2021].
@@
 Different training/testing set combinations were used to address different goals; here we describe the general procedure of how a classifier was derived from a training set.
-Random forest models [@Breiman2001] were trained to predict the human coded body position label for each window from the set of 436 motion features using the *randomForest* package in R [@LiawWiener2002].
+Random forest models [@Breiman2001] were trained to predict the human coded body position label for each window from the set of 436 motion features using the *randomForest* package [@LiawWiener2002] in R version 4.1.2 [@R42].
 Random forest models use random subsets of features and random subsets of data across many iterations; this random subsetting creates models that are less likely to be overfit.

### 2023-03-14 manuscript.tex
R citations

@@
 
-Infants were recruited in one of two age groups: \emph{Younger} infants were between 4 and 7 months and \emph{Older} infants were between 11 and 14 months.
+Infants were recruited in one of two age groups: \emph{Younger} infants were between 4 and 7 months and \emph{older} infants were between 11 and 14 months.
 There were 8 infants in the 4-7 month group (X female) and 14 in the 11-14 month group (X female).
 Families were recruited through social media advertisements and from community events in Southern California.
-The ethnicity for infants was reported as X Hispanix/Latinx and X for not Hispanix/Latinx.
+Caregivers reported the ethnicity of infants as X Hispanix/Latinx and X for not Hispanix/Latinx.
 Race was reported as X Caucasian, X Asian, X \textasciitilde\textasciitilde\textasciitilde{} and X others.
-Parents were compensated \$30 in cash for each visit they completed.
+Families were compensated \$30 for every home recording session they completed.
 The BLINDED Institutional Review Board reviewed and approved all procedures associated with the study.
-All participants gave their informed consent to participate.
+All caregivers gave their informed consent before the start of the study.
 
@@
 Across the two age groups, we report data on a total of 33 sessions, with 13 sessions from younger infants and 20 sessions from older infants.
-Across sessions, younger infants' age ranged from 3.84 to 7.16 (\emph{M} = 5.03) and older infants' age ranged from 10.74 to 14.23 (\emph{M} = 11.75).
+Across sessions, younger infants' age ranged from 3.8 to 7.2 months (\emph{M} = 5.0) and older infants' age ranged from 10.7 to 14.2 months (\emph{M} = 11.7) .
 
@@
 Four inertial movement units (IMUs) were used to record infant movement across the day (MC10 Biostamp).
-A custom garment was made to hold the IMUs: Internal pockets were sewn into a snug-fitting pair of infant leggings so that IMUs would stay close to the body and so that infants could not pull out the sensors.
-On each side of the body there was a pocket over the hip and a pocket just above the ankle.
-Each sensor recorded accelerometer and gyroscope data at 62.5 Hz throughout the day, with sufficient battery and on-device storage to record for approximately 12 hours.
+A custom garment was designed to hold the IMUs: Internal pockets were sewn into a snug-fitting pair of infant leggings so that IMUs would stay close to the body (reducing vibration), which also prevented infants from pulling out the sensors.
+A pocket over the hip and a pocket just above the ankle were sewn on the outside of the right and left legs of the garment.
+Each sensor recorded accelerometer and gyroscope data at 62.5 Hz throughout the day, with sufficient battery and onboard storage to record for approximately 12 hours.
 Infants also wore a LENA® recorder throughout the day in the front pocket of a LENA® shirt, located near the infant's chest.
@@
 The proprietary ``Boosted Battery Base'' attached to the action camera to allow for a total of 3 hours of recording.
-However, as described below, this created two video files temporally separated by a gap of approximately 40-45 s.
+As described below, the video recording was not continuous.
+The camera created two video files temporally separated by a gap of approximately 40-45 s.
 Caregivers also received a log sheet to record times that infants napped as well as times that the sensor garment was removed from the infant (e.g., baths, diaper changes).
@@
 
-\caption{Example Timeline Caption.}\label{fig:exemplar-timeline}
+\caption{Timeline from an exemplar participant (older infant 15).
+The top row shows the model-predicted body position across the entire recording period.
+Annotations indicate when the video camera was turned on by the experimenter when arriving at the house, when the sensors were first placed on the infant, when the guided call took place, and when the video files were recorded.
+Gray areas on the timeline indicate naps, and white areas indicate times when the sensors were removed.
+The bottom row shows a zoomed-in view of the video period during which ground truth data were available.
+The top timeline shows human-coded body position and the bottom row shows model-predicted body position; these were the data used for validation.
+The first part of the video period was the proximal comparison, when video and motion data were highly synchronized.
+The second part of the video recording was the distal comparison that had coarser synchronization.
+Accuracy data for the distal comparison are provided overall and during 10-minute bins, marked by vertical dashed lines.}\label{fig:exemplar-timeline}
 \end{figure}
@@
 On the day of the visit, a researcher arrived at the participant's home in the morning between and prepared all the equipment at the doorstep.
-In order to synchronize all three recording devices, the researcher began by first turning on the video camera and the LENA audio device (the IMUs were already configured and placed at arrival because they required a proprietary sensor dock to begin recording in the laboratory).
+In order to synchronize all three recording devices, the researcher began by first turning on the video camera and the LENA audio device (the IMUs were already configured and placed before arrival because they required a proprietary sensor dock to begin recording in the laboratory).
 To create an easily recognizable synchronization point between the video recording and IMU data, the researcher dropped or struck the leggings (containing the IMUs) on a surface in view of the camera, as in Franchak et al.
@@
 All the equipment---once recording and with synchronization information recorded---was placed inside a large bucket and left outside the family's front door.
-
-The researcher then called the caregiver on the phone and walked them through a set of procedures needed to properly set up the equipment and record video for classifier training and testing.
+The researcher then called the caregiver on the phone and walked them through a set of procedures needed to properly set up the equipment and record video for ground truth human annotation of body position.
 At the start of this ``guided call'', the caregiver was instructed to place the camera in an area that captured the majority of the room.
@@
 
-Afterwards, they were to go about their day as usual with the infant wearing the sensor garment until their bedtime, only taking off the sensor for naps, baths, and diaper changes.
+Afterwards, they were asked to go about their day as usual with the infant wearing the sensor garment until their bedtime, only taking off the sensor for naps, baths, and diaper changes.
 The caregiver logged the times the sensors were removed (blank areas in the timeline in Figure \ref{fig:exemplar-timeline})) or the child took a nap (gray areas in the timeline in Figure \ref{fig:exemplar-timeline})) so that those times could be excluded from analysis.
-The following day a researcher picked up the equipment, verified the paperwork was signed, and compensated the participant.
-
-Because the camera only had the battery life to record for \textasciitilde3 hours (divided into two 90-minute video files), this divided the day into different periods for analysis.
+The following day a researcher picked up the equipment, verified that the paperwork was signed, and compensated the participant.
+
+Because the camera only had the battery life to record for \textasciitilde3 hours (split into two 90-minute video files), this divided the day into different periods for analysis.
 As seen in the bottom of Figure \ref{fig:exemplar-timeline}), the \emph{video period} comprised the first three hours of recording starting from the researcher's arrival when they turned on the camera.
 The first 90-minute video file, termed the \emph{proximal comparison}, contained the activities during the guided call followed by a period of infants and caregivers resuming their normal activities.
-Because this video contained the synchronization point, the data in this period had temporal synchrony between IMU and video data that contained errors of no more than 30-60 ms (1-2 video frames).
+Because this video contained the synchronization point, the data in this period had high temporal synchrony between IMU and video data.
+Synchrony errors were less than 30-60 ms (1-2 video frames).
 When the first video file ended, a second video file was recorded, termed the \emph{distal comparison}.
@@
 
-The proximal and distal comparison videos for each participant were annotated by trained human coders to identify infant \emph{body position} into one of 5 mutually-exclusive categories: supine, prone, sitting, upright, or held by caregiver.
+The proximal and distal comparison videos for each participant were annotated by trained human coders to identify infant \emph{body position} into one of 5 mutually-exclusive categories following the definitions in prior work (Franchak et al., 2021): supine, prone, sitting, upright, or held by caregiver.
 All coding was done using Datavyu software (datavyu.org).
 
-Supine was coded when the infant was lying on their back or was reclined up to a 45 degrees angle.
+Supine was coded when the infant was lying on their back or was reclined up to a 45 degree angle.
 Supine was also coded in the rare cases when the infant was laying on their side.
-Prone was coded when the infant was lying on their stomach, was on all fours in a downward dog position, or was crawling.
+Prone was coded when the infant was lying on their stomach, was stationary supported by the hands/knees or the hands/feet, or was crawling.
 We scored sitting to include any form of the following seated positions: 1) infants sat with their buttocks on a surface, such as on the floor or a caregiver's lap, 2) infant was in a kneeling-sit position, in which their knees were on the ground with their legs tucked underneath the buttocks, and 3) infant was in a seating device, such as a high chair, that kept the torso oriented perpendicular to the ground (a reclined position, such as in a young infant's car seat, would be counted as supine).
@@
 Different training/testing set combinations were used to address different goals; here we describe the general procedure of how a classifier was derived from a training set.
-Random forest models (Breiman, 2001) were trained to predict the human coded body position label for each window from the set of 436 motion features using the \emph{randomForest} package in R (Liaw et al., 2002).
+Random forest models (Breiman, 2001) were trained to predict the human coded body position label for each window from the set of 436 motion features using the \emph{randomForest} package (Liaw et al., 2002) in R version 4.1.2 (R Core Team, 2021).
 Random forest models use random subsets of features and random subsets of data across many iterations; this random subsetting creates models that are less likely to be overfit.
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-R42}{}}%
+R Core Team. (2021). \emph{R: A language and environment for statistical computing}. {R Foundation for Statistical Computing}.
+
 \leavevmode\vadjust pre{\hypertarget{ref-RasanenSeshadri2020}{}}%

### 2023-03-14 manuscript.Rmd
revised everything

@@
 Using the synchronization point, human-coded body annotations from video were linked to the corresponding times in the IMU timeseries data.
-A single, merged dataset was created with synchronized accelerometer signals (in three orientations: X, Y, and Z) and gyroscope signals (in three orientations: roll, pitch, and yaw) for each of the four sensors (left hip, right hip, left ankle, right ankle) with the corresponding timestamp and body position code.
+A single, merged dataset was created with synchronized accelerometer signals (in three orientations: X, Y, and Z) and gyroscope signals (in three orientations: roll, pitch, and yaw) for each of the four sensors (left hip, right hip, left ankle, right ankle) with the corresponding timestamp and body position code using the *timetk* package [@DanchoVaughan2023] and the *lubridate* package [@GrolemundWickham2011] in R version 4.1.2 [@R42].
 
@@
 
-Windowed datasets were used for machine learning classification and validation.
+Datasets containing one row per 4-s window were used for machine learning classification and validation.
 For each analysis reported in the results, a subset of data were defined as a "training" set and another, independent portion of the data were defined as a "testing" set.
 Different training/testing set combinations were used to address different goals; here we describe the general procedure of how a classifier was derived from a training set.
-Random forest models [@Breiman2001] were trained to predict the human coded body position label for each window from the set of 436 motion features using the *randomForest* package [@LiawWiener2002] in R version 4.1.2 [@R42].
+Random forest models [@Breiman2001] were trained to predict the human coded body position label for each window from the set of 436 motion features using the *randomForest* package [@LiawWiener2002].
 Random forest models use random subsets of features and random subsets of data across many iterations; this random subsetting creates models that are less likely to be overfit.
@@
 
-We report four sets of results based on `r length(unique(ppts$id_uni))` full-day testing sessions resulting in a total of `r sum(duration$duration_hours)` hours of movement recording.
-First, we focus on the accuracy of group and individual models trained and validated on the proximal comparison period.
+We report three sets of results based on `r length(unique(ppts$id_uni))` full-day testing sessions resulting in a total of `r round(sum(duration$duration_hours))` hours of movement recording.
+First, we focus on the accuracy of group and individual models trained and validated from data during the proximal comparison period.
 This period is unique in having the highest degree of temporal synchrony between video and motion data, allowing us to assess the accuracy of individual body position events.
 Second, we applied models trained from data during the proximal comparison to predict infant behavior during the distal comparison, when coarser synchrony between video and motion data was available.
-This distal comparison is crucial, because it provides the first ever test of how well body classification models predict behavior beyond the initial, supervised period.
-Data from the first two sets of analyses confirm that body position classification models are accurate across a long period of time, suggesting that data from the entire day should be valid.
-In the third set of results we examine the data quality of full-day recordings, and in the fourth set of results we test whether full-day recordings capture well-established age differences in body position experiences.
+This distal comparison is critical because it provides the first-ever test of how well infant body classification models predict behavior beyond the initial, supervised period.
+Data from the first two sets of analyses confirm that body position classification models are accurate across a long period of time, suggesting that data from the entire day will be valid.
+In the third set of results we examine the data quality of full-day recordings and test whether full-day recordings capture well-established age differences in body position experiences.
 
@@
 }
-# metric_labels <- c("Overall Accuracy", "Balanced Accuracy", "F1", "Sensitivity", "Pos Pred Value", "Kappa")
-# title_labels <- c("A. Overall Accuracy", "B. Balanced Accuracy", "C. F1 Score", "D. Sensitivity", "E. Pos. Pred. Value", "F. Kappa")
+
 metric_labels <- c("Overall Accuracy", "Kappa", "Sensitivity", "Pos Pred Value")
@@
 (metric_plots[[1]] + metric_plots[[2]])
-# (metric_plots[[1]] + metric_plots[[2]])/(metric_plots[[3]]+metric_plots[[4]])/(metric_plots[[5]] + metric_plots[[6]])
 
@@
 
-The first set of analyses use data from the proximal comparison, in which video annotations and motion data were tightly synchronized (within 30-60 ms).
-This high degree of synchronization makes it possible to link human-coded body position annotations to each 4-s window of motion data, providing ground truth data for model training and testing.
+The first set of analyses use data from the proximal comparison to determine the "best case" accuracy of the models, training and testing on similar types of data.
+The high degree of temporal synchronization between video and motion data during this period makes it possible to link human-coded body position annotations to each 4-s window of motion data, providing ground truth data for model training and testing.
 As in past work [@FranchakScott2021], we compared two types of models: *group models* and *individual models*.
-To assess the accuracy of body position classification for each recording session, we held back the last 25% of a session's proximal comparison data as the testing set.
+To assess the accuracy of body position classification for each recording session, we reserved the last 25% of a session's proximal comparison data as the testing set.
 The testing set was never used as training data.
+The two types of models differed based on what data were used during training, but the same random forest modeling procedure was used in both cases.
 A group model was trained for each session using an aggregated dataset containing the first 75% of all **other** sessions' proximal comparison data.
 This leave-one-out cross-validation tested the generalization of the model to a recording session that was not used at all in the training set.
-In contrast, individual models were trained using the first 75% of a recording session, then tested against the last 25%.
+In contrast, individual models were trained using the first 75% of a recording session, then tested against the last 25% of the same session.
 The individual model tests whether earlier training data generalize to later testing data within an individual participant's recording.
@@
 High sensitivity means that events are unlikely to be missed.
-In contrast, positive predictive value (PPV) refers to the proportion of events classified for a given position actually belonged to that position (e.g., if the model said a baby was upright during 100 windows, how many of those windows were indeed human-coded sitting events?).
+In contrast, positive predictive value (PPV) refers to the proportion of events classified for a given position that actually belonged to that position (e.g., if the model said a baby was upright during 100 windows, how many of those windows were indeed human-coded sitting events?).
 High PPV means that we can be confident in the event label.
@@
 In the first set of results, we showed that group and individual models trained from data during the proximal comparison period were accurate at classifying body position in the same recording.
-In the next analysis, we examine long-term performance by testing how accurately models trained from the proximal period could predict body position during the distal period---the 90-minute long recordings that followed the proximal recordings (Figure \@ref(fig:exemplar-timeline)).
+In the next analysis, we examine long-term performance by testing how accurately models trained from the proximal period could predict body position during the distal period (Figure \@ref(fig:exemplar-timeline)).
 A single group model was created using all sessions' proximal period training data (rather than group models leaving out a single session); the same individual models were used.
@@
 This misalignment precluded calculating accuracy based on the proportion of matching events, so we used another technique following past work [@YaoPlotz2019; @FranchakScott2021].
-For each distal comparison, we summed the amount of time infants were predicted to be in each of the 5 body position categories from the model and compared that to the summed time for the body positions based on human coding.
+For each session, we summed the amount of time infants were predicted to be in each of the 5 body position categories from the model and compared that to the summed time for the body positions based on human coding.
 Below, we report correlations between the model-predicted and human-coded aggregated body position time across the entire distal period as well as within finer, 10-minute bins.
 
-Whereas the first set of analyses used all `r length(unique(ppts$id_uni))` sessions, this was not possible in the distal comparison.
+Whereas the proximal analyses used all `r length(unique(ppts$id_uni))` sessions, this was not possible in the distal comparison.
 Because the start of the visit was scheduled during a time the infant was awake, it was common for a nap to be needed following the proximal period.
 Nine sessions were excluded because the infant was either napping or otherwise not on camera during the entire 90-minute distal recording.
-Three additional sessions were excluded because a caregiver accidentally turned off the video camera (*n* = 1) or left the house (*n* = 2).
+Three additional sessions were excluded because a caregiver accidentally turned off the video camera (*n* = 1) or (purposefully) left the house (*n* = 2).
 This left `r length(unique(agree$group$file))` sessions with usable distal comparison data.
@@
 Reviewing the video indicated that this participant spent a long period of time in a seating device that was reclined almost exactly at 45 degrees, making it difficult to determine if the infant was sitting or supine.
-The infant also spent a long time in mother's arms in an ambiguous supine/sitting position that was hard to classify.
+The infant also spent a long time in the mother's arms in an ambiguous supine/sitting position that was hard to classify.
 This participant's proximal accuracy was also poor because similar ambiguous appeared during the initial recording.
@@
 Reviewing the video showed that all disagreements occurred when the infant was in a baby walker; human coders scored this as "upright" but the models predicted it as "held".
-Most likely, the infant's movements in the baby walker were more similar to how a baby moved while carried, and unlike how most infant's moved while walking upright.
+Most likely, the infant's movements in the baby walker were more similar to how a baby moved while carried, and unlike how most infants moved while walking upright.
 
@@
     mutate(sample = row_number(), bin = floor(sample/300)) %>% 
-    add_count(file, bin) %>% ungroup() %>% filter(n > 200) %>% group_by(file) %>% mutate(num_bins = length(unique(bin))) %>% ungroup()
+    add_count(file, bin) %>% ungroup() %>% filter(n > 210) %>% group_by(file) %>% mutate(num_bins = length(unique(bin))) %>% ungroup()
   model <- ds %>% count(file, bin, num_bins, pos) %>% add_count(file, bin,num_bins,  wt = n, name = "total") %>% mutate(prop_model = n/total)
@@
 Although overall aggregate agreement in the distal comparison was strong, it is important to show that similarly strong agreement is found within a shorter timescale.
-We repeated agreement analysis, correlating human-coded body position prevalence with model-predicted prevalence, after dividing the distal comparison period into nine 10-minute bins (marked by vertical dashed lines in Figure \@ref(fig:exemplar-timeline)).
+We repeated the agreement analysis (correlating human-coded body position prevalence with model-predicted prevalence) after dividing the distal comparison period into nine 10-minute bins (marked by vertical dashed lines in Figure \@ref(fig:exemplar-timeline)).
 Infants had varying numbers of 10-minute bins depending on much time they were awake and on camera.
-Bins were included only if there was > X minutes of usable data.
-<!-- RANGE OF BINSm USABLE DATA -->
+Bins were included only if there was > 7 minutes of usable data.
 
-Table \@ref(tab:pt2binstable) shows the agreement correlation coefficients for group and individual models, including and excluding the outliers identified in the previous section.
+Table \@ref(tab:pt2binstable) shows the agreement correlation coefficients for group and individual models, including and excluding the two outliers identified in the previous section.
 Performance at a short timescale was similar to performance overall: Overall agreement after excluding outliers was excellent for both group (*r* = `r group_agree_cor_noout`) and individual models (*r* = `r split_agree_cor_noout`).
@@
 Nap times reported during the recording period ranged from `r apa_num(min(coverage$naptime))`-`r apa_num(max(coverage$naptime))` hours with a median of `r apa_num(median(coverage$naptime))` hours.
-Body position data was available to describe `r apa_num(min(coverage$prop_accounted))`-`r apa_num(max(coverage$prop_accounted))`% of the awake portion of the recording period (median = `r apa_num(median(coverage$prop_accounted))`%).
+Body position data was available to describe `r apa_num(min(coverage$prop_accounted)*100, digits = 1)`-`r apa_num(max(coverage$prop_accounted)*100, digits = 1)`% of the awake portion of the recording period (median = `r apa_num(median(coverage$prop_accounted)*100, digits = 1)`%).
 Younger infant \#13 had the least portion of the waking day accounted for; the family left the house to run errands during the majority of the recording period, and the infant napped during much of the remaining time at home.
@@
 The most straightforward comparisons between the two investigations are upright and prone, because they were defined identically.
-In @Survey, upright time increased from 0.6-5.5% of the time for infants 3-6 months, increasing to 22.0% at 12 months.
-Similarly, younger infants' sitting time estimated from full-day group models was `r pos_sum_lookup["Upright", "Younger_Group"]`% compared with `r pos_sum_lookup["Upright", "Older_Group"]`% for the older group.
-In @Survey, prone time increased ranged from 2.9-9.2% of the time for infants 3-6 months, and was 7.2% at 12 months.
-In the current study, younger infants' prone time was was `r pos_sum_lookup["Prone", "Younger_Group"]`% and `r pos_sum_lookup["Prone", "Older_Group"]`% for older infants.
-Held, sitting, and supine are more difficult to compare because prior work [@Survey] considered held any time that infants were in caregivers' arms or on their laps, however, we only counted laying or sitting on caregivers' laps as supine and sitting, respectively.
-Further complicating the comparison, @Survey included a reclined category that we would classify as either supine or sitting depending on the specifics.
-Although the numbers are less directly comparable, our results are similar in showing that held and supine time was less for older infants than younger infants.
+In @Survey, upright time was 0.6-5.5% of the time for infants 3-6 months, increasing to 22.0% at 12 months.
+Results from the current study were similar: younger infants' sitting time estimated from full-day group models was `r pos_sum_lookup["Upright", "Younger_Group"]`% compared with `r pos_sum_lookup["Upright", "Older_Group"]`% for the older group.
+In @Survey, prone time was 2.9-9.2% of the time for infants 3-6 months, and was 7.2% at 12 months.
+In the current study, prone time was `r pos_sum_lookup["Prone", "Younger_Group"]`% for younger infants and `r pos_sum_lookup["Prone", "Older_Group"]`% for older infants.
+Held, sitting, and supine are more difficult to compare because prior work [@Survey] considered held as any time that infants were in caregivers' arms or on their laps, however, we counted laying or sitting on caregivers' laps as supine and sitting, respectively.
+Further complicating the comparison, @Survey included a reclined category that we would classify as either supine or sitting (whichever it was closer to).
+Although the numbers are less directly comparable, our results are similar in showing that held and supine time was less for older infants than younger infants (Table \@ref(tab:agetable)).
 Sitting was greater for older infants (`r pos_sum_lookup["Sitting", "Older_Group"]`%) compared with younger infants (`r pos_sum_lookup["Sitting", "Younger_Group"]`%).

### 2023-03-14 manuscript.tex
revised everything

@@
 Using the synchronization point, human-coded body annotations from video were linked to the corresponding times in the IMU timeseries data.
-A single, merged dataset was created with synchronized accelerometer signals (in three orientations: X, Y, and Z) and gyroscope signals (in three orientations: roll, pitch, and yaw) for each of the four sensors (left hip, right hip, left ankle, right ankle) with the corresponding timestamp and body position code.
+A single, merged dataset was created with synchronized accelerometer signals (in three orientations: X, Y, and Z) and gyroscope signals (in three orientations: roll, pitch, and yaw) for each of the four sensors (left hip, right hip, left ankle, right ankle) with the corresponding timestamp and body position code using the \emph{timetk} package (Dancho \& Vaughan, 2023) and the \emph{lubridate} package (Grolemund \& Wickham, 2011) in R version 4.1.2 (R Core Team, 2021).
 
@@
 
-Windowed datasets were used for machine learning classification and validation.
+Datasets containing one row per 4-s window were used for machine learning classification and validation.
 For each analysis reported in the results, a subset of data were defined as a ``training'' set and another, independent portion of the data were defined as a ``testing'' set.
 Different training/testing set combinations were used to address different goals; here we describe the general procedure of how a classifier was derived from a training set.
-Random forest models (Breiman, 2001) were trained to predict the human coded body position label for each window from the set of 436 motion features using the \emph{randomForest} package (Liaw et al., 2002) in R version 4.1.2 (R Core Team, 2021).
+Random forest models (Breiman, 2001) were trained to predict the human coded body position label for each window from the set of 436 motion features using the \emph{randomForest} package (Liaw et al., 2002).
 Random forest models use random subsets of features and random subsets of data across many iterations; this random subsetting creates models that are less likely to be overfit.
@@
 
-We report four sets of results based on 33 full-day testing sessions resulting in a total of 294.07 hours of movement recording.
-First, we focus on the accuracy of group and individual models trained and validated on the proximal comparison period.
+We report three sets of results based on 33 full-day testing sessions resulting in a total of 294 hours of movement recording.
+First, we focus on the accuracy of group and individual models trained and validated from data during the proximal comparison period.
 This period is unique in having the highest degree of temporal synchrony between video and motion data, allowing us to assess the accuracy of individual body position events.
 Second, we applied models trained from data during the proximal comparison to predict infant behavior during the distal comparison, when coarser synchrony between video and motion data was available.
-This distal comparison is crucial, because it provides the first ever test of how well body classification models predict behavior beyond the initial, supervised period.
-Data from the first two sets of analyses confirm that body position classification models are accurate across a long period of time, suggesting that data from the entire day should be valid.
-In the third set of results we examine the data quality of full-day recordings, and in the fourth set of results we test whether full-day recordings capture well-established age differences in body position experiences.
+This distal comparison is critical because it provides the first-ever test of how well infant body classification models predict behavior beyond the initial, supervised period.
+Data from the first two sets of analyses confirm that body position classification models are accurate across a long period of time, suggesting that data from the entire day will be valid.
+In the third set of results we examine the data quality of full-day recordings and test whether full-day recordings capture well-established age differences in body position experiences.
 
@@
 
-The first set of analyses use data from the proximal comparison, in which video annotations and motion data were tightly synchronized (within 30-60 ms).
-This high degree of synchronization makes it possible to link human-coded body position annotations to each 4-s window of motion data, providing ground truth data for model training and testing.
+The first set of analyses use data from the proximal comparison to determine the ``best case'' accuracy of the models, training and testing on similar types of data.
+The high degree of temporal synchronization between video and motion data during this period makes it possible to link human-coded body position annotations to each 4-s window of motion data, providing ground truth data for model training and testing.
 As in past work (Franchak et al., 2021), we compared two types of models: \emph{group models} and \emph{individual models}.
-To assess the accuracy of body position classification for each recording session, we held back the last 25\% of a session's proximal comparison data as the testing set.
+To assess the accuracy of body position classification for each recording session, we reserved the last 25\% of a session's proximal comparison data as the testing set.
 The testing set was never used as training data.
+The two types of models differed based on what data were used during training, but the same random forest modeling procedure was used in both cases.
 A group model was trained for each session using an aggregated dataset containing the first 75\% of all \textbf{other} sessions' proximal comparison data.
 This leave-one-out cross-validation tested the generalization of the model to a recording session that was not used at all in the training set.
-In contrast, individual models were trained using the first 75\% of a recording session, then tested against the last 25\%.
+In contrast, individual models were trained using the first 75\% of a recording session, then tested against the last 25\% of the same session.
 The individual model tests whether earlier training data generalize to later testing data within an individual participant's recording.
@@
 In the first set of results, we showed that group and individual models trained from data during the proximal comparison period were accurate at classifying body position in the same recording.
-In the next analysis, we examine long-term performance by testing how accurately models trained from the proximal period could predict body position during the distal period---the 90-minute long recordings that followed the proximal recordings (Figure \ref{fig:exemplar-timeline}).
+In the next analysis, we examine long-term performance by testing how accurately models trained from the proximal period could predict body position during the distal period (Figure \ref{fig:exemplar-timeline}).
 A single group model was created using all sessions' proximal period training data (rather than group models leaving out a single session); the same individual models were used.
@@
 This misalignment precluded calculating accuracy based on the proportion of matching events, so we used another technique following past work (Franchak et al., 2021; Yao et al., 2019).
-For each distal comparison, we summed the amount of time infants were predicted to be in each of the 5 body position categories from the model and compared that to the summed time for the body positions based on human coding.
+For each session, we summed the amount of time infants were predicted to be in each of the 5 body position categories from the model and compared that to the summed time for the body positions based on human coding.
 Below, we report correlations between the model-predicted and human-coded aggregated body position time across the entire distal period as well as within finer, 10-minute bins.
 
-Whereas the first set of analyses used all 33 sessions, this was not possible in the distal comparison.
+Whereas the proximal analyses used all 33 sessions, this was not possible in the distal comparison.
 Because the start of the visit was scheduled during a time the infant was awake, it was common for a nap to be needed following the proximal period.
 Nine sessions were excluded because the infant was either napping or otherwise not on camera during the entire 90-minute distal recording.
-Three additional sessions were excluded because a caregiver accidentally turned off the video camera (\emph{n} = 1) or left the house (\emph{n} = 2).
+Three additional sessions were excluded because a caregiver accidentally turned off the video camera (\emph{n} = 1) or (purposefully) left the house (\emph{n} = 2).
 This left 22 sessions with usable distal comparison data.
@@
 Reviewing the video showed that all disagreements occurred when the infant was in a baby walker; human coders scored this as ``upright'' but the models predicted it as ``held''.
-Most likely, the infant's movements in the baby walker were more similar to how a baby moved while carried, and unlike how most infant's moved while walking upright.
+Most likely, the infant's movements in the baby walker were more similar to how a baby moved while carried, and unlike how most infants moved while walking upright.
 
@@
 Although overall aggregate agreement in the distal comparison was strong, it is important to show that similarly strong agreement is found within a shorter timescale.
-We repeated agreement analysis, correlating human-coded body position prevalence with model-predicted prevalence, after dividing the distal comparison period into nine 10-minute bins (marked by vertical dashed lines in Figure \ref{fig:exemplar-timeline}).
+We repeated the agreement analysis (correlating human-coded body position prevalence with model-predicted prevalence) after dividing the distal comparison period into nine 10-minute bins (marked by vertical dashed lines in Figure \ref{fig:exemplar-timeline}).
 Infants had varying numbers of 10-minute bins depending on much time they were awake and on camera.
-Bins were included only if there was \textgreater{} X minutes of usable data.
-
-Table \ref{tab:pt2binstable} shows the agreement correlation coefficients for group and individual models, including and excluding the outliers identified in the previous section.
+Bins were included only if there was \textgreater{} 7 minutes of usable data.
+
+Table \ref{tab:pt2binstable} shows the agreement correlation coefficients for group and individual models, including and excluding the two outliers identified in the previous section.
 Performance at a short timescale was similar to performance overall: Overall agreement after excluding outliers was excellent for both group (\emph{r} = 0.92) and individual models (\emph{r} = 0.94).
@@
 The vast majority of session-averaged predictions fall within 1 minute of error without a clear bias towards overestimation or underestimation (with the previously-identified outliers as exceptions).
-For group models, we calculated the percentage of 10-min bins across participants that had errors \textless{} 1 minute: 94.68\% for held, 80.85\% for supine, 94.68\% for prone, 77.66\% for sitting, and 94.68\% for upright.
-For individual models, the percent of 10-min bins with \textless{} 1 minute of error was: 92.94\% for held, 88.24\% for supine, 88.24\% for prone, 83.53\% for sitting, and 96.47\% for upright.
+For group models, we calculated the percentage of 10-min bins across participants that had errors \textless{} 1 minute: 94.62\% for held, 80.65\% for supine, 94.62\% for prone, 77.42\% for sitting, and 94.62\% for upright.
+For individual models, the percent of 10-min bins with \textless{} 1 minute of error was: 92.86\% for held, 88.10\% for supine, 88.10\% for prone, 83.33\% for sitting, and 96.43\% for upright.
 Taken together, these results suggest that most estimates across models, sessions, and body positions had small amounts of prediction error even after a long delay separating the period that models were trained and tested.
@@
 Nap times reported during the recording period ranged from 0.00-4.99 hours with a median of 2.22 hours.
-Body position data was available to describe 0.18-1.00\% of the awake portion of the recording period (median = 1.00\%).
+Body position data was available to describe 18.3-100.0\% of the awake portion of the recording period (median = 100.0\%).
 Younger infant \#13 had the least portion of the waking day accounted for; the family left the house to run errands during the majority of the recording period, and the infant napped during much of the remaining time at home.
@@
 The most straightforward comparisons between the two investigations are upright and prone, because they were defined identically.
-In Franchak (2019), upright time increased from 0.6-5.5\% of the time for infants 3-6 months, increasing to 22.0\% at 12 months.
-Similarly, younger infants' sitting time estimated from full-day group models was 7.72\% compared with 18.57\% for the older group.
-In Franchak (2019), prone time increased ranged from 2.9-9.2\% of the time for infants 3-6 months, and was 7.2\% at 12 months.
-In the current study, younger infants' prone time was was 14.46\% and 14.45\% for older infants.
-Held, sitting, and supine are more difficult to compare because prior work (Franchak, 2019) considered held any time that infants were in caregivers' arms or on their laps, however, we only counted laying or sitting on caregivers' laps as supine and sitting, respectively.
-Further complicating the comparison, Franchak (2019) included a reclined category that we would classify as either supine or sitting depending on the specifics.
-Although the numbers are less directly comparable, our results are similar in showing that held and supine time was less for older infants than younger infants.
+In Franchak (2019), upright time was 0.6-5.5\% of the time for infants 3-6 months, increasing to 22.0\% at 12 months.
+Results from the current study were similar: younger infants' sitting time estimated from full-day group models was 7.72\% compared with 18.57\% for the older group.
+In Franchak (2019), prone time was 2.9-9.2\% of the time for infants 3-6 months, and was 7.2\% at 12 months.
+In the current study, prone time was 14.46\% for younger infants and 14.45\% for older infants.
+Held, sitting, and supine are more difficult to compare because prior work (Franchak, 2019) considered held as any time that infants were in caregivers' arms or on their laps, however, we counted laying or sitting on caregivers' laps as supine and sitting, respectively.
+Further complicating the comparison, Franchak (2019) included a reclined category that we would classify as either supine or sitting (whichever it was closer to).
+Although the numbers are less directly comparable, our results are similar in showing that held and supine time was less for older infants than younger infants (Table \ref{tab:agetable}).
 Sitting was greater for older infants (44.42\%) compared with younger infants (25.37\%).
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-DanchoVaughan2023}{}}%
+Dancho, M., \& Vaughan, D. (2023). \emph{Timetk: A tool kit for working with time series in r}.
+
 \leavevmode\vadjust pre{\hypertarget{ref-DengTrujillo-Priego2019}{}}%
@@
 \emph{Infant Behavior and Development}, \emph{62}, 101530.
+
+\leavevmode\vadjust pre{\hypertarget{ref-GrolemundWickham2011}{}}%
+Grolemund, G., \& Wickham, H. (2011). Dates and times made easy with {lubridate}. \emph{Journal of Statistical Software}, \emph{40}(3), 1--25. \url{https://www.jstatsoft.org/v40/i03/}
 

### 2023-03-15 manuscript.Rmd
figure captions

@@
 
-```{r metrics, warning=FALSE, fig.cap="Metrics", fig.align="center", fig.width=6, fig.height=3, cache=TRUE, cache.extra=tools::md5sum("data/group_split_metrics.csv")}
+```{r metrics, warning=FALSE, fig.cap="Metrics of agreement between human-annotated body position and model predictions of body position from the proximal comparison period.
+Overall accuracy (A) and Cohen's Kappa (B) are plotted separately for group models and individual models.
+Each blue circle represents that accuracy for each recording session.
+Horizontal black bars indicates the mean across sessions.", fig.align="center", fig.width=6, fig.height=3, cache=TRUE, cache.extra=tools::md5sum("data/group_split_metrics.csv")}
 metrics <- read_csv("data/group_split_metrics.csv") %>% 
@@
   apa_table(metrics_summary, col.names = c("Metric", "Median", "Mean", "SD", "Median", "Mean", "SD"),    col_spanners = list("Group" = c(2, 4), "Individual" = c(5, 7)),
-          caption = "Summary statistics for model performance metrics shown separately for group and individual models.")
+          caption = "Model performance metrics (overall accuracy, Cohen's Kappa, sensitivity, positive predictive value) from the proximal comparison period.
+Descriptive statistics are shown separately for group and individual models.")
 
@@
           midrules = c(5, 10),
-          caption = "Model performance metrics for each body position category, shown separately for group and individual models.")
+          caption = "Model performance metrics for each body position (supine, prone, sitting, upright, and held) during the proximal comparison period, shown separately for group and individual models.")
 metrics_long_summary <- metrics_long_summary %>% 
@@
 Agreement for group models is shown in (A) and agreement for individual models is shown in (B).
-Plots are shown separately for each body position with a reference line that indicates perfect agreement; each point in a plot represent data for a single participant.
-The three outlier participants are plotted in dark gray, with a different shape marking each individual.", fig.align="center", fig.width=7, fig.height=6.5, cache=TRUE, cache.extra=tools::md5sum("data/compiled_agreement_position.csv"), cache.extra=tools::md5sum("data/compiled_agreement_split.csv")}
+Plots are shown separately for each body position with a reference line that indicates perfect agreement; each point in a plot represents data for a single session.
+The two outlier participants are plotted in dark gray, with a different shape marking each individual.
+", fig.align="center", fig.width=7, fig.height=6.5, cache=TRUE, cache.extra=tools::md5sum("data/compiled_agreement_position.csv"), cache.extra=tools::md5sum("data/compiled_agreement_split.csv")}
 
@@
 apa_table(pt2corr, col.names = c("Position", "Group", "Individual", "Group", "Individual"), col_spanners = list(`With Outliers` = c(2, 3), `Without Outliers` = c(4, 5)),
-          midrules = 5, caption = "Correlations between human-coded and model-predicted body position durations across the entire long delay period.
-Correlations are provided within each posture and overall, and computed separately using group and individual models with and without outlier participants.")
+          midrules = 5, caption = "Correlations between human-coded and model-predicted body position durations across the entire distal comparison period.
+Correlations are provided within each body position and overall.
+Correlations are presented separately for group and individual models with and without the two outlier participants.")
 ```
@@
 
-```{r part2bins, warning=FALSE, fig.cap="Prediction difference (difference in minutes between human-coded and model-predicted body position) for 10-minute bins in the distal comparison.
-Each point shows the mean and SE for a single participant for each body position, summarizing the prediction difference for each of their 10-minute bins.
+```{r part2bins, warning=FALSE, fig.cap="Prediction difference (difference in minutes between human-coded and model-predicted body position) for 10-minute bins in the distal comparison period.
+Each point shows the mean and SE for a single participant's recording session for each body position, summarizing the prediction difference for each of their 10-minute bins.
 Points falling within the gray shaded region indicate that average prediction errors were less than 1 minute.
 Performance is plotted separately for (A) group models and (B) individual models.
-The three outlier participants are plotted in dark gray, with a different shape marking each individual.", fig.align="center", fig.width=8, fig.height=4.5, cache=TRUE, cache.extra=tools::md5sum("data/compiled_agreement_position.csv"), cache.extra=tools::md5sum("data/compiled_agreement_split.csv")}
+The two outlier participants are plotted in dark gray, with a different shape marking each individual.", fig.align="center", fig.width=8, fig.height=4.5, cache=TRUE, cache.extra=tools::md5sum("data/compiled_agreement_position.csv"), cache.extra=tools::md5sum("data/compiled_agreement_split.csv")}
 calculate_agreement_bins <- function(ds) {
@@
 apa_table(pt2corr_bins, col.names = c("Position", "Group", "Individual", "Group", "Individual"), col_spanners = list(`With Outliers` = c(2, 3), `Without Outliers` = c(4, 5)),
-          midrules = 5, caption = "Correlations between human-coded and model-predicted body position durations using 10-minute bins during the distal comparison.
+          midrules = 5, caption = "Correlations between human-coded and model-predicted body position durations using 10-minute bins during the distal comparison period.
 Correlations are provided within each posture and overall, and computed separately using group and individual models with and without outlier participants.")
@@
 
-```{r timelines, warning=FALSE, fig.cap="Timelines", fig.align="center", fig.width=7.75, fig.height=8, cache=TRUE, cache.extra=tools::md5sum("data/imu-compiled.csv")}
+```{r timelines, warning=FALSE, fig.cap="Full-day timelines for each individual recording session, split by (A) younger infants and (B) older infants.
+Each participant's timeline shows a stacked bar graph with the proportion of time spent in each of the body position for every 5-minute period throughout the day, based on group model predictions of body position.
+The x-axis shows time of day.
+Caregiver-reported naps are marked by gray bars; blank gaps indicate caregiver-reported times that the sensor garment was removed for diaper changes, baths, or trips out of the house.", fig.align="center", fig.width=7.75, fig.height=8, cache=TRUE, cache.extra=tools::md5sum("data/imu-compiled.csv")}
 timeline_data <- read_csv("data/imu-compiled.csv") %>% 
@@
 
-```{r age, warning=FALSE, fig.cap="Age trends", fig.align="center", fig.width=7.5, fig.height=7, cache=TRUE, cache.extra=tools::md5sum("data/imu-compiled.csv")}
+```{r age, warning=FALSE, fig.cap="Age differences in daily body position predicted from (A) group models and (B) individual models.
+Each circle represents one full-day recording session's proportion of time in each body position (y-axis) against age in months (x-axis).
+Black lines indicate best fit regression lines indicating decreases in supine time and increases in sitting and upright time with age.", fig.align="center", fig.width=7.5, fig.height=7, cache=TRUE, cache.extra=tools::md5sum("data/imu-compiled.csv")}
 fullday_group <- read_csv("data/imu-compiled.csv") %>% 

### 2023-03-15 manuscript.tex
figure captions

@@
 
-\caption{Metrics}\label{fig:metrics}
+\caption{Metrics of agreement between human-annotated body position and model predictions of body position from the proximal comparison period.
+Overall accuracy (A) and Cohen's Kappa (B) are plotted separately for group models and individual models.
+Each blue circle represents that accuracy for each recording session.
+Horizontal black bars indicates the mean across sessions.}\label{fig:metrics}
 \end{figure}
@@
 
-\caption{\label{tab:metrics-table}Summary statistics for model performance metrics shown separately for group and individual models.}
+\caption{\label{tab:metrics-table}Model performance metrics (overall accuracy, Cohen's Kappa, sensitivity, positive predictive value) from the proximal comparison period.
+Descriptive statistics are shown separately for group and individual models.}
 
@@
 
-\caption{\label{tab:metrics-by-class}Model performance metrics for each body position category, shown separately for group and individual models.}
+\caption{\label{tab:metrics-by-class}Model performance metrics for each body position (supine, prone, sitting, upright, and held) during the proximal comparison period, shown separately for group and individual models.}
 
@@
 Agreement for group models is shown in (A) and agreement for individual models is shown in (B).
-Plots are shown separately for each body position with a reference line that indicates perfect agreement; each point in a plot represent data for a single participant.
-The three outlier participants are plotted in dark gray, with a different shape marking each individual.}\label{fig:part2overall}
+Plots are shown separately for each body position with a reference line that indicates perfect agreement; each point in a plot represents data for a single session.
+The two outlier participants are plotted in dark gray, with a different shape marking each individual.
+}\label{fig:part2overall}
 \end{figure}
@@
 
-\caption{\label{tab:pt2overalltable}Correlations between human-coded and model-predicted body position durations across the entire long delay period.
-Correlations are provided within each posture and overall, and computed separately using group and individual models with and without outlier participants.}
+\caption{\label{tab:pt2overalltable}Correlations between human-coded and model-predicted body position durations across the entire distal comparison period.
+Correlations are provided within each body position and overall.
+Correlations are presented separately for group and individual models with and without the two outlier participants.}
 
@@
 
-\caption{Prediction difference (difference in minutes between human-coded and model-predicted body position) for 10-minute bins in the distal comparison.
-Each point shows the mean and SE for a single participant for each body position, summarizing the prediction difference for each of their 10-minute bins.
+\caption{Prediction difference (difference in minutes between human-coded and model-predicted body position) for 10-minute bins in the distal comparison period.
+Each point shows the mean and SE for a single participant's recording session for each body position, summarizing the prediction difference for each of their 10-minute bins.
 Points falling within the gray shaded region indicate that average prediction errors were less than 1 minute.
 Performance is plotted separately for (A) group models and (B) individual models.
-The three outlier participants are plotted in dark gray, with a different shape marking each individual.}\label{fig:part2bins}
+The two outlier participants are plotted in dark gray, with a different shape marking each individual.}\label{fig:part2bins}
 \end{figure}
@@
 
-\caption{\label{tab:pt2binstable}Correlations between human-coded and model-predicted body position durations using 10-minute bins during the distal comparison.
+\caption{\label{tab:pt2binstable}Correlations between human-coded and model-predicted body position durations using 10-minute bins during the distal comparison period.
 Correlations are provided within each posture and overall, and computed separately using group and individual models with and without outlier participants.}
@@
 
-\caption{Timelines}\label{fig:timelines}
+\caption{Full-day timelines for each individual recording session, split by (A) younger infants and (B) older infants.
+Each participant's timeline shows a stacked bar graph with the proportion of time spent in each of the body position for every 5-minute period throughout the day, based on group model predictions of body position.
+The x-axis shows time of day.
+Caregiver-reported naps are marked by gray bars; blank gaps indicate caregiver-reported times that the sensor garment was removed for diaper changes, baths, or trips out of the house.}\label{fig:timelines}
 \end{figure}
@@
 
-\caption{Age trends}\label{fig:age}
+\caption{Age differences in daily body position predicted from (A) group models and (B) individual models.
+Each circle represents one full-day recording session's proportion of time in each body position (y-axis) against age in months (x-axis).
+Black lines indicate best fit regression lines indicating decreases in supine time and increases in sitting and upright time with age.}\label{fig:age}
 \end{figure}

### 2023-03-15 manuscript.Rmd
discussion outline

@@
 
+## Performance Verdict: It Works
+
+Accuracy good compared to typical metrics of agreement, comparable to human agreement, in both proximal and distal comparisons.
+
+Outlier participants give us more confidence that it actually works as intended.
+Disagreements limited to some edge cases, not all events/classes.
+Even w/ outliers (which are they on timelines), most of the rest of the data seems typical for the age group.
+Some outlier-ish data in age comparisons, but nothing out of the ordinary (hard to know if it's the model or the behavior)
+
+Nice to have video to confirm. Best practice in LENA is to still do some transcription yourself to see what cases it works. Still helpful to have other sources to triangulate against.
+
+## Group and Individual Models
+
+No clear winner.
+Higher raw accuracy/kappa for individual models; some (but not all) better by-class performance for Kappa's.
+But group models better for prone; more sensitive for prone and for held; barring outliers; group has some better part 2 correlations depending on the class.
+Notably, worst outlier in aggregate age difference data was in an individual model.
+
+Big advantages of group model are that 1) trained on more data (so likely a bit more robust), 2) less dependent on getting the perfect training set, 3) between individual differences are due to behavior, not due to choice of model, 4) can apply to any infant, 5) can apply to an infant for whom we have no video.
+Individual models might capture some idiosyncracies of movement (or device use) that are hard to get from big model that homogenizes behavior.
+Individual models are a nice choice for rapid prototyping/piloting/feasibility.
+Can get proof of concept on a single participant or a new kind of code (loc; stroller use; etc.).
+Clinical cases too.
+
+Tell some things we did but didn't report.
+Tried modeling each age separately but there was no overall improvement in performance.
+Categories are so broad and include less/more mature forms of behavior.
+Not independent sitting, but all kinds of sitting.
+Tried other ML algorithms.
+Tried fewer sensors.
+
+## What does this get us?
+
+Other things look a bit wacky, but within-day heterogeniety is the point.
+Shouldn't be surprised to find it.
+Some can confirm on video.
+Activity are a big shaper of behavior, have enough behavior to actually see it.
+
+See how other aspects of behavior align w/ movement in real time. See how new motor skills relate to chg in daily behavior.
+
+Efficiency. Hours saved.
+
+More easily shareable data (non-identifiable, text based)
+
+## Limitations
+- No validation at the end of the day (practically this would be hard to do, syncing a challenge)
+- Limited to within the home. Driving/strollers are motion signals we don't have training data for, but this could be added in the future. 
+- Relying on CG reports of naps/diaper changes.
+
+## Conclusions
+
 \newpage

### 2023-03-15 manuscript.Rmd
started writing discussion

@@
 ---
-title             : "Full-day, in home validation of infant body position measurements from inertial sensors"
+title             : "Long-form recording of infant body position in the home using wearable inertial sensors"
 shorttitle        : "Infant position in the home"
@@
 
-### How Well did Recordings Capture Infants' Entire Day?
+### How Well Did Recordings Capture Infants' Entire Day?
 
@@
 
+Across three analyses, we demonstrated the validity of long-form recordings of infant body position using wearable inertial sensors.
+Both group and individual models trained during the proximal comparison period performed well on testing data collected from the same recording period.
+Most important, accuracy was consistently strong later in the recording when behavior was completely unsupervised: Human-coded and model-predicted durations of body positions during the distal comparison period were highly correlated, even when narrowing to the scale of 10 minutes.
+Examining full recordings showed that the new method allows us to capture more data (median = `r apa_num(median(coverage$hours_recorded))` hours of awake behavior) than is typical with video methods (1-2 hours) while simultaneously reducing the human labor needed to annotate it.
+Ultimately, age differences in body position mirrored past findings that employed other methods, suggesting that the end product is suitable for describing developmental changes in motor behavior.
+
 ## Performance Verdict: It Works
@@
 
-Nice to have video to confirm. Best practice in LENA is to still do some transcription yourself to see what cases it works. Still helpful to have other sources to triangulate against.
+Nice to have video to confirm.
+Best practice in LENA is to still do some transcription yourself to see what cases it works.
+Still helpful to have other sources to triangulate against.
+Maybe don't use it for a holding detector.
 
@@
 - Relying on CG reports of naps/diaper changes. 
+- Bias in training data
 

### 2023-03-15 manuscript.Rmd
first part of discussion

@@
 
-## Performance Verdict: It Works
-
-Accuracy good compared to typical metrics of agreement, comparable to human agreement, in both proximal and distal comparisons.
+## Accurate Results in Challenging Circumstances
+
+How did classification in the current investigation compare to prior work [@FranchakScott2021; @AiraksinenRasanen2020; @AiraksinenGallen2022]?
+Across body positions, median accuracy was `r metrics_lookup["Overall Accuracy", "median_Group"]`% for group models (Kappa = `r metrics_lookup["Kappa", "median_Group"]`) and `r metrics_lookup["Overall Accuracy", "median_Individual"]`% for individual models (Kappa = `r metrics_lookup["Kappa", "median_Individual"]`) during the proximal comparison.
+Although these accuracy and kappa values are slightly lower than in past work (accuracy 95-98%, kappas .93-.95), we note that those past values were obtained under ideal circumstances.
+A researcher applied the sensors in the lab [@FranchakScott2021; @AiraksinenRasanen2020] or home [@AiraksinenGallen2022] and set the stage for how infants and caregivers should interact.
+For example, both @FranchakScott2021 and @AiraksinenGallen2022 instructed the caregivers that the recorded sessions should mimic a "playtime", which would encourage more common activities that would be easier to classify (e.g., crawling and sitting on the floor) and discourage idiosyncratic and potentially more challenging activities (e.g., positioning in infant-specific furniture like exersaucers and walkers, sitting or lying for long periods while eating or nursing).
+Examining the timelines in Figure \@ref(fig:timelines) shows how much variability there is between and within sessions in the earlier parts of the recording that were used for training and testing.
+With the exception of the brief guided session at the start of the recording (which ensured that every infant spent at least a short amount of time in each body position), the majority of the recorded behavior was truly natural and "in the wild".
+
+The challenge only increased in the distal comparison period.
+Since sessions were scheduled at convenient times to for the infant and caregiver, the transition from the proximal to distal period increased the odds that infants might need a nap, a snack/meal, or engage in a less typical activity (e.g., watching TV).
+Indeed, the distal comparison in Figure \@ref(fig:exemplar-timeline) contains a long period (almost 50 minutes) where the infant sat in a high chair eating lunch.
+With such a long delay in time from when the sensor garment was first put on, there is an increasing chance that sensors moved from their ideal positions.
+Regardless of these challenges, accuracy in the distal comparison continued to be strong.
+Barring two outliers (which we will discuss further), agreement was high for group (*r* = `r group_agree_cor_noout`) and individual models (*r* = `r split_agree_cor_noout`).
+Even at a finer timescale, most errors within 10-minute bins were less than 1 minute in total for all five body positions.
+
+To put this in context, the reliability of human coders on the body position code was ~90%, putting model-predicted accuracy on a similar level to human coders.
+Acceptable inter-rater reliability in developmental psychology studies of motor behavior is greater than 80%; both group and individual model results exceed that standard for most sessions.
+<!-- Reliability standard? -->
 

### 2023-03-15 manuscript.Rmd
more discussion

@@
 How did classification in the current investigation compare to prior work [@FranchakScott2021; @AiraksinenRasanen2020; @AiraksinenGallen2022]?
-Across body positions, median accuracy was `r metrics_lookup["Overall Accuracy", "median_Group"]`% for group models (Kappa = `r metrics_lookup["Kappa", "median_Group"]`) and `r metrics_lookup["Overall Accuracy", "median_Individual"]`% for individual models (Kappa = `r metrics_lookup["Kappa", "median_Individual"]`) during the proximal comparison.
+Across body positions, median accuracy was `r metrics_lookup["Overall Accuracy", "median_Group"]*100`% for group models (Kappa = `r metrics_lookup["Kappa", "median_Group"]*100`) and `r metrics_lookup["Overall Accuracy", "median_Individual"]`% for individual models (Kappa = `r metrics_lookup["Kappa", "median_Individual"]`) during the proximal comparison.
 Although these accuracy and kappa values are slightly lower than in past work (accuracy 95-98%, kappas .93-.95), we note that those past values were obtained under ideal circumstances.
@@
 To put this in context, the reliability of human coders on the body position code was ~90%, putting model-predicted accuracy on a similar level to human coders.
-Acceptable inter-rater reliability in developmental psychology studies of motor behavior is greater than 80%; both group and individual model results exceed that standard for most sessions.
+Acceptable inter-rater reliability in developmental psychology studies of motor behavior is greater than 80%; both group and individual model results exceed that standard for most recorded sessions.
+Another comparison we can make is to the accuracy of the LENA&reg;'s automated metrics.
+Correlations between human and model frequencies were *r* = .698 for adult word count and *r* = .649 for child word count [@CristiaLavechin2020].
+Even including outliers, correlations between human and model-predicted body position were *r* = `r group_agree_cor` (group) and *r* = `r split_agree_cor` (individual), easily surpassing the LENA&reg; device's most reliable metrics.
 <!-- Reliability standard? -->

### 2023-03-15 manuscript.tex
more discussion

@@
 \hypersetup{
-  pdftitle={Full-day, in home validation of infant body position measurements from inertial sensors},
+  pdftitle={Long-form recording of infant body position in the home using wearable inertial sensors},
   pdfauthor={John M. Franchak1, Maximilian Tang1, Hailey Rousey1, \& Chuan Luo1},
@@
 
-\title{Full-day, in home validation of infant body position measurements from inertial sensors}
+\title{Long-form recording of infant body position in the home using wearable inertial sensors}
 \author{John M. Franchak\textsuperscript{1}, Maximilian Tang\textsuperscript{1}, Hailey Rousey\textsuperscript{1}, \& Chuan Luo\textsuperscript{1}}
@@
 \hypertarget{how-well-did-recordings-capture-infants-entire-day}{%
-\subsubsection{How Well did Recordings Capture Infants' Entire Day?}\label{how-well-did-recordings-capture-infants-entire-day}}
+\subsubsection{How Well Did Recordings Capture Infants' Entire Day?}\label{how-well-did-recordings-capture-infants-entire-day}}
 
@@
 \section{Discussion}\label{discussion}}
+
+Across three analyses, we demonstrated the validity of long-form recordings of infant body position using wearable inertial sensors.
+Both group and individual models trained during the proximal comparison period performed well on testing data collected from the same recording period.
+Most important, accuracy was consistently strong later in the recording when behavior was completely unsupervised: Human-coded and model-predicted durations of body positions during the distal comparison period were highly correlated, even when narrowing to the scale of 10 minutes.
+Examining full recordings showed that the new method allows us to capture more data (median = 6.15 hours of awake behavior) than is typical with video methods (1-2 hours) while simultaneously reducing the human labor needed to annotate it.
+Ultimately, age differences in body position mirrored past findings that employed other methods, suggesting that the end product is suitable for describing developmental changes in motor behavior.
+
+\hypertarget{accurate-results-in-challenging-circumstances}{%
+\subsection{Accurate Results in Challenging Circumstances}\label{accurate-results-in-challenging-circumstances}}
+
+How did classification in the current investigation compare to prior work (Airaksinen et al., 2022, 2020; Franchak et al., 2021)?
+Across body positions, median accuracy was 89.40\% for group models (Kappa = 76.80) and 0.93\% for individual models (Kappa = 0.85) during the proximal comparison.
+Although these accuracy and kappa values are slightly lower than in past work (accuracy 95-98\%, kappas .93-.95), we note that those past values were obtained under ideal circumstances.
+A researcher applied the sensors in the lab (Airaksinen et al., 2020; Franchak et al., 2021) or home (Airaksinen et al., 2022) and set the stage for how infants and caregivers should interact.
+For example, both Franchak et al.
+(2021) and Airaksinen et al.
+(2022) instructed the caregivers that the recorded sessions should mimic a ``playtime'', which would encourage more common activities that would be easier to classify (e.g., crawling and sitting on the floor) and discourage idiosyncratic and potentially more challenging activities (e.g., positioning in infant-specific furniture like exersaucers and walkers, sitting or lying for long periods while eating or nursing).
+Examining the timelines in Figure \ref{fig:timelines} shows how much variability there is between and within sessions in the earlier parts of the recording that were used for training and testing.
+With the exception of the brief guided session at the start of the recording (which ensured that every infant spent at least a short amount of time in each body position), the majority of the recorded behavior was truly natural and ``in the wild''.
+
+The challenge only increased in the distal comparison period.
+Since sessions were scheduled at convenient times to for the infant and caregiver, the transition from the proximal to distal period increased the odds that infants might need a nap, a snack/meal, or engage in a less typical activity (e.g., watching TV).
+Indeed, the distal comparison in Figure \ref{fig:exemplar-timeline} contains a long period (almost 50 minutes) where the infant sat in a high chair eating lunch.
+With such a long delay in time from when the sensor garment was first put on, there is an increasing chance that sensors moved from their ideal positions.
+Regardless of these challenges, accuracy in the distal comparison continued to be strong.
+Barring two outliers (which we will discuss further), agreement was high for group (\emph{r} = 0.92) and individual models (\emph{r} = 0.94).
+Even at a finer timescale, most errors within 10-minute bins were less than 1 minute in total for all five body positions.
+
+To put this in context, the reliability of human coders on the body position code was \textasciitilde90\%, putting model-predicted accuracy on a similar level to human coders.
+Acceptable inter-rater reliability in developmental psychology studies of motor behavior is greater than 80\%; both group and individual model results exceed that standard for most recorded sessions.
+Another comparison we can make is to the accuracy of the LENA®'s automated metrics.
+Correlations between human and model frequencies were \emph{r} = .698 for adult word count and \emph{r} = .649 for child word count (Cristia et al., 2020).
+Even including outliers, correlations between human and model-predicted body position were \emph{r} = 0.80 (group) and \emph{r} = 0.94 (individual), easily surpassing the LENA® device's most reliable metrics.
+
+Outlier participants give us more confidence that it actually works as intended.
+Disagreements limited to some edge cases, not all events/classes.
+Even w/ outliers (which are they on timelines), most of the rest of the data seems typical for the age group.
+Some outlier-ish data in age comparisons, but nothing out of the ordinary (hard to know if it's the model or the behavior)
+
+Nice to have video to confirm.
+Best practice in LENA is to still do some transcription yourself to see what cases it works.
+Still helpful to have other sources to triangulate against.
+Maybe don't use it for a holding detector.
+
+\hypertarget{group-and-individual-models}{%
+\subsection{Group and Individual Models}\label{group-and-individual-models}}
+
+No clear winner.
+Higher raw accuracy/kappa for individual models; some (but not all) better by-class performance for Kappa's.
+But group models better for prone; more sensitive for prone and for held; barring outliers; group has some better part 2 correlations depending on the class.
+Notably, worst outlier in aggregate age difference data was in an individual model.
+
+Big advantages of group model are that 1) trained on more data (so likely a bit more robust), 2) less dependent on getting the perfect training set, 3) between individual differences are due to behavior, not due to choice of model, 4) can apply to any infant, 5) can apply to an infant for whom we have no video.
+Individual models might capture some idiosyncracies of movement (or device use) that are hard to get from big model that homogenizes behavior.
+Individual models are a nice choice for rapid prototyping/piloting/feasibility.
+Can get proof of concept on a single participant or a new kind of code (loc; stroller use; etc.).
+Clinical cases too.
+
+Tell some things we did but didn't report.
+Tried modeling each age separately but there was no overall improvement in performance.
+Categories are so broad and include less/more mature forms of behavior.
+Not independent sitting, but all kinds of sitting.
+Tried other ML algorithms.
+Tried fewer sensors.
+
+\hypertarget{what-does-this-get-us}{%
+\subsection{What does this get us?}\label{what-does-this-get-us}}
+
+Other things look a bit wacky, but within-day heterogeniety is the point.
+Shouldn't be surprised to find it.
+Some can confirm on video.
+Activity are a big shaper of behavior, have enough behavior to actually see it.
+
+See how other aspects of behavior align w/ movement in real time. See how new motor skills relate to chg in daily behavior.
+
+Efficiency. Hours saved.
+
+More easily shareable data (non-identifiable, text based)
+
+\hypertarget{limitations}{%
+\subsection{Limitations}\label{limitations}}
+
+\begin{itemize}
+\tightlist
+\item
+  No validation at the end of the day (practically this would be hard to do, syncing a challenge)
+\item
+  Limited to within the home. Driving/strollers are motion signals we don't have training data for, but this could be added in the future.
+\item
+  Relying on CG reports of naps/diaper changes.
+\item
+  Bias in training data
+\end{itemize}
+
+\hypertarget{conclusions}{%
+\subsection{Conclusions}\label{conclusions}}
 

### 2023-03-15 manuscript.Rmd
more discussion

@@
 How did classification in the current investigation compare to prior work [@FranchakScott2021; @AiraksinenRasanen2020; @AiraksinenGallen2022]?
-Across body positions, median accuracy was `r metrics_lookup["Overall Accuracy", "median_Group"]*100`% for group models (Kappa = `r metrics_lookup["Kappa", "median_Group"]*100`) and `r metrics_lookup["Overall Accuracy", "median_Individual"]`% for individual models (Kappa = `r metrics_lookup["Kappa", "median_Individual"]`) during the proximal comparison.
+Across body positions, median accuracy was `r metrics_lookup["Overall Accuracy", "median_Group"]*100`% for group models (Kappa = `r metrics_lookup["Kappa", "median_Group"]`) and `r metrics_lookup["Overall Accuracy", "median_Individual"]*100`% for individual models (Kappa = `r metrics_lookup["Kappa", "median_Individual"]`) during the proximal comparison.
 Although these accuracy and kappa values are slightly lower than in past work (accuracy 95-98%, kappas .93-.95), we note that those past values were obtained under ideal circumstances.
@@
 Correlations between human and model frequencies were *r* = .698 for adult word count and *r* = .649 for child word count [@CristiaLavechin2020].
-Even including outliers, correlations between human and model-predicted body position were *r* = `r group_agree_cor` (group) and *r* = `r split_agree_cor` (individual), easily surpassing the LENA&reg; device's most reliable metrics.
+Even including outliers, correlations between human and model-predicted body position were *r* = `r group_agree_cor` (group) and *r* = `r split_agree_cor` (individual), stronger than the LENA&reg; automated counts.
+Note, however, that the LENA&reg; validation study used independently-collected data, whereas we are measuring performance for infants whose earlier data comprised the training set.
+More work is needed to know how well group models would predict a different set of infants' behavior in a different environmental and/or cultural context.
 <!-- Reliability standard? -->
 
-Outlier participants give us more confidence that it actually works as intended.
-Disagreements limited to some edge cases, not all events/classes.
-Even w/ outliers (which are they on timelines), most of the rest of the data seems typical for the age group.
-Some outlier-ish data in age comparisons, but nothing out of the ordinary (hard to know if it's the model or the behavior)
-
-Nice to have video to confirm.
-Best practice in LENA is to still do some transcription yourself to see what cases it works.
-Still helpful to have other sources to triangulate against.
-Maybe don't use it for a holding detector.
+In their validation study, @CristiaLavechin2020 make the recommendation that researchers using LENA&reg; manually annotate a portion of data collected for their particular use case to ensure that automated LENA&reg; metrics are reliable for the participants and environment pertinent to the research question.
+Similarly, collecting some video data for each participant in future long-form motor behavior recordings provides a way to check reliability and identify errant predictions.
+Although predictions were accurate for most sessions, we noted two outliers for whom model predictions failed to match human annotations.
+Fortunately, we could reference video recordings to understand why the errors occurred.
+Idiosyncrasies in activities and device use make it challenging to decide *a priori* whether odd looking classifications are possible.
+For example, younger infant \#8 on Figure \@ref(fig:timelines) spent an unusual amount of time upright at an age where infants cannot yet stand.
+Inspecting the video revealed that the infant spent long periods of time suspended in a jumper that supported their body in an upright position.
+Although it is not feasible to collect large amounts of video data to check model predictions, interviewing caregivers about common activities and devices may provide a way to understand unexpected predictions.
+
+Counterintuitively, the two outlier participants increased our confidence that the body position classification method works as intended.
+At first glance, seeing low agreement rates for two sessions would suggest that the models perform poorly at predicting some *infants'* behavior.
+Instead, we found that for those infants, confusion was restricted to two positions during a single, long-lasting event and the other three body positions continued to be classified correctly.
+In other words, the models failed to predict a particular *event* for those two infants.
+Unfortunately, those events lasted a long time.
+In the case of confusion between supine and sitting for the infant who was reclined, the confusion likely stems from the difficult of categorizing behaviors that do not always fit neatly into discrete classes.
+Even to a human observer, there are often "edge" cases when scoring infant behavior, which is why inter-rater reliability is never perfect.
 
@@
 See how other aspects of behavior align w/ movement in real time. See how new motor skills relate to chg in daily behavior. 
+LINK TO AGE DIFFS. Some outlier-ish data in age comparisons, but nothing out of the ordinary (hard to know if it's the model or the behavior)
 

### 2023-03-15 manuscript.tex
more discussion

@@
 How did classification in the current investigation compare to prior work (Airaksinen et al., 2022, 2020; Franchak et al., 2021)?
-Across body positions, median accuracy was 89.40\% for group models (Kappa = 76.80) and 0.93\% for individual models (Kappa = 0.85) during the proximal comparison.
+Across body positions, median accuracy was 89.40\% for group models (Kappa = 0.77) and 93.30\% for individual models (Kappa = 0.85) during the proximal comparison.
 Although these accuracy and kappa values are slightly lower than in past work (accuracy 95-98\%, kappas .93-.95), we note that those past values were obtained under ideal circumstances.
@@
 Acceptable inter-rater reliability in developmental psychology studies of motor behavior is greater than 80\%; both group and individual model results exceed that standard for most recorded sessions.
-Another comparison we can make is to the accuracy of the LENA®'s automated metrics.
+Another comparison we can make is to the accuracy of the LENA®`s automated metrics.
 Correlations between human and model frequencies were \emph{r} = .698 for adult word count and \emph{r} = .649 for child word count (Cristia et al., 2020).
-Even including outliers, correlations between human and model-predicted body position were \emph{r} = 0.80 (group) and \emph{r} = 0.94 (individual), easily surpassing the LENA® device's most reliable metrics.
-
-Outlier participants give us more confidence that it actually works as intended.
-Disagreements limited to some edge cases, not all events/classes.
-Even w/ outliers (which are they on timelines), most of the rest of the data seems typical for the age group.
-Some outlier-ish data in age comparisons, but nothing out of the ordinary (hard to know if it's the model or the behavior)
-
-Nice to have video to confirm.
-Best practice in LENA is to still do some transcription yourself to see what cases it works.
-Still helpful to have other sources to triangulate against.
-Maybe don't use it for a holding detector.
+Even including outliers, correlations between human and model-predicted body position were \emph{r} = 0.80 (group) and \emph{r} = 0.94 (individual), stronger than the LENA® automated counts.
+Note, however, that the LENA® validation study used independently-collected data, whereas we are measuring performance for infants whose earlier data comprised the training set.
+More work is needed to know how well group models would predict a different set of infants' behavior in a different environmental and/or cultural context.
+
+In their validation study, Cristia et al.
+(2020) make the recommendation that researchers using LENA® manually annotate a portion of data collected for their particular use case to ensure that automated LENA® metrics are reliable for the participants and environment pertinent to the research question.
+Similarly, collecting some video data for each participant in future long-form motor behavior recordings provides a way to check reliability and identify errant predictions.
+Although predictions were accurate for most sessions, we noted two outliers for whom model predictions failed to match human annotations.
+Fortunately, we could reference video recordings to understand why the errors occurred.
+Idiosyncrasies in activities and device use make it challenging to decide \emph{a priori} whether odd looking classifications are possible.
+For example, younger infant \#8 on Figure \ref{fig:timelines} spent an unusual amount of time upright at an age where infants cannot yet stand.
+Inspecting the video revealed that the infant spent long periods of time suspended in a jumper that supported their body in an upright position.
+Although it is not feasible to collect large amounts of video data to check model predictions, interviewing caregivers about common activities and devices may provide a way to understand unexpected predictions.
+
+Counterintuitively, the two outlier participants increased our confidence that the body position classification method works as intended.
+At first glance, seeing low agreement rates for two sessions would suggest that the models perform poorly at predicting some \emph{infants'} behavior.
+Instead, we found that for those infants, confusion was restricted to two positions during a single, long-lasting event and the other three body positions continued to be classified correctly.
+In other words, the models failed to predict a particular \emph{event} for those two infants.
+Unfortunately, those events lasted a long time.
+In the case of confusion between supine and sitting for the infant who was reclined, the confusion likely stems from the difficult of categorizing behaviors that do not always fit neatly into discrete classes.
+Even to a human observer, there are often ``edge'' cases when scoring infant behavior, which is why inter-rater reliability is never perfect.
 
@@
 See how other aspects of behavior align w/ movement in real time. See how new motor skills relate to chg in daily behavior.
+LINK TO AGE DIFFS. Some outlier-ish data in age comparisons, but nothing out of the ordinary (hard to know if it's the model or the behavior)
 

### 2023-03-16 manuscript.Rmd
more discussion

@@
 Examining full recordings showed that the new method allows us to capture more data (median = `r apa_num(median(coverage$hours_recorded))` hours of awake behavior) than is typical with video methods (1-2 hours) while simultaneously reducing the human labor needed to annotate it.
-Ultimately, age differences in body position mirrored past findings that employed other methods, suggesting that the end product is suitable for describing developmental changes in motor behavior.
+Ultimately, age differences in body position mirrored past findings that employed other methods, suggesting that the outcome of full-day body position recordings is suitable for describing developmental changes in motor behavior.
 
@@
 At first glance, seeing low agreement rates for two sessions would suggest that the models perform poorly at predicting some *infants'* behavior.
-Instead, we found that for those infants, confusion was restricted to two positions during a single, long-lasting event and the other three body positions continued to be classified correctly.
-In other words, the models failed to predict a particular *event* for those two infants.
-Unfortunately, those events lasted a long time.
-In the case of confusion between supine and sitting for the infant who was reclined, the confusion likely stems from the difficult of categorizing behaviors that do not always fit neatly into discrete classes.
-Even to a human observer, there are often "edge" cases when scoring infant behavior, which is why inter-rater reliability is never perfect.
-
-## Group and Individual Models
-
-No clear winner.
-Higher raw accuracy/kappa for individual models; some (but not all) better by-class performance for Kappa's.
-But group models better for prone; more sensitive for prone and for held; barring outliers; group has some better part 2 correlations depending on the class.
-Notably, worst outlier in aggregate age difference data was in an individual model.
-
-Big advantages of group model are that 1) trained on more data (so likely a bit more robust), 2) less dependent on getting the perfect training set, 3) between individual differences are due to behavior, not due to choice of model, 4) can apply to any infant, 5) can apply to an infant for whom we have no video.
-Individual models might capture some idiosyncracies of movement (or device use) that are hard to get from big model that homogenizes behavior.
-Individual models are a nice choice for rapid prototyping/piloting/feasibility.
-Can get proof of concept on a single participant or a new kind of code (loc; stroller use; etc.).
-Clinical cases too.
-
-Tell some things we did but didn't report.
-Tried modeling each age separately but there was no overall improvement in performance.
-Categories are so broad and include less/more mature forms of behavior.
-Not independent sitting, but all kinds of sitting.
-Tried other ML algorithms.
-Tried fewer sensors.
+Instead, we found that for those infants, errors were restricted to two positions during a single, long-lasting event, while the other three body positions continued to be classified correctly.
+In other words, the models failed to predict a particular *event* for each of the two infants.
+Unfortunately, those events lasted a long time leading to poor overall agreement.
+In the case of confusion between supine and sitting for the infant who was reclined, the confusion likely stems from the difficult of categorizing behaviors that do not always fit neatly into discrete classes (e.g., continuous body position angles categorized as either supine or sitting).
+
+## Benefits and Drawbacks of Different Modeling Approaches
+
+Throughout the paper, we compared results from two modeling approaches: group models that included all participants' data in the training set versus individual models that used only one participant's data.
+When considering different metrics, testing periods, body positions, as well as the logistical benefits of each, there is no clear winner.
+Below, we discuss the pros and cons to each approach so that researchers might decide what works best for their intended application.
+
+Overall accuracy and kappa values were better in individual models compared with group models when collapsed across body positions in the proximal period.
+However, within-class performance did not always favor individual models---prone predictions were better for group models, and held predictions were almost identical.
+In the distal comparison (after removing outliers), overall correlations were nearly identical.
+Within classes, group models had an advantage for prone, upright, and held, whereas individual models had an advantage for sitting and supine.
+Possibly, individual models are better suited for capturing unique aspects of the devices used for sitting and reclining that lead to better performance in those classes.
+Finally, age effects (either by group or by age correlation) were nearly identical across the two methods, suggesting that either method would lead to the correct conclusion regarding developmental changes in body position.
+Outliers were present in both models' predictions: group models produced three sessions with the worst overall accuracy, but individual models produced the single worst kappa.
+The outliers for the distal comparison period appeared in both models' predictions.
+For aggregated full-day body position, the most blatant outlier (a young infant held > 60% of the day) resulted from an individual model.
+The choice of model might depend on which behavior is most important to capture for a given research question.
+
+Aside from differences in validity, researchers may favor one modeling approach based on logistical concerns.
+Group models have several major advantages.
+First, by definition they are trained on more data, which might make predictions more consistent.
+Second, group models remove the need to get optimal training data from each participant.
+It can be difficult to elicit every behavior of interest in each infant; individual models require that every behavior that will be later predicted was displayed by each individual infant.
+In fact, group models remove the need to get *any* training data for a particular infant, meaning that they can be applied to any infant regardless of whether they have training data.
+Finally, applying a group model means that individual differences between infants are due to their behavior being classified in a consistent way, rather than having a separate set of rules for predicting each individual.
+However, given the variety in infant behavior, individual models are more agile in that they can capture the unique and unexpected.
+Likewise, individual models can be an excellent choice for rapid prototyping and pilot testing.
+A researcher can get proof-of-concept data from a single participant or a new kind of classification scheme (i.e., locomotion, restrain in device) without annotating an entire sample.
+Individual models may also be applied in clinical cases where infants vary widely in their motor abilities and/or use of assistive devices.
+
+Beyond group and individual models, there were many degrees of freedom in our choice of how best to model body position.
+For the sake of brevity, we did not report every possible variation, only the best approach.
+However, our openly shared data mean that researchers can experiment with different sets of motion features, different sets of training data, different machine learning algorithms, and different sets of hyperparameters.
+As in past work [@AiraksinenRasanen2020], performance is best when using all four sensors and degrades when using any subset of two (i.e., hips only, left leg only).
+We tested but did not report models that used either only accelerometer data and only gyroscope data; in both cases, performance is degraded compared with models that use both types of motion.
+We also tried different machine learning algorithms (XGBoost, Tabnet), but found that random forest models performed the best.
+Finally, we tried modeling each age group separately, but found no boost in performance by tailoring to the particular age group.
+<!-- DATA CITE -->
 

### 2023-03-16 manuscript.tex
more discussion

@@
 Examining full recordings showed that the new method allows us to capture more data (median = 6.15 hours of awake behavior) than is typical with video methods (1-2 hours) while simultaneously reducing the human labor needed to annotate it.
-Ultimately, age differences in body position mirrored past findings that employed other methods, suggesting that the end product is suitable for describing developmental changes in motor behavior.
+Ultimately, age differences in body position mirrored past findings that employed other methods, suggesting that the outcome of full-day body position recordings is suitable for describing developmental changes in motor behavior.
 
@@
 At first glance, seeing low agreement rates for two sessions would suggest that the models perform poorly at predicting some \emph{infants'} behavior.
-Instead, we found that for those infants, confusion was restricted to two positions during a single, long-lasting event and the other three body positions continued to be classified correctly.
-In other words, the models failed to predict a particular \emph{event} for those two infants.
-Unfortunately, those events lasted a long time.
-In the case of confusion between supine and sitting for the infant who was reclined, the confusion likely stems from the difficult of categorizing behaviors that do not always fit neatly into discrete classes.
-Even to a human observer, there are often ``edge'' cases when scoring infant behavior, which is why inter-rater reliability is never perfect.
-
-\hypertarget{group-and-individual-models}{%
-\subsection{Group and Individual Models}\label{group-and-individual-models}}
-
-No clear winner.
-Higher raw accuracy/kappa for individual models; some (but not all) better by-class performance for Kappa's.
-But group models better for prone; more sensitive for prone and for held; barring outliers; group has some better part 2 correlations depending on the class.
-Notably, worst outlier in aggregate age difference data was in an individual model.
-
-Big advantages of group model are that 1) trained on more data (so likely a bit more robust), 2) less dependent on getting the perfect training set, 3) between individual differences are due to behavior, not due to choice of model, 4) can apply to any infant, 5) can apply to an infant for whom we have no video.
-Individual models might capture some idiosyncracies of movement (or device use) that are hard to get from big model that homogenizes behavior.
-Individual models are a nice choice for rapid prototyping/piloting/feasibility.
-Can get proof of concept on a single participant or a new kind of code (loc; stroller use; etc.).
-Clinical cases too.
-
-Tell some things we did but didn't report.
-Tried modeling each age separately but there was no overall improvement in performance.
-Categories are so broad and include less/more mature forms of behavior.
-Not independent sitting, but all kinds of sitting.
-Tried other ML algorithms.
-Tried fewer sensors.
+Instead, we found that for those infants, errors were restricted to two positions during a single, long-lasting event, while the other three body positions continued to be classified correctly.
+In other words, the models failed to predict a particular \emph{event} for each of the two infants.
+Unfortunately, those events lasted a long time leading to poor overall agreement.
+In the case of confusion between supine and sitting for the infant who was reclined, the confusion likely stems from the difficult of categorizing behaviors that do not always fit neatly into discrete classes (e.g., continuous body position angles categorized as either supine or sitting).
+
+\hypertarget{benefits-and-drawbacks-of-different-modeling-approaches}{%
+\subsection{Benefits and Drawbacks of Different Modeling Approaches}\label{benefits-and-drawbacks-of-different-modeling-approaches}}
+
+Throughout the paper, we compared results from two modeling approaches: group models that included all participants' data in the training set versus individual models that used only one participant's data.
+When considering different metrics, testing periods, body positions, as well as the logistical benefits of each, there is no clear winner.
+Below, we discuss the pros and cons to each approach so that researchers might decide what works best for their intended application.
+
+Overall accuracy and kappa values were better in individual models compared with group models when collapsed across body positions in the proximal period.
+However, within-class performance did not always favor individual models---prone predictions were better for group models, and held predictions were almost identical.
+In the distal comparison (after removing outliers), overall correlations were nearly identical.
+Within classes, group models had an advantage for prone, upright, and held, whereas individual models had an advantage for sitting and supine.
+Possibly, individual models are better suited for capturing unique aspects of the devices used for sitting and reclining that lead to better performance in those classes.
+Finally, age effects (either by group or by age correlation) were nearly identical across the two methods, suggesting that either method would lead to the correct conclusion regarding developmental changes in body position.
+Outliers were present in both models' predictions: group models produced three sessions with the worst overall accuracy, but individual models produced the single worst kappa.
+The outliers for the distal comparison period appeared in both models' predictions.
+For aggregated full-day body position, the most blatant outlier (a young infant held \textgreater{} 60\% of the day) resulted from an individual model.
+The choice of model might depend on which behavior is most important to capture for a given research question.
+
+Aside from differences in validity, researchers may favor one modeling approach based on logistical concerns.
+Group models have several major advantages.
+First, by definition they are trained on more data, which might make predictions more consistent.
+Second, group models remove the need to get optimal training data from each participant.
+It can be difficult to elicit every behavior of interest in each infant; individual models require that every behavior that will be later predicted was displayed by each individual infant.
+In fact, group models remove the need to get \emph{any} training data for a particular infant, meaning that they can be applied to any infant regardless of whether they have training data.
+Finally, applying a group model means that individual differences between infants are due to their behavior being classified in a consistent way, rather than having a separate set of rules for predicting each individual.
+However, given the variety in infant behavior, individual models are more agile in that they can capture the unique and unexpected.
+Likewise, individual models can be an excellent choice for rapid prototyping and pilot testing.
+A researcher can get proof-of-concept data from a single participant or a new kind of classification scheme (i.e., locomotion, restrain in device) without annotating an entire sample.
+Individual models may also be applied in clinical cases where infants vary widely in their motor abilities and/or use of assistive devices.
+
+Beyond group and individual models, there were many degrees of freedom in our choice of how best to model body position.
+For the sake of brevity, we did not report every possible variation, only the best approach.
+However, our openly shared data mean that researchers can experiment with different sets of motion features, different sets of training data, different machine learning algorithms, and different sets of hyperparameters.
+As in past work (Airaksinen et al., 2020), performance is best when using all four sensors and degrades when using any subset of two (i.e., hips only, left leg only).
+We tested but did not report models that used either only accelerometer data and only gyroscope data; in both cases, performance is degraded compared with models that use both types of motion.
+We also tried different machine learning algorithms (XGBoost, Tabnet), but found that random forest models performed the best.
+Finally, we tried modeling each age group separately, but found no boost in performance by tailoring to the particular age group.
 

### 2023-03-16 manuscript.Rmd
get new research possibilities going

@@
 We also tried different machine learning algorithms (XGBoost, Tabnet), but found that random forest models performed the best.
-Finally, we tried modeling each age group separately, but found no boost in performance by tailoring to the particular age group.
+Finally, we tried modeling each age group separately, but found no benefit to performance by tailoring to the particular age group.
 <!-- DATA CITE -->
 
-## What does this get us?
+## Novel Research Possibilities
+
+Long-form recordings of motor behavior bring about new research possibilities.
+Although past work using video [@Karasik2011; @ChenSchneider2022; @HerzbergFletcher2021] produced real-time data, such data were limited to a relatively small part of the infants' day.
+And although survey methods [@Survey] could capture moments scattered throughout the day, they do not produce real-time data.
+The combination of real-time, full-day data about infant motor behavior is unprecedented and offers new opportunities for understanding infants' everyday experiences.
+Collecting dense data over the entire variety of daily experiences helps to more accurately measure infants' experiences in aggregate (as in Figure \@ref(fig:age)) without biasing results from a particular type of activity (e.g., play).
+But as Figure \@ref(fig:timelines) shows, aggregating responses masks the rich information that is available from the entire individual timelines provided by long-form recordings.
+
+In the past decade, developmental scientists have discovered that the distributional structure of infants' experiences matter for learning [@ClerkinHart2017; @KachergisYu2017; @RazAbney2019].
 
@@
 Activity are a big shaper of behavior, have enough behavior to actually see it.
+LINK TO AGE DIFFS. Some outlier-ish data in age comparisons, but nothing out of the ordinary (hard to know if it's the model or the behavior)
 
 See how other aspects of behavior align w/ movement in real time. See how new motor skills relate to chg in daily behavior.
-LINK TO AGE DIFFS. Some outlier-ish data in age comparisons, but nothing out of the ordinary (hard to know if it's the model or the behavior)
 

### 2023-03-16 manuscript.Rmd
up to limitations section

@@
 Collecting dense data over the entire variety of daily experiences helps to more accurately measure infants' experiences in aggregate (as in Figure \@ref(fig:age)) without biasing results from a particular type of activity (e.g., play).
-But as Figure \@ref(fig:timelines) shows, aggregating responses masks the rich information that is available from the entire individual timelines provided by long-form recordings.
-
-In the past decade, developmental scientists have discovered that the distributional structure of infants' experiences matter for learning [@ClerkinHart2017; @KachergisYu2017; @RazAbney2019].
-
-Other things look a bit wacky, but within-day heterogeniety is the point.
-Shouldn't be surprised to find it.
-Some can confirm on video.
-Activity are a big shaper of behavior, have enough behavior to actually see it.
-LINK TO AGE DIFFS. Some outlier-ish data in age comparisons, but nothing out of the ordinary (hard to know if it's the model or the behavior)
-
-See how other aspects of behavior align w/ movement in real time. See how new motor skills relate to chg in daily behavior.
-
-Efficiency. Hours saved.
-
-More easily shareable data (non-identifiable, text based)
+Long-form recordings also have the potential measure clinically-relevant outcomes, such as whether a physical therapy intervention to improve infant sitting skill leads to increases in everyday sitting time.
+But as Figure \@ref(fig:timelines) shows, aggregating responses masks the rich temporal structure contained within long-form recordings.
+
+In the past decade, developmental scientists have discovered that distributional information about infants' experiences matter for how infants learn [@ClerkinHart2017; @KachergisYu2017; @RazAbney2019].
+Skewed distributions that favor the repetition of a small subset of experiences facilitate learning.
+For example, recordings from wearable head cameras in the home indicate that infants see a small subset of objects with high frequency [@ClerkinHart2017], but most objects are seen infrequently.
+Heterogeneity is also found in how experiences are distributed in time---burstiness and clustering are seen in infants' daily experiences seeing objects and hearing music.
+Long-form motor recordings provide a novel opportunity to measure the temporal structure of motor experiences.
+How skewed are infants' experiences with different motor behaviors, and how are they clustered in time?
+Linking the distributional structure of a particular motor experience in a real day, such as time spent sitting, might be predicted by sitting skill and/or might infants' future sitting ability.
+<!-- CITES FOR BURSTINESS CASILLAS FAUSEY -->
+
+Moreover, real-time recording of motor behavior provides a way to measure how infant motor behavior links to other types of experiences in the moment in daily life.
+As we reviewed, @MalachowskiSalo2023 used a combination of LENA&reg; and EMA surveys to find that infants heard less adult speech when restrained in seating devices.
+Combining LENA&reg; and long-form motor recordings can take this a step further by measuring more precisely how speech and body position covary within a day.
+Synchronizing long form motor recordings to other continuous monitors, such as wearable ECG [@WassPhillips2022] can provide information about how infant arousal relates to motor behavior.
+"TotTag"---a wearable device for measuring infant-caregiver distance [@SaloPannuto2022]---or caregiver worn activity trackers/physiology monitors could be paired with body position classification to better understand how moment-to-moment interpersonal dynamics relate to infant motor behavior.
+
+What underpins these novel research applications is that body position classifications can be applied automatically at scale.
+Annotating a video corpus of infant behavior from a moderately sized sample---such as 40 infants recorded for 2 hours each in 2 separate visits [@HerzbergFletcher2021]---is incredibly labor intensive.
+Annotating a larger, more representative sample of *hundreds of infants* while simultaneously scoring *full-day data* would be prohibitive.
+With sufficiently accurate group models, researchers could annotate a moderate video sample to then apply the model to hundreds of full-day recordings.
+The added cost of more recording time is low, meaning that future studies could sample across multiple days of behavior to better understand intra- versus inter-day variability.
+Prior diary methods show that infants inconsistently display new motor skills on a day-to-day basis [@AdolphRobinson2008; @AdolphRobinson2011].
+Multi-day recordings could further uncover how infants' experiences vary over timescales from seconds to days.
+
+<!-- More easily shareable data (non-identifiable, text based) -->
 

### 2023-03-16 manuscript.tex
up to limitations section

@@
 We also tried different machine learning algorithms (XGBoost, Tabnet), but found that random forest models performed the best.
-Finally, we tried modeling each age group separately, but found no boost in performance by tailoring to the particular age group.
-
-\hypertarget{what-does-this-get-us}{%
-\subsection{What does this get us?}\label{what-does-this-get-us}}
-
-Other things look a bit wacky, but within-day heterogeniety is the point.
-Shouldn't be surprised to find it.
-Some can confirm on video.
-Activity are a big shaper of behavior, have enough behavior to actually see it.
-
-See how other aspects of behavior align w/ movement in real time. See how new motor skills relate to chg in daily behavior.
-LINK TO AGE DIFFS. Some outlier-ish data in age comparisons, but nothing out of the ordinary (hard to know if it's the model or the behavior)
-
-Efficiency. Hours saved.
+Finally, we tried modeling each age group separately, but found no benefit to performance by tailoring to the particular age group.
+
+\hypertarget{novel-research-possibilities}{%
+\subsection{Novel Research Possibilities}\label{novel-research-possibilities}}
+
+Long-form recordings of motor behavior bring about new research possibilities.
+Although past work using video (Chen et al., 2022; Herzberg et al., 2021; Karasik et al., 2011) produced real-time data, such data were limited to a relatively small part of the infants' day.
+And although survey methods (Franchak, 2019) could capture moments scattered throughout the day, they do not produce real-time data.
+The combination of real-time, full-day data about infant motor behavior is unprecedented and offers new opportunities for understanding infants' everyday experiences.
+Collecting dense data over the entire variety of daily experiences helps to more accurately measure infants' experiences in aggregate (as in Figure \ref{fig:age}) without biasing results from a particular type of activity (e.g., play).
+Long-form recordings also have the potential measure clinically-relevant outcomes, such as whether a physical therapy intervention to improve infant sitting skill leads to increases in everyday sitting time.
+But as Figure \ref{fig:timelines} shows, aggregating responses masks the rich temporal structure contained within long-form recordings.
+
+In the past decade, developmental scientists have discovered that distributional information about infants' experiences matter for how infants learn (Clerkin et al., 2017; Kachergis et al., 2017; Raz et al., 2019).
+Skewed distributions that favor the repetition of a small subset of experiences facilitate learning.
+For example, recordings from wearable head cameras in the home indicate that infants see a small subset of objects with high frequency (Clerkin et al., 2017), but most objects are seen infrequently.
+Heterogeneity is also found in how experiences are distributed in time---burstiness and clustering are seen in infants' daily experiences seeing objects and hearing music.
+Long-form motor recordings provide a novel opportunity to measure the temporal structure of motor experiences.
+How skewed are infants' experiences with different motor behaviors, and how are they clustered in time?
+Linking the distributional structure of a particular motor experience in a real day, such as time spent sitting, might be predicted by sitting skill and/or might infants' future sitting ability.
+
+Moreover, real-time recording of motor behavior provides a way to measure how infant motor behavior links to other types of experiences in the moment in daily life.
+As we reviewed, Malachowski et al.
+(2023) used a combination of LENA® and EMA surveys to find that infants heard less adult speech when restrained in seating devices.
+Combining LENA® and long-form motor recordings can take this a step further by measuring more precisely how speech and body position covary within a day.
+Synchronizing long form motor recordings to other continuous monitors, such as wearable ECG (Wass et al., 2022) can provide information about how infant arousal relates to motor behavior.
+``TotTag''---a wearable device for measuring infant-caregiver distance (Salo et al., 2022)---or caregiver worn activity trackers/physiology monitors could be paired with body position classification to better understand how moment-to-moment interpersonal dynamics relate to infant motor behavior.
+
+What underpins these novel research applications is that body position classifications can be applied automatically at scale.
+Annotating a video corpus of infant behavior from a moderately sized sample---such as 40 infants recorded for 2 hours each in 2 separate visits (Herzberg et al., 2021)---is incredibly labor intensive.
+Annotating a larger, more representative sample of \emph{hundreds of infants} while simultaneously scoring \emph{full-day data} would be prohibitive.
+With sufficiently accurate group models, researchers could annotate a moderate video sample to then apply the model to hundreds of full-day recordings.
+The added cost of more recording time is low, meaning that future studies could sample across multiple days of behavior to better understand intra- versus inter-day variability.
+Prior diary methods show that infants inconsistently display new motor skills on a day-to-day basis (Adolph et al., 2008; Adolph \& Robinson, 2011).
+Multi-day recordings could further uncover how infants' experiences vary over timescales from seconds to days.
 
@@
 \begin{CSLReferences}{1}{0}
+\leavevmode\vadjust pre{\hypertarget{ref-AdolphRobinson2011}{}}%
+Adolph, K. E., \& Robinson, S. R. (2011). Sampling development. \emph{Journal of Cognition and Development}, \emph{12}, 411--423. \url{https://doi.org/10.1080/15248372.2011.608190}
+
+\leavevmode\vadjust pre{\hypertarget{ref-AdolphRobinson2008}{}}%
+Adolph, K.
+E., Robinson, S.
+R., Young, J.
+W., \& Gill-Alvarez, F.
+(2008).
+What is the shape of developmental change?
+\emph{Psychological Review}, \emph{115}, 527--543.
+\url{https://doi.org/10.1037/0033-295X.115.3.527}
+
 \leavevmode\vadjust pre{\hypertarget{ref-AdolphCTL2014}{}}%
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-ClerkinHart2017}{}}%
+Clerkin, E.
+M., Hart, E., Rehg, J.
+M., Yu, C., \& Smith, L.
+B.
+(2017).
+Real-world visual statistics and infants' first-learned object names.
+\emph{Philosophical Transactions of the Royal Society B}, \emph{372}, 20160055.
+
 \leavevmode\vadjust pre{\hypertarget{ref-CliffReilly2009}{}}%
@@
 \emph{Child Development}, \emph{93}(1), 150--164.
+
+\leavevmode\vadjust pre{\hypertarget{ref-KachergisYu2017}{}}%
+Kachergis, G., Yu, C., \& Shiffrin, R. M. (2017). A bootstrapping model of frequency and context effects in word learning. \emph{Cognitive Science}, \emph{41}(3), 590--622.
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-RazAbney2019}{}}%
+Raz, H.
+K., Abney, D.
+H., Crandall, D., Yu, C., \& Smith, L.
+B.
+(2019).
+How do infants start learning object names in a sea of clutter?
+\emph{Annual Conference of the Cognitive Science Society}, 521--526.
+
 \leavevmode\vadjust pre{\hypertarget{ref-RenDing2016}{}}%
@@
 \emph{Applied Intelligence}, \emph{45}(2), 512--529.
+
+\leavevmode\vadjust pre{\hypertarget{ref-SaloPannuto2022}{}}%
+Salo, V.
+C., Pannuto, P., Hedgecock, W., Biri, A., Russo, D.
+A., Piersiak, H.
+A., \& Humphreys, K.
+L.
+(2022).
+Measuring naturalistic proximity as a window into caregiver--child interaction patterns.
+\emph{Behavior Research Methods}, \emph{54}(4), 1580--1594.
+\url{https://doi.org/10.3758/s13428-021-01681-8}
 

### 2023-03-16 manuscript.Rmd
FULL DRAFT

@@
 As we reviewed, @MalachowskiSalo2023 used a combination of LENA&reg; and EMA surveys to find that infants heard less adult speech when restrained in seating devices.
-Combining LENA&reg; and long-form motor recordings can take this a step further by measuring more precisely how speech and body position covary within a day.
+Combining LENA&reg; and long-form motor recordings can take this a step further by measuring more precisely how speech and body position co-vary within a day.
 Synchronizing long form motor recordings to other continuous monitors, such as wearable ECG [@WassPhillips2022] can provide information about how infant arousal relates to motor behavior.
@@
 ## Limitations
-- No validation at the end of the day (practically this would be hard to do, syncing a challenge)
-- Limited to within the home. Driving/strollers are motion signals we don't have training data for, but this could be added in the future. 
-- Relying on CG reports of naps/diaper changes. 
-- Bias in training data
+
+We acknowledge several limitations in the current approach.
+First, a stronger test of full-day accuracy would be to compare accuracy at the start of the recording period to accuracy at the end, rather than using the distal comparison video that followed the first 90 minutes of the study.
+As we described in the procedure section, synchronizing video cameras and inertial sensors required capturing a synchronization point on video.
+We completed this procedure prior to giving equipment to caregivers so that they were not responsible for synchronization; in fact, caregivers never touched a button on the video cameras.
+An end-of-day video would require a lot on the part of caregivers, and any mistakes in the procedure would lead to misaligned and unusable data.
+Given that most recordings ended at infants' bedtimes, having an experimenter visit families in the evening to video record would be intrusive.
+
+Second, we relied on caregiver logs of infants' naps and times when sensors were removed to separate usable data (times when infants wore the sensors while awake) from unusable data.
+Although most caregivers were diligent about completing logs, there were a few cases that caregivers may have failed to report naps (younger infants \#1 and \#4).
+In the future, algorithms can be developed to automatically identify periods of sleep and times when sensors are off the body.
+Such algorithms exist for adults using wrist-worn sensors, however, we did not use them because they have not been validated for infant participants wearing sensors on hips/ankles.
+Differences in infants' positioning while asleep (held in caregivers arms, laying in cribs, reclined in strollers and carseats) might make classifying sleep more difficult compared with adults based on movement.
+<!-- NEED SLEEP AND WEAR TIME CITES> -->
+<!-- Algorithms for infant sleep? -->
+
+Third, although the goal was to collect fully naturalistic behavior, we opted to restrict recording to times when participants were in the home.
+Inertial sensors can travel with participants, and there is no doubt that infants could have worn the sensors out of the home on errands.
+However, we do not yet have training data to detect periods of transportation (e.g., moving in a car or in a stroller) that might add noise and/or lead to misclassification of body positions.
+Future work should aim to collect data in these scenarios so that body position can be captured in or out of the home.
 
 ## Conclusions
+
+In summary, the current study demonstrates the validity of long-form recordings of infant motor behavior in the home.
+Our analyses show that body position classifications---whether infants are supine, prone, sitting, upright, or held---are accurate immediately and even following a substantial delay.
+Most important, we find substantial agreement between human-coded and model-predicted body position for data collected during truly everyday, unsupervised activities that create the most challenging cases for automatic classification.
+In most cases, model prediction accuracy approached human reliability, suggesting that model predictions can be confidently used in analyses of full-day infant behavior.
+Examining the resulting corpus of `r apa_num(sum(coverage$hours_recorded))` hours of real-time infant body position showed the feasibility of capturing data covering the majority of infants' awake time and the variety of activities contained therein.
+A rudimentary analysis of aggregated data showed that full-day position estimates conformed to expectations about age differences in motor behavior.
+Future work employing the method can go beyond aggregated measures of behavior to uncover the temporal structure in infants' daily motor experiences.
 

### 2023-03-16 manuscript.tex
FULL DRAFT

@@
 (2023) used a combination of LENA® and EMA surveys to find that infants heard less adult speech when restrained in seating devices.
-Combining LENA® and long-form motor recordings can take this a step further by measuring more precisely how speech and body position covary within a day.
+Combining LENA® and long-form motor recordings can take this a step further by measuring more precisely how speech and body position co-vary within a day.
 Synchronizing long form motor recordings to other continuous monitors, such as wearable ECG (Wass et al., 2022) can provide information about how infant arousal relates to motor behavior.
@@
 
-More easily shareable data (non-identifiable, text based)
-
 \hypertarget{limitations}{%
@@
 
-\begin{itemize}
-\tightlist
-\item
-  No validation at the end of the day (practically this would be hard to do, syncing a challenge)
-\item
-  Limited to within the home. Driving/strollers are motion signals we don't have training data for, but this could be added in the future.
-\item
-  Relying on CG reports of naps/diaper changes.
-\item
-  Bias in training data
-\end{itemize}
+We acknowledge several limitations in the current approach.
+First, a stronger test of full-day accuracy would be to compare accuracy at the start of the recording period to accuracy at the end, rather than using the distal comparison video that followed the first 90 minutes of the study.
+As we described in the procedure section, synchronizing video cameras and inertial sensors required capturing a synchronization point on video.
+We completed this procedure prior to giving equipment to caregivers so that they were not responsible for synchronization; in fact, caregivers never touched a button on the video cameras.
+An end-of-day video would require a lot on the part of caregivers, and any mistakes in the procedure would lead to misaligned and unusable data.
+Given that most recordings ended at infants' bedtimes, having an experimenter visit families in the evening to video record would be intrusive.
+
+Second, we relied on caregiver logs of infants' naps and times when sensors were removed to separate usable data (times when infants wore the sensors while awake) from unusable data.
+Although most caregivers were diligent about completing logs, there were a few cases that caregivers may have failed to report naps (younger infants \#1 and \#4).
+In the future, algorithms can be developed to automatically identify periods of sleep and times when sensors are off the body.
+Such algorithms exist for adults using wrist-worn sensors, however, we did not use them because they have not been validated for infant participants wearing sensors on hips/ankles.
+Differences in infants' positioning while asleep (held in caregivers arms, laying in cribs, reclined in strollers and carseats) might make classifying sleep more difficult compared with adults based on movement.
+
+Third, although the goal was to collect fully naturalistic behavior, we opted to restrict recording to times when participants were in the home.
+Inertial sensors can travel with participants, and there is no doubt that infants could have worn the sensors out of the home on errands.
+However, we do not yet have training data to detect periods of transportation (e.g., moving in a car or in a stroller) that might add noise and/or lead to misclassification of body positions.
+Future work should aim to collect data in these scenarios so that body position can be captured in or out of the home.
 
@@
 \subsection{Conclusions}\label{conclusions}}
+
+In summary, the current study demonstrates the validity of long-form recordings of infant motor behavior in the home.
+Our analyses show that body position classifications---whether infants are supine, prone, sitting, upright, or held---are accurate immediately and even following a substantial delay.
+Most important, we find substantial agreement between human-coded and model-predicted body position for data collected during truly everyday, unsupervised activities that create the most challenging cases for automatic classification.
+In most cases, model prediction accuracy approached human reliability, suggesting that model predictions can be confidently used in analyses of full-day infant behavior.
+Examining the resulting corpus of 201.06 hours of real-time infant body position showed the feasibility of capturing data covering the majority of infants' awake time and the variety of activities contained therein.
+A rudimentary analysis of aggregated data showed that full-day position estimates conformed to expectations about age differences in motor behavior.
+Future work employing the method can go beyond aggregated measures of behavior to uncover the temporal structure in infants' daily motor experiences.
 

### 2023-03-24 manuscript.Rmd
Updated reliability

@@
 Interrater reliability was based on the proportion of video frames that the two coders chose the same body position code.
-Overall agreement averaged 90.5% across video files, ranging from 68.4%-100%.
-Cohen's *kappa* averaged 85.5% across video files, ranging from 31.0%-100%.
-<!-- UPDATE THESE BEFORE SUBMITTING -->
+Overall agreement averaged 90.9% across video files, ranging from 68.4%-100%.
+Cohen's *kappa* averaged 86.1% across video files, ranging from 31.0%-100%.
 
@@
 Overall accuracy for group models  (*M* = `r metrics_lookup["Overall Accuracy", "mean_Group"]`) was slightly lower than accuracy for individual models (*M* = `r metrics_lookup["Overall Accuracy", "mean_Individual"]`).
-Although overall accuracy from our unsupervised, in-home data collection did not match the near-perfect accuracy  (.95-.98) found in prior in-lab studies [@FranchakScott2021; @AiraksinenRasanen2020], both models approached the level of agreement found between two human coders (*M* = .905).
+Although overall accuracy from our unsupervised, in-home data collection did not match the near-perfect accuracy  (.95-.98) found in prior in-lab studies [@FranchakScott2021; @AiraksinenRasanen2020], both models approached the level of agreement found between two human coders (*M* = .906).
 Most likely, lower accuracy in the current study results from the more variable and complex behavior observed in an unsupervised setting rather than from a difference in the quality of the classification model.
@@
   scale_shape_manual(guide = "none", values = 22:23) + 
-  scale_x_continuous(breaks = c(0,.5,1), labels = c(0, 50, 100), limits = c(-.07, 1.07), name = "Actual Prevalence (%)") + 
-  scale_y_continuous(breaks = c(0,.5,1), labels = c(0, 50, 100), limits = c(-.07, 1.07), name = "Predicted Prevalence (%)") + 
+  scale_x_continuous(breaks = c(0,.5,1), labels = c(0, 50, 100), limits = c(-.07, 1.07), name = "Actual Time (%)") + 
+  scale_y_continuous(breaks = c(0,.5,1), labels = c(0, 50, 100), limits = c(-.07, 1.07), name = "Predicted Time (%)") + 
   facet_wrap("pos", ncol = 2, as.table = T) +

### 2023-03-24 manuscript.tex
Updated reliability

@@
 Interrater reliability was based on the proportion of video frames that the two coders chose the same body position code.
-Overall agreement averaged 90.5\% across video files, ranging from 68.4\%-100\%.
-Cohen's \emph{kappa} averaged 85.5\% across video files, ranging from 31.0\%-100\%.
+Overall agreement averaged 90.9\% across video files, ranging from 68.4\%-100\%.
+Cohen's \emph{kappa} averaged 86.1\% across video files, ranging from 31.0\%-100\%.
 
@@
 Overall accuracy for group models (\emph{M} = 0.85) was slightly lower than accuracy for individual models (\emph{M} = 0.92).
-Although overall accuracy from our unsupervised, in-home data collection did not match the near-perfect accuracy (.95-.98) found in prior in-lab studies (Airaksinen et al., 2020; Franchak et al., 2021), both models approached the level of agreement found between two human coders (\emph{M} = .905).
+Although overall accuracy from our unsupervised, in-home data collection did not match the near-perfect accuracy (.95-.98) found in prior in-lab studies (Airaksinen et al., 2020; Franchak et al., 2021), both models approached the level of agreement found between two human coders (\emph{M} = .906).
 Most likely, lower accuracy in the current study results from the more variable and complex behavior observed in an unsupervised setting rather than from a difference in the quality of the classification model.

### 2023-03-24 manuscript.Rmd
fixed timeline data inclusion

@@
          nap_time = ifelse(is.na(sit_time), nap_time, NA)) %>% 
-  filter(id_uni != 10401) %>% 
   pivot_longer(cols = nap_time:upright_time, names_to = "position", values_to = "prop") %>% 
@@
 Body position data was available to describe `r apa_num(min(coverage$prop_accounted)*100, digits = 1)`-`r apa_num(max(coverage$prop_accounted)*100, digits = 1)`% of the awake portion of the recording period (median = `r apa_num(median(coverage$prop_accounted)*100, digits = 1)`%).
-Younger infant \#13 had the least portion of the waking day accounted for; the family left the house to run errands during the majority of the recording period, and the infant napped during much of the remaining time at home.
+Younger infant \#14 had the least portion of the waking day accounted for; the family left the house to run errands during the majority of the recording period, and the infant napped during much of the remaining time at home.
 Overall, the recordings produced a median of `r apa_num(median(coverage$hours_recorded))` hours of motion data, with the entire dataset totaling `r apa_num(sum(coverage$hours_recorded))` hours.
@@
          group = factor(ifelse(age <= 8, "Younger", "Older")),
-         nap_time = ifelse(is.na(sit_time), nap_time, NA)) %>% 
-  filter(id_uni != 10401)
+         nap_time = ifelse(is.na(sit_time), nap_time, NA))
+
 fullday_split <- read_csv("data/imu-compiled-split.csv") %>% 
@@
          group = factor(ifelse(age <= 8, "Younger", "Older")),
-         nap_time = ifelse(is.na(sit_time), nap_time, NA)) %>% 
-  filter(id_uni != 10401)
+         nap_time = ifelse(is.na(sit_time), nap_time, NA))
+
 fullday <- list("group" = fullday_group, "split" = fullday_split)

### 2023-03-24 manuscript.tex
fixed timeline data inclusion

@@
 
-Most participants were tested in a single session (\emph{n} = 16), but 6 participants contributed between 2-4 sessions as part of an ongoing longitudinal study.
+Most participants were tested in a single session (\emph{n} = 15), but 7 participants contributed between 2-4 sessions as part of an ongoing longitudinal study.
 Only 1 session was excluded due to a technical error---one of the four IMU sensors failed to record, resulting in an unusable set of data for classification.
-Across the two age groups, we report data on a total of 33 sessions, with 13 sessions from younger infants and 20 sessions from older infants.
-Across sessions, younger infants' age ranged from 3.8 to 7.2 months (\emph{M} = 5.0) and older infants' age ranged from 10.7 to 14.2 months (\emph{M} = 11.7) .
+Across the two age groups, we report data on a total of 34 sessions, with 14 sessions from younger infants and 20 sessions from older infants.
+Across sessions, younger infants' age ranged from 3.8 to 7.2 months (\emph{M} = 5.2) and older infants' age ranged from 10.7 to 14.2 months (\emph{M} = 11.7) .
 
@@
 
-We report three sets of results based on 33 full-day testing sessions resulting in a total of 294 hours of movement recording.
+We report three sets of results based on 34 full-day testing sessions resulting in a total of 302 hours of movement recording.
 First, we focus on the accuracy of group and individual models trained and validated from data during the proximal comparison period.
@@
 
-Whereas the proximal analyses used all 33 sessions, this was not possible in the distal comparison.
+Whereas the proximal analyses used all 34 sessions, this was not possible in the distal comparison.
 Because the start of the visit was scheduled during a time the infant was awake, it was common for a nap to be needed following the proximal period.
@@
 Predictions from the group model were used to ensure that motion data were classified consistently across all sessions.
-Session start times ranged from 08:55 to 13:20 with a median of 10:20.
+Session start times ranged from 09:15 to 13:20 with a median of 10:25.
 Although the equipment was always dropped off in the morning, sometimes infants were unexpectedly asleep, leading to a few sessions in which infants began wearing the sensors later than intended (such as \#11 in the younger group).
@@
 Among participants who wore the sensors from morning to bedtime, the majority of the time during the day either resulted in usable body position data or was excluded due to caregiver-reported naps (gray periods in Figure \ref{fig:timelines}).
-The total length of the recording period ranged from 6.92-11.75 hours with a median of 9.08 hours.
-Nap times reported during the recording period ranged from 0.00-4.99 hours with a median of 2.22 hours.
-Body position data was available to describe 18.3-100.0\% of the awake portion of the recording period (median = 100.0\%).
-Younger infant \#13 had the least portion of the waking day accounted for; the family left the house to run errands during the majority of the recording period, and the infant napped during much of the remaining time at home.
-Overall, the recordings produced a median of 6.15 hours of motion data, with the entire dataset totaling 201.06 hours.
+The total length of the recording period ranged from 6.92-11.75 hours with a median of 9.04 hours.
+Nap times reported during the recording period ranged from 0.00-4.99 hours with a median of 2.25 hours.
+Body position data was available to describe 37.3-100.0\% of the awake portion of the recording period (median = 100.0\%).
+Younger infant \#14 had the least portion of the waking day accounted for; the family left the house to run errands during the majority of the recording period, and the infant napped during much of the remaining time at home.
+Overall, the recordings produced a median of 6.14 hours of motion data, with the entire dataset totaling 206.66 hours.
 Even the shortest full recording (3.06 hours) exceeded the longest observational sessions in past work in which an experimenter operated a camera.
 
-Considering that body position annotation takes 2-5 hours of labor per 1 hour of behavior, the 201 hours of data we recorded would have taken 402-1005 hours of labor to annotate from video.
-However, only an estimated 62-155 hours of labor was needed to annotate the initial hour of each session's video that was used to train the machine learning models---an immense savings in the human cost of annotation that allows the method to scale to larger sample sizes and recording durations.
+Considering that body position annotation takes 2-5 hours of labor per 1 hour of behavior, the 207 hours of data we recorded would have taken 413-1033 hours of labor to annotate from video.
+However, only an estimated 64-160 hours of labor was needed to annotate the initial hour of each session's video that was used to train the machine learning models---an immense savings in the human cost of annotation that allows the method to scale to larger sample sizes and recording durations.
 
@@
 \midrule
-Supine & 38.5\% (24.0) & 14.0\% (8.4) & 40.2\% (31.6) & 10.0\% (9.5)\\
-Prone & 14.5\% (13.8) & 14.4\% (6.0) & 12.5\% (10.2) & 16.9\% (10.6)\\
-Sitting & 25.4\% (12.4) & 44.4\% (10.1) & 18.8\% (16.3) & 46.9\% (13.3)\\
-Upright & 7.7\% (9.3) & 18.6\% (7.4) & 10.4\% (13.6) & 18.7\% (8.4)\\
-Held & 13.1\% (7.4) & 8.5\% (5.4) & 17.1\% (20.8) & 7.4\% (7.6)\\
+Supine & 38.6\% (24.1) & 14.0\% (8.1) & 43.1\% (32.0) & 10.5\% (9.6)\\
+Prone & 14.0\% (14.0) & 15.1\% (6.3) & 11.1\% (11.4) & 16.6\% (10.2)\\
+Sitting & 28.9\% (15.1) & 44.4\% (9.4) & 20.3\% (14.8) & 46.6\% (12.4)\\
+Upright & 5.4\% (7.1) & 18.2\% (7.4) & 9.5\% (13.3) & 18.4\% (7.8)\\
+Held & 13.0\% (7.7) & 8.3\% (5.2) & 16.0\% (20.1) & 8.0\% (7.6)\\
 \bottomrule
@@
 In Franchak (2019), upright time was 0.6-5.5\% of the time for infants 3-6 months, increasing to 22.0\% at 12 months.
-Results from the current study were similar: younger infants' sitting time estimated from full-day group models was 7.72\% compared with 18.57\% for the older group.
+Results from the current study were similar: younger infants' sitting time estimated from full-day group models was 5.41\% compared with 18.24\% for the older group.
 In Franchak (2019), prone time was 2.9-9.2\% of the time for infants 3-6 months, and was 7.2\% at 12 months.
-In the current study, prone time was 14.46\% for younger infants and 14.45\% for older infants.
+In the current study, prone time was 14.03\% for younger infants and 15.09\% for older infants.
 Held, sitting, and supine are more difficult to compare because prior work (Franchak, 2019) considered held as any time that infants were in caregivers' arms or on their laps, however, we counted laying or sitting on caregivers' laps as supine and sitting, respectively.
@@
 Although the numbers are less directly comparable, our results are similar in showing that held and supine time was less for older infants than younger infants (Table \ref{tab:agetable}).
-Sitting was greater for older infants (44.42\%) compared with younger infants (25.37\%).
+Sitting was greater for older infants (44.39\%) compared with younger infants (28.91\%).
 Moreover, sitting was the most frequent body position followed by upright in 12-month-olds (Franchak, 2019), just as we observed in the current study.
@@
 Figure \ref{fig:age} shows that both group and individual model predictions captured age-related differences in body position.
-Age in months and upright time were significantly, positively correlated using group (\emph{r} = 0.59, \emph{p} \textless{} .001) and individual models (\emph{r} = 0.50, \emph{p} = .005).
-Similarly, age predicted increases in sitting time using group (\emph{r} = 0.68, \emph{p} \textless{} .001) and individual models (\emph{r} = 0.72, \emph{p} \textless{} .001).
-In contrast, age predicted decreases in supine time using group (\emph{r} = -0.66, \emph{p} \textless{} .001) and individual models (\emph{r} = -0.61, \emph{p} \textless{} .001).
+Age in months and upright time were significantly, positively correlated using group (\emph{r} = 0.65, \emph{p} \textless{} .001) and individual models (\emph{r} = 0.52, \emph{p} = .002).
+Similarly, age predicted increases in sitting time using group (\emph{r} = 0.57, \emph{p} \textless{} .001) and individual models (\emph{r} = 0.72, \emph{p} \textless{} .001).
+In contrast, age predicted decreases in supine time using group (\emph{r} = -0.61, \emph{p} \textless{} .001) and individual models (\emph{r} = -0.61, \emph{p} \textless{} .001).
 No clear age trends were found for prone time.
-Age was not significantly related to held time for group models (\emph{r} = -0.23, \emph{p} = .200), but individual models showed a significant negative correlation (\emph{r} = -0.42, \emph{p} = .021) resulting from an outlier in the younger group.
+Age was not significantly related to held time for group models (\emph{r} = -0.25, \emph{p} = .148), but individual models showed a significant negative correlation (\emph{r} = -0.37, \emph{p} = .039) resulting from an outlier in the younger group.
 
@@
 Most important, accuracy was consistently strong later in the recording when behavior was completely unsupervised: Human-coded and model-predicted durations of body positions during the distal comparison period were highly correlated, even when narrowing to the scale of 10 minutes.
-Examining full recordings showed that the new method allows us to capture more data (median = 6.15 hours of awake behavior) than is typical with video methods (1-2 hours) while simultaneously reducing the human labor needed to annotate it.
+Examining full recordings showed that the new method allows us to capture more data (median = 6.14 hours of awake behavior) than is typical with video methods (1-2 hours) while simultaneously reducing the human labor needed to annotate it.
 Ultimately, age differences in body position mirrored past findings that employed other methods, suggesting that the outcome of full-day body position recordings is suitable for describing developmental changes in motor behavior.
@@
 In most cases, model prediction accuracy approached human reliability, suggesting that model predictions can be confidently used in analyses of full-day infant behavior.
-Examining the resulting corpus of 201.06 hours of real-time infant body position showed the feasibility of capturing data covering the majority of infants' awake time and the variety of activities contained therein.
+Examining the resulting corpus of 206.66 hours of real-time infant body position showed the feasibility of capturing data covering the majority of infants' awake time and the variety of activities contained therein.
 A rudimentary analysis of aggregated data showed that full-day position estimates conformed to expectations about age differences in motor behavior.

### 2023-03-27 manuscript.Rmd
Lots of small edits

@@
 Infants struggle to manipulate objects while supine and prone, but sitting affords object exploration [@SoskaAdolph2014].
-Upright walking changes how infants and caregivers interact; while walking infants move farther away from caregivers and share toys in different ways compared to when crawling in a prone position [@ChenSchneider2022; @Karasik2011].
+Upright walking changes how infants and caregivers interact; walking infants move farther away from caregivers and share toys in different ways compared to infants crawling in a prone position [@ChenSchneider2022; @Karasik2011].
 As infants grow older and acquire new abilities, such as independent sitting and walking, they spend more time sitting and upright and less time held, supine, and prone [@Survey; @Thurman2017; @Freeplay; @AdolphCTL2014].
-Caregivers provide different learning input to infants depending on infants' body position [@KretchKoziol2022; @WestFletcher2022; @WestIverson2021; @Karasik2014].
+Caregivers provide different learning input to infants depending on infants' body position [@KretchKoziol2022; @WestIverson2021; @Karasik2014].
+For example, infants' walking movements spur caregivers to talk about actions rather than just describe objects.
 Thus, characterizing individual differences in the day-to-day accumulation of body position experiences informs developmental theory by revealing differential opportunities for learning [@CurrentOpinion].
 
-In this paper, we present an inertial sensing method to classify infant body position from moment-to-moment across an entire day, and validate its accuracy using over 100 hours of video recorded across 34 in-home data collection sessions.
+In this paper, we present an inertial sensing method to classify infant body position from moment to moment across an entire day, and validate its accuracy using over 100 hours of video recorded across 34 in-home data collection sessions.
 Our method takes inspiration from a more mature technology: Long-form audio recordings of infants' language experiences using wearable audio recorders.
@@
 For example, infants' daily experiences hearing music are clustered in time, with "bursty" episodes of music presence separated by relatively long periods during which music is absent [@MendozaFausey2022].
-Synchronizing audio recordings with other data sources extends researchers' ability to characterize daily experiences.
+In addition, synchronizing audio recordings with other data sources extends researchers' ability to characterize daily experiences.
 Linking LENA&reg; speech measurements to repeated, time-stamped text-message surveys about infant device placement revealed that infants heard less caregiver speech during periods that they were restrained in devices such as swings, exersaucers, and car seats [@MalachowskiSalo2023].
@@
 Third and fourth, recordings capture *real-time data* over a *full day*.
-The ability to record real-time data is vital for making inferences about processes that happen on the timescale of minutes or even seconds within in an individual as opposed to comparisons of aggregated data between infants.
+The ability to record real-time data is vital for making inferences about processes that happen on the timescale of minutes or even seconds within an individual as opposed to comparisons of aggregated data between infants.
 Synchronizing real-time data to other data streams helps to reveal sources of variability within an individual [e.g., @WassPhillips2022].
@@
 For example, correlations between human transcribed counts of adult words and child vocalizations against LENA&reg;'s automatic counts were strong, *r* = .698 and *r* = .649, respectively [@CristiaLavechin2020].
-However, poor agreement was found for other metrics, such as the number of "conversational turns" between the child and communicative partners and the identification of male speakers.
+For other metrics, such as the number of "conversational turns" between the child and communicative partners and the identification of male speakers, agreement was poor.
 <!-- EVENTUALLY CITE NEW EMA HERE -->
@@
 Even if full day videos were available, the lack of suitable automatic classification tools means that the human cost of annotation makes it difficult for video methods to scale to testing large numbers of participants.
-Our annotation of body position takes approximately 2-5 hours to complete for every hour of video (depending on how often infants switch positions), meaning that a full "waking day" of approximately 11 hours for a 12-month-old [@GallandTaylor2012] could take 22-55 hours of labor to full annotate.
+Our annotation of body position takes approximately 2-5 hours to complete for every hour of video (depending on how often infants switch positions), meaning that a full "waking day" of approximately 11 hours for a 12-month-old [@GallandTaylor2012] could take 22-55 hours of labor to fully annotate.
 
@@
 Ecological momentary assessment uses text-message or app-based notifications to prompt caregivers to make repeated estimates of infants' behavior over the course of a day [@Survey; @KadookaCaufield2021].
-And although the responses are valuable in aggregate, survey methods lack the real-time temporal resolution to describe moment-to-moment changes in behavior.
+Although the EMA responses are valuable in aggregate, survey methods lack the real-time temporal resolution to describe moment-to-moment changes in behavior.
 For diaries and interviews, limits on caregivers' memory mean that they will report what was most frequent, but cannot remember events that happen on the scale of seconds and minutes.
@@
 Unlike motion tracking systems that might be used in a lab, IMUs do not provide data about the position of the body in space.
-Thus, data processing algorithms are needed to classify the raw sensor data (i.e., linear and angular acceleration timeseries) into meaningful categories (e.g., supine, prone, sitting, upright).
+Thus, data processing algorithms are needed to classify the raw sensor data (i.e., linear and angular acceleration time series) into meaningful categories (e.g., supine, prone, sitting, upright).
 The difficulty of the classification task depends on the categories of interest.
@@
 An upright infant can be standing still or can be walking briskly across the room.
-A prone infant and be stationary in "tummy time", or they can crawl in myriad ways [@AdolphVereijken1998].
+A prone infant can be stationary in "tummy time", or they can crawl in a myriad of ways [@AdolphVereijken1998].
 Moreover, the configuration of the arms, legs, and torso within a body position can vary greatly in everyday contexts.
-Infants can sit on the floor in a tripod position with support from an arm, in a "V" position with legs fully extended, in a "W" position with knees bent.
-Sitting on a caregivers' lap without the need to maintain balance means that the legs can dangle and the torso can lean in different directions.
+Infants can sit on the floor in a tripod position with support from an arm, in a "V" position with legs fully extended, or in a "W" position with knees bent.
+Sitting on a caregiver's lap without the need to maintain balance means that the legs can dangle and the torso can lean in different directions.
 Sitting in a high chair or car seat reduces the magnitude of postural sway within sitting, and creates even more possibilities for how the arms and legs may move relative to the torso.
@@
 
-Three prior investigations have used machine learning to categorize infant body position from IMUs towards the goal of collecting full-day data.
+Three prior investigations have used different machine learning techniques to categorize infant body position from IMUs towards the goal of collecting full-day data.
 @AiraksinenRasanen2020 tested 4- to 8-month-olds in a laboratory visit, and found 95% accuracy in distinguishing between body position categories that crawling infants could perform on the floor (excluding times that infants were held by caregivers).
@@
 Supervised recordings from past work [@FranchakScott2021; @AiraksinenRasanen2020; @AiraksinenGallen2022], whether in the home or in the lab, let researchers set up the situation to encourage or restrict certain behaviors.
-Usually, caregivers were asked to play with the infant.
+Prior work focused on "free play", in which caregivers were asked to play with the infant without restraining the infant or otherwise shaping how they could move.
 However, in a real day, non-play activities (e.g., eating lunch in a high chair) create challenging situations for applying automated classification of body position.
+Indeed different types of sitting---independently on the floor, supported on a caregiver's lap, or restrained in a high chair---all need to be counted as equal.
+Moreover, models must be able to new variants of behaviors that occur over the course of natural behavior that did not appear in the training data.
 Will models trained on video-recorded observations at the beginning of the day generalize to predict behavior at a later time?
@@
 A custom garment was designed to hold the IMUs: Internal pockets were sewn into a snug-fitting pair of infant leggings so that IMUs would stay close to the body (reducing vibration), which also prevented infants from pulling out the sensors.
-A pocket over the hip and a pocket just above the ankle were sewn on the outside of the right and left legs of the garment.
+A pocket over the hip and a pocket just above the ankle were sewn on the lateral surface of the right and left legs of the garment.
 Each sensor recorded accelerometer and gyroscope data at 62.5 Hz throughout the day, with sufficient battery and onboard storage to record for approximately 12 hours.
@@
 Figure \@ref(fig:exemplar-timeline)) shows an exemplar timeline of the entire procedure and recording periods for a single participant.
-On the day of the visit, a researcher arrived at the participant’s home in the morning between and prepared all the equipment at the doorstep.
+On the day of the visit, a researcher arrived at the participant’s home in the morning and prepared all the equipment at the doorstep.
 In order to synchronize all three recording devices, the researcher began by first turning on the video camera and the LENA audio device (the IMUs were already configured and placed before arrival because they required a proprietary sensor dock to begin recording in the laboratory).
@@
 
-Afterwards, they were asked to go about their day as usual with the infant wearing the sensor garment until their bedtime, only taking off the sensor for naps, baths, and diaper changes.
-The caregiver logged the times the sensors were removed (blank areas in the timeline in Figure \@ref(fig:exemplar-timeline))) or the child took a nap (gray areas in the timeline in Figure \@ref(fig:exemplar-timeline))) so that those times could be excluded from analysis.
+Afterwards, they were asked to go about their day as usual with the infant wearing the sensor garment until their bedtime, only taking off the sensor for naps, baths, diaper changes, and trips out of the house.
+The caregiver logged the times the sensors were removed (blank areas in the timeline in Figure \@ref(fig:exemplar-timeline)) or the child took a nap (gray areas in the timeline in Figure \@ref(fig:exemplar-timeline)) so that those times could be excluded from analysis.
 The following day a researcher picked up the equipment, verified that the paperwork was signed, and compensated the participant.
@@
 The same machine learning classification process was used as in prior work [@FranchakScott2021].
-Using the synchronization point, human-coded body annotations from video were linked to the corresponding times in the IMU timeseries data.
+Using the synchronization point, human-coded body annotations from video were linked to the corresponding times in the IMU time series data.
 A single, merged dataset was created with synchronized accelerometer signals (in three orientations: X, Y, and Z) and gyroscope signals (in three orientations: roll, pitch, and yaw) for each of the four sensors (left hip, right hip, left ankle, right ankle) with the corresponding timestamp and body position code using the *timetk* package [@DanchoVaughan2023] and the *lubridate* package [@GrolemundWickham2011] in R version 4.1.2 [@R42].
@@
 <!-- WHAT CAN WE SHARE AND HOW -->
+https://nyu.databrary.org/volume/1580 - all raw data files and annotations for one participant
+
+This reproducible manuscript showing how we accumulated
 

### 2023-03-27 manuscript.tex
Lots of small edits

@@
 Infants struggle to manipulate objects while supine and prone, but sitting affords object exploration (Soska \& Adolph, 2014).
-Upright walking changes how infants and caregivers interact; while walking infants move farther away from caregivers and share toys in different ways compared to when crawling in a prone position (Chen et al., 2022; Karasik et al., 2011).
+Upright walking changes how infants and caregivers interact; walking infants move farther away from caregivers and share toys in different ways compared to infants crawling in a prone position (Chen et al., 2022; Karasik et al., 2011).
 As infants grow older and acquire new abilities, such as independent sitting and walking, they spend more time sitting and upright and less time held, supine, and prone (Adolph \& Tamis-LeMonda, 2014; Franchak et al., 2018; Franchak, 2019; Thurman \& Corbetta, 2017).
-Caregivers provide different learning input to infants depending on infants' body position (Karasik et al., 2014; Kretch et al., 2022; West et al., 2022; West \& Iverson, 2021).
+Caregivers provide different learning input to infants depending on infants' body position (Karasik et al., 2014; Kretch et al., 2022; West \& Iverson, 2021).
+For example, infants' walking movements spur caregivers to talk about actions rather than just describe objects.
 Thus, characterizing individual differences in the day-to-day accumulation of body position experiences informs developmental theory by revealing differential opportunities for learning (Franchak, 2020).
 
-In this paper, we present an inertial sensing method to classify infant body position from moment-to-moment across an entire day, and validate its accuracy using over 100 hours of video recorded across 34 in-home data collection sessions.
+In this paper, we present an inertial sensing method to classify infant body position from moment to moment across an entire day, and validate its accuracy using over 100 hours of video recorded across 34 in-home data collection sessions.
 Our method takes inspiration from a more mature technology: Long-form audio recordings of infants' language experiences using wearable audio recorders.
@@
 For example, infants' daily experiences hearing music are clustered in time, with ``bursty'' episodes of music presence separated by relatively long periods during which music is absent (Mendoza \& Fausey, 2022).
-Synchronizing audio recordings with other data sources extends researchers' ability to characterize daily experiences.
+In addition, synchronizing audio recordings with other data sources extends researchers' ability to characterize daily experiences.
 Linking LENA® speech measurements to repeated, time-stamped text-message surveys about infant device placement revealed that infants heard less caregiver speech during periods that they were restrained in devices such as swings, exersaucers, and car seats (Malachowski et al., 2023).
@@
 Third and fourth, recordings capture \emph{real-time data} over a \emph{full day}.
-The ability to record real-time data is vital for making inferences about processes that happen on the timescale of minutes or even seconds within in an individual as opposed to comparisons of aggregated data between infants.
+The ability to record real-time data is vital for making inferences about processes that happen on the timescale of minutes or even seconds within an individual as opposed to comparisons of aggregated data between infants.
 Synchronizing real-time data to other data streams helps to reveal sources of variability within an individual (e.g., Wass et al., 2022).
@@
 For example, correlations between human transcribed counts of adult words and child vocalizations against LENA®'s automatic counts were strong, \emph{r} = .698 and \emph{r} = .649, respectively (Cristia et al., 2020).
-However, poor agreement was found for other metrics, such as the number of ``conversational turns'' between the child and communicative partners and the identification of male speakers.
+For other metrics, such as the number of ``conversational turns'' between the child and communicative partners and the identification of male speakers, agreement was poor.
 
@@
 Even if full day videos were available, the lack of suitable automatic classification tools means that the human cost of annotation makes it difficult for video methods to scale to testing large numbers of participants.
-Our annotation of body position takes approximately 2-5 hours to complete for every hour of video (depending on how often infants switch positions), meaning that a full ``waking day'' of approximately 11 hours for a 12-month-old (Galland et al., 2012) could take 22-55 hours of labor to full annotate.
+Our annotation of body position takes approximately 2-5 hours to complete for every hour of video (depending on how often infants switch positions), meaning that a full ``waking day'' of approximately 11 hours for a 12-month-old (Galland et al., 2012) could take 22-55 hours of labor to fully annotate.
 
@@
 Ecological momentary assessment uses text-message or app-based notifications to prompt caregivers to make repeated estimates of infants' behavior over the course of a day (Franchak, 2019; Kadooka et al., 2021, April).
-And although the responses are valuable in aggregate, survey methods lack the real-time temporal resolution to describe moment-to-moment changes in behavior.
+Although the EMA responses are valuable in aggregate, survey methods lack the real-time temporal resolution to describe moment-to-moment changes in behavior.
 For diaries and interviews, limits on caregivers' memory mean that they will report what was most frequent, but cannot remember events that happen on the scale of seconds and minutes.
@@
 An upright infant can be standing still or can be walking briskly across the room.
-A prone infant and be stationary in ``tummy time'', or they can crawl in myriad ways (Adolph et al., 1998).
+A prone infant can be stationary in ``tummy time'', or they can crawl in a myriad of ways (Adolph et al., 1998).
 Moreover, the configuration of the arms, legs, and torso within a body position can vary greatly in everyday contexts.
-Infants can sit on the floor in a tripod position with support from an arm, in a ``V'' position with legs fully extended, in a ``W'' position with knees bent.
-Sitting on a caregivers' lap without the need to maintain balance means that the legs can dangle and the torso can lean in different directions.
+Infants can sit on the floor in a tripod position with support from an arm, in a ``V'' position with legs fully extended, or in a ``W'' position with knees bent.
+Sitting on a caregiver's lap without the need to maintain balance means that the legs can dangle and the torso can lean in different directions.
 Sitting in a high chair or car seat reduces the magnitude of postural sway within sitting, and creates even more possibilities for how the arms and legs may move relative to the torso.
@@
 
-Three prior investigations have used machine learning to categorize infant body position from IMUs towards the goal of collecting full-day data.
+Three prior investigations have used different machine learning techniques to categorize infant body position from IMUs towards the goal of collecting full-day data.
 Airaksinen et al.
@@
 Supervised recordings from past work (Airaksinen et al., 2022, 2020; Franchak et al., 2021), whether in the home or in the lab, let researchers set up the situation to encourage or restrict certain behaviors.
-Usually, caregivers were asked to play with the infant.
+Prior work focused on ``free play'', in which caregivers were asked to play with the infant without restraining the infant or otherwise shaping how they could move.
 However, in a real day, non-play activities (e.g., eating lunch in a high chair) create challenging situations for applying automated classification of body position.
@@
 The same machine learning classification process was used as in prior work (Franchak et al., 2021).
-Using the synchronization point, human-coded body annotations from video were linked to the corresponding times in the IMU timeseries data.
+Using the synchronization point, human-coded body annotations from video were linked to the corresponding times in the IMU time series data.
 A single, merged dataset was created with synchronized accelerometer signals (in three orientations: X, Y, and Z) and gyroscope signals (in three orientations: roll, pitch, and yaw) for each of the four sensors (left hip, right hip, left ankle, right ankle) with the corresponding timestamp and body position code using the \emph{timetk} package (Dancho \& Vaughan, 2023) and the \emph{lubridate} package (Grolemund \& Wickham, 2011) in R version 4.1.2 (R Core Team, 2021).
@@
 Validation analyses below will compare the human-coded annotations and random forest model predictions for testing data that are independent from training data, providing a test of whether models can successfully generalize.
+
+\url{https://nyu.databrary.org/volume/1580} - all raw data files and annotations for one participant
+
+This reproducible manuscript showing how we accumulated
 
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-WestFletcher2022}{}}%
-West, K.
-L., Fletcher, K.
-K., Adolph, K.
-E., \& Tamis-LeMonda, C.
-S.
-(2022).
-Mothers talk about infants' actions: How verbs correspond to infants' real-time behavior.
-\emph{Developmental Psychology}, \emph{58}(3), 405--416.
-\url{https://doi.org/10.1037/dev0001285}
-
 \leavevmode\vadjust pre{\hypertarget{ref-WestIverson2021}{}}%

### 2023-03-27 manuscript.Rmd
Finished minor lab edits

@@
 We scored sitting to include any form of the following seated positions: 1) infants sat with their buttocks on a surface, such as on the floor or a caregiver’s lap, 2) infant was in a kneeling-sit position, in which their knees were on the ground with their legs tucked underneath the buttocks, and 3) infant was in a seating device, such as a high chair, that kept the torso oriented perpendicular to the ground (a reclined position, such as in a young infant's car seat, would be counted as supine).
-Upright was coded when the infant was standing on the ground with two feet or walking (regardless of whether infants' balance was assisted by a caregiver or with their hands holding onto something for support).
+Upright was coded when the infant was standing or squatting on the ground with two feet or walking (regardless of whether infants' balance was assisted by a caregiver or with their hands holding onto something for support).
 When an infant was carried by a caregiver, held was coded.
@@
 
-Datasets containing one row per 4-s window were used for machine learning classification and validation.
+The resulting windowed datasets were used for machine learning classification and validation.
 For each analysis reported in the results, a subset of data were defined as a "training" set and another, independent portion of the data were defined as a "testing" set.
@@
 The *tidymodels* set of packages in R [@tidymodels] was used to tune the "mtry" and "ntrees" hyperparameters, which control the number of columns sampled and the total number of trees created when modeling.
-We use an mtry parameter of 44 and and ntree parameter of 550, however, changing hyperparameters had little impact on model performance.
+We use an mtry parameter of 44 and and ntree parameter of 550; however, changing hyperparameters had little impact on model performance.
 Just like a linear regression model, the resulting random forest model could later be applied to a set of testing data with the *predict* function.
@@
 Each blue circle represents that accuracy for each recording session.
-Horizontal black bars indicates the mean across sessions.", fig.align="center", fig.width=6, fig.height=3, cache=TRUE, cache.extra=tools::md5sum("data/group_split_metrics.csv")}
+Horizontal black bars indicate the mean across sessions.", fig.align="center", fig.width=6, fig.height=3, cache=TRUE, cache.extra=tools::md5sum("data/group_split_metrics.csv")}
 metrics <- read_csv("data/group_split_metrics.csv") %>% 
@@
 This leave-one-out cross-validation tested the generalization of the model to a recording session that was not used at all in the training set.
-In contrast, individual models were trained using the first 75% of a recording session, then tested against the last 25% of the same session.
+In contrast, individual models were trained using the first 75% of a recording session, then tested against the (temporally) last 25% of the same session.
 The individual model tests whether earlier training data generalize to later testing data within an individual participant's recording.
@@
 
-<!-- GRAPHIC FOR TWO KINDS OF MODELS? -->
-
 ### Overall Accuracy
@@
 Overall accuracy for group models  (*M* = `r metrics_lookup["Overall Accuracy", "mean_Group"]`) was slightly lower than accuracy for individual models (*M* = `r metrics_lookup["Overall Accuracy", "mean_Individual"]`).
-Although overall accuracy from our unsupervised, in-home data collection did not match the near-perfect accuracy  (.95-.98) found in prior in-lab studies [@FranchakScott2021; @AiraksinenRasanen2020], both models approached the level of agreement found between two human coders (*M* = .906).
+Although overall accuracy from our semi-supervised, in-home data collection did not match the near-perfect accuracy  (.95-.98) found in prior in-lab studies [@FranchakScott2021; @AiraksinenRasanen2020], both models approached the level of agreement found between two human coders (*M* = .906).
 Most likely, lower accuracy in the current study results from the more variable and complex behavior observed in an unsupervised setting rather than from a difference in the quality of the classification model.
 Visual inspection of Figure \@ref(fig:metrics)A shows that accuracy values were heavily skewed, with most approaching perfect accuracy but a few participants with very poor accuracy.
-Looking at the median performance suggests that the difference between models was not as wide for the typical participant (group median accuracy = `r metrics_lookup["Overall Accuracy", "median_Group"]`; individual median accuracy = `r metrics_lookup["Overall Accuracy", "median_Individual"]`.
+Looking at the median performance suggests that the difference between models was not as wide for the typical participant (group median accuracy = `r metrics_lookup["Overall Accuracy", "median_Group"]`; individual median accuracy = `r metrics_lookup["Overall Accuracy", "median_Individual"]`).
 Indeed, it is notable that the worst-case accuracy from group models (minimum = `r metrics_lookup["Overall Accuracy", "min_Group"]`) was considerably lower than in individual models (minimum = `r metrics_lookup["Overall Accuracy", "min_Individual"]`).
@@
 Strong overall accuracy can be misleading when class prevalence is unbalanced.
-For example, most infants spent longer sitting than prone, so good performance for sitting classification could mask poorer performance in prone classification.
+For example, most infants spent more time sitting than prone, so good performance for sitting classification could mask poorer performance in prone classification.
 Accordingly, we report Cohen's kappa, a commonly-used metric that penalizes missing rare events (Figure \@ref(fig:metrics)B), and provide classification metrics for each individual body position (Table \@ref(tab:metrics-by-class)) to account for imbalance in body position rates within and between individuals.
@@
 High sensitivity means that events are unlikely to be missed.
-In contrast, positive predictive value (PPV) refers to the proportion of events classified for a given position that actually belonged to that position (e.g., if the model said a baby was upright during 100 windows, how many of those windows were indeed human-coded sitting events?).
+In contrast, positive predictive value (PPV) refers to the proportion of events classified for a given position that actually belonged to that position (e.g., if the model said a baby was upright during 100 windows, how many of those windows were indeed human-coded upright events?).
 High PPV means that we can be confident in the event label.
@@
 For group models, sensitivity and PPV were similar: They were highest for supine, prone, and sitting (the most accurately identified class) and lowest for upright and held.
-All values were sufficient (> .8).
 Results for individual models were similar to group models, with the exception of a somewhat lower sensitivity score for held.
@@
 In the first set of results, we showed that group and individual models trained from data during the proximal comparison period were accurate at classifying body position in the same recording.
-In the next analysis, we examine long-term performance by testing how accurately models trained from the proximal period could predict body position during the distal period (Figure \@ref(fig:exemplar-timeline)).
+In the next analysis, we examine long-term performance by testing how accurately models trained from the proximal period could predict body position during the distal comparison period.
 A single group model was created using all sessions' proximal period training data (rather than group models leaving out a single session); the same individual models were used.
@@
 Whereas the proximal analyses used all `r length(unique(ppts$id_uni))` sessions, this was not possible in the distal comparison.
-Because the start of the visit was scheduled during a time the infant was awake, it was common for a nap to be needed following the proximal period.
+Because the start of the visit was scheduled during a time when the infant was awake, it was common for a nap to be needed following the proximal period.
 Nine sessions were excluded because the infant was either napping or otherwise not on camera during the entire 90-minute distal recording.
@@
 The infant also spent a long time in the mother's arms in an ambiguous supine/sitting position that was hard to classify.
-This participant's proximal accuracy was also poor because similar ambiguous appeared during the initial recording.
+This participant's proximal accuracy was also poor because similar ambiguities appeared during the initial recording.
 In contrast, the "gray diamond outlier" had strong proximal accuracy, with confusion only arising in the distal period between upright and held categories.
@@
 ```{r part2bins, warning=FALSE, fig.cap="Prediction difference (difference in minutes between human-coded and model-predicted body position) for 10-minute bins in the distal comparison period.
-Each point shows the mean and SE for a single participant's recording session for each body position, summarizing the prediction difference for each of their 10-minute bins.
+Each point shows the mean and SE for a single recording session for each body position, summarizing the prediction difference for each of their 10-minute bins.
 Points falling within the gray shaded region indicate that average prediction errors were less than 1 minute.
@@
 We repeated the agreement analysis (correlating human-coded body position prevalence with model-predicted prevalence) after dividing the distal comparison period into nine 10-minute bins (marked by vertical dashed lines in Figure \@ref(fig:exemplar-timeline)).
-Infants had varying numbers of 10-minute bins depending on much time they were awake and on camera.
+Infants had varying numbers of 10-minute bins depending on how much time they were awake and on camera.
 Bins were included only if there was > 7 minutes of usable data.
@@
 ```{r timelines, warning=FALSE, fig.cap="Full-day timelines for each individual recording session, split by (A) younger infants and (B) older infants.
-Each participant's timeline shows a stacked bar graph with the proportion of time spent in each of the body position for every 5-minute period throughout the day, based on group model predictions of body position.
+Each participant's timeline shows a stacked bar graph with the proportion of time spent in each of the body positions for every 5-minute period throughout the day, based on group model predictions of body position.
 The x-axis shows time of day.
@@
 Predictions from the group model were used to ensure that motion data were classified consistently across all sessions.
+Moreover, for infants who did not display all 5 body positions in the training period, group models were necessary to be able to predict those behaviors across the entire day.
 Session start times ranged from `r start_times[1]` to `r start_times[length(start_times)]` with a median of `r start_times[floor(length(start_times)/2)]`.
 Although the equipment was always dropped off in the morning, sometimes infants were unexpectedly asleep, leading to a few sessions in which infants began wearing the sensors later than intended (such as \#11 in the younger group).
-With two exceptions, recordings lasted until the infants' bed time.
+With two exceptions, recordings lasted until the infants' bedtime.
 Older participant \#1 had the equipment picked up on the same day rather than the next day, so the recording ended at 17:00.
@@
 Considering that body position annotation takes 2-5 hours of labor per 1 hour of behavior, the `r round(sum(coverage$hours_recorded))` hours of data we recorded would have taken `r round(sum(coverage$hours_recorded)*2)`-`r round(sum(coverage$hours_recorded)*5)` hours of labor to annotate from video.
-However, only an estimated `r round(nrow(coverage)*2)`-`r round(nrow(coverage)*5)` hours of labor was needed to annotate the initial hour of each session's video that was used to train the machine learning models---an immense savings in the human cost of annotation that allows the method to scale to larger sample sizes and recording durations.
-<!-- CITES -->
+However, only an estimated `r round(nrow(coverage)*2)`-`r round(nrow(coverage)*5)` hours of labor was needed to annotate the initial hour of each session's video that was used to train the machine learning models---an immense savings in the human labor cost of annotation that allows the method to scale to larger sample sizes and recording durations.
 
@@
 
-In the last analysis, we demonstrate how the full-day recordings can reveal individual differences in body position according to age.
+In the last analysis, we demonstrate how the full-day recordings can reveal individual differences in body position according to age, replicating age differences revealed in past work [@Survey].
+Table \@ref(tab:agetable) summarizes the percent of time that younger and older infants spent in each body position out of their awake samples, as predicted by group and individual models.
 Past work using sparse EMA surveys across the day [@Survey] found that from 3-12 months, infants spent less time held, reclined, and supine but spent increasingly more time sitting and upright.
 Prone time changed little.
-Table \@ref(tab:agetable) summarizes the percent of time that younger and older infants spent in each body position out of their awake samples, as predicted by group and individual models.
-
 The most straightforward comparisons between the two investigations are upright and prone, because they were defined identically.
 In @Survey, upright time was 0.6-5.5% of the time for infants 3-6 months, increasing to 22.0% at 12 months.
-Results from the current study were similar: younger infants' sitting time estimated from full-day group models was `r pos_sum_lookup["Upright", "Younger_Group"]`% compared with `r pos_sum_lookup["Upright", "Older_Group"]`% for the older group.
+Results from the current study were similar: younger infants' upright time estimated from full-day group models was `r pos_sum_lookup["Upright", "Younger_Group"]`% compared with `r pos_sum_lookup["Upright", "Older_Group"]`% for the older group.
 In @Survey, prone time was 2.9-9.2% of the time for infants 3-6 months, and was 7.2% at 12 months.
@@
 Examining full recordings showed that the new method allows us to capture more data (median = `r apa_num(median(coverage$hours_recorded))` hours of awake behavior) than is typical with video methods (1-2 hours) while simultaneously reducing the human labor needed to annotate it.
-Ultimately, age differences in body position mirrored past findings that employed other methods, suggesting that the outcome of full-day body position recordings is suitable for describing developmental changes in motor behavior.
+Ultimately, age differences in body position mirrored past findings that employed other methods, suggesting that the outcome of full-day body position recordings is suitable for capturing individual differences and describing developmental changes in motor behavior.
 
@@
 The challenge only increased in the distal comparison period.
-Since sessions were scheduled at convenient times to for the infant and caregiver, the transition from the proximal to distal period increased the odds that infants might need a nap, a snack/meal, or engage in a less typical activity (e.g., watching TV).
+Since sessions were scheduled at convenient times for the infant and caregiver, the transition from the proximal to distal period increased the odds that infants might need a nap, a snack/meal, or engage in a less typical activity (e.g., watching TV).
 Indeed, the distal comparison in Figure \@ref(fig:exemplar-timeline) contains a long period (almost 50 minutes) where the infant sat in a high chair eating lunch.
@@
 Acceptable inter-rater reliability in developmental psychology studies of motor behavior is greater than 80%; both group and individual model results exceed that standard for most recorded sessions.
-Another comparison we can make is to the accuracy of the LENA&reg;'s automated metrics.
+Another comparison we can make is to the accuracy of the LENA&reg; device's automated metrics.
 Correlations between human and model frequencies were *r* = .698 for adult word count and *r* = .649 for child word count [@CristiaLavechin2020].
@@
 Unfortunately, those events lasted a long time leading to poor overall agreement.
-In the case of confusion between supine and sitting for the infant who was reclined, the confusion likely stems from the difficult of categorizing behaviors that do not always fit neatly into discrete classes (e.g., continuous body position angles categorized as either supine or sitting).
+In the case of confusion between supine and sitting for the infant who was reclined, the confusion likely stems from the difficulty of categorizing behaviors that do not always fit neatly into discrete classes (e.g., continuous body position angles categorized as either supine or sitting).
 
@@
 Beyond group and individual models, there were many degrees of freedom in our choice of how best to model body position.
-For the sake of brevity, we did not report every possible variation, only the best approach.
-However, our openly shared data mean that researchers can experiment with different sets of motion features, different sets of training data, different machine learning algorithms, and different sets of hyperparameters.
+For the sake of brevity, we report the best approach, not every possible variation in modeling.
+However, our openly shared data and code allow researcher to experiment with different sets of motion features, different sets of training data, different machine learning algorithms, and different hyperparameter values.
 As in past work [@AiraksinenRasanen2020], performance is best when using all four sensors and degrades when using any subset of two (i.e., hips only, left leg only).
@@
 Collecting dense data over the entire variety of daily experiences helps to more accurately measure infants' experiences in aggregate (as in Figure \@ref(fig:age)) without biasing results from a particular type of activity (e.g., play).
-Long-form recordings also have the potential measure clinically-relevant outcomes, such as whether a physical therapy intervention to improve infant sitting skill leads to increases in everyday sitting time.
-But as Figure \@ref(fig:timelines) shows, aggregating responses masks the rich temporal structure contained within long-form recordings.
+Long-form recordings also have the potential to measure clinically-relevant outcomes, such as whether a physical therapy intervention to improve infant sitting skill leads to increases in everyday sitting time.
 
 In the past decade, developmental scientists have discovered that distributional information about infants' experiences matter for how infants learn [@ClerkinHart2017; @KachergisYu2017; @RazAbney2019].
-Skewed distributions that favor the repetition of a small subset of experiences facilitate learning.
+Skewed distributions---those that favor the repetition of a small subset of experiences---facilitate learning.
 For example, recordings from wearable head cameras in the home indicate that infants see a small subset of objects with high frequency [@ClerkinHart2017], but most objects are seen infrequently.
 Heterogeneity is also found in how experiences are distributed in time---burstiness and clustering are seen in infants' daily experiences seeing objects and hearing music.
-Long-form motor recordings provide a novel opportunity to measure the temporal structure of motor experiences.
+Long-form motor recordings provide a novel opportunity to measure the temporal structure of motor experiences that can be seen in Figure \@ref(fig:timelines).
 How skewed are infants' experiences with different motor behaviors, and how are they clustered in time?
-Linking the distributional structure of a particular motor experience in a real day, such as time spent sitting, might be predicted by sitting skill and/or might infants' future sitting ability.
+Linking the distributional structure of a particular motor experience in a real day, such as time spent sitting, might be predicted by sitting skill and/or might predict infants' future sitting ability.
 <!-- CITES FOR BURSTINESS CASILLAS FAUSEY -->

### 2023-03-27 manuscript.tex
Finished minor lab edits

@@
 However, in a real day, non-play activities (e.g., eating lunch in a high chair) create challenging situations for applying automated classification of body position.
+Indeed different types of sitting---independently on the floor, supported on a caregiver's lap, or restrained in a high chair---all need to be counted as equal.
+Moreover, models must be able to new variants of behaviors that occur over the course of natural behavior that did not appear in the training data.
 Will models trained on video-recorded observations at the beginning of the day generalize to predict behavior at a later time?
@@
 Figure \ref{fig:exemplar-timeline}) shows an exemplar timeline of the entire procedure and recording periods for a single participant.
-On the day of the visit, a researcher arrived at the participant's home in the morning between and prepared all the equipment at the doorstep.
+On the day of the visit, a researcher arrived at the participant's home in the morning and prepared all the equipment at the doorstep.
 In order to synchronize all three recording devices, the researcher began by first turning on the video camera and the LENA audio device (the IMUs were already configured and placed before arrival because they required a proprietary sensor dock to begin recording in the laboratory).
@@
 
-Afterwards, they were asked to go about their day as usual with the infant wearing the sensor garment until their bedtime, only taking off the sensor for naps, baths, and diaper changes.
-The caregiver logged the times the sensors were removed (blank areas in the timeline in Figure \ref{fig:exemplar-timeline})) or the child took a nap (gray areas in the timeline in Figure \ref{fig:exemplar-timeline})) so that those times could be excluded from analysis.
+Afterwards, they were asked to go about their day as usual with the infant wearing the sensor garment until their bedtime, only taking off the sensor for naps, baths, diaper changes, and trips out of the house.
+The caregiver logged the times the sensors were removed (blank areas in the timeline in Figure \ref{fig:exemplar-timeline}) or the child took a nap (gray areas in the timeline in Figure \ref{fig:exemplar-timeline}) so that those times could be excluded from analysis.
 The following day a researcher picked up the equipment, verified that the paperwork was signed, and compensated the participant.
@@
 We scored sitting to include any form of the following seated positions: 1) infants sat with their buttocks on a surface, such as on the floor or a caregiver's lap, 2) infant was in a kneeling-sit position, in which their knees were on the ground with their legs tucked underneath the buttocks, and 3) infant was in a seating device, such as a high chair, that kept the torso oriented perpendicular to the ground (a reclined position, such as in a young infant's car seat, would be counted as supine).
-Upright was coded when the infant was standing on the ground with two feet or walking (regardless of whether infants' balance was assisted by a caregiver or with their hands holding onto something for support).
+Upright was coded when the infant was standing or squatting on the ground with two feet or walking (regardless of whether infants' balance was assisted by a caregiver or with their hands holding onto something for support).
 When an infant was carried by a caregiver, held was coded.
@@
 
-Datasets containing one row per 4-s window were used for machine learning classification and validation.
+The resulting windowed datasets were used for machine learning classification and validation.
 For each analysis reported in the results, a subset of data were defined as a ``training'' set and another, independent portion of the data were defined as a ``testing'' set.
@@
 The \emph{tidymodels} set of packages in R (Kuhn \& Wickham, 2020) was used to tune the ``mtry'' and ``ntrees'' hyperparameters, which control the number of columns sampled and the total number of trees created when modeling.
-We use an mtry parameter of 44 and and ntree parameter of 550, however, changing hyperparameters had little impact on model performance.
+We use an mtry parameter of 44 and and ntree parameter of 550; however, changing hyperparameters had little impact on model performance.
 Just like a linear regression model, the resulting random forest model could later be applied to a set of testing data with the \emph{predict} function.
@@
 Each blue circle represents that accuracy for each recording session.
-Horizontal black bars indicates the mean across sessions.}\label{fig:metrics}
+Horizontal black bars indicate the mean across sessions.}\label{fig:metrics}
 \end{figure}
@@
 This leave-one-out cross-validation tested the generalization of the model to a recording session that was not used at all in the training set.
-In contrast, individual models were trained using the first 75\% of a recording session, then tested against the last 25\% of the same session.
+In contrast, individual models were trained using the first 75\% of a recording session, then tested against the (temporally) last 25\% of the same session.
 The individual model tests whether earlier training data generalize to later testing data within an individual participant's recording.
@@
 Overall accuracy for group models (\emph{M} = 0.85) was slightly lower than accuracy for individual models (\emph{M} = 0.92).
-Although overall accuracy from our unsupervised, in-home data collection did not match the near-perfect accuracy (.95-.98) found in prior in-lab studies (Airaksinen et al., 2020; Franchak et al., 2021), both models approached the level of agreement found between two human coders (\emph{M} = .906).
+Although overall accuracy from our semi-supervised, in-home data collection did not match the near-perfect accuracy (.95-.98) found in prior in-lab studies (Airaksinen et al., 2020; Franchak et al., 2021), both models approached the level of agreement found between two human coders (\emph{M} = .906).
 Most likely, lower accuracy in the current study results from the more variable and complex behavior observed in an unsupervised setting rather than from a difference in the quality of the classification model.
 Visual inspection of Figure \ref{fig:metrics}A shows that accuracy values were heavily skewed, with most approaching perfect accuracy but a few participants with very poor accuracy.
-Looking at the median performance suggests that the difference between models was not as wide for the typical participant (group median accuracy = 0.89; individual median accuracy = 0.93.
+Looking at the median performance suggests that the difference between models was not as wide for the typical participant (group median accuracy = 0.89; individual median accuracy = 0.93).
 Indeed, it is notable that the worst-case accuracy from group models (minimum = 0.50) was considerably lower than in individual models (minimum = 0.73).
@@
 Strong overall accuracy can be misleading when class prevalence is unbalanced.
-For example, most infants spent longer sitting than prone, so good performance for sitting classification could mask poorer performance in prone classification.
+For example, most infants spent more time sitting than prone, so good performance for sitting classification could mask poorer performance in prone classification.
 Accordingly, we report Cohen's kappa, a commonly-used metric that penalizes missing rare events (Figure \ref{fig:metrics}B), and provide classification metrics for each individual body position (Table \ref{tab:metrics-by-class}) to account for imbalance in body position rates within and between individuals.
@@
 For group models, sensitivity and PPV were similar: They were highest for supine, prone, and sitting (the most accurately identified class) and lowest for upright and held.
-All values were sufficient (\textgreater{} .8).
 Results for individual models were similar to group models, with the exception of a somewhat lower sensitivity score for held.
@@
 In the first set of results, we showed that group and individual models trained from data during the proximal comparison period were accurate at classifying body position in the same recording.
-In the next analysis, we examine long-term performance by testing how accurately models trained from the proximal period could predict body position during the distal period (Figure \ref{fig:exemplar-timeline}).
+In the next analysis, we examine long-term performance by testing how accurately models trained from the proximal period could predict body position during the distal comparison period.
 A single group model was created using all sessions' proximal period training data (rather than group models leaving out a single session); the same individual models were used.
@@
 Whereas the proximal analyses used all 34 sessions, this was not possible in the distal comparison.
-Because the start of the visit was scheduled during a time the infant was awake, it was common for a nap to be needed following the proximal period.
+Because the start of the visit was scheduled during a time when the infant was awake, it was common for a nap to be needed following the proximal period.
 Nine sessions were excluded because the infant was either napping or otherwise not on camera during the entire 90-minute distal recording.
@@
 The infant also spent a long time in the mother's arms in an ambiguous supine/sitting position that was hard to classify.
-This participant's proximal accuracy was also poor because similar ambiguous appeared during the initial recording.
+This participant's proximal accuracy was also poor because similar ambiguities appeared during the initial recording.
 In contrast, the ``gray diamond outlier'' had strong proximal accuracy, with confusion only arising in the distal period between upright and held categories.
@@
 \caption{Prediction difference (difference in minutes between human-coded and model-predicted body position) for 10-minute bins in the distal comparison period.
-Each point shows the mean and SE for a single participant's recording session for each body position, summarizing the prediction difference for each of their 10-minute bins.
+Each point shows the mean and SE for a single recording session for each body position, summarizing the prediction difference for each of their 10-minute bins.
 Points falling within the gray shaded region indicate that average prediction errors were less than 1 minute.
@@
 We repeated the agreement analysis (correlating human-coded body position prevalence with model-predicted prevalence) after dividing the distal comparison period into nine 10-minute bins (marked by vertical dashed lines in Figure \ref{fig:exemplar-timeline}).
-Infants had varying numbers of 10-minute bins depending on much time they were awake and on camera.
+Infants had varying numbers of 10-minute bins depending on how much time they were awake and on camera.
 Bins were included only if there was \textgreater{} 7 minutes of usable data.
@@
 \caption{Full-day timelines for each individual recording session, split by (A) younger infants and (B) older infants.
-Each participant's timeline shows a stacked bar graph with the proportion of time spent in each of the body position for every 5-minute period throughout the day, based on group model predictions of body position.
+Each participant's timeline shows a stacked bar graph with the proportion of time spent in each of the body positions for every 5-minute period throughout the day, based on group model predictions of body position.
 The x-axis shows time of day.
@@
 Predictions from the group model were used to ensure that motion data were classified consistently across all sessions.
+Moreover, for infants who did not display all 5 body positions in the training period, group models were necessary to be able to predict those behaviors across the entire day.
 Session start times ranged from 09:15 to 13:20 with a median of 10:25.
 Although the equipment was always dropped off in the morning, sometimes infants were unexpectedly asleep, leading to a few sessions in which infants began wearing the sensors later than intended (such as \#11 in the younger group).
-With two exceptions, recordings lasted until the infants' bed time.
+With two exceptions, recordings lasted until the infants' bedtime.
 Older participant \#1 had the equipment picked up on the same day rather than the next day, so the recording ended at 17:00.
@@
 Considering that body position annotation takes 2-5 hours of labor per 1 hour of behavior, the 207 hours of data we recorded would have taken 413-1033 hours of labor to annotate from video.
-However, only an estimated 64-160 hours of labor was needed to annotate the initial hour of each session's video that was used to train the machine learning models---an immense savings in the human cost of annotation that allows the method to scale to larger sample sizes and recording durations.
+However, only an estimated 64-160 hours of labor was needed to annotate the initial hour of each session's video that was used to train the machine learning models---an immense savings in the human labor cost of annotation that allows the method to scale to larger sample sizes and recording durations.
 
@@
 
-In the last analysis, we demonstrate how the full-day recordings can reveal individual differences in body position according to age.
+In the last analysis, we demonstrate how the full-day recordings can reveal individual differences in body position according to age, replicating age differences revealed in past work (Franchak, 2019).
+Table \ref{tab:agetable} summarizes the percent of time that younger and older infants spent in each body position out of their awake samples, as predicted by group and individual models.
 Past work using sparse EMA surveys across the day (Franchak, 2019) found that from 3-12 months, infants spent less time held, reclined, and supine but spent increasingly more time sitting and upright.
 Prone time changed little.
-Table \ref{tab:agetable} summarizes the percent of time that younger and older infants spent in each body position out of their awake samples, as predicted by group and individual models.
-
 The most straightforward comparisons between the two investigations are upright and prone, because they were defined identically.
 In Franchak (2019), upright time was 0.6-5.5\% of the time for infants 3-6 months, increasing to 22.0\% at 12 months.
-Results from the current study were similar: younger infants' sitting time estimated from full-day group models was 5.41\% compared with 18.24\% for the older group.
+Results from the current study were similar: younger infants' upright time estimated from full-day group models was 5.41\% compared with 18.24\% for the older group.
 In Franchak (2019), prone time was 2.9-9.2\% of the time for infants 3-6 months, and was 7.2\% at 12 months.
@@
 Examining full recordings showed that the new method allows us to capture more data (median = 6.14 hours of awake behavior) than is typical with video methods (1-2 hours) while simultaneously reducing the human labor needed to annotate it.
-Ultimately, age differences in body position mirrored past findings that employed other methods, suggesting that the outcome of full-day body position recordings is suitable for describing developmental changes in motor behavior.
+Ultimately, age differences in body position mirrored past findings that employed other methods, suggesting that the outcome of full-day body position recordings is suitable for capturing individual differences and describing developmental changes in motor behavior.
 
@@
 The challenge only increased in the distal comparison period.
-Since sessions were scheduled at convenient times to for the infant and caregiver, the transition from the proximal to distal period increased the odds that infants might need a nap, a snack/meal, or engage in a less typical activity (e.g., watching TV).
+Since sessions were scheduled at convenient times for the infant and caregiver, the transition from the proximal to distal period increased the odds that infants might need a nap, a snack/meal, or engage in a less typical activity (e.g., watching TV).
 Indeed, the distal comparison in Figure \ref{fig:exemplar-timeline} contains a long period (almost 50 minutes) where the infant sat in a high chair eating lunch.
@@
 Acceptable inter-rater reliability in developmental psychology studies of motor behavior is greater than 80\%; both group and individual model results exceed that standard for most recorded sessions.
-Another comparison we can make is to the accuracy of the LENA®`s automated metrics.
+Another comparison we can make is to the accuracy of the LENA® device's automated metrics.
 Correlations between human and model frequencies were \emph{r} = .698 for adult word count and \emph{r} = .649 for child word count (Cristia et al., 2020).
@@
 Beyond group and individual models, there were many degrees of freedom in our choice of how best to model body position.
-For the sake of brevity, we did not report every possible variation, only the best approach.
-However, our openly shared data mean that researchers can experiment with different sets of motion features, different sets of training data, different machine learning algorithms, and different sets of hyperparameters.
+For the sake of brevity, we report the best approach, not every possible variation in modeling.
+However, our openly shared data and code allow researcher to experiment with different sets of motion features, different sets of training data, different machine learning algorithms, and different hyperparameter values.
 As in past work (Airaksinen et al., 2020), performance is best when using all four sensors and degrades when using any subset of two (i.e., hips only, left leg only).
@@
 Collecting dense data over the entire variety of daily experiences helps to more accurately measure infants' experiences in aggregate (as in Figure \ref{fig:age}) without biasing results from a particular type of activity (e.g., play).
-Long-form recordings also have the potential measure clinically-relevant outcomes, such as whether a physical therapy intervention to improve infant sitting skill leads to increases in everyday sitting time.
-But as Figure \ref{fig:timelines} shows, aggregating responses masks the rich temporal structure contained within long-form recordings.
+Long-form recordings also have the potential to measure clinically-relevant outcomes, such as whether a physical therapy intervention to improve infant sitting skill leads to increases in everyday sitting time.
 
 In the past decade, developmental scientists have discovered that distributional information about infants' experiences matter for how infants learn (Clerkin et al., 2017; Kachergis et al., 2017; Raz et al., 2019).
-Skewed distributions that favor the repetition of a small subset of experiences facilitate learning.
+Skewed distributions---those that favor the repetition of a small subset of experiences---facilitate learning.
 For example, recordings from wearable head cameras in the home indicate that infants see a small subset of objects with high frequency (Clerkin et al., 2017), but most objects are seen infrequently.
 Heterogeneity is also found in how experiences are distributed in time---burstiness and clustering are seen in infants' daily experiences seeing objects and hearing music.
-Long-form motor recordings provide a novel opportunity to measure the temporal structure of motor experiences.
+Long-form motor recordings provide a novel opportunity to measure the temporal structure of motor experiences that can be seen in Figure \ref{fig:timelines}.
 How skewed are infants' experiences with different motor behaviors, and how are they clustered in time?
-Linking the distributional structure of a particular motor experience in a real day, such as time spent sitting, might be predicted by sitting skill and/or might infants' future sitting ability.
+Linking the distributional structure of a particular motor experience in a real day, such as time spent sitting, might be predicted by sitting skill and/or might predict infants' future sitting ability.
 

### 2023-03-27 manuscript.Rmd
Clean through to methods

@@
 authornote: |
-  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.
-
-  Enter author note here.
+  We are grateful to Vanessa Scott, Tasnia Haider, and Ishapreet Kaur for their help in collecting data for the present study and to the research assistants of the UCR Perception, Action, and Development Lab for annotating videos.
+
+  This work was funded by National Science Foundation Grant BCS #1941449 to the first author.
 
@@
 
-# SUBMITTED DATA IS BASED ON THE FOLLOWING IDS
+# SUBMITTED DATA IS BASED ON THE FOLLOWING SESSION IDS
 #9908  9909  9910  9911 10201 10203 10204 10301 10501 10601 10701 10702 10703 10704 10802 10901 11001 11002 11004 11102 11103 11104 11202 11501 11601 11901 12001 12002 12101 12302 12303 12501 12502 13001
-
 ```
@@
 Finally, we discuss the advantages of using inertial sensing to classify motor behavior.
-Despite promising past results in brief, supervised sessions [@FranchakScott2021; @AiraksinenRasanen2020; @AiraksinenGallen2022], the current investigation takes a needed step forward by testing accuracy over long, unsupervised recordings.
+Despite promising past results in brief, supervised sessions [@FranchakScott2021; @AiraksinenRasanen2020; @AiraksinenGallen2022], the current investigation takes a needed step forward by testing accuracy over long, unsupervised recordings in what we term a *distal comparison*.
 Although in the current investigation we focus on classification of body position, this approach can be extended to categorize other aspects of movement, such as locomotion [@AiraksinenRasanen2020] and infant-caregiver contact [@YaoPlotz2019].
-<!-- Accompanying our paper, we have created a database [CITE] containing synchronized video and IMU data, allowing researchers to annotate and train models for behaviors beyond body position.  -->
+We provide openly shared data, materials, and analysis code to facilitate other researchers' adopting this method and applying it to populations and behaviors (see Data Sharing and Transparency below).
 
@@
 
-The LENA&reg; recorder is a commercial device that is worn by infants in a custom shirt pocket; the recorder has sufficient battery life and storage to record for an entire day.
+The LENA&reg; recorder is a commercial device worn in a custom shirt pocket; the recorder has sufficient battery life and storage to record for an entire day.
 Closed-source LENA&reg; algorithms analyze the audio recordings to provide automatic counts of useful metrics, such as the number of words spoken by adults in the vicinity of the infant.
@@
 Full-day language recording synchronized with other data sources allows researchers to identify how auditory input and vocal production interact with other processes.
-Beyond individual differences in aggregated data, long-form recordings can be used to determine the temporal schedule of experiences.
+Beyond individual differences in aggregated data, long-form recordings reveal the temporal schedule of experiences.
 For example, infants' daily experiences hearing music are clustered in time, with "bursty" episodes of music presence separated by relatively long periods during which music is absent [@MendozaFausey2022].
@@
 The ability to record real-time data is vital for making inferences about processes that happen on the timescale of minutes or even seconds within an individual as opposed to comparisons of aggregated data between infants.
-Synchronizing real-time data to other data streams helps to reveal sources of variability within an individual [e.g., @WassPhillips2022].
+Synchronizing real-time data to other data streams helps to reveal sources of variability within an individual [e.g., @WassPhillips2022; @MalachowskiSalo2023].
 Full-day recordings are essential for capturing experiences across the heterogeneity of daily routines that moderate behavior (e.g., play, feeding, errands) [@KadookaCaufield2021; @Tamis-LeMondaCustode2018].
@@
 Fifth, *automatic classification* means that the approach can scale to analyze large numbers of participants over long recordings without the bottleneck of manual annotation/transcription.
+
 But automatic classification can only replace human annotation if it is sufficiently accurate and unbiased.
-An independent assessment of the LENA&reg; algorithms found mixed results about the accuracy of different outcomes.
+An independent assessment of the LENA&reg; algorithms found mixed results about the accuracy of different measurements.
 For example, correlations between human transcribed counts of adult words and child vocalizations against LENA&reg;'s automatic counts were strong, *r* = .698 and *r* = .649, respectively [@CristiaLavechin2020].
 For other metrics, such as the number of "conversational turns" between the child and communicative partners and the identification of male speakers, agreement was poor.
-<!-- EVENTUALLY CITE NEW EMA HERE -->
-
 Thus, for some use cases (and for some metrics), long form audio recordings provide a mobile, unobtrusive way to automatically score real-time data over a full day.
@@
 
-Video and survey methods are the current state-of-the-art in assessing motor behavior.
+Video and survey methods are the current state-of-the-art in assessing infants' gross motor behavior in naturalistic tasks.
 Although each method has complementary advantages and disadvantages for characterizing infants' everyday motor experiences, neither method on its own can provide comparable data to the long form audio recordings reviewed in the previous section.
@@
 However, in a real day, non-play activities (e.g., eating lunch in a high chair) create challenging situations for applying automated classification of body position.
-Indeed different types of sitting---independently on the floor, supported on a caregiver's lap, or restrained in a high chair---all need to be counted as equal.
-Moreover, models must be able to new variants of behaviors that occur over the course of natural behavior that did not appear in the training data.
-Will models trained on video-recorded observations at the beginning of the day generalize to predict behavior at a later time?
-Assessing the validity of temporally *distal* periods is a crucial step to establish whether automatic classification can be used to measure body position across a day.
+For example different types of sitting---independently on the floor, supported on a caregiver's lap, or restrained in a high chair---all need to scored as the same behavior.
+Moreover, classifiers must be able to detect new variants of behaviors that might arise over the course of a real day; it is impossible for researchers to gather training data for every possible variation that might occur.
+Thus, our central question is whether models trained on video-recorded observations at the beginning of the day generalize to predict behavior at a later time?
+Assessing the validity of temporally *distal* periods---when infants and caregivers are unsupervised and free to follow their everyday routines---is a crucial step to establish whether automatic classification can be used to measure body position across a day.
 
@@
 A *proximal comparison* period began when participants received the equipment and completed a guided phone call during which caregivers were asked to elicit different body positions based on prompts from the experimenter.
-Although not directly supervised, this period was most similar to previous recordings because it occurred during a convenient time for the infant and caregiver to play while they received instructions from the experimenter.
+This "semi-supervised" period was most similar to previous recordings because it occurred during a convenient time for the infant and caregiver to play while they received instructions from the experimenter.
 The **first goal** of the current study was to determine the accuracy of body position classification during the proximal comparison period using this novel, semi-supervised procedure in participants' homes.
@@
 A second, *distal comparison* period followed the proximal comparison period and captured approximately 90 minutes of home behavior that was completely unsupervised.
-Caregivers and infants could (and did) do whatever they wished.
+Caregivers and infants could (and did) do whatever they wished, and no researcher was present.
 Because this recording happened a considerable amount of time after the initial setup and instructions from the experimenter, accuracy could decline if caregivers or infants moved the garment or sensors.
@@
 Could body position be successfully recorded over the desired period?
-The novel full-day dataset affords us a unique chance to ask whether the estimated time infants spent in different body positions align with past results using video and survey methods.
-Thus, we determined whether full-day body position measurements conformed to expected age differences in body position.
+If full-day classifications of infant behavior are accurate, they should demonstrate convergent validity with other methods.
+Thus, we determined whether full-day body position measurements conformed to expected age differences in body position that have been identified from previous video and survey methods.
 Based on past results [@Survey], infants should spend increasingly more time sitting and upright but less time supine over the age range tested (4 to 14 months).
@@
 Infants were recruited in one of two age groups: *Younger* infants were between 4 and 7 months and *older* infants were between 11 and 14 months.
-There were `r num_younger` infants in the 4-7 month group (X female) and `r num_older` in the 11-14 month group (X female).
+There were `r num_younger` infants in the 4-7 month group and `r num_older` in the 11-14 month group.
+Ten infants were female and 12 were male.
 Families were recruited through social media advertisements and from community events in Southern California.
-Caregivers reported the ethnicity of infants as X Hispanix/Latinx and X for not Hispanix/Latinx.
-Race was reported as X Caucasian, X Asian, X ~~~ and X others.
+Most infants were reported as either Hispanic and White or Non-Hispanic and White.
 Families were compensated \$30 for every home recording session they completed.
@@
 All caregivers gave their informed consent before the start of the study.
-<!-- TODO RACE ETHNICITY SEX, COUNT BY HAND -->
 
@@
 
-<!-- ## Data Sharing and Transparency -->
+## Data Sharing and Transparency
 <!-- WHAT CAN WE SHARE AND HOW -->

### 2023-03-27 manuscript.tex
Clean through to methods

@@
 
-Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.
-
-Enter author note here.
+We are grateful to Vanessa Scott, Tasnia Haider, and Ishapreet Kaur for their help in collecting data for the present study and to the research assistants of the UCR Perception, Action, and Development Lab for annotating videos.
+
+This work was funded by National Science Foundation Grant BCS \#1941449 to the first author.
 
@@
 Finally, we discuss the advantages of using inertial sensing to classify motor behavior.
-Despite promising past results in brief, supervised sessions (Airaksinen et al., 2022, 2020; Franchak et al., 2021), the current investigation takes a needed step forward by testing accuracy over long, unsupervised recordings.
+Despite promising past results in brief, supervised sessions (Airaksinen et al., 2022, 2020; Franchak et al., 2021), the current investigation takes a needed step forward by testing accuracy over long, unsupervised recordings in what we term a \emph{distal comparison}.
 Although in the current investigation we focus on classification of body position, this approach can be extended to categorize other aspects of movement, such as locomotion (Airaksinen et al., 2020) and infant-caregiver contact (Yao et al., 2019).
+We provide openly shared data, materials, and analysis code to facilitate other researchers' adopting this method and applying it to populations and behaviors (see Data Sharing and Transparency below).
 
@@
 
-The LENA® recorder is a commercial device that is worn by infants in a custom shirt pocket; the recorder has sufficient battery life and storage to record for an entire day.
+The LENA® recorder is a commercial device worn in a custom shirt pocket; the recorder has sufficient battery life and storage to record for an entire day.
 Closed-source LENA® algorithms analyze the audio recordings to provide automatic counts of useful metrics, such as the number of words spoken by adults in the vicinity of the infant.
@@
 Full-day language recording synchronized with other data sources allows researchers to identify how auditory input and vocal production interact with other processes.
-Beyond individual differences in aggregated data, long-form recordings can be used to determine the temporal schedule of experiences.
+Beyond individual differences in aggregated data, long-form recordings reveal the temporal schedule of experiences.
 For example, infants' daily experiences hearing music are clustered in time, with ``bursty'' episodes of music presence separated by relatively long periods during which music is absent (Mendoza \& Fausey, 2022).
@@
 The ability to record real-time data is vital for making inferences about processes that happen on the timescale of minutes or even seconds within an individual as opposed to comparisons of aggregated data between infants.
-Synchronizing real-time data to other data streams helps to reveal sources of variability within an individual (e.g., Wass et al., 2022).
+Synchronizing real-time data to other data streams helps to reveal sources of variability within an individual (Malachowski et al., 2023; e.g., Wass et al., 2022).
 Full-day recordings are essential for capturing experiences across the heterogeneity of daily routines that moderate behavior (e.g., play, feeding, errands) (Kadooka et al., 2021, April; Tamis-LeMonda et al., 2018).
@@
 Fifth, \emph{automatic classification} means that the approach can scale to analyze large numbers of participants over long recordings without the bottleneck of manual annotation/transcription.
+
 But automatic classification can only replace human annotation if it is sufficiently accurate and unbiased.
@@
 For other metrics, such as the number of ``conversational turns'' between the child and communicative partners and the identification of male speakers, agreement was poor.
-
 Thus, for some use cases (and for some metrics), long form audio recordings provide a mobile, unobtrusive way to automatically score real-time data over a full day.
@@
 Infants were recruited in one of two age groups: \emph{Younger} infants were between 4 and 7 months and \emph{older} infants were between 11 and 14 months.
-There were 8 infants in the 4-7 month group (X female) and 14 in the 11-14 month group (X female).
+There were 8 infants in the 4-7 month group and 14 in the 11-14 month group.
+Ten infants were female and 12 were male.
 Families were recruited through social media advertisements and from community events in Southern California.
-Caregivers reported the ethnicity of infants as X Hispanix/Latinx and X for not Hispanix/Latinx.
-Race was reported as X Caucasian, X Asian, X \textasciitilde\textasciitilde\textasciitilde{} and X others.
+Most infants were reported as either Hispanic and White or Non-Hispanic and White.
 Families were compensated \$30 for every home recording session they completed.
@@
 Validation analyses below will compare the human-coded annotations and random forest model predictions for testing data that are independent from training data, providing a test of whether models can successfully generalize.
+
+\hypertarget{data-sharing-and-transparency}{%
+\subsection{Data Sharing and Transparency}\label{data-sharing-and-transparency}}
 

### 2023-03-27 manuscript.Rmd
up to results

@@
 Families were recruited through social media advertisements and from community events in Southern California.
-Most infants were reported as either Hispanic and White or Non-Hispanic and White.
+Most infants were reported to be either Hispanic and White or Non-Hispanic and White.
 Families were compensated \$30 for every home recording session they completed.
@@
 Across the two age groups, we report data on a total of `r length(unique(ppts$id_uni))` sessions, with `r num_sessions %>% filter(group == "Younger") %>% pull(n) %>% sum` sessions from younger infants and `r num_sessions %>% filter(group == "Older") %>% pull(n) %>% sum` sessions from older infants.
-Across sessions, younger infants' age ranged from `r apa_num(age["Younger","min"], digits = 1)` to `r apa_num(age["Younger","max"], digits = 1)` months (*M* = `r apa_num(age["Younger","mean"], digits = 1)`) and older infants' age ranged from `r apa_num(age["Older","min"], digits = 1)` to `r apa_num(age["Older","max"], digits = 1)` months (*M* = `r apa_num(age["Older","mean"], digits = 1)`) .
+Across sessions, younger infants' age ranged from `r apa_num(age["Younger","min"], digits = 1)` to `r apa_num(age["Younger","max"], digits = 1)` months (*M* = `r apa_num(age["Younger","mean"], digits = 1)`) and older infants' age ranged from `r apa_num(age["Older","min"], digits = 1)` to `r apa_num(age["Older","max"], digits = 1)` months (*M* = `r apa_num(age["Older","mean"], digits = 1)`).
 
@@
 The camera created two video files temporally separated by a gap of approximately 40-45 s.
-Caregivers also received a log sheet to record times that infants napped as well as times that the sensor garment was removed from the infant (e.g., baths, diaper changes).
+Caregivers also received a log sheet to record times that infants napped as well as times that the sensor garment was removed from the infant (e.g., baths, diaper changes, errands).
 
@@
 
-Figure \@ref(fig:exemplar-timeline)) shows an exemplar timeline of the entire procedure and recording periods for a single participant.
+Figure \@ref(fig:exemplar-timeline) shows an exemplar timeline of the entire procedure and recording periods for a single participant.
 On the day of the visit, a researcher arrived at the participant’s home in the morning and prepared all the equipment at the doorstep.
@@
 As seen in the bottom of Figure \@ref(fig:exemplar-timeline)), the *video period* comprised the first three hours of recording starting from the researcher's arrival when they turned on the camera.
-The first 90-minute video file, termed the *proximal comparison*, contained the activities during the guided call followed by a period of infants and caregivers resuming their normal activities.
+The first 90-minute video file, termed the *proximal comparison*, contained the activities during the guided call followed by a period during which infants and caregivers resumed their normal activities.
 Because this video contained the synchronization point, the data in this period had high temporal synchrony between IMU and video data.
-Synchrony errors were less than 30-60 ms (1-2 video frames).
-When the first video file ended, a second video file was recorded, termed the *distal comparison*.
+Synchronization errors were estimated to be less than 30-60 ms (1-2 video frames).
+When the first video file ended, a second video file was automatically recorded, termed the *distal comparison*.
 This video recorded the next 90 minutes of natural activity.
-However, because there was a variable gap of ~40 s between the two videos, temporal synchrony in the distal comparison video was approximate containing offsets of ~ 5 s in either direction.
+However, because there was a variable gap of ~40 s between the two videos, synchronization in the distal comparison video was coarser, with estimated temporal offsets of ~5 s in either direction.
 
@@
 All coding was done using Datavyu software (datavyu.org).
+A Databrary repository contains the entire video recording, coding files, and raw IMU data for a single participant (https://nyu.databrary.org/volume/1580).
+<!-- CODE OCEAN HERE -->
 
@@
 Times during the video when the infant was out of view were excluded.
-Periods when the leggings were being adjusted or taken off the infant were also excluded, as were transitions between body positions.
+Periods when the sensor garment was adjusted or taken off the infant were also excluded, as were transitions between body positions.
 
@@
 Interrater reliability was based on the proportion of video frames that the two coders chose the same body position code.
-Overall agreement averaged 90.9% across video files, ranging from 68.4%-100%.
-Cohen's *kappa* averaged 86.1% across video files, ranging from 31.0%-100%.
+Overall agreement averaged 90.9% across video files, ranging from 68.4%-100% for individual video files.
+Cohen's kappa averaged 86.1% across video files, ranging from 31.0%-100% for individual video files.
 
@@
 A single, merged dataset was created with synchronized accelerometer signals (in three orientations: X, Y, and Z) and gyroscope signals (in three orientations: roll, pitch, and yaw) for each of the four sensors (left hip, right hip, left ankle, right ankle) with the corresponding timestamp and body position code using the *timetk* package [@DanchoVaughan2023] and the *lubridate* package [@GrolemundWickham2011] in R version 4.1.2 [@R42].
+<!-- CODE OCEAN HERE -->
 
@@
 The 436 total motion features corresponded to a single body annotation code for each 4-s window.
-Windows were only used for training/testing if they contained a single body position for > 75% (3 s) of time within the window to ensure that motion signals could be tied to a clear example of each behavior.
-
-The resulting windowed datasets were used for machine learning classification and validation.
+Windows were only used for training/testing if they contained a single body position for > 75% (3 s) of time within the window to ensure that motion signals could be linked to an unambiguous example of each behavior.
+
+The resulting windowed dataset was used for machine learning classification and validation.
 For each analysis reported in the results, a subset of data were defined as a "training" set and another, independent portion of the data were defined as a "testing" set.
@@
 ## Data Sharing and Transparency
-<!-- WHAT CAN WE SHARE AND HOW -->
-https://nyu.databrary.org/volume/1580 - all raw data files and annotations for one participant
-
-This reproducible manuscript showing how we accumulated
+
+Three online repositories contain openly shared data, materials, and analysis code.
+A Databrary repository (https://nyu.databrary.org/volume/1580) includes an exemplar participant's recording session, with the raw video data files, the Datavyu annotations of those video files, a log file with machine-readable synchronization points and nap/diaper change times, and accelerometer and gyroscope data for each of the 4 sensors.
+A CodeOcean capsule (LINK) hosts an R environment with the exemplar participant's data and source code to: 1) synchronize IMU and video annotations, 2) calculate windowed motion features for their data, and 3) train and test the body position classifier using an "individual model".
+The chosen participant's classification accuracy was at the median level.
+Because of the overall size of the full data set and computational power/time required to synchronize and create windowed datasets, it would not be feasible to create a reproducible environment to process the entire dataset.
+However, in a second CodeOcean capsule (LINK) we share the entire dataset of windowed motion features with corresponding body position codes used to validate the method (described in the results below), along with the processed datafiles used for each set of analyses.
+This reproducible manuscript created in *papaja* [@papaja] can be executed in the CodeOcean environment for full computational transparency and can be downloaded from Github (LINK).
+<!-- FINAL LINKS -->
 

### 2023-03-27 manuscript.tex
up to results

@@
 But automatic classification can only replace human annotation if it is sufficiently accurate and unbiased.
-An independent assessment of the LENA® algorithms found mixed results about the accuracy of different outcomes.
+An independent assessment of the LENA® algorithms found mixed results about the accuracy of different measurements.
 For example, correlations between human transcribed counts of adult words and child vocalizations against LENA®'s automatic counts were strong, \emph{r} = .698 and \emph{r} = .649, respectively (Cristia et al., 2020).
@@
 However, in a real day, non-play activities (e.g., eating lunch in a high chair) create challenging situations for applying automated classification of body position.
-Indeed different types of sitting---independently on the floor, supported on a caregiver's lap, or restrained in a high chair---all need to be counted as equal.
-Moreover, models must be able to new variants of behaviors that occur over the course of natural behavior that did not appear in the training data.
-Will models trained on video-recorded observations at the beginning of the day generalize to predict behavior at a later time?
-Assessing the validity of temporally \emph{distal} periods is a crucial step to establish whether automatic classification can be used to measure body position across a day.
+For example different types of sitting---independently on the floor, supported on a caregiver's lap, or restrained in a high chair---all need to scored as the same behavior.
+Moreover, classifiers must be able to detect new variants of behaviors that might arise over the course of a real day; it is impossible for researchers to gather training data for every possible variation that might occur.
+Thus, our central question is whether models trained on video-recorded observations at the beginning of the day generalize to predict behavior at a later time?
+Assessing the validity of temporally \emph{distal} periods---when infants and caregivers are unsupervised and free to follow their everyday routines---is a crucial step to establish whether automatic classification can be used to measure body position across a day.
 
@@
 A \emph{proximal comparison} period began when participants received the equipment and completed a guided phone call during which caregivers were asked to elicit different body positions based on prompts from the experimenter.
-Although not directly supervised, this period was most similar to previous recordings because it occurred during a convenient time for the infant and caregiver to play while they received instructions from the experimenter.
+This ``semi-supervised'' period was most similar to previous recordings because it occurred during a convenient time for the infant and caregiver to play while they received instructions from the experimenter.
 The \textbf{first goal} of the current study was to determine the accuracy of body position classification during the proximal comparison period using this novel, semi-supervised procedure in participants' homes.
@@
 A second, \emph{distal comparison} period followed the proximal comparison period and captured approximately 90 minutes of home behavior that was completely unsupervised.
-Caregivers and infants could (and did) do whatever they wished.
+Caregivers and infants could (and did) do whatever they wished, and no researcher was present.
 Because this recording happened a considerable amount of time after the initial setup and instructions from the experimenter, accuracy could decline if caregivers or infants moved the garment or sensors.
@@
 Could body position be successfully recorded over the desired period?
-The novel full-day dataset affords us a unique chance to ask whether the estimated time infants spent in different body positions align with past results using video and survey methods.
-Thus, we determined whether full-day body position measurements conformed to expected age differences in body position.
+If full-day classifications of infant behavior are accurate, they should demonstrate convergent validity with other methods.
+Thus, we determined whether full-day body position measurements conformed to expected age differences in body position that have been identified from previous video and survey methods.
 Based on past results (Franchak, 2019), infants should spend increasingly more time sitting and upright but less time supine over the age range tested (4 to 14 months).
@@
 Across the two age groups, we report data on a total of 34 sessions, with 14 sessions from younger infants and 20 sessions from older infants.
-Across sessions, younger infants' age ranged from 3.8 to 7.2 months (\emph{M} = 5.2) and older infants' age ranged from 10.7 to 14.2 months (\emph{M} = 11.7) .
+Across sessions, younger infants' age ranged from 3.8 to 7.2 months (\emph{M} = 5.2) and older infants' age ranged from 10.7 to 14.2 months (\emph{M} = 11.7).
 
@@
 
-Figure \ref{fig:exemplar-timeline}) shows an exemplar timeline of the entire procedure and recording periods for a single participant.
+Figure \ref{fig:exemplar-timeline} shows an exemplar timeline of the entire procedure and recording periods for a single participant.
 On the day of the visit, a researcher arrived at the participant's home in the morning and prepared all the equipment at the doorstep.
@@
 As seen in the bottom of Figure \ref{fig:exemplar-timeline}), the \emph{video period} comprised the first three hours of recording starting from the researcher's arrival when they turned on the camera.
-The first 90-minute video file, termed the \emph{proximal comparison}, contained the activities during the guided call followed by a period of infants and caregivers resuming their normal activities.
+The first 90-minute video file, termed the \emph{proximal comparison}, contained the activities during the guided call followed by a period during which infants and caregivers resumed their normal activities.
 Because this video contained the synchronization point, the data in this period had high temporal synchrony between IMU and video data.
-Synchrony errors were less than 30-60 ms (1-2 video frames).
-When the first video file ended, a second video file was recorded, termed the \emph{distal comparison}.
+Synchronization errors were estimated to be less than 30-60 ms (1-2 video frames).
+When the first video file ended, a second video file was automatically recorded, termed the \emph{distal comparison}.
 This video recorded the next 90 minutes of natural activity.
-However, because there was a variable gap of \textasciitilde40 s between the two videos, temporal synchrony in the distal comparison video was approximate containing offsets of \textasciitilde{} 5 s in either direction.
+However, because there was a variable gap of \textasciitilde40 s between the two videos, synchronization in the distal comparison video was coarser, with estimated temporal offsets of \textasciitilde5 s in either direction.
 
@@
 All coding was done using Datavyu software (datavyu.org).
+A Databrary repository contains the entire video recording, coding files, and raw IMU data for a single participant (\url{https://nyu.databrary.org/volume/1580}).
 
@@
 Interrater reliability was based on the proportion of video frames that the two coders chose the same body position code.
-Overall agreement averaged 90.9\% across video files, ranging from 68.4\%-100\%.
-Cohen's \emph{kappa} averaged 86.1\% across video files, ranging from 31.0\%-100\%.
+Overall agreement averaged 90.9\% across video files, ranging from 68.4\%-100\% for individual video files.
+Cohen's kappa averaged 86.1\% across video files, ranging from 31.0\%-100\% for individual video files.
 
@@
 The 436 total motion features corresponded to a single body annotation code for each 4-s window.
-Windows were only used for training/testing if they contained a single body position for \textgreater{} 75\% (3 s) of time within the window to ensure that motion signals could be tied to a clear example of each behavior.
-
-The resulting windowed datasets were used for machine learning classification and validation.
+Windows were only used for training/testing if they contained a single body position for \textgreater{} 75\% (3 s) of time within the window to ensure that motion signals could be linked to an unambiguous example of each behavior.
+
+The resulting windowed dataset was used for machine learning classification and validation.
 For each analysis reported in the results, a subset of data were defined as a ``training'' set and another, independent portion of the data were defined as a ``testing'' set.
@@
 
-\url{https://nyu.databrary.org/volume/1580} - all raw data files and annotations for one participant
-
-This reproducible manuscript showing how we accumulated
+Three online repositories contain openly shared data, materials, and analysis code.
+A Databrary repository (\url{https://nyu.databrary.org/volume/1580}) includes an exemplar participant's recording session, with the raw video data files, the Datavyu annotations of those video files, a log file with machine-readable synchronization points and nap/diaper change times, and accelerometer and gyroscope data for each of the 4 sensors.
+A CodeOcean capsule (LINK) hosts an R environment with the exemplar participant's data and source code to: 1) synchronize IMU and video annotations, 2) calculate windowed motion features for their data, and 3) train and test the body position classifier using an ``individual model''.
+The chosen participant's classification accuracy was at the median level.
+Because of the overall size of the full data set and computational power/time required to synchronize and create windowed datasets, it would not be feasible to create a reproducible environment to process the entire dataset.
+However, in a second CodeOcean capsule (link) we share the entire dataset of windowed motion features with corresponding body position codes used to validate the method (described in the results below), along with the processed datafiles used for each set of analyses.
+This reproducible manuscript created in \emph{papaja} (Aust \& Barth, 2022) can be executed in the CodeOcean environment for full computational transparency.
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-papaja}{}}%
+Aust, F., \& Barth, M. (2022). \emph{{papaja: Prepare reproducible APA journal articles with R Markdown}}. \url{https://github.com/crsh/papaja}
+
 \leavevmode\vadjust pre{\hypertarget{ref-Barbaro2019}{}}%

### 2023-03-27 manuscript.Rmd
revised whole paper

@@
 The chosen participant's classification accuracy was at the median level.
-Because of the overall size of the full data set and computational power/time required to synchronize and create windowed datasets, it would not be feasible to create a reproducible environment to process the entire dataset.
-However, in a second CodeOcean capsule (LINK) we share the entire dataset of windowed motion features with corresponding body position codes used to validate the method (described in the results below), along with the processed datafiles used for each set of analyses.
-This reproducible manuscript created in *papaja* [@papaja] can be executed in the CodeOcean environment for full computational transparency and can be downloaded from Github (LINK).
+Because of the overall size of the full dataset and the computational power/time required to synchronize and create windowed datasets for each session, it would not be feasible to create a reproducible environment to process all 34 sessions.
+However, in a second CodeOcean capsule (LINK) we share the entire result: The dataset of windowed motion features with corresponding body position codes used to validate the method (described in the results below), along with the processed intermediary files used for each set of analyses.
+This reproducible manuscript created in *papaja* [@papaja] can be executed in the CodeOcean environment for full computational transparency, and can be downloaded from Github (LINK).
 <!-- FINAL LINKS -->
@@
 Second, we applied models trained from data during the proximal comparison to predict infant behavior during the distal comparison, when coarser synchrony between video and motion data was available.
-This distal comparison is critical because it provides the first-ever test of how well infant body classification models predict behavior beyond the initial, supervised period.
+This distal comparison is critical because it provides the first-ever test of how well infant body classification models predict behavior that is variable, unrestricted, and unsupervised.
 Data from the first two sets of analyses confirm that body position classification models are accurate across a long period of time, suggesting that data from the entire day will be valid.
@@
 As in past work [@FranchakScott2021], we compared two types of models: *group models* and *individual models*.
-To assess the accuracy of body position classification for each recording session, we reserved the last 25% of a session's proximal comparison data as the testing set.
+To assess the accuracy of body position classification for each recording session, we reserved the (temporally) last 25% of a session's proximal comparison data as the testing set.
 The testing set was never used as training data.
@@
 Although overall accuracy from our semi-supervised, in-home data collection did not match the near-perfect accuracy  (.95-.98) found in prior in-lab studies [@FranchakScott2021; @AiraksinenRasanen2020], both models approached the level of agreement found between two human coders (*M* = .906).
-Most likely, lower accuracy in the current study results from the more variable and complex behavior observed in an unsupervised setting rather than from a difference in the quality of the classification model.
-Visual inspection of Figure \@ref(fig:metrics)A shows that accuracy values were heavily skewed, with most approaching perfect accuracy but a few participants with very poor accuracy.
-Looking at the median performance suggests that the difference between models was not as wide for the typical participant (group median accuracy = `r metrics_lookup["Overall Accuracy", "median_Group"]`; individual median accuracy = `r metrics_lookup["Overall Accuracy", "median_Individual"]`).
-Indeed, it is notable that the worst-case accuracy from group models (minimum = `r metrics_lookup["Overall Accuracy", "min_Group"]`) was considerably lower than in individual models (minimum = `r metrics_lookup["Overall Accuracy", "min_Individual"]`).
+Most likely, lower accuracy in the current study results from the more variable and complex behavior observed in a a semi-supervised setting rather than from a difference in the quality of the classification model; although the first part of the proximal period was guided by the experimenter during a brief phone call, the remainder of the proximal period included natural behavior.
+Visual inspection of Figure \@ref(fig:metrics)A shows that accuracy values were heavily skewed, with many approaching perfect accuracy but a few sessions with very poor accuracy.
+Looking at the median performance suggests that the difference between models was not considerable for the typical participant (group median accuracy = `r metrics_lookup["Overall Accuracy", "median_Group"]`; individual median accuracy = `r metrics_lookup["Overall Accuracy", "median_Individual"]`).
+Indeed, it is notable that the worst-case accuracy from group models (minimum = `r metrics_lookup["Overall Accuracy", "min_Group"]`) was far lower than in individual models (minimum = `r metrics_lookup["Overall Accuracy", "min_Individual"]`).
 Possibly, individual models accounted for idiosyncrasies in behavior or inconsistencies in sensor placement for those sessions that group models could not generalize to.
@@
 ### Cohen's Kappa 
-Strong overall accuracy can be misleading when class prevalence is unbalanced.
+Strong overall accuracy can be misleading when the relative frequency of different classes is unbalanced.
 For example, most infants spent more time sitting than prone, so good performance for sitting classification could mask poorer performance in prone classification.
-Accordingly, we report Cohen's kappa, a commonly-used metric that penalizes missing rare events (Figure \@ref(fig:metrics)B), and provide classification metrics for each individual body position (Table \@ref(tab:metrics-by-class)) to account for imbalance in body position rates within and between individuals.
+Accordingly, we report Cohen's kappa, a commonly-used metric that penalizes missing rare events (Figure \@ref(fig:metrics)B), and we provide classification metrics for each individual body position (Table \@ref(tab:metrics-by-class)) to account for imbalance in body position rates within and between individuals.
 Similar to overall accuracy, kappa values were strong for both model types with group kappas (*M* = `r metrics_lookup["Kappa", "mean_Group"]`) somewhat worse compared with individual kappas (*M* = `r metrics_lookup["Kappa", "mean_Individual"]`).
@@
 Figure \@ref(fig:part2overall) and Table \@ref(tab:pt2overalltable) summarize the overall agreement during the distal comparison period.
-For each session, we calculated the actual prevalence of each body position as a percentage of the time that the infant was on video and awake using human annotated body position (x-axis on Figure \@ref(fig:part2overall)).
-Predicted prevalence was calculated the same way for group and individual model predictions, omitting the off-camera and nap periods to make a direct comparison.
-Overall---across participants and across body position classes---agreement was strong: The correlation between group model predictions and human-coded prevalence was *r* = `r group_agree_cor`, and the correlation between individual model predictions and human-coded prevalence was *r* = `r split_agree_cor`.
+For each session, we calculated the actual time spent in each body position (out of times the infant was visible on camera and awake) using human annotated body position (x-axis on Figure \@ref(fig:part2overall)).
+Predicted time was calculated the same way for group and individual model predictions, omitting the off-camera and nap periods to make a direct comparison.
+Overall---across participants and across body position classes---agreement was strong: The correlation between group model predictions and human-coded time was *r* = `r group_agree_cor`, and the correlation between individual model predictions and human-coded time was *r* = `r split_agree_cor`.
 As in the proximal comparison, agreement varied somewhat between body positions; in particular, agreement for held was poor.
@@
 This participant's proximal accuracy was also poor because similar ambiguities appeared during the initial recording.
-In contrast, the "gray diamond outlier" had strong proximal accuracy, with confusion only arising in the distal period between upright and held categories.
+In contrast, the "gray diamond" outlier had strong proximal accuracy, with confusion only arising in the distal period between upright and held categories.
 Reviewing the video showed that all disagreements occurred when the infant was in a baby walker; human coders scored this as "upright" but the models predicted it as "held".
@@
 Although overall aggregate agreement in the distal comparison was strong, it is important to show that similarly strong agreement is found within a shorter timescale.
-We repeated the agreement analysis (correlating human-coded body position prevalence with model-predicted prevalence) after dividing the distal comparison period into nine 10-minute bins (marked by vertical dashed lines in Figure \@ref(fig:exemplar-timeline)).
+We repeated the agreement analysis (correlating human-coded body position time with model-predicted time) after dividing the distal comparison period into nine 10-minute bins (marked by vertical dashed lines in Figure \@ref(fig:exemplar-timeline)).
 Infants had varying numbers of 10-minute bins depending on how much time they were awake and on camera.
@@
 The first two sets of results show that accuracy was consistently high across the proximal and distal recordings, providing confidence that predictions over the remainder of the day would continue to be accurate after the video recordings stopped.
-This leads to two final questions---how successfully did recordings capture infants' entire day, and how suitable were day-long recordings for capturing individual differences in body position?
+This leads to two final questions---how successfully did recordings capture infants' entire day, and how well do findings from full-day classification converge with findings that use other methodologies?
 
@@
 Nap times reported during the recording period ranged from `r apa_num(min(coverage$naptime))`-`r apa_num(max(coverage$naptime))` hours with a median of `r apa_num(median(coverage$naptime))` hours.
-Body position data was available to describe `r apa_num(min(coverage$prop_accounted)*100, digits = 1)`-`r apa_num(max(coverage$prop_accounted)*100, digits = 1)`% of the awake portion of the recording period (median = `r apa_num(median(coverage$prop_accounted)*100, digits = 1)`%).
+Body position data were available to describe `r apa_num(min(coverage$prop_accounted)*100, digits = 1)`-`r apa_num(max(coverage$prop_accounted)*100, digits = 1)`% of the awake portion of the recording period (median = `r apa_num(median(coverage$prop_accounted)*100, digits = 1)`%).
 Younger infant \#14 had the least portion of the waking day accounted for; the family left the house to run errands during the majority of the recording period, and the infant napped during much of the remaining time at home.
@@
 
-In the last analysis, we demonstrate how the full-day recordings can reveal individual differences in body position according to age, replicating age differences revealed in past work [@Survey].
+In order to show the convergent validity of the classification method, we demonstrate how the full-day recordings can reveal individual differences in body position according to age, replicating age differences revealed in past work [@Survey].
 Table \@ref(tab:agetable) summarizes the percent of time that younger and older infants spent in each body position out of their awake samples, as predicted by group and individual models.
@@
 To put this in context, the reliability of human coders on the body position code was ~90%, putting model-predicted accuracy on a similar level to human coders.
-Acceptable inter-rater reliability in developmental psychology studies of motor behavior is greater than 80%; both group and individual model results exceed that standard for most recorded sessions.
 Another comparison we can make is to the accuracy of the LENA&reg; device's automated metrics.
@@
 More work is needed to know how well group models would predict a different set of infants' behavior in a different environmental and/or cultural context.
-<!-- Reliability standard? -->
 
@@
 Fortunately, we could reference video recordings to understand why the errors occurred.
-Idiosyncrasies in activities and device use make it challenging to decide *a priori* whether odd looking classifications are possible.
+Idiosyncrasies in activities and device use make it challenging to decide *a priori* whether "odd" classifications (from the expectation of the researcher) are possible.
 For example, younger infant \#8 on Figure \@ref(fig:timelines) spent an unusual amount of time upright at an age where infants cannot yet stand.
@@
 Likewise, individual models can be an excellent choice for rapid prototyping and pilot testing.
-A researcher can get proof-of-concept data from a single participant or a new kind of classification scheme (i.e., locomotion, restrain in device) without annotating an entire sample.
+A researcher can get proof-of-concept data from a single participant or a new kind of classification scheme (i.e., locomotion, restraint in device) without annotating an entire sample.
 Individual models may also be applied in clinical cases where infants vary widely in their motor abilities and/or use of assistive devices.
@@
 Beyond group and individual models, there were many degrees of freedom in our choice of how best to model body position.
-For the sake of brevity, we report the best approach, not every possible variation in modeling.
+For the sake of brevity, we chose to report the best approach, not every possible variation in modeling.
 However, our openly shared data and code allow researcher to experiment with different sets of motion features, different sets of training data, different machine learning algorithms, and different hyperparameter values.
-As in past work [@AiraksinenRasanen2020], performance is best when using all four sensors and degrades when using any subset of two (i.e., hips only, left leg only).
-We tested but did not report models that used either only accelerometer data and only gyroscope data; in both cases, performance is degraded compared with models that use both types of motion.
+As in past work [@AiraksinenRasanen2020], performance was best when using all four sensors and degraded when using any subset of two (i.e., hips only, left leg only).
+We tested but did not report models that used either only accelerometer data and only gyroscope data; in both cases, performance was degraded compared with models that use both types of motion.
 We also tried different machine learning algorithms (XGBoost, Tabnet), but found that random forest models performed the best.
 Finally, we tried modeling each age group separately, but found no benefit to performance by tailoring to the particular age group.
-<!-- DATA CITE -->
 
@@
 Collecting dense data over the entire variety of daily experiences helps to more accurately measure infants' experiences in aggregate (as in Figure \@ref(fig:age)) without biasing results from a particular type of activity (e.g., play).
-Long-form recordings also have the potential to measure clinically-relevant outcomes, such as whether a physical therapy intervention to improve infant sitting skill leads to increases in everyday sitting time.
+Long-form recordings also have the potential to measure clinically-relevant outcomes for infants with motor delay or other pediatric concerns.
 
@@
 For example, recordings from wearable head cameras in the home indicate that infants see a small subset of objects with high frequency [@ClerkinHart2017], but most objects are seen infrequently.
-Heterogeneity is also found in how experiences are distributed in time---burstiness and clustering are seen in infants' daily experiences seeing objects and hearing music.
-Long-form motor recordings provide a novel opportunity to measure the temporal structure of motor experiences that can be seen in Figure \@ref(fig:timelines).
+Heterogeneity is also found in how experiences are distributed in time---burstiness and clustering are seen in infants' daily experiences seeing objects and hearing music [@MendozaFausey2022; @CasillasElliott2021].
+Long-form motor recordings provide a novel opportunity to measure the temporal structure in the sequences of body positions seen in Figure \@ref(fig:timelines).
 How skewed are infants' experiences with different motor behaviors, and how are they clustered in time?
 Linking the distributional structure of a particular motor experience in a real day, such as time spent sitting, might be predicted by sitting skill and/or might predict infants' future sitting ability.
-<!-- CITES FOR BURSTINESS CASILLAS FAUSEY -->
 
@@
 Annotating a larger, more representative sample of *hundreds of infants* while simultaneously scoring *full-day data* would be prohibitive.
-With sufficiently accurate group models, researchers could annotate a moderate video sample to then apply the model to hundreds of full-day recordings.
-The added cost of more recording time is low, meaning that future studies could sample across multiple days of behavior to better understand intra- versus inter-day variability.
+With sufficiently accurate group models, researchers could annotate a moderately-sized video corpus to then apply the model to hundreds of full-day recordings.
+Unlike with video, the added cost of more IMU sensor recording time is low, meaning that future studies could sample across multiple days of behavior to better understand intra- versus inter-day variability.
 Prior diary methods show that infants inconsistently display new motor skills on a day-to-day basis [@AdolphRobinson2008; @AdolphRobinson2011].
 Multi-day recordings could further uncover how infants' experiences vary over timescales from seconds to days.
-
 <!-- More easily shareable data (non-identifiable, text based) -->
@@
 As we described in the procedure section, synchronizing video cameras and inertial sensors required capturing a synchronization point on video.
-We completed this procedure prior to giving equipment to caregivers so that they were not responsible for synchronization; in fact, caregivers never touched a button on the video cameras.
-An end-of-day video would require a lot on the part of caregivers, and any mistakes in the procedure would lead to misaligned and unusable data.
+We completed this procedure prior to giving equipment to caregivers so that they were not responsible for synchronization; in fact, caregivers never (purposefully) touched a button on the video camera.
+An end-of-day video would require a lot of effort and compliance on the part of caregivers, and any mistakes in the procedure would lead to misaligned and unusable data.
 Given that most recordings ended at infants' bedtimes, having an experimenter visit families in the evening to video record would be intrusive.
@@
 In most cases, model prediction accuracy approached human reliability, suggesting that model predictions can be confidently used in analyses of full-day infant behavior.
-Examining the resulting corpus of `r apa_num(sum(coverage$hours_recorded))` hours of real-time infant body position showed the feasibility of capturing data covering the majority of infants' awake time and the variety of activities contained therein.
+Examining the resulting corpus of > 200 hours of real-time infant body position showed the feasibility of capturing data covering the majority of infants' awake time and the variety of activities contained therein.
 A rudimentary analysis of aggregated data showed that full-day position estimates conformed to expectations about age differences in motor behavior.

### 2023-03-27 manuscript.tex
revised whole paper

@@
 The chosen participant's classification accuracy was at the median level.
-Because of the overall size of the full data set and computational power/time required to synchronize and create windowed datasets, it would not be feasible to create a reproducible environment to process the entire dataset.
-However, in a second CodeOcean capsule (link) we share the entire dataset of windowed motion features with corresponding body position codes used to validate the method (described in the results below), along with the processed datafiles used for each set of analyses.
-This reproducible manuscript created in \emph{papaja} (Aust \& Barth, 2022) can be executed in the CodeOcean environment for full computational transparency.
+Because of the overall size of the full dataset and the computational power/time required to synchronize and create windowed datasets for each session, it would not be feasible to create a reproducible environment to process all 34 sessions.
+However, in a second CodeOcean capsule (LINK) we share the entire result: The dataset of windowed motion features with corresponding body position codes used to validate the method (described in the results below), along with the processed intermediary files used for each set of analyses.
+This reproducible manuscript created in \emph{papaja} (Aust \& Barth, 2022) can be executed in the CodeOcean environment for full computational transparency, and can be downloaded from Github (LINK).
 
@@
 As in past work (Franchak et al., 2021), we compared two types of models: \emph{group models} and \emph{individual models}.
-To assess the accuracy of body position classification for each recording session, we reserved the last 25\% of a session's proximal comparison data as the testing set.
+To assess the accuracy of body position classification for each recording session, we reserved the (temporally) last 25\% of a session's proximal comparison data as the testing set.
 The testing set was never used as training data.
@@
 Although overall accuracy from our semi-supervised, in-home data collection did not match the near-perfect accuracy (.95-.98) found in prior in-lab studies (Airaksinen et al., 2020; Franchak et al., 2021), both models approached the level of agreement found between two human coders (\emph{M} = .906).
-Most likely, lower accuracy in the current study results from the more variable and complex behavior observed in an unsupervised setting rather than from a difference in the quality of the classification model.
-Visual inspection of Figure \ref{fig:metrics}A shows that accuracy values were heavily skewed, with most approaching perfect accuracy but a few participants with very poor accuracy.
-Looking at the median performance suggests that the difference between models was not as wide for the typical participant (group median accuracy = 0.89; individual median accuracy = 0.93).
-Indeed, it is notable that the worst-case accuracy from group models (minimum = 0.50) was considerably lower than in individual models (minimum = 0.73).
+Most likely, lower accuracy in the current study results from the more variable and complex behavior observed in a a semi-supervised setting rather than from a difference in the quality of the classification model; although the first part of the proximal period was guided by the experimenter during a brief phone call, the remainder of the proximal period included natural behavior.
+Visual inspection of Figure \ref{fig:metrics}A shows that accuracy values were heavily skewed, with many approaching perfect accuracy but a few sessions with very poor accuracy.
+Looking at the median performance suggests that the difference between models was not considerable for the typical participant (group median accuracy = 0.89; individual median accuracy = 0.93).
+Indeed, it is notable that the worst-case accuracy from group models (minimum = 0.50) was far lower than in individual models (minimum = 0.73).
 Possibly, individual models accounted for idiosyncrasies in behavior or inconsistencies in sensor placement for those sessions that group models could not generalize to.
@@
 
-Strong overall accuracy can be misleading when class prevalence is unbalanced.
+Strong overall accuracy can be misleading when the relative frequency of different classes is unbalanced.
 For example, most infants spent more time sitting than prone, so good performance for sitting classification could mask poorer performance in prone classification.
-Accordingly, we report Cohen's kappa, a commonly-used metric that penalizes missing rare events (Figure \ref{fig:metrics}B), and provide classification metrics for each individual body position (Table \ref{tab:metrics-by-class}) to account for imbalance in body position rates within and between individuals.
+Accordingly, we report Cohen's kappa, a commonly-used metric that penalizes missing rare events (Figure \ref{fig:metrics}B), and we provide classification metrics for each individual body position (Table \ref{tab:metrics-by-class}) to account for imbalance in body position rates within and between individuals.
 Similar to overall accuracy, kappa values were strong for both model types with group kappas (\emph{M} = 0.75) somewhat worse compared with individual kappas (\emph{M} = 0.82).
@@
 Figure \ref{fig:part2overall} and Table \ref{tab:pt2overalltable} summarize the overall agreement during the distal comparison period.
-For each session, we calculated the actual prevalence of each body position as a percentage of the time that the infant was on video and awake using human annotated body position (x-axis on Figure \ref{fig:part2overall}).
-Predicted prevalence was calculated the same way for group and individual model predictions, omitting the off-camera and nap periods to make a direct comparison.
-Overall---across participants and across body position classes---agreement was strong: The correlation between group model predictions and human-coded prevalence was \emph{r} = 0.80, and the correlation between individual model predictions and human-coded prevalence was \emph{r} = 0.91.
+For each session, we calculated the actual time spent in each body position (out of times the infant was visible on camera and awake) using human annotated body position (x-axis on Figure \ref{fig:part2overall}).
+Predicted time was calculated the same way for group and individual model predictions, omitting the off-camera and nap periods to make a direct comparison.
+Overall---across participants and across body position classes---agreement was strong: The correlation between group model predictions and human-coded time was \emph{r} = 0.80, and the correlation between individual model predictions and human-coded time was \emph{r} = 0.91.
 As in the proximal comparison, agreement varied somewhat between body positions; in particular, agreement for held was poor.
@@
 This participant's proximal accuracy was also poor because similar ambiguities appeared during the initial recording.
-In contrast, the ``gray diamond outlier'' had strong proximal accuracy, with confusion only arising in the distal period between upright and held categories.
+In contrast, the ``gray diamond'' outlier had strong proximal accuracy, with confusion only arising in the distal period between upright and held categories.
 Reviewing the video showed that all disagreements occurred when the infant was in a baby walker; human coders scored this as ``upright'' but the models predicted it as ``held''.
@@
 Although overall aggregate agreement in the distal comparison was strong, it is important to show that similarly strong agreement is found within a shorter timescale.
-We repeated the agreement analysis (correlating human-coded body position prevalence with model-predicted prevalence) after dividing the distal comparison period into nine 10-minute bins (marked by vertical dashed lines in Figure \ref{fig:exemplar-timeline}).
+We repeated the agreement analysis (correlating human-coded body position time with model-predicted time) after dividing the distal comparison period into nine 10-minute bins (marked by vertical dashed lines in Figure \ref{fig:exemplar-timeline}).
 Infants had varying numbers of 10-minute bins depending on how much time they were awake and on camera.
@@
 Nap times reported during the recording period ranged from 0.00-4.99 hours with a median of 2.25 hours.
-Body position data was available to describe 37.3-100.0\% of the awake portion of the recording period (median = 100.0\%).
+Body position data were available to describe 37.3-100.0\% of the awake portion of the recording period (median = 100.0\%).
 Younger infant \#14 had the least portion of the waking day accounted for; the family left the house to run errands during the majority of the recording period, and the infant napped during much of the remaining time at home.
@@
 
-In the last analysis, we demonstrate how the full-day recordings can reveal individual differences in body position according to age, replicating age differences revealed in past work (Franchak, 2019).
+In order to show the convergent validity of the classification method, we demonstrate how the full-day recordings can reveal individual differences in body position according to age, replicating age differences revealed in past work (Franchak, 2019).
 Table \ref{tab:agetable} summarizes the percent of time that younger and older infants spent in each body position out of their awake samples, as predicted by group and individual models.
@@
 To put this in context, the reliability of human coders on the body position code was \textasciitilde90\%, putting model-predicted accuracy on a similar level to human coders.
-Acceptable inter-rater reliability in developmental psychology studies of motor behavior is greater than 80\%; both group and individual model results exceed that standard for most recorded sessions.
 Another comparison we can make is to the accuracy of the LENA® device's automated metrics.
@@
 Fortunately, we could reference video recordings to understand why the errors occurred.
-Idiosyncrasies in activities and device use make it challenging to decide \emph{a priori} whether odd looking classifications are possible.
+Idiosyncrasies in activities and device use make it challenging to decide \emph{a priori} whether ``odd'' classifications (from the expectation of the researcher) are possible.
 For example, younger infant \#8 on Figure \ref{fig:timelines} spent an unusual amount of time upright at an age where infants cannot yet stand.
@@
 Beyond group and individual models, there were many degrees of freedom in our choice of how best to model body position.
-For the sake of brevity, we report the best approach, not every possible variation in modeling.
+For the sake of brevity, we chose to report the best approach, not every possible variation in modeling.
 However, our openly shared data and code allow researcher to experiment with different sets of motion features, different sets of training data, different machine learning algorithms, and different hyperparameter values.
-As in past work (Airaksinen et al., 2020), performance is best when using all four sensors and degrades when using any subset of two (i.e., hips only, left leg only).
-We tested but did not report models that used either only accelerometer data and only gyroscope data; in both cases, performance is degraded compared with models that use both types of motion.
+As in past work (Airaksinen et al., 2020), performance was best when using all four sensors and degraded when using any subset of two (i.e., hips only, left leg only).
+We tested but did not report models that used either only accelerometer data and only gyroscope data; in both cases, performance was degraded compared with models that use both types of motion.
 We also tried different machine learning algorithms (XGBoost, Tabnet), but found that random forest models performed the best.
@@
 Collecting dense data over the entire variety of daily experiences helps to more accurately measure infants' experiences in aggregate (as in Figure \ref{fig:age}) without biasing results from a particular type of activity (e.g., play).
-Long-form recordings also have the potential to measure clinically-relevant outcomes, such as whether a physical therapy intervention to improve infant sitting skill leads to increases in everyday sitting time.
+Long-form recordings also have the potential to measure clinically-relevant outcomes for infants with motor delay or other pediatric concerns.
 
@@
 For example, recordings from wearable head cameras in the home indicate that infants see a small subset of objects with high frequency (Clerkin et al., 2017), but most objects are seen infrequently.
-Heterogeneity is also found in how experiences are distributed in time---burstiness and clustering are seen in infants' daily experiences seeing objects and hearing music.
-Long-form motor recordings provide a novel opportunity to measure the temporal structure of motor experiences that can be seen in Figure \ref{fig:timelines}.
+Heterogeneity is also found in how experiences are distributed in time---burstiness and clustering are seen in infants' daily experiences seeing objects and hearing music (Casillas \& Elliott, 2021; Mendoza \& Fausey, 2022).
+Long-form motor recordings provide a novel opportunity to measure the temporal structure in the sequences of body positions seen in Figure \ref{fig:timelines}.
 How skewed are infants' experiences with different motor behaviors, and how are they clustered in time?
@@
 \emph{International Journal of Behavioral Nutrition and Physical Activity}, \emph{17}(1), 14.
+
+\leavevmode\vadjust pre{\hypertarget{ref-CasillasElliott2021}{}}%
+Casillas, M., \& Elliott, M. (2021). \emph{Cross-cultural differences in children's object handling at home}.
 

### 2023-03-27 manuscript.Rmd
abstract

@@
 abstract: |
-  Abstract
+  Long form audio recordings have had a transformational effect on the study of infant language acquisition by using mobile, unobtrusive devices to gather full-day, real-time data that can be automatically scored.
+How can we produce similar data in service of measuring infants' everyday motor behaviors, such as body position?
+The aim of the current study was to validate long form recordings of infant position (supine, prone, sitting, upright, held by caregiver) based on machine learning classification of data from inertial sensors worn on infants' ankles and hips.
+Using over 100 hours of video recordings synchronized with inertial sensor data from infants in their homes, we demonstrate that body position classifications are sufficiently accurate to measure infant behavior.
+Moreover, classification remained accurate when predicting behavior later in the session when infants and caregivers were unsupervised and went about their normal activities, showing that the method can handle the challenge of measuring unconstrained, natural activity.
+Next, we show that the inertial sensing method has convergent validity by replicating age differences in body position found using other methods with full-day data captured from inertial sensors.
+We end the paper with a discussion of the novel opportunities that long form motor recordings afford for understanding infant learning and development.
 
 keywords          : "body position, motor development, everyday experiences, sitting, machine learning"
-wordcount         : "X"
+wordcount         : "12,030"
 

### 2023-03-27 manuscript.tex
abstract

@@
 {}{\PatchFailed}
-\keywords{body position, motor development, everyday experiences, sitting, machine learning\newline\indent Word count: X}
+\keywords{body position, motor development, everyday experiences, sitting, machine learning\newline\indent Word count: 12,030}
 \DeclareDelayedFloatFlavor{ThreePartTable}{table}
@@
 \abstract{%
-Abstract
+Long form audio recordings have had a transformational effect on the study of infant language acquisition by using mobile, unobtrusive devices to gather full-day, real-time data that can be automatically scored.
+How can we produce similar data in service of measuring infants' everyday motor behaviors, such as body position?
+The aim of the current study was to validate long form recordings of infant position (supine, prone, sitting, upright, held by caregiver) based on machine learning classification of data from inertial sensors worn on infants' ankles and hips.
+Using over 100 hours of video recordings synchronized with inertial sensor data from infants in their homes, we demonstrate that body position classifications are sufficiently accurate to measure infant behavior.
+Moreover, classification remained accurate when predicting behavior later in the session when infants and caregivers were unsupervised and went about their normal activities, showing that the method can handle the challenge of measuring unconstrained, natural activity.
+Next, we show that the inertial sensing method has convergent validity by replicating age differences in body position found using other methods with full-day data captured from inertial sensors.
+We end the paper with a discussion of the novel opportunities that long form motor recordings afford for understanding infant learning and development.
 }
@@
 Annotating a larger, more representative sample of \emph{hundreds of infants} while simultaneously scoring \emph{full-day data} would be prohibitive.
-With sufficiently accurate group models, researchers could annotate a moderate video sample to then apply the model to hundreds of full-day recordings.
-The added cost of more recording time is low, meaning that future studies could sample across multiple days of behavior to better understand intra- versus inter-day variability.
+With sufficiently accurate group models, researchers could annotate a moderately-sized video corpus to then apply the model to hundreds of full-day recordings.
+Unlike with video, the added cost of more IMU sensor recording time is low, meaning that future studies could sample across multiple days of behavior to better understand intra- versus inter-day variability.
 Prior diary methods show that infants inconsistently display new motor skills on a day-to-day basis (Adolph et al., 2008; Adolph \& Robinson, 2011).
@@
 In most cases, model prediction accuracy approached human reliability, suggesting that model predictions can be confidently used in analyses of full-day infant behavior.
-Examining the resulting corpus of 206.66 hours of real-time infant body position showed the feasibility of capturing data covering the majority of infants' awake time and the variety of activities contained therein.
+Examining the resulting corpus of \textgreater{} 200 hours of real-time infant body position showed the feasibility of capturing data covering the majority of infants' awake time and the variety of activities contained therein.
 A rudimentary analysis of aggregated data showed that full-day position estimates conformed to expectations about age differences in motor behavior.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-CasillasElliott2021}{}}%
-Casillas, M., \& Elliott, M. (2021). \emph{Cross-cultural differences in children's object handling at home}.
+Casillas, M., \& Elliott, M. (2021). Cross-cultural differences in children's object handling at home. \emph{PsyArXiv}.
 
@@
 \leavevmode\vadjust pre{\hypertarget{ref-DanchoVaughan2023}{}}%
-Dancho, M., \& Vaughan, D. (2023). \emph{Timetk: A tool kit for working with time series in r}.
+Dancho, M., \& Vaughan, D. (2023). \emph{Timetk: A tool kit for working with time series in {R}}.
 

### 2023-03-28 manuscript.Rmd
updated ignore for LFS

@@
 A Databrary repository (https://nyu.databrary.org/volume/1580) includes an exemplar participant's recording session, with the raw video data files, the Datavyu annotations of those video files, a log file with machine-readable synchronization points and nap/diaper change times, and accelerometer and gyroscope data for each of the 4 sensors.
-A CodeOcean capsule (LINK) hosts an R environment with the exemplar participant's data and source code to: 1) synchronize IMU and video annotations, 2) calculate windowed motion features for their data, and 3) train and test the body position classifier using an "individual model".
+A GitHub repository (LINK) contains the exemplar participant's data and source code to: 1) synchronize IMU and video annotations, 2) calculate windowed motion features for their data, and 3) train and test the body position classifier using an "individual model".
 The chosen participant's classification accuracy was at the median level.
-Because of the overall size of the full dataset and the computational power/time required to synchronize and create windowed datasets for each session, it would not be feasible to create a reproducible environment to process all 34 sessions.
-However, in a second CodeOcean capsule (LINK) we share the entire result: The dataset of windowed motion features with corresponding body position codes used to validate the method (described in the results below), along with the processed intermediary files used for each set of analyses.
-This reproducible manuscript created in *papaja* [@papaja] can be executed in the CodeOcean environment for full computational transparency, and can be downloaded from Github (LINK).
+Because of the overall size of the full dataset and the computational power/time required to synchronize and create windowed datasets for each session, it would not be feasible to reproduce the calculations for all 34 sessions.
+However, in a second Github repository (LINK) we share the full results of those computations: The dataset of windowed motion features with corresponding body position codes used to validate the method (described in the results below), along with the processed intermediary files used for each set of analyses.
+This reproducible manuscript created in RMarkdown and *papaja* [@papaja] is generated from those data files for computational transparency.
 <!-- FINAL LINKS -->

### 2023-03-28 group_split_comparison/group_split.R
Added comments to group_split

@@ -1,13 +1,16 @@
+# This script calculates the first set of results in the paper based on the proximal period
+# compiled_data_lite.RData contains the windowed motion feature data for all sessions
+# The data are partitioned into training and testing sets
+# Group models and individual ("split") models are trained, and then predictions are made on the testing set
+# Agreement metrics are saved to /data to be reported in the RMarkdown manuscript
+# Depending on hardware, this script can take several hours to run
+
 library(here)
 library(randomForest)
 library(cvms)
 library(caret)
 library(tidyverse)
 library(rstatix)
-# library(furrr)
-# 
-# plan("multisession", workers = 4)
-# plan("sequential")
 
 #LOAD DATA
 load( here("group_split_comparison","compiled_data_lite.RData"))
@@ -21,12 +24,10 @@ total_samples <- slide_filt %>% drop_na(code) %>% group_by(id) %>% summarize(n_s
 training <- slide_filt %>% group_by(id, code) %>% slice_head(prop = .75) %>% ungroup 
 testing <- slide_filt %>% group_by(id, code) %>% slice_tail(prop = .25) %>% ungroup
 
-bad_ids <- c(9909, 10204, 12502, 9911, 10701, 10802, 12002, 9910, 11104, 12101, 11501, 11601)
-
-group_model <- function(temp_id, training, ntree = 50, bad_ids = NULL) {
+group_model <- function(temp_id, training, ntree = 50) {
   print(temp_id)
   temp_group_training <- training %>% 
-    filter(id != temp_id, !(id %in% bad_ids)) %>% 
+    filter(id != temp_id) %>% 
     select_if(not_all_na) %>% 
     mutate(code = fct_drop(code)) %>% 
     select(-id)
@@ -46,11 +47,7 @@ split_model <- function(temp_id, training, ntree = 50) {
 
 ids <- unique(training$id)
 group_mods <-map(ids, ~group_model(.x, training)) %>% set_names(ids)
-# group_select_mods <- map(ids, ~group_model(.x, training, bad_ids = bad_ids)) %>% set_names(ids)
 split_mods <- map(ids, ~split_model(.x, training)) %>% set_names(ids)
-
-# save(group_mods, split_mods, file = "group_split_comparison/models.RData")
[…]

### 2023-03-28 manuscript.Rmd
session info

@@
 keywords          : "body position, motor development, everyday experiences, sitting, machine learning"
-wordcount         : "12,030"
+wordcount         : "12,022"
 
@@
 ```{r analysis-preferences}
-# Seed for random number generation
 set.seed(42)
@@
 A Databrary repository contains the entire video recording, coding files, and raw IMU data for a single participant (https://nyu.databrary.org/volume/1580).
-<!-- CODE OCEAN HERE -->
 
@@
 A single, merged dataset was created with synchronized accelerometer signals (in three orientations: X, Y, and Z) and gyroscope signals (in three orientations: roll, pitch, and yaw) for each of the four sensors (left hip, right hip, left ankle, right ankle) with the corresponding timestamp and body position code using the *timetk* package [@DanchoVaughan2023] and the *lubridate* package [@GrolemundWickham2011] in R version 4.1.2 [@R42].
-<!-- CODE OCEAN HERE -->
 
@@
 A Databrary repository (https://nyu.databrary.org/volume/1580) includes an exemplar participant's recording session, with the raw video data files, the Datavyu annotations of those video files, a log file with machine-readable synchronization points and nap/diaper change times, and accelerometer and gyroscope data for each of the 4 sensors.
-A GitHub repository (LINK) contains the exemplar participant's data and source code to: 1) synchronize IMU and video annotations, 2) calculate windowed motion features for their data, and 3) train and test the body position classifier using an "individual model".
+A GitHub repository (https://github.com/JohnFranchak/body_position_classification_ms.git) contains the exemplar participant's data and source code to: 1) synchronize IMU and video annotations, 2) calculate windowed motion features for their data, and 3) train and test the body position classifier using an "individual model".
 The chosen participant's classification accuracy was at the median level.
 Because of the overall size of the full dataset and the computational power/time required to synchronize and create windowed datasets for each session, it would not be feasible to reproduce the calculations for all 34 sessions.
-However, in a second Github repository (LINK) we share the full results of those computations: The dataset of windowed motion features with corresponding body position codes used to validate the method (described in the results below), along with the processed intermediary files used for each set of analyses.
-This reproducible manuscript created in RMarkdown and *papaja* [@papaja] is generated from those data files for computational transparency.
-<!-- FINAL LINKS -->
+However, in a second Github repository (https://github.com/JohnFranchak/body_position_classification_ms.git) we share the full results of those computations: The dataset of windowed motion features with corresponding body position codes used to validate the method (described in the results below), along with the processed intermediary files used for each set of analyses.
+This reproducible manuscript created in RMarkdown and *papaja* [@papaja] can be regenerated from those data files for full computational transparency.
 
@@
 
-# outlier_ids <- c(10701, 10401, 10601)
-# fullday %>% cor_mat(age, sit_time:upright_time) %>% cor_mark_significant()
-# fullday %>% filter(!(id_uni %in% outlier_ids)) %>% cor_mat(age, sit_time:upright_time) %>% cor_mark_significant()
-# 
-# fullday_long %>% group_by(position, group) %>% get_summary_stats(prop)
-
 ```
@@
 Multi-day recordings could further uncover how infants' experiences vary over timescales from seconds to days.
-<!-- More easily shareable data (non-identifiable, text based) -->
 
@@
 Differences in infants' positioning while asleep (held in caregivers arms, laying in cribs, reclined in strollers and carseats) might make classifying sleep more difficult compared with adults based on movement.
-<!-- NEED SLEEP AND WEAR TIME CITES> -->
-<!-- Algorithms for infant sleep? -->
 
@@
 
+```{r results="hide"}
+devtools::session_info() %>% jsonlite::write_json("session_info.json")
+```
+
 \newpage

### 2023-03-28 manuscript.tex
session info

@@
 {}{\PatchFailed}
-\keywords{body position, motor development, everyday experiences, sitting, machine learning\newline\indent Word count: 12,030}
+\keywords{body position, motor development, everyday experiences, sitting, machine learning\newline\indent Word count: 12,022}
 \DeclareDelayedFloatFlavor{ThreePartTable}{table}
@@
 A Databrary repository (\url{https://nyu.databrary.org/volume/1580}) includes an exemplar participant's recording session, with the raw video data files, the Datavyu annotations of those video files, a log file with machine-readable synchronization points and nap/diaper change times, and accelerometer and gyroscope data for each of the 4 sensors.
-A CodeOcean capsule (LINK) hosts an R environment with the exemplar participant's data and source code to: 1) synchronize IMU and video annotations, 2) calculate windowed motion features for their data, and 3) train and test the body position classifier using an ``individual model''.
+A GitHub repository (\url{https://github.com/JohnFranchak/body_position_classification_ms.git}) contains the exemplar participant's data and source code to: 1) synchronize IMU and video annotations, 2) calculate windowed motion features for their data, and 3) train and test the body position classifier using an ``individual model''.
 The chosen participant's classification accuracy was at the median level.
-Because of the overall size of the full dataset and the computational power/time required to synchronize and create windowed datasets for each session, it would not be feasible to create a reproducible environment to process all 34 sessions.
-However, in a second CodeOcean capsule (LINK) we share the entire result: The dataset of windowed motion features with corresponding body position codes used to validate the method (described in the results below), along with the processed intermediary files used for each set of analyses.
-This reproducible manuscript created in \emph{papaja} (Aust \& Barth, 2022) can be executed in the CodeOcean environment for full computational transparency, and can be downloaded from Github (LINK).
+Because of the overall size of the full dataset and the computational power/time required to synchronize and create windowed datasets for each session, it would not be feasible to reproduce the calculations for all 34 sessions.
+However, in a second Github repository (\url{https://github.com/JohnFranchak/body_position_classification_ms.git}) we share the full results of those computations: The dataset of windowed motion features with corresponding body position codes used to validate the method (described in the results below), along with the processed intermediary files used for each set of analyses.
+This reproducible manuscript created in RMarkdown and \emph{papaja} (Aust \& Barth, 2022) can be regenerated from those data files for full computational transparency.
 

### 2023-03-28 manuscript.Rmd
updated author note

@@
 
-  This work was funded by National Science Foundation Grant BCS #1941449 to the first author.
+  This work was funded by National Science Foundation Grant BCS #1941449 to the first author. The authors declare no conflict of interest.
 
@@
 keywords          : "body position, motor development, everyday experiences, sitting, machine learning"
-wordcount         : "12,022"
+wordcount         : "11,995"
 
@@
 A Databrary repository (https://nyu.databrary.org/volume/1580) includes an exemplar participant's recording session, with the raw video data files, the Datavyu annotations of those video files, a log file with machine-readable synchronization points and nap/diaper change times, and accelerometer and gyroscope data for each of the 4 sensors.
-A GitHub repository (https://github.com/JohnFranchak/body_position_classification_ms.git) contains the exemplar participant's data and source code to: 1) synchronize IMU and video annotations, 2) calculate windowed motion features for their data, and 3) train and test the body position classifier using an "individual model".
-The chosen participant's classification accuracy was at the median level.
+A GitHub repository (https://github.com/JohnFranchak/body_position_classification_example) contains the exemplar participant's data and source code to: 1) synchronize IMU and video annotations, 2) calculate windowed motion features for their data, and 3) train and test the body position classifier using an "individual model".
 Because of the overall size of the full dataset and the computational power/time required to synchronize and create windowed datasets for each session, it would not be feasible to reproduce the calculations for all 34 sessions.
-However, in a second Github repository (https://github.com/JohnFranchak/body_position_classification_ms.git) we share the full results of those computations: The dataset of windowed motion features with corresponding body position codes used to validate the method (described in the results below), along with the processed intermediary files used for each set of analyses.
+However, in a second Github repository (https://github.com/JohnFranchak/body_position_classification_ms) we share the full results of those computations: The dataset of windowed motion features with corresponding body position codes used to validate the method.
 This reproducible manuscript created in RMarkdown and *papaja* [@papaja] can be regenerated from those data files for full computational transparency.
@@
 ```{r results="hide"}
-devtools::session_info() %>% jsonlite::write_json("session_info.json")
+writeLines(capture.output(devtools::session_info()), "session_info.txt")
 ```

### 2023-03-28 manuscript.tex
updated author note

@@
 {}{\PatchFailed}
-\keywords{body position, motor development, everyday experiences, sitting, machine learning\newline\indent Word count: 12,022}
+\keywords{body position, motor development, everyday experiences, sitting, machine learning\newline\indent Word count: 11,995}
 \DeclareDelayedFloatFlavor{ThreePartTable}{table}
@@
 
-This work was funded by National Science Foundation Grant BCS \#1941449 to the first author.
+This work was funded by National Science Foundation Grant BCS \#1941449 to the first author. The authors declare no conflict of interest.
 
@@
 A Databrary repository (\url{https://nyu.databrary.org/volume/1580}) includes an exemplar participant's recording session, with the raw video data files, the Datavyu annotations of those video files, a log file with machine-readable synchronization points and nap/diaper change times, and accelerometer and gyroscope data for each of the 4 sensors.
-A GitHub repository (\url{https://github.com/JohnFranchak/body_position_classification_ms.git}) contains the exemplar participant's data and source code to: 1) synchronize IMU and video annotations, 2) calculate windowed motion features for their data, and 3) train and test the body position classifier using an ``individual model''.
-The chosen participant's classification accuracy was at the median level.
+A GitHub repository (\url{https://github.com/JohnFranchak/body_position_classification_example}) contains the exemplar participant's data and source code to: 1) synchronize IMU and video annotations, 2) calculate windowed motion features for their data, and 3) train and test the body position classifier using an ``individual model''.
 Because of the overall size of the full dataset and the computational power/time required to synchronize and create windowed datasets for each session, it would not be feasible to reproduce the calculations for all 34 sessions.
-However, in a second Github repository (\url{https://github.com/JohnFranchak/body_position_classification_ms.git}) we share the full results of those computations: The dataset of windowed motion features with corresponding body position codes used to validate the method (described in the results below), along with the processed intermediary files used for each set of analyses.
+However, in a second Github repository (\url{https://github.com/JohnFranchak/body_position_classification_ms}) we share the full results of those computations: The dataset of windowed motion features with corresponding body position codes used to validate the method.
 This reproducible manuscript created in RMarkdown and \emph{papaja} (Aust \& Barth, 2022) can be regenerated from those data files for full computational transparency.

### 2023-04-24 manuscript.Rmd
updated for BRM

@@
 
+# Declarations
+
+## Funding
+
+This work was funded by National Science Foundation Grant BCS #1941449.
+
+## Competing Interests
+
+The authors have no competing interests to declare that are relevant to the content of this article.
+
+## Ethics Approval
+
+The study was performed in accordance with the ethical standards as laid down in the 1964 Declaration of Helsinki.
+The study procedures were approved by the Institutional Review Board of the University of California, Riverside, Protocol HS-15-050.
+
+## Consent
+
+Consent to participate: All caregivers provided written informed consent prior to the start of the study. 
+Consent to publish: Additional written consent was obtained by caregivers for data sharing of audio and video data.
+
+## Data, Material and Code availability
+
+A Databrary repository (https://nyu.databrary.org/volume/1580) includes an exemplar participant's recording session, with the raw video data files, the Datavyu annotations of those video files, a log file with machine-readable synchronization points and nap/diaper change times, and accelerometer and gyroscope data for each of the 4 sensors.
+A GitHub repository (https://github.com/JohnFranchak/body_position_classification_example) contains the exemplar participant's data and source code to: 1) synchronize IMU and video annotations, 2) calculate windowed motion features for their data, and 3) train and test the body position classifier using an "individual model".
+Because of the overall size of the full dataset and the computational power/time required to synchronize and create windowed datasets for each session, it would not be feasible to reproduce the calculations for all 34 sessions.
+However, in a second Github repository (https://github.com/JohnFranchak/body_position_classification_ms) we share the full results of those computations: The dataset of windowed motion features with corresponding body position codes used to validate the method.
+
+\newpage
+
 # References

### 2023-04-24 manuscript.tex
updated for BRM

@@
 
+\hypertarget{declarations}{%
+\section{Declarations}\label{declarations}}
+
+\hypertarget{funding}{%
+\subsection{Funding}\label{funding}}
+
+This work was funded by National Science Foundation Grant BCS \#1941449.
+
+\hypertarget{competing-interests}{%
+\subsection{Competing Interests}\label{competing-interests}}
+
+The authors have no competing interests to declare that are relevant to the content of this article.
+
+\hypertarget{ethics-approval}{%
+\subsection{Ethics Approval}\label{ethics-approval}}
+
+The study was performed in accordance with the ethical standards as laid down in the 1964 Declaration of Helsinki.
+The study procedures were approved by the Institutional Review Board of the University of California, Riverside, Protocol HS-15-050.
+
+\hypertarget{consent}{%
+\subsection{Consent}\label{consent}}
+
+Consent to participate: All caregivers provided written informed consent prior to the start of the study.
+Consent to publish: Additional written consent was obtained by caregivers for data sharing of audio and video data.
+
+\hypertarget{data-material-and-code-availability}{%
+\subsection{Data, Material and Code availability}\label{data-material-and-code-availability}}
+
+A Databrary repository (\url{https://nyu.databrary.org/volume/1580}) includes an exemplar participant's recording session, with the raw video data files, the Datavyu annotations of those video files, a log file with machine-readable synchronization points and nap/diaper change times, and accelerometer and gyroscope data for each of the 4 sensors.
+A GitHub repository (\url{https://github.com/JohnFranchak/body_position_classification_example}) contains the exemplar participant's data and source code to: 1) synchronize IMU and video annotations, 2) calculate windowed motion features for their data, and 3) train and test the body position classifier using an ``individual model''.
+Because of the overall size of the full dataset and the computational power/time required to synchronize and create windowed datasets for each session, it would not be feasible to reproduce the calculations for all 34 sessions.
+However, in a second Github repository (\url{https://github.com/JohnFranchak/body_position_classification_ms}) we share the full results of those computations: The dataset of windowed motion features with corresponding body position codes used to validate the method.
+
+\newpage
+
 \hypertarget{references}{%

### 2023-06-13 manuscript.Rmd
revisions

@@
 Synchronized video with ground-truth human annotations creates training data for a machine learning algorithm, such as a random forest model, that can be later used to predict categories from IMU data that was not annotated.
-Crucially, this allows automatic classification to scale to full day recording by relying on a relatively smaller set of video annotation.
+Crucially, this allows automatic classification to scale to full day recording by relying on a relatively smaller set of video annotations.
 
@@
 A pocket over the hip and a pocket just above the ankle were sewn on the lateral surface of the right and left legs of the garment.
-Each sensor recorded accelerometer and gyroscope data at 62.5 Hz throughout the day, with sufficient battery and onboard storage to record for approximately 12 hours.
+Each sensor recorded accelerometer and gyroscope data throughout the day, with sufficient battery and onboard storage to record for approximately 12 hours.
+We chose a sampling rate of 62.5 Hz (on of several device presets) based on prior work that used rates of 50-64 Hz [@FranchakScott2021; @AiraksinenRasanen2020; @YaoPlotz2019].
 Infants also wore a LENA&reg; recorder throughout the day in the front pocket of a LENA&reg; shirt, located near the infant’s chest.
 
-Video recordings were captured using an action camera on a miniature tripod (Insta360 ONE R) that caregivers placed in the room that the infant was in.
+Video recordings were captured at 30 Hz using an action camera on a miniature tripod (Insta360 ONE R) that caregivers placed in the room that the infant was in.
 The proprietary "Boosted Battery Base" attached to the action camera to allow for a total of 3 hours of recording.
@@
 
-Because the camera only had the battery life to record for ~3 hours (split into two 90-minute video files), this divided the day into different periods for analysis.
+Because the camera only had the battery life to record for ~3 hours (automatically split into two 90-minute video files), this divided the day into different periods for analysis.
 As seen in the bottom of Figure \@ref(fig:exemplar-timeline)), the *video period* comprised the first three hours of recording starting from the researcher's arrival when they turned on the camera.
@@
 This video recorded the next 90 minutes of natural activity.
-However, because there was a variable gap of ~40 s between the two videos, synchronization in the distal comparison video was coarser, with estimated temporal offsets of ~5 s in either direction.
+However, because of a limitation in the camera, there was a variable gap of ~40 s between the two videos, synchronization in the distal comparison video was coarser, with estimated temporal offsets of ~5 s in either direction.
 
@@
 Second, we applied models trained from data during the proximal comparison to predict infant behavior during the distal comparison, when coarser synchrony between video and motion data was available.
-This distal comparison is critical because it provides the first-ever test of how well infant body classification models predict behavior that is variable, unrestricted, and unsupervised.
+This distal comparison is critical because it provides, to our knowledge, the first-ever test of how well infant body classification models predict behavior that is variable, unrestricted, and unsupervised.
 Data from the first two sets of analyses confirm that body position classification models are accurate across a long period of time, suggesting that data from the entire day will be valid.
@@
 After the distal comparison video ended, caregivers were instructed to keep the sensors on their infants for the remainder of the day until infants went to bed, removing the sensors for naps, diaper changes, and trips out of the house.
+The algorithm was not designed to classify behavior during transportation (e.g., strollers, automobiles) so data collection was restricted to in-home behavior.
 The first two sets of results show that accuracy was consistently high across the proximal and distal recordings, providing confidence that predictions over the remainder of the day would continue to be accurate after the video recordings stopped.

### 2023-06-14 manuscript.Rmd
some revision edits

@@
 When sitting and upright, infants have a better view of faces and distant objects compared to their view while in a prone position [@Freeplay; @LuoFranchak2020; @CWW].
-Infants struggle to manipulate objects while supine and prone, but sitting affords object exploration [@SoskaAdolph2014].
+Learning to sit facilitates object exploration [@SoskaAdolph2014].
 Upright walking changes how infants and caregivers interact; walking infants move farther away from caregivers and share toys in different ways compared to infants crawling in a prone position [@ChenSchneider2022; @Karasik2011].
@@
 Despite promising past results in brief, supervised sessions [@FranchakScott2021; @AiraksinenRasanen2020; @AiraksinenGallen2022], the current investigation takes a needed step forward by testing accuracy over long, unsupervised recordings in what we term a *distal comparison*.
-Although in the current investigation we focus on classification of body position, this approach can be extended to categorize other aspects of movement, such as locomotion [@AiraksinenRasanen2020] and infant-caregiver contact [@YaoPlotz2019].
-We provide openly shared data, materials, and analysis code to facilitate other researchers' adopting this method and applying it to populations and behaviors (see Data Sharing and Transparency below).
 
@@
 Data are recorded to onboard device memory, so participants do not need to be in range of a receiver.
-Second, wearable audio recording is *unobtrusive*.
-Participants' reactivity to observation, such as from a video camera, may influence behavior.
+Second, wearable audio recording is more *unobtrusive*.
+Participants' reactivity to observation, such as from a video camera, may influence behavior more compared with a sensor that records only motion or audio data.
 For example, caregivers spoke more frequently to infants during a video-recorded portion of a home recording compared with audio-only segments captured by a LENA&reg; device [@Bergelson2019Input].
@@
 
-Third, although the goal was to collect fully naturalistic behavior, we opted to restrict recording to times when participants were in the home.
+Third, although the goal was to collect fully naturalistic behavior, our recording protocol led to some changes that might have affected (or missed) some behaviors.
+We opted to restrict recording to times when participants were in the home.
 Inertial sensors can travel with participants, and there is no doubt that infants could have worn the sensors out of the home on errands.
 However, we do not yet have training data to detect periods of transportation (e.g., moving in a car or in a stroller) that might add noise and/or lead to misclassification of body positions.
-Future work should aim to collect data in these scenarios so that body position can be captured in or out of the home.
+It is also unknown whether participants might have behaved differently had they not worn the sensors.
+The added bulk in the garment from the sensors (including the LENA&reg; recorder) might have made some positions more uncomfortable, such as lying prone with the LENA&reg; recorder worn on the chest.
 

### 2023-06-14 manuscript.Rmd
wear time revision

@@
 We chose a sampling rate of 62.5 Hz (on of several device presets) based on prior work that used rates of 50-64 Hz [@FranchakScott2021; @AiraksinenRasanen2020; @YaoPlotz2019].
-Infants also wore a LENA&reg; recorder throughout the day in the front pocket of a LENA&reg; shirt, located near the infant’s chest.
+Infants also wore a LENA&reg; recorder throughout the day in the front pocket of a LENA&reg; shirt, located near the infant’s chest, to determine whether data could be simultaneously recorded from the LENA&reg; and IMU sensors.
+However, LENA&reg; data were not analyzed as part of the current study.
 
@@
 
-Among participants who wore the sensors from morning to bedtime, the majority of the time during the day either resulted in usable body position data or was excluded due to caregiver-reported naps (gray periods in Figure \@ref(fig:timelines)).
+Among participants who wore the sensors from morning to bedtime, all infants wore the sensor garment during 100% of the intended recording period, excluding the times caregivers were asked to remove the garment (naps, diaper changes, trips out of the house), based on caregiver logs of wear time.
+No caregiver mentioned needing to remove the garment for any other reason (such as infant discomfort), meaning that the majority of the time during the day either resulted in usable body position data or was excluded due to caregiver-reported naps (gray periods in Figure \@ref(fig:timelines)).
 The total length of the recording period ranged from `r apa_num(min(coverage$total))`-`r apa_num(max(coverage$total))` hours with a median of `r apa_num(median(coverage$total))` hours.
@@
 Second, we relied on caregiver logs of infants' naps and times when sensors were removed to separate usable data (times when infants wore the sensors while awake) from unusable data.
-Although most caregivers were diligent about completing logs, there were a few cases that caregivers may have failed to report naps (younger infants \#1 and \#4).
+Future work should aim to further assess caregivers' perceptions about the usability and comfort of the garment.
+Although infants wore the sensor garment throughout all of the desired times, we did not collect independent data (aside from caregivers' logs) to verify wear time.
+Most caregivers were diligent about completing logs, however, there were a few cases that caregivers may have failed to report naps (younger infants \#1 and \#4).
 In the future, algorithms can be developed to automatically identify periods of sleep and times when sensors are off the body.
@@
 It is also unknown whether participants might have behaved differently had they not worn the sensors.
-The added bulk in the garment from the sensors (including the LENA&reg; recorder) might have made some positions more uncomfortable, such as lying prone with the LENA&reg; recorder worn on the chest.
+The added bulk in the garment from the sensors (including the LENA&reg; recorder) might have made some positions more uncomfortable, such as lying prone with the LENA&reg; recorder worn on the chest, and/or influenced how caregivers chose to position infants throughout the day.
 

### 2023-06-14 manuscript.Rmd
typos

@@
 For example, correlations between human transcribed counts of adult words and child vocalizations against LENA&reg;'s automatic counts were strong, *r* = .698 and *r* = .649, respectively [@CristiaLavechin2020].
-For other metrics, such as the number of "conversational turns" between the child and communicative partners and the identification of male speakers, agreement was poor.
+For other metrics, such as the number of "conversational turns" between the child and communicative partners, agreement was poor (*r* = .364).
 Thus, for some use cases (and for some metrics), long form audio recordings provide a mobile, unobtrusive way to automatically score real-time data over a full day.
@@
 Most participants were tested in a single session (*n* = `r nrow(num_sessions %>% filter(n == 1))`), but `r nrow(num_sessions %>% filter(n > 1))` participants contributed between 2-4 sessions as part of an ongoing longitudinal study.
+Although including more data from some participants could over-represent their characteristics in the model, we reasoned that this drawback was outweighed by having more available data to use for training.
 Only 1 session was excluded due to a technical error---one of the four IMU sensors failed to record, resulting in an unusable set of data for classification.
@@
 Across sessions, younger infants' age ranged from `r apa_num(age["Younger","min"], digits = 1)` to `r apa_num(age["Younger","max"], digits = 1)` months (*M* = `r apa_num(age["Younger","mean"], digits = 1)`) and older infants' age ranged from `r apa_num(age["Older","min"], digits = 1)` to `r apa_num(age["Older","max"], digits = 1)` months (*M* = `r apa_num(age["Older","mean"], digits = 1)`).
+The total number of recording sessions (`r length(unique(ppts$id_uni))`) exceeded the number of sessions employed in comparable past work: 10 in @NamPark2013, 15 in @FranchakScott2021, 22 in @AiraksinenRasanen2020, and 33 in @YaoPlotz2019.
 
@@
 Each sensor recorded accelerometer and gyroscope data throughout the day, with sufficient battery and onboard storage to record for approximately 12 hours.
-We chose a sampling rate of 62.5 Hz (on of several device presets) based on prior work that used rates of 50-64 Hz [@FranchakScott2021; @AiraksinenRasanen2020; @YaoPlotz2019].
+We chose a sampling rate of 62.5 Hz (one of the available preset rates) based on prior work that used rates of 50-64 Hz [@FranchakScott2021; @AiraksinenRasanen2020; @YaoPlotz2019].
 Infants also wore a LENA&reg; recorder throughout the day in the front pocket of a LENA&reg; shirt, located near the infant’s chest, to determine whether data could be simultaneously recorded from the LENA&reg; and IMU sensors.

### 2023-06-14 manuscript.tex
typos

@@
 When sitting and upright, infants have a better view of faces and distant objects compared to their view while in a prone position (Franchak et al., 2018; Kretch et al., 2014; Luo \& Franchak, 2020).
-Infants struggle to manipulate objects while supine and prone, but sitting affords object exploration (Soska \& Adolph, 2014).
+Learning to sit facilitates object exploration (Soska \& Adolph, 2014).
 Upright walking changes how infants and caregivers interact; walking infants move farther away from caregivers and share toys in different ways compared to infants crawling in a prone position (Chen et al., 2022; Karasik et al., 2011).
@@
 Despite promising past results in brief, supervised sessions (Airaksinen et al., 2022, 2020; Franchak et al., 2021), the current investigation takes a needed step forward by testing accuracy over long, unsupervised recordings in what we term a \emph{distal comparison}.
-Although in the current investigation we focus on classification of body position, this approach can be extended to categorize other aspects of movement, such as locomotion (Airaksinen et al., 2020) and infant-caregiver contact (Yao et al., 2019).
-We provide openly shared data, materials, and analysis code to facilitate other researchers' adopting this method and applying it to populations and behaviors (see Data Sharing and Transparency below).
 
@@
 Data are recorded to onboard device memory, so participants do not need to be in range of a receiver.
-Second, wearable audio recording is \emph{unobtrusive}.
-Participants' reactivity to observation, such as from a video camera, may influence behavior.
+Second, wearable audio recording is more \emph{unobtrusive}.
+Participants' reactivity to observation, such as from a video camera, may influence behavior more compared with a sensor that records only motion or audio data.
 For example, caregivers spoke more frequently to infants during a video-recorded portion of a home recording compared with audio-only segments captured by a LENA® device (Bergelson, Amatuni, et al., 2019).
@@
 For example, correlations between human transcribed counts of adult words and child vocalizations against LENA®'s automatic counts were strong, \emph{r} = .698 and \emph{r} = .649, respectively (Cristia et al., 2020).
-For other metrics, such as the number of ``conversational turns'' between the child and communicative partners and the identification of male speakers, agreement was poor.
+For other metrics, such as the number of ``conversational turns'' between the child and communicative partners, agreement was poor (\emph{r} = .364).
 Thus, for some use cases (and for some metrics), long form audio recordings provide a mobile, unobtrusive way to automatically score real-time data over a full day.
@@
 Most participants were tested in a single session (\emph{n} = 15), but 7 participants contributed between 2-4 sessions as part of an ongoing longitudinal study.
+Although including more data from some participants could over-represent their characteristics in the model, we reasoned that this drawback was outweighed by having more available data to use for training.
 Only 1 session was excluded due to a technical error---one of the four IMU sensors failed to record, resulting in an unusable set of data for classification.
@@
 Across sessions, younger infants' age ranged from 3.8 to 7.2 months (\emph{M} = 5.2) and older infants' age ranged from 10.7 to 14.2 months (\emph{M} = 11.7).
+The total number of recording sessions (34) exceeded the number of sessions employed in comparable past work: 10 in @Nam and Park (2013), 15 in Franchak et al.
+(2021), 22 in Airaksinen et al.
+(2020), and 33 in Yao et al.
+(2019).
 
@@
 A pocket over the hip and a pocket just above the ankle were sewn on the lateral surface of the right and left legs of the garment.
-Each sensor recorded accelerometer and gyroscope data at 62.5 Hz throughout the day, with sufficient battery and onboard storage to record for approximately 12 hours.
-Infants also wore a LENA® recorder throughout the day in the front pocket of a LENA® shirt, located near the infant's chest.
-
-Video recordings were captured using an action camera on a miniature tripod (Insta360 ONE R) that caregivers placed in the room that the infant was in.
+Each sensor recorded accelerometer and gyroscope data throughout the day, with sufficient battery and onboard storage to record for approximately 12 hours.
+We chose a sampling rate of 62.5 Hz (on of several device presets) based on prior work that used rates of 50-64 Hz (Airaksinen et al., 2020; Franchak et al., 2021; Yao et al., 2019).
+Infants also wore a LENA® recorder throughout the day in the front pocket of a LENA® shirt, located near the infant's chest, to determine whether data could be simultaneously recorded from the LENA® and IMU sensors.
+However, LENA® data were not analyzed as part of the current study.
+
+Video recordings were captured at 30 Hz using an action camera on a miniature tripod (Insta360 ONE R) that caregivers placed in the room that the infant was in.
 The proprietary ``Boosted Battery Base'' attached to the action camera to allow for a total of 3 hours of recording.
@@
 
-Because the camera only had the battery life to record for \textasciitilde3 hours (split into two 90-minute video files), this divided the day into different periods for analysis.
+Because the camera only had the battery life to record for \textasciitilde3 hours (automatically split into two 90-minute video files), this divided the day into different periods for analysis.
 As seen in the bottom of Figure \ref{fig:exemplar-timeline}), the \emph{video period} comprised the first three hours of recording starting from the researcher's arrival when they turned on the camera.
@@
 This video recorded the next 90 minutes of natural activity.
-However, because there was a variable gap of \textasciitilde40 s between the two videos, synchronization in the distal comparison video was coarser, with estimated temporal offsets of \textasciitilde5 s in either direction.
+However, because of a limitation in the camera, there was a variable gap of \textasciitilde40 s between the two videos, synchronization in the distal comparison video was coarser, with estimated temporal offsets of \textasciitilde5 s in either direction.
 
@@
 
-Among participants who wore the sensors from morning to bedtime, the majority of the time during the day either resulted in usable body position data or was excluded due to caregiver-reported naps (gray periods in Figure \ref{fig:timelines}).
+Among participants who wore the sensors from morning to bedtime, all infants wore the sensor garment during 100\% of the intended recording period, excluding the times caregivers were asked to remove the garment (naps, diaper changes, trips out of the house), based on caregiver logs of wear time.
+No caregiver mentioned needing to remove the garment for any other reason (such as infant discomfort), meaning that the majority of the time during the day either resulted in usable body position data or was excluded due to caregiver-reported naps (gray periods in Figure \ref{fig:timelines}).
 The total length of the recording period ranged from 6.92-11.75 hours with a median of 9.04 hours.
@@
 
-Third, although the goal was to collect fully naturalistic behavior, we opted to restrict recording to times when participants were in the home.
+Third, although the goal was to collect fully naturalistic behavior, our recording protocol led to some changes that might have affected (or missed) some behaviors.
+We opted to restrict recording to times when participants were in the home.
 Inertial sensors can travel with participants, and there is no doubt that infants could have worn the sensors out of the home on errands.
 However, we do not yet have training data to detect periods of transportation (e.g., moving in a car or in a stroller) that might add noise and/or lead to misclassification of body positions.
-Future work should aim to collect data in these scenarios so that body position can be captured in or out of the home.
+It is also unknown whether participants might have behaved differently had they not worn the sensors.
+The added bulk in the garment from the sensors (including the LENA® recorder) might have made some positions more uncomfortable, such as lying prone with the LENA® recorder worn on the chest, and/or influenced how caregivers chose to position infants throughout the day.
 

### 2023-06-16 manuscript.Rmd
group_split_sensor_reduce

@@
 footnotelist      : no
-csl               : apa.csl
+biblio-style      : "apa"
 classoption       : "man"
-output            : papaja::apa6_pdf
+output:
+  papaja::apa6_pdf:
+    citation_package: biblatex
 header-includes:

### 2023-06-16 manuscript.tex
group_split_sensor_reduce

@@
 \fi
+\usepackage[style=apa,]{biblatex}
+\addbibresource{master.bib}
 \IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
@@
 Changes in \emph{body position}---whether infants are supine on their backs, prone on their bellies, sitting, upright, or held by a caregiver---have immediate consequences for vision, object exploration, and social interaction.
-When sitting and upright, infants have a better view of faces and distant objects compared to their view while in a prone position (Franchak et al., 2018; Kretch et al., 2014; Luo \& Franchak, 2020).
-Learning to sit facilitates object exploration (Soska \& Adolph, 2014).
-Upright walking changes how infants and caregivers interact; walking infants move farther away from caregivers and share toys in different ways compared to infants crawling in a prone position (Chen et al., 2022; Karasik et al., 2011).
-As infants grow older and acquire new abilities, such as independent sitting and walking, they spend more time sitting and upright and less time held, supine, and prone (Adolph \& Tamis-LeMonda, 2014; Franchak et al., 2018; Franchak, 2019; Thurman \& Corbetta, 2017).
-Caregivers provide different learning input to infants depending on infants' body position (Karasik et al., 2014; Kretch et al., 2022; West \& Iverson, 2021).
+When sitting and upright, infants have a better view of faces and distant objects compared to their view while in a prone position \autocite{Freeplay,LuoFranchak2020,CWW}.
+Learning to sit facilitates object exploration \autocite{SoskaAdolph2014}.
+Upright walking changes how infants and caregivers interact; walking infants move farther away from caregivers and share toys in different ways compared to infants crawling in a prone position \autocite{ChenSchneider2022,Karasik2011}.
+As infants grow older and acquire new abilities, such as independent sitting and walking, they spend more time sitting and upright and less time held, supine, and prone \autocite{Survey,Thurman2017,Freeplay,AdolphCTL2014}.
+Caregivers provide different learning input to infants depending on infants' body position \autocite{KretchKoziol2022,WestIverson2021,Karasik2014}.
 For example, infants' walking movements spur caregivers to talk about actions rather than just describe objects.
-Thus, characterizing individual differences in the day-to-day accumulation of body position experiences informs developmental theory by revealing differential opportunities for learning (Franchak, 2020).
+Thus, characterizing individual differences in the day-to-day accumulation of body position experiences informs developmental theory by revealing differential opportunities for learning \autocite{CurrentOpinion}.
 
@@
 Finally, we discuss the advantages of using inertial sensing to classify motor behavior.
-Despite promising past results in brief, supervised sessions (Airaksinen et al., 2022, 2020; Franchak et al., 2021), the current investigation takes a needed step forward by testing accuracy over long, unsupervised recordings in what we term a \emph{distal comparison}.
+Despite promising past results in brief, supervised sessions \autocite{FranchakScott2021,AiraksinenRasanen2020,AiraksinenGallen2022}, the current investigation takes a needed step forward by testing accuracy over long, unsupervised recordings in what we term a \emph{distal comparison}.
 
@@
 Closed-source LENA® algorithms analyze the audio recordings to provide automatic counts of useful metrics, such as the number of words spoken by adults in the vicinity of the infant.
-Other long-form audio methods rely on custom-built recorders (Wass et al., 2022), apply alternative classification algorithms to LENA® data (Micheletti et al., 2022; Räsänen et al., 2020), or manually transcribe audio recorded by LENA® devices to improve accuracy or identify behaviors beyond the built-in categories (Bergelson, Casillas, et al., 2019; Mendoza \& Fausey, 2021).
+Other long-form audio methods rely on custom-built recorders \autocite{WassPhillips2022}, apply alternative classification algorithms to LENA® data \autocite{RasanenSeshadri2020,MichelettiYao2022}, or manually transcribe audio recorded by LENA® devices to improve accuracy or identify behaviors beyond the built-in categories \autocite{Bergelson2019Corpus,MendozaFausey2021}.
 
 Long-form audio recordings have had a transformational impact on language development research by allowing researchers to characterize opportunities for learning in daily life.
-For example, measuring the amount of speech heard by infants in the home (Weisleder \& Fernald, 2013) or in a daycare setting (Perry et al., 2018) revealed how individual differences in speech input predict later vocabulary.
+For example, measuring the amount of speech heard by infants in the home \autocite{Weisleder2013} or in a daycare setting \autocite{PerryPrince2018} revealed how individual differences in speech input predict later vocabulary.
 Full-day language recording synchronized with other data sources allows researchers to identify how auditory input and vocal production interact with other processes.
 Beyond individual differences in aggregated data, long-form recordings reveal the temporal schedule of experiences.
-For example, infants' daily experiences hearing music are clustered in time, with ``bursty'' episodes of music presence separated by relatively long periods during which music is absent (Mendoza \& Fausey, 2022).
+For example, infants' daily experiences hearing music are clustered in time, with ``bursty'' episodes of music presence separated by relatively long periods during which music is absent \autocite{MendozaFausey2022}.
 In addition, synchronizing audio recordings with other data sources extends researchers' ability to characterize daily experiences.
-Linking LENA® speech measurements to repeated, time-stamped text-message surveys about infant device placement revealed that infants heard less caregiver speech during periods that they were restrained in devices such as swings, exersaucers, and car seats (Malachowski et al., 2023).
-A custom-built wearable ECG and audio recorder allowed Wass et al.
-(2022) to discover that infant arousal increases the likelihood of infant vocalization across the day.
+Linking LENA® speech measurements to repeated, time-stamped text-message surveys about infant device placement revealed that infants heard less caregiver speech during periods that they were restrained in devices such as swings, exersaucers, and car seats \autocite{MalachowskiSalo2023}.
+A custom-built wearable ECG and audio recorder allowed \textcite{WassPhillips2022} to discover that infant arousal increases the likelihood of infant vocalization across the day.
 
@@
 Participants' reactivity to observation, such as from a video camera, may influence behavior more compared with a sensor that records only motion or audio data.
-For example, caregivers spoke more frequently to infants during a video-recorded portion of a home recording compared with audio-only segments captured by a LENA® device (Bergelson, Amatuni, et al., 2019).
+For example, caregivers spoke more frequently to infants during a video-recorded portion of a home recording compared with audio-only segments captured by a LENA® device \autocite{Bergelson2019Input}.
 Third and fourth, recordings capture \emph{real-time data} over a \emph{full day}.
 The ability to record real-time data is vital for making inferences about processes that happen on the timescale of minutes or even seconds within an individual as opposed to comparisons of aggregated data between infants.
-Synchronizing real-time data to other data streams helps to reveal sources of variability within an individual (Malachowski et al., 2023; e.g., Wass et al., 2022).
-Full-day recordings are essential for capturing experiences across the heterogeneity of daily routines that moderate behavior (e.g., play, feeding, errands) (Kadooka et al., 2021, April; Tamis-LeMonda et al., 2018).
-``Burstiness'' of behavior means that long recordings are needed to capture clusters of events amid long periods in which they may be absent (Barbaro \& Fausey, 2022; Warlaumont et al., 2021).
+Synchronizing real-time data to other data streams helps to reveal sources of variability within an individual \autocites[e.g.,][]{WassPhillips2022,MalachowskiSalo2023}.
+Full-day recordings are essential for capturing experiences across the heterogeneity of daily routines that moderate behavior (e.g., play, feeding, errands) \autocite{KadookaCaufield2021,Tamis-LeMondaCustode2018}.
+``Burstiness'' of behavior means that long recordings are needed to capture clusters of events amid long periods in which they may be absent \autocite{BarbaroFausey2022,WarlaumontSobowale2021}.
 Fifth, \emph{automatic classification} means that the approach can scale to analyze large numbers of participants over long recordings without the bottleneck of manual annotation/transcription.
@@
 An independent assessment of the LENA® algorithms found mixed results about the accuracy of different measurements.
-For example, correlations between human transcribed counts of adult words and child vocalizations against LENA®'s automatic counts were strong, \emph{r} = .698 and \emph{r} = .649, respectively (Cristia et al., 2020).
+For example, correlations between human transcribed counts of adult words and child vocalizations against LENA®'s automatic counts were strong, \emph{r} = .698 and \emph{r} = .649, respectively \autocite{CristiaLavechin2020}.
 For other metrics, such as the number of ``conversational turns'' between the child and communicative partners, agreement was poor (\emph{r} = .364).
@@
 Video observation is the most common way of measuring infant motor behavior in home recordings.
-Most often, an experimenter with a handheld camera follows infants from room to room to ensure that their movements are visible throughout the recording session (Chen et al., 2022; Herzberg et al., 2021; Karasik et al., 2011).
+Most often, an experimenter with a handheld camera follows infants from room to room to ensure that their movements are visible throughout the recording session \autocite{Karasik2011,ChenSchneider2022,HerzbergFletcher2021}.
 The primary advantage of video recording is that it captures real-time behavior.
@@
 Logistically, an experimenter cannot follow behind infants to record their behavior from morning to night (and were they to do so, they would likely alter infants' and caregivers' behavior).
-Typical video recording sessions last 45-120 minutes (Chen et al., 2022; Herzberg et al., 2021; Karasik et al., 2011), far short of capturing the variety of activities across the full daily routine.
+Typical video recording sessions last 45-120 minutes \autocite{Karasik2011,ChenSchneider2022,HerzbergFletcher2021}, far short of capturing the variety of activities across the full daily routine.
 Even if full day videos were available, the lack of suitable automatic classification tools means that the human cost of annotation makes it difficult for video methods to scale to testing large numbers of participants.
-Our annotation of body position takes approximately 2-5 hours to complete for every hour of video (depending on how often infants switch positions), meaning that a full ``waking day'' of approximately 11 hours for a 12-month-old (Galland et al., 2012) could take 22-55 hours of labor to fully annotate.
+Our annotation of body position takes approximately 2-5 hours to complete for every hour of video (depending on how often infants switch positions), meaning that a full ``waking day'' of approximately 11 hours for a 12-month-old \autocite{GallandTaylor2012} could take 22-55 hours of labor to fully annotate.
 
 In contrast, survey methods such as daily diaries/inventories or ecological momentary assessment (EMA) are mobile, unobtrusive, can be applied across an entire day, and do not need laborious annotation.
-Diary studies provide caregivers with logs or structured interviews to estimate from memory how much time infants spend in particular activities (Karasik et al., 2022; Majnemer \& Barr, 2005).
-Ecological momentary assessment uses text-message or app-based notifications to prompt caregivers to make repeated estimates of infants' behavior over the course of a day (Franchak, 2019; Kadooka et al., 2021, April).
+Diary studies provide caregivers with logs or structured interviews to estimate from memory how much time infants spend in particular activities \autocite{MajnemerBarr2005,KarasikKuchirko2022}.
+Ecological momentary assessment uses text-message or app-based notifications to prompt caregivers to make repeated estimates of infants' behavior over the course of a day \autocite{Survey,KadookaCaufield2021}.
 Although the EMA responses are valuable in aggregate, survey methods lack the real-time temporal resolution to describe moment-to-moment changes in behavior.
@@
 
-Measuring infant movement with inertial movement units (IMUs) is a promising avenue for long form recordings of motor behavior in the home (Barbaro, 2019; Bruijns et al., 2020; Cliff et al., 2009; Lobo et al., 2019).
+Measuring infant movement with inertial movement units (IMUs) is a promising avenue for long form recordings of motor behavior in the home \autocite{CliffReilly2009,Barbaro2019,BruijnsTruelove2020,LoboHall2019}.
 Lightweight sensors (10-30 g) can be embedded in garments to make recordings fully \emph{mobile}, and they are \emph{unobtrusive} because they do not require a researcher's presence nor do they record sensitive audio/video that might influence participants' behavior.
 Many commercially-available and inexpensive IMUs have \textgreater{} 12 hour battery life with onboard storage to record \emph{real-time}, \emph{full-day} motion data at a sampling rate that is higher than typical video (e.g., 50-100 Hz).
-Indeed, past work has successfully recorded the rate of leg kicks (e.g., Deng et al., 2019) and activity intensity (Schneller et al., 2017) in infants and children across multiple days.
+Indeed, past work has successfully recorded the rate of leg kicks \autocite[e.g.,][]{DengTrujillo-Priego2019} and activity intensity \autocite{SchnellerBentsen2017} in infants and children across multiple days.
 
@@
 The difficulty of the classification task depends on the categories of interest.
-More basic aspects of movement, such as overall activity intensity, can be identified by taking the magnitude of acceleration (irrespective of direction) and applying thresholds or cut-points to define when a movement has occurred at a particular intensity (Armstrong et al., 2019; e.g., Hager et al., 2017).
+More basic aspects of movement, such as overall activity intensity, can be identified by taking the magnitude of acceleration (irrespective of direction) and applying thresholds or cut-points to define when a movement has occurred at a particular intensity \autocites[e.g.,][]{HagerTilton2017,ArmstrongCovington2019}.
 
@@
 An upright infant can be standing still or can be walking briskly across the room.
-A prone infant can be stationary in ``tummy time'', or they can crawl in a myriad of ways (Adolph et al., 1998).
+A prone infant can be stationary in ``tummy time'', or they can crawl in a myriad of ways \autocite{AdolphVereijken1998}.
 Moreover, the configuration of the arms, legs, and torso within a body position can vary greatly in everyday contexts.
@@
 Indeed, creating an all-encompassing set of rules for how to annotate sitting from video is no trivial task because of the various ways that sitting can occur in daily life.
-Finally, caregivers frequently pick up and transport infants, creating motion signals that need to be differentiated from independent activity (Kwon et al., 2019; Patel et al., 2019).
-Although Greenspan et al.
-(2021) were able to use cut-point definitions about infant trunk angle to categorize body positions, this method was not able to differentiate times that infants were held by caregivers off the ground.
-Thus, modern approaches to human activity recognition have used machine learning to classify activity categories based on features derived from IMU data in adults (Arif \& Kattan, 2015; Preece et al., 2009), children (Nam \& Park, 2013; Ren et al., 2016; Stewart et al., 2018), and infants (Airaksinen et al., 2020; Franchak et al., 2021; Yao et al., 2019).
+Finally, caregivers frequently pick up and transport infants, creating motion signals that need to be differentiated from independent activity \autocite{KwonZavos2019,PatelShi2019}.
+Although \textcite{GreenspanCunha2021} were able to use cut-point definitions about infant trunk angle to categorize body positions, this method was not able to differentiate times that infants were held by caregivers off the ground.
+Thus, modern approaches to human activity recognition have used machine learning to classify activity categories based on features derived from IMU data in adults \autocite{Preece2009,Arif2015}, children \autocite{NamPark2013,RenDing2016,StewartNarayanan2018}, and infants \autocite{YaoPlotz2019,FranchakScott2021,AiraksinenRasanen2020}.
 Synchronized video with ground-truth human annotations creates training data for a machine learning algorithm, such as a random forest model, that can be later used to predict categories from IMU data that was not annotated.
@@
 Three prior investigations have used different machine learning techniques to categorize infant body position from IMUs towards the goal of collecting full-day data.
-Airaksinen et al.
-(2020) tested 4- to 8-month-olds in a laboratory visit, and found 95\% accuracy in distinguishing between body position categories that crawling infants could perform on the floor (excluding times that infants were held by caregivers).
-Using a wider age range of 6-18 months, Franchak et al.
-(2021) found 98\% accuracy (\emph{kappa} = 95\%) in a laboratory validation study in categorizing body position that included infants who could both crawl and walk and also included a category for caregiver holding.
-Most recently, Airaksinen et al.
-(2022) conducted a validation study of body position classification in either a home or clinic testing in 4- to 19-month-olds, refining their previous method to detect moments that infants were carried by caregivers.
+\textcite{AiraksinenRasanen2020} tested 4- to 8-month-olds in a laboratory visit, and found 95\% accuracy in distinguishing between body position categories that crawling infants could perform on the floor (excluding times that infants were held by caregivers).
+Using a wider age range of 6-18 months, \textcite{FranchakScott2021} found 98\% accuracy (\emph{kappa} = 95\%) in a laboratory validation study in categorizing body position that included infants who could both crawl and walk and also included a category for caregiver holding.
+Most recently, \textcite{AiraksinenGallen2022} conducted a validation study of body position classification in either a home or clinic testing in 4- to 19-month-olds, refining their previous method to detect moments that infants were carried by caregivers.
 Classification accuracy did not vary between lab and home settings, and was generally high (95\%, \emph{kappa} = .93).
@@
 Accordingly, the overarching goal of the current study is to test the validity of long-form body position recording in the home during unsupervised, everyday behavior.
-Supervised recordings from past work (Airaksinen et al., 2022, 2020; Franchak et al., 2021), whether in the home or in the lab, let researchers set up the situation to encourage or restrict certain behaviors.
+Supervised recordings from past work \autocite{FranchakScott2021,AiraksinenRasanen2020,AiraksinenGallen2022}, whether in the home or in the lab, let researchers set up the situation to encourage or restrict certain behaviors.
 Prior work focused on ``free play'', in which caregivers were asked to play with the infant without restraining the infant or otherwise shaping how they could move.
@@
 The \textbf{first goal} of the current study was to determine the accuracy of body position classification during the proximal comparison period using this novel, semi-supervised procedure in participants' homes.
-Past work found better performance using ``individual models''---models that were trained on one participant's data to predict their later behavior---compared with a ``group model'' that aggregated data from all infants to create a single body position classifier (Franchak et al., 2021), so we compared both modeling approaches in the current investigation.
+Past work found better performance using ``individual models''---models that were trained on one participant's data to predict their later behavior---compared with a ``group model'' that aggregated data from all infants to create a single body position classifier \autocite{FranchakScott2021}, so we compared both modeling approaches in the current investigation.
 
@@
 Thus, we determined whether full-day body position measurements conformed to expected age differences in body position that have been identified from previous video and survey methods.
-Based on past results (Franchak, 2019), infants should spend increasingly more time sitting and upright but less time supine over the age range tested (4 to 14 months).
+Based on past results \autocite{Survey}, infants should spend increasingly more time sitting and upright but less time supine over the age range tested (4 to 14 months).
 
@@
 Across sessions, younger infants' age ranged from 3.8 to 7.2 months (\emph{M} = 5.2) and older infants' age ranged from 10.7 to 14.2 months (\emph{M} = 11.7).
-The total number of recording sessions (34) exceeded the number of sessions employed in comparable past work: 10 in @Nam and Park (2013), 15 in Franchak et al.
-(2021), 22 in Airaksinen et al.
-(2020), and 33 in Yao et al.
-(2019).
+The total number of recording sessions (34) exceeded the number of sessions employed in comparable past work: 10 in \textcite{NamPark2013}, 15 in \textcite{FranchakScott2021}, 22 in \textcite{AiraksinenRasanen2020}, and 33 in \textcite{YaoPlotz2019}.
 
@@
 Each sensor recorded accelerometer and gyroscope data throughout the day, with sufficient battery and onboard storage to record for approximately 12 hours.
-We chose a sampling rate of 62.5 Hz (on of several device presets) based on prior work that used rates of 50-64 Hz (Airaksinen et al., 2020; Franchak et al., 2021; Yao et al., 2019).
+We chose a sampling rate of 62.5 Hz (one of the available preset rates) based on prior work that used rates of 50-64 Hz \autocite{FranchakScott2021,AiraksinenRasanen2020,YaoPlotz2019}.
 Infants also wore a LENA® recorder throughout the day in the front pocket of a LENA® shirt, located near the infant's chest, to determine whether data could be simultaneously recorded from the LENA® and IMU sensors.
@@
 In order to synchronize all three recording devices, the researcher began by first turning on the video camera and the LENA audio device (the IMUs were already configured and placed before arrival because they required a proprietary sensor dock to begin recording in the laboratory).
-To create an easily recognizable synchronization point between the video recording and IMU data, the researcher dropped or struck the leggings (containing the IMUs) on a surface in view of the camera, as in Franchak et al.
-(2021).
+To create an easily recognizable synchronization point between the video recording and IMU data, the researcher dropped or struck the leggings (containing the IMUs) on a surface in view of the camera, as in \textcite{FranchakScott2021}.
 All the equipment---once recording and with synchronization information recorded---was placed inside a large bucket and left outside the family's front door.
@@
 
-The proximal and distal comparison videos for each participant were annotated by trained human coders to identify infant \emph{body position} into one of 5 mutually-exclusive categories following the definitions in prior work (Franchak et al., 2021): supine, prone, sitting, upright, or held by caregiver.
+The proximal and distal comparison videos for each participant were annotated by trained human coders to identify infant \emph{body position} into one of 5 mutually-exclusive categories following the definitions in prior work \autocite{FranchakScott2021}: supine, prone, sitting, upright, or held by caregiver.
 All coding was done using Datavyu software (datavyu.org).
@@
 
-The same machine learning classification process was used as in prior work (Franchak et al., 2021).
+The same machine learning classification process was used as in prior work \autocite{FranchakScott2021}.
 Using the synchronization point, human-coded body annotations from video were linked to the corresponding times in the IMU time series data.
-A single, merged dataset was created with synchronized accelerometer signals (in three orientations: X, Y, and Z) and gyroscope signals (in three orientations: roll, pitch, and yaw) for each of the four sensors (left hip, right hip, left ankle, right ankle) with the corresponding timestamp and body position code using the \emph{timetk} package (Dancho \& Vaughan, 2023) and the \emph{lubridate} package (Grolemund \& Wickham, 2011) in R version 4.1.2 (R Core Team, 2021).
+A single, merged dataset was created with synchronized accelerometer signals (in three orientations: X, Y, and Z) and gyroscope signals (in three orientations: roll, pitch, and yaw) for each of the four sensors (left hip, right hip, left ankle, right ankle) with the corresponding timestamp and body position code using the \emph{timetk} package \autocite{DanchoVaughan2023} and the \emph{lubridate} package \autocite{GrolemundWickham2011} in R version 4.1.2 \autocite{R42}.
 
 Classification training and prediction was conducted on a windowed dataset that summarized the raw, 62.5 Hz motion signals within 4-s windows.
-Overlapping moving windows captured 4-s (250 samples) of data starting each second, which is a common unit of analysis in prior studies of human activity classification (Airaksinen et al., 2020; Franchak et al., 2021; Nam \& Park, 2013).
+Overlapping moving windows captured 4-s (250 samples) of data starting each second, which is a common unit of analysis in prior studies of human activity classification \autocite{FranchakScott2021,NamPark2013,AiraksinenRasanen2020}.
 The 250 samples of data in each window were aggregated into a variety of motion features---summary statistics that could be fed into the machine learning model.
@@
 Different training/testing set combinations were used to address different goals; here we describe the general procedure of how a classifier was derived from a training set.
-Random forest models (Breiman, 2001) were trained to predict the human coded body position label for each window from the set of 436 motion features using the \emph{randomForest} package (Liaw et al., 2002).
+Random forest models \autocite{Breiman2001} were trained to predict the human coded body position label for each window from the set of 436 motion features using the \emph{randomForest} package \autocite{LiawWiener2002}.
 Random forest models use random subsets of features and random subsets of data across many iterations; this random subsetting creates models that are less likely to be overfit.
-The \emph{tidymodels} set of packages in R (Kuhn \& Wickham, 2020) was used to tune the ``mtry'' and ``ntrees'' hyperparameters, which control the number of columns sampled and the total number of trees created when modeling.
+The \emph{tidymodels} set of packages in R \autocite{tidymodels} was used to tune the ``mtry'' and ``ntrees'' hyperparameters, which control the number of columns sampled and the total number of trees created when modeling.
 We use an mtry parameter of 44 and and ntree parameter of 550; however, changing hyperparameters had little impact on model performance.
@@
 However, in a second Github repository (\url{https://github.com/JohnFranchak/body_position_classification_ms}) we share the full results of those computations: The dataset of windowed motion features with corresponding body position codes used to validate the method.
-This reproducible manuscript created in RMarkdown and \emph{papaja} (Aust \& Barth, 2022) can be regenerated from those data files for full computational transparency.
+This reproducible manuscript created in RMarkdown and \emph{papaja} \autocite{papaja} can be regenerated from those data files for full computational transparency.
 
@@
 The high degree of temporal synchronization between video and motion data during this period makes it possible to link human-coded body position annotations to each 4-s window of motion data, providing ground truth data for model training and testing.
-As in past work (Franchak et al., 2021), we compared two types of models: \emph{group models} and \emph{individual models}.
+As in past work \autocite{FranchakScott2021}, we compared two types of models: \emph{group models} and \emph{individual models}.
 To assess the accuracy of body position classification for each recording session, we reserved the (temporally) last 25\% of a session's proximal comparison data as the testing set.
@@
 Overall accuracy for group models (\emph{M} = 0.85) was slightly lower than accuracy for individual models (\emph{M} = 0.92).
-Although overall accuracy from our semi-supervised, in-home data collection did not match the near-perfect accuracy (.95-.98) found in prior in-lab studies (Airaksinen et al., 2020; Franchak et al., 2021), both models approached the level of agreement found between two human coders (\emph{M} = .906).
+Although overall accuracy from our semi-supervised, in-home data collection did not match the near-perfect accuracy (.95-.98) found in prior in-lab studies \autocite{FranchakScott2021,AiraksinenRasanen2020}, both models approached the level of agreement found between two human coders (\emph{M} = .906).
 Most likely, lower accuracy in the current study results from the more variable and complex behavior observed in a a semi-supervised setting rather than from a difference in the quality of the classification model; although the first part of the proximal period was guided by the experimenter during a brief phone call, the remainder of the proximal period included natural behavior.
@@
 Similar to overall accuracy, kappa values were strong for both model types with group kappas (\emph{M} = 0.75) somewhat worse compared with individual kappas (\emph{M} = 0.82).
-Guidelines for interpreting kappa statistics (Landis \& Koch, 1977) consider 0.81--1.00 ``Almost Perfect,'' 0.61--0.80 ``Substantial,'' 0.41--0.60 ``Moderate,'' 0.21-- 0.40 ``Fair,'' and 0--0.20 ``Slight to Poor'', indicating that agreement for most group and individual model predictions fell in the Substantial to Almost Perfect range.
-
-As in past work (Airaksinen et al., 2020; Franchak et al., 2021), all body positions were accurately classified even though performance varied somewhat between positions.
+Guidelines for interpreting kappa statistics \autocite{LandisKoch1977} consider 0.81--1.00 ``Almost Perfect,'' 0.61--0.80 ``Substantial,'' 0.41--0.60 ``Moderate,'' 0.21-- 0.40 ``Fair,'' and 0--0.20 ``Slight to Poor'', indicating that agreement for most group and individual model predictions fell in the Substantial to Almost Perfect range.
+
+As in past work \autocite{FranchakScott2021,AiraksinenRasanen2020}, all body positions were accurately classified even though performance varied somewhat between positions.
 As Table \ref{tab:metrics-by-class} shows, mean kappa statistics were strongest for prone (group \emph{M} = 0.860, individual \emph{M} = 0.841) and supine (group \emph{M} = 0.764, individual \emph{M} = 0.912).
@@
 As previously mentioned, distal videos had only coarse temporal synchrony with motion recordings (misalignment of up to \textasciitilde{} 5 s in either direction).
-This misalignment precluded calculating accuracy based on the proportion of matching events, so we used another technique following past work (Franchak et al., 2021; Yao et al., 2019).
+This misalignment precluded calculating accuracy based on the proportion of matching events, so we used another technique following past work \autocite{YaoPlotz2019,FranchakScott2021}.
 For each session, we summed the amount of time infants were predicted to be in each of the 5 body position categories from the model and compared that to the summed time for the body positions based on human coding.
@@
 
-In order to show the convergent validity of the classification method, we demonstrate how the full-day recordings can reveal individual differences in body position according to age, replicating age differences revealed in past work (Franchak, 2019).
+In order to show the convergent validity of the classification method, we demonstrate how the full-day recordings can reveal individual differences in body position according to age, replicating age differences revealed in past work \autocite{Survey}.
 Table \ref{tab:agetable} summarizes the percent of time that younger and older infants spent in each body position out of their awake samples, as predicted by group and individual models.
-Past work using sparse EMA surveys across the day (Franchak, 2019) found that from 3-12 months, infants spent less time held, reclined, and supine but spent increasingly more time sitting and upright.
+Past work using sparse EMA surveys across the day \autocite{Survey} found that from 3-12 months, infants spent less time held, reclined, and supine but spent increasingly more time sitting and upright.
 Prone time changed little.
 The most straightforward comparisons between the two investigations are upright and prone, because they were defined identically.
-In Franchak (2019), upright time was 0.6-5.5\% of the time for infants 3-6 months, increasing to 22.0\% at 12 months.
+In \textcite{Survey}, upright time was 0.6-5.5\% of the time for infants 3-6 months, increasing to 22.0\% at 12 months.
 Results from the current study were similar: younger infants' upright time estimated from full-day group models was 5.41\% compared with 18.24\% for the older group.
-In Franchak (2019), prone time was 2.9-9.2\% of the time for infants 3-6 months, and was 7.2\% at 12 months.
+In \textcite{Survey}, prone time was 2.9-9.2\% of the time for infants 3-6 months, and was 7.2\% at 12 months.
 In the current study, prone time was 14.03\% for younger infants and 15.09\% for older infants.
-Held, sitting, and supine are more difficult to compare because prior work (Franchak, 2019) considered held as any time that infants were in caregivers' arms or on their laps, however, we counted laying or sitting on caregivers' laps as supine and sitting, respectively.
-Further complicating the comparison, Franchak (2019) included a reclined category that we would classify as either supine or sitting (whichever it was closer to).
+Held, sitting, and supine are more difficult to compare because prior work \autocite{Survey} considered held as any time that infants were in caregivers' arms or on their laps, however, we counted laying or sitting on caregivers' laps as supine and sitting, respectively.
+Further complicating the comparison, \textcite{Survey} included a reclined category that we would classify as either supine or sitting (whichever it was closer to).
 Although the numbers are less directly comparable, our results are similar in showing that held and supine time was less for older infants than younger infants (Table \ref{tab:agetable}).
 Sitting was greater for older infants (44.39\%) compared with younger infants (28.91\%).
-Moreover, sitting was the most frequent body position followed by upright in 12-month-olds (Franchak, 2019), just as we observed in the current study.
+Moreover, sitting was the most frequent body position followed by upright in 12-month-olds \autocite{Survey}, just as we observed in the current study.
 Overall, the estimates from the current investigation are consistent with age differences found using a different methodology.
@@
 
-How did classification in the current investigation compare to prior work (Airaksinen et al., 2022, 2020; Franchak et al., 2021)?
+How did classification in the current investigation compare to prior work \autocite{FranchakScott2021,AiraksinenRasanen2020,AiraksinenGallen2022}?
 Across body positions, median accuracy was 89.40\% for group models (Kappa = 0.77) and 93.30\% for individual models (Kappa = 0.85) during the proximal comparison.
 Although these accuracy and kappa values are slightly lower than in past work (accuracy 95-98\%, kappas .93-.95), we note that those past values were obtained under ideal circumstances.
-A researcher applied the sensors in the lab (Airaksinen et al., 2020; Franchak et al., 2021) or home (Airaksinen et al., 2022) and set the stage for how infants and caregivers should interact.
-For example, both Franchak et al.
-(2021) and Airaksinen et al.
-(2022) instructed the caregivers that the recorded sessions should mimic a ``playtime'', which would encourage more common activities that would be easier to classify (e.g., crawling and sitting on the floor) and discourage idiosyncratic and potentially more challenging activities (e.g., positioning in infant-specific furniture like exersaucers and walkers, sitting or lying for long periods while eating or nursing).
+A researcher applied the sensors in the lab \autocite{FranchakScott2021,AiraksinenRasanen2020} or home \autocite{AiraksinenGallen2022} and set the stage for how infants and caregivers should interact.
+For example, both \textcite{FranchakScott2021} and \textcite{AiraksinenGallen2022} instructed the caregivers that the recorded sessions should mimic a ``playtime'', which would encourage more common activities that would be easier to classify (e.g., crawling and sitting on the floor) and discourage idiosyncratic and potentially more challenging activities (e.g., positioning in infant-specific furniture like exersaucers and walkers, sitting or lying for long periods while eating or nursing).
 Examining the timelines in Figure \ref{fig:timelines} shows how much variability there is between and within sessions in the earlier parts of the recording that were used for training and testing.
@@
 Another comparison we can make is to the accuracy of the LENA® device's automated metrics.
-Correlations between human and model frequencies were \emph{r} = .698 for adult word count and \emph{r} = .649 for child word count (Cristia et al., 2020).
+Correlations between human and model frequencies were \emph{r} = .698 for adult word count and \emph{r} = .649 for child word count \autocite{CristiaLavechin2020}.
 Even including outliers, correlations between human and model-predicted body position were \emph{r} = 0.80 (group) and \emph{r} = 0.94 (individual), stronger than the LENA® automated counts.
@@
 
-In their validation study, Cristia et al.
-(2020) make the recommendation that researchers using LENA® manually annotate a portion of data collected for their particular use case to ensure that automated LENA® metrics are reliable for the participants and environment pertinent to the research question.
+In their validation study, \textcite{CristiaLavechin2020} make the recommendation that researchers using LENA® manually annotate a portion of data collected for their particular use case to ensure that automated LENA® metrics are reliable for the participants and environment pertinent to the research question.
 Similarly, collecting some video data for each participant in future long-form motor behavior recordings provides a way to check reliability and identify errant predictions.
@@
 However, our openly shared data and code allow researcher to experiment with different sets of motion features, different sets of training data, different machine learning algorithms, and different hyperparameter values.
-As in past work (Airaksinen et al., 2020), performance was best when using all four sensors and degraded when using any subset of two (i.e., hips only, left leg only).
+As in past work \autocite{AiraksinenRasanen2020}, performance was best when using all four sensors and degraded when using any subset of two (i.e., hips only, left leg only).
 We tested but did not report models that used either only accelerometer data and only gyroscope data; in both cases, performance was degraded compared with models that use both types of motion.
@@
 Long-form recordings of motor behavior bring about new research possibilities.
-Although past work using video (Chen et al., 2022; Herzberg et al., 2021; Karasik et al., 2011) produced real-time data, such data were limited to a relatively small part of the infants' day.
-And although survey methods (Franchak, 2019) could capture moments scattered throughout the day, they do not produce real-time data.
+Although past work using video \autocite{Karasik2011,ChenSchneider2022,HerzbergFletcher2021} produced real-time data, such data were limited to a relatively small part of the infants' day.
+And although survey methods \autocite{Survey} could capture moments scattered throughout the day, they do not produce real-time data.
 The combination of real-time, full-day data about infant motor behavior is unprecedented and offers new opportunities for understanding infants' everyday experiences.
@@
 
-In the past decade, developmental scientists have discovered that distributional information about infants' experiences matter for how infants learn (Clerkin et al., 2017; Kachergis et al., 2017; Raz et al., 2019).
+In the past decade, developmental scientists have discovered that distributional information about infants' experiences matter for how infants learn \autocite{ClerkinHart2017,KachergisYu2017,RazAbney2019}.
 Skewed distributions---those that favor the repetition of a small subset of experiences---facilitate learning.
-For example, recordings from wearable head cameras in the home indicate that infants see a small subset of objects with high frequency (Clerkin et al., 2017), but most objects are seen infrequently.
-Heterogeneity is also found in how experiences are distributed in time---burstiness and clustering are seen in infants' daily experiences seeing objects and hearing music (Casillas \& Elliott, 2021; Mendoza \& Fausey, 2022).
+For example, recordings from wearable head cameras in the home indicate that infants see a small subset of objects with high frequency \autocite{ClerkinHart2017}, but most objects are seen infrequently.
+Heterogeneity is also found in how experiences are distributed in time---burstiness and clustering are seen in infants' daily experiences seeing objects and hearing music \autocite{MendozaFausey2022,CasillasElliott2021}.
 Long-form motor recordings provide a novel opportunity to measure the temporal structure in the sequences of body positions seen in Figure \ref{fig:timelines}.
@@
 Moreover, real-time recording of motor behavior provides a way to measure how infant motor behavior links to other types of experiences in the moment in daily life.
-As we reviewed, Malachowski et al.
-(2023) used a combination of LENA® and EMA surveys to find that infants heard less adult speech when restrained in seating devices.
+As we reviewed, \textcite{MalachowskiSalo2023} used a combination of LENA® and EMA surveys to find that infants heard less adult speech when restrained in seating devices.
 Combining LENA® and long-form motor recordings can take this a step further by measuring more precisely how speech and body position co-vary within a day.
-Synchronizing long form motor recordings to other continuous monitors, such as wearable ECG (Wass et al., 2022) can provide information about how infant arousal relates to motor behavior.
-``TotTag''---a wearable device for measuring infant-caregiver distance (Salo et al., 2022)---or caregiver worn activity trackers/physiology monitors could be paired with body position classification to better understand how moment-to-moment interpersonal dynamics relate to infant motor behavior.
+Synchronizing long form motor recordings to other continuous monitors, such as wearable ECG \autocite{WassPhillips2022} can provide information about how infant arousal relates to motor behavior.
+``TotTag''---a wearable device for measuring infant-caregiver distance \autocite{SaloPannuto2022}---or caregiver worn activity trackers/physiology monitors could be paired with body position classification to better understand how moment-to-moment interpersonal dynamics relate to infant motor behavior.
 
 What underpins these novel research applications is that body position classifications can be applied automatically at scale.
-Annotating a video corpus of infant behavior from a moderately sized sample---such as 40 infants recorded for 2 hours each in 2 separate visits (Herzberg et al., 2021)---is incredibly labor intensive.
+Annotating a video corpus of infant behavior from a moderately sized sample---such as 40 infants recorded for 2 hours each in 2 separate visits \autocite{HerzbergFletcher2021}---is incredibly labor intensive.
 Annotating a larger, more representative sample of \emph{hundreds of infants} while simultaneously scoring \emph{full-day data} would be prohibitive.
@@
 Unlike with video, the added cost of more IMU sensor recording time is low, meaning that future studies could sample across multiple days of behavior to better understand intra- versus inter-day variability.
-Prior diary methods show that infants inconsistently display new motor skills on a day-to-day basis (Adolph et al., 2008; Adolph \& Robinson, 2011).
+Prior diary methods show that infants inconsistently display new motor skills on a day-to-day basis \autocite{AdolphRobinson2008,AdolphRobinson2011}.
 Multi-day recordings could further uncover how infants' experiences vary over timescales from seconds to days.
@@
 \hypertarget{refs}{}
-\begin{CSLReferences}{1}{0}
-\leavevmode\vadjust pre{\hypertarget{ref-AdolphRobinson2011}{}}%
-Adolph, K. E., \& Robinson, S. R. (2011). Sampling development. \emph{Journal of Cognition and Development}, \emph{12}, 411--423. \url{https://doi.org/10.1080/15248372.2011.608190}
-
-\leavevmode\vadjust pre{\hypertarget{ref-AdolphRobinson2008}{}}%
-Adolph, K.
-E., Robinson, S.
-R., Young, J.
-W., \& Gill-Alvarez, F.
-(2008).
-What is the shape of developmental change?
-\emph{Psychological Review}, \emph{115}, 527--543.
-\url{https://doi.org/10.1037/0033-295X.115.3.527}
-
-\leavevmode\vadjust pre{\hypertarget{ref-AdolphCTL2014}{}}%
-Adolph, K.
-E., \& Tamis-LeMonda, C.
-S.
-(2014).
-The costs and benefits of development: The transition from crawling to walking.
-\emph{Child Development Perspectives}, \emph{8}, 187--192.
-\url{https://doi.org/10.1111/cdep.12085}
-
-\leavevmode\vadjust pre{\hypertarget{ref-AdolphVereijken1998}{}}%
-Adolph, K. E., Vereijken, B., \& Denny, M. A. (1998). Learning to crawl. \emph{Child Development}, \emph{69}, 1299--1312. \url{https://doi.org/10.1111/j.1467-8624.1998.tb06213.x}
-
-\leavevmode\vadjust pre{\hypertarget{ref-AiraksinenGallen2022}{}}%
-Airaksinen, M., Gallen, A., Kivi, A., Vijayakrishnan, P., Häyrinen, T., Ilén, E., Räsänen, O., Haataja, L.
-M., \& Vanhatalo, S.
-(2022).
-Intelligent wearable allows out-of-the-lab tracking of developing motor abilities in infants.
-\emph{Communications Medicine}, \emph{2}(1).
-\url{https://doi.org/10.1038/s43856-022-00131-6}
-
-\leavevmode\vadjust pre{\hypertarget{ref-AiraksinenRasanen2020}{}}%
-Airaksinen, M., Räsänen, O., Ilén, E., Häyrinen, T., Kivi, A., Marchi, V., Gallen, A., Blom, S., Varhe, A., Kaartinen, N., et al.
-(2020).
-Automatic posture and movement tracking of infants with wearable movement sensors.
-\emph{Scientific Reports}, \emph{10}(1), 1--13.
-
-\leavevmode\vadjust pre{\hypertarget{ref-Arif2015}{}}%
-Arif, M., \& Kattan, A. (2015). Physical activities monitoring using wearable acceleration sensors attached to the body. \emph{{PLoS ONE}}, \emph{10}, e0130851.
-
-\leavevmode\vadjust pre{\hypertarget{ref-ArmstrongCovington2019}{}}%
-Armstrong, B., Covington, L.
-B., Hager, E.
-R., \& Black, M.
-M.
-(2019).
-Objective sleep and physical activity using 24-hour ankle-worn accelerometry among toddlers from low-income families.
-\emph{Sleep Health}, \emph{5}(5), 459--465.
-
-\leavevmode\vadjust pre{\hypertarget{ref-papaja}{}}%
-Aust, F., \& Barth, M. (2022). \emph{{papaja: Prepare reproducible APA journal articles with R Markdown}}. \url{https://github.com/crsh/papaja}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Barbaro2019}{}}%
-Barbaro, K. de. (2019). Automated sensing of daily activity: A new lens into development. \emph{Developmental Psychobiology}, \emph{61}(3), 444--464.
-
-\leavevmode\vadjust pre{\hypertarget{ref-BarbaroFausey2022}{}}%
-Barbaro, K.
-de, \& Fausey, C.
-M.
-(2022).
-Ten lessons about infants' everyday experiences.
-\emph{Current Directions in Psychological Science}, \emph{31}(1), 28--33.
-\url{https://doi.org/10.1177/09637214211059536}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Bergelson2019Input}{}}%
-Bergelson, E., Amatuni, A., Dailey, S., Koorathota, S., \& Tor, S. (2019). Day by day, hour by hour: Naturalistic language input to infants. \emph{Developmental Science}, \emph{22}, e12715.
-
-\leavevmode\vadjust pre{\hypertarget{ref-Bergelson2019Corpus}{}}%
-Bergelson, E., Casillas, M., Soderstrom, M., Seidl, A., Warlaumont, A.
-S., \& Amatuni, A.
-(2019).
-What do {N}orth {A}merican babies hear?
-{A} large-scale cross-corpus analysis.
-\emph{Developmental Science}, \emph{22}, e12724.
-
-\leavevmode\vadjust pre{\hypertarget{ref-Breiman2001}{}}%
-Breiman, L. (2001). Random forests. \emph{Machine Learning}, \emph{45}, 5--32.
-
-\leavevmode\vadjust pre{\hypertarget{ref-BruijnsTruelove2020}{}}%
-Bruijns, B.
-A., Truelove, S., Johnson, A.
-M., Gilliland, J., \& Tucker, P.
-(2020).
-Infants' and toddlers' physical activity and sedentary time as measured by accelerometry: A systematic review and meta-analysis.
-\emph{International Journal of Behavioral Nutrition and Physical Activity}, \emph{17}(1), 14.
-
-\leavevmode\vadjust pre{\hypertarget{ref-CasillasElliott2021}{}}%
-Casillas, M., \& Elliott, M. (2021). Cross-cultural differences in children's object handling at home. \emph{PsyArXiv}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-ChenSchneider2022}{}}%
-Chen, Q., Schneider, J.
-L., West, K.
-L., \& Iverson, J.
-M.
-(2022).
-Infant locomotion shapes proximity to adults during everyday play in the u.s.
-\emph{Infancy}.
-\url{https://doi.org/10.1111/infa.12503}
-
-\leavevmode\vadjust pre{\hypertarget{ref-ClerkinHart2017}{}}%
-Clerkin, E.
-M., Hart, E., Rehg, J.
-M., Yu, C., \& Smith, L.
-B.
-(2017).
-Real-world visual statistics and infants' first-learned object names.
-\emph{Philosophical Transactions of the Royal Society B}, \emph{372}, 20160055.
-
-\leavevmode\vadjust pre{\hypertarget{ref-CliffReilly2009}{}}%
-Cliff, D.
-P., Reilly, J.
-J., \& Okely, A.
-D.
-(2009).
-Methodological considerations in using accelerometers to assess habitual physical activity in children aged 0--5 years.
-\emph{Journal of Science and Medicine in Sport}, \emph{12}(5), 557--567.
-
-\leavevmode\vadjust pre{\hypertarget{ref-CristiaLavechin2020}{}}%
-Cristia, A., Lavechin, M., Scaff, C., Soderstrom, M., Rowland, C., Räsänen, O., Bunce, J., \& Bergelson, E.
-(2020).
-A thorough evaluation of the language environment analysis ({LENA}) system.
-\emph{Behavior Research Methods}, \emph{53}(2), 467--486.
-\url{https://doi.org/10.3758/s13428-020-01393-5}
-
-\leavevmode\vadjust pre{\hypertarget{ref-DanchoVaughan2023}{}}%
-Dancho, M., \& Vaughan, D. (2023). \emph{Timetk: A tool kit for working with time series in {R}}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-DengTrujillo-Priego2019}{}}%
-Deng, W., Trujillo-Priego, I.
-A., \& Smith, B.
-A.
-(2019).
-How many days are necessary to represent an infant{\textquotesingle}s typical daily leg movement behavior using wearable sensors?
-\emph{Physical Therapy}, \emph{99}(6), 730--738.
-\url{https://doi.org/10.1093/ptj/pzz036}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Survey}{}}%
-Franchak, J. M. (2019). Changing opportunities for learning in everyday life: Infant body position over the first year. \emph{Infancy}, \emph{24}, 187--209.
-
-\leavevmode\vadjust pre{\hypertarget{ref-CurrentOpinion}{}}%
-Franchak, J. M. (2020). The ecology of infants' perceptual-motor exploration. \emph{Current Opinion in Psychology}, \emph{32}, 110--114.
-
-\leavevmode\vadjust pre{\hypertarget{ref-Freeplay}{}}%
-Franchak, J. M., Kretch, K. S., \& Adolph, K. E. (2018). See and be seen: Infant-caregiver social looking during locomotor free play. \emph{Developmental Science}, \emph{21}, e12626.
-
-\leavevmode\vadjust pre{\hypertarget{ref-FranchakScott2021}{}}%
-Franchak, J.
-M., Scott, V., \& Luo, C.
-(2021).
-A contactless method for measuring full-day, naturalistic motor behavior using wearable inertial sensors.
-\emph{Frontiers in Psychology}, \emph{12}.
-\url{https://doi.org/10.3389/fpsyg.2021.701343}
-
-\leavevmode\vadjust pre{\hypertarget{ref-GallandTaylor2012}{}}%
-Galland, B.
-C., Taylor, B.
-J., Elder, D.
-E., \& Herbison, P.
-(2012).
-Normal sleep patterns in infants and children: A systematic review of observational studies.
-\emph{Sleep Medicine Reviews}, \emph{16}(3), 213--222.
-\url{https://doi.org/10.1016/j.smrv.2011.06.001}
-
-\leavevmode\vadjust pre{\hypertarget{ref-GreenspanCunha2021}{}}%
-Greenspan, B., Cunha, A.
-B., \& Lobo, M.
-A.
-(2021).
-Design and validation of a smart garment to measure positioning practices of parents with young infants.
-\emph{Infant Behavior and Development}, \emph{62}, 101530.
-
-\leavevmode\vadjust pre{\hypertarget{ref-GrolemundWickham2011}{}}%
-Grolemund, G., \& Wickham, H. (2011). Dates and times made easy with {lubridate}. \emph{Journal of Statistical Software}, \emph{40}(3), 1--25. \url{https://www.jstatsoft.org/v40/i03/}
-
-\leavevmode\vadjust pre{\hypertarget{ref-HagerTilton2017}{}}%
-Hager, E., Tilton, N., Wang, Y., Kapur, N., Arbaiza, R., Merry, B., \& Black, M.
-(2017).
-The home environment and toddler physical activity: An ecological momentary assessment study.
-\emph{Pediatric Obesity}, \emph{12}(1), 1--9.
-
-\leavevmode\vadjust pre{\hypertarget{ref-HerzbergFletcher2021}{}}%
-Herzberg, O., Fletcher, K.
-K., Schatz, J.
-L., Adolph, K.
-E., \& Tamis-LeMonda, C.
-S.
-(2021).
-Infant exuberant object play at home: Immense amounts of time-distributed, variable practice.
-\emph{Child Development}, \emph{93}(1), 150--164.
-
-\leavevmode\vadjust pre{\hypertarget{ref-KachergisYu2017}{}}%
-Kachergis, G., Yu, C., \& Shiffrin, R. M. (2017). A bootstrapping model of frequency and context effects in word learning. \emph{Cognitive Science}, \emph{41}(3), 590--622.
-
-\leavevmode\vadjust pre{\hypertarget{ref-KadookaCaufield2021}{}}%
-Kadooka, K., Caufield, M., Fausey, C.
-M., \& Franchak, J.
-M.
-(2021, April).
-Visuomotor learning opportunities are nested within everyday activities.
-\emph{Paper Presented at the Biennial Meeting of the Society for Research in Child Development}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-KarasikKuchirko2022}{}}%
-Karasik, L.
-B., Kuchirko, Y., Dodojonova, R.
-M., \& Elison, J.
-T.
-(2022).
-Comparison of {U.S.} And {Tajik} infants' time in containment devices.
-\emph{Infant and Child Development}, \emph{31}(4).
-\url{https://doi.org/10.1002/icd.2340}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Karasik2011}{}}%
-Karasik, L.
-B., Tamis-LeMonda, C.
-S., \& Adolph, K.
-E.
-(2011).
-Transition from crawling to walking and infants' actions with objects and people.
-\emph{Child Development}, \emph{82}, 1199--1209.
-\url{https://doi.org/10.1111/j.1467-8624.2011.01595.x}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Karasik2014}{}}%
-Karasik, L.
-B., Tamis-LeMonda, C.
-S., \& Adolph, K.
-E.
-(2014).
-Crawling and walking infants elicit different verbal responses from mothers.
-\emph{Developmental Science}, \emph{17}, 388--395.
-\url{https://doi.org/10.1111/desc.12129}
-
-\leavevmode\vadjust pre{\hypertarget{ref-CWW}{}}%
-Kretch, K.
-S., Franchak, J.
-M., \& Adolph, K.
-E.
-(2014).
-Crawling and walking infants see the world differently.
-\emph{Child Development}, \emph{85}, 1503--1518.
-\url{https://doi.org/10.1111/cdev.12206}
-
-\leavevmode\vadjust pre{\hypertarget{ref-KretchKoziol2022}{}}%
-Kretch, K.
-S., Koziol, N.
-A., Marcinowski, E.
-C., Kane, A.
-E., Inamdar, K., Brown, E.
-D., Bovaird, J.
-A., Harbourne, R.
-T., Hsu, L.-Y., Lobo, M.
-A., \& Dusing, S.
-C.
-(2022).
-Infant posture and caregiver-provided cognitive opportunities in typically developing infants and infants with motor delay.
-\emph{Developmental Psychobiology}, \emph{64}(1).
-\url{https://doi.org/10.1002/dev.22233}
-
-\leavevmode\vadjust pre{\hypertarget{ref-tidymodels}{}}%
-Kuhn, M., \& Wickham, H. (2020). Tidymodels: A collection of packages for modeling and machine learning using tidyverse principles. \emph{Boston, MA, USA}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-KwonZavos2019}{}}%
-Kwon, S., Zavos, P., Nickele, K., Sugianto, A., \& Albert, M.
-V.
-(2019).
-Hip and wrist-worn accelerometer data analysis for toddler activities.
-\emph{International Journal of Environmental Research and Public Health}, \emph{16}(14), 2598.
-
-\leavevmode\vadjust pre{\hypertarget{ref-LandisKoch1977}{}}%
-Landis, J. R., \& Koch, G. G. (1977). The measurement of observer agreement for categorical data. \emph{Biometrics}, \emph{33}(1), 159. \url{https://doi.org/10.2307/2529310}
-
-\leavevmode\vadjust pre{\hypertarget{ref-LiawWiener2002}{}}%
-Liaw, A., Wiener, M., et al. (2002). Classification and regression by randomForest. \emph{R News}, \emph{2}(3), 18--22.
-
-\leavevmode\vadjust pre{\hypertarget{ref-LoboHall2019}{}}%
-Lobo, M.
-A., Hall, M.
-L., Greenspan, B., Rohloff, P., Prosser, L.
-A., \& Smith, B.
-A.
-(2019).
-Wearables for pediatric rehabilitation: How to optimally design and use products to meet the needs of users.
-\emph{Physical Therapy}, \emph{99}(6), 647--657.
-
-\leavevmode\vadjust pre{\hypertarget{ref-LuoFranchak2020}{}}%
-Luo, C., \& Franchak, J. M. (2020). Head and body structure infants' visual experiences during mobile, naturalistic play. \emph{{PLoS ONE}}, \emph{15}, e0242009.
-
-\leavevmode\vadjust pre{\hypertarget{ref-MajnemerBarr2005}{}}%
-Majnemer, A., \& Barr, R. G. (2005). Influence of supine sleep positioning on early motor milestone acquisition. \emph{Developmental Medicine and Child Neurology}, \emph{47}, 370--376.
-
-\leavevmode\vadjust pre{\hypertarget{ref-MalachowskiSalo2023}{}}%
-Malachowski, L.
-G., Salo, V.
-C., Needham, A.
-W., \& Humphreys, K.
-L.
-(2023).
-Infant placement and language exposure in daily life.
-\emph{Infant and Child Development}.
-\url{https://doi.org/10.1002/icd.2405}
-
-\leavevmode\vadjust pre{\hypertarget{ref-MendozaFausey2021}{}}%
-Mendoza, J. K., \& Fausey, C. M. (2021). Everyday music in infancy. \emph{Developmental Science}, \emph{24}(6). \url{https://doi.org/10.1111/desc.13122}
-
-\leavevmode\vadjust pre{\hypertarget{ref-MendozaFausey2022}{}}%
-Mendoza, J.
-K., \& Fausey, C.
-M.
-(2022).
-Everyday parameters for episode-to-episode dynamics in the daily music of infancy.
-\emph{Cognitive Science}, \emph{46}(8).
-\url{https://doi.org/10.1111/cogs.13178}
-
-\leavevmode\vadjust pre{\hypertarget{ref-MichelettiYao2022}{}}%
-Micheletti, M., Yao, X., Johnson, M., \& Barbaro, K.
-de.
-(2022).
-Validating a model to detect infant crying from naturalistic audio.
-\emph{Behavior Research Methods}.
-\url{https://doi.org/10.3758/s13428-022-01961-x}
-
-\leavevmode\vadjust pre{\hypertarget{ref-NamPark2013}{}}%
-Nam, Y., \& Park, J.
-W.
-(2013).
-Child activity recognition based on cooperative fusion model of a triaxial accelerometer and a barometric pressure sensor.
-\emph{IEEE Journal of Biomedical and Health Informatics}, \emph{17}, 420--426.
-
-\leavevmode\vadjust pre{\hypertarget{ref-PatelShi2019}{}}%
-Patel, P., Shi, Y., Hajiaghajani, F., Biswas, S., \& Lee, M.-H.
-(2019).
-A novel two-body sensor system to study spontaneous movements in infants during caregiver physical contact.
-\emph{Infant Behavior and Development}, \emph{57}, 101383.
-\url{https://doi.org/10.1016/j.infbeh.2019.101383}
-
-\leavevmode\vadjust pre{\hypertarget{ref-PerryPrince2018}{}}%
-Perry, L.
-K., Prince, E.
-B., Valtierra, A.
-M., Rivero-Fernandez, C., Ullery, M.
-A., Katz, L.
-F., Laursen, B., \& Messinger, D.
-S.
-(2018).
-A year in words: The dynamics and consequences of language experiences in an intervention classroom.
-\emph{{PLOS} {ONE}}, \emph{13}(7), e0199893.
-\url{https://doi.org/10.1371/journal.pone.0199893}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Preece2009}{}}%
-Preece, S.
-J., Goulermas, J.
-Y., Kenney, L.
-P.
-J., \& Howard, D.
-(2009).
-A comparison of feature extraction methods for the classification of dynamic activities from accelerometer data.
-\emph{IEEE Transactions on Biomedical Engineering}, \emph{56}, 871--879.
-
-\leavevmode\vadjust pre{\hypertarget{ref-R42}{}}%
-R Core Team. (2021). \emph{R: A language and environment for statistical computing}. {R Foundation for Statistical Computing}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-RasanenSeshadri2020}{}}%
-Räsänen, O., Seshadri, S., Lavechin, M., Cristia, A., \& Casillas, M.
-(2020).
-{ALICE}: An open-source tool for automatic measurement of phoneme, syllable, and word counts from child-centered daylong recordings.
-\emph{Behavior Research Methods}, \emph{53}(2), 818--835.
-\url{https://doi.org/10.3758/s13428-020-01460-x}
-
-\leavevmode\vadjust pre{\hypertarget{ref-RazAbney2019}{}}%
-Raz, H.
-K., Abney, D.
-H., Crandall, D., Yu, C., \& Smith, L.
-B.
-(2019).
-How do infants start learning object names in a sea of clutter?
-\emph{Annual Conference of the Cognitive Science Society}, 521--526.
-
-\leavevmode\vadjust pre{\hypertarget{ref-RenDing2016}{}}%
-Ren, X., Ding, W., Crouter, S.
-E., Mu, Y., \& Xie, R.
-(2016).
-Activity recognition and intensity estimation in youth from accelerometer data aided by machine learning.
-\emph{Applied Intelligence}, \emph{45}(2), 512--529.
-
-\leavevmode\vadjust pre{\hypertarget{ref-SaloPannuto2022}{}}%
-Salo, V.
-C., Pannuto, P., Hedgecock, W., Biri, A., Russo, D.
-A., Piersiak, H.
-A., \& Humphreys, K.
-L.
-(2022).
-Measuring naturalistic proximity as a window into caregiver--child interaction patterns.
-\emph{Behavior Research Methods}, \emph{54}(4), 1580--1594.
-\url{https://doi.org/10.3758/s13428-021-01681-8}
-
-\leavevmode\vadjust pre{\hypertarget{ref-SchnellerBentsen2017}{}}%
-Schneller, M.
-B., Bentsen, P., Nielsen, G., Brønd, J.
-C., Ried-Larsen, M., Mygind, E., \& Schipperijn, J.
-(2017).
-Measuring children's physical activity.
-\emph{Medicine \& Science in Sports \& Exercise}, \emph{49}(6), 1261--1269.
-\url{https://doi.org/10.1249/mss.0000000000001222}
-
-\leavevmode\vadjust pre{\hypertarget{ref-SoskaAdolph2014}{}}%
-Soska, K. C., \& Adolph, K. E. (2014). Postural position constrains multimodal object exploration in infants. \emph{Infancy}, \emph{19}, 138--161. \url{https://doi.org/10.1111/infa.12039}
-
-\leavevmode\vadjust pre{\hypertarget{ref-StewartNarayanan2018}{}}%
-Stewart, T., Narayanan, A., Hedayatrad, L., Neville, J., Mackay, L., \& Duncan, S.
-(2018).
-A dual-accelerometer system for classifying physical activity in children and adults.
-\emph{Medicine and Science in Sports and Exercise}, \emph{50}(12), 2595--2602.
-
-\leavevmode\vadjust pre{\hypertarget{ref-Tamis-LeMondaCustode2018}{}}%
-Tamis-LeMonda, C.
-S., Custode, S., Kuchirko, Y., Escobar, K., \& Lo, T.
-(2018).
-Routine language: Speech directed to infants during home activities.
-\emph{Child Development}, \emph{90}(6), 2135--2152.
-\url{https://doi.org/10.1111/cdev.13089}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Thurman2017}{}}%
-Thurman, S. L., \& Corbetta, D. (2017). Spatial exploration and changes in infant-mother dyads around transitions in infant locomotion. \emph{Developmental Psychology}, \emph{53}, 1207--1221.
-
-\leavevmode\vadjust pre{\hypertarget{ref-WarlaumontSobowale2021}{}}%
-Warlaumont, A.
-S., Sobowale, K., \& Fausey, C.
-M.
-(2021).
-Daylong mobile audio recordings reveal multitimescale dynamics in infants' vocal productions and auditory experiences.
-\emph{Current Directions in Psychological Science}, \emph{31}(1), 12--19.
-\url{https://doi.org/10.1177/09637214211058166}
-
-\leavevmode\vadjust pre{\hypertarget{ref-WassPhillips2022}{}}%
-Wass, S., Phillips, E., Smith, C., Fatimehin, E.
-O., \& Goupil, L.
-(2022).
-Vocal communication is tied to interpersonal arousal coupling in caregiver-infant dyads.
-\emph{{eLife}}, \emph{11}.
-\url{https://doi.org/10.7554/elife.77399}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Weisleder2013}{}}%
-Weisleder, A., \& Fernald, A. (2013). Talking to children matters: Early language experience strengthens processing and builds vocabulary. \emph{Psychological Science}, \emph{24}, 2143--2152.
-
-\leavevmode\vadjust pre{\hypertarget{ref-WestIverson2021}{}}%
-West, K. L., \& Iverson, J. M. (2021). Communication changes when infants begin to walk. \emph{Developmental Science}, \emph{24}(5). \url{https://doi.org/10.1111/desc.13102}
-
-\leavevmode\vadjust pre{\hypertarget{ref-YaoPlotz2019}{}}%
-Yao, X., Plötz, T., Johnson, M., \& Barbaro, K.
-de.
-(2019).
-Automated detection of infant holding using wearable sensing: Implications for developmental science and intervention.
-\emph{Proceedings of the ACM on Interactive, Mobile, Wearable and Ubiquitous Technologies}, \emph{3}(2), 1--17.
-
+\begin{CSLReferences}{0}{0}
 \end{CSLReferences}
 
+\printbibliography
+
 \end{document}

### 2023-06-16 group_split_comparison/group_split_sensor_reduce.R

@@ -1,0 +1,82 @@
+# This script calculates the first set of results in the paper based on the proximal period
+# compiled_data_lite.RData contains the windowed motion feature data for all sessions
+# The data are partitioned into training and testing sets
+# Group models and individual ("split") models are trained, and then predictions are made on the testing set
+# Agreement metrics are saved to /data to be reported in the RMarkdown manuscript
+# Depending on hardware, this script can take several hours to run
+
+library(here)
+library(randomForest)
+library(cvms)
+library(caret)
+library(tidyverse)
+library(rstatix)
+
+#LOAD DATA
+load( here("group_split_comparison","compiled_data_lite.RData"))
+ntree = 550
+mtry = 44
+not_all_na <- function(x) any(!is.na(x))
+
+total_time <- session %>% mutate(coded_time = end_time_coded - start_time_coded) %>% arrange(coded_time)
+total_samples <- slide_filt %>% drop_na(code) %>% group_by(id) %>% summarize(n_samples = n()) %>% arrange(n_samples)
+
+training <- slide_filt %>% group_by(id, code) %>% slice_head(prop = .75) %>% ungroup 
+testing <- slide_filt %>% group_by(id, code) %>% slice_tail(prop = .25) %>% ungroup
+
+group_model <- function(temp_id, training, ntree = 50) {
+  print(temp_id)
+  temp_group_training <- training %>% 
+    filter(id != temp_id) %>% 
+    select_if(not_all_na) %>% 
+    mutate(code = fct_drop(code)) %>% 
+    select(-id)
+  group <- randomForest(code ~ ., data = temp_group_training, 
+                        localImp = TRUE, proximity = FALSE, ntree = ntree, mtry = mtry)
+}
+
+split_model <- function(temp_id, training, ntree = 50) {
+  temp_split_training <- training %>% 
+    filter(id == temp_id) %>% 
+    select_if(not_all_na) %>% 
+    mutate(code = fct_drop(code)) %>% 
+    select(-id)
+  split <- randomForest(code ~ ., data = temp_split_training, 
+                        localImp = TRUE, proximity = FALSE, ntree = ntree, mtry = mtry)
+}
+
+ids <- unique(training$id)
+group_mods <-map(ids, ~group_model(.x, training)) %>% set_names(ids)
[…]

### 2023-06-16 group_split_comparison/group_split_sensor_reduce.R
sensor reduce

@@ -12,17 +12,33 @@ library(caret)
 library(tidyverse)
 library(rstatix)
 
+feature_list <- read_csv("group_split_comparison/feature_names.csv")
+
+ reduce_type <- "left_side"
+# reduce_type <- "right_side"
+# reduce_type <- "left_ankle"
+# reduce_type <- "right_ankle"
+# reduce_type <- "left_hip"
+# reduce_type <- "right_hip"
+
 #LOAD DATA
 load( here("group_split_comparison","compiled_data_lite.RData"))
 ntree = 550
 mtry = 44
 not_all_na <- function(x) any(!is.na(x))
 
-total_time <- session %>% mutate(coded_time = end_time_coded - start_time_coded) %>% arrange(coded_time)
-total_samples <- slide_filt %>% drop_na(code) %>% group_by(id) %>% summarize(n_samples = n()) %>% arrange(n_samples)
-
 training <- slide_filt %>% group_by(id, code) %>% slice_head(prop = .75) %>% ungroup 
 testing <- slide_filt %>% group_by(id, code) %>% slice_tail(prop = .25) %>% ungroup
+
+
+if (reduce_type == "left_side")   reduce_names <- feature_list %>% filter(left_side == T) %>% pull(names)
+if (reduce_type == "right_side")  reduce_names <- feature_list %>% filter(right_side == T) %>% pull(names)
+if (reduce_type == "left_ankle")  reduce_names <- feature_list %>% filter(left_ankle_only == T) %>% pull(names)
+if (reduce_type == "right_ankle") reduce_names <- feature_list %>% filter(right_ankle_only == T) %>% pull(names)
+if (reduce_type == "left_hip")    reduce_names <- feature_list %>% filter(left_hip_only == T) %>% pull(names)
+if (reduce_type == "right_hip")   reduce_names <- feature_list %>% filter(right_hip_only == T) %>% pull(names)
+      
+training <- training %>% select(id, all_of(reduce_names))
 
 group_model <- function(temp_id, training, ntree = 50) {
   print(temp_id)
@@ -67,7 +83,7 @@ res_split %>% get_summary_stats(-ids)
 res_split %>% arrange(`Overall Accuracy`)
 
 ds <- bind_rows(res_split, res_group) 
-write_csv(ds, file = "data/group_split_metrics.csv")
+write_csv(ds, file = str_glue("data/reduce/reduce_metrics{reduce_type}.csv"))
 
[…]

### 2023-06-16 manuscript.Rmd
made the table

@@
 Table \@ref(tab:metrics-table) shows the overall sensitivity and PPV across classes.
+
+### Sensor Configuration
+<!-- data/reduce, duplicates of the group_split metrics but performed on reductions of sensor features from group_split_sensor_reduce.R -->
+
+```{r metrics-reduce, cache=TRUE, cache.extra=tools::md5sum("data/group_split_metrics.csv")}
+all_files = list.files("data/reduce", full.names = T, pattern = ".csv")
+overall_files = all_files %>% discard(str_detect(., pattern = "class"))
+sensor_comp <- read_csv(overall_files, id = "type")
+sensor_comp <- sensor_comp %>% mutate(type = str_remove(type, "data/reduce/reduce_metrics_"),
+                              type = str_remove(type, ".csv"),
+                              type = factor(type,
+                              levels = c("all", "left_side", "right_side", "hips", "ankles",
+                                         "left_hip","left_ankle", "right_hip", "right_ankle"),
+                              labels = c("All", "Left Hip/Ankle", "Right Hip/Ankle", "Both Hips", "Both Ankles", 
+                                         "Left Hip", "Left Ankle", "Right Hip", "Right Ankle")),
+                              Model = factor(model, levels = c("group", "split", labels = c("Group", "Individual"))))
+
+metrics_sensor <- sensor_comp %>% group_by(Model, type) %>% 
+  get_summary_stats(`Overall Accuracy`) %>% 
+  select(-(variable:max), -(q1:mad), -se, -ci) %>% 
+  pivot_wider(names_from = Model, values_from = c("median", "mean", "sd")) %>% 
+  relocate(mean_group, .after = median_group) %>% 
+  relocate(sd_group, .after = mean_group) %>% 
+  mutate(across(-type, ~ apa_num(.x, digits = 3)))
+
+apa_table(metrics_sensor, col.names = c("Sensors", "Median", "Mean", "SD", "Median", "Mean", "SD"),    col_spanners = list("Group" = c(2, 4), "Individual" = c(5, 7)), midrules = c(1,5),
+          caption = "Median, mean, and SD of overall accuracy calculated with different sets of sensor features.
+The top row shows performance using all features calculated from the four sensors.
+Rows 2-5 show accuracy using pairs of sensors (left hip and ankle, right hip and ankle, left and right hip, left and right ankle), and rows 6-9 show accuracy using each individual sensor.")
+```
+
+\renewcommand{\arraystretch}{.75}
 

### 2023-06-16 manuscript.tex
made the table

@@
 
+\hypertarget{sensor-configuration}{%
+\subsubsection{Sensor Configuration}\label{sensor-configuration}}
+
+\begin{table}[tbp]
+
+\begin{center}
+\begin{threeparttable}
+
+\caption{\label{tab:metrics-reduce}Median, mean, and SD of overall accuracy calculated with different sets of sensor features.
+The top row shows performance using all features calculated from the four sensors.
+Rows 2-5 show accuracy using pairs of sensors (left hip and ankle, right hip and ankle, left and right hip, left and right ankle), and rows 6-9 show accuracy using each individual sensor.}
+
+\begin{tabular}{lllllll}
+\toprule
+ & \multicolumn{3}{c}{Group} & \multicolumn{3}{c}{Individual} \\
+\cmidrule(r){2-4} \cmidrule(r){5-7}
+Sensors & Median & Mean & SD & Median & Mean & SD\\
+\midrule
+All & 0.894 & 0.846 & 0.131 & 0.933 & 0.916 & 0.072\\ \midrule
+Left Hip/Ankle & 0.891 & 0.824 & 0.177 & 0.915 & 0.901 & 0.071\\
+Right Hip/Ankle & 0.842 & 0.800 & 0.167 & 0.897 & 0.898 & 0.086\\
+Both Hips & 0.859 & 0.791 & 0.195 & 0.907 & 0.898 & 0.072\\
+Both Ankles & 0.797 & 0.719 & 0.196 & 0.864 & 0.844 & 0.108\\ \midrule
+Left Hip & 0.842 & 0.762 & 0.208 & 0.889 & 0.872 & 0.077\\
+Left Ankle & 0.717 & 0.667 & 0.208 & 0.851 & 0.825 & 0.117\\
+Right Hip & 0.846 & 0.772 & 0.178 & 0.896 & 0.876 & 0.092\\
+Right Ankle & 0.704 & 0.683 & 0.160 & 0.830 & 0.819 & 0.114\\
+\bottomrule
+\end{tabular}
+
+\end{threeparttable}
+\end{center}
+
+\end{table}
+
+\renewcommand{\arraystretch}{.75}
+
 \hypertarget{measure-the-distal-accuracy-of-body-position-classification-models}{%

### 2023-06-16 group_split_comparison/group_split_sensor_reduce.R

@@ -14,7 +14,9 @@ library(rstatix)
 
 feature_list <- read_csv("group_split_comparison/feature_names.csv")
 
- reduce_type <- "left_side"
+ reduce_type <- "hips"
+# reduce_type <- "ankles"
+# reduce_type <- "left_side"
 # reduce_type <- "right_side"
 # reduce_type <- "left_ankle"
 # reduce_type <- "right_ankle"
@@ -30,7 +32,8 @@ not_all_na <- function(x) any(!is.na(x))
 training <- slide_filt %>% group_by(id, code) %>% slice_head(prop = .75) %>% ungroup 
 testing <- slide_filt %>% group_by(id, code) %>% slice_tail(prop = .25) %>% ungroup
 
-
+if (reduce_type == "hips")        reduce_names <- feature_list %>% filter(hips == T) %>% pull(names)
+if (reduce_type == "ankles")      reduce_names <- feature_list %>% filter(ankles == T) %>% pull(names)
 if (reduce_type == "left_side")   reduce_names <- feature_list %>% filter(left_side == T) %>% pull(names)
 if (reduce_type == "right_side")  reduce_names <- feature_list %>% filter(right_side == T) %>% pull(names)
 if (reduce_type == "left_ankle")  reduce_names <- feature_list %>% filter(left_ankle_only == T) %>% pull(names)
@@ -95,4 +98,7 @@ res_split_class <-  map2(split_mods, ids, ~metrics(.x, filter(testing, id == .y)
              mutate(id = .y, model = "split"))
 
 ds_class <- bind_rows(res_split_class, res_group_class) 
-write_csv(ds_class, file =  str_glue("data/reduce/reduce_metrics{reduce_type}_class.csv"))
+write_csv(ds_class, file =  str_glue("data/reduce/reduce_metrics_{reduce_type}_class.csv"))
+
+
+

### 2023-06-16 manuscript.Rmd
sensor comparison done

--- 2023-06-16 manuscript.Rmd
@@
 
+<!-- data/reduce, duplicates of the group_split metrics but performed on reductions of sensor features from group_split_sensor_reduce.R -->
+
+```{r metrics-reduce, cache=TRUE, cache.extra=tools::md5sum("data/reduce/reduce_metrics_right_side.csv")}
+all_files = list.files("data/reduce", full.names = T, pattern = ".csv")
+overall_files = all_files %>% discard(str_detect(., pattern = "class"))
+sensor_comp <- read_csv(overall_files, id = "type")
+sensor_comp <- sensor_comp %>% mutate(type = str_remove(type, "data/reduce/reduce_metrics_"),
+                              type = str_remove(type, ".csv"),
+                              type = factor(type,
+                              levels = c("all", "left_side", "right_side", "hips", "ankles",
+                                         "left_hip","left_ankle", "right_hip", "right_ankle"),
+                              labels = c("All", "Left Hip/Ankle", "Right Hip/Ankle", "Both Hips", "Both Ankles", 
+                                         "Left Hip", "Left Ankle", "Right Hip", "Right Ankle")),
+                              Model = factor(model, levels = c("group", "split", labels = c("Group", "Individual"))))
+
+metrics_sensor <- sensor_comp %>% group_by(Model, type) %>% 
+  get_summary_stats(`Overall Accuracy`) %>% 
+  select(-(variable:max), -(q1:mad), -se, -ci) %>% 
+  pivot_wider(names_from = Model, values_from = c("median", "mean", "sd")) %>% 
+  relocate(mean_group, .after = median_group) %>% 
+  relocate(sd_group, .after = mean_group) %>% 
+  mutate(across(-type, ~ apa_num(.x, digits = 3)))
+
+apa_table(metrics_sensor, col.names = c("Sensors", "Median", "Mean", "SD", "Median", "Mean", "SD"),    col_spanners = list("Group" = c(2, 4), "Individual" = c(5, 7)), midrules = c(1,5),
+          caption = "Median, mean, and SD of overall accuracy calculated with different sets of sensor features.
+The top row shows performance using all features calculated from the four sensors.
+Rows 2-5 show accuracy using pairs of sensors (left hip and ankle, right hip and ankle, left and right hip, left and right ankle), and rows 6-9 show accuracy using each individual sensor.")
+metrics_sensor <- metrics_sensor %>% column_to_rownames("type")
+```
+
 ### Overall Accuracy
@@
 Possibly, individual models accounted for idiosyncrasies in behavior or inconsistencies in sensor placement for those sessions that group models could not generalize to.
+
+Similar to past work [@AiraksinenRasanen2020], overall accuracy decreased when fewer sensors were used.
+Table \@ref(tab:metrics-reduce) compared overall accuracy for all four sensors (top row), pairs of 2 sensors (rows 2-5), and single sensors (rows 6-9).
+The highest accuracy was observed when using all four sensors (group model *M* = `r metrics_sensor["All", "mean_group"]`; individual model *M* = `r metrics_sensor["All", "mean_split"]`), and the lowest when using only a single sensor (left ankle group model *M* = `r metrics_sensor["Left Ankle", "mean_group"]`; right ankle individual model *M* = `r metrics_sensor["Right Ankle", "mean_split"]`).
+Some sensor pairs showed comparable accuracy to using all four sensors: Left ankle and hip has group model accuracy of *M* = `r metrics_sensor["Left Hip/Ankle", "mean_group"]` and individual model accuracy of *M* = `r metrics_sensor["Left Hip/Ankle", "mean_split"]`, only 2.2% and 1.5% worse than using all four sensors.
+Other pairings were less accurate; notably, using both ankles resulted in group model accuracy of *M* = `r metrics_sensor["Both Ankles", "mean_group"]` and individual model accuracy of *M* = `r metrics_sensor["Both Ankles", "mean_split"]`, 12.7% and 7.2% worse than using all four sensors.
 
@@
 
-Beyond group and individual models, there were many degrees of freedom in our choice of how best to model body position.
+As in past work [@AiraksinenRasanen2020], performance was best when using all four sensors and degraded when using any subset of two (i.e., hips only, left leg only).
+Small decrements and accuracy were seen for some pairs (particularly when using the left ankle and hip), but decrements for other pairs (both ankles) or individual sensors were substantial.
+Using a hip and ankle pair instead of four sensors would be more cost effective without a large decrease in accuracy, but in some applications the increased accuracy might be worth the cost.
+Beyond group and individual models and sensor pairings, there were many degrees of freedom in our choice of how best to model body position.
 For the sake of brevity, we chose to report the best approach, not every possible variation in modeling.
 However, our openly shared data and code allow researcher to experiment with different sets of motion features, different sets of training data, different machine learning algorithms, and different hyperparameter values.
-As in past work [@AiraksinenRasanen2020], performance was best when using all four sensors and degraded when using any subset of two (i.e., hips only, left leg only).
 We tested but did not report models that used either only accelerometer data and only gyroscope data; in both cases, performance was degraded compared with models that use both types of motion.

### 2023-06-16 manuscript.tex
sensor comparison done

@@
 
-\hypertarget{overall-accuracy}{%
-\subsubsection{Overall Accuracy}\label{overall-accuracy}}
-
-Overall accuracy (Figure \ref{fig:metrics}A) represents the proportion of 4-s windows in the testing set in which the model prediction matched the human annotation of body position.
-Overall accuracy for group models (\emph{M} = 0.85) was slightly lower than accuracy for individual models (\emph{M} = 0.92).
-Although overall accuracy from our semi-supervised, in-home data collection did not match the near-perfect accuracy (.95-.98) found in prior in-lab studies \autocite{FranchakScott2021,AiraksinenRasanen2020}, both models approached the level of agreement found between two human coders (\emph{M} = .906).
-Most likely, lower accuracy in the current study results from the more variable and complex behavior observed in a a semi-supervised setting rather than from a difference in the quality of the classification model; although the first part of the proximal period was guided by the experimenter during a brief phone call, the remainder of the proximal period included natural behavior.
-Visual inspection of Figure \ref{fig:metrics}A shows that accuracy values were heavily skewed, with many approaching perfect accuracy but a few sessions with very poor accuracy.
-Looking at the median performance suggests that the difference between models was not considerable for the typical participant (group median accuracy = 0.89; individual median accuracy = 0.93).
-Indeed, it is notable that the worst-case accuracy from group models (minimum = 0.50) was far lower than in individual models (minimum = 0.73).
-Possibly, individual models accounted for idiosyncrasies in behavior or inconsistencies in sensor placement for those sessions that group models could not generalize to.
-
-\hypertarget{cohens-kappa}{%
-\subsubsection{Cohen's Kappa}\label{cohens-kappa}}
-
-Strong overall accuracy can be misleading when the relative frequency of different classes is unbalanced.
-For example, most infants spent more time sitting than prone, so good performance for sitting classification could mask poorer performance in prone classification.
-Accordingly, we report Cohen's kappa, a commonly-used metric that penalizes missing rare events (Figure \ref{fig:metrics}B), and we provide classification metrics for each individual body position (Table \ref{tab:metrics-by-class}) to account for imbalance in body position rates within and between individuals.
-Similar to overall accuracy, kappa values were strong for both model types with group kappas (\emph{M} = 0.75) somewhat worse compared with individual kappas (\emph{M} = 0.82).
-Guidelines for interpreting kappa statistics \autocite{LandisKoch1977} consider 0.81--1.00 ``Almost Perfect,'' 0.61--0.80 ``Substantial,'' 0.41--0.60 ``Moderate,'' 0.21-- 0.40 ``Fair,'' and 0--0.20 ``Slight to Poor'', indicating that agreement for most group and individual model predictions fell in the Substantial to Almost Perfect range.
-
-As in past work \autocite{FranchakScott2021,AiraksinenRasanen2020}, all body positions were accurately classified even though performance varied somewhat between positions.
-As Table \ref{tab:metrics-by-class} shows, mean kappa statistics were strongest for prone (group \emph{M} = 0.860, individual \emph{M} = 0.841) and supine (group \emph{M} = 0.764, individual \emph{M} = 0.912).
-Sitting performance fell in the middle, and was considerably worse for group models than individual models (group \emph{M} = 0.702, individual \emph{M} = 0.887).
-Held (group \emph{M} = 0.726, individual \emph{M} = 0.727) and upright (group \emph{M} = 0.673, individual \emph{M} = 0.741) performance was the least accurate, however, average performance was still within the ``Substantial'' range.
-
-\hypertarget{sensitivity-and-positive-predictive-value}{%
-\subsubsection{Sensitivity and Positive Predictive Value}\label{sensitivity-and-positive-predictive-value}}
-
-Beyond accuracy, it is important to establish that classification is neither too sensitive nor too conservative in choosing one position versus another.
-Sensitivity refers to the proportion of events of a given position that were correctly identified (e.g., out of 100 human-coded sitting windows, how many of those windows did the model correctly classify as sitting?).
-High sensitivity means that events are unlikely to be missed.
-In contrast, positive predictive value (PPV) refers to the proportion of events classified for a given position that actually belonged to that position (e.g., if the model said a baby was upright during 100 windows, how many of those windows were indeed human-coded upright events?).
-High PPV means that we can be confident in the event label.
-Table \ref{tab:metrics-by-class} shows the sensitivity and PPV by body position class for group and individual models.
-For group models, sensitivity and PPV were similar: They were highest for supine, prone, and sitting (the most accurately identified class) and lowest for upright and held.
-Results for individual models were similar to group models, with the exception of a somewhat lower sensitivity score for held.
-Overall, the results suggest a reasonable balance between sensitivity and PPV among different body position classes for both model types.
-Table \ref{tab:metrics-table} shows the overall sensitivity and PPV across classes.
-
-\hypertarget{sensor-configuration}{%
-\subsubsection{Sensor Configuration}\label{sensor-configuration}}
-
 \begin{table}[tbp]
@@
 
-\renewcommand{\arraystretch}{.75}
+\hypertarget{overall-accuracy}{%
+\subsubsection{Overall Accuracy}\label{overall-accuracy}}
+
+Overall accuracy (Figure \ref{fig:metrics}A) represents the proportion of 4-s windows in the testing set in which the model prediction matched the human annotation of body position.
+Overall accuracy for group models (\emph{M} = 0.85) was slightly lower than accuracy for individual models (\emph{M} = 0.92).
+Although overall accuracy from our semi-supervised, in-home data collection did not match the near-perfect accuracy (.95-.98) found in prior in-lab studies \autocite{FranchakScott2021,AiraksinenRasanen2020}, both models approached the level of agreement found between two human coders (\emph{M} = .906).
+Most likely, lower accuracy in the current study results from the more variable and complex behavior observed in a a semi-supervised setting rather than from a difference in the quality of the classification model; although the first part of the proximal period was guided by the experimenter during a brief phone call, the remainder of the proximal period included natural behavior.
+Visual inspection of Figure \ref{fig:metrics}A shows that accuracy values were heavily skewed, with many approaching perfect accuracy but a few sessions with very poor accuracy.
+Looking at the median performance suggests that the difference between models was not considerable for the typical participant (group median accuracy = 0.89; individual median accuracy = 0.93).
+Indeed, it is notable that the worst-case accuracy from group models (minimum = 0.50) was far lower than in individual models (minimum = 0.73).
+Possibly, individual models accounted for idiosyncrasies in behavior or inconsistencies in sensor placement for those sessions that group models could not generalize to.
+
+Similar to past work \autocite{AiraksinenRasanen2020}, overall accuracy decreased when fewer sensors were used.
+Table \ref{tab:metrics-reduce} compared overall accuracy for all four sensors (top row), pairs of 2 sensors (rows 2-5), and single sensors (rows 6-9).
+The highest accuracy was observed when using all four sensors (group model \emph{M} = 0.846; individual model \emph{M} = 0.916), and the lowest when using only a single sensor (left ankle group model \emph{M} = 0.667; right ankle individual model \emph{M} = 0.819).
+Some sensor pairs showed comparable accuracy to using all four sensors: Left ankle and hip has group model accuracy of \emph{M} = 0.824 and individual model accuracy of \emph{M} = 0.901, only 2.2\% and 1.5\% worse than using all four sensors.
+Other pairings were less accurate; notably, using both ankles resulted in group model accuracy of \emph{M} = 0.719 and individual model accuracy of \emph{M} = 0.844, 12.7\% and 7.2\% worse than using all four sensors.
+
+\hypertarget{cohens-kappa}{%
+\subsubsection{Cohen's Kappa}\label{cohens-kappa}}
+
+Strong overall accuracy can be misleading when the relative frequency of different classes is unbalanced.
+For example, most infants spent more time sitting than prone, so good performance for sitting classification could mask poorer performance in prone classification.
+Accordingly, we report Cohen's kappa, a commonly-used metric that penalizes missing rare events (Figure \ref{fig:metrics}B), and we provide classification metrics for each individual body position (Table \ref{tab:metrics-by-class}) to account for imbalance in body position rates within and between individuals.
+Similar to overall accuracy, kappa values were strong for both model types with group kappas (\emph{M} = 0.75) somewhat worse compared with individual kappas (\emph{M} = 0.82).
+Guidelines for interpreting kappa statistics \autocite{LandisKoch1977} consider 0.81--1.00 ``Almost Perfect,'' 0.61--0.80 ``Substantial,'' 0.41--0.60 ``Moderate,'' 0.21-- 0.40 ``Fair,'' and 0--0.20 ``Slight to Poor'', indicating that agreement for most group and individual model predictions fell in the Substantial to Almost Perfect range.
+
+As in past work \autocite{FranchakScott2021,AiraksinenRasanen2020}, all body positions were accurately classified even though performance varied somewhat between positions.
+As Table \ref{tab:metrics-by-class} shows, mean kappa statistics were strongest for prone (group \emph{M} = 0.860, individual \emph{M} = 0.841) and supine (group \emph{M} = 0.764, individual \emph{M} = 0.912).
+Sitting performance fell in the middle, and was considerably worse for group models than individual models (group \emph{M} = 0.702, individual \emph{M} = 0.887).
+Held (group \emph{M} = 0.726, individual \emph{M} = 0.727) and upright (group \emph{M} = 0.673, individual \emph{M} = 0.741) performance was the least accurate, however, average performance was still within the ``Substantial'' range.
+
+\hypertarget{sensitivity-and-positive-predictive-value}{%
+\subsubsection{Sensitivity and Positive Predictive Value}\label{sensitivity-and-positive-predictive-value}}
+
+Beyond accuracy, it is important to establish that classification is neither too sensitive nor too conservative in choosing one position versus another.
+Sensitivity refers to the proportion of events of a given position that were correctly identified (e.g., out of 100 human-coded sitting windows, how many of those windows did the model correctly classify as sitting?).
+High sensitivity means that events are unlikely to be missed.
+In contrast, positive predictive value (PPV) refers to the proportion of events classified for a given position that actually belonged to that position (e.g., if the model said a baby was upright during 100 windows, how many of those windows were indeed human-coded upright events?).
+High PPV means that we can be confident in the event label.
+Table \ref{tab:metrics-by-class} shows the sensitivity and PPV by body position class for group and individual models.
+For group models, sensitivity and PPV were similar: They were highest for supine, prone, and sitting (the most accurately identified class) and lowest for upright and held.
+Results for individual models were similar to group models, with the exception of a somewhat lower sensitivity score for held.
+Overall, the results suggest a reasonable balance between sensitivity and PPV among different body position classes for both model types.
+Table \ref{tab:metrics-table} shows the overall sensitivity and PPV across classes.
 
@@
 
-Beyond group and individual models, there were many degrees of freedom in our choice of how best to model body position.
+As in past work \autocite{AiraksinenRasanen2020}, performance was best when using all four sensors and degraded when using any subset of two (i.e., hips only, left leg only).
+Small decrements and accuracy were seen for some pairs (particularly when using the left ankle and hip), but decrements for other pairs (both ankles) or individual sensors were substantial.
+Using a hip and ankle pair instead of four sensors would be more cost effective without a large decrease in accuracy, but in some applications the increased accuracy might be worth the cost.
+Beyond group and individual models and sensor pairings, there were many degrees of freedom in our choice of how best to model body position.
 For the sake of brevity, we chose to report the best approach, not every possible variation in modeling.
 However, our openly shared data and code allow researcher to experiment with different sets of motion features, different sets of training data, different machine learning algorithms, and different hyperparameter values.
-As in past work \autocite{AiraksinenRasanen2020}, performance was best when using all four sensors and degraded when using any subset of two (i.e., hips only, left leg only).
 We tested but did not report models that used either only accelerometer data and only gyroscope data; in both cases, performance was degraded compared with models that use both types of motion.

### 2023-06-19 manuscript.Rmd
garment details

@@
 How can we produce similar data in service of measuring infants' everyday motor behaviors, such as body position?
-The aim of the current study was to validate long form recordings of infant position (supine, prone, sitting, upright, held by caregiver) based on machine learning classification of data from inertial sensors worn on infants' ankles and hips.
+The aim of the current study was to validate long form recordings of infant position (supine, prone, sitting, upright, held by caregiver) based on machine learning classification of data from inertial sensors worn on infants' ankles and thighs.
 Using over 100 hours of video recordings synchronized with inertial sensor data from infants in their homes, we demonstrate that body position classifications are sufficiently accurate to measure infant behavior.
@@
 In the current study, we report the feasibility and validity of body position classification over the full day in the home based on 34 testing sessions from 22 infants aged 4-14 months.
-Participants received a custom pair of infant leggings embedded with 4 IMUs (one on each ankle and one on each hip) and a video camera to collect ground truth data about infant body position.
+Participants received a custom pair of infant leggings embedded with 4 IMUs (one on each ankle and one on each thigh) and a video camera to collect ground truth data about infant body position.
 A *proximal comparison* period began when participants received the equipment and completed a guided phone call during which caregivers were asked to elicit different body positions based on prompts from the experimenter.
@@
 A custom garment was designed to hold the IMUs: Internal pockets were sewn into a snug-fitting pair of infant leggings so that IMUs would stay close to the body (reducing vibration), which also prevented infants from pulling out the sensors.
-A pocket over the hip and a pocket just above the ankle were sewn on the lateral surface of the right and left legs of the garment.
+A pocket over the thigh and a pocket just above the ankle were sewn on the lateral surface of the right and left legs of the garment.
+Multiple sizes of the garment were created (modeled on US 3-6 mo, 6-9 mo, 9-12 mo, and 12-18 mo sizing).
+Caregivers indicated in advance which size would fit their infant, and to "size down" if on the cusp to ensure a snug fit and reduce sensor movement.
+Each garment had a distinct pattern on the seat of the pants so that caregivers could identify front versus back and place the garment in the correct orientation.
+
 Each sensor recorded accelerometer and gyroscope data throughout the day, with sufficient battery and onboard storage to record for approximately 12 hours.
@@
 However, LENA&reg; data were not analyzed as part of the current study.
+
+```{r garment, out.width="0.75\\linewidth", include=TRUE, fig.align="center", fig.cap="Sensor garment worn by infant participant.
+Four IMUs were placed in interior pockets sewn into a tightly-fitting pair of infant leggings.
+White dashed rectangles mark the approximate locations of each sensor pocket (above the left and right ankles and on the left and right thighs, just below the thighs).
+A white dashed rectangle also marks the LENA audio recorder worn in a pocket on the infant's shirt.", echo=FALSE}
+knitr::include_graphics("figures/garment.jpg")
+```
 
@@
 Using the synchronization point, human-coded body annotations from video were linked to the corresponding times in the IMU time series data.
-A single, merged dataset was created with synchronized accelerometer signals (in three orientations: X, Y, and Z) and gyroscope signals (in three orientations: roll, pitch, and yaw) for each of the four sensors (left hip, right hip, left ankle, right ankle) with the corresponding timestamp and body position code using the *timetk* package [@DanchoVaughan2023] and the *lubridate* package [@GrolemundWickham2011] in R version 4.1.2 [@R42].
+A single, merged dataset was created with synchronized accelerometer signals (in three orientations: X, Y, and Z) and gyroscope signals (in three orientations: roll, pitch, and yaw) for each of the four sensors (left thigh, right thigh, left ankle, right ankle) with the corresponding timestamp and body position code using the *timetk* package [@DanchoVaughan2023] and the *lubridate* package [@GrolemundWickham2011] in R version 4.1.2 [@R42].
 
@@
 The 250 samples of data in each window were aggregated into a variety of motion features---summary statistics that could be fed into the machine learning model.
-The minimum, maximum, 25th percentile, 75th percentile, mean, median, skew, kurtosis, standard deviation, and sum were computed for each signal (e.g., right hip X linear acceleration, left ankle pitch angular acceleration).
+The minimum, maximum, 25th percentile, 75th percentile, mean, median, skew, kurtosis, standard deviation, and sum were computed for each signal (e.g., right thigh X linear acceleration, left ankle pitch angular acceleration).
 The 10 summary statistics and 24 sensor signals generated 240 columns of motion features that described movement within each window.
@@
                                          "left_hip","left_ankle", "right_hip", "right_ankle"),
-                              labels = c("All", "Left Hip/Ankle", "Right Hip/Ankle", "Both Hips", "Both Ankles", 
-                                         "Left Hip", "Left Ankle", "Right Hip", "Right Ankle")),
+                              labels = c("All", "Left Hip/Ankle", "Right Thigh/Ankle", "Both Thighs", "Both Ankles", 
+                                         "Left Thigh", "Left Ankle", "Right Thigh", "Right Ankle")),
                               Model = factor(model, levels = c("group", "split", labels = c("Group", "Individual"))))
@@
 The top row shows performance using all features calculated from the four sensors.
-Rows 2-5 show accuracy using pairs of sensors (left hip and ankle, right hip and ankle, left and right hip, left and right ankle), and rows 6-9 show accuracy using each individual sensor.")
+Rows 2-5 show accuracy using pairs of sensors (left thigh and ankle, right thigh and ankle, left and right thigh, left and right ankle), and rows 6-9 show accuracy using each individual sensor.")
 metrics_sensor <- metrics_sensor %>% column_to_rownames("type")
@@
 The highest accuracy was observed when using all four sensors (group model *M* = `r metrics_sensor["All", "mean_group"]`; individual model *M* = `r metrics_sensor["All", "mean_split"]`), and the lowest when using only a single sensor (left ankle group model *M* = `r metrics_sensor["Left Ankle", "mean_group"]`; right ankle individual model *M* = `r metrics_sensor["Right Ankle", "mean_split"]`).
-Some sensor pairs showed comparable accuracy to using all four sensors: Left ankle and hip has group model accuracy of *M* = `r metrics_sensor["Left Hip/Ankle", "mean_group"]` and individual model accuracy of *M* = `r metrics_sensor["Left Hip/Ankle", "mean_split"]`, only 2.2% and 1.5% worse than using all four sensors.
+Some sensor pairs showed comparable accuracy to using all four sensors: Left ankle and thigh has group model accuracy of *M* = `r metrics_sensor["Left Thigh/Ankle", "mean_group"]` and individual model accuracy of *M* = `r metrics_sensor["Left Thigh/Ankle", "mean_split"]`, only 2.2% and 1.5% worse than using all four sensors.
 Other pairings were less accurate; notably, using both ankles resulted in group model accuracy of *M* = `r metrics_sensor["Both Ankles", "mean_group"]` and individual model accuracy of *M* = `r metrics_sensor["Both Ankles", "mean_split"]`, 12.7% and 7.2% worse than using all four sensors.
@@
 
-As in past work [@AiraksinenRasanen2020], performance was best when using all four sensors and degraded when using any subset of two (i.e., hips only, left leg only).
-Small decrements and accuracy were seen for some pairs (particularly when using the left ankle and hip), but decrements for other pairs (both ankles) or individual sensors were substantial.
-Using a hip and ankle pair instead of four sensors would be more cost effective without a large decrease in accuracy, but in some applications the increased accuracy might be worth the cost.
+As in past work [@AiraksinenRasanen2020], performance was best when using all four sensors and degraded when using any subset of two (i.e., thighs only, left leg only).
+Small decrements and accuracy were seen for some pairs (particularly when using the left ankle and thigh), but decrements for other pairs (both ankles) or individual sensors were substantial.
+Using a thigh and ankle pair instead of four sensors would be more cost effective without a large decrease in accuracy, but in some applications the increased accuracy might be worth the cost.
 Beyond group and individual models and sensor pairings, there were many degrees of freedom in our choice of how best to model body position.
@@
 In the future, algorithms can be developed to automatically identify periods of sleep and times when sensors are off the body.
-Such algorithms exist for adults using wrist-worn sensors, however, we did not use them because they have not been validated for infant participants wearing sensors on hips/ankles.
+Such algorithms exist for adults using wrist-worn sensors, however, we did not use them because they have not been validated for infant participants wearing sensors on thighs/ankles.
 Differences in infants' positioning while asleep (held in caregivers arms, laying in cribs, reclined in strollers and carseats) might make classifying sleep more difficult compared with adults based on movement.

### 2023-06-19 manuscript.tex
garment details

--- 2023-06-16 manuscript.tex
@@
 In the current study, we report the feasibility and validity of body position classification over the full day in the home based on 34 testing sessions from 22 infants aged 4-14 months.
-Participants received a custom pair of infant leggings embedded with 4 IMUs (one on each ankle and one on each hip) and a video camera to collect ground truth data about infant body position.
+Participants received a custom pair of infant leggings embedded with 4 IMUs (one on each ankle and one on each thigh) and a video camera to collect ground truth data about infant body position.
 A \emph{proximal comparison} period began when participants received the equipment and completed a guided phone call during which caregivers were asked to elicit different body positions based on prompts from the experimenter.
@@
 A custom garment was designed to hold the IMUs: Internal pockets were sewn into a snug-fitting pair of infant leggings so that IMUs would stay close to the body (reducing vibration), which also prevented infants from pulling out the sensors.
-A pocket over the hip and a pocket just above the ankle were sewn on the lateral surface of the right and left legs of the garment.
+A pocket over the thigh and a pocket just above the ankle were sewn on the lateral surface of the right and left legs of the garment.
+Multiple sizes of the garment were created (modeled on US 3-6 mo, 6-9 mo, 9-12 mo, and 12-18 mo sizing).
+Caregivers indicated in advance which size would fit their infant, and to ``size down'' if on the cusp to ensure a snug fit and reduce sensor movement.
+Each garment had a distinct pattern on the seat of the pants so that caregivers could identify front versus back and place the garment in the correct orientation.
+
 Each sensor recorded accelerometer and gyroscope data throughout the day, with sufficient battery and onboard storage to record for approximately 12 hours.
@@
 However, LENA® data were not analyzed as part of the current study.
+
+\begin{figure}
+
+{\centering \includegraphics[width=0.75\linewidth]{figures/garment}
+
+}
+
+\caption{Sensor garment worn by infant participant.
+Four IMUs were placed in interior pockets sewn into a tightly-fitting pair of infant leggings.
+White dashed rectangles mark the approximate locations of each sensor pocket (above the left and right ankles and on the left and right thighs, just below the thighs).
+A white dashed rectangle also marks the LENA audio recorder worn in a pocket on the infant's shirt.}\label{fig:garment}
+\end{figure}
 
@@
 Using the synchronization point, human-coded body annotations from video were linked to the corresponding times in the IMU time series data.
-A single, merged dataset was created with synchronized accelerometer signals (in three orientations: X, Y, and Z) and gyroscope signals (in three orientations: roll, pitch, and yaw) for each of the four sensors (left hip, right hip, left ankle, right ankle) with the corresponding timestamp and body position code using the \emph{timetk} package \autocite{DanchoVaughan2023} and the \emph{lubridate} package \autocite{GrolemundWickham2011} in R version 4.1.2 \autocite{R42}.
+A single, merged dataset was created with synchronized accelerometer signals (in three orientations: X, Y, and Z) and gyroscope signals (in three orientations: roll, pitch, and yaw) for each of the four sensors (left thigh, right thigh, left ankle, right ankle) with the corresponding timestamp and body position code using the \emph{timetk} package \autocite{DanchoVaughan2023} and the \emph{lubridate} package \autocite{GrolemundWickham2011} in R version 4.1.2 \autocite{R42}.
 
@@
 
+\begin{table}[tbp]
+
+\begin{center}
+\begin{threeparttable}
+
+\caption{\label{tab:metrics-reduce}Median, mean, and SD of overall accuracy calculated with different sets of sensor features.
+The top row shows performance using all features calculated from the four sensors.
+Rows 2-5 show accuracy using pairs of sensors (left thigh and ankle, right thigh and ankle, left and right thigh, left and right ankle), and rows 6-9 show accuracy using each individual sensor.}
+
+\begin{tabular}{lllllll}
+\toprule
+ & \multicolumn{3}{c}{Group} & \multicolumn{3}{c}{Individual} \\
+\cmidrule(r){2-4} \cmidrule(r){5-7}
+Sensors & Median & Mean & SD & Median & Mean & SD\\
+\midrule
+All & 0.894 & 0.846 & 0.131 & 0.933 & 0.916 & 0.072\\ \midrule
+Left Hip/Ankle & 0.891 & 0.824 & 0.177 & 0.915 & 0.901 & 0.071\\
+Right Thigh/Ankle & 0.842 & 0.800 & 0.167 & 0.897 & 0.898 & 0.086\\
+Both Thighs & 0.859 & 0.791 & 0.195 & 0.907 & 0.898 & 0.072\\
+Both Ankles & 0.797 & 0.719 & 0.196 & 0.864 & 0.844 & 0.108\\ \midrule
+Left Thigh & 0.842 & 0.762 & 0.208 & 0.889 & 0.872 & 0.077\\
+Left Ankle & 0.717 & 0.667 & 0.208 & 0.851 & 0.825 & 0.117\\
+Right Thigh & 0.846 & 0.772 & 0.178 & 0.896 & 0.876 & 0.092\\
+Right Ankle & 0.704 & 0.683 & 0.160 & 0.830 & 0.819 & 0.114\\
+\bottomrule
+\end{tabular}
+
+\end{threeparttable}
+\end{center}
+
+\end{table}
+
 \hypertarget{overall-accuracy}{%
@@
 Possibly, individual models accounted for idiosyncrasies in behavior or inconsistencies in sensor placement for those sessions that group models could not generalize to.
+
+Similar to past work \autocite{AiraksinenRasanen2020}, overall accuracy decreased when fewer sensors were used.
+Table \ref{tab:metrics-reduce} compared overall accuracy for all four sensors (top row), pairs of 2 sensors (rows 2-5), and single sensors (rows 6-9).
+The highest accuracy was observed when using all four sensors (group model \emph{M} = 0.846; individual model \emph{M} = 0.916), and the lowest when using only a single sensor (left ankle group model \emph{M} = 0.667; right ankle individual model \emph{M} = 0.819).
+Some sensor pairs showed comparable accuracy to using all four sensors: Left ankle and thigh has group model accuracy of \emph{M} = NA and individual model accuracy of \emph{M} = NA, only 2.2\% and 1.5\% worse than using all four sensors.
+Other pairings were less accurate; notably, using both ankles resulted in group model accuracy of \emph{M} = 0.719 and individual model accuracy of \emph{M} = 0.844, 12.7\% and 7.2\% worse than using all four sensors.
 
@@
 
-Beyond group and individual models, there were many degrees of freedom in our choice of how best to model body position.
+As in past work \autocite{AiraksinenRasanen2020}, performance was best when using all four sensors and degraded when using any subset of two (i.e., thighs only, left leg only).
+Small decrements and accuracy were seen for some pairs (particularly when using the left ankle and thigh), but decrements for other pairs (both ankles) or individual sensors were substantial.
+Using a thigh and ankle pair instead of four sensors would be more cost effective without a large decrease in accuracy, but in some applications the increased accuracy might be worth the cost.
+Beyond group and individual models and sensor pairings, there were many degrees of freedom in our choice of how best to model body position.
 For the sake of brevity, we chose to report the best approach, not every possible variation in modeling.
 However, our openly shared data and code allow researcher to experiment with different sets of motion features, different sets of training data, different machine learning algorithms, and different hyperparameter values.
-As in past work \autocite{AiraksinenRasanen2020}, performance was best when using all four sensors and degraded when using any subset of two (i.e., hips only, left leg only).
 We tested but did not report models that used either only accelerometer data and only gyroscope data; in both cases, performance was degraded compared with models that use both types of motion.

### 2023-06-20 manuscript.Rmd
cutting

@@
 
-From moment to moment, infants' movements facilitate and constrain how they can interact with their surroundings.
-Changes in *body position*---whether infants are supine on their backs, prone on their bellies, sitting, upright, or held by a caregiver---have immediate consequences for vision, object exploration, and social interaction.
-When sitting and upright, infants have a better view of faces and distant objects compared to their view while in a prone position [@Freeplay; @LuoFranchak2020; @CWW].
-Learning to sit facilitates object exploration [@SoskaAdolph2014].
-Upright walking changes how infants and caregivers interact; walking infants move farther away from caregivers and share toys in different ways compared to infants crawling in a prone position [@ChenSchneider2022; @Karasik2011].
+Infants' movements facilitate and constrain how they can interact with their surroundings.
+Changes in *body position*---whether infants are supine on their backs, prone on their bellies, sitting, upright, or held by a caregiver---have in-the-moment consequences for vision, object exploration, and social interaction.
+When sitting and upright, infants have a better view of faces and distant objects compared to their view while prone [@Freeplay; @LuoFranchak2020; @CWW].
+While walking upright, infants move farther away from caregivers and share toys in different ways compared to infants crawling in a prone position [@ChenSchneider2022; @Karasik2011].
 As infants grow older and acquire new abilities, such as independent sitting and walking, they spend more time sitting and upright and less time held, supine, and prone [@Survey; @Thurman2017; @Freeplay; @AdolphCTL2014].
-Caregivers provide different learning input to infants depending on infants' body position [@KretchKoziol2022; @WestIverson2021; @Karasik2014].
-For example, infants' walking movements spur caregivers to talk about actions rather than just describe objects.
 Thus, characterizing individual differences in the day-to-day accumulation of body position experiences informs developmental theory by revealing differential opportunities for learning [@CurrentOpinion].
 
-In this paper, we present an inertial sensing method to classify infant body position from moment to moment across an entire day, and validate its accuracy using over 100 hours of video recorded across 34 in-home data collection sessions.
-Our method takes inspiration from a more mature technology: Long-form audio recordings of infants' language experiences using wearable audio recorders.
-We begin by describing the impact of long-form audio recordings on research in developmental psychology, and identify the key features that should be replicated in long-form recordings of motor behavior.
+In this paper, we present an inertial sensing method to classify infants' full-day, real-time body position.
+Our method takes inspiration from a more mature technology: Long-form audio recordings of infants' language experiences.
+We begin by identifying the key features of wearable audio recorders that should be replicated in long-form recordings of motor behavior.
 Next, we review the current state-of-the-art in measuring infant motor behavior---video and survey data---and their limitations in capturing real-time, full-day behavior.
@@
 Closed-source LENA&reg; algorithms analyze the audio recordings to provide automatic counts of useful metrics, such as the number of words spoken by adults in the vicinity of the infant.
-Other long-form audio methods rely on custom-built recorders [@WassPhillips2022], apply alternative classification algorithms to LENA&reg; data [@RasanenSeshadri2020; @MichelettiYao2022], or manually transcribe audio recorded by LENA&reg; devices to improve accuracy or identify behaviors beyond the built-in categories [@Bergelson2019Corpus; @MendozaFausey2021].
-
 Long-form audio recordings have had a transformational impact on language development research by allowing researchers to characterize opportunities for learning in daily life.
@@
 For example, infants' daily experiences hearing music are clustered in time, with "bursty" episodes of music presence separated by relatively long periods during which music is absent [@MendozaFausey2022].
-In addition, synchronizing audio recordings with other data sources extends researchers' ability to characterize daily experiences.
-Linking LENA&reg; speech measurements to repeated, time-stamped text-message surveys about infant device placement revealed that infants heard less caregiver speech during periods that they were restrained in devices such as swings, exersaucers, and car seats [@MalachowskiSalo2023].
-A custom-built wearable ECG and audio recorder allowed @WassPhillips2022 to discover that infant arousal increases the likelihood of infant vocalization across the day.
 
@@
 Third and fourth, recordings capture *real-time data* over a *full day*.
-The ability to record real-time data is vital for making inferences about processes that happen on the timescale of minutes or even seconds within an individual as opposed to comparisons of aggregated data between infants.
+Real-time data are vital for identifying processes that unfold over minutes or seconds within an individual as opposed to comparisons of aggregated data between infants.
 Synchronizing real-time data to other data streams helps to reveal sources of variability within an individual [e.g., @WassPhillips2022; @MalachowskiSalo2023].
@@
 
-But automatic classification can only replace human annotation if it is sufficiently accurate and unbiased.
-An independent assessment of the LENA&reg; algorithms found mixed results about the accuracy of different measurements.
-For example, correlations between human transcribed counts of adult words and child vocalizations against LENA&reg;'s automatic counts were strong, *r* = .698 and *r* = .649, respectively [@CristiaLavechin2020].
+But automatic classification can only replace human annotation if it is sufficiently accurate.
+An independent assessment of the LENA&reg; algorithms found mixed results about classification accuracy.
+Correlations between human transcribed counts of adult words and child vocalizations against LENA&reg;'s automatic counts were strong, *r* = .698 and *r* = .649, respectively [@CristiaLavechin2020].
 For other metrics, such as the number of "conversational turns" between the child and communicative partners, agreement was poor (*r* = .364).
 Thus, for some use cases (and for some metrics), long form audio recordings provide a mobile, unobtrusive way to automatically score real-time data over a full day.
-In the remainder of the paper, we turn to the question of how to replicate those key features in long form recordings of motor behavior.
+<!-- In the remainder of the paper, we turn to the question of how to replicate those key features in long form recordings of motor behavior. -->
 
@@
 Video and survey methods are the current state-of-the-art in assessing infants' gross motor behavior in naturalistic tasks.
-Although each method has complementary advantages and disadvantages for characterizing infants' everyday motor experiences, neither method on its own can provide comparable data to the long form audio recordings reviewed in the previous section.
+Although each method has complementary advantages and disadvantages for characterizing infants' everyday motor experiences, neither method fulfills all five key features of long form audio recordings reviewed in the previous section.
 
@@
 Most often, an experimenter with a handheld camera follows infants from room to room to ensure that their movements are visible throughout the recording session [@Karasik2011; @ChenSchneider2022; @HerzbergFletcher2021].
-The primary advantage of video recording is that it captures real-time behavior.
-Infants' body position, locomotion, and reaching comprise events that occur on the timescale of seconds, so standard video recording is adequate to score gross motor behavior.
+The primary advantage of video recording is that it captures real-time behavior.Standard 30 Hz video recording is adequate to capture changes in infant body position that occur on the timescale of seconds.
 However, requiring an experimenter to operate a camera is obtrusive, whereas relying on a stationary camera means that infants will be absent from view as they move from place to place.
 Moreover, video observation cannot easily scale to long durations or large numbers of participants.
-Logistically, an experimenter cannot follow behind infants to record their behavior from morning to night (and were they to do so, they would likely alter infants' and caregivers' behavior).
-Typical video recording sessions last 45-120 minutes [@Karasik2011; @ChenSchneider2022; @HerzbergFletcher2021], far short of capturing the variety of activities across the full daily routine.
-Even if full day videos were available, the lack of suitable automatic classification tools means that the human cost of annotation makes it difficult for video methods to scale to testing large numbers of participants.
-Our annotation of body position takes approximately 2-5 hours to complete for every hour of video (depending on how often infants switch positions), meaning that a full "waking day" of approximately 11 hours for a 12-month-old [@GallandTaylor2012] could take 22-55 hours of labor to fully annotate.
+An experimenter cannot follow behind infants to record their behavior from morning to night; typical video recording sessions last 45-120 minutes [@Karasik2011; @ChenSchneider2022; @HerzbergFletcher2021], far short of capturing the variety of activities across the full daily routine.
+Even if full-day videos were available, the lack of suitable automatic classification tools means that the human cost of annotation would be immense.
+Our annotation of body position takes approximately 2-5 hours to complete for every hour of video (depending on how often infants switch positions), meaning that a full "waking day" of approximately 11 hours for a 12-month-old [@GallandTaylor2012] could take 22-55 hours of labor to annotate.
 
 In contrast, survey methods such as daily diaries/inventories or ecological momentary assessment (EMA) are mobile, unobtrusive, can be applied across an entire day, and do not need laborious annotation.
-Diary studies provide caregivers with logs or structured interviews to estimate from memory how much time infants spend in particular activities [@MajnemerBarr2005; @KarasikKuchirko2022].
-Ecological momentary assessment uses text-message or app-based notifications to prompt caregivers to make repeated estimates of infants' behavior over the course of a day [@Survey; @KadookaCaufield2021].
-Although the EMA responses are valuable in aggregate, survey methods lack the real-time temporal resolution to describe moment-to-moment changes in behavior.
-For diaries and interviews, limits on caregivers' memory mean that they will report what was most frequent, but cannot remember events that happen on the scale of seconds and minutes.
+Diary studies provide caregivers with logs or structured interviews record activities [@MajnemerBarr2005; @KarasikKuchirko2022].
+Ecological momentary assessment uses smartphone notifications to prompt caregivers to make repeated estimates of behaviors throughout the day [@Survey; @KadookaCaufield2021].
+Although such survey responses are valuable in aggregate, they  lack the real-time temporal resolution to describe moment-to-moment changes in behavior.
 At best, EMA surveys prompt caregivers to make hourly observations; increasing the number of surveys per day would be too burdensome for the respondent.

### 2023-06-20 manuscript.tex
cutting

@@
 
-From moment to moment, infants' movements facilitate and constrain how they can interact with their surroundings.
-Changes in \emph{body position}---whether infants are supine on their backs, prone on their bellies, sitting, upright, or held by a caregiver---have immediate consequences for vision, object exploration, and social interaction.
-When sitting and upright, infants have a better view of faces and distant objects compared to their view while in a prone position \autocite{Freeplay,LuoFranchak2020,CWW}.
-Learning to sit facilitates object exploration \autocite{SoskaAdolph2014}.
-Upright walking changes how infants and caregivers interact; walking infants move farther away from caregivers and share toys in different ways compared to infants crawling in a prone position \autocite{ChenSchneider2022,Karasik2011}.
+Infants' movements facilitate and constrain how they can interact with their surroundings.
+Changes in \emph{body position}---whether infants are supine on their backs, prone on their bellies, sitting, upright, or held by a caregiver---have in-the-moment consequences for vision, object exploration, and social interaction.
+When sitting and upright, infants have a better view of faces and distant objects compared to their view while prone \autocite{Freeplay,LuoFranchak2020,CWW}.
+While walking upright, infants move farther away from caregivers and share toys in different ways compared to infants crawling in a prone position \autocite{ChenSchneider2022,Karasik2011}.
 As infants grow older and acquire new abilities, such as independent sitting and walking, they spend more time sitting and upright and less time held, supine, and prone \autocite{Survey,Thurman2017,Freeplay,AdolphCTL2014}.
-Caregivers provide different learning input to infants depending on infants' body position \autocite{KretchKoziol2022,WestIverson2021,Karasik2014}.
-For example, infants' walking movements spur caregivers to talk about actions rather than just describe objects.
 Thus, characterizing individual differences in the day-to-day accumulation of body position experiences informs developmental theory by revealing differential opportunities for learning \autocite{CurrentOpinion}.
 
-In this paper, we present an inertial sensing method to classify infant body position from moment to moment across an entire day, and validate its accuracy using over 100 hours of video recorded across 34 in-home data collection sessions.
-Our method takes inspiration from a more mature technology: Long-form audio recordings of infants' language experiences using wearable audio recorders.
-We begin by describing the impact of long-form audio recordings on research in developmental psychology, and identify the key features that should be replicated in long-form recordings of motor behavior.
+In this paper, we present an inertial sensing method to classify infants' full-day, real-time body position.
+Our method takes inspiration from a more mature technology: Long-form audio recordings of infants' language experiences.
+We begin by identifying the key features of wearable audio recorders that should be replicated in long-form recordings of motor behavior.
 Next, we review the current state-of-the-art in measuring infant motor behavior---video and survey data---and their limitations in capturing real-time, full-day behavior.
@@
 Closed-source LENA® algorithms analyze the audio recordings to provide automatic counts of useful metrics, such as the number of words spoken by adults in the vicinity of the infant.
-Other long-form audio methods rely on custom-built recorders \autocite{WassPhillips2022}, apply alternative classification algorithms to LENA® data \autocite{RasanenSeshadri2020,MichelettiYao2022}, or manually transcribe audio recorded by LENA® devices to improve accuracy or identify behaviors beyond the built-in categories \autocite{Bergelson2019Corpus,MendozaFausey2021}.
-
 Long-form audio recordings have had a transformational impact on language development research by allowing researchers to characterize opportunities for learning in daily life.
@@
 For example, infants' daily experiences hearing music are clustered in time, with ``bursty'' episodes of music presence separated by relatively long periods during which music is absent \autocite{MendozaFausey2022}.
-In addition, synchronizing audio recordings with other data sources extends researchers' ability to characterize daily experiences.
-Linking LENA® speech measurements to repeated, time-stamped text-message surveys about infant device placement revealed that infants heard less caregiver speech during periods that they were restrained in devices such as swings, exersaucers, and car seats \autocite{MalachowskiSalo2023}.
-A custom-built wearable ECG and audio recorder allowed \textcite{WassPhillips2022} to discover that infant arousal increases the likelihood of infant vocalization across the day.
 
@@
 Third and fourth, recordings capture \emph{real-time data} over a \emph{full day}.
-The ability to record real-time data is vital for making inferences about processes that happen on the timescale of minutes or even seconds within an individual as opposed to comparisons of aggregated data between infants.
+Real-time data are vital for identifying processes that unfold over minutes or seconds within an individual as opposed to comparisons of aggregated data between infants.
 Synchronizing real-time data to other data streams helps to reveal sources of variability within an individual \autocites[e.g.,][]{WassPhillips2022,MalachowskiSalo2023}.
@@
 
-But automatic classification can only replace human annotation if it is sufficiently accurate and unbiased.
-An independent assessment of the LENA® algorithms found mixed results about the accuracy of different measurements.
-For example, correlations between human transcribed counts of adult words and child vocalizations against LENA®'s automatic counts were strong, \emph{r} = .698 and \emph{r} = .649, respectively \autocite{CristiaLavechin2020}.
+But automatic classification can only replace human annotation if it is sufficiently accurate.
+An independent assessment of the LENA® algorithms found mixed results about classification accuracy.
+Correlations between human transcribed counts of adult words and child vocalizations against LENA®'s automatic counts were strong, \emph{r} = .698 and \emph{r} = .649, respectively \autocite{CristiaLavechin2020}.
 For other metrics, such as the number of ``conversational turns'' between the child and communicative partners, agreement was poor (\emph{r} = .364).
 Thus, for some use cases (and for some metrics), long form audio recordings provide a mobile, unobtrusive way to automatically score real-time data over a full day.
-In the remainder of the paper, we turn to the question of how to replicate those key features in long form recordings of motor behavior.
 
@@
 Most often, an experimenter with a handheld camera follows infants from room to room to ensure that their movements are visible throughout the recording session \autocite{Karasik2011,ChenSchneider2022,HerzbergFletcher2021}.
-The primary advantage of video recording is that it captures real-time behavior.
-Infants' body position, locomotion, and reaching comprise events that occur on the timescale of seconds, so standard video recording is adequate to score gross motor behavior.
+The primary advantage of video recording is that it captures real-time behavior.Standard 30 Hz video recording is adequate to capture changes in infant body position that occur on the timescale of seconds.
 However, requiring an experimenter to operate a camera is obtrusive, whereas relying on a stationary camera means that infants will be absent from view as they move from place to place.
 Moreover, video observation cannot easily scale to long durations or large numbers of participants.
-Logistically, an experimenter cannot follow behind infants to record their behavior from morning to night (and were they to do so, they would likely alter infants' and caregivers' behavior).
-Typical video recording sessions last 45-120 minutes \autocite{Karasik2011,ChenSchneider2022,HerzbergFletcher2021}, far short of capturing the variety of activities across the full daily routine.
-Even if full day videos were available, the lack of suitable automatic classification tools means that the human cost of annotation makes it difficult for video methods to scale to testing large numbers of participants.
-Our annotation of body position takes approximately 2-5 hours to complete for every hour of video (depending on how often infants switch positions), meaning that a full ``waking day'' of approximately 11 hours for a 12-month-old \autocite{GallandTaylor2012} could take 22-55 hours of labor to fully annotate.
+An experimenter cannot follow behind infants to record their behavior from morning to night; typical video recording sessions last 45-120 minutes \autocite{Karasik2011,ChenSchneider2022,HerzbergFletcher2021}, far short of capturing the variety of activities across the full daily routine.
+Even if full-day videos were available, the lack of suitable automatic classification tools means that the human cost of annotation would be immense.
+Our annotation of body position takes approximately 2-5 hours to complete for every hour of video (depending on how often infants switch positions), meaning that a full ``waking day'' of approximately 11 hours for a 12-month-old \autocite{GallandTaylor2012} could take 22-55 hours of labor to annotate.
 
 In contrast, survey methods such as daily diaries/inventories or ecological momentary assessment (EMA) are mobile, unobtrusive, can be applied across an entire day, and do not need laborious annotation.
-Diary studies provide caregivers with logs or structured interviews to estimate from memory how much time infants spend in particular activities \autocite{MajnemerBarr2005,KarasikKuchirko2022}.
-Ecological momentary assessment uses text-message or app-based notifications to prompt caregivers to make repeated estimates of infants' behavior over the course of a day \autocite{Survey,KadookaCaufield2021}.
-Although the EMA responses are valuable in aggregate, survey methods lack the real-time temporal resolution to describe moment-to-moment changes in behavior.
-For diaries and interviews, limits on caregivers' memory mean that they will report what was most frequent, but cannot remember events that happen on the scale of seconds and minutes.
+Diary studies provide caregivers with logs or structured interviews record activities \autocite{MajnemerBarr2005,KarasikKuchirko2022}.
+Ecological momentary assessment uses smartphone notifications to prompt caregivers to make repeated estimates of behaviors throughout the day \autocite{Survey,KadookaCaufield2021}.
+Although such survey responses are valuable in aggregate, they lack the real-time temporal resolution to describe moment-to-moment changes in behavior.
 At best, EMA surveys prompt caregivers to make hourly observations; increasing the number of surveys per day would be too burdensome for the respondent.

### 2023-06-20 manuscript.Rmd
more cuts

@@
 Measuring infant movement with inertial movement units (IMUs) is a promising avenue for long form recordings of motor behavior in the home [@CliffReilly2009; @Barbaro2019; @BruijnsTruelove2020; @LoboHall2019].
-Lightweight sensors (10-30 g) can be embedded in garments to make recordings fully *mobile*, and they are *unobtrusive* because they do not require a researcher's presence nor do they record sensitive audio/video that might influence participants' behavior.
-Many commercially-available and inexpensive IMUs have > 12 hour battery life with onboard storage to record *real-time*, *full-day* motion data at a sampling rate that is higher than typical video (e.g., 50-100 Hz).
-Indeed, past work has successfully recorded the rate of leg kicks [e.g., @DengTrujillo-Priego2019] and activity intensity [@SchnellerBentsen2017] in infants and children across multiple days.
+Lightweight sensors (10-30 g) can be embedded in garments to make recordings fully *mobile*, and they are *unobtrusive* because they do not require a researcher to follow with a camcorder.
+Many commercially-available IMUs have > 12 hour battery life with onboard storage to record *real-time*, *full-day* motion data at a high sampling rate (e.g., 50-100 Hz).
 
 The open question is whether *automatic classification* is sufficiently accurate to measure movement categories that are relevant to developmental and clinical research, and whether measurements continue to be accurate over long recording periods.
-IMUs typically contain accelerometers that measure linear acceleration paired with gyroscopes that measure angular acceleration.
-Unlike motion tracking systems that might be used in a lab, IMUs do not provide data about the position of the body in space.
-Thus, data processing algorithms are needed to classify the raw sensor data (i.e., linear and angular acceleration time series) into meaningful categories (e.g., supine, prone, sitting, upright).
-The difficulty of the classification task depends on the categories of interest.
-More basic aspects of movement, such as overall activity intensity, can be identified by taking the magnitude of acceleration (irrespective of direction) and applying thresholds or cut-points to define when a movement has occurred at a particular intensity [e.g., @HagerTilton2017; @ArmstrongCovington2019].
-
-Categorizing body position---supine on the back, prone on the belly, sitting, upright, or held off the ground by a caregiver---is too complex for cut-point definitions to be accurate.
-First, the magnitude of movement can vary greatly *within* a body position.
+Data processing algorithms are needed to classify the raw sensor data (i.e., linear and angular acceleration time series) into meaningful categories.
+Categorizing body position---supine on the back, prone on the belly, sitting, upright, or held off the ground by a caregiver---is a complex because movement can vary greatly *within* a body position.
 An upright infant can be standing still or can be walking briskly across the room.
@@
 Sitting in a high chair or car seat reduces the magnitude of postural sway within sitting, and creates even more possibilities for how the arms and legs may move relative to the torso.
-Indeed, creating an all-encompassing set of rules for how to annotate sitting from video is no trivial task because of the various ways that sitting can occur in daily life.
 Finally, caregivers frequently pick up and transport infants, creating motion signals that need to be differentiated from independent activity [@KwonZavos2019; @PatelShi2019].
-Although @GreenspanCunha2021 were able to use cut-point definitions about infant trunk angle to categorize body positions, this method was not able to differentiate times that infants were held by caregivers off the ground.
-Thus, modern approaches to human activity recognition have used machine learning to classify activity categories based on features derived from IMU data in adults [@Preece2009; @Arif2015], children [@NamPark2013; @RenDing2016; @StewartNarayanan2018], and infants [@YaoPlotz2019; @FranchakScott2021; @AiraksinenRasanen2020].
-Synchronized video with ground-truth human annotations creates training data for a machine learning algorithm, such as a random forest model, that can be later used to predict categories from IMU data that was not annotated.
-Crucially, this allows automatic classification to scale to full day recording by relying on a relatively smaller set of video annotations.
-
+
+Modern approaches to human activity recognition have used machine learning to classify activity categories based on features derived from IMU data in adults [@Preece2009; @Arif2015], children [@NamPark2013; @RenDing2016; @StewartNarayanan2018], and infants [@YaoPlotz2019; @FranchakScott2021; @AiraksinenRasanen2020].
 Three prior investigations have used different machine learning techniques to categorize infant body position from IMUs towards the goal of collecting full-day data.

### 2023-06-20 manuscript.tex
more cuts

@@
 Measuring infant movement with inertial movement units (IMUs) is a promising avenue for long form recordings of motor behavior in the home \autocite{CliffReilly2009,Barbaro2019,BruijnsTruelove2020,LoboHall2019}.
-Lightweight sensors (10-30 g) can be embedded in garments to make recordings fully \emph{mobile}, and they are \emph{unobtrusive} because they do not require a researcher's presence nor do they record sensitive audio/video that might influence participants' behavior.
-Many commercially-available and inexpensive IMUs have \textgreater{} 12 hour battery life with onboard storage to record \emph{real-time}, \emph{full-day} motion data at a sampling rate that is higher than typical video (e.g., 50-100 Hz).
-Indeed, past work has successfully recorded the rate of leg kicks \autocite[e.g.,][]{DengTrujillo-Priego2019} and activity intensity \autocite{SchnellerBentsen2017} in infants and children across multiple days.
+Lightweight sensors (10-30 g) can be embedded in garments to make recordings fully \emph{mobile}, and they are \emph{unobtrusive} because they do not require a researcher to follow with a camcorder.
+Many commercially-available IMUs have \textgreater{} 12 hour battery life with onboard storage to record \emph{real-time}, \emph{full-day} motion data at a high sampling rate (e.g., 50-100 Hz).
 
 The open question is whether \emph{automatic classification} is sufficiently accurate to measure movement categories that are relevant to developmental and clinical research, and whether measurements continue to be accurate over long recording periods.
-IMUs typically contain accelerometers that measure linear acceleration paired with gyroscopes that measure angular acceleration.
-Unlike motion tracking systems that might be used in a lab, IMUs do not provide data about the position of the body in space.
-Thus, data processing algorithms are needed to classify the raw sensor data (i.e., linear and angular acceleration time series) into meaningful categories (e.g., supine, prone, sitting, upright).
-The difficulty of the classification task depends on the categories of interest.
-More basic aspects of movement, such as overall activity intensity, can be identified by taking the magnitude of acceleration (irrespective of direction) and applying thresholds or cut-points to define when a movement has occurred at a particular intensity \autocites[e.g.,][]{HagerTilton2017,ArmstrongCovington2019}.
-
-Categorizing body position---supine on the back, prone on the belly, sitting, upright, or held off the ground by a caregiver---is too complex for cut-point definitions to be accurate.
-First, the magnitude of movement can vary greatly \emph{within} a body position.
+Data processing algorithms are needed to classify the raw sensor data (i.e., linear and angular acceleration time series) into meaningful categories.
+Categorizing body position---supine on the back, prone on the belly, sitting, upright, or held off the ground by a caregiver---is a complex because movement can vary greatly \emph{within} a body position.
 An upright infant can be standing still or can be walking briskly across the room.
@@
 Sitting in a high chair or car seat reduces the magnitude of postural sway within sitting, and creates even more possibilities for how the arms and legs may move relative to the torso.
-Indeed, creating an all-encompassing set of rules for how to annotate sitting from video is no trivial task because of the various ways that sitting can occur in daily life.
 Finally, caregivers frequently pick up and transport infants, creating motion signals that need to be differentiated from independent activity \autocite{KwonZavos2019,PatelShi2019}.
-Although \textcite{GreenspanCunha2021} were able to use cut-point definitions about infant trunk angle to categorize body positions, this method was not able to differentiate times that infants were held by caregivers off the ground.
-Thus, modern approaches to human activity recognition have used machine learning to classify activity categories based on features derived from IMU data in adults \autocite{Preece2009,Arif2015}, children \autocite{NamPark2013,RenDing2016,StewartNarayanan2018}, and infants \autocite{YaoPlotz2019,FranchakScott2021,AiraksinenRasanen2020}.
-Synchronized video with ground-truth human annotations creates training data for a machine learning algorithm, such as a random forest model, that can be later used to predict categories from IMU data that was not annotated.
-Crucially, this allows automatic classification to scale to full day recording by relying on a relatively smaller set of video annotations.
-
+
+Modern approaches to human activity recognition have used machine learning to classify activity categories based on features derived from IMU data in adults \autocite{Preece2009,Arif2015}, children \autocite{NamPark2013,RenDing2016,StewartNarayanan2018}, and infants \autocite{YaoPlotz2019,FranchakScott2021,AiraksinenRasanen2020}.
 Three prior investigations have used different machine learning techniques to categorize infant body position from IMUs towards the goal of collecting full-day data.

### 2023-06-20 manuscript.Rmd
more cuts

@@
 Three prior investigations have used different machine learning techniques to categorize infant body position from IMUs towards the goal of collecting full-day data.
-@AiraksinenRasanen2020 tested 4- to 8-month-olds in a laboratory visit, and found 95% accuracy in distinguishing between body position categories that crawling infants could perform on the floor (excluding times that infants were held by caregivers).
-Using a wider age range of 6-18 months, @FranchakScott2021 found 98% accuracy (*kappa* = 95%) in a laboratory validation study in categorizing body position that included infants who could both crawl and walk and also included a category for caregiver holding.
-Most recently, @AiraksinenGallen2022 conducted a validation study of body position classification in either a home or clinic testing in 4- to 19-month-olds, refining their previous method to detect moments that infants were carried by caregivers.
+@AiraksinenRasanen2020 tested 4- to 8-month-olds in a laboratory visit with a 4-sensor array (one on each thigh and one below each shoulder), and found 95% accuracy in distinguishing between body position categories that crawling infants could perform on the floor (excluding times that infants were held by caregivers).
+Using a wider age range of 6-18 months, @FranchakScott2021 found 98% accuracy (*kappa* = 95%) in a laboratory validation study with a 3-sensor array (ankle, knee, thigh on a single leg) in categorizing body position that included infants who could both crawl and walk and also included a category for caregiver holding.
+Most recently, @AiraksinenGallen2022 conducted a validation study of body position classification in either a home or clinic testing 4- to 19-month-olds with their 4-sensor system, refining their previous method to detect moments that infants were carried by caregivers.
 Classification accuracy did not vary between lab and home settings, and was generally high (95%, *kappa* = .93).
@@
 However, in a real day, non-play activities (e.g., eating lunch in a high chair) create challenging situations for applying automated classification of body position.
-For example different types of sitting---independently on the floor, supported on a caregiver's lap, or restrained in a high chair---all need to scored as the same behavior.
+For example different types of sitting---independently on the floor, supported on a caregiver's lap, or restrained in a high chair---all need to be scored as sitting.
 Moreover, classifiers must be able to detect new variants of behaviors that might arise over the course of a real day; it is impossible for researchers to gather training data for every possible variation that might occur.
@@
 Thus, the **third goal** of the study was to examine the quality of the full-day data.
-Could body position be successfully recorded over the desired period?
+Could infants wear the sensor garment throughout the desired period?
 If full-day classifications of infant behavior are accurate, they should demonstrate convergent validity with other methods.
-Thus, we determined whether full-day body position measurements conformed to expected age differences in body position that have been identified from previous video and survey methods.
+Thus, we determined whether full-day body position measurements conformed to expected age differences in body position.
 Based on past results [@Survey], infants should spend increasingly more time sitting and upright but less time supine over the age range tested (4 to 14 months).
@@
 Four inertial movement units (IMUs) were used to record infant movement across the day (MC10 Biostamp).
-A custom garment was designed to hold the IMUs: Internal pockets were sewn into a snug-fitting pair of infant leggings so that IMUs would stay close to the body (reducing vibration), which also prevented infants from pulling out the sensors.
+A custom garment was designed to hold the IMUs: Internal pockets were sewn into snug-fitting infant leggings so that IMUs would stay close to the body (reducing vibration).
 A pocket over the thigh and a pocket just above the ankle were sewn on the lateral surface of the right and left legs of the garment.
 Multiple sizes of the garment were created (modeled on US 3-6 mo, 6-9 mo, 9-12 mo, and 12-18 mo sizing).
-Caregivers indicated in advance which size would fit their infant, and to "size down" if on the cusp to ensure a snug fit and reduce sensor movement.
+Caregivers indicated in advance which size would fit their infant, and to "size down" if between sizes to ensure a snug fit and minimize sensor movement.
 Each garment had a distinct pattern on the seat of the pants so that caregivers could identify front versus back and place the garment in the correct orientation.
 
-Each sensor recorded accelerometer and gyroscope data throughout the day, with sufficient battery and onboard storage to record for approximately 12 hours.
-We chose a sampling rate of 62.5 Hz (one of the available preset rates) based on prior work that used rates of 50-64 Hz [@FranchakScott2021; @AiraksinenRasanen2020; @YaoPlotz2019].
-Infants also wore a LENA&reg; recorder throughout the day in the front pocket of a LENA&reg; shirt, located near the infant’s chest, to determine whether data could be simultaneously recorded from the LENA&reg; and IMU sensors.
-However, LENA&reg; data were not analyzed as part of the current study.
+Each sensor had sufficient battery and onboard torage to record accelerometer and gyroscope data for approximately 12 hours.
+We chose a sampling rate of 62.5 Hz (one of the available presets) based on prior work that used rates of 50-64 Hz [@FranchakScott2021; @AiraksinenRasanen2020; @YaoPlotz2019].
+Infants also wore a LENA&reg; recorder throughout the day in the front pocket of a LENA&reg; shirt, located near the infant’s chest, to determine whether data could be simultaneously recorded from the LENA&reg; and IMU sensors (LENA&reg; data were not analyzed in the current study).
 
@@
 
-Video recordings were captured at 30 Hz using an action camera on a miniature tripod (Insta360 ONE R) that caregivers placed in the room that the infant was in.
-The proprietary "Boosted Battery Base" attached to the action camera to allow for a total of 3 hours of recording.
-As described below, the video recording was not continuous.
-The camera created two video files temporally separated by a gap of approximately 40-45 s.
-Caregivers also received a log sheet to record times that infants napped as well as times that the sensor garment was removed from the infant (e.g., baths, diaper changes, errands).
+Videos were captured at 30 Hz using an action camera on a miniature tripod (Insta360 ONE R) that caregivers placed in the same room as the infant.
+Although recordings lasted 3 hours, they were divided into two video files of approximately 90 minutes temporally separated by a gap of 40-45 s.
+Caregivers received a log sheet to record times that infants napped and times that the sensor garment was removed from the infant (e.g., baths, diaper changes, errands).
 
@@
 Figure \@ref(fig:exemplar-timeline) shows an exemplar timeline of the entire procedure and recording periods for a single participant.
-On the day of the visit, a researcher arrived at the participant’s home in the morning and prepared all the equipment at the doorstep.
-In order to synchronize all three recording devices, the researcher began by first turning on the video camera and the LENA audio device (the IMUs were already configured and placed before arrival because they required a proprietary sensor dock to begin recording in the laboratory).
-To create an easily recognizable synchronization point between the video recording and IMU data, the researcher dropped or struck the leggings (containing the IMUs) on a surface in view of the camera, as in @FranchakScott2021.
+A researcher arrived at the participant’s home in the morning and set each device to record while at the doorstep.
+To create a recognizable synchronization point between the video recording and IMU data, the researcher dropped the sensor garment containing the IMUs on a surface in view of the camera, as in @FranchakScott2021.
 All the equipment---once recording and with synchronization information recorded---was placed inside a large bucket and left outside the family's front door.
@@
 The researcher kept time to ensure at least 1 minute of behavior for each activity.
-Once completed, the caregiver was then instructed to play with their infant for 10 minutes within view of the camera to collect additional training data with the infant in positions that would be typical of play.
+Once completed, the caregiver was then instructed to play with their infant for 10 minutes within view of the camera to collect additional ground truth data.
 
@@
 The caregiver logged the times the sensors were removed (blank areas in the timeline in Figure \@ref(fig:exemplar-timeline)) or the child took a nap (gray areas in the timeline in Figure \@ref(fig:exemplar-timeline)) so that those times could be excluded from analysis.
-The following day a researcher picked up the equipment, verified that the paperwork was signed, and compensated the participant.
+The following day a researcher picked up the bucket of equipment.
 
@@
 Synchronization errors were estimated to be less than 30-60 ms (1-2 video frames).
-When the first video file ended, a second video file was automatically recorded, termed the *distal comparison*.
+The second video file comprised the *distal comparison*.
 This video recorded the next 90 minutes of natural activity.
-However, because of a limitation in the camera, there was a variable gap of ~40 s between the two videos, synchronization in the distal comparison video was coarser, with estimated temporal offsets of ~5 s in either direction.
+However, because of a limitation in the camera, there was a variable gap of ~40 s between the two videos, so synchronization in the distal comparison video was coarser, with estimated temporal offsets of ~5 s in either direction.
 
@@
 
-Supine was coded when the infant was lying on their back or was reclined up to a 45 degree angle.
-Supine was also coded in the rare cases when the infant was laying on their side.
-Prone was coded when the infant was lying on their stomach, was stationary supported by the hands/knees or the hands/feet, or was crawling.
+Supine was coded when the infant was lying on their back, on their side, or was reclined up to a 45  Prone was coded when the infant was lying on their stomach, was stationary supported by the hands/knees or the hands/feet, or was crawling.
 We scored sitting to include any form of the following seated positions: 1) infants sat with their buttocks on a surface, such as on the floor or a caregiver’s lap, 2) infant was in a kneeling-sit position, in which their knees were on the ground with their legs tucked underneath the buttocks, and 3) infant was in a seating device, such as a high chair, that kept the torso oriented perpendicular to the ground (a reclined position, such as in a young infant's car seat, would be counted as supine).
 Upright was coded when the infant was standing or squatting on the ground with two feet or walking (regardless of whether infants' balance was assisted by a caregiver or with their hands holding onto something for support).
-When an infant was carried by a caregiver, held was coded.
+Our goal in creating a category for "held by caregiver" was to separate times when infants were in control of their body position from times when they were suspended in the air (rather than resting on furniture or a surface).
+Held was coded when infants were carried off of the groun.
 However, when the caregiver was sitting with the infant in their lap the infant’s body position was coded as if the caregiver was a surface (e.g., if the infant was sitting on the caregiver’s lap this was coded as sitting).
@@
 Classification training and prediction was conducted on a windowed dataset that summarized the raw, 62.5 Hz motion signals within 4-s windows.
-Overlapping moving windows captured 4-s (250 samples) of data starting each second, which is a common unit of analysis in prior studies of human activity classification [@FranchakScott2021; @NamPark2013; @AiraksinenRasanen2020].
-The 250 samples of data in each window were aggregated into a variety of motion features---summary statistics that could be fed into the machine learning model.
+Data were reduced in time by creating overlapping moving windows (4-s long, comprising 250 samples) starting each second, which is a common unit of analysis in prior studies of human activity classification [@FranchakScott2021; @NamPark2013; @AiraksinenRasanen2020].
+For each 4-s window, we aggregated the 250 samples to create single scores for a variety of motion features---summary statistics that could be fed into the machine learning model.
 The minimum, maximum, 25th percentile, 75th percentile, mean, median, skew, kurtosis, standard deviation, and sum were computed for each signal (e.g., right thigh X linear acceleration, left ankle pitch angular acceleration).
@@
 The resulting windowed dataset was used for machine learning classification and validation.
-For each analysis reported in the results, a subset of data were defined as a "training" set and another, independent portion of the data were defined as a "testing" set.
-Different training/testing set combinations were used to address different goals; here we describe the general procedure of how a classifier was derived from a training set.
-Random forest models [@Breiman2001] were trained to predict the human coded body position label for each window from the set of 436 motion features using the *randomForest* package [@LiawWiener2002].
-Random forest models use random subsets of features and random subsets of data across many iterations; this random subsetting creates models that are less likely to be overfit.
-The *tidymodels* set of packages in R [@tidymodels] was used to tune the "mtry" and "ntrees" hyperparameters, which control the number of columns sampled and the total number of trees created when modeling.
-We use an mtry parameter of 44 and and ntree parameter of 550; however, changing hyperparameters had little impact on model performance.
-Just like a linear regression model, the resulting random forest model could later be applied to a set of testing data with the *predict* function.
-Validation analyses below will compare the human-coded annotations and random forest model predictions for testing data that are independent from training data, providing a test of whether models can successfully generalize.
+For each analysis reported in the results, a subset of data were defined as a "training" set and another, independent portion of the data were defined as a "testing" set (further described below).
+Random forest models [@Breiman2001] used the training set to learn the body position label for each window from the set of 436 motion features using the *randomForest* package [@LiawWiener2002].
+The resulting random forest model could later be applied to a set of testing data with the *predict* function.
 
@@
 We report three sets of results based on `r length(unique(ppts$id_uni))` full-day testing sessions resulting in a total of `r round(sum(duration$duration_hours))` hours of movement recording.
-First, we focus on the accuracy of group and individual models trained and validated from data during the proximal comparison period.
+
+<!-- First, we focus on the accuracy of group and individual models trained and validated from data during the proximal comparison period.
 This period is unique in having the highest degree of temporal synchrony between video and motion data, allowing us to assess the accuracy of individual body position events.
@@
 In the third set of results we examine the data quality of full-day recordings and test whether full-day recordings capture well-established age differences in body position experiences.
+-->
 
@@
 The testing set was never used as training data.
-The two types of models differed based on what data were used during training, but the same random forest modeling procedure was used in both cases.
-A group model was trained for each session using an aggregated dataset containing the first 75% of all **other** sessions' proximal comparison data.
+The two types of models differed based on what data were used during training: A group model was trained for each session using an aggregated dataset containing the first 75% of all **other** sessions' proximal comparison data.
 This leave-one-out cross-validation tested the generalization of the model to a recording session that was not used at all in the training set.
@@
 Figure \@ref(fig:metrics), Table \@ref(tab:metrics-table), and Table \@ref(tab:metrics-by-class) summarize the performance of group and individual models using standard metrics for classification.
-
 <!-- data/reduce, duplicates of the group_split metrics but performed on reductions of sensor features from group_split_sensor_reduce.R -->
@@
 Looking at the median performance suggests that the difference between models was not considerable for the typical participant (group median accuracy = `r metrics_lookup["Overall Accuracy", "median_Group"]`; individual median accuracy = `r metrics_lookup["Overall Accuracy", "median_Individual"]`).
-Indeed, it is notable that the worst-case accuracy from group models (minimum = `r metrics_lookup["Overall Accuracy", "min_Group"]`) was far lower than in individual models (minimum = `r metrics_lookup["Overall Accuracy", "min_Individual"]`).
+
+<!-- Indeed, it is notable that the worst-case accuracy from group models (minimum = `r metrics_lookup["Overall Accuracy", "min_Group"]`) was far lower than in individual models (minimum = `r metrics_lookup["Overall Accuracy", "min_Individual"]`).
 Possibly, individual models accounted for idiosyncrasies in behavior or inconsistencies in sensor placement for those sessions that group models could not generalize to.
-
-Similar to past work [@AiraksinenRasanen2020], overall accuracy decreased when fewer sensors were used.
+-->
+
+As in @AiraksinenRasanen2020, overall accuracy decreased when fewer sensors were used.
 Table \@ref(tab:metrics-reduce) compared overall accuracy for all four sensors (top row), pairs of 2 sensors (rows 2-5), and single sensors (rows 6-9).
 The highest accuracy was observed when using all four sensors (group model *M* = `r metrics_sensor["All", "mean_group"]`; individual model *M* = `r metrics_sensor["All", "mean_split"]`), and the lowest when using only a single sensor (left ankle group model *M* = `r metrics_sensor["Left Ankle", "mean_group"]`; right ankle individual model *M* = `r metrics_sensor["Right Ankle", "mean_split"]`).
-Some sensor pairs showed comparable accuracy to using all four sensors: Left ankle and thigh has group model accuracy of *M* = `r metrics_sensor["Left Thigh/Ankle", "mean_group"]` and individual model accuracy of *M* = `r metrics_sensor["Left Thigh/Ankle", "mean_split"]`, only 2.2% and 1.5% worse than using all four sensors.
+Accuracy for some pairs approached 4-sensor accuracy: Left ankle and thigh had group model accuracy of *M* = `r metrics_sensor["Left Thigh/Ankle", "mean_group"]` and individual model accuracy of *M* = `r metrics_sensor["Left Thigh/Ankle", "mean_split"]`, only 2.2% and 1.5% worse than using all four sensors.
 Other pairings were less accurate; notably, using both ankles resulted in group model accuracy of *M* = `r metrics_sensor["Both Ankles", "mean_group"]` and individual model accuracy of *M* = `r metrics_sensor["Both Ankles", "mean_split"]`, 12.7% and 7.2% worse than using all four sensors.
@@
 Strong overall accuracy can be misleading when the relative frequency of different classes is unbalanced.
-For example, most infants spent more time sitting than prone, so good performance for sitting classification could mask poorer performance in prone classification.
 Accordingly, we report Cohen's kappa, a commonly-used metric that penalizes missing rare events (Figure \@ref(fig:metrics)B), and we provide classification metrics for each individual body position (Table \@ref(tab:metrics-by-class)) to account for imbalance in body position rates within and between individuals.
@@
 
-Beyond accuracy, it is important to establish that classification is neither too sensitive nor too conservative in choosing one position versus another.
 Sensitivity refers to the proportion of events of a given position that were correctly identified (e.g., out of 100 human-coded sitting windows, how many of those windows did the model correctly classify as sitting?).

### 2023-06-20 manuscript.tex
more cuts

@@
 Three prior investigations have used different machine learning techniques to categorize infant body position from IMUs towards the goal of collecting full-day data.
-\textcite{AiraksinenRasanen2020} tested 4- to 8-month-olds in a laboratory visit, and found 95\% accuracy in distinguishing between body position categories that crawling infants could perform on the floor (excluding times that infants were held by caregivers).
-Using a wider age range of 6-18 months, \textcite{FranchakScott2021} found 98\% accuracy (\emph{kappa} = 95\%) in a laboratory validation study in categorizing body position that included infants who could both crawl and walk and also included a category for caregiver holding.
-Most recently, \textcite{AiraksinenGallen2022} conducted a validation study of body position classification in either a home or clinic testing in 4- to 19-month-olds, refining their previous method to detect moments that infants were carried by caregivers.
+\textcite{AiraksinenRasanen2020} tested 4- to 8-month-olds in a laboratory visit with a 4-sensor array (one on each thigh and one below each shoulder), and found 95\% accuracy in distinguishing between body position categories that crawling infants could perform on the floor (excluding times that infants were held by caregivers).
+Using a wider age range of 6-18 months, \textcite{FranchakScott2021} found 98\% accuracy (\emph{kappa} = 95\%) in a laboratory validation study with a 3-sensor array (ankle, knee, thigh on a single leg) in categorizing body position that included infants who could both crawl and walk and also included a category for caregiver holding.
+Most recently, \textcite{AiraksinenGallen2022} conducted a validation study of body position classification in either a home or clinic testing 4- to 19-month-olds with their 4-sensor system, refining their previous method to detect moments that infants were carried by caregivers.
 Classification accuracy did not vary between lab and home settings, and was generally high (95\%, \emph{kappa} = .93).
@@
 Thus, the \textbf{third goal} of the study was to examine the quality of the full-day data.
-Could body position be successfully recorded over the desired period?
+Could infants wear the sensor garment throughout the desired period?
 If full-day classifications of infant behavior are accurate, they should demonstrate convergent validity with other methods.
-Thus, we determined whether full-day body position measurements conformed to expected age differences in body position that have been identified from previous video and survey methods.
+Thus, we determined whether full-day body position measurements conformed to expected age differences in body position.
 Based on past results \autocite{Survey}, infants should spend increasingly more time sitting and upright but less time supine over the age range tested (4 to 14 months).
@@
 Four inertial movement units (IMUs) were used to record infant movement across the day (MC10 Biostamp).
-A custom garment was designed to hold the IMUs: Internal pockets were sewn into a snug-fitting pair of infant leggings so that IMUs would stay close to the body (reducing vibration), which also prevented infants from pulling out the sensors.
+A custom garment was designed to hold the IMUs: Internal pockets were sewn into snug-fitting infant leggings so that IMUs would stay close to the body (reducing vibration).
 A pocket over the thigh and a pocket just above the ankle were sewn on the lateral surface of the right and left legs of the garment.
 Multiple sizes of the garment were created (modeled on US 3-6 mo, 6-9 mo, 9-12 mo, and 12-18 mo sizing).
-Caregivers indicated in advance which size would fit their infant, and to ``size down'' if on the cusp to ensure a snug fit and reduce sensor movement.
+Caregivers indicated in advance which size would fit their infant, and to ``size down'' if between sizes to ensure a snug fit and minimize sensor movement.
 Each garment had a distinct pattern on the seat of the pants so that caregivers could identify front versus back and place the garment in the correct orientation.
 
-Each sensor recorded accelerometer and gyroscope data throughout the day, with sufficient battery and onboard storage to record for approximately 12 hours.
-We chose a sampling rate of 62.5 Hz (one of the available preset rates) based on prior work that used rates of 50-64 Hz \autocite{FranchakScott2021,AiraksinenRasanen2020,YaoPlotz2019}.
-Infants also wore a LENA® recorder throughout the day in the front pocket of a LENA® shirt, located near the infant's chest, to determine whether data could be simultaneously recorded from the LENA® and IMU sensors.
-However, LENA® data were not analyzed as part of the current study.
+Each sensor had sufficient battery and onboard torage to record accelerometer and gyroscope data for approximately 12 hours.
+We chose a sampling rate of 62.5 Hz (one of the available presets) based on prior work that used rates of 50-64 Hz \autocite{FranchakScott2021,AiraksinenRasanen2020,YaoPlotz2019}.
+Infants also wore a LENA® recorder throughout the day in the front pocket of a LENA® shirt, located near the infant's chest, to determine whether data could be simultaneously recorded from the LENA® and IMU sensors (LENA® data were not analyzed in the current study).
 
@@
 
-Video recordings were captured at 30 Hz using an action camera on a miniature tripod (Insta360 ONE R) that caregivers placed in the room that the infant was in.
-The proprietary ``Boosted Battery Base'' attached to the action camera to allow for a total of 3 hours of recording.
-As described below, the video recording was not continuous.
-The camera created two video files temporally separated by a gap of approximately 40-45 s.
-Caregivers also received a log sheet to record times that infants napped as well as times that the sensor garment was removed from the infant (e.g., baths, diaper changes, errands).
+Videos were captured at 30 Hz using an action camera on a miniature tripod (Insta360 ONE R) that caregivers placed in the same room as the infant.
+Although recordings lasted 3 hours, they were divided into two video files of approximately 90 minutes temporally separated by a gap of 40-45 s.
+Caregivers received a log sheet to record times that infants napped and times that the sensor garment was removed from the infant (e.g., baths, diaper changes, errands).
 
@@
 Figure \ref{fig:exemplar-timeline} shows an exemplar timeline of the entire procedure and recording periods for a single participant.
-On the day of the visit, a researcher arrived at the participant's home in the morning and prepared all the equipment at the doorstep.
-In order to synchronize all three recording devices, the researcher began by first turning on the video camera and the LENA audio device (the IMUs were already configured and placed before arrival because they required a proprietary sensor dock to begin recording in the laboratory).
-To create an easily recognizable synchronization point between the video recording and IMU data, the researcher dropped or struck the leggings (containing the IMUs) on a surface in view of the camera, as in \textcite{FranchakScott2021}.
+A researcher arrived at the participant's home in the morning and set each device to record while at the doorstep.
+To create a recognizable synchronization point between the video recording and IMU data, the researcher dropped the sensor garment containing the IMUs on a surface in view of the camera, as in \textcite{FranchakScott2021}.
 All the equipment---once recording and with synchronization information recorded---was placed inside a large bucket and left outside the family's front door.
@@
 The caregiver logged the times the sensors were removed (blank areas in the timeline in Figure \ref{fig:exemplar-timeline}) or the child took a nap (gray areas in the timeline in Figure \ref{fig:exemplar-timeline}) so that those times could be excluded from analysis.
-The following day a researcher picked up the equipment, verified that the paperwork was signed, and compensated the participant.
+The following day a researcher picked up the bucket of equipment.
 
@@
 Synchronization errors were estimated to be less than 30-60 ms (1-2 video frames).
-When the first video file ended, a second video file was automatically recorded, termed the \emph{distal comparison}.
+The second video file comprised the \emph{distal comparison}.
 This video recorded the next 90 minutes of natural activity.
-However, because of a limitation in the camera, there was a variable gap of \textasciitilde40 s between the two videos, synchronization in the distal comparison video was coarser, with estimated temporal offsets of \textasciitilde5 s in either direction.
+However, because of a limitation in the camera, there was a variable gap of \textasciitilde40 s between the two videos, so synchronization in the distal comparison video was coarser, with estimated temporal offsets of \textasciitilde5 s in either direction.
 
@@
 
-Supine was coded when the infant was lying on their back or was reclined up to a 45 degree angle.
-Supine was also coded in the rare cases when the infant was laying on their side.
-Prone was coded when the infant was lying on their stomach, was stationary supported by the hands/knees or the hands/feet, or was crawling.
+Supine was coded when the infant was lying on their back, on their side, or was reclined up to a 45 Prone was coded when the infant was lying on their stomach, was stationary supported by the hands/knees or the hands/feet, or was crawling.
 We scored sitting to include any form of the following seated positions: 1) infants sat with their buttocks on a surface, such as on the floor or a caregiver's lap, 2) infant was in a kneeling-sit position, in which their knees were on the ground with their legs tucked underneath the buttocks, and 3) infant was in a seating device, such as a high chair, that kept the torso oriented perpendicular to the ground (a reclined position, such as in a young infant's car seat, would be counted as supine).
 Upright was coded when the infant was standing or squatting on the ground with two feet or walking (regardless of whether infants' balance was assisted by a caregiver or with their hands holding onto something for support).
-When an infant was carried by a caregiver, held was coded.
+Our goal in creating a category for ``held by caregiver'' was to separate times when infants were in control of their body position from times when they were suspended in the air (rather than resting on furniture or a surface).
+Held was coded when infants were carried off of the groun.
 However, when the caregiver was sitting with the infant in their lap the infant's body position was coded as if the caregiver was a surface (e.g., if the infant was sitting on the caregiver's lap this was coded as sitting).
@@
 Classification training and prediction was conducted on a windowed dataset that summarized the raw, 62.5 Hz motion signals within 4-s windows.
-Overlapping moving windows captured 4-s (250 samples) of data starting each second, which is a common unit of analysis in prior studies of human activity classification \autocite{FranchakScott2021,NamPark2013,AiraksinenRasanen2020}.
-The 250 samples of data in each window were aggregated into a variety of motion features---summary statistics that could be fed into the machine learning model.
+Data were reduced in time by creating overlapping moving windows (4-s long, comprising 250 samples) starting each second, which is a common unit of analysis in prior studies of human activity classification \autocite{FranchakScott2021,NamPark2013,AiraksinenRasanen2020}.
+For each 4-s window, we aggregated the 250 samples to create single scores for a variety of motion features---summary statistics that could be fed into the machine learning model.
 The minimum, maximum, 25th percentile, 75th percentile, mean, median, skew, kurtosis, standard deviation, and sum were computed for each signal (e.g., right thigh X linear acceleration, left ankle pitch angular acceleration).
@@
 The resulting windowed dataset was used for machine learning classification and validation.
-For each analysis reported in the results, a subset of data were defined as a ``training'' set and another, independent portion of the data were defined as a ``testing'' set.
-Different training/testing set combinations were used to address different goals; here we describe the general procedure of how a classifier was derived from a training set.
-Random forest models \autocite{Breiman2001} were trained to predict the human coded body position label for each window from the set of 436 motion features using the \emph{randomForest} package \autocite{LiawWiener2002}.
-Random forest models use random subsets of features and random subsets of data across many iterations; this random subsetting creates models that are less likely to be overfit.
-The \emph{tidymodels} set of packages in R \autocite{tidymodels} was used to tune the ``mtry'' and ``ntrees'' hyperparameters, which control the number of columns sampled and the total number of trees created when modeling.
-We use an mtry parameter of 44 and and ntree parameter of 550; however, changing hyperparameters had little impact on model performance.
-Just like a linear regression model, the resulting random forest model could later be applied to a set of testing data with the \emph{predict} function.
-Validation analyses below will compare the human-coded annotations and random forest model predictions for testing data that are independent from training data, providing a test of whether models can successfully generalize.
+For each analysis reported in the results, a subset of data were defined as a ``training'' set and another, independent portion of the data were defined as a ``testing'' set (further described below).
+Random forest models \autocite{Breiman2001} used the training set to learn the body position label for each window from the set of 436 motion features using the \emph{randomForest} package \autocite{LiawWiener2002}.
+The resulting random forest model could later be applied to a set of testing data with the \emph{predict} function.
 
@@
 We report three sets of results based on 34 full-day testing sessions resulting in a total of 302 hours of movement recording.
-First, we focus on the accuracy of group and individual models trained and validated from data during the proximal comparison period.
-This period is unique in having the highest degree of temporal synchrony between video and motion data, allowing us to assess the accuracy of individual body position events.
-Second, we applied models trained from data during the proximal comparison to predict infant behavior during the distal comparison, when coarser synchrony between video and motion data was available.
-This distal comparison is critical because it provides, to our knowledge, the first-ever test of how well infant body classification models predict behavior that is variable, unrestricted, and unsupervised.
-Data from the first two sets of analyses confirm that body position classification models are accurate across a long period of time, suggesting that data from the entire day will be valid.
-In the third set of results we examine the data quality of full-day recordings and test whether full-day recordings capture well-established age differences in body position experiences.
 
@@
 The testing set was never used as training data.
-The two types of models differed based on what data were used during training, but the same random forest modeling procedure was used in both cases.
-A group model was trained for each session using an aggregated dataset containing the first 75\% of all \textbf{other} sessions' proximal comparison data.
+The two types of models differed based on what data were used during training: A group model was trained for each session using an aggregated dataset containing the first 75\% of all \textbf{other} sessions' proximal comparison data.
 This leave-one-out cross-validation tested the generalization of the model to a recording session that was not used at all in the training set.
@@
 Looking at the median performance suggests that the difference between models was not considerable for the typical participant (group median accuracy = 0.89; individual median accuracy = 0.93).
-Indeed, it is notable that the worst-case accuracy from group models (minimum = 0.50) was far lower than in individual models (minimum = 0.73).
-Possibly, individual models accounted for idiosyncrasies in behavior or inconsistencies in sensor placement for those sessions that group models could not generalize to.
-
-Similar to past work \autocite{AiraksinenRasanen2020}, overall accuracy decreased when fewer sensors were used.
+
+As in \textcite{AiraksinenRasanen2020}, overall accuracy decreased when fewer sensors were used.
 Table \ref{tab:metrics-reduce} compared overall accuracy for all four sensors (top row), pairs of 2 sensors (rows 2-5), and single sensors (rows 6-9).
 The highest accuracy was observed when using all four sensors (group model \emph{M} = 0.846; individual model \emph{M} = 0.916), and the lowest when using only a single sensor (left ankle group model \emph{M} = 0.667; right ankle individual model \emph{M} = 0.819).
-Some sensor pairs showed comparable accuracy to using all four sensors: Left ankle and thigh has group model accuracy of \emph{M} = NA and individual model accuracy of \emph{M} = NA, only 2.2\% and 1.5\% worse than using all four sensors.
+Accuracy for some pairs approached 4-sensor accuracy: Left ankle and thigh had group model accuracy of \emph{M} = NA and individual model accuracy of \emph{M} = NA, only 2.2\% and 1.5\% worse than using all four sensors.
 Other pairings were less accurate; notably, using both ankles resulted in group model accuracy of \emph{M} = 0.719 and individual model accuracy of \emph{M} = 0.844, 12.7\% and 7.2\% worse than using all four sensors.
@@
 Strong overall accuracy can be misleading when the relative frequency of different classes is unbalanced.
-For example, most infants spent more time sitting than prone, so good performance for sitting classification could mask poorer performance in prone classification.
 Accordingly, we report Cohen's kappa, a commonly-used metric that penalizes missing rare events (Figure \ref{fig:metrics}B), and we provide classification metrics for each individual body position (Table \ref{tab:metrics-by-class}) to account for imbalance in body position rates within and between individuals.

### 2023-06-20 manuscript.Rmd
responded to all requests

@@
 To assess the accuracy of body position classification for each recording session, we reserved the (temporally) last 25% of a session's proximal comparison data as the testing set.
-The testing set was never used as training data.
-The two types of models differed based on what data were used during training: A group model was trained for each session using an aggregated dataset containing the first 75% of all **other** sessions' proximal comparison data.
+The testing set was never used as training data, and was the same for both modeling approaches to facilitate direct comparisons between the models.
+We generated two different training datasets relative to each testing set.
+For the group model training set, we aggregated the first 75% of all **other** sessions' proximal comparison data.
 This leave-one-out cross-validation tested the generalization of the model to a recording session that was not used at all in the training set.
-In contrast, individual models were trained using the first 75% of a recording session, then tested against the (temporally) last 25% of the same session.
+The individual model training set used the first 75% of data from the testing set's session.
 The individual model tests whether earlier training data generalize to later testing data within an individual participant's recording.
-In both cases, the testing set was the same, allowing for direct comparison between the two modeling approaches.
 Figure \@ref(fig:metrics), Table \@ref(tab:metrics-table), and Table \@ref(tab:metrics-by-class) summarize the performance of group and individual models using standard metrics for classification.
@@
 Overall, the results suggest a reasonable balance between sensitivity and PPV among different body position classes for both model types.
-Table \@ref(tab:metrics-table) shows the overall sensitivity and PPV across classes.
 
@@
 
-In the first set of results, we showed that group and individual models trained from data during the proximal comparison period were accurate at classifying body position in the same recording.
+The first set of results showed that group and individual models trained from data during the proximal comparison period were within the proximal period.
 In the next analysis, we examine long-term performance by testing how accurately models trained from the proximal period could predict body position during the distal comparison period.
 A single group model was created using all sessions' proximal period training data (rather than group models leaving out a single session); the same individual models were used.
-As previously mentioned, distal videos had only coarse temporal synchrony with motion recordings (misalignment of up to ~ 5 s in either direction).
-This misalignment precluded calculating accuracy based on the proportion of matching events, so we used another technique following past work [@YaoPlotz2019; @FranchakScott2021].
-For each session, we summed the amount of time infants were predicted to be in each of the 5 body position categories from the model and compared that to the summed time for the body positions based on human coding.
-Below, we report correlations between the model-predicted and human-coded aggregated body position time across the entire distal period as well as within finer, 10-minute bins.
+Distal videos had only coarse temporal synchrony with motion recordings which precluded calculating accuracy based on the proportion of matching events.
+Instead, we summed the amount of time infants were predicted to be in each of the 5 body position categories from the model and compared that to the summed time for the body positions based on human coding [@YaoPlotz2019; @FranchakScott2021].
+
+<!-- Below, we report correlations between the model-predicted and human-coded aggregated body position time across the entire distal period as well as within finer, 10-minute bins.  -->
 
 Whereas the proximal analyses used all `r length(unique(ppts$id_uni))` sessions, this was not possible in the distal comparison.
-Because the start of the visit was scheduled during a time when the infant was awake, it was common for a nap to be needed following the proximal period.
+Because the start of the visit was scheduled during a time when the infant was awake, it was common for a nap to follow the proximal period.
 Nine sessions were excluded because the infant was either napping or otherwise not on camera during the entire 90-minute distal recording.
@@
 Visual inspection of Figure \@ref(fig:part2overall) indicated two extreme outliers, which we marked by a gray square and a gray diamond.
-We closely investigated each of the two outliers to understand why their agreement was below the standard of the other sessions.
 The "gray square" outlier had significant confusion between sitting and supine classification.
 Reviewing the video indicated that this participant spent a long period of time in a seating device that was reclined almost exactly at 45 degrees, making it difficult to determine if the infant was sitting or supine.
-The infant also spent a long time in the mother's arms in an ambiguous supine/sitting position that was hard to classify.
+The infant also spent a long time in the mother's arms in an ambiguous supine/sitting position.
 This participant's proximal accuracy was also poor because similar ambiguities appeared during the initial recording.
@@
 Overall agreement among the non-outlier sessions was excellent for both group models (*r* = `r group_agree_cor_noout`) and individual models (*r* = `r split_agree_cor_noout`).
-With the exception of held, within-class agreement was strong for the other four body positions.
 
@@
 
-Collecting real-time data that can be synchronized to other sources is a key goal in our development of long-form body position classification.
 Although overall aggregate agreement in the distal comparison was strong, it is important to show that similarly strong agreement is found within a shorter timescale.
-We repeated the agreement analysis (correlating human-coded body position time with model-predicted time) after dividing the distal comparison period into nine 10-minute bins (marked by vertical dashed lines in Figure \@ref(fig:exemplar-timeline)).
+We repeated the agreement analysis after dividing the distal comparison period into nine 10-minute bins (marked by vertical dashed lines in Figure \@ref(fig:exemplar-timeline)).
 Infants had varying numbers of 10-minute bins depending on how much time they were awake and on camera.
 Bins were included only if there was > 7 minutes of usable data.
-
 Table \@ref(tab:pt2binstable) shows the agreement correlation coefficients for group and individual models, including and excluding the two outliers identified in the previous section.
@@
 The gray shaded area marks ± 1 minute of prediction error.
-The vast majority of session-averaged predictions fall within 1 minute of error without a clear bias towards overestimation or underestimation (with the previously-identified outliers as exceptions).
+Most session-averaged predictions fall within 1 minute of error without a clear bias towards overestimation or underestimation.
 For group models, we calculated the percentage of 10-min bins across participants that had errors < 1 minute: `r group_diff["Held","prop"]`% for held, `r group_diff["Supine","prop"]`% for supine, `r group_diff["Prone","prop"]`% for prone, `r group_diff["Sitting","prop"]`% for sitting, and `r group_diff["Upright","prop"]`% for upright.
 For individual models, the percent of 10-min bins with < 1 minute of error was: `r ind_diff["Held","prop"]`% for held, `r ind_diff["Supine","prop"]`% for supine, `r ind_diff["Prone","prop"]`% for prone, `r ind_diff["Sitting","prop"]`% for sitting, and `r ind_diff["Upright","prop"]`% for upright.
-Taken together, these results suggest that most estimates across models, sessions, and body positions had small amounts of prediction error even after a long delay separating the period that models were trained and tested.
 
@@
 Predictions from the group model were used to ensure that motion data were classified consistently across all sessions.
-Moreover, for infants who did not display all 5 body positions in the training period, group models were necessary to be able to predict those behaviors across the entire day.
+Moreover, for infants who did not display all 5 body positions in the training period, group models were necessary to predict those behaviors across the entire day.
 Session start times ranged from `r start_times[1]` to `r start_times[length(start_times)]` with a median of `r start_times[floor(length(start_times)/2)]`.
-Although the equipment was always dropped off in the morning, sometimes infants were unexpectedly asleep, leading to a few sessions in which infants began wearing the sensors later than intended (such as \#11 in the younger group).
+Sometimes infants were unexpectedly asleep at the scheduled time, leading to a few sessions in which infants began wearing the sensors later than intended (such as \#11 in the younger group).
 With two exceptions, recordings lasted until the infants' bedtime.
@@
 Among participants who wore the sensors from morning to bedtime, all infants wore the sensor garment during 100% of the intended recording period, excluding the times caregivers were asked to remove the garment (naps, diaper changes, trips out of the house), based on caregiver logs of wear time.
-No caregiver mentioned needing to remove the garment for any other reason (such as infant discomfort), meaning that the majority of the time during the day either resulted in usable body position data or was excluded due to caregiver-reported naps (gray periods in Figure \@ref(fig:timelines)).
+No caregiver reporting removing the garment for any other reason (such as infant discomfort), meaning that the majority of the time during the day either resulted in usable body position data or was excluded due to caregiver-reported naps (gray periods in Figure \@ref(fig:timelines)).
 The total length of the recording period ranged from `r apa_num(min(coverage$total))`-`r apa_num(max(coverage$total))` hours with a median of `r apa_num(median(coverage$total))` hours.
@@
 Table \@ref(tab:agetable) summarizes the percent of time that younger and older infants spent in each body position out of their awake samples, as predicted by group and individual models.
-Past work using sparse EMA surveys across the day [@Survey] found that from 3-12 months, infants spent less time held, reclined, and supine but spent increasingly more time sitting and upright.
-Prone time changed little.
+EMA surveys [@Survey] found that from 3-12 months, infants spent less time held, reclined, and supine but spent increasingly more time sitting and upright.
 The most straightforward comparisons between the two investigations are upright and prone, because they were defined identically.
@@
 In the current study, prone time was `r pos_sum_lookup["Prone", "Younger_Group"]`% for younger infants and `r pos_sum_lookup["Prone", "Older_Group"]`% for older infants.
-Held, sitting, and supine are more difficult to compare because prior work [@Survey] considered held as any time that infants were in caregivers' arms or on their laps, however, we counted laying or sitting on caregivers' laps as supine and sitting, respectively.
-Further complicating the comparison, @Survey included a reclined category that we would classify as either supine or sitting (whichever it was closer to).
-Although the numbers are less directly comparable, our results are similar in showing that held and supine time was less for older infants than younger infants (Table \@ref(tab:agetable)).
+Furthermore, our results are similar in showing that held and supine time was less for older infants than younger infants (Table \@ref(tab:agetable)).
 Sitting was greater for older infants (`r pos_sum_lookup["Sitting", "Older_Group"]`%) compared with younger infants (`r pos_sum_lookup["Sitting", "Younger_Group"]`%).
 Moreover, sitting was the most frequent body position followed by upright in 12-month-olds [@Survey], just as we observed in the current study.
-Overall, the estimates from the current investigation are consistent with age differences found using a different methodology.
 
@@
 
-Across three analyses, we demonstrated the validity of long-form recordings of infant body position using wearable inertial sensors.
-Both group and individual models trained during the proximal comparison period performed well on testing data collected from the same recording period.
+Here, we demonstrated the validity of long-form recordings of infant body position using wearable inertial sensors.
+Models trained during the proximal comparison period performed well on testing data collected from the same recording period.
 Most important, accuracy was consistently strong later in the recording when behavior was completely unsupervised: Human-coded and model-predicted durations of body positions during the distal comparison period were highly correlated, even when narrowing to the scale of 10 minutes.
 Examining full recordings showed that the new method allows us to capture more data (median = `r apa_num(median(coverage$hours_recorded))` hours of awake behavior) than is typical with video methods (1-2 hours) while simultaneously reducing the human labor needed to annotate it.
-Ultimately, age differences in body position mirrored past findings that employed other methods, suggesting that the outcome of full-day body position recordings is suitable for capturing individual differences and describing developmental changes in motor behavior.
+Ultimately, age differences in body position mirrored past findings that employed other methods, suggesting that the outcome of full-day body position recordings is suitable for describing developmental changes in motor behavior.
 
@@
 A researcher applied the sensors in the lab [@FranchakScott2021; @AiraksinenRasanen2020] or home [@AiraksinenGallen2022] and set the stage for how infants and caregivers should interact.
-For example, both @FranchakScott2021 and @AiraksinenGallen2022 instructed the caregivers that the recorded sessions should mimic a "playtime", which would encourage more common activities that would be easier to classify (e.g., crawling and sitting on the floor) and discourage idiosyncratic and potentially more challenging activities (e.g., positioning in infant-specific furniture like exersaucers and walkers, sitting or lying for long periods while eating or nursing).
+For example, both @FranchakScott2021 and @AiraksinenGallen2022 instructed the caregivers that the recorded sessions should mimic "playtime", which would encourage more common activities that would be easier to classify (e.g., crawling and sitting on the floor) and discourage idiosyncratic and potentially more challenging activities (e.g., positioning in infant-specific furniture like exersaucers and walkers, sitting or lying for long periods while eating or nursing).
 Examining the timelines in Figure \@ref(fig:timelines) shows how much variability there is between and within sessions in the earlier parts of the recording that were used for training and testing.
-With the exception of the brief guided session at the start of the recording (which ensured that every infant spent at least a short amount of time in each body position), the majority of the recorded behavior was truly natural and "in the wild".
-
-The challenge only increased in the distal comparison period.
+
+The challenge increased in the distal comparison period.
 Since sessions were scheduled at convenient times for the infant and caregiver, the transition from the proximal to distal period increased the odds that infants might need a nap, a snack/meal, or engage in a less typical activity (e.g., watching TV).
 Indeed, the distal comparison in Figure \@ref(fig:exemplar-timeline) contains a long period (almost 50 minutes) where the infant sat in a high chair eating lunch.
-With such a long delay in time from when the sensor garment was first put on, there is an increasing chance that sensors moved from their ideal positions.
 Regardless of these challenges, accuracy in the distal comparison continued to be strong.
@@
 Even at a finer timescale, most errors within 10-minute bins were less than 1 minute in total for all five body positions.
-
-To put this in context, the reliability of human coders on the body position code was ~90%, putting model-predicted accuracy on a similar level to human coders.
-Another comparison we can make is to the accuracy of the LENA&reg; device's automated metrics.
-Correlations between human and model frequencies were *r* = .698 for adult word count and *r* = .649 for child word count [@CristiaLavechin2020].
-Even including outliers, correlations between human and model-predicted body position were *r* = `r group_agree_cor` (group) and *r* = `r split_agree_cor` (individual), stronger than the LENA&reg; automated counts.
-Note, however, that the LENA&reg; validation study used independently-collected data, whereas we are measuring performance for infants whose earlier data comprised the training set.
-More work is needed to know how well group models would predict a different set of infants' behavior in a different environmental and/or cultural context.
-
-In their validation study, @CristiaLavechin2020 make the recommendation that researchers using LENA&reg; manually annotate a portion of data collected for their particular use case to ensure that automated LENA&reg; metrics are reliable for the participants and environment pertinent to the research question.
-Similarly, collecting some video data for each participant in future long-form motor behavior recordings provides a way to check reliability and identify errant predictions.
-Although predictions were accurate for most sessions, we noted two outliers for whom model predictions failed to match human annotations.
-Fortunately, we could reference video recordings to understand why the errors occurred.
+For comparison, the reliability of human coders on the body position code was ~90%, putting model-predicted accuracy on a similar level to human coders.
+
 Idiosyncrasies in activities and device use make it challenging to decide *a priori* whether "odd" classifications (from the expectation of the researcher) are possible.
@@
 Although it is not feasible to collect large amounts of video data to check model predictions, interviewing caregivers about common activities and devices may provide a way to understand unexpected predictions.
-
 Counterintuitively, the two outlier participants increased our confidence that the body position classification method works as intended.
@@
 In other words, the models failed to predict a particular *event* for each of the two infants.
-Unfortunately, those events lasted a long time leading to poor overall agreement.
-In the case of confusion between supine and sitting for the infant who was reclined, the confusion likely stems from the difficulty of categorizing behaviors that do not always fit neatly into discrete classes (e.g., continuous body position angles categorized as either supine or sitting).
 
@@
 It can be difficult to elicit every behavior of interest in each infant; individual models require that every behavior that will be later predicted was displayed by each individual infant.
-In fact, group models remove the need to get *any* training data for a particular infant, meaning that they can be applied to any infant regardless of whether they have training data.
+In fact, group models remove the need to get *any* training data for a particular infant.
 Finally, applying a group model means that individual differences between infants are due to their behavior being classified in a consistent way, rather than having a separate set of rules for predicting each individual.
-However, given the variety in infant behavior, individual models are more agile in that they can capture the unique and unexpected.
+However, given the variety in infant behavior, individual models are more agile in capturing the unique and unexpected.
 Likewise, individual models can be an excellent choice for rapid prototyping and pilot testing.
@@
 Linking the distributional structure of a particular motor experience in a real day, such as time spent sitting, might be predicted by sitting skill and/or might predict infants' future sitting ability.
-
-Moreover, real-time recording of motor behavior provides a way to measure how infant motor behavior links to other types of experiences in the moment in daily life.
-As we reviewed, @MalachowskiSalo2023 used a combination of LENA&reg; and EMA surveys to find that infants heard less adult speech when restrained in seating devices.
+Moreover, real-time recording of motor behavior provides a way to measure how infant motor behavior links to other types of experiences in the moment in daily life.@MalachowskiSalo2023 used a combination of LENA&reg; and EMA surveys to find that infants heard less adult speech when restrained in seating devices.
 Combining LENA&reg; and long-form motor recordings can take this a step further by measuring more precisely how speech and body position co-vary within a day.
-Synchronizing long form motor recordings to other continuous monitors, such as wearable ECG [@WassPhillips2022] can provide information about how infant arousal relates to motor behavior.
-"TotTag"---a wearable device for measuring infant-caregiver distance [@SaloPannuto2022]---or caregiver worn activity trackers/physiology monitors could be paired with body position classification to better understand how moment-to-moment interpersonal dynamics relate to infant motor behavior.
 

### 2023-06-20 manuscript.tex
responded to all requests

@@
 Overall, the results suggest a reasonable balance between sensitivity and PPV among different body position classes for both model types.
-Table \ref{tab:metrics-table} shows the overall sensitivity and PPV across classes.
 
@@
 
-In the first set of results, we showed that group and individual models trained from data during the proximal comparison period were accurate at classifying body position in the same recording.
+The first set of results showed that group and individual models trained from data during the proximal comparison period were within the proximal period.
 In the next analysis, we examine long-term performance by testing how accurately models trained from the proximal period could predict body position during the distal comparison period.
 A single group model was created using all sessions' proximal period training data (rather than group models leaving out a single session); the same individual models were used.
-As previously mentioned, distal videos had only coarse temporal synchrony with motion recordings (misalignment of up to \textasciitilde{} 5 s in either direction).
-This misalignment precluded calculating accuracy based on the proportion of matching events, so we used another technique following past work \autocite{YaoPlotz2019,FranchakScott2021}.
-For each session, we summed the amount of time infants were predicted to be in each of the 5 body position categories from the model and compared that to the summed time for the body positions based on human coding.
-Below, we report correlations between the model-predicted and human-coded aggregated body position time across the entire distal period as well as within finer, 10-minute bins.
+Distal videos had only coarse temporal synchrony with motion recordings which precluded calculating accuracy based on the proportion of matching events.
+Instead, we summed the amount of time infants were predicted to be in each of the 5 body position categories from the model and compared that to the summed time for the body positions based on human coding \autocite{YaoPlotz2019,FranchakScott2021}.
 
 Whereas the proximal analyses used all 34 sessions, this was not possible in the distal comparison.
-Because the start of the visit was scheduled during a time when the infant was awake, it was common for a nap to be needed following the proximal period.
+Because the start of the visit was scheduled during a time when the infant was awake, it was common for a nap to follow the proximal period.
 Nine sessions were excluded because the infant was either napping or otherwise not on camera during the entire 90-minute distal recording.
@@
 Visual inspection of Figure \ref{fig:part2overall} indicated two extreme outliers, which we marked by a gray square and a gray diamond.
-We closely investigated each of the two outliers to understand why their agreement was below the standard of the other sessions.
 The ``gray square'' outlier had significant confusion between sitting and supine classification.
 Reviewing the video indicated that this participant spent a long period of time in a seating device that was reclined almost exactly at 45 degrees, making it difficult to determine if the infant was sitting or supine.
-The infant also spent a long time in the mother's arms in an ambiguous supine/sitting position that was hard to classify.
+The infant also spent a long time in the mother's arms in an ambiguous supine/sitting position.
 This participant's proximal accuracy was also poor because similar ambiguities appeared during the initial recording.
@@
 Overall agreement among the non-outlier sessions was excellent for both group models (\emph{r} = 0.95) and individual models (\emph{r} = 0.96).
-With the exception of held, within-class agreement was strong for the other four body positions.
 
@@
 
-Collecting real-time data that can be synchronized to other sources is a key goal in our development of long-form body position classification.
 Although overall aggregate agreement in the distal comparison was strong, it is important to show that similarly strong agreement is found within a shorter timescale.
-We repeated the agreement analysis (correlating human-coded body position time with model-predicted time) after dividing the distal comparison period into nine 10-minute bins (marked by vertical dashed lines in Figure \ref{fig:exemplar-timeline}).
+We repeated the agreement analysis after dividing the distal comparison period into nine 10-minute bins (marked by vertical dashed lines in Figure \ref{fig:exemplar-timeline}).
 Infants had varying numbers of 10-minute bins depending on how much time they were awake and on camera.
 Bins were included only if there was \textgreater{} 7 minutes of usable data.
-
 Table \ref{tab:pt2binstable} shows the agreement correlation coefficients for group and individual models, including and excluding the two outliers identified in the previous section.
@@
 The gray shaded area marks ± 1 minute of prediction error.
-The vast majority of session-averaged predictions fall within 1 minute of error without a clear bias towards overestimation or underestimation (with the previously-identified outliers as exceptions).
+Most session-averaged predictions fall within 1 minute of error without a clear bias towards overestimation or underestimation.
 For group models, we calculated the percentage of 10-min bins across participants that had errors \textless{} 1 minute: 94.62\% for held, 80.65\% for supine, 94.62\% for prone, 77.42\% for sitting, and 94.62\% for upright.
 For individual models, the percent of 10-min bins with \textless{} 1 minute of error was: 92.86\% for held, 88.10\% for supine, 88.10\% for prone, 83.33\% for sitting, and 96.43\% for upright.
-Taken together, these results suggest that most estimates across models, sessions, and body positions had small amounts of prediction error even after a long delay separating the period that models were trained and tested.
 
@@
 Predictions from the group model were used to ensure that motion data were classified consistently across all sessions.
-Moreover, for infants who did not display all 5 body positions in the training period, group models were necessary to be able to predict those behaviors across the entire day.
+Moreover, for infants who did not display all 5 body positions in the training period, group models were necessary to predict those behaviors across the entire day.
 Session start times ranged from 09:15 to 13:20 with a median of 10:25.
-Although the equipment was always dropped off in the morning, sometimes infants were unexpectedly asleep, leading to a few sessions in which infants began wearing the sensors later than intended (such as \#11 in the younger group).
+Sometimes infants were unexpectedly asleep at the scheduled time, leading to a few sessions in which infants began wearing the sensors later than intended (such as \#11 in the younger group).
 With two exceptions, recordings lasted until the infants' bedtime.
@@
 Among participants who wore the sensors from morning to bedtime, all infants wore the sensor garment during 100\% of the intended recording period, excluding the times caregivers were asked to remove the garment (naps, diaper changes, trips out of the house), based on caregiver logs of wear time.
-No caregiver mentioned needing to remove the garment for any other reason (such as infant discomfort), meaning that the majority of the time during the day either resulted in usable body position data or was excluded due to caregiver-reported naps (gray periods in Figure \ref{fig:timelines}).
+No caregiver reporting removing the garment for any other reason (such as infant discomfort), meaning that the majority of the time during the day either resulted in usable body position data or was excluded due to caregiver-reported naps (gray periods in Figure \ref{fig:timelines}).
 The total length of the recording period ranged from 6.92-11.75 hours with a median of 9.04 hours.
@@
 Table \ref{tab:agetable} summarizes the percent of time that younger and older infants spent in each body position out of their awake samples, as predicted by group and individual models.
-Past work using sparse EMA surveys across the day \autocite{Survey} found that from 3-12 months, infants spent less time held, reclined, and supine but spent increasingly more time sitting and upright.
-Prone time changed little.
+EMA surveys \autocite{Survey} found that from 3-12 months, infants spent less time held, reclined, and supine but spent increasingly more time sitting and upright.
 The most straightforward comparisons between the two investigations are upright and prone, because they were defined identically.
@@
 In the current study, prone time was 14.03\% for younger infants and 15.09\% for older infants.
-Held, sitting, and supine are more difficult to compare because prior work \autocite{Survey} considered held as any time that infants were in caregivers' arms or on their laps, however, we counted laying or sitting on caregivers' laps as supine and sitting, respectively.
-Further complicating the comparison, \textcite{Survey} included a reclined category that we would classify as either supine or sitting (whichever it was closer to).
-Although the numbers are less directly comparable, our results are similar in showing that held and supine time was less for older infants than younger infants (Table \ref{tab:agetable}).
+Furthermore, our results are similar in showing that held and supine time was less for older infants than younger infants (Table \ref{tab:agetable}).
 Sitting was greater for older infants (44.39\%) compared with younger infants (28.91\%).
 Moreover, sitting was the most frequent body position followed by upright in 12-month-olds \autocite{Survey}, just as we observed in the current study.
-Overall, the estimates from the current investigation are consistent with age differences found using a different methodology.
 
@@
 
-Across three analyses, we demonstrated the validity of long-form recordings of infant body position using wearable inertial sensors.
-Both group and individual models trained during the proximal comparison period performed well on testing data collected from the same recording period.
+Here, we demonstrated the validity of long-form recordings of infant body position using wearable inertial sensors.
+Models trained during the proximal comparison period performed well on testing data collected from the same recording period.
 Most important, accuracy was consistently strong later in the recording when behavior was completely unsupervised: Human-coded and model-predicted durations of body positions during the distal comparison period were highly correlated, even when narrowing to the scale of 10 minutes.
 Examining full recordings showed that the new method allows us to capture more data (median = 6.14 hours of awake behavior) than is typical with video methods (1-2 hours) while simultaneously reducing the human labor needed to annotate it.
-Ultimately, age differences in body position mirrored past findings that employed other methods, suggesting that the outcome of full-day body position recordings is suitable for capturing individual differences and describing developmental changes in motor behavior.
+Ultimately, age differences in body position mirrored past findings that employed other methods, suggesting that the outcome of full-day body position recordings is suitable for describing developmental changes in motor behavior.
 
@@
 A researcher applied the sensors in the lab \autocite{FranchakScott2021,AiraksinenRasanen2020} or home \autocite{AiraksinenGallen2022} and set the stage for how infants and caregivers should interact.
-For example, both \textcite{FranchakScott2021} and \textcite{AiraksinenGallen2022} instructed the caregivers that the recorded sessions should mimic a ``playtime'', which would encourage more common activities that would be easier to classify (e.g., crawling and sitting on the floor) and discourage idiosyncratic and potentially more challenging activities (e.g., positioning in infant-specific furniture like exersaucers and walkers, sitting or lying for long periods while eating or nursing).
+For example, both \textcite{FranchakScott2021} and \textcite{AiraksinenGallen2022} instructed the caregivers that the recorded sessions should mimic ``playtime'', which would encourage more common activities that would be easier to classify (e.g., crawling and sitting on the floor) and discourage idiosyncratic and potentially more challenging activities (e.g., positioning in infant-specific furniture like exersaucers and walkers, sitting or lying for long periods while eating or nursing).
 Examining the timelines in Figure \ref{fig:timelines} shows how much variability there is between and within sessions in the earlier parts of the recording that were used for training and testing.
-With the exception of the brief guided session at the start of the recording (which ensured that every infant spent at least a short amount of time in each body position), the majority of the recorded behavior was truly natural and ``in the wild''.
-
-The challenge only increased in the distal comparison period.
+
+The challenge increased in the distal comparison period.
 Since sessions were scheduled at convenient times for the infant and caregiver, the transition from the proximal to distal period increased the odds that infants might need a nap, a snack/meal, or engage in a less typical activity (e.g., watching TV).
 Indeed, the distal comparison in Figure \ref{fig:exemplar-timeline} contains a long period (almost 50 minutes) where the infant sat in a high chair eating lunch.
-With such a long delay in time from when the sensor garment was first put on, there is an increasing chance that sensors moved from their ideal positions.
 Regardless of these challenges, accuracy in the distal comparison continued to be strong.
@@
 Even at a finer timescale, most errors within 10-minute bins were less than 1 minute in total for all five body positions.
-
-To put this in context, the reliability of human coders on the body position code was \textasciitilde90\%, putting model-predicted accuracy on a similar level to human coders.
-Another comparison we can make is to the accuracy of the LENA® device's automated metrics.
-Correlations between human and model frequencies were \emph{r} = .698 for adult word count and \emph{r} = .649 for child word count \autocite{CristiaLavechin2020}.
-Even including outliers, correlations between human and model-predicted body position were \emph{r} = 0.80 (group) and \emph{r} = 0.94 (individual), stronger than the LENA® automated counts.
-Note, however, that the LENA® validation study used independently-collected data, whereas we are measuring performance for infants whose earlier data comprised the training set.
-More work is needed to know how well group models would predict a different set of infants' behavior in a different environmental and/or cultural context.
-
-In their validation study, \textcite{CristiaLavechin2020} make the recommendation that researchers using LENA® manually annotate a portion of data collected for their particular use case to ensure that automated LENA® metrics are reliable for the participants and environment pertinent to the research question.
-Similarly, collecting some video data for each participant in future long-form motor behavior recordings provides a way to check reliability and identify errant predictions.
-Although predictions were accurate for most sessions, we noted two outliers for whom model predictions failed to match human annotations.
-Fortunately, we could reference video recordings to understand why the errors occurred.
+For comparison, the reliability of human coders on the body position code was \textasciitilde90\%, putting model-predicted accuracy on a similar level to human coders.
+
 Idiosyncrasies in activities and device use make it challenging to decide \emph{a priori} whether ``odd'' classifications (from the expectation of the researcher) are possible.
@@
 In other words, the models failed to predict a particular \emph{event} for each of the two infants.
-Unfortunately, those events lasted a long time leading to poor overall agreement.
-In the case of confusion between supine and sitting for the infant who was reclined, the confusion likely stems from the difficulty of categorizing behaviors that do not always fit neatly into discrete classes (e.g., continuous body position angles categorized as either supine or sitting).
 
@@
 It can be difficult to elicit every behavior of interest in each infant; individual models require that every behavior that will be later predicted was displayed by each individual infant.
-In fact, group models remove the need to get \emph{any} training data for a particular infant, meaning that they can be applied to any infant regardless of whether they have training data.
+In fact, group models remove the need to get \emph{any} training data for a particular infant.
 Finally, applying a group model means that individual differences between infants are due to their behavior being classified in a consistent way, rather than having a separate set of rules for predicting each individual.
-However, given the variety in infant behavior, individual models are more agile in that they can capture the unique and unexpected.
+However, given the variety in infant behavior, individual models are more agile in capturing the unique and unexpected.
 Likewise, individual models can be an excellent choice for rapid prototyping and pilot testing.
@@
 Linking the distributional structure of a particular motor experience in a real day, such as time spent sitting, might be predicted by sitting skill and/or might predict infants' future sitting ability.
-
-Moreover, real-time recording of motor behavior provides a way to measure how infant motor behavior links to other types of experiences in the moment in daily life.
-As we reviewed, \textcite{MalachowskiSalo2023} used a combination of LENA® and EMA surveys to find that infants heard less adult speech when restrained in seating devices.
+Moreover, real-time recording of motor behavior provides a way to measure how infant motor behavior links to other types of experiences in the moment in daily life.@MalachowskiSalo2023 used a combination of LENA® and EMA surveys to find that infants heard less adult speech when restrained in seating devices.
 Combining LENA® and long-form motor recordings can take this a step further by measuring more precisely how speech and body position co-vary within a day.
-Synchronizing long form motor recordings to other continuous monitors, such as wearable ECG \autocite{WassPhillips2022} can provide information about how infant arousal relates to motor behavior.
-``TotTag''---a wearable device for measuring infant-caregiver distance \autocite{SaloPannuto2022}---or caregiver worn activity trackers/physiology monitors could be paired with body position classification to better understand how moment-to-moment interpersonal dynamics relate to infant motor behavior.
 

### 2023-06-21 manuscript.tex
submitted

@@
 To assess the accuracy of body position classification for each recording session, we reserved the (temporally) last 25\% of a session's proximal comparison data as the testing set.
-The testing set was never used as training data.
-The two types of models differed based on what data were used during training: A group model was trained for each session using an aggregated dataset containing the first 75\% of all \textbf{other} sessions' proximal comparison data.
+The testing set was never used as training data, and was the same for both modeling approaches to facilitate direct comparisons between the models.
+We generated two different training datasets relative to each testing set.
+For the group model training set, we aggregated the first 75\% of all \textbf{other} sessions' proximal comparison data.
 This leave-one-out cross-validation tested the generalization of the model to a recording session that was not used at all in the training set.
-In contrast, individual models were trained using the first 75\% of a recording session, then tested against the (temporally) last 25\% of the same session.
+The individual model training set used the first 75\% of data from the testing set's session.
 The individual model tests whether earlier training data generalize to later testing data within an individual participant's recording.
-In both cases, the testing set was the same, allowing for direct comparison between the two modeling approaches.
 Figure \ref{fig:metrics}, Table \ref{tab:metrics-table}, and Table \ref{tab:metrics-by-class} summarize the performance of group and individual models using standard metrics for classification.
@@
 
-\hypertarget{references}{%
-\section{References}\label{references}}
-
 \hypertarget{refs}{}

### 2023-08-14 manuscript.Rmd
Working on revision

@@
 Most often, an experimenter with a handheld camera follows infants from room to room to ensure that their movements are visible throughout the recording session [@Karasik2011; @ChenSchneider2022; @HerzbergFletcher2021].
-The primary advantage of video recording is that it captures real-time behavior.Standard 30 Hz video recording is adequate to capture changes in infant body position that occur on the timescale of seconds.
+The primary advantage of video recording is that it captures real-time behavior.
+Standard 30 Hz video recording is adequate to capture changes in infant body position that occur on the timescale of seconds.
 However, requiring an experimenter to operate a camera is obtrusive, whereas relying on a stationary camera means that infants will be absent from view as they move from place to place.
@@
 Data processing algorithms are needed to classify the raw sensor data (i.e., linear and angular acceleration time series) into meaningful categories.
-Categorizing body position---supine on the back, prone on the belly, sitting, upright, or held off the ground by a caregiver---is a complex because movement can vary greatly *within* a body position.
+Categorizing body position---supine on the back, prone on the belly, sitting, upright, or held off the ground by a caregiver---is complex because movement can vary greatly *within* a body position.
 An upright infant can be standing still or can be walking briskly across the room.
@@
 Moreover, classifiers must be able to detect new variants of behaviors that might arise over the course of a real day; it is impossible for researchers to gather training data for every possible variation that might occur.
-Thus, our central question is whether models trained on video-recorded observations at the beginning of the day generalize to predict behavior at a later time?
+Thus, our central question is whether models trained on video-recorded observations at the beginning of the day generalize to predict behavior at a later time.
 Assessing the validity of temporally *distal* periods---when infants and caregivers are unsupervised and free to follow their everyday routines---is a crucial step to establish whether automatic classification can be used to measure body position across a day.
@@
 Four inertial movement units (IMUs) were used to record infant movement across the day (MC10 Biostamp).
-A custom garment was designed to hold the IMUs: Internal pockets were sewn into snug-fitting infant leggings so that IMUs would stay close to the body (reducing vibration).
+A custom garment was designed to hold the IMUs (Figure \@ref(fig:garment)).
+Internal pockets were sewn into snug-fitting infant leggings so that IMUs would stay close to the body (reducing vibration).
 A pocket over the thigh and a pocket just above the ankle were sewn on the lateral surface of the right and left legs of the garment.
@@
 
-Each sensor had sufficient battery and onboard torage to record accelerometer and gyroscope data for approximately 12 hours.
+Each sensor had sufficient battery and onboard storage to record accelerometer and gyroscope data for approximately 12 hours.
 We chose a sampling rate of 62.5 Hz (one of the available presets) based on prior work that used rates of 50-64 Hz [@FranchakScott2021; @AiraksinenRasanen2020; @YaoPlotz2019].
@@
 
-Supine was coded when the infant was lying on their back, on their side, or was reclined up to a 45  Prone was coded when the infant was lying on their stomach, was stationary supported by the hands/knees or the hands/feet, or was crawling.
+Supine was coded when the infant was lying on their back, on their side, or was reclined up to a 45 degree angle.
+Prone was coded when the infant was lying on their stomach, was stationary supported by the hands/knees or the hands/feet, or was crawling.
 We scored sitting to include any form of the following seated positions: 1) infants sat with their buttocks on a surface, such as on the floor or a caregiver’s lap, 2) infant was in a kneeling-sit position, in which their knees were on the ground with their legs tucked underneath the buttocks, and 3) infant was in a seating device, such as a high chair, that kept the torso oriented perpendicular to the ground (a reclined position, such as in a young infant's car seat, would be counted as supine).
@@
 Our goal in creating a category for "held by caregiver" was to separate times when infants were in control of their body position from times when they were suspended in the air (rather than resting on furniture or a surface).
-Held was coded when infants were carried off of the groun.
+Held was coded when infants were carried off of the ground.
 However, when the caregiver was sitting with the infant in their lap the infant’s body position was coded as if the caregiver was a surface (e.g., if the infant was sitting on the caregiver’s lap this was coded as sitting).
@@
 For each 4-s window, we aggregated the 250 samples to create single scores for a variety of motion features---summary statistics that could be fed into the machine learning model.
-The minimum, maximum, 25th percentile, 75th percentile, mean, median, skew, kurtosis, standard deviation, and sum were computed for each signal (e.g., right thigh X linear acceleration, left ankle pitch angular acceleration).
+The minimum, maximum, 25th percentile, 75th percentile, mean, median, skew, kurtosis, standard deviation, and sum were computed for each signal (e.g., right thigh linear acceleration along the X-axis, left ankle pitch angular acceleration).
 The 10 summary statistics and 24 sensor signals generated 240 columns of motion features that described movement within each window.
@@
 Although overall accuracy from our semi-supervised, in-home data collection did not match the near-perfect accuracy  (.95-.98) found in prior in-lab studies [@FranchakScott2021; @AiraksinenRasanen2020], both models approached the level of agreement found between two human coders (*M* = .906).
-Most likely, lower accuracy in the current study results from the more variable and complex behavior observed in a a semi-supervised setting rather than from a difference in the quality of the classification model; although the first part of the proximal period was guided by the experimenter during a brief phone call, the remainder of the proximal period included natural behavior.
+Most likely, lower accuracy in the current study results from the more variable and complex behavior observed in a semi-supervised setting rather than from a difference in the quality of the classification model; although the first part of the proximal period was guided by the experimenter during a brief phone call, the remainder of the proximal period included natural behavior.
 Visual inspection of Figure \@ref(fig:metrics)A shows that accuracy values were heavily skewed, with many approaching perfect accuracy but a few sessions with very poor accuracy.
@@
 Among participants who wore the sensors from morning to bedtime, all infants wore the sensor garment during 100% of the intended recording period, excluding the times caregivers were asked to remove the garment (naps, diaper changes, trips out of the house), based on caregiver logs of wear time.
-No caregiver reporting removing the garment for any other reason (such as infant discomfort), meaning that the majority of the time during the day either resulted in usable body position data or was excluded due to caregiver-reported naps (gray periods in Figure \@ref(fig:timelines)).
+No caregiver reported removing the garment for any other reason (such as infant discomfort), meaning that the majority of the time during the day either resulted in usable body position data or was excluded due to caregiver-reported naps (gray periods in Figure \@ref(fig:timelines)).
 The total length of the recording period ranged from `r apa_num(min(coverage$total))`-`r apa_num(max(coverage$total))` hours with a median of `r apa_num(median(coverage$total))` hours.
@@
 Age in months and upright time were significantly, positively correlated using group (*r* = `r age_cor_group["age","upright_time"]`, *p* `r apa_p(age_cor_group_pval["age","upright_time"], add_equals = T)`) and individual models (*r* = `r age_cor_ind["age","upright_time"]`, *p* `r apa_p(age_cor_ind_pval["age","upright_time"], add_equals = T)`).
-Similarly, age predicted increases in sitting time using group (*r* = `r age_cor_group["age","sit_time"]`, *p* `r apa_p(age_cor_group_pval["age","sit_time"], add_equals = T)`) and individual models (*r* = `r age_cor_ind["age","sit_time"]`, *p* `r apa_p(age_cor_ind_pval["age","sit_time"], add_equals = T)`).
-In contrast, age predicted decreases in supine time using group (*r* = `r age_cor_group["age","supine_time"]`, *p* `r apa_p(age_cor_group_pval["age","supine_time"], add_equals = T)`) and individual models (*r* = `r age_cor_ind["age","supine_time"]`, *p* `r apa_p(age_cor_ind_pval["age","supine_time"], add_equals = T)`).
+Similarly, sitting time increased with age based on predictions from both group (*r* = `r age_cor_group["age","sit_time"]`, *p* `r apa_p(age_cor_group_pval["age","sit_time"], add_equals = T)`) and individual models (*r* = `r age_cor_ind["age","sit_time"]`, *p* `r apa_p(age_cor_ind_pval["age","sit_time"], add_equals = T)`).
+In contrast, supine time decreased with age using predictions from both group (*r* = `r age_cor_group["age","supine_time"]`, *p* `r apa_p(age_cor_group_pval["age","supine_time"], add_equals = T)`) and individual models (*r* = `r age_cor_ind["age","supine_time"]`, *p* `r apa_p(age_cor_ind_pval["age","supine_time"], add_equals = T)`).
 No clear age trends were found for prone time.
@@
 Linking the distributional structure of a particular motor experience in a real day, such as time spent sitting, might be predicted by sitting skill and/or might predict infants' future sitting ability.
-Moreover, real-time recording of motor behavior provides a way to measure how infant motor behavior links to other types of experiences in the moment in daily life.@MalachowskiSalo2023 used a combination of LENA&reg; and EMA surveys to find that infants heard less adult speech when restrained in seating devices.
+Moreover, real-time recording of motor behavior provides a way to measure how infant motor behavior links to other types of experiences in the moment in daily life.
+@MalachowskiSalo2023 used a combination of LENA&reg; and EMA surveys to find that infants heard less adult speech when restrained in seating devices.
 Combining LENA&reg; and long-form motor recordings can take this a step further by measuring more precisely how speech and body position co-vary within a day.
@@
 
-# References
+<!-- # References -->
 

### 2023-08-14 manuscript.tex
Working on revision

@@
 \usepackage{amsmath,amssymb}
-\usepackage{lmodern}
 \usepackage{iftex}
@@
 \else % if luatex or xetex
-  \usepackage{unicode-math}
+  \usepackage{unicode-math} % this also loads fontspec
   \defaultfontfeatures{Scale=MatchLowercase}
   \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
+\fi
+\usepackage{lmodern}
+\ifPDFTeX\else
+  % xetex/luatex font selection
 \fi
@@
 \fi
-\newlength{\cslhangindent}
-\setlength{\cslhangindent}{1.5em}
-\newlength{\csllabelwidth}
-\setlength{\csllabelwidth}{3em}
-\newlength{\cslentryspacingunit} % times entry-spacing
-\setlength{\cslentryspacingunit}{\parskip}
-\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
- {% don't indent paragraphs
-  \setlength{\parindent}{0pt}
-  % turn on hanging indent if param 1 is 1
-  \ifodd #1
-  \let\oldpar\par
-  \def\par{\hangindent=\cslhangindent\oldpar}
-  \fi
-  % set entry spacing
-  \setlength{\parskip}{#2\cslentryspacingunit}
- }%
- {}
-\usepackage{calc}
-\newcommand{\CSLBlock}[1]{#1\hfill\break}
-\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
-\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
-\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
 \ifLuaTeX
@@
 \IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
-\urlstyle{same} % disable monospaced font for URLs
+\urlstyle{same}
 \hypersetup{
@@
 Most often, an experimenter with a handheld camera follows infants from room to room to ensure that their movements are visible throughout the recording session \autocite{Karasik2011,ChenSchneider2022,HerzbergFletcher2021}.
-The primary advantage of video recording is that it captures real-time behavior.Standard 30 Hz video recording is adequate to capture changes in infant body position that occur on the timescale of seconds.
+The primary advantage of video recording is that it captures real-time behavior.
+Standard 30 Hz video recording is adequate to capture changes in infant body position that occur on the timescale of seconds.
 However, requiring an experimenter to operate a camera is obtrusive, whereas relying on a stationary camera means that infants will be absent from view as they move from place to place.
@@
 Data processing algorithms are needed to classify the raw sensor data (i.e., linear and angular acceleration time series) into meaningful categories.
-Categorizing body position---supine on the back, prone on the belly, sitting, upright, or held off the ground by a caregiver---is a complex because movement can vary greatly \emph{within} a body position.
+Categorizing body position---supine on the back, prone on the belly, sitting, upright, or held off the ground by a caregiver---is complex because movement can vary greatly \emph{within} a body position.
 An upright infant can be standing still or can be walking briskly across the room.
@@
 Moreover, classifiers must be able to detect new variants of behaviors that might arise over the course of a real day; it is impossible for researchers to gather training data for every possible variation that might occur.
-Thus, our central question is whether models trained on video-recorded observations at the beginning of the day generalize to predict behavior at a later time?
+Thus, our central question is whether models trained on video-recorded observations at the beginning of the day generalize to predict behavior at a later time.
 Assessing the validity of temporally \emph{distal} periods---when infants and caregivers are unsupervised and free to follow their everyday routines---is a crucial step to establish whether automatic classification can be used to measure body position across a day.
@@
 Four inertial movement units (IMUs) were used to record infant movement across the day (MC10 Biostamp).
-A custom garment was designed to hold the IMUs: Internal pockets were sewn into snug-fitting infant leggings so that IMUs would stay close to the body (reducing vibration).
+A custom garment was designed to hold the IMUs (Figure \ref{fig:garment}).
+Internal pockets were sewn into snug-fitting infant leggings so that IMUs would stay close to the body (reducing vibration).
 A pocket over the thigh and a pocket just above the ankle were sewn on the lateral surface of the right and left legs of the garment.
@@
 
-Each sensor had sufficient battery and onboard torage to record accelerometer and gyroscope data for approximately 12 hours.
+Each sensor had sufficient battery and onboard storage to record accelerometer and gyroscope data for approximately 12 hours.
 We chose a sampling rate of 62.5 Hz (one of the available presets) based on prior work that used rates of 50-64 Hz \autocite{FranchakScott2021,AiraksinenRasanen2020,YaoPlotz2019}.
@@
 
-Supine was coded when the infant was lying on their back, on their side, or was reclined up to a 45 Prone was coded when the infant was lying on their stomach, was stationary supported by the hands/knees or the hands/feet, or was crawling.
+Supine was coded when the infant was lying on their back, on their side, or was reclined up to a 45 degree angle.
+Prone was coded when the infant was lying on their stomach, was stationary supported by the hands/knees or the hands/feet, or was crawling.
 We scored sitting to include any form of the following seated positions: 1) infants sat with their buttocks on a surface, such as on the floor or a caregiver's lap, 2) infant was in a kneeling-sit position, in which their knees were on the ground with their legs tucked underneath the buttocks, and 3) infant was in a seating device, such as a high chair, that kept the torso oriented perpendicular to the ground (a reclined position, such as in a young infant's car seat, would be counted as supine).
@@
 Our goal in creating a category for ``held by caregiver'' was to separate times when infants were in control of their body position from times when they were suspended in the air (rather than resting on furniture or a surface).
-Held was coded when infants were carried off of the groun.
+Held was coded when infants were carried off of the ground.
 However, when the caregiver was sitting with the infant in their lap the infant's body position was coded as if the caregiver was a surface (e.g., if the infant was sitting on the caregiver's lap this was coded as sitting).
@@
 Although overall accuracy from our semi-supervised, in-home data collection did not match the near-perfect accuracy (.95-.98) found in prior in-lab studies \autocite{FranchakScott2021,AiraksinenRasanen2020}, both models approached the level of agreement found between two human coders (\emph{M} = .906).
-Most likely, lower accuracy in the current study results from the more variable and complex behavior observed in a a semi-supervised setting rather than from a difference in the quality of the classification model; although the first part of the proximal period was guided by the experimenter during a brief phone call, the remainder of the proximal period included natural behavior.
+Most likely, lower accuracy in the current study results from the more variable and complex behavior observed in a semi-supervised setting rather than from a difference in the quality of the classification model; although the first part of the proximal period was guided by the experimenter during a brief phone call, the remainder of the proximal period included natural behavior.
 Visual inspection of Figure \ref{fig:metrics}A shows that accuracy values were heavily skewed, with many approaching perfect accuracy but a few sessions with very poor accuracy.
@@
 Among participants who wore the sensors from morning to bedtime, all infants wore the sensor garment during 100\% of the intended recording period, excluding the times caregivers were asked to remove the garment (naps, diaper changes, trips out of the house), based on caregiver logs of wear time.
-No caregiver reporting removing the garment for any other reason (such as infant discomfort), meaning that the majority of the time during the day either resulted in usable body position data or was excluded due to caregiver-reported naps (gray periods in Figure \ref{fig:timelines}).
+No caregiver reported removing the garment for any other reason (such as infant discomfort), meaning that the majority of the time during the day either resulted in usable body position data or was excluded due to caregiver-reported naps (gray periods in Figure \ref{fig:timelines}).
 The total length of the recording period ranged from 6.92-11.75 hours with a median of 9.04 hours.
@@
 Age in months and upright time were significantly, positively correlated using group (\emph{r} = 0.65, \emph{p} \textless{} .001) and individual models (\emph{r} = 0.52, \emph{p} = .002).
-Similarly, age predicted increases in sitting time using group (\emph{r} = 0.57, \emph{p} \textless{} .001) and individual models (\emph{r} = 0.72, \emph{p} \textless{} .001).
-In contrast, age predicted decreases in supine time using group (\emph{r} = -0.61, \emph{p} \textless{} .001) and individual models (\emph{r} = -0.61, \emph{p} \textless{} .001).
+Similarly, sitting time increased with age based on predictions from both group (\emph{r} = 0.57, \emph{p} \textless{} .001) and individual models (\emph{r} = 0.72, \emph{p} \textless{} .001).
+In contrast, supine time decreased with age using predictions from both group (\emph{r} = -0.61, \emph{p} \textless{} .001) and individual models (\emph{r} = -0.61, \emph{p} \textless{} .001).
 No clear age trends were found for prone time.
@@
 Linking the distributional structure of a particular motor experience in a real day, such as time spent sitting, might be predicted by sitting skill and/or might predict infants' future sitting ability.
-Moreover, real-time recording of motor behavior provides a way to measure how infant motor behavior links to other types of experiences in the moment in daily life.@MalachowskiSalo2023 used a combination of LENA® and EMA surveys to find that infants heard less adult speech when restrained in seating devices.
+Moreover, real-time recording of motor behavior provides a way to measure how infant motor behavior links to other types of experiences in the moment in daily life.
+\textcite{MalachowskiSalo2023} used a combination of LENA® and EMA surveys to find that infants heard less adult speech when restrained in seating devices.
 Combining LENA® and long-form motor recordings can take this a step further by measuring more precisely how speech and body position co-vary within a day.
@@
 
-\hypertarget{refs}{}
-\begin{CSLReferences}{0}{0}
-\end{CSLReferences}
-
 \printbibliography

### 2023-08-14 manuscript.Rmd
Working on revision

@@
 #9908  9909  9910  9911 10201 10203 10204 10301 10501 10601 10701 10702 10703 10704 10802 10901 11001 11002 11004 11102 11103 11104 11202 11501 11601 11901 12001 12002 12101 12302 12303 12501 12502 13001
+# repeat sessions - younger = 3
+# repeat sessions - older = 4
 ```
@@
 Sitting on a caregiver's lap without the need to maintain balance means that the legs can dangle and the torso can lean in different directions.
-Sitting in a high chair or car seat reduces the magnitude of postural sway within sitting, and creates even more possibilities for how the arms and legs may move relative to the torso.
+Sitting in a high chair or car seat likely restricts the range of torso orientations compared with sitting independently or on a caregiver's lap.
 Finally, caregivers frequently pick up and transport infants, creating motion signals that need to be differentiated from independent activity [@KwonZavos2019; @PatelShi2019].
@@
 age <- ppts %>% group_by(group) %>% get_summary_stats(age) %>% column_to_rownames("group")
+
 ```
@@
 Families were recruited through social media advertisements and from community events in Southern California.
-Most infants were reported to be either Hispanic and White or Non-Hispanic and White.
+Most infants were reported to be either Hispanic and White (*n* = 8) or Non-Hispanic and White (*n* = 7).
 Families were compensated \$30 for every home recording session they completed.
@@
 
-Most participants were tested in a single session (*n* = `r nrow(num_sessions %>% filter(n == 1))`), but `r nrow(num_sessions %>% filter(n > 1))` participants contributed between 2-4 sessions as part of an ongoing longitudinal study.
+Most participants were tested in a single session (*n* = `r nrow(num_sessions %>% filter(n == 1))`), but `r nrow(num_sessions %>% filter(n > 1))` participants contributed between 2-4 sessions as part of an ongoing longitudinal study (three from the younger group and four from the older group).
 Although including more data from some participants could over-represent their characteristics in the model, we reasoned that this drawback was outweighed by having more available data to use for training.
@@
 The highest accuracy was observed when using all four sensors (group model *M* = `r metrics_sensor["All", "mean_group"]`; individual model *M* = `r metrics_sensor["All", "mean_split"]`), and the lowest when using only a single sensor (left ankle group model *M* = `r metrics_sensor["Left Ankle", "mean_group"]`; right ankle individual model *M* = `r metrics_sensor["Right Ankle", "mean_split"]`).
-Accuracy for some pairs approached 4-sensor accuracy: Left ankle and thigh had group model accuracy of *M* = `r metrics_sensor["Left Thigh/Ankle", "mean_group"]` and individual model accuracy of *M* = `r metrics_sensor["Left Thigh/Ankle", "mean_split"]`, only 2.2% and 1.5% worse than using all four sensors.
+Accuracy for some pairs approached 4-sensor accuracy: Left ankle and thigh had group model accuracy of *M* = `r metrics_sensor["Left Hip/Ankle", "mean_group"]` and individual model accuracy of *M* = `r metrics_sensor["Left Hip/Ankle", "mean_split"]`, only 2.2% and 1.5% worse than using all four sensors.
 Other pairings were less accurate; notably, using both ankles resulted in group model accuracy of *M* = `r metrics_sensor["Both Ankles", "mean_group"]` and individual model accuracy of *M* = `r metrics_sensor["Both Ankles", "mean_split"]`, 12.7% and 7.2% worse than using all four sensors.
@@
   ds %>% drop_na(prop) %>% 
-  ggplot() + facet_wrap("position", as.table = T, ncol = 2) + 
+  ggplot() + facet_wrap("position", as.table = T, ncol = 2, scales = "free_x") + 
   geom_smooth(aes(x = age, y = prop), color = "#191919", method = "lm", se = T) + 
@@
   scale_x_continuous(name = "Age (months)", breaks = seq(3, 15, 3), limits = c(3,15)) + 
-  ylim(0,1) + ylab("Proportion of waking day") + 
+    scale_y_continuous(name = "Proportion of waking day", breaks = seq(0, 1, .25), limits = c(0,1)) + 
   theme(panel.spacing.x = unit(.5, "lines")) +

### 2023-08-14 manuscript.tex
Working on revision

@@
 Sitting on a caregiver's lap without the need to maintain balance means that the legs can dangle and the torso can lean in different directions.
-Sitting in a high chair or car seat reduces the magnitude of postural sway within sitting, and creates even more possibilities for how the arms and legs may move relative to the torso.
+Sitting in a high chair or car seat likely restricts the range of torso orientations compared with sitting independently or on a caregiver's lap.
 Finally, caregivers frequently pick up and transport infants, creating motion signals that need to be differentiated from independent activity \autocite{KwonZavos2019,PatelShi2019}.
@@
 Families were recruited through social media advertisements and from community events in Southern California.
-Most infants were reported to be either Hispanic and White or Non-Hispanic and White.
+Most infants were reported to be either Hispanic and White (\emph{n} = 8) or Non-Hispanic and White (\emph{n} = 7).
 Families were compensated \$30 for every home recording session they completed.
@@
 
-Most participants were tested in a single session (\emph{n} = 15), but 7 participants contributed between 2-4 sessions as part of an ongoing longitudinal study.
+Most participants were tested in a single session (\emph{n} = 15), but 7 participants contributed between 2-4 sessions as part of an ongoing longitudinal study (three from the younger group and four from the older group).
 Although including more data from some participants could over-represent their characteristics in the model, we reasoned that this drawback was outweighed by having more available data to use for training.
@@
 The highest accuracy was observed when using all four sensors (group model \emph{M} = 0.846; individual model \emph{M} = 0.916), and the lowest when using only a single sensor (left ankle group model \emph{M} = 0.667; right ankle individual model \emph{M} = 0.819).
-Accuracy for some pairs approached 4-sensor accuracy: Left ankle and thigh had group model accuracy of \emph{M} = NA and individual model accuracy of \emph{M} = NA, only 2.2\% and 1.5\% worse than using all four sensors.
+Accuracy for some pairs approached 4-sensor accuracy: Left ankle and thigh had group model accuracy of \emph{M} = 0.824 and individual model accuracy of \emph{M} = 0.901, only 2.2\% and 1.5\% worse than using all four sensors.
 Other pairings were less accurate; notably, using both ankles resulted in group model accuracy of \emph{M} = 0.719 and individual model accuracy of \emph{M} = 0.844, 12.7\% and 7.2\% worse than using all four sensors.

### 2023-08-14 manuscript.Rmd
Working on revision

@@
 abstract: |
-  Long form audio recordings have had a transformational effect on the study of infant language acquisition by using mobile, unobtrusive devices to gather full-day, real-time data that can be automatically scored.
+  Long-form audio recordings have had a transformational effect on the study of infant language acquisition by using mobile, unobtrusive devices to gather full-day, real-time data that can be automatically scored.
 How can we produce similar data in service of measuring infants' everyday motor behaviors, such as body position?
-The aim of the current study was to validate long form recordings of infant position (supine, prone, sitting, upright, held by caregiver) based on machine learning classification of data from inertial sensors worn on infants' ankles and thighs.
+The aim of the current study was to validate long-form recordings of infant position (supine, prone, sitting, upright, held by caregiver) based on machine learning classification of data from inertial sensors worn on infants' ankles and thighs.
 Using over 100 hours of video recordings synchronized with inertial sensor data from infants in their homes, we demonstrate that body position classifications are sufficiently accurate to measure infant behavior.
@@
 Next, we show that the inertial sensing method has convergent validity by replicating age differences in body position found using other methods with full-day data captured from inertial sensors.
-We end the paper with a discussion of the novel opportunities that long form motor recordings afford for understanding infant learning and development.
+We end the paper with a discussion of the novel opportunities that long-form motor recordings afford for understanding infant learning and development.
 
@@
 Finally, we discuss the advantages of using inertial sensing to classify motor behavior.
-Despite promising past results in brief, supervised sessions [@FranchakScott2021; @AiraksinenRasanen2020; @AiraksinenGallen2022], the current investigation takes a needed step forward by testing accuracy over long, unsupervised recordings in what we term a *distal comparison*.
+Despite promising past results in brief, supervised sessions [@FranchakScott2021; @AiraksinenRasanen2020; @GreenspanCunha2021; @AiraksinenGallen2022], the current investigation takes a needed step forward by testing accuracy over long, unsupervised recordings in what we term a *distal comparison*.
 
@@
 For other metrics, such as the number of "conversational turns" between the child and communicative partners, agreement was poor (*r* = .364).
-Thus, for some use cases (and for some metrics), long form audio recordings provide a mobile, unobtrusive way to automatically score real-time data over a full day.
+Thus, for some use cases (and for some metrics), long-form audio recordings provide a mobile, unobtrusive way to automatically score real-time data over a full day.
 <!-- In the remainder of the paper, we turn to the question of how to replicate those key features in long form recordings of motor behavior. -->
@@
 Video and survey methods are the current state-of-the-art in assessing infants' gross motor behavior in naturalistic tasks.
-Although each method has complementary advantages and disadvantages for characterizing infants' everyday motor experiences, neither method fulfills all five key features of long form audio recordings reviewed in the previous section.
+Although each method has complementary advantages and disadvantages for characterizing infants' everyday motor experiences, neither method fulfills all five key features of long-form audio recordings reviewed in the previous section.
 
@@
 
-Measuring infant movement with inertial movement units (IMUs) is a promising avenue for long form recordings of motor behavior in the home [@CliffReilly2009; @Barbaro2019; @BruijnsTruelove2020; @LoboHall2019].
+Measuring infant movement with inertial movement units (IMUs) is a promising avenue for long-form recordings of motor behavior in the home [@CliffReilly2009; @Barbaro2019; @BruijnsTruelove2020; @LoboHall2019].
 Lightweight sensors (10-30 g) can be embedded in garments to make recordings fully *mobile*, and they are *unobtrusive* because they do not require a researcher to follow with a camcorder.
@@
 Modern approaches to human activity recognition have used machine learning to classify activity categories based on features derived from IMU data in adults [@Preece2009; @Arif2015], children [@NamPark2013; @RenDing2016; @StewartNarayanan2018], and infants [@YaoPlotz2019; @FranchakScott2021; @AiraksinenRasanen2020].
-Three prior investigations have used different machine learning techniques to categorize infant body position from IMUs towards the goal of collecting full-day data.
+Three prior investigations have used different machine learning techniques to categorize infant body position from multiple IMUs towards the goal of collecting full-day data.
 @AiraksinenRasanen2020 tested 4- to 8-month-olds in a laboratory visit with a 4-sensor array (one on each thigh and one below each shoulder), and found 95% accuracy in distinguishing between body position categories that crawling infants could perform on the floor (excluding times that infants were held by caregivers).
 Using a wider age range of 6-18 months, @FranchakScott2021 found 98% accuracy (*kappa* = 95%) in a laboratory validation study with a 3-sensor array (ankle, knee, thigh on a single leg) in categorizing body position that included infants who could both crawl and walk and also included a category for caregiver holding.
-Most recently, @AiraksinenGallen2022 conducted a validation study of body position classification in either a home or clinic testing 4- to 19-month-olds with their 4-sensor system, refining their previous method to detect moments that infants were carried by caregivers.
+Most recently, @AiraksinenGallen2022 conducted a validation study of body position classification in either a home or clinic testing 4- to 19-month-olds with a 4-sensor system, refining their previous method to detect moments that infants were carried by caregivers.
 Classification accuracy did not vary between lab and home settings, and was generally high (95%, *kappa* = .93).
-Although all three studies yielded promising classification accuracy, accuracy was assessed in brief (15-60 minute) sessions supervised by a researcher, leaving the open question of how well body position classification will scale to testing across an entire day of natural home life.
+In contrast to the three machine learning studies that used 3-4 sensors [@AiraksinenRasanen2020; @FranchakScott2021; @AiraksinenGallen2022], @GreenspanCunha2021 used orientation from a single hip-worn sensor to measure body position with a high degree of accuracy (*kappa* = .84).
+Although all four studies yielded promising classification accuracy, accuracy was assessed in brief (15-60 minute) sessions supervised by a researcher, leaving the open question of how well body position classification will scale to testing across an entire day of natural home life.
 
@@
 Families were compensated \$30 for every home recording session they completed.
-The BLINDED Institutional Review Board reviewed and approved all procedures associated with the study.
+The University of California, Riverside Institutional Review Board reviewed and approved all procedures associated with the study.
 All caregivers gave their informed consent before the start of the study.
 
-Most participants were tested in a single session (*n* = `r nrow(num_sessions %>% filter(n == 1))`), but `r nrow(num_sessions %>% filter(n > 1))` participants contributed between 2-4 sessions as part of an ongoing longitudinal study (three from the younger group and four from the older group).
+Most participants were tested in a single session (*n* = `r nrow(num_sessions %>% filter(n == 1))`), but `r nrow(num_sessions %>% filter(n > 1))` participants contributed between 2-4 sessions as part of an ongoing longitudinal study (3 from the younger group and 4 from the older group).
 Although including more data from some participants could over-represent their characteristics in the model, we reasoned that this drawback was outweighed by having more available data to use for training.
@@
 Across sessions, younger infants' age ranged from `r apa_num(age["Younger","min"], digits = 1)` to `r apa_num(age["Younger","max"], digits = 1)` months (*M* = `r apa_num(age["Younger","mean"], digits = 1)`) and older infants' age ranged from `r apa_num(age["Older","min"], digits = 1)` to `r apa_num(age["Older","max"], digits = 1)` months (*M* = `r apa_num(age["Older","mean"], digits = 1)`).
-The total number of recording sessions (`r length(unique(ppts$id_uni))`) exceeded the number of sessions employed in comparable past work: 10 in @NamPark2013, 15 in @FranchakScott2021, 22 in @AiraksinenRasanen2020, and 33 in @YaoPlotz2019.
+The total number of recording sessions (`r length(unique(ppts$id_uni))`) exceeded the number of sessions employed in comparable past work: 10 in @NamPark2013, 15 in @FranchakScott2021, 22 in @AiraksinenRasanen2020, 23 in @GreenspanCunha2021, and 33 in @YaoPlotz2019.
 
@@
 Because the camera only had the battery life to record for ~3 hours (automatically split into two 90-minute video files), this divided the day into different periods for analysis.
-As seen in the bottom of Figure \@ref(fig:exemplar-timeline)), the *video period* comprised the first three hours of recording starting from the researcher's arrival when they turned on the camera.
+As seen in the bottom of Figure \@ref(fig:exemplar-timeline), the *video period* comprised the first three hours of recording starting from the researcher's arrival when they turned on the camera.
 The first 90-minute video file, termed the *proximal comparison*, contained the activities during the guided call followed by a period during which infants and caregivers resumed their normal activities.
@@
 
-The proximal and distal comparison videos for each participant were annotated by trained human coders to identify infant *body position* into one of 5 mutually-exclusive categories following the definitions in prior work [@FranchakScott2021]: supine, prone, sitting, upright, or held by caregiver.
+The proximal and distal comparison videos for each participant were annotated by trained human coders to classify infant *body position* into one of 5 mutually-exclusive categories following the definitions in prior work [@FranchakScott2021]: supine, prone, sitting, upright, or held by caregiver.
 All coding was done using Datavyu software (datavyu.org).
@@
                                          "left_hip","left_ankle", "right_hip", "right_ankle"),
-                              labels = c("All", "Left Hip/Ankle", "Right Thigh/Ankle", "Both Thighs", "Both Ankles", 
+                              labels = c("All", "Left Thigh/Ankle", "Right Thigh/Ankle", "Both Thighs", "Both Ankles", 
                                          "Left Thigh", "Left Ankle", "Right Thigh", "Right Ankle")),
@@
 The highest accuracy was observed when using all four sensors (group model *M* = `r metrics_sensor["All", "mean_group"]`; individual model *M* = `r metrics_sensor["All", "mean_split"]`), and the lowest when using only a single sensor (left ankle group model *M* = `r metrics_sensor["Left Ankle", "mean_group"]`; right ankle individual model *M* = `r metrics_sensor["Right Ankle", "mean_split"]`).
-Accuracy for some pairs approached 4-sensor accuracy: Left ankle and thigh had group model accuracy of *M* = `r metrics_sensor["Left Hip/Ankle", "mean_group"]` and individual model accuracy of *M* = `r metrics_sensor["Left Hip/Ankle", "mean_split"]`, only 2.2% and 1.5% worse than using all four sensors.
+Accuracy for some pairs approached 4-sensor accuracy: Left ankle and thigh had group model accuracy of *M* = `r metrics_sensor["Left Thigh/Ankle", "mean_group"]` and individual model accuracy of *M* = `r metrics_sensor["Left Thigh/Ankle", "mean_split"]`, only 2.2% and 1.5% worse than using all four sensors.
 Other pairings were less accurate; notably, using both ankles resulted in group model accuracy of *M* = `r metrics_sensor["Both Ankles", "mean_group"]` and individual model accuracy of *M* = `r metrics_sensor["Both Ankles", "mean_split"]`, 12.7% and 7.2% worse than using all four sensors.
@@
 
-The first set of results showed that group and individual models trained from data during the proximal comparison period were within the proximal period.
+The first set of results showed that group and individual models trained from data during the proximal comparison period were accurate during the proximal period.
 In the next analysis, we examine long-term performance by testing how accurately models trained from the proximal period could predict body position during the distal comparison period.
@@
 Predicted time was calculated the same way for group and individual model predictions, omitting the off-camera and nap periods to make a direct comparison.
-Overall---across participants and across body position classes---agreement was strong: The correlation between group model predictions and human-coded time was *r* = `r group_agree_cor`, and the correlation between individual model predictions and human-coded time was *r* = `r split_agree_cor`.
+Agreement was strong across participants and across body position classes: The correlation between group model predictions and human-coded time was *r* = `r group_agree_cor`, and the correlation between individual model predictions and human-coded time was *r* = `r split_agree_cor`.
 As in the proximal comparison, agreement varied somewhat between body positions; in particular, agreement for held was poor.
@@
 The challenge increased in the distal comparison period.
-Since sessions were scheduled at convenient times for the infant and caregiver, the transition from the proximal to distal period increased the odds that infants might need a nap, a snack/meal, or engage in a less typical activity (e.g., watching TV).
+Since sessions were scheduled at convenient times for the infant and caregiver, the transition from the proximal to distal period increased the odds that infants might need a nap, eat a snack/meal, or engage in a less typical activity (e.g., watching TV).
 Indeed, the distal comparison in Figure \@ref(fig:exemplar-timeline) contains a long period (almost 50 minutes) where the infant sat in a high chair eating lunch.
@@
 Long-form recordings of motor behavior bring about new research possibilities.
-Although past work using video [@Karasik2011; @ChenSchneider2022; @HerzbergFletcher2021] produced real-time data, such data were limited to a relatively small part of the infants' day.
+Although past work using video [@Karasik2011; @ChenSchneider2022; @HerzbergFletcher2021] produced real-time data, such data were limited to a relatively small part of the day.
 And although survey methods [@Survey] could capture moments scattered throughout the day, they do not produce real-time data.
@@
 We acknowledge several limitations in the current approach.
-First, a stronger test of full-day accuracy would be to compare accuracy at the start of the recording period to accuracy at the end, rather than using the distal comparison video that followed the first 90 minutes of the study.
+First, despite the large amount of data collected per infant, the sample of infants was small (22 unique participants).
+Since the study combined data from participants who completed single sessions and participants who completed multiple sessions, the seven infants with multiple sessions are over-represented in the dataset.
+A larger, more representative sample would be needed to determine whether the method would generalize  to the broader population.
+
+Second, a stronger test of full-day accuracy would be to compare accuracy at the start of the recording period to accuracy at the end, rather than using the distal comparison video that followed the first 90 minutes of the study.
 As we described in the procedure section, synchronizing video cameras and inertial sensors required capturing a synchronization point on video.
@@
 
-Second, we relied on caregiver logs of infants' naps and times when sensors were removed to separate usable data (times when infants wore the sensors while awake) from unusable data.
+Third, we relied on caregiver logs of infants' naps and times when sensors were removed to separate usable data (times when infants wore the sensors while awake) from unusable data.
 Future work should aim to further assess caregivers' perceptions about the usability and comfort of the garment.
@@
 
-Third, although the goal was to collect fully naturalistic behavior, our recording protocol led to some changes that might have affected (or missed) some behaviors.
+Fourth, although the goal was to collect fully naturalistic behavior, our recording protocol led to some changes that might have affected (or missed) some behaviors.
 We opted to restrict recording to times when participants were in the home.

### 2023-08-14 manuscript.tex
Working on revision

@@
 \abstract{%
-Long form audio recordings have had a transformational effect on the study of infant language acquisition by using mobile, unobtrusive devices to gather full-day, real-time data that can be automatically scored.
+Long-form audio recordings have had a transformational effect on the study of infant language acquisition by using mobile, unobtrusive devices to gather full-day, real-time data that can be automatically scored.
 How can we produce similar data in service of measuring infants' everyday motor behaviors, such as body position?
-The aim of the current study was to validate long form recordings of infant position (supine, prone, sitting, upright, held by caregiver) based on machine learning classification of data from inertial sensors worn on infants' ankles and thighs.
+The aim of the current study was to validate long-form recordings of infant position (supine, prone, sitting, upright, held by caregiver) based on machine learning classification of data from inertial sensors worn on infants' ankles and thighs.
 Using over 100 hours of video recordings synchronized with inertial sensor data from infants in their homes, we demonstrate that body position classifications are sufficiently accurate to measure infant behavior.
@@
 Next, we show that the inertial sensing method has convergent validity by replicating age differences in body position found using other methods with full-day data captured from inertial sensors.
-We end the paper with a discussion of the novel opportunities that long form motor recordings afford for understanding infant learning and development.
+We end the paper with a discussion of the novel opportunities that long-form motor recordings afford for understanding infant learning and development.
 }
@@
 Finally, we discuss the advantages of using inertial sensing to classify motor behavior.
-Despite promising past results in brief, supervised sessions \autocite{FranchakScott2021,AiraksinenRasanen2020,AiraksinenGallen2022}, the current investigation takes a needed step forward by testing accuracy over long, unsupervised recordings in what we term a \emph{distal comparison}.
+Despite promising past results in brief, supervised sessions \autocite{FranchakScott2021,AiraksinenRasanen2020,GreenspanCunha2021,AiraksinenGallen2022}, the current investigation takes a needed step forward by testing accuracy over long, unsupervised recordings in what we term a \emph{distal comparison}.
 
@@
 For other metrics, such as the number of ``conversational turns'' between the child and communicative partners, agreement was poor (\emph{r} = .364).
-Thus, for some use cases (and for some metrics), long form audio recordings provide a mobile, unobtrusive way to automatically score real-time data over a full day.
+Thus, for some use cases (and for some metrics), long-form audio recordings provide a mobile, unobtrusive way to automatically score real-time data over a full day.
 
@@
 
-Measuring infant movement with inertial movement units (IMUs) is a promising avenue for long form recordings of motor behavior in the home \autocite{CliffReilly2009,Barbaro2019,BruijnsTruelove2020,LoboHall2019}.
+Measuring infant movement with inertial movement units (IMUs) is a promising avenue for long-form recordings of motor behavior in the home \autocite{CliffReilly2009,Barbaro2019,BruijnsTruelove2020,LoboHall2019}.
 Lightweight sensors (10-30 g) can be embedded in garments to make recordings fully \emph{mobile}, and they are \emph{unobtrusive} because they do not require a researcher to follow with a camcorder.
@@
 Modern approaches to human activity recognition have used machine learning to classify activity categories based on features derived from IMU data in adults \autocite{Preece2009,Arif2015}, children \autocite{NamPark2013,RenDing2016,StewartNarayanan2018}, and infants \autocite{YaoPlotz2019,FranchakScott2021,AiraksinenRasanen2020}.
-Three prior investigations have used different machine learning techniques to categorize infant body position from IMUs towards the goal of collecting full-day data.
+Three prior investigations have used different machine learning techniques to categorize infant body position from multiple IMUs towards the goal of collecting full-day data.
 \textcite{AiraksinenRasanen2020} tested 4- to 8-month-olds in a laboratory visit with a 4-sensor array (one on each thigh and one below each shoulder), and found 95\% accuracy in distinguishing between body position categories that crawling infants could perform on the floor (excluding times that infants were held by caregivers).
 Using a wider age range of 6-18 months, \textcite{FranchakScott2021} found 98\% accuracy (\emph{kappa} = 95\%) in a laboratory validation study with a 3-sensor array (ankle, knee, thigh on a single leg) in categorizing body position that included infants who could both crawl and walk and also included a category for caregiver holding.
-Most recently, \textcite{AiraksinenGallen2022} conducted a validation study of body position classification in either a home or clinic testing 4- to 19-month-olds with their 4-sensor system, refining their previous method to detect moments that infants were carried by caregivers.
+Most recently, \textcite{AiraksinenGallen2022} conducted a validation study of body position classification in either a home or clinic testing 4- to 19-month-olds with a 4-sensor system, refining their previous method to detect moments that infants were carried by caregivers.
 Classification accuracy did not vary between lab and home settings, and was generally high (95\%, \emph{kappa} = .93).
-Although all three studies yielded promising classification accuracy, accuracy was assessed in brief (15-60 minute) sessions supervised by a researcher, leaving the open question of how well body position classification will scale to testing across an entire day of natural home life.
+In contrast to the three machine learning studies that used 3-4 sensors \autocite{AiraksinenRasanen2020,FranchakScott2021,AiraksinenGallen2022}, \textcite{GreenspanCunha2021} used orientation from a single hip-worn sensor to measure body position with a high degree of accuracy (\emph{kappa} = .84).
+Although all four studies yielded promising classification accuracy, accuracy was assessed in brief (15-60 minute) sessions supervised by a researcher, leaving the open question of how well body position classification will scale to testing across an entire day of natural home life.
 
@@
 Families were compensated \$30 for every home recording session they completed.
-The BLINDED Institutional Review Board reviewed and approved all procedures associated with the study.
+The University of California, Riverside Institutional Review Board reviewed and approved all procedures associated with the study.
 All caregivers gave their informed consent before the start of the study.
 
-Most participants were tested in a single session (\emph{n} = 15), but 7 participants contributed between 2-4 sessions as part of an ongoing longitudinal study (three from the younger group and four from the older group).
+Most participants were tested in a single session (\emph{n} = 15), but 7 participants contributed between 2-4 sessions as part of an ongoing longitudinal study (3 from the younger group and 4 from the older group).
 Although including more data from some participants could over-represent their characteristics in the model, we reasoned that this drawback was outweighed by having more available data to use for training.
@@
 Across sessions, younger infants' age ranged from 3.8 to 7.2 months (\emph{M} = 5.2) and older infants' age ranged from 10.7 to 14.2 months (\emph{M} = 11.7).
-The total number of recording sessions (34) exceeded the number of sessions employed in comparable past work: 10 in \textcite{NamPark2013}, 15 in \textcite{FranchakScott2021}, 22 in \textcite{AiraksinenRasanen2020}, and 33 in \textcite{YaoPlotz2019}.
+The total number of recording sessions (34) exceeded the number of sessions employed in comparable past work: 10 in \textcite{NamPark2013}, 15 in \textcite{FranchakScott2021}, 22 in \textcite{AiraksinenRasanen2020}, 23 in \textcite{GreenspanCunha2021}, and 33 in \textcite{YaoPlotz2019}.
 
@@
 Because the camera only had the battery life to record for \textasciitilde3 hours (automatically split into two 90-minute video files), this divided the day into different periods for analysis.
-As seen in the bottom of Figure \ref{fig:exemplar-timeline}), the \emph{video period} comprised the first three hours of recording starting from the researcher's arrival when they turned on the camera.
+As seen in the bottom of Figure \ref{fig:exemplar-timeline}, the \emph{video period} comprised the first three hours of recording starting from the researcher's arrival when they turned on the camera.
 The first 90-minute video file, termed the \emph{proximal comparison}, contained the activities during the guided call followed by a period during which infants and caregivers resumed their normal activities.
@@
 
-The proximal and distal comparison videos for each participant were annotated by trained human coders to identify infant \emph{body position} into one of 5 mutually-exclusive categories following the definitions in prior work \autocite{FranchakScott2021}: supine, prone, sitting, upright, or held by caregiver.
+The proximal and distal comparison videos for each participant were annotated by trained human coders to classify infant \emph{body position} into one of 5 mutually-exclusive categories following the definitions in prior work \autocite{FranchakScott2021}: supine, prone, sitting, upright, or held by caregiver.
 All coding was done using Datavyu software (datavyu.org).
@@
 All & 0.894 & 0.846 & 0.131 & 0.933 & 0.916 & 0.072\\ \midrule
-Left Hip/Ankle & 0.891 & 0.824 & 0.177 & 0.915 & 0.901 & 0.071\\
+Left Thigh/Ankle & 0.891 & 0.824 & 0.177 & 0.915 & 0.901 & 0.071\\
 Right Thigh/Ankle & 0.842 & 0.800 & 0.167 & 0.897 & 0.898 & 0.086\\
@@
 Predicted time was calculated the same way for group and individual model predictions, omitting the off-camera and nap periods to make a direct comparison.
-Overall---across participants and across body position classes---agreement was strong: The correlation between group model predictions and human-coded time was \emph{r} = 0.80, and the correlation between individual model predictions and human-coded time was \emph{r} = 0.91.
+Agreement was strong across participants and across body position classes: The correlation between group model predictions and human-coded time was \emph{r} = 0.80, and the correlation between individual model predictions and human-coded time was \emph{r} = 0.91.
 As in the proximal comparison, agreement varied somewhat between body positions; in particular, agreement for held was poor.
@@
 The challenge increased in the distal comparison period.
-Since sessions were scheduled at convenient times for the infant and caregiver, the transition from the proximal to distal period increased the odds that infants might need a nap, a snack/meal, or engage in a less typical activity (e.g., watching TV).
+Since sessions were scheduled at convenient times for the infant and caregiver, the transition from the proximal to distal period increased the odds that infants might need a nap, eat a snack/meal, or engage in a less typical activity (e.g., watching TV).
 Indeed, the distal comparison in Figure \ref{fig:exemplar-timeline} contains a long period (almost 50 minutes) where the infant sat in a high chair eating lunch.
@@
 Long-form recordings of motor behavior bring about new research possibilities.
-Although past work using video \autocite{Karasik2011,ChenSchneider2022,HerzbergFletcher2021} produced real-time data, such data were limited to a relatively small part of the infants' day.
+Although past work using video \autocite{Karasik2011,ChenSchneider2022,HerzbergFletcher2021} produced real-time data, such data were limited to a relatively small part of the day.
 And although survey methods \autocite{Survey} could capture moments scattered throughout the day, they do not produce real-time data.
@@
 We acknowledge several limitations in the current approach.
-First, a stronger test of full-day accuracy would be to compare accuracy at the start of the recording period to accuracy at the end, rather than using the distal comparison video that followed the first 90 minutes of the study.
+First, despite the large amount of data collected per infant, the sample of infants was small (22 unique participants).
+Since the study combined data from participants who completed single sessions and participants who completed multiple sessions, the seven infants with multiple sessions are over-represented in the dataset.
+A larger, more representative sample would be needed to determine whether the method would generalize to the broader population.
+
+Second, a stronger test of full-day accuracy would be to compare accuracy at the start of the recording period to accuracy at the end, rather than using the distal comparison video that followed the first 90 minutes of the study.
 As we described in the procedure section, synchronizing video cameras and inertial sensors required capturing a synchronization point on video.

### 2023-08-17 manuscript.Rmd
Hailey and Max edits

@@
 
-But automatic classification can only replace human annotation if it is sufficiently accurate.
+However, automatic classification can only replace human annotation if it is sufficiently accurate.
 An independent assessment of the LENA&reg; algorithms found mixed results about classification accuracy.
@@
 In contrast, survey methods such as daily diaries/inventories or ecological momentary assessment (EMA) are mobile, unobtrusive, can be applied across an entire day, and do not need laborious annotation.
-Diary studies provide caregivers with logs or structured interviews record activities [@MajnemerBarr2005; @KarasikKuchirko2022].
+Diary studies provide caregivers with logs or structured interviews to record activities [@MajnemerBarr2005; @KarasikKuchirko2022].
 Ecological momentary assessment uses smartphone notifications to prompt caregivers to make repeated estimates of behaviors throughout the day [@Survey; @KadookaCaufield2021].
@@
 Overall accuracy (A) and Cohen's Kappa (B) are plotted separately for group models and individual models.
-Each blue circle represents that accuracy for each recording session.
+Each blue circle represents the accuracy for each recording session.
 Horizontal black bars indicate the mean across sessions.", fig.align="center", fig.width=6, fig.height=3, cache=TRUE, cache.extra=tools::md5sum("data/group_split_metrics.csv")}
@@
 Each circle represents one full-day recording session's proportion of time in each body position (y-axis) against age in months (x-axis).
-Black lines indicate best fit regression lines indicating decreases in supine time and increases in sitting and upright time with age.", fig.align="center", fig.width=7.5, fig.height=7, cache=TRUE, cache.extra=tools::md5sum("data/imu-compiled.csv")}
+Black lines indicate best fit regression lines, which show decreases in supine time and increases in sitting and upright time with age.", fig.align="center", fig.width=7.5, fig.height=7, cache=TRUE, cache.extra=tools::md5sum("data/imu-compiled.csv")}
 fullday_group <- read_csv("data/imu-compiled.csv") %>% 
@@
 Models trained during the proximal comparison period performed well on testing data collected from the same recording period.
-Most important, accuracy was consistently strong later in the recording when behavior was completely unsupervised: Human-coded and model-predicted durations of body positions during the distal comparison period were highly correlated, even when narrowing to the scale of 10 minutes.
+Most important, accuracy was consistently strong later in the distal recording period when behavior was completely unsupervised: Human-coded and model-predicted durations of body positions during the distal comparison period were highly correlated, even when narrowing to the scale of 10 minutes.
 Examining full recordings showed that the new method allows us to capture more data (median = `r apa_num(median(coverage$hours_recorded))` hours of awake behavior) than is typical with video methods (1-2 hours) while simultaneously reducing the human labor needed to annotate it.
@@
 How skewed are infants' experiences with different motor behaviors, and how are they clustered in time?
-Linking the distributional structure of a particular motor experience in a real day, such as time spent sitting, might be predicted by sitting skill and/or might predict infants' future sitting ability.
+The distributional structure of a particular motor experience in a real day, such as time spent sitting, might be predicted by concurrent sitting skill and/or might predict future sitting skill.
 Moreover, real-time recording of motor behavior provides a way to measure how infant motor behavior links to other types of experiences in the moment in daily life.

### 2023-08-17 manuscript.tex
Hailey and Max edits

@@
 
-But automatic classification can only replace human annotation if it is sufficiently accurate.
+However, automatic classification can only replace human annotation if it is sufficiently accurate.
 An independent assessment of the LENA® algorithms found mixed results about classification accuracy.
@@
 In contrast, survey methods such as daily diaries/inventories or ecological momentary assessment (EMA) are mobile, unobtrusive, can be applied across an entire day, and do not need laborious annotation.
-Diary studies provide caregivers with logs or structured interviews record activities \autocite{MajnemerBarr2005,KarasikKuchirko2022}.
+Diary studies provide caregivers with logs or structured interviews to record activities \autocite{MajnemerBarr2005,KarasikKuchirko2022}.
 Ecological momentary assessment uses smartphone notifications to prompt caregivers to make repeated estimates of behaviors throughout the day \autocite{Survey,KadookaCaufield2021}.
@@
 Overall accuracy (A) and Cohen's Kappa (B) are plotted separately for group models and individual models.
-Each blue circle represents that accuracy for each recording session.
+Each blue circle represents the accuracy for each recording session.
 Horizontal black bars indicate the mean across sessions.}\label{fig:metrics}
@@
 Each circle represents one full-day recording session's proportion of time in each body position (y-axis) against age in months (x-axis).
-Black lines indicate best fit regression lines indicating decreases in supine time and increases in sitting and upright time with age.}\label{fig:age}
+Black lines indicate best fit regression lines, which show decreases in supine time and increases in sitting and upright time with age.}\label{fig:age}
 \end{figure}
@@
 Models trained during the proximal comparison period performed well on testing data collected from the same recording period.
-Most important, accuracy was consistently strong later in the recording when behavior was completely unsupervised: Human-coded and model-predicted durations of body positions during the distal comparison period were highly correlated, even when narrowing to the scale of 10 minutes.
+Most important, accuracy was consistently strong later in the distal recording period when behavior was completely unsupervised: Human-coded and model-predicted durations of body positions during the distal comparison period were highly correlated, even when narrowing to the scale of 10 minutes.
 Examining full recordings showed that the new method allows us to capture more data (median = 6.14 hours of awake behavior) than is typical with video methods (1-2 hours) while simultaneously reducing the human labor needed to annotate it.

### 2023-09-08 rev1-manuscript/manuscript.Rmd
revision

--- 2023-08-14 manuscript.Rmd
@@
 #9908  9909  9910  9911 10201 10203 10204 10301 10501 10601 10701 10702 10703 10704 10802 10901 11001 11002 11004 11102 11103 11104 11202 11501 11601 11901 12001 12002 12101 12302 12303 12501 12502 13001
-# repeat sessions - younger = 3
-# repeat sessions - older = 4
 ```
@@
 Most often, an experimenter with a handheld camera follows infants from room to room to ensure that their movements are visible throughout the recording session [@Karasik2011; @ChenSchneider2022; @HerzbergFletcher2021].
-The primary advantage of video recording is that it captures real-time behavior.
-Standard 30 Hz video recording is adequate to capture changes in infant body position that occur on the timescale of seconds.
+The primary advantage of video recording is that it captures real-time behavior.Standard 30 Hz video recording is adequate to capture changes in infant body position that occur on the timescale of seconds.
 However, requiring an experimenter to operate a camera is obtrusive, whereas relying on a stationary camera means that infants will be absent from view as they move from place to place.
@@
 Data processing algorithms are needed to classify the raw sensor data (i.e., linear and angular acceleration time series) into meaningful categories.
-Categorizing body position---supine on the back, prone on the belly, sitting, upright, or held off the ground by a caregiver---is complex because movement can vary greatly *within* a body position.
+Categorizing body position---supine on the back, prone on the belly, sitting, upright, or held off the ground by a caregiver---is a complex because movement can vary greatly *within* a body position.
 An upright infant can be standing still or can be walking briskly across the room.
@@
 Sitting on a caregiver's lap without the need to maintain balance means that the legs can dangle and the torso can lean in different directions.
-Sitting in a high chair or car seat likely restricts the range of torso orientations compared with sitting independently or on a caregiver's lap.
+Sitting in a high chair or car seat reduces the magnitude of postural sway within sitting, and creates even more possibilities for how the arms and legs may move relative to the torso.
 Finally, caregivers frequently pick up and transport infants, creating motion signals that need to be differentiated from independent activity [@KwonZavos2019; @PatelShi2019].
@@
 Moreover, classifiers must be able to detect new variants of behaviors that might arise over the course of a real day; it is impossible for researchers to gather training data for every possible variation that might occur.
-Thus, our central question is whether models trained on video-recorded observations at the beginning of the day generalize to predict behavior at a later time.
+Thus, our central question is whether models trained on video-recorded observations at the beginning of the day generalize to predict behavior at a later time?
 Assessing the validity of temporally *distal* periods---when infants and caregivers are unsupervised and free to follow their everyday routines---is a crucial step to establish whether automatic classification can be used to measure body position across a day.
@@
 age <- ppts %>% group_by(group) %>% get_summary_stats(age) %>% column_to_rownames("group")
-
 ```
@@
 Families were recruited through social media advertisements and from community events in Southern California.
-Most infants were reported to be either Hispanic and White (*n* = 8) or Non-Hispanic and White (*n* = 7).
+Most infants were reported to be either Hispanic and White or Non-Hispanic and White.
 Families were compensated \$30 for every home recording session they completed.
@@
 
-Most participants were tested in a single session (*n* = `r nrow(num_sessions %>% filter(n == 1))`), but `r nrow(num_sessions %>% filter(n > 1))` participants contributed between 2-4 sessions as part of an ongoing longitudinal study (three from the younger group and four from the older group).
+Most participants were tested in a single session (*n* = `r nrow(num_sessions %>% filter(n == 1))`), but `r nrow(num_sessions %>% filter(n > 1))` participants contributed between 2-4 sessions as part of an ongoing longitudinal study.
 Although including more data from some participants could over-represent their characteristics in the model, we reasoned that this drawback was outweighed by having more available data to use for training.
@@
 Four inertial movement units (IMUs) were used to record infant movement across the day (MC10 Biostamp).
-A custom garment was designed to hold the IMUs (Figure \@ref(fig:garment)).
-Internal pockets were sewn into snug-fitting infant leggings so that IMUs would stay close to the body (reducing vibration).
+A custom garment was designed to hold the IMUs: Internal pockets were sewn into snug-fitting infant leggings so that IMUs would stay close to the body (reducing vibration).
 A pocket over the thigh and a pocket just above the ankle were sewn on the lateral surface of the right and left legs of the garment.
@@
 
-Each sensor had sufficient battery and onboard storage to record accelerometer and gyroscope data for approximately 12 hours.
+Each sensor had sufficient battery and onboard torage to record accelerometer and gyroscope data for approximately 12 hours.
 We chose a sampling rate of 62.5 Hz (one of the available presets) based on prior work that used rates of 50-64 Hz [@FranchakScott2021; @AiraksinenRasanen2020; @YaoPlotz2019].
@@
 
-Supine was coded when the infant was lying on their back, on their side, or was reclined up to a 45 degree angle.
-Prone was coded when the infant was lying on their stomach, was stationary supported by the hands/knees or the hands/feet, or was crawling.
+Supine was coded when the infant was lying on their back, on their side, or was reclined up to a 45  Prone was coded when the infant was lying on their stomach, was stationary supported by the hands/knees or the hands/feet, or was crawling.
 We scored sitting to include any form of the following seated positions: 1) infants sat with their buttocks on a surface, such as on the floor or a caregiver’s lap, 2) infant was in a kneeling-sit position, in which their knees were on the ground with their legs tucked underneath the buttocks, and 3) infant was in a seating device, such as a high chair, that kept the torso oriented perpendicular to the ground (a reclined position, such as in a young infant's car seat, would be counted as supine).
@@
 Our goal in creating a category for "held by caregiver" was to separate times when infants were in control of their body position from times when they were suspended in the air (rather than resting on furniture or a surface).
-Held was coded when infants were carried off of the ground.
+Held was coded when infants were carried off of the groun.
 However, when the caregiver was sitting with the infant in their lap the infant’s body position was coded as if the caregiver was a surface (e.g., if the infant was sitting on the caregiver’s lap this was coded as sitting).
@@
 For each 4-s window, we aggregated the 250 samples to create single scores for a variety of motion features---summary statistics that could be fed into the machine learning model.
-The minimum, maximum, 25th percentile, 75th percentile, mean, median, skew, kurtosis, standard deviation, and sum were computed for each signal (e.g., right thigh linear acceleration along the X-axis, left ankle pitch angular acceleration).
+The minimum, maximum, 25th percentile, 75th percentile, mean, median, skew, kurtosis, standard deviation, and sum were computed for each signal (e.g., right thigh X linear acceleration, left ankle pitch angular acceleration).
 The 10 summary statistics and 24 sensor signals generated 240 columns of motion features that described movement within each window.
@@
 Although overall accuracy from our semi-supervised, in-home data collection did not match the near-perfect accuracy  (.95-.98) found in prior in-lab studies [@FranchakScott2021; @AiraksinenRasanen2020], both models approached the level of agreement found between two human coders (*M* = .906).
-Most likely, lower accuracy in the current study results from the more variable and complex behavior observed in a semi-supervised setting rather than from a difference in the quality of the classification model; although the first part of the proximal period was guided by the experimenter during a brief phone call, the remainder of the proximal period included natural behavior.
+Most likely, lower accuracy in the current study results from the more variable and complex behavior observed in a a semi-supervised setting rather than from a difference in the quality of the classification model; although the first part of the proximal period was guided by the experimenter during a brief phone call, the remainder of the proximal period included natural behavior.
 Visual inspection of Figure \@ref(fig:metrics)A shows that accuracy values were heavily skewed, with many approaching perfect accuracy but a few sessions with very poor accuracy.
@@
 Among participants who wore the sensors from morning to bedtime, all infants wore the sensor garment during 100% of the intended recording period, excluding the times caregivers were asked to remove the garment (naps, diaper changes, trips out of the house), based on caregiver logs of wear time.
-No caregiver reported removing the garment for any other reason (such as infant discomfort), meaning that the majority of the time during the day either resulted in usable body position data or was excluded due to caregiver-reported naps (gray periods in Figure \@ref(fig:timelines)).
+No caregiver reporting removing the garment for any other reason (such as infant discomfort), meaning that the majority of the time during the day either resulted in usable body position data or was excluded due to caregiver-reported naps (gray periods in Figure \@ref(fig:timelines)).
 The total length of the recording period ranged from `r apa_num(min(coverage$total))`-`r apa_num(max(coverage$total))` hours with a median of `r apa_num(median(coverage$total))` hours.
@@
   ds %>% drop_na(prop) %>% 
-  ggplot() + facet_wrap("position", as.table = T, ncol = 2, scales = "free_x") + 
+  ggplot() + facet_wrap("position", as.table = T, ncol = 2) + 
   geom_smooth(aes(x = age, y = prop), color = "#191919", method = "lm", se = T) + 
@@
   scale_x_continuous(name = "Age (months)", breaks = seq(3, 15, 3), limits = c(3,15)) + 
-    scale_y_continuous(name = "Proportion of waking day", breaks = seq(0, 1, .25), limits = c(0,1)) + 
+  ylim(0,1) + ylab("Proportion of waking day") + 
   theme(panel.spacing.x = unit(.5, "lines")) +
@@
 Age in months and upright time were significantly, positively correlated using group (*r* = `r age_cor_group["age","upright_time"]`, *p* `r apa_p(age_cor_group_pval["age","upright_time"], add_equals = T)`) and individual models (*r* = `r age_cor_ind["age","upright_time"]`, *p* `r apa_p(age_cor_ind_pval["age","upright_time"], add_equals = T)`).
-Similarly, sitting time increased with age based on predictions from both group (*r* = `r age_cor_group["age","sit_time"]`, *p* `r apa_p(age_cor_group_pval["age","sit_time"], add_equals = T)`) and individual models (*r* = `r age_cor_ind["age","sit_time"]`, *p* `r apa_p(age_cor_ind_pval["age","sit_time"], add_equals = T)`).
-In contrast, supine time decreased with age using predictions from both group (*r* = `r age_cor_group["age","supine_time"]`, *p* `r apa_p(age_cor_group_pval["age","supine_time"], add_equals = T)`) and individual models (*r* = `r age_cor_ind["age","supine_time"]`, *p* `r apa_p(age_cor_ind_pval["age","supine_time"], add_equals = T)`).
+Similarly, age predicted increases in sitting time using group (*r* = `r age_cor_group["age","sit_time"]`, *p* `r apa_p(age_cor_group_pval["age","sit_time"], add_equals = T)`) and individual models (*r* = `r age_cor_ind["age","sit_time"]`, *p* `r apa_p(age_cor_ind_pval["age","sit_time"], add_equals = T)`).
+In contrast, age predicted decreases in supine time using group (*r* = `r age_cor_group["age","supine_time"]`, *p* `r apa_p(age_cor_group_pval["age","supine_time"], add_equals = T)`) and individual models (*r* = `r age_cor_ind["age","supine_time"]`, *p* `r apa_p(age_cor_ind_pval["age","supine_time"], add_equals = T)`).
 No clear age trends were found for prone time.
@@
 Linking the distributional structure of a particular motor experience in a real day, such as time spent sitting, might be predicted by sitting skill and/or might predict infants' future sitting ability.
-Moreover, real-time recording of motor behavior provides a way to measure how infant motor behavior links to other types of experiences in the moment in daily life.
-@MalachowskiSalo2023 used a combination of LENA&reg; and EMA surveys to find that infants heard less adult speech when restrained in seating devices.
+Moreover, real-time recording of motor behavior provides a way to measure how infant motor behavior links to other types of experiences in the moment in daily life.@MalachowskiSalo2023 used a combination of LENA&reg; and EMA surveys to find that infants heard less adult speech when restrained in seating devices.
 Combining LENA&reg; and long-form motor recordings can take this a step further by measuring more precisely how speech and body position co-vary within a day.

### 2023-09-08 rev1-manuscript/manuscript.tex
revision

--- 2023-08-14 manuscript.tex
@@
 Most often, an experimenter with a handheld camera follows infants from room to room to ensure that their movements are visible throughout the recording session \autocite{Karasik2011,ChenSchneider2022,HerzbergFletcher2021}.
-The primary advantage of video recording is that it captures real-time behavior.
-Standard 30 Hz video recording is adequate to capture changes in infant body position that occur on the timescale of seconds.
+The primary advantage of video recording is that it captures real-time behavior.Standard 30 Hz video recording is adequate to capture changes in infant body position that occur on the timescale of seconds.
 However, requiring an experimenter to operate a camera is obtrusive, whereas relying on a stationary camera means that infants will be absent from view as they move from place to place.
@@
 Data processing algorithms are needed to classify the raw sensor data (i.e., linear and angular acceleration time series) into meaningful categories.
-Categorizing body position---supine on the back, prone on the belly, sitting, upright, or held off the ground by a caregiver---is complex because movement can vary greatly \emph{within} a body position.
+Categorizing body position---supine on the back, prone on the belly, sitting, upright, or held off the ground by a caregiver---is a complex because movement can vary greatly \emph{within} a body position.
 An upright infant can be standing still or can be walking briskly across the room.
@@
 Sitting on a caregiver's lap without the need to maintain balance means that the legs can dangle and the torso can lean in different directions.
-Sitting in a high chair or car seat likely restricts the range of torso orientations compared with sitting independently or on a caregiver's lap.
+Sitting in a high chair or car seat reduces the magnitude of postural sway within sitting, and creates even more possibilities for how the arms and legs may move relative to the torso.
 Finally, caregivers frequently pick up and transport infants, creating motion signals that need to be differentiated from independent activity \autocite{KwonZavos2019,PatelShi2019}.
@@
 Moreover, classifiers must be able to detect new variants of behaviors that might arise over the course of a real day; it is impossible for researchers to gather training data for every possible variation that might occur.
-Thus, our central question is whether models trained on video-recorded observations at the beginning of the day generalize to predict behavior at a later time.
+Thus, our central question is whether models trained on video-recorded observations at the beginning of the day generalize to predict behavior at a later time?
 Assessing the validity of temporally \emph{distal} periods---when infants and caregivers are unsupervised and free to follow their everyday routines---is a crucial step to establish whether automatic classification can be used to measure body position across a day.
@@
 Families were recruited through social media advertisements and from community events in Southern California.
-Most infants were reported to be either Hispanic and White (\emph{n} = 8) or Non-Hispanic and White (\emph{n} = 7).
+Most infants were reported to be either Hispanic and White or Non-Hispanic and White.
 Families were compensated \$30 for every home recording session they completed.
@@
 
-Most participants were tested in a single session (\emph{n} = 15), but 7 participants contributed between 2-4 sessions as part of an ongoing longitudinal study (three from the younger group and four from the older group).
+Most participants were tested in a single session (\emph{n} = 15), but 7 participants contributed between 2-4 sessions as part of an ongoing longitudinal study.
 Although including more data from some participants could over-represent their characteristics in the model, we reasoned that this drawback was outweighed by having more available data to use for training.
@@
 Four inertial movement units (IMUs) were used to record infant movement across the day (MC10 Biostamp).
-A custom garment was designed to hold the IMUs (Figure \ref{fig:garment}).
-Internal pockets were sewn into snug-fitting infant leggings so that IMUs would stay close to the body (reducing vibration).
+A custom garment was designed to hold the IMUs: Internal pockets were sewn into snug-fitting infant leggings so that IMUs would stay close to the body (reducing vibration).
 A pocket over the thigh and a pocket just above the ankle were sewn on the lateral surface of the right and left legs of the garment.
@@
 
-Each sensor had sufficient battery and onboard storage to record accelerometer and gyroscope data for approximately 12 hours.
+Each sensor had sufficient battery and onboard torage to record accelerometer and gyroscope data for approximately 12 hours.
 We chose a sampling rate of 62.5 Hz (one of the available presets) based on prior work that used rates of 50-64 Hz \autocite{FranchakScott2021,AiraksinenRasanen2020,YaoPlotz2019}.
@@
 
-Supine was coded when the infant was lying on their back, on their side, or was reclined up to a 45 degree angle.
-Prone was coded when the infant was lying on their stomach, was stationary supported by the hands/knees or the hands/feet, or was crawling.
+Supine was coded when the infant was lying on their back, on their side, or was reclined up to a 45 Prone was coded when the infant was lying on their stomach, was stationary supported by the hands/knees or the hands/feet, or was crawling.
 We scored sitting to include any form of the following seated positions: 1) infants sat with their buttocks on a surface, such as on the floor or a caregiver's lap, 2) infant was in a kneeling-sit position, in which their knees were on the ground with their legs tucked underneath the buttocks, and 3) infant was in a seating device, such as a high chair, that kept the torso oriented perpendicular to the ground (a reclined position, such as in a young infant's car seat, would be counted as supine).
@@
 Our goal in creating a category for ``held by caregiver'' was to separate times when infants were in control of their body position from times when they were suspended in the air (rather than resting on furniture or a surface).
-Held was coded when infants were carried off of the ground.
+Held was coded when infants were carried off of the groun.
 However, when the caregiver was sitting with the infant in their lap the infant's body position was coded as if the caregiver was a surface (e.g., if the infant was sitting on the caregiver's lap this was coded as sitting).
@@
 Although overall accuracy from our semi-supervised, in-home data collection did not match the near-perfect accuracy (.95-.98) found in prior in-lab studies \autocite{FranchakScott2021,AiraksinenRasanen2020}, both models approached the level of agreement found between two human coders (\emph{M} = .906).
-Most likely, lower accuracy in the current study results from the more variable and complex behavior observed in a semi-supervised setting rather than from a difference in the quality of the classification model; although the first part of the proximal period was guided by the experimenter during a brief phone call, the remainder of the proximal period included natural behavior.
+Most likely, lower accuracy in the current study results from the more variable and complex behavior observed in a a semi-supervised setting rather than from a difference in the quality of the classification model; although the first part of the proximal period was guided by the experimenter during a brief phone call, the remainder of the proximal period included natural behavior.
 Visual inspection of Figure \ref{fig:metrics}A shows that accuracy values were heavily skewed, with many approaching perfect accuracy but a few sessions with very poor accuracy.
@@
 The highest accuracy was observed when using all four sensors (group model \emph{M} = 0.846; individual model \emph{M} = 0.916), and the lowest when using only a single sensor (left ankle group model \emph{M} = 0.667; right ankle individual model \emph{M} = 0.819).
-Accuracy for some pairs approached 4-sensor accuracy: Left ankle and thigh had group model accuracy of \emph{M} = 0.824 and individual model accuracy of \emph{M} = 0.901, only 2.2\% and 1.5\% worse than using all four sensors.
+Accuracy for some pairs approached 4-sensor accuracy: Left ankle and thigh had group model accuracy of \emph{M} = NA and individual model accuracy of \emph{M} = NA, only 2.2\% and 1.5\% worse than using all four sensors.
 Other pairings were less accurate; notably, using both ankles resulted in group model accuracy of \emph{M} = 0.719 and individual model accuracy of \emph{M} = 0.844, 12.7\% and 7.2\% worse than using all four sensors.
@@
 Among participants who wore the sensors from morning to bedtime, all infants wore the sensor garment during 100\% of the intended recording period, excluding the times caregivers were asked to remove the garment (naps, diaper changes, trips out of the house), based on caregiver logs of wear time.
-No caregiver reported removing the garment for any other reason (such as infant discomfort), meaning that the majority of the time during the day either resulted in usable body position data or was excluded due to caregiver-reported naps (gray periods in Figure \ref{fig:timelines}).
+No caregiver reporting removing the garment for any other reason (such as infant discomfort), meaning that the majority of the time during the day either resulted in usable body position data or was excluded due to caregiver-reported naps (gray periods in Figure \ref{fig:timelines}).
 The total length of the recording period ranged from 6.92-11.75 hours with a median of 9.04 hours.
@@
 Age in months and upright time were significantly, positively correlated using group (\emph{r} = 0.65, \emph{p} \textless{} .001) and individual models (\emph{r} = 0.52, \emph{p} = .002).
-Similarly, sitting time increased with age based on predictions from both group (\emph{r} = 0.57, \emph{p} \textless{} .001) and individual models (\emph{r} = 0.72, \emph{p} \textless{} .001).
-In contrast, supine time decreased with age using predictions from both group (\emph{r} = -0.61, \emph{p} \textless{} .001) and individual models (\emph{r} = -0.61, \emph{p} \textless{} .001).
+Similarly, age predicted increases in sitting time using group (\emph{r} = 0.57, \emph{p} \textless{} .001) and individual models (\emph{r} = 0.72, \emph{p} \textless{} .001).
+In contrast, age predicted decreases in supine time using group (\emph{r} = -0.61, \emph{p} \textless{} .001) and individual models (\emph{r} = -0.61, \emph{p} \textless{} .001).
 No clear age trends were found for prone time.
@@
 Linking the distributional structure of a particular motor experience in a real day, such as time spent sitting, might be predicted by sitting skill and/or might predict infants' future sitting ability.
-Moreover, real-time recording of motor behavior provides a way to measure how infant motor behavior links to other types of experiences in the moment in daily life.
-\textcite{MalachowskiSalo2023} used a combination of LENA® and EMA surveys to find that infants heard less adult speech when restrained in seating devices.
+Moreover, real-time recording of motor behavior provides a way to measure how infant motor behavior links to other types of experiences in the moment in daily life.@MalachowskiSalo2023 used a combination of LENA® and EMA surveys to find that infants heard less adult speech when restrained in seating devices.
 Combining LENA® and long-form motor recordings can take this a step further by measuring more precisely how speech and body position co-vary within a day.

### 2023-10-17 manuscript.Rmd
Hailey and Max edits
lab meeting notes and examples

@@
 shorttitle        : "Infant position in the home"
-
 author: 
@@
   We are grateful to Vanessa Scott, Tasnia Haider, and Ishapreet Kaur for their help in collecting data for the present study and to the research assistants of the UCR Perception, Action, and Development Lab for annotating videos.
-
   This work was funded by National Science Foundation Grant BCS #1941449 to the first author. The authors declare no conflict of interest.
-
 abstract: |
@@
 wordcount         : "11,995"
-
 bibliography      : "master.bib"
-
 floatsintext      : no
@@
 The first set of results showed that group and individual models trained from data during the proximal comparison period were accurate during the proximal period.
+This suggests that the immediate accuracy of body position predictions early in the recording session was strong, but does not address how accurately predictions will be later in the recording.
 In the next analysis, we examine long-term performance by testing how accurately models trained from the proximal period could predict body position during the distal comparison period.

### 2023-10-17 Lab_Meeting_.R

@@ -1,0 +1,37 @@
+library(tidyverse)
+install.packages("ggforce")
+library(ggforce)
+
+ds <- read_csv("data/group_split_metrics_class.csv")
+
+#id in this data refers to participant id and model is the 2 types of models we can look at (split and group)
+
+#Prevalence is how much was that thing observed (e.g., how prevalent was held, prone, sitting, etc.)
+
+# The equal sign is used for many ways; one equal sign is making the value change to that whereas double equal means youre just seeing what is true or a comparison 
+#Only keep the models where it says "group" and not include "split" using the double == 
+ds <- filter(ds, model == "group")
+
+#mapping is what things are on the X axis, what is on the Y and how to represent data graphically 
+#in this example we are looking at prevalence and class aka the different body positions
+ggplot(data = ds, mapping = aes(x = Class, y = Prevalence)) +
+  geom_boxplot()
+
+#you can add more things like labels for each axis or x and y lims
+ggplot(data = ds, mapping = aes(x = Class, y = Kappa)) +
+  geom_boxplot() + ylab("Reliability (Kappa)") + xlab("")
+
+#when looking at scatter plots you can use geom_point
+geom_point(size = 2, alpha = .5)
+
+ggplot(data = ds, mapping = aes(x = Class, y = Kappa, size = Prevalence)) +
+  geom_point(alpha = .5) + ylim(0, 1) + ylab("Reliability (Kappa)") + xlab("")
+
+
+ggplot(ds, aes(x = Kappa, fill = Class)) + geom_histogram()
+
+ggplot(ds, aes(x = Kappa, fill = Class)) + geom_histogram() + facet_wrap(~ Class)
+
+#stat summary 
+
+#theme_minimal
